### 2020-06-19 grove-paper.tex
initial outline meeting

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\begin{document}

%% Title information
\title[Short Title]{Grove: A Convergent Collaborative Structure-Editor Calculus}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011008</concept_id>
%<concept_desc>Software and its engineering~General programming languages</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003456.10003457.10003521.10003525</concept_id>
%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}
%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
\label{sec:introduction}

Motivation: 
- collaborative editing (both synchronous ala Google Docs and asynchronous version control) 
  is good and important as computing grows
- semantic structure editing is good because it solves the gap problem (semantic editor services 
  are always available) -- cite Hazelnut papers (talk about holes)
- previous approaches to collaborative editing have limitations
  - diff/merge based approaches (trying to solve the inverse problem based on final states -- 
    you lose the actual actions that were performed, and have to reconstruct them or an approx.
    of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
    some papers have started to look at that, but fundamentally we don't want to throw away the
    knowledge we have about the edits!)
  - operational transforms (complexity, you have to patch previous actions based on new actions)
  - CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
    because it is relatively simple: you just send all the edits to all the replicas and they are 
    convergent by design
- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
  the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility 
  allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
  using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
  the semantic editor services will be disabled because the program is going to be broken in multiple places
  transiently)

  this is tricky because:
  - some edits might be conflicting -- solve this with "conflict holes"
  - adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
- since we are commutative, we solve both synchronus and async collaborative editing
  - and this resolves issues around merges and conflicts
- contribution of this paper is to solve these problems from type-theoretic first principles:
  - ...
- Hazel

\subsection{Contributions and Paper Organization}
\label{sec:paper-organization}

\section{Grove By Example}
\label{sec:grove-by-example}

TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise

\subsection{Scenario A: Editing Different Parts of the Code}
A + B make edits and transmit to C, C is able to resolve without any assistance, 
order that the edits are received doesn't matter you get the same program state
at the end

\subsection{Scenario B: Edits to Nested Parts of the Code}
A edits part of the code

B moves the code surrounding the edits of A to another part of the code

\subsection{Scenario C: Direct Conflicts}

A and B both insert different expressions into the same hole

(think about critiques re: diff)

(compare to what happens when there are conflicts in a VCS, you get 
diff markers inserted into files and the semantic editor services 
no longer understand the code)

\subsection{Scenario D: Deletion of Code Being Edited}
B deletes code that A is editing (introduce notion of identity)

(orphans)

Figure shared between D and E that shows the deleted nodes and restore button

\subsection{Scenario E: Restoring Deleted Code}
If only one restore, then it does what you expect (basically a cut and paste)

If multiple replicas restore to different places, then you get some complexity:

 -- restore isn't a copy, it is a move

Multi-parent

Cycles

(Maybe not in initial grove by example) The A and B diamond example

\section{Formalism}
\label{sec:formalism}

\subsection{Syntax}

\subsection{Convergent Graphs}

Graph CRDT description

Mapping of syntax into the graph representation

Graph actions

holes are not explicit in the graph

\begin{theorem}[Commutativity (convergency?)]

\end{theorem}

\subsection{Type System}
\label{sub:type-system}

conflicts (multiple children in the same spot)

multi-parent

cycles

non-empty holes

\subsection{Semantic Actions}
Mapping to graph actions

Sensibility

\subsection{Agda Mechanization}
\label{sub:agda-mechanization}

\section{Implementation}
\label{sec:implementation}
GRV -- how it is implemented, how it connects to the formalism, describe the graph,

\section{Related Work}
\label{sec:related-work}

Hazel

CRDTs

Pidjul

Etherpad

Live Share

Git

Darcs

Unision?

\section{Discussion and Conclusion}
\label{sec:conclusion}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2020-07-01 grove-paper.tex
Paper meta-data and some figures

@@
 \acmVolume{1}
-\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
-\acmYear{2018}
+\acmYear{2021}
 \acmMonth{1}
@@
 
-%% Bibliography style
 \bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
+\citestyle{acmauthoryear}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\renewcommand{\topfraction}{1} % Allow floats to take up the page
+\renewcommand{\textfraction}{0}
 
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
+%% Packages
+\usepackage{booktabs}
+\usepackage{subcaption}
+\usepackage{tikz}
+
+%%%%%%%%
+% TikZ Stuff
+%\usepackage{etex} % Fix "No room for new \dimen" error
+\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
+\usepackage{tikz}
+\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
+
+\usetikzlibrary{external}
+%\tikzexternalize % activate!
+%\usepackage{lua-visual-debug}
+
+\usetikzlibrary{arrows.meta} % Arrow Tips
+\tikzset{>=Stealth}
+%\tikzset{<=stealth}
+%\tikzset{arrows={-Stealth[scale=50]}}
+%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
+%\tikzset{wideline/.style={line width=0.7pt}}
+%\tikzset{boldline/.style={color=black,line width=1.0pt}}
+
+\usetikzlibrary{
+  graphs,       % Graph *notation*
+  graphdrawing, % Graph *layout*
+}
+
+\usegdlibrary{
+  trees,
+}
 
@@
 %% Title information
-\title[Short Title]{Grove: A Convergent Collaborative Structure-Editor Calculus}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-% \subtitle{Subtitle}                     %% \subtitle is optional
-%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
+\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
+%\subtitle{Subtitle}
 
 %% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
+\author{Michael D. Adams}
+\orcid{0000-0003-3160-6972}
 
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\author{Eric Griffis}
+\orcid{nnnn-nnnn-nnnn-nnnn}
+
+\author{Cyrus Omar}
+\orcid{0000-0003-4502-7971}
 \affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
+  \position{Assistant Research Scientist}
+  \department[0]{Computer Science and Engineering}
+  \department[1]{Electrical Engineering and Computer Science}
+  \department[2]{College of Engineering}
+  \institution{University of Michigan}
+  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
+  \city{Ann Arbor}
+  \state{MI}
+  \postcode{48109-2121}
+  \country{USA}
 }
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
 
@@
 
+\subsection{User Interface}
+
+TODO: include mocked-up screenshots
+
+Say: "might not be a popup"
+
 \subsection{Scenario A: Editing Different Parts of the Code}
+
+\begin{figure}
+\centering
+TODO: commutativity diagram of screenshots with two cursors
+\caption{TODO:Caption}
+\label{fig:my_label}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+% TODO: texttt
+% TODO: position the different graphs
+% TODO: ballanced trees
+% TODO: use tikz graph notation? tree notation?
+\path (0cm,0cm) graph[tree layout] {
+  + -> {
+    3[as={\texttt{3}}], % TODO: put a cursor (green) here
+    4 % TODO: put a cursor (red) here
+  }
+};
+\path (-3cm,-2cm) graph[tree layout] {
+  + -> {
+    13, % TODO: put a cursor here
+    4
+  }
+};
+\path (3cm,-2cm) graph[tree layout] {
+  + -> {
+    3,
+    14 % TODO: put a cursor here
+  }
+};
+\path (0cm,-4cm) graph[tree layout] {
+  + -> {
+    13, % TODO: put a cursor here
+    14 % TODO: put a cursor here
+  }
+};
+\end{tikzpicture}
+\caption{Scenario A: Editing Different Parts of the Code}
+\label{fig:Scenario-A}
+\end{figure}
+
 A + B make edits and transmit to C, C is able to resolve without any assistance, 
@@
 \subsection{Scenario B: Edits to Nested Parts of the Code}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+% TODO: texttt
+% TODO: position the different graphs
+% TODO: ballanced trees
+% TODO: use tikz graph notation? tree notation?
+\path (0cm,0cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    * -> {
+      2,
+      x
+    }
+  }
+};
+\path (-3cm,-2cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    + -> {
+      1,
+      * -> {
+        2,
+        x
+      }
+    }
+  }
+};
+\path (3cm,-2cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    * -> {
+      3,
+      x
+    }
+  }
+};
+\path (0cm,-4cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    + -> {
+      1,
+      * -> {
+        3,
+        x
+      }
+    }
+  }
+};
+\end{tikzpicture}
+\caption{Scenario B: Edits to Nested Parts of the Code}
+\label{fig:Schenario-B}
+\end{figure}
+
 A edits part of the code
@@
 \subsection{Scenario C: Direct Conflicts}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+% TODO: texttt
+% TODO: position the different graphs
+% TODO: ballanced trees
+% TODO: use tikz graph notation? tree notation?
+\path (0cm,0cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    x
+  }
+};
+\path (-3cm,-2cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    + -> {
+      1,
+      x
+    }
+  }
+};
+\path (3cm,-2cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    * -> {
+      2,
+      x
+    }
+  }
+};
+\path (2cm,-4cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    + -> {
+      1,
+      * -> {
+        3,
+        x
+      }
+    }
+  }
+};
+\path (-2cm,-4cm) graph[tree layout] {
+  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+    * -> {
+      3,
+      + -> {
+        1,
+        x
+      }
+    }
+  }
+};
+\end{tikzpicture}
+\caption{Scenario C: Direct Conflicts}
+\label{fig:Scenario-C}
+\end{figure}
+
+DIAGRAM: lam x. x
+
+DIAGRAM: lam x. 1 + x
+
+DIAGRAM: lam x. 2 * x
+
+DIAGRAM: Merge: diamond with "1 +" and "2 *"
+
+DIAGRAM: Resolution \#1: lam x. 1 + 2 * x
+
+DIAGRAM: Resolution \#2: lam x. 2 * (1 + x)
 
@@
 \subsection{Scenario D: Deletion of Code Being Edited}
-B deletes code that A is editing (introduce notion of identity)
+
+DIAGRAM: lam x. 2 + x
+
+DIAGRAM: (lam x. HOLE) (2 + x)
+
+DIAGRAM: lam x. 3 + x
+
+DIAGRAM: Merge: (lam x. HOLE) (3 + x)
+
+A deletes code that B is editing (introduce notion of identity)
 
@@
 
-\begin{theorem}[Commutativity (convergency?)]
+\begin{theorem}[Commutativity (convergence?)]
 
@@
 GRV -- how it is implemented, how it connects to the formalism, describe the graph,
+
+We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
+
+Compared to the core language, our implementation lacks support for lists.
 

### 2020-07-03 grove-paper.tex
Updates from Overleaf

@@
 \usepackage{subcaption}
-\usepackage{tikz}
+\usepackage[pdf]{graphviz}
 
@@
 
-\usetikzlibrary{external}
-%\tikzexternalize % activate!
+\usetikzlibrary{external} % Needs to be separately enabled
+%\tikzexternalize % Enable externalization
 %\usepackage{lua-visual-debug}
@@
 
+TODO: gray-scale unchanged parts of diagrams
+
 TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise
 
-\subsection{User Interface}
+TODO: annotate vertexes and edges with ids (gray superscript?)
 
-TODO: include mocked-up screenshots
+Single User Intuition:
+ - manipulating trees,
+ - holes
 
-Say: "might not be a popup"
+Multi User Intuition:
+ - vertexes have unique ids
+ - edits create or delete edges in the tree,
+ - vertexes are automatically created when an edge goes to it
+
+Deeper intuition:
+ - edges have unique ids
+ - once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
+
+\subsection{Scenario X: Holes in the Tree}
+\label{sub:Holes in the Tree}
+
+\begin{figure}
+\centering
+\digraph[]{simple-tree-with-hole}{
+  node [shape=Mrecord,fontsize=11,ranksep=0];
+  edge [arrowhead=vee,fontsize=11,weight=2];
+  rankdir = LR;
+  {rank=min; n0 [shape=point]};
+  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
+  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
+  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
+  n3:left -> n6 [color=black,label="7",edgeURL="#",edgetooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6",labeltooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6"];
+  n0:root -> n3 [color=black,label="4",edgeURL="#",edgetooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3",labeltooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3"]}
+\begin{tikzpicture}
+% TODO: texttt
+% TODO: position the different graphs
+\path (0cm,0cm) graph[tree layout] {
+  + -> {
+    2[as={\texttt{2}}],
+    % <empty>
+  }
+};
+\end{tikzpicture}
+\caption{Scenario X: Holes in the Tree}
+\label{fig:Scenario X: Holes in the Tree}
+\end{figure}
+
+DIAGRAM: 2 + HOLE
+
+NEEDS UI EXAMPLE
 
@@
 \caption{TODO:Caption}
-\label{fig:my_label}
+\label{fig:Screenshot0}
 \end{figure}
@@
 % TODO: position the different graphs
-% TODO: ballanced trees
-% TODO: use tikz graph notation? tree notation?
 \path (0cm,0cm) graph[tree layout] {
@@
 \caption{Scenario A: Editing Different Parts of the Code}
-\label{fig:Scenario-A}
+\label{fig:Scenario A: Editing Different Parts of the Code}
 \end{figure}
@@
 % TODO: position the different graphs
-% TODO: ballanced trees
-% TODO: use tikz graph notation? tree notation?
 \path (0cm,0cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     * -> {
@@
 \path (-3cm,-2cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     + -> {
@@
 \path (3cm,-2cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     * -> {
@@
 \path (0cm,-4cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     + -> {
@@
 
+TODO: REDO to MOVE a SUBTREE
+
+NEEDS UI MOCKUP
+
 A edits part of the code
@@
 B moves the code surrounding the edits of A to another part of the code
+
+Contrast with a traditional diff
 
@@
 % TODO: position the different graphs
-% TODO: ballanced trees
-% TODO: use tikz graph notation? tree notation?
 \path (0cm,0cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     x
@@
 \path (3cm,-2cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     * -> {
-      2,
-      x
+      x,
+      2
     }
@@
 };
-\path (2cm,-4cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+\path (0cm,-4cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
+    + -> {
+      1,
+      x
+    },
+    * -> {
+      x,
+      2
+    }
+  }
+};
+\path (2cm,-8cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     + -> {
@@
       * -> {
-        3,
-        x
+        x,
+        2
       }
@@
 };
-\path (-2cm,-4cm) graph[tree layout] {
-  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
+\path (-2cm,-8cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
     * -> {
-      3,
       + -> {
@@
         x
-      }
+      },
+      2
     }
@@
 
+TODO: SHOW intermediate Steps (maybe no graph)
+  Show user's perspective
+  Accept one side = delete others (select this)
+  Delete x in *
+  Move + to x
+
+UI Level Move = Delete and Add
+
 DIAGRAM: Resolution \#1: lam x. 1 + 2 * x
@@
 DIAGRAM: Resolution \#2: lam x. 2 * (1 + x)
+
+NEEDS UI MOCKUP
+
+Say: "might not be a popup"
 
@@
 
+\begin{figure}
+\centering
+\begin{tikzpicture}
+% TODO: texttt
+% TODO: position the different graphs
+\path (0cm,0cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
+    + -> {
+      2,
+      x
+    }
+  }
+};
+\path (-3cm,-2cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
+    + -> {
+      3,
+      x
+    }
+  }
+};
+\path (3cm,-2cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
+    hole
+    },
+  + -> {
+    2,
+    x
+  }
+};
+\path (0cm,-4cm) graph[tree layout] {
+  lam[as={$\lambda_{\texttt{x}}$}] -> {
+    hole
+    },
+  + -> {
+    3,
+    x
+  }
+};
+\end{tikzpicture}
+\caption{Scenario D: Deletion of Code Being Edited}
+\label{fig:Scenario-D}
+\end{figure}
+
+TODO: move into Scenario B
+
 DIAGRAM: lam x. 2 + x
@@
 
+NEEDS UI MOCKUP
+
 A deletes code that B is editing (introduce notion of identity)
@@
 \subsection{Scenario E: Restoring Deleted Code}
+
+Combine with Scenario B (moving code):
+  delete
+  restore
+  cut/paste (i.e., move) = delete then restore
+  copy structure vs copy id
+  paste structure vs paste id
+    (show two or three options at paste time)
+
+DIAGRAM: 1 * 2 + 3 + HOLE
+
+DIAGRAM: HOLE + 3 + HOLE (1 * 2)
+
+DIAGRAM: HOLE + 3 + 1 * 2
+
+NEEDS UI MOCKUP
+
 If only one restore, then it does what you expect (basically a cut and paste)
@@
 (Maybe not in initial grove by example) The A and B diamond example
+
+\subsection{Scenario F: Cycles}
+
+Synthetic cycles:
+
+DIAGRAM: restoring into yourself (paste before delete?)
+
+Organic cycles:
+
+DIAGRAM:
+
+NEEDS UI MOCKUP
 
@@
 
-Compared to the core language, our implementation lacks support for lists.
+Compared to the core language, our implementation lacks support for lists and case expressions.
+
+Optimizations: least fixed point
 

### 2020-07-03 grove-paper.tex
Debugging GraphViz

@@
 \centering
-\digraph[]{simple-tree-with-hole}{
+\digraph[]{simple-tree-with-holeX}{
   node [shape=Mrecord,fontsize=11,ranksep=0];

### 2020-07-03 grove-paper.tex
Get graphviz to play nice with latexmk

--- 2020-07-03 grove-paper.tex
@@
 \usepackage{subcaption}
-\usepackage[pdf]{graphviz}
+\usepackage{graphicx}
+
+% Support \includegraphics of .dot files
+\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
 
@@
 \centering 
-\digraph[]{simple-tree-with-hole}{
-  node [shape=Mrecord,fontsize=11,ranksep=0];
-  edge [arrowhead=vee,fontsize=11,weight=2];
-  rankdir = LR;
-  {rank=min; n0 [shape=point]};
-  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
-  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
-  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
-  n3:left -> n6 [color=black,label="7",edgeURL="#",edgetooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6",labeltooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6"];
-  n0:root -> n3 [color=black,label="4",edgeURL="#",edgetooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3",labeltooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3"]}
+\includegraphics{simpletreewithholeX.dot}
+%\digraph[]{simpletreewithholeX}{
+%  node [shape=Mrecord,fontsize=11,ranksep=0];
+%  edge [arrowhead=vee,fontsize=11,weight=2];
+%  rankdir = LR;
+%  {rank=min; n0 [shape=point]};
+%  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
+%  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
+%  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
+%  n3:left -> n6 [color=black,label="7",edgeURL="#"];
+%  n0:root -> n3 [color=black,label="4",edgeURL="#"]}
 \begin{tikzpicture}

### 2020-07-03 grove-paper.tex
Updates from Overleaf

@@
 \usepackage{subcaption}
-\usepackage{graphicx}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
+\usepackage[pdf]{graphviz}
 
@@
 \centering
-\includegraphics{simpletreewithholeX.dot}
-%\digraph[]{simpletreewithholeX}{
-%  node [shape=Mrecord,fontsize=11,ranksep=0];
-%  edge [arrowhead=vee,fontsize=11,weight=2];
-%  rankdir = LR;
-%  {rank=min; n0 [shape=point]};
-%  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
-%  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
-%  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
-%  n3:left -> n6 [color=black,label="7",edgeURL="#"];
-%  n0:root -> n3 [color=black,label="4",edgeURL="#"]}
+\digraph[]{simpletreewithholeX}{
+  node [shape=Mrecord,fontsize=11,ranksep=0];
+  edge [arrowhead=vee,fontsize=11,weight=2];
+  rankdir = LR;
+  {rank=min; n0 [shape=point]};
+  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
+  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
+  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
+  n3:left -> n6 [color=black,label="7",edgeURL="#"];
+  n0:root -> n3 [color=black,label="4",edgeURL="#"]}
 \begin{tikzpicture}

### 2020-07-03 grove-paper.tex
Trying to get preview working in Overleaf
Make files be editable in Overleaf

--- 2020-07-03 grove-paper.tex
@@
 \centering 
-\includegraphics{simpletreewithholeX.dot}
-%\digraph[]{simpletreewithholeX}{
-%  node [shape=Mrecord,fontsize=11,ranksep=0];
-%  edge [arrowhead=vee,fontsize=11,weight=2];
-%  rankdir = LR;
-%  {rank=min; n0 [shape=point]};
-%  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
-%  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
-%  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
-%  n3:left -> n6 [color=black,label="7",edgeURL="#"];
-%  n0:root -> n3 [color=black,label="4",edgeURL="#"]}
-\begin{tikzpicture}
-% TODO: texttt
-% TODO: position the different graphs
-\path (0cm,0cm) graph[tree layout] {
-  + -> {
-    2[as={\texttt{2}}],
-    % <empty>
-  }
-};
-\end{tikzpicture}
+\includegraphics[scale=0.75]{graphviz/simple-tree-with-hole.dot}
 \caption{Scenario X: Holes in the Tree}
@@
 \centering
-\begin{tikzpicture}
-% TODO: texttt
-% TODO: position the different graphs
-\path (0cm,0cm) graph[tree layout] {
-  + -> {
-    3[as={\texttt{3}}], % TODO: put a cursor (green) here
-    4 % TODO: put a cursor (red) here
-  }
-};
-\path (-3cm,-2cm) graph[tree layout] {
-  + -> {
-    13, % TODO: put a cursor here
-    4
-  }
-};
-\path (3cm,-2cm) graph[tree layout] {
-  + -> {
-    3,
-    14 % TODO: put a cursor here
-  }
-};
-\path (0cm,-4cm) graph[tree layout] {
-  + -> {
-    13, % TODO: put a cursor here
-    14 % TODO: put a cursor here
-  }
-};
-\end{tikzpicture}
+\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-1.dot}
+\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-2.dot}
+\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-3.dot}
+\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-4.dot}
 \caption{Scenario A: Editing Different Parts of the Code}

### 2020-07-13 grove-paper.tex
Updates from Overleaf

@@
 }
+
+%%%%%%%%
+% Lang components
+
+% sets
+\def\E{\mathcal{E}}
+\def\G{\mathcal{G}}
+\def\I{\mathcal{I}}
+\def\K{\mathcal{K}}
+\def\U{\mathcal{U}}
+\def\V{\mathcal{V}}
+
+% constructors
+\def\ExpVar{\mathsf{Exp\_var}}
+\def\ExpLam{\mathsf{Exp\_lam}}
+\def\ExpApp{\mathsf{Exp\_app}}
+\def\ExpNum{\mathsf{Exp\_num}}
+\def\ExpPlus{\mathsf{Exp\_plus}}
+\def\PatVar{\mathsf{Pat\_var}}
+\def\TypNum{\mathsf{Typ\_num}}
+\def\TypArrow{\mathsf{Typ\_arrow}}
+
+% indexes
+\def\Root{\mathsf{Root}}
+\def\LamParam{\mathsf{Lam\_param}}
+\def\LamType{\mathsf{Lam\_type}}
+\def\LamBody{\mathsf{Lam\_body}}
+\def\AppFun{\mathsf{App\_fun}}
+\def\AppArg{\mathsf{App\_arg}}
+\def\PlusLeft{\mathsf{Plus\_left}}
+\def\PlusRight{\mathsf{Plus\_right}}
+\def\ArrowArg{\mathsf{Arrow\_arg}}
+\def\ArrowResult{\mathsf{Arrow\_result}}
+
+% relations
+\DeclareMathOperator{\arity}{\text{arity}}
+
+\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
 
@@
 
-\subsection{Scenario X: Holes in the Tree}
-\label{sub:Holes in the Tree}
+For presentational simplicity we omit param\_type in all lambdas
+
+\subsection{Structure Editing}
+
+We present this at two levels.
+One is the underlying graph model, which
+involves explicit identifiers, graph edges, indexes, and add and delete
+operations.
+
+The other is the interface a user might see,
+which omits these details unless needed.
+
+Throughout these examples the graph model is presented
+as a graph with.
+
+\subsubsection{Trees and Holes}
+\label{sub:Trees and Holes}
+
+TODO: explain holes.
 
@@
 \centering 
-\includegraphics[scale=0.75]{graphviz/simple-tree-with-hole.dot}
-\caption{Scenario X: Holes in the Tree}
-\label{fig:Scenario X: Holes in the Tree}
+\includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
+\caption{A simple tree with a hole}
+\label{fig:A simple tree with a hole}
+\end{figure}
+NEEDS UI EXAMPLE
+
+Consider the tree in Figure~\ref{fig:A simple tree with a hole},
+which represents the code \texttt{2 + \hole}.
+
+Note that this tree contains three vertexes.
+The first is depicted with a simple dot and is the root vertex.
+The other two are labeled \texttt{plus} and \texttt{num(2)}
+and represent the plus operator and the numeric literal 2, respectively.
+
+Each vertex has a unique identifier and a list of positions
+for child vertexes.
+In this figure, the \texttt{plus} and \texttt{num(2)} vertexes have
+identifiers 3 and 6 respectively,
+the \texttt{plus} vertex has positions for its left and right children
+and the \texttt{num(2)} vertex has no positions for any children.
+The root vertex has the distinguished identifier 0
+and has one position for its children.
+
+Also each edge has a unique identifier. In Figure~\ref{fig:A simple tree with a hole},
+these edges have idenifiers 4 and 7.  These are used
+when adding and deleting edges as discussed in the
+next subsubsection.
+
+In this paper we represent variables and numeric constants
+as leaves and represent their names or values as
+arguments to those constructors (e.g., \texttt{2} in \texttt{num(2)}).
+(In TODO:REF, we discuss how to allow edits within these.)
+
+TODO: web page Svg ?
+
+Declarative sentences
+
+Outline refinement
+
+The edge coming out of it to the TODO:word of the expression.
+We call this the \emph{rooted} vertex.
+
+\subsubsection{Delete and Add}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/delete-and-add/1.dot}};
+\path (0,-4cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/2.dot}};
+\path (0,-8cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/3.dot}};
+\end{tikzpicture}
+\caption{Delete and Add}
+\label{fig:Delete and Add}
+\end{figure}
+NEEDS UI EXAMPLE
+
+We model edits to the code as sequences of deletions and additions of edges to this graph.
+For example, consider the first graph in Figure~\ref{fig:Delete and Add}, which
+represents the expression \texttt{1 + 2}.
+Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
+becomes \textt{1 + 3}.
+We model this by first deleting edge~7.
+This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
+represents the expression \texttt{1 + \hole}.
+Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
+Vertex~9 is just not reachable from the root.
+
+After that deletion, we then add a new vertex to represent~\texttt{3} and an
+edge to it.  This results in the bottom graph in Figure~\ref{fig:Delete and Add}.
+
+Note that this edge has a new identifier distinct from the one before.
+An edge once deleted cannot be restored.  Instead a new edge (with a new id) is created.
+
+Explain orphans
+
+sNote that we will omit orphans from diagrams if they are not relevant
+(e.g., they are deleted and are not added back later)
+
+\subsubsection{Move}
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
+\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
+\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
+\end{tikzpicture}
+\caption{Move}
+\label{fig:Move}
 \end{figure}
 
-DIAGRAM: 2 + HOLE
+\subsubsection{Cut and Paste}
 
-NEEDS UI EXAMPLE
+cut/paste (i.e., move) = delete then add
+copy structure vs copy id
+paste structure vs paste id
+  (show two or three options at paste time)
 
-\subsection{Scenario A: Editing Different Parts of the Code}
+If only one restore, then it does what you expect (basically a cut and paste)
+
+If multiple replicas restore to different places, then you get some complexity:
+
+Restore isn't a copy, it is a move
+
+\subsection{Collaborative Structure Editing}
+
+\subsubsection{Editing Different Parts of the Code}
+
+\begin{figure}
+\centering
+TODO: add screenshot of code text to each node
+TODO: remove orphans from diagram
+\begin{tikzpicture}
+\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
+\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
+\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
+\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
+\end{tikzpicture}
+\caption{Editing Different Parts of the Code}
+\label{fig:Editing Different Parts of the Code}
+\end{figure}
+
+A + B make edits and transmit to C, C is able to resolve without any assistance, 
+order that the edits are received doesn't matter you get the same program state
+at the end
+
+commutativity
 
@@
 
+\subsubsection{Editing Nested Parts of the Code}
+
 \begin{figure}
 \centering
-\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-1.dot}
-\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-2.dot}
-\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-3.dot}
-\includegraphics[scale=0.75]{graphviz/editing-different-parts-of-the-code-4.dot}
-\caption{Scenario A: Editing Different Parts of the Code}
-\label{fig:Scenario A: Editing Different Parts of the Code}
+TODO: add screenshot of code text to each node
+TODO: remove orphans from diagram
+TODO: annotate edges with delete and restore
+\begin{tikzpicture}
+\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
+\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
+\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
+\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
+\end{tikzpicture}
+\caption{Editing Nested Parts of the Code}
+\label{fig:Editing Nested Parts of the Code}
 \end{figure}
 
-A + B make edits and transmit to C, C is able to resolve without any assistance, 
-order that the edits are received doesn't matter you get the same program state
-at the end
+A deletes code that B is editing
+Then A pastes it somewhere else
 
-\subsection{Scenario B: Edits to Nested Parts of the Code}
+Explain that can continue editing orphans
+
+Contrast with a traditional diff
+
+\subsubsection{Direct conflicts}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
+\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
+\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
+\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
+\end{tikzpicture}
+\caption{Direct Conflicts}
+\label{fig:Direct Conflicts}
+\end{figure}
+
+A and B both insert different expressions into the same hole
+
+Formal conflict doesn't nessisarily mean a UI conflict.
+Mention resolution algorithms and that we discuss idioms
+as automatic quick fixes
+
+Multiple adds cause conflicts
+(Not usually presented in the UI)
+
+Resolve by deleting one or the other or deleting both and adding replacement
+
+(think about critiques re: diff)
+
+(compare to what happens when there are conflicts in a VCS, you get 
+diff markers inserted into files and the semantic editor services 
+no longer understand the code)
+
+\subsubsection{Wrapping Code}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
+\path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
+\path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
+\path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
+\path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
+\end{tikzpicture}
+\caption{Wrapping}
+\label{fig:Wrapping}
+\end{figure}
+
+\subsubsection{Conflicting wrap operations}
+Multi-parent
 
@@
 % TODO: position the different graphs
-\path (0cm,0cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    * -> {
-      2,
-      x
-    }
-  }
-};
-\path (-3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      1,
-      * -> {
-        2,
-        x
-      }
-    }
-  }
-};
-\path (3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    * -> {
-      3,
-      x
-    }
-  }
-};
-\path (0cm,-4cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      1,
-      * -> {
-        3,
-        x
-      }
-    }
-  }
-};
+\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
+\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
+\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
+\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
+% \path (0cm,0cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     x
+%   }
+% };
+% \path (-3cm,-2cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     + -> {
+%       1,
+%       x
+%     }
+%   }
+% };
+% \path (3cm,-2cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     * -> {
+%       x,
+%       2
+%     }
+%   }
+% };
+% \path (0cm,-4cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     + -> {
+%       1,
+%       x
+%     },
+%     * -> {
+%       x,
+%       2
+%     }
+%   }
+% };
+% \path (2cm,-8cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     + -> {
+%       1,
+%       * -> {
+%         x,
+%         2
+%       }
+%     }
+%   }
+% };
+% \path (-2cm,-8cm) graph[tree layout] {
+%   lam[as={$\lambda_{\texttt{x}}$}] -> {
+%     * -> {
+%       + -> {
+%         1,
+%         x
+%       },
+%       2
+%     }
+%   }
+% };
 \end{tikzpicture}
-\caption{Scenario B: Edits to Nested Parts of the Code}
-\label{fig:Schenario-B}
+\caption{Conflicting wraps}
+\label{fig:Conflicting wraps}
 \end{figure}
 
-TODO: REDO to MOVE a SUBTREE
-
-NEEDS UI MOCKUP
-
-A edits part of the code
-
-B moves the code surrounding the edits of A to another part of the code
-
-Contrast with a traditional diff
-
-\subsection{Scenario C: Direct Conflicts}
+TODO: SHOW intermediate Steps (maybe no graph)
+  Show user's perspective
+  Accept one side = delete others (select this)
+  Delete x in *
+  Move + to x
 
@@
 \begin{tikzpicture}
-% TODO: texttt
-% TODO: position the different graphs
 \path (0cm,0cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    x
-  }
-};
-\path (-3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      1,
-      x
-    }
-  }
-};
-\path (3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    * -> {
-      x,
-      2
-    }
-  }
-};
-\path (0cm,-4cm) graph[tree layout] {
   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 };
-\path (2cm,-8cm) graph[tree layout] {
+\path (2cm,-4cm) graph[tree layout] {
   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 };
-\path (-2cm,-8cm) graph[tree layout] {
+\path (-2cm,-4cm) graph[tree layout] {
   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 \end{tikzpicture}
-\caption{Scenario C: Direct Conflicts}
-\label{fig:Scenario-C}
+\caption{Wrap Resolution}
+\label{fig:Wrap Resolution}
 \end{figure}
-
-DIAGRAM: lam x. x
-
-DIAGRAM: lam x. 1 + x
-
-DIAGRAM: lam x. 2 * x
-
-DIAGRAM: Merge: diamond with "1 +" and "2 *"
-
-TODO: SHOW intermediate Steps (maybe no graph)
-  Show user's perspective
-  Accept one side = delete others (select this)
-  Delete x in *
-  Move + to x
-
-UI Level Move = Delete and Add
-
-DIAGRAM: Resolution \#1: lam x. 1 + 2 * x
-
-DIAGRAM: Resolution \#2: lam x. 2 * (1 + x)
 
@@
 
-A and B both insert different expressions into the same hole
-
-(think about critiques re: diff)
-
-(compare to what happens when there are conflicts in a VCS, you get 
-diff markers inserted into files and the semantic editor services 
-no longer understand the code)
-
-\subsection{Scenario D: Deletion of Code Being Edited}
-
-\begin{figure}
-\centering
-\begin{tikzpicture}
-% TODO: texttt
-% TODO: position the different graphs
-\path (0cm,0cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      2,
-      x
-    }
-  }
-};
-\path (-3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      3,
-      x
-    }
-  }
-};
-\path (3cm,-2cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    hole
-    },
-  + -> {
-    2,
-    x
-  }
-};
-\path (0cm,-4cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    hole
-    },
-  + -> {
-    3,
-    x
-  }
-};
-\end{tikzpicture}
-\caption{Scenario D: Deletion of Code Being Edited}
-\label{fig:Scenario-D}
-\end{figure}
-
-TODO: move into Scenario B
-
-DIAGRAM: lam x. 2 + x
-
-DIAGRAM: (lam x. HOLE) (2 + x)
-
-DIAGRAM: lam x. 3 + x
-
-DIAGRAM: Merge: (lam x. HOLE) (3 + x)
-
-NEEDS UI MOCKUP
-
-A deletes code that B is editing (introduce notion of identity)
-
-(orphans)
-
-Figure shared between D and E that shows the deleted nodes and restore button
-
-\subsection{Scenario E: Restoring Deleted Code}
-
-Combine with Scenario B (moving code):
-  delete
-  restore
-  cut/paste (i.e., move) = delete then restore
-  copy structure vs copy id
-  paste structure vs paste id
-    (show two or three options at paste time)
-
-DIAGRAM: 1 * 2 + 3 + HOLE
-
-DIAGRAM: HOLE + 3 + HOLE (1 * 2)
-
-DIAGRAM: HOLE + 3 + 1 * 2
-
-NEEDS UI MOCKUP
-
-If only one restore, then it does what you expect (basically a cut and paste)
-
-If multiple replicas restore to different places, then you get some complexity:
-
- -- restore isn't a copy, it is a move
-
-Multi-parent
-
-Cycles
-
-(Maybe not in initial grove by example) The A and B diamond example
-
-\subsection{Scenario F: Cycles}
+\subsubsection{Cycles}
 
@@
 
-DIAGRAM:
+DIAGRAM: if cycles between "1 +" and "2 *" in previous diagrams
 
 NEEDS UI MOCKUP
+
+(Maybe not in initial grove by example) The A and B diamond example
 
@@
 
+% $\arraycolsep=2pt\begin{array}{llcl}
+% \mathsf{Typ} & \tau & ::= & t ~\vert~ \aparr{\tau}{\tau} ~\vert~ \aall{t}{\tau} ~\vert~ \arec{t}{\tau} ~\vert~ \aprod{\labelset}{\mapschema{\tau}{i}{\labelset}} ~\vert~ \asum{\labelset}{\mapschema{\tau}{i}{\labelset}}\\
+% \mathsf{Exp} & e & ::= & x ~\vert~ \aelam{\tau}{x}{e} ~\vert~ \aeap{e}{e} ~\vert~ \aetlam{t}{e} ~\vert~ \aetap{e}{\tau} ~\vert~ \aefold{e} ~\vert~ \aetpl{\labelset}{\mapschema{e}{i}{\labelset}} ~\vert~  \aein{\ell}{e} \\
+% & & \vert & \aematchwith{n}{e}{\seqschemaX{r}}\\
+% \mathsf{Rule} & r & ::= & \aematchrule{p}{e}\\
+% \mathsf{Pat} & p & ::= & x  ~\vert~ \aewildp ~\vert~ \aefoldp{p} ~\vert~ \aetplp{\labelset}{\mapschema{p}{i}{\labelset}} ~\vert~ \aeinjp{\ell}{p}
+% \end{array}$
+
+\begin{figure}
+\[
+\arraycolsep=0pt
+\begin{array}{rlll}
+ p    & {}\in Pat & {}\mathrel{::=}{} & x \mid \_ \\
+ e    & {}\in Exp & {}\mathrel{::=}{} & \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
+ \tau & {}\in Typ & {}\mathrel{::=}{} & \tau \rightarrow \tau \mid Num \mid \_
+\end{array}
+\]
+\caption{Formal syntax (as a grammar) for the lambda calculus that we are considering.}
+\label{fig:formal-syntax}
+\end{figure}
+
+Turn that into a constructor set K + index set I + arity function (has type: K -> $\wp$(I)).
+
+$\K = \{ \Root, \ExpVar, \ExpLam, \ExpApp, \ExpNum, \ExpPlus,$
+
+$\qquad\ \ \PatVar, \TypNum, \TypArrow \}$
+
+$\I = \{ \Root, \LamParam, \LamType, \LamBody, \AppFun, \AppArg,$
+
+$\qquad\ \ \PlusLeft, \PlusRight, \ArrowArg, \ArrowResult \}$
+
+\vspace{2ex}
+
+$\arity : \K \rightarrow \wp(\I)$
+
+\vspace{2ex}
+
+\begin{tabular}{ll}
+    $\K$ & $\wp(\I)$ \\
+    \hline
+    $\Root$ & $\{ \Root \}$ \\
+    $\ExpLam$ & $\{ \LamParam, \LamType, \LamBody \}$ \\
+    $\ExpApp$ & $\{ \AppFun, \AppArg \}$ \\
+    $\ExpPlus$ & $\{ \PlusLeft, \PlusRight \}$ \\
+\end{tabular}
+
 \subsection{Convergent Graphs}
 
-Graph CRDT description
+\subsubsection{Graphs}
 
-Mapping of syntax into the graph representation
+Graph is a function from edges to edge states.
 
-Graph actions
+$\G : \E \rightarrow \{ \bot, +, - \}$
 
-holes are not explicit in the graph
+An edge is a uuid * source vertex * index * target vertex.  (We are assuming you can generate globally unique IDs.)
 
-\begin{theorem}[Commutativity (convergence?)]
+\def\e{\varepsilon}
+
+An edge $\e \in \E = \U \times \V \times \I \times \V$ is ...
+
+A vertex is a pair of a uuid and a constructor. (uuid + constructor is the identity)
+
+$\V = \U \times \K$
+
+An edge state is bottom + - . Define the total ordering (forms a lattice).
+
+Define $\sqcup_{edge state}$ (i.e. join)
+
+lattice ordering $\bot < + < -$ LUB e.g., join A B = C
+
+You can map from syntax to graph by selecting
+unique IDs (for both edges and vertexes).
+    - holes are not explicit in the graph
+
+\subsubsection{Graph Actions}
+A graph action is a +E or -E.
+
+Graph action semantics is a transition system
+between graphs. (join with whatever it was before)
+
+G --action--> G'
+
+\subsection{Agda Mechanization}
+\label{sub:agda-mechanization}
+
+Discuss uniqueness of uuid
+
+\begin{theorem}[Commutativity]
+
+G  -- a1 a2 --> G'
+
+------------------
+
+G' -- a2 a1 --> G
 
@@
 \label{sub:type-system}
+
+editor states:
 
@@
 
-\subsection{Semantic Actions}
+----------------
+
+User must supply edge selection for multi-parent
+
+\subsection{Edit Actions}
 Mapping to graph actions
 
-Sensibility
+Sensibility -- mapping back from graphs
+to well-typed states that fixes holes
 
-\subsection{Agda Mechanization}
-\label{sub:agda-mechanization}
+\section{TODO}
+TODO: leaves
+
+TODO: our model supports treating these as a cons-list of characters
 
@@
 
+\subsection{Variable names, strings, and numbers}
+
+NOTE: we could implement each digit as a separate characters
+
+GUI for string conflicts: use popups
+
 %% Acknowledgments

### 2020-07-14 grove-paper.tex
Updates from Overleaf

@@
 Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
-becomes \textt{1 + 3}.
+becomes \texttt{1 + 3}.
 We model this by first deleting edge~7.
@@
 
-sNote that we will omit orphans from diagrams if they are not relevant
+Note that we will omit orphans from diagrams if they are not relevant
 (e.g., they are deleted and are not added back later)
@@
 \subsubsection{Move}
+
 \begin{figure}
@@
 
-\subsubsection{Cut and Paste}
+An interesting consequence of our design is that a move operation
+does not have to delete code from one position and create a copy
+in another position.
+Instead, by deleting just the edge to the code to be moved and then
+adding a new edge to that code somewhere else, the actual code is
+moved, not just a copy.
 
-cut/paste (i.e., move) = delete then add
-copy structure vs copy id
-paste structure vs paste id
-  (show two or three options at paste time)
+For example, consider Figure~\ref{fig:Move}, which represents
+the movement of \texttt{1 * 2} from the left child of \texttt{plus}
+to its right child.  The top graph is the code before the move.
+First, edge~7 is deleted, which results in the middle graph
+where the rooted code is now \texttt{\hole{} + \hole},
+and vertex~6 is orphaned.
+
+Next edge~16 is added, which results in the bottom graph
+where the rooted code is now \texttt{\hole{} + 1 * 2},
+and vertex~6 is no longer orphaned.
+
+This move semantics gives us a richer structure than
+when treating code as a list of lines.  We exploit
+this in Section~REF:TODO in order merge edits that involve moving code.
+
+Of course, if the user wants a genuine copy
+and not just a move, that can easily be achieved
+(TODO: figure for this?)
+
+This is similar to the paste with formatting versus
+paste without formatting.
 
@@
 
+When editing different parts of the code
+
+For example, consider the graphs in Figure~\ref{fig:Editing Different Parts of the Code}.
+The
+
 A + B make edits and transmit to C, C is able to resolve without any assistance, 
@@
 \subsubsection{Editing Nested Parts of the Code}
+
+(Similar to how git merged across moves.
+A notable difference though is that \texttt{git} detects
+file moves by structural similarity rather than
+our nominal equality.
+(We exploit this in Subsection~REF:TODO).
 
@@
 
-%% Bibliography
-%\bibliography{bibfile}
-
 %% Appendix
@@
 
+%% Bibliography
+%\bibliography{bibfile}
+
 \end{document}

### 2020-07-15 grove-paper.tex
Updates from Overleaf

@@
 
+Note that we do not explicitly delete or add vertexes.
+Our only actions are deleting or adding \textit{edges}.
+Vertexes are implicitly created whenever an edge
+is created that points to it.
+Deleting vertexes is not necessary, as we can simply
+remove the edge pointing to it.
+
 \subsubsection{Move}
@@
 Of course, if the user wants a genuine copy
-and not just a move, that can easily be achieved
+and not just a move, that can be achieved
+by creating new vertexes with fresh identifiers.
 (TODO: figure for this?)
@@
 If multiple replicas restore to different places, then you get some complexity:
-
-Restore isn't a copy, it is a move
 
@@
 
-For example, consider the graphs in Figure~\ref{fig:Editing Different Parts of the Code}.
-The
+For example, suppose there are users, A and B,
+that are synchronized with each other and
+have the code \texttt{3 + 4}.
+That is to say, they both have
+the top graphs in Figure~\ref{fig:Editing Different Parts of the Code}.
 
-A + B make edits and transmit to C, C is able to resolve without any assistance, 
-order that the edits are received doesn't matter you get the same program state
-at the end
+Next, suppose user A changes the \texttt{3} to a \texttt{13},
+and user B changes the \texttt{4} to a \texttt{14}.
+Before transmitting their edits to each other,
+user A thus has left graph and user B has the right graph.
+For user A, this involves deleting edge~7 and creating edge~14.
+For user B, this involves deleting edge~10 and creating edge~18.
 
-commutativity
+To merge these two graphs each user simply replays the edge deletions and additions
+that the other user performed.
+
+Both user then have the bottom graph and are again synchronized with each other.
+
+Note that aside from transmitting the edge deletions and additions,
+no coordination between A and B is needed to coordinate the merge.
+For example, suppose that A and B did not transmit their edits to each other
+but to a third user, C, that has the original graph at the top of Figure~\ref{fig:Editing Different Parts of the Code}.
+User C can replay the edits from A and B in any order and will always get
+the graph at the bottom of Figure~\ref{fig:Editing Different Parts of the Code}.
+
+Stated more formally, our edit model obeys commutativity
+as shown in Figure~\ref{fig:Commutativity}.
+For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
+applying them in either order produces the same results.
 
@@
 TODO: commutativity diagram of screenshots with two cursors
-\caption{TODO:Caption}
-\label{fig:Screenshot0}
+\caption{TODO:Commutativity}
+\label{fig:Commutativity}
 \end{figure}

### 2020-07-16 grove-paper.tex
Updates from Overleaf

@@
 
-A deletes code that B is editing
-Then A pastes it somewhere else
+Consider Figure~\ref{fig:Editing Nested Parts of the Code}.
+Alice and Bob both start with the top graph, which represents the code \texttt{2 * x + \hole{}}.
+Alice changes the \texttt{2} to a \texttt{3}, which results in the left graph.
+Bob moves the \texttt{2 * x} from the left of the plus to the right of the plus
+by deleting edge~7 and adding edge~16, which results in the right graph.
+Since vertex and edge identifiers are preserved by these,
+both Alice's and Bob's edits can be applied to the other's graph,
+which results in both Alice and Bob having the bottom graph.
+
+As with Subsection~SEC:TODO, this merge does not require coordination
+between Alice and Bob.  They could independently send their edits
+to a third party, Charley, who could replay their edits in any order
+and will always get the bottom graph.
+
+Supporting this kind of behavior is
 

### 2020-07-20 grove-paper.tex
Updates from Overleaf

@@
 
-Supporting this kind of behavior is
+A consequence of this design is that we allow users to edit
+parts of the code that have been deleted.  For example,
+Alice might receive Bob's deletion of edge~7 before making her
+edits and before receiving Bob's adding of edge~16.
+Her edits would then be within the code deleted by Bob.
 
-Explain that can continue editing orphans
+TODO:figure
+
+Contrast this
 
@@
 
-A and B both insert different expressions into the same hole
+Finally, we address genuine conflicts between edits.
+Consider the top graph of Figure~\ref{fig:Direct Conflicts}.
+Suppose Alice inserts the number literal \texttt{2} into the body of the lambda
+while Bob inserts the function application \texttt{f x} also into the body of the lambda.
+These results in the left and right graphs of \ref{fig:Direct Conflicts} respectively.
+In particular pay attention to the edges going out of the \texttt{body} index of node~3.
+In the left graph it is edge~10, which points to a number literal,
+and in the right graph it is edge~13, which points to a function application.
 
-Formal conflict doesn't nessisarily mean a UI conflict.
-Mention resolution algorithms and that we discuss idioms
-as automatic quick fixes
+We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
+of node~3.
+This results in the bottom graph of Figure~\ref{fig:Direct Conflicts}.
+The conflict between Alice's and Bob's changes is
+represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
+index of node~3).
+In the graph representation, we flag this conflict by highlighting its edges
+in red.
+In the textual representation, we flag this conflict with the notation \texttt{\{ 2 | f x \}}.
+This is a conflict that much eventually be resolved by the user by deleting
+one of the edges and editing the other to be the merged value.
+For example, one could delete edge~10 and then edit the \texttt{x}
+to be \texttt{x + 2}.
+Note that users can continue to edit the code while in this conflicted state.
+It just must be resolved before the the code can be run.\footnote{
+It may be possible to develop evaluation models that allow these sorts of conflicts,
+but that is beyond the scope of this paper.}.
+
+Note that as a convenience to the user, certain simple
+conflicts might be automatically resolved,
+but we consider this a higher-level, user-interface consideration.
+
+TODO: need example of fully-automatic resolution and semi-automatic resolution
 

### 2020-07-21 grove-paper.tex
Updates from Overleaf

@@
 
+Compare this to what happens when there are conflicts in a traditional,
+line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
+Those systems annotate files with difference markers (e.g., sequences of \texttt{<}, \texttt{>}, or \texttt{=})
+when a merge conflict happens.
+These show alternate versions versions of the code, and it is up to the user
+to replace those alternates with the desired merged result.
+
+Our tree-based model is similar in that it presents alternate versions
+of the code, and it is up to the user to replace those alternates with the desired merged result.
+Our model is different, in that merge conflicts in the line-based model
+prevent some semantic editor services from running
+that could continue to run in the tree-based model because
+code inside and outside the conflict is still valid code.
+It is just the point where they meet the poses a potential problem for editor services.
+Another place our model is different
+is that changes are explicit rather than being inferred.
+Finally, in our model, since all changes are in terms of
+the structure of the code rather than lines,
+all changes respect that structure.
+This is unlike the line-based model where there is no guarantee
+that a merge conflict follows the structure of the code.
+Finally, our model differs from the traditional model
+in that we have explict add and remove actions
+where the traditional model infers these from a differencing algorithm.
+
 Note that as a convenience to the user, certain simple
@@
 (Not usually presented in the UI)
-
-Resolve by deleting one or the other or deleting both and adding replacement
-
-(think about critiques re: diff)
-
-(compare to what happens when there are conflicts in a VCS, you get 
-diff markers inserted into files and the semantic editor services 
-no longer understand the code)
 
@@
 \end{figure}
+
+Before further exploring how to merge changes from multiple users,
+we want to point out a common idiom
+
+Traditional: no relation (and indent might change)
 

### 2020-07-22 grove-paper.tex
Updates from Overleaf

@@
 
-Contrast this
-
 Contrast with a traditional diff
@@
 \centering
+TODO: naive add that creates a conflict
 \begin{tikzpicture}
@@
 Before further exploring how to merge changes from multiple users,
-we want to point out a common idiom
+we want to point out a common idiom that we call \emph{wrapping}.
+Consider the first graph in Figure~\ref{fig:Wrapping}, which
+represents the code \texttt{$\lambda$ x . x}, and suppose
+the user's cursor is in the body field of the lambda (i.e., around the variable
+reference to \texttt{x} in the body of the lambda).
+
+If the user requests that a plus be inserted,
+the editor may notice that there is something already in that hole.
+A naive editor could certainly add the plus by creating a new
+edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
+This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
+as shown in Figure~REF:TODO.
+A better user experience
+would be interpret the request for a plus to mean to
+take whatever (if anything) is in the cursor
+and place it as a child of the plus.
+This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
+and results in the code \texttt{$\lambda$ x . x + \hole}.
+(This is similar to the behavior in other structure editors such as in CITE:TODO.)
+Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
+Note that this is a layer above
 

### 2020-07-27 grove-paper.tex
Updates from Overleaf

@@
 
+TODO: explain that multiple delete or add are idempotent
+
+TODO: explain that delete overrides add
+
 \begin{figure}
@@
 
-\subsubsection{Direct conflicts}
+\subsubsection{Multi-child conflicts}
 
@@
 It may be possible to develop evaluation models that allow these sorts of conflicts,
-but that is beyond the scope of this paper.}.
+but that is beyond the scope of this paper.}
 
@@
 
+\subsubsection{Multi-parent conflicts}
+
+In addition to conflicts due multiple edges coming \emph{out of} an index of a particular vertex,
+conflicts can manifest as multiple edges going \emph{into} a particular vertex.
+We call these \emph{multi-parent conflicts}.
+
+\begin{figure}
+\centering
+
+TODO TOP: +(\hole, +(\hole, x))
+
+TODO LEFT: +(x, +(\hole, \hole))
+
+TODO RIGHT: +(\hole, +(x, \hole))
+
+TODO BOTTOM: Two parents pointing to x
+
+\caption{Multi-parent conflict}
+\label{fig:Multi-parent conflict}
+\end{figure}
+
+An example of this is shown in Figure~\ref{fig:Multi-parent conflict},
+which starts with the top graph, which represents the expression \texttt{\hole + (\hole + x)}.
+Suppose Alice moves the variable \texttt{x} to the left-most hole.
+This results in the left graph, which represents the expression \texttt{x + (\hole + \hole)}.
+At the same time, suppose Bob moves the variable \texttt{x} to the right-most hole.
+This results in the right graph, which represents the expression \texttt{\hole + (x + \hole)}.
+In both cases, these edits are achieved with one edge delete and one edge add.
+Alice deletes edge TODO and adds edge TODO, while Bob deletes edge TODO and adds edge TODO.
+As explained in Subsection~TODO:REF, deletes are idempotent so both Alice and Bob deleting
+edge TODO is not a problem.
+However, both edge TODO (added by Alice) and edge TODO (added by Bob) point to the same vertex.
+This generates the bottom graph, in which vertex TODO as two parents pointing to it.
+We call this situation a multi-parent conflict.
+
 \subsubsection{Wrapping Code}
@@
 
-Before further exploring how to merge changes from multiple users,
+Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
 we want to point out a common idiom that we call \emph{wrapping}.
@@
 \subsubsection{Conflicting wrap operations}
-Multi-parent
 
@@
 \end{figure}
+
+If two different 
+As seen in REF:SEC:TODO, conflicting wraps gener
+
+Multi-parent
 

### 2020-07-29 grove-paper.tex
Updates from Overleaf

@@
 
-\subsubsection{Wrapping Code}
+% \subsubsection{Wrapping Code}
 
-\begin{figure}
-\centering
-TODO: naive add that creates a conflict
-\begin{tikzpicture}
-\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
-\path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
-\path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
-\path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
-\path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
-\end{tikzpicture}
-\caption{Wrapping}
-\label{fig:Wrapping}
-\end{figure}
+% \begin{figure}
+% \centering
+% TODO: naive add that creates a conflict
+% \begin{tikzpicture}
+% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
+% \path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
+% \path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
+% \path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
+% \path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
+% \end{tikzpicture}
+% \caption{Wrapping}
+% \label{fig:Wrapping}
+% \end{figure}
 
-Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
-we want to point out a common idiom that we call \emph{wrapping}.
-Consider the first graph in Figure~\ref{fig:Wrapping}, which
-represents the code \texttt{$\lambda$ x . x}, and suppose
-the user's cursor is in the body field of the lambda (i.e., around the variable
-reference to \texttt{x} in the body of the lambda).
+% Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
+% we want to point out a common idiom that we call \emph{wrapping}.
+% Consider the first graph in Figure~\ref{fig:Wrapping}, which
+% represents the code \texttt{$\lambda$ x . x}, and suppose
+% the user's cursor is in the body field of the lambda (i.e., around the variable
+% reference to \texttt{x} in the body of the lambda).
 
-If the user requests that a plus be inserted,
-the editor may notice that there is something already in that hole.
-A naive editor could certainly add the plus by creating a new
-edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
-This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
-as shown in Figure~REF:TODO.
-A better user experience
-would be interpret the request for a plus to mean to
-take whatever (if anything) is in the cursor
-and place it as a child of the plus.
-This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
-and results in the code \texttt{$\lambda$ x . x + \hole}.
-(This is similar to the behavior in other structure editors such as in CITE:TODO.)
-Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
-Note that this is a layer above
+% If the user requests that a plus be inserted,
+% the editor may notice that there is something already in that hole.
+% A naive editor could certainly add the plus by creating a new
+% edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
+% This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
+% as shown in Figure~REF:TODO.
+% A better user experience
+% would be interpret the request for a plus to mean to
+% take whatever (if anything) is in the cursor
+% and place it as a child of the plus.
+% This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
+% and results in the code \texttt{$\lambda$ x . x + \hole}.
+% (This is similar to the behavior in other structure editors such as in CITE:TODO.)
+% Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
+% Note that this is a layer above
 
-Traditional: no relation (and indent might change)
+% Traditional: no relation (and indent might change)
 
-\subsubsection{Conflicting wrap operations}
+% \subsubsection{Conflicting wrap operations}
 
-\begin{figure}
-\centering
-\begin{tikzpicture}
-% TODO: texttt
-% TODO: position the different graphs
-\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
-\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
-\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
-\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}
+% % TODO: texttt
+% % TODO: position the different graphs
+% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
+% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
+% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
+% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
+% % \path (0cm,0cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     x
+% %   }
+% % };
+% % \path (-3cm,-2cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     + -> {
+% %       1,
+% %       x
+% %     }
+% %   }
+% % };
+% % \path (3cm,-2cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     * -> {
+% %       x,
+% %       2
+% %     }
+% %   }
+% % };
+% % \path (0cm,-4cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     + -> {
+% %       1,
+% %       x
+% %     },
+% %     * -> {
+% %       x,
+% %       2
+% %     }
+% %   }
+% % };
+% % \path (2cm,-8cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     + -> {
+% %       1,
+% %       * -> {
+% %         x,
+% %         2
+% %       }
+% %     }
+% %   }
+% % };
+% % \path (-2cm,-8cm) graph[tree layout] {
+% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
+% %     * -> {
+% %       + -> {
+% %         1,
+% %         x
+% %       },
+% %       2
+% %     }
+% %   }
+% % };
+% \end{tikzpicture}
+% \caption{Conflicting wraps}
+% \label{fig:Conflicting wraps}
+% \end{figure}
+
+% If two different 
+% As seen in REF:SEC:TODO, conflicting wraps gener
+
+% Multi-parent
+
+% TODO: SHOW intermediate Steps (maybe no graph)
+%   Show user's perspective
+%   Accept one side = delete others (select this)
+%   Delete x in *
+%   Move + to x
+
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}
 % \path (0cm,0cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     x
-%   }
-% };
-% \path (-3cm,-2cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     + -> {
-%       1,
-%       x
-%     }
-%   }
-% };
-% \path (3cm,-2cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     * -> {
-%       x,
-%       2
-%     }
-%   }
-% };
-% \path (0cm,-4cm) graph[tree layout] {
 %   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 % };
-% \path (2cm,-8cm) graph[tree layout] {
+% \path (2cm,-4cm) graph[tree layout] {
 %   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 % };
-% \path (-2cm,-8cm) graph[tree layout] {
+% \path (-2cm,-4cm) graph[tree layout] {
 %   lam[as={$\lambda_{\texttt{x}}$}] -> {
@@
 % };
-\end{tikzpicture}
-\caption{Conflicting wraps}
-\label{fig:Conflicting wraps}
-\end{figure}
+% \end{tikzpicture}
+% \caption{Wrap Resolution}
+% \label{fig:Wrap Resolution}
+% \end{figure}
 
-If two different 
-As seen in REF:SEC:TODO, conflicting wraps gener
+% NEEDS UI MOCKUP
 
-Multi-parent
-
-TODO: SHOW intermediate Steps (maybe no graph)
-  Show user's perspective
-  Accept one side = delete others (select this)
-  Delete x in *
-  Move + to x
-
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      1,
-      x
-    },
-    * -> {
-      x,
-      2
-    }
-  }
-};
-\path (2cm,-4cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    + -> {
-      1,
-      * -> {
-        x,
-        2
-      }
-    }
-  }
-};
-\path (-2cm,-4cm) graph[tree layout] {
-  lam[as={$\lambda_{\texttt{x}}$}] -> {
-    * -> {
-      + -> {
-        1,
-        x
-      },
-      2
-    }
-  }
-};
-\end{tikzpicture}
-\caption{Wrap Resolution}
-\label{fig:Wrap Resolution}
-\end{figure}
-
-NEEDS UI MOCKUP
-
-Say: "might not be a popup"
+% Say: "might not be a popup"
 

### 2020-07-31 grove-paper.tex
Updates from Overleaf

@@
 
-\subsection{Syntax}
-
 % $\arraycolsep=2pt\begin{array}{llcl}
@@
 \arraycolsep=0pt
-\begin{array}{rlll}
- p    & {}\in Pat & {}\mathrel{::=}{} & x \mid \_ \\
- e    & {}\in Exp & {}\mathrel{::=}{} & \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
- \tau & {}\in Typ & {}\mathrel{::=}{} & \tau \rightarrow \tau \mid Num \mid \_
+\begin{array}{lrlll}
+ \textrm{Patterns:}          & p    & {}\in Pat & {}\mathrel{::=}{} & x \mid \_ \\
+ \textrm{Expressions:\qquad} & e    & {}\in Exp & {}\mathrel{::=}{} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
+ \textrm{Types:}             & \tau & {}\in Typ & {}\mathrel{::=}{} & \tau \rightarrow \tau \mid Num \mid \_ \\
 \end{array}
 \]
-\caption{Formal syntax (as a grammar) for the lambda calculus that we are considering.}
-\label{fig:formal-syntax}
+\caption{Syntax as a grammar for the lambda calculus that we are considering.}
+\label{fig:Syntax as a grammar}
 \end{figure}
+
+\subsection{Syntax}
+
+The language we are modeling is shown in Figure~\ref{fig:Syntax as a grammar}.
+Standard for a simply-typed lambda calculus except that
+all terms have holes (represented by $\_$)
+and the argument to a lambda is its own term.
+The latter of which allows us to express lambdas with a hole for the bound variable.
+
+In order to model this we define
 
@@
 
-$\K = \{ \Root, \ExpVar, \ExpLam, \ExpApp, \ExpNum, \ExpPlus,$
-
-$\qquad\ \ \PatVar, \TypNum, \TypArrow \}$
-
-$\I = \{ \Root, \LamParam, \LamType, \LamBody, \AppFun, \AppArg,$
-
-$\qquad\ \ \PlusLeft, \PlusRight, \ArrowArg, \ArrowResult \}$
-
-\vspace{2ex}
-
-$\arity : \K \rightarrow \wp(\I)$
-
-\vspace{2ex}
-
-\begin{tabular}{ll}
-    $\K$ & $\wp(\I)$ \\
+\begin{figure}
+\[
+\arraycolsep=0pt
+\begin{array}{ll}
+ \arity : \K \rightarrow & \wp(\I) \\
  \hline
-    $\Root$ & $\{ \Root \}$ \\
-    $\ExpLam$ & $\{ \LamParam, \LamType, \LamBody \}$ \\
-    $\ExpApp$ & $\{ \AppFun, \AppArg \}$ \\
-    $\ExpPlus$ & $\{ \PlusLeft, \PlusRight \}$ \\
-\end{tabular}
+ \arity(\Root)={}     & \left\{ \Root \right\} \\
+ \arity(\PatVar)={}   & \left\{ \right\} \\
+ \arity(\ExpVar)={}   & \left\{ \right\} \\
+ \arity(\ExpLam)={}   & \left\{ \LamParam, \LamType, \LamBody \right\} \\
+ \arity(\ExpApp)={}   & \left\{ \AppFun, \AppArg \right\} \\
+ \arity(\ExpPlus)={}  & \left\{ \PlusLeft, \PlusRight \right\} \\
+ \arity(\TypArrow)={} & \left\{ \ArrowArg, \ArrowResult \right\} \\
+ \arity(\TypNum)={}   & \left\{ \right\} \\
+\end{array}
+\]
+\caption{Constructors, Indexes and Arity}
+\label{fig:Constructors, Indexes and Arity}
+\end{figure}
 

### 2020-08-03 grove-paper.tex
Updates from Overleaf

@@
 
+\subsubsection{Wrapping Code}
+
+TODO: move subsubsection elsewhere?
+
+\begin{figure}
+\centering
+TODO: naive add that creates a conflict
+\begin{tikzpicture}
+\path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
+\path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
+\path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
+\path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
+\path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
+\end{tikzpicture}
+\caption{Wrapping}
+\label{fig:Wrapping}
+\end{figure}
+
+Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
+we want to point out a common idiom that we call \emph{wrapping}.
+Consider the first graph in Figure~\ref{fig:Wrapping}, which
+represents the code \texttt{$\lambda$ x . x}, and suppose
+the user's cursor is in the body field of the lambda (i.e., around the variable
+reference to \texttt{x} in the body of the lambda).
+
+If the user requests that a plus be inserted,
+the editor may notice that there is something already in that hole.
+A naive editor could certainly add the plus by creating a new
+edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
+This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
+as shown in Figure~REF:TODO.
+A better user experience
+would be interpret the request for a plus to mean to
+take whatever (if anything) is in the cursor
+and place it as a child of the plus.
+This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
+and results in the code \texttt{$\lambda$ x . x + \hole}.
+(This is similar to the behavior in other structure editors such as in CITE:TODO.)
+Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
+Note that this is a layer above
+
+Traditional: no relation (and indent might change)
+
 \subsection{Collaborative Structure Editing}
@@
 We call this situation a multi-parent conflict.
-
-% \subsubsection{Wrapping Code}
-
-% \begin{figure}
-% \centering
-% TODO: naive add that creates a conflict
-% \begin{tikzpicture}
-% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
-% \path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
-% \path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
-% \path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
-% \path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
-% \end{tikzpicture}
-% \caption{Wrapping}
-% \label{fig:Wrapping}
-% \end{figure}
-
-% Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
-% we want to point out a common idiom that we call \emph{wrapping}.
-% Consider the first graph in Figure~\ref{fig:Wrapping}, which
-% represents the code \texttt{$\lambda$ x . x}, and suppose
-% the user's cursor is in the body field of the lambda (i.e., around the variable
-% reference to \texttt{x} in the body of the lambda).
-
-% If the user requests that a plus be inserted,
-% the editor may notice that there is something already in that hole.
-% A naive editor could certainly add the plus by creating a new
-% edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
-% This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
-% as shown in Figure~REF:TODO.
-% A better user experience
-% would be interpret the request for a plus to mean to
-% take whatever (if anything) is in the cursor
-% and place it as a child of the plus.
-% This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
-% and results in the code \texttt{$\lambda$ x . x + \hole}.
-% (This is similar to the behavior in other structure editors such as in CITE:TODO.)
-% Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
-% Note that this is a layer above
-
-% Traditional: no relation (and indent might change)
 
@@
 
-Synthetic cycles:
-
-DIAGRAM: restoring into yourself (paste before delete?)
-
-Organic cycles:
-
-DIAGRAM: if cycles between "1 +" and "2 *" in previous diagrams
+\begin{figure}
+\centering
+lam x. app(\#1, \hole)
+\caption{Single-User Cycles}
+\label{fig:Single-User Cycles}
+\end{figure}
+
+Since we are representing programs as graphs,
+we have to consider cycles.
+
+Cycles can occur in a both single-user and multi-user contexts.
+
+In a single user context, a cycle can appear when a user pastes
+a vertex as a child of itself.
+An example of this is shown in Figure~\ref{fig:Single-User Cycles},
+where the user has pasted vertex~TODO as a child of vertex~TODO.
+
+As a user interface consideration, it might be best to either
+disallow such edits to at least warn users that their
+edits would create a cycle.
+
+While a single-user edits can always be checked to see
+whether they create a cycle and thus the system
+multi-user edits may not manifest cycles until after
+the edits are merged.
+there are much more practical reasons why cycles
+may appear when multiple users are editing part of the code.
+
+For example, consider the situation shown in Figure~\ref{fig:Multi-User Cycles}.
+The top graph represents code that is synchronized between Alice and Bob
+and represents the code \texttt{(1 * \hole) * (2 * \hole)}.
+The left graph shows
+
+\begin{figure}
+\centering
+
+TOP: *(*(1, \hole), *(2, \hole))
+
+LEFT: *(*(1, *(2, \hole)), \hole)
+
+RIGHT: *(\hole, *(2, *(1, \hole)))
+
+BOTTOM: *(\hole, \hole) *(1, *(2, \#1))
+
+\caption{Multi-User Cycles}
+\label{fig:Multi-User Cycles}
+\end{figure}
 

### 2020-08-10 grove-paper.tex
Work on Single-User Edits

@@
 
+This section introduces collaborative structure editing in Grove by example. 
+We begin in Section~\ref{sub:Program Representation} by introducing basic edit actions
+and the underlying graph representation of expressions with holes.
+We will then discuss scenarios that arise in the collaborative setting in 
+Section~\ref{sub:Collaborative Structure Editing}.
+
 TODO: gray-scale unchanged parts of diagrams
 
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise
-
-TODO: annotate vertexes and edges with ids (gray superscript?)
-
-Single User Intuition:
- - manipulating trees,
- - holes
-
-Multi User Intuition:
- - vertexes have unique ids
- - edits create or delete edges in the tree,
- - vertexes are automatically created when an edge goes to it
-
-Deeper intuition:
- - edges have unique ids
- - once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
-
-For presentational simplicity we omit param\_type in all lambdas
-
-\subsection{Structure Editing}
-
-We present this at two levels.
+TODO: typewritter font in figure graphs
+
 One is the underlying graph model, which
@@
 
+We present this at two levels.
 The other is the interface a user might see,
 which omits these details unless needed.
-
 Throughout these examples the graph model is presented
@@
 
-\subsubsection{Trees and Holes}
-\label{sub:Trees and Holes}
-
-TODO: explain holes.
+TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
+
+\subsection{Program Representation}
+\label{sub:Program Representation}
 
@@
 \centering
+TODO: UI
+
 \includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
@@
 \end{figure}
-NEEDS UI EXAMPLE
-
-Consider the tree in Figure~\ref{fig:A simple tree with a hole},
-which represents the code \texttt{2 + \hole}.
-
-Note that this tree contains three vertexes.
-The first is depicted with a simple dot and is the root vertex.
-The other two are labeled \texttt{plus} and \texttt{num(2)}
-and represent the plus operator and the numeric literal 2, respectively.
-
-Each vertex has a unique identifier and a list of positions
-for child vertexes.
-In this figure, the \texttt{plus} and \texttt{num(2)} vertexes have
-identifiers 3 and 6 respectively,
-the \texttt{plus} vertex has positions for its left and right children
-and the \texttt{num(2)} vertex has no positions for any children.
-The root vertex has the distinguished identifier 0
-and has one position for its children.
-
-Also each edge has a unique identifier. In Figure~\ref{fig:A simple tree with a hole},
-these edges have idenifiers 4 and 7.  These are used
-when adding and deleting edges as discussed in the
-next subsubsection.
-
-In this paper we represent variables and numeric constants
-as leaves and represent their names or values as
-arguments to those constructors (e.g., \texttt{2} in \texttt{num(2)}).
-(In TODO:REF, we discuss how to allow edits within these.)
-
-TODO: web page Svg ?
-
-Declarative sentences
-
-Outline refinement
-
-The edge coming out of it to the TODO:word of the expression.
-We call this the \emph{rooted} vertex.
-
-\subsubsection{Delete and Add}
+
+The editor state for each replica in Grove consists of an expression with holes.
+For example, Figure~TODO:1(a) shows the simple arithmetic expression \texttt{2 + \hole}, which
+has a hole on the right side of the plus operator.
+
+Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
+Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
+In this figure, vertices~3 and~6 have the constructors \texttt{plus} and \texttt{num(2)}, respectively.
+Each term constructor has an associated set of child positions.
+The \texttt{plus} constructor has positions for \texttt{left} and \texttt{right} children
+and the \texttt{num(n)} constructors have no child positions.
+The root vertex has a distinguished identifier and a constructor with a single child position.
+For simplicity, we visually represent the root with a dot.
+
+Each edge is labeled with a globally unique identifier (e.g.,~4 and~7 in the figure) and a child position.
+An edge indicates that the destination vertex is a child of the origin vertex at the given position.
+Visually, we indicate the position of an edge by the location of its origin.
+For example, Figure~TODO:1(b) shows the graph corresponding to the term in Figure~TODO:1(a).
+
+TODO: holes
+
+% In this paper we represent variables and numeric constants
+% as leaves and represent their names or values as
+% arguments to those constructors (e.g., \texttt{2} in \texttt{num(2)}).
+% (In TODO:REF, we discuss how to allow edits within these.)
+
+\subsection{Single User Edits}
+\label{sub:Single User Edits}
+
+TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
+
+TODO: remove cursor from Fig 1 and Fig 2
+
+(though as we show in Section~\ref{sec:formalism} creating a vertex is implicit in edge creation)
+
+Continuity of examples: Add -> Change -> Wrap -> Move
+
+\subsubsection{Add}
+TODO: explain add to the right
+
+TODO: continuity of examples
+
+\subsubsection{Change}
 
@@
 TODO: explain that delete overrides add
+
+TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
 
@@
 NEEDS UI EXAMPLE
+
+\footnote{For the purposes of this paper, we represent identifiers and number literals
+indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
+
+TODO: 2 -> 22
 
@@
 \subsection{Collaborative Structure Editing}
+\label{sub:Collaborative Structure Editing}
 
@@
 \centering
-lam x. app(\#1, \hole)
+TODO: lam x. app(\#1, \hole)
 \caption{Single-User Cycles}
@@
 whether they create a cycle and thus the system
+can immediately warn the user,
 multi-user edits may not manifest cycles until after
 the edits are merged.
-there are much more practical reasons why cycles
-may appear when multiple users are editing part of the code.
 
@@
 and represents the code \texttt{(1 * \hole) * (2 * \hole)}.
-The left graph shows
+In the left graph, Alice moves vertex~TODO to the \texttt{right}
+index of vertex~TODO.
+The resulting graph represents the code \texttt{(1 * (2 * \hole)) * \hole}.
+In the right graph, Bob moves vertex~TODO to the \texttt{right} index of vertex~TODO.
+The resulting graph represents the code \texttt{\hole * (2 * (1 * \hole))}.
+On their own, neither edit creates a cycle.
+However, when these edits are merged,
+not only is a cycle created, but
+the vertexes in the cycle
+
+Can also be detected due to neigther original leaving them disconnected.
 

### 2020-08-13 grove-paper.tex
Rough draft of "Add" section and TikZ figures for "Add" and "Change"

@@
 
+TODO: Screenshot
+
 \begin{figure}
@@
 \label{fig:A simple tree with a hole}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[tree layout] {
+ root[as={\textbullet}] -> {
+  +[as={$\texttt{+}_1$}] -> [edge label={left}] {
+   22[as={$\texttt{num(22)}_2$}],
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{A simple tree with a hole (Alternate figure)}
+\label{fig:A simple tree with a hole (Alternate figure)}
 \end{figure}
@@
 
+Now that we have discussed adding holes
+
+TODO: num(22) -> var(x)
+
 \subsubsection{Add}
-TODO: explain add to the right
-
-TODO: continuity of examples
+
+\tikzset{graphs/graph style/.style={
+ tree layout,
+ level sep=1cm,
+ nodes={draw, rounded corners=1mm},
+ edges={draw, sloped, font={\small}}
+}}
+\tikzset{graphs/root style/.style={
+ draw=none,
+ as={\textbullet}
+}}
+%\tikzset{graphs/every edge/.style={}
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  +[as={$\texttt{+}_1$}] -> {
+   x [> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_2$}],
+   y [> edge label={right}, > edge label'={3}, as={$\texttt{var(y)}_3$}]
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Addition}
+\label{fig:Addition}
+\end{figure}
+
+When users insert code into a hole, our graph model simply adds a corresponding a vertex and edge.
+For example, suppose the user adds \texttt{y} as the right child of vertex~1 in Figure~\ref{fig:A simple tree with a hole (Alternate figure)}.
+This corresponds to creating a vertex for the variable reference \texttt{y} and adding an edge
+to that vertex from vertex~1.
+Figure~\ref{fig:Addition} shows the result, where the new vertex has identifier~3.
+
+Note that adding an edge implicitly creates.
 
@@
 
+In order to change an existing vertex, we delete the existing edge to that vertex and add the new value.
+For example, in
+
+Note that while we
+Note that the vertex continues to exist (this is essential for further manipulation)
+Once created a vertex is never deleted.
+
 TODO: explain that multiple delete or add are idempotent
@@
 TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
+
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  +[as={$\texttt{+}_1$}] -> {
+   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
+  }
+ }
+};
+\path (1cm,0cm) graph[graph style] {
+ y [as={$\texttt{var(y)}_3$}]
+};
+\end{tikzpicture}
+\caption{TODO}
+\label{fig:Change-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  +[as={$\texttt{+}_1$}] -> {
+   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
+   z [> edge label={right}, as={$\texttt{var(z)}_4$}]
+  }
+ }
+};
+\path (1cm,0cm) graph[graph style] {
+ y [as={$\texttt{var(y)}_3$}]
+};
+\end{tikzpicture}
+\caption{TODO}
+\label{fig:Change-b}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Change}
+\label{fig:Change}
+\end{figure}
 
@@
 
+\subsection{Cursors}
+\label{sub:Cursors}
+
+TODO: cursors are informational
+
 \section{Formalism}

### 2020-08-14 grove-paper.tex
Work on Figures for "Addition" and "Change" section

@@
 }
+
+\tikzset{graphs/graph style/.style={
+ tree layout,
+ level sep=1cm,
+ nodes={draw, rounded corners=1mm},
+ edges={draw, sloped, font={\small}}
+}}
+\tikzset{graphs/root style/.style={
+ draw=none,
+ as={\textbullet}
+}}
+%\tikzset{graphs/every edge/.style={}
 
@@
 
-TODO: Screenshot
-
-\begin{figure}
-\centering
-TODO: UI
-
-\includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
-\caption{A simple tree with a hole}
-\label{fig:A simple tree with a hole}
-\end{figure}
-
-\begin{figure}
+% \begin{figure}
+% \centering
+% TODO: UI
+% \includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
+% \caption{A simple tree with a hole}
+% \label{fig:A simple tree with a hole}
+% \end{figure}
+
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+TODO
+\caption{User interface}
+\label{fig:A simple tree with a hole-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
 \centering
 \begin{tikzpicture}
-\path (0cm,0cm) graph[tree layout] {
- root[as={\textbullet}] -> {
-  +[as={$\texttt{+}_1$}] -> [edge label={left}] {
-   22[as={$\texttt{num(22)}_2$}],
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  +[as={$\texttt{+}_1$}] -> [] {
+   22[> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_3$}],
    {}
@@
 \end{tikzpicture}
-\caption{A simple tree with a hole (Alternate figure)}
-\label{fig:A simple tree with a hole (Alternate figure)}
+\caption{Graph representation}
+\label{fig:A simple tree with a hole-b}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{A simple tree with a hole}
+\label{fig:A simple tree with a hole}
 \end{figure}
@@
 The editor state for each replica in Grove consists of an expression with holes.
-For example, Figure~TODO:1(a) shows the simple arithmetic expression \texttt{2 + \hole}, which
+For example, Figure~\ref{fig:A simple tree with a hole} shows the simple arithmetic expression \texttt{2 + \hole}, which
 has a hole on the right side of the plus operator.
@@
 
+For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
+
 Each edge is labeled with a globally unique identifier (e.g.,~4 and~7 in the figure) and a child position.
@@
 
-TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
-
-TODO: remove cursor from Fig 1 and Fig 2
+% TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
+
+% TODO: remove cursor from Fig 1 and Fig 2
 
@@
 
-Continuity of examples: Add -> Change -> Wrap -> Move
-
 Now that we have discussed adding holes
 
-TODO: num(22) -> var(x)
-
 \subsubsection{Add}
 
-\tikzset{graphs/graph style/.style={
- tree layout,
- level sep=1cm,
- nodes={draw, rounded corners=1mm},
- edges={draw, sloped, font={\small}}
-}}
-\tikzset{graphs/root style/.style={
- draw=none,
- as={\textbullet}
-}}
-%\tikzset{graphs/every edge/.style={}
-\begin{figure}
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+TODO
+\caption{User interface}
+\label{fig:Addition-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
 \centering
@@
 \end{tikzpicture}
+\caption{Graph representation}
+\label{fig:Addition-b}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Addition}
@@
 \end{figure}
+
+We model edits to the code as sequences of additions and deletions of edges to this graph.
+For example, consider the graph in Figure~\ref{fig:Delete and Add}, which
+represents the expression \texttt{1 + 2}.
+Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
+becomes \texttt{1 + 3}.
 
@@
 };
-\path (1cm,0cm) graph[graph style] {
+\path (2cm,-2cm) graph[graph style] {
  y [as={$\texttt{var(y)}_3$}]
@@
 
-TODO: 2 -> 22
-
-We model edits to the code as sequences of deletions and additions of edges to this graph.
-For example, consider the first graph in Figure~\ref{fig:Delete and Add}, which
-represents the expression \texttt{1 + 2}.
-Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
-becomes \texttt{1 + 3}.
 We model this by first deleting edge~7.
@@
 remove the edge pointing to it.
-
-\subsubsection{Move}
-
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
-\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
-\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
-\end{tikzpicture}
-\caption{Move}
-\label{fig:Move}
-\end{figure}
-
-An interesting consequence of our design is that a move operation
-does not have to delete code from one position and create a copy
-in another position.
-Instead, by deleting just the edge to the code to be moved and then
-adding a new edge to that code somewhere else, the actual code is
-moved, not just a copy.
-
-For example, consider Figure~\ref{fig:Move}, which represents
-the movement of \texttt{1 * 2} from the left child of \texttt{plus}
-to its right child.  The top graph is the code before the move.
-First, edge~7 is deleted, which results in the middle graph
-where the rooted code is now \texttt{\hole{} + \hole},
-and vertex~6 is orphaned.
-
-Next edge~16 is added, which results in the bottom graph
-where the rooted code is now \texttt{\hole{} + 1 * 2},
-and vertex~6 is no longer orphaned.
-
-This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-Of course, if the user wants a genuine copy
-and not just a move, that can be achieved
-by creating new vertexes with fresh identifiers.
-(TODO: figure for this?)
-
-This is similar to the paste with formatting versus
-paste without formatting.
-
-If only one restore, then it does what you expect (basically a cut and paste)
-
-If multiple replicas restore to different places, then you get some complexity:
 
@@
 Traditional: no relation (and indent might change)
+
+\subsubsection{Move}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
+\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
+\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
+\end{tikzpicture}
+\caption{Move}
+\label{fig:Move}
+\end{figure}
+
+An interesting consequence of our design is that a move operation
+does not have to delete code from one position and create a copy
+in another position.
+Instead, by deleting just the edge to the code to be moved and then
+adding a new edge to that code somewhere else, the actual code is
+moved, not just a copy.
+
+For example, consider Figure~\ref{fig:Move}, which represents
+the movement of \texttt{1 * 2} from the left child of \texttt{plus}
+to its right child.  The top graph is the code before the move.
+First, edge~7 is deleted, which results in the middle graph
+where the rooted code is now \texttt{\hole{} + \hole},
+and vertex~6 is orphaned.
+
+Next edge~16 is added, which results in the bottom graph
+where the rooted code is now \texttt{\hole{} + 1 * 2},
+and vertex~6 is no longer orphaned.
+
+This move semantics gives us a richer structure than
+when treating code as a list of lines.  We exploit
+this in Section~REF:TODO in order merge edits that involve moving code.
+
+Of course, if the user wants a genuine copy
+and not just a move, that can be achieved
+by creating new vertexes with fresh identifiers.
+(TODO: figure for this?)
+
+This is similar to the paste with formatting versus
+paste without formatting.
+
+If only one restore, then it does what you expect (basically a cut and paste)
+
+If multiple replicas restore to different places, then you get some complexity:
 

### 2020-08-14 grove-paper.tex
More work on "Add" and "Change" sections and some TikZ improvements

@@
   graphdrawing, % Graph *layout*
+  quotes,       % Quote syntax (e.g., "foo")
 }
@@
  root[root style] -> {
-  +[as={$\texttt{+}_1$}] -> [] {
-   22[> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_3$}],
+  +[as={$\texttt{+}_1$}] -> {
+   "$\texttt{var(x)}_3$" [> "left", > "2"'],
    {}
@@
 
-Now that we have discussed adding holes
-
-\subsubsection{Add}
+\subsubsection{Additions}
 
@@
   +[as={$\texttt{+}_1$}] -> {
-   x [> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_2$}],
-   y [> edge label={right}, > edge label'={3}, as={$\texttt{var(y)}_3$}]
+   "$\texttt{var(x)}_3$" [> "left", > "2"'],
+   "$\texttt{var(y)}_5$" [> "right", > "4"']
   }
@@
 We model edits to the code as sequences of additions and deletions of edges to this graph.
-For example, consider the graph in Figure~\ref{fig:Delete and Add}, which
-represents the expression \texttt{1 + 2}.
-Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
-becomes \texttt{1 + 3}.
-
-When users insert code into a hole, our graph model simply adds a corresponding a vertex and edge.
-For example, suppose the user adds \texttt{y} as the right child of vertex~1 in Figure~\ref{fig:A simple tree with a hole (Alternate figure)}.
+For example, suppose the user adds \texttt{y} as the right child of vertex~1 in Figure~\ref{fig:A simple tree with a hole (Alternate figure)}
+so that the expression becomes \texttt{x + y}.
 This corresponds to creating a vertex for the variable reference \texttt{y} and adding an edge
@@
 
-Note that adding an edge implicitly creates.
-
-\subsubsection{Change}
-
-In order to change an existing vertex, we delete the existing edge to that vertex and add the new value.
-For example, in
-
-Note that while we
-Note that the vertex continues to exist (this is essential for further manipulation)
-Once created a vertex is never deleted.
-
-TODO: explain that multiple delete or add are idempotent
-
-TODO: explain that delete overrides add
-
-TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
+TODO: Note that adding an edge implicitly creates.
+
+For the purposes of this paper, we represent identifiers and number literals
+indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.
+
+\subsubsection{Changes}
 
@@
   +[as={$\texttt{+}_1$}] -> {
-   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
+   "$\texttt{var(x)}_3$" [> "left", > "2"'],
+   {}
   }
@@
   +[as={$\texttt{+}_1$}] -> {
-   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
-   z [> edge label={right}, as={$\texttt{var(z)}_4$}]
+   "$\texttt{var(x)}_3$" [> "left", > "2"'],
+   "$\texttt{var(z)}_7$" [> "right", > "6"']
   }
@@
 \path (2cm,-2cm) graph[graph style] {
- y [as={$\texttt{var(y)}_3$}]
+ "$\texttt{var(y)}_5$" []
 };
@@
 
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/delete-and-add/1.dot}};
-\path (0,-4cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/2.dot}};
-\path (0,-8cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/3.dot}};
-\end{tikzpicture}
-\caption{Delete and Add}
-\label{fig:Delete and Add}
-\end{figure}
-NEEDS UI EXAMPLE
-
-\footnote{For the purposes of this paper, we represent identifiers and number literals
-indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-
-We model this by first deleting edge~7.
-This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
-represents the expression \texttt{1 + \hole}.
-Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
-Vertex~9 is just not reachable from the root.
-
-After that deletion, we then add a new vertex to represent~\texttt{3} and an
-edge to it.  This results in the bottom graph in Figure~\ref{fig:Delete and Add}.
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}
+% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/delete-and-add/1.dot}};
+% \path (0,-4cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/2.dot}};
+% \path (0,-8cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/3.dot}};
+% \end{tikzpicture}
+% \caption{Delete and Add}
+% \label{fig:Delete and Add}
+% \end{figure}
+
+Next suppose we want to replace \texttt{y} with \texttt{z} so that the code becomes \texttt{x + z}.
+This is accomplished by first deleting edge~4.
+This results in Figure~\ref{fig:Change-a}, which represents the expression \texttt{x + \hole{}}.
+After that deletion, we add a new vertex to represent~\texttt{z} and an
+edge from vertex~1 to that vertex.
+This results in the bottom graph in Figure~\ref{fig:Change-b},
+which represents the expression \texttt{x + z}.
 
@@
 An edge once deleted cannot be restored.  Instead a new edge (with a new id) is created.
+
+Note that vertex~3 continues to exist, and if it had any children, those children could remain connected to it.
+In our system, once a vertex is created it is never deleted.
+This allows further manipulations of those vertexes as shown later in this section.
+
+TODO: explain that multiple delete or add are idempotent
+
+TODO: explain that delete overrides add
+
+TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
 
@@
 \subsubsection{Wrapping Code}
-
-TODO: move subsubsection elsewhere?
 

### 2020-08-14 grove-paper.tex
Establish continuity in the single-user figures

@@
 \tikzset{graphs/root style/.style={
- draw=none,
- as={\textbullet}
+ %draw=none,
+ as={\textbullet$_0$}
 }}
@@
  root[root style] -> {
-  +[as={$\texttt{+}_1$}] -> {
-   "$\texttt{var(x)}_3$" [> "left", > "2"'],
+  "$\texttt{*}_2$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_4$" [> "left", > "3"'],
    {}
@@
 The editor state for each replica in Grove consists of an expression with holes.
-For example, Figure~\ref{fig:A simple tree with a hole} shows the simple arithmetic expression \texttt{2 + \hole}, which
+For example, Figure~\ref{fig:A simple tree with a hole} shows the simple arithmetic expression \texttt{x * \hole}, which
 has a hole on the right side of the plus operator.
@@
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In this figure, vertices~3 and~6 have the constructors \texttt{plus} and \texttt{num(2)}, respectively.
+In this figure, vertices~2 and~4 have the constructors \texttt{times} and \texttt{var(x)}, respectively.
 Each term constructor has an associated set of child positions.
-The \texttt{plus} constructor has positions for \texttt{left} and \texttt{right} children
-and the \texttt{num(n)} constructors have no child positions.
+The \texttt{*} constructor has positions for \texttt{left} and \texttt{right} children
+and the \texttt{var} constructors have no child positions but takes an argument naming the particular variable referenced
+(e.g., \texttt{var(x)} for the variable \texttt{x}).
 The root vertex has a distinguished identifier and a constructor with a single child position.
@@
 
-Each edge is labeled with a globally unique identifier (e.g.,~4 and~7 in the figure) and a child position.
+Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position.
 An edge indicates that the destination vertex is a child of the origin vertex at the given position.
@@
  root[root style] -> {
-  +[as={$\texttt{+}_1$}] -> {
-   "$\texttt{var(x)}_3$" [> "left", > "2"'],
-   "$\texttt{var(y)}_5$" [> "right", > "4"']
+  "$\texttt{*}_2$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_4$" [> "left", > "3"'],
+   "$\texttt{var(y)}_6$" [> "right", > "5"']
   }
@@
 \end{tikzpicture}
-\caption{Graph representation}
+\caption{Figure~\ref{fig:Addition-b} after adding vertex~6 and edge~5}
 \label{fig:Addition-b}
@@
 We model edits to the code as sequences of additions and deletions of edges to this graph.
-For example, suppose the user adds \texttt{y} as the right child of vertex~1 in Figure~\ref{fig:A simple tree with a hole (Alternate figure)}
-so that the expression becomes \texttt{x + y}.
+For example, suppose the user adds \texttt{y} as the right child of vertex~2 in
+Figure~\ref{fig:A simple tree with a hole}
+so that the expression becomes \texttt{x * y}.
 This corresponds to creating a vertex for the variable reference \texttt{y} and adding an edge
-to that vertex from vertex~1.
-Figure~\ref{fig:Addition} shows the result, where the new vertex has identifier~3.
+to that vertex from vertex~2.
+Figure~\ref{fig:Addition} shows the result, where the new vertex has identifier~6.
 
@@
 \path (2cm,-2cm) graph[graph style] {
- y [as={$\texttt{var(y)}_3$}]
+ y [as={$\texttt{var(y)}_6$}]
 };
 \end{tikzpicture}
-\caption{TODO}
+\caption{Figure~\ref{fig:Addition-b} after deleting edge~5}
 \label{fig:Change-a}
@@
  root[root style] -> {
-  +[as={$\texttt{+}_1$}] -> {
-   "$\texttt{var(x)}_3$" [> "left", > "2"'],
-   "$\texttt{var(z)}_7$" [> "right", > "6"']
+  "$\texttt{*}_2$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_4$" [> "left", > "3"'],
+   "$\texttt{var(z)}_8$" [> "right", > "7"']
   }
@@
 \path (2cm,-2cm) graph[graph style] {
- "$\texttt{var(y)}_5$" []
+ "$\texttt{var(y)}_6$"
 };
 \end{tikzpicture}
-\caption{TODO}
+\caption{Figure~\ref{fig:Change-a} after adding vertex~8 and edge~7}
 \label{fig:Change-b}
@@
 
-Next suppose we want to replace \texttt{y} with \texttt{z} so that the code becomes \texttt{x + z}.
-This is accomplished by first deleting edge~4.
-This results in Figure~\ref{fig:Change-a}, which represents the expression \texttt{x + \hole{}}.
+Next suppose we want to replace \texttt{y} with \texttt{z} so that the code becomes \texttt{x * z}.
+This is accomplished by first deleting edge~5.
+This results in Figure~\ref{fig:Change-a}, which represents the expression \texttt{x * \hole{}}.
 After that deletion, we add a new vertex to represent~\texttt{z} and an
-edge from vertex~1 to that vertex.
+edge from vertex~2 to that vertex.
 This results in the bottom graph in Figure~\ref{fig:Change-b},
-which represents the expression \texttt{x + z}.
+which represents the expression \texttt{x * z}.
 
@@
 \begin{figure}
-\centering
-TODO: naive add that creates a conflict
+% TODO: naive add that creates a conflict
+% \begin{tikzpicture}
+% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
+% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
+% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
+% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
+% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
+% \end{tikzpicture}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
 \begin{tikzpicture}
-\path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
-\path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
-\path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
-\path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
-\path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
+\path (0cm,0cm) graph[graph style] {
+ root[root style]
+};
+\path (-1cm,-3cm) graph[graph style] {
+ "$\texttt{*}_2$" -> {
+  "$\texttt{var(x)}_4$" [> "left", > "3"'],
+  "$\texttt{var(z)}_6$" [> "right", > "5"']
+ }
+};
 \end{tikzpicture}
+\caption{Figure~\ref{fig:Change-b} after deleting edge~1}
+\label{fig:Wrapping-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_8$" [> "root", > "7"']
+ }
+};
+\path (-1cm,-3cm) graph[graph style] {
+ "$\texttt{*}_2$" -> {
+  "$\texttt{var(x)}_4$" [> "left", > "3"'],
+  "$\texttt{var(z)}_6$" [> "right", > "5"']
+ }
+};
+\end{tikzpicture}
+\caption{Figure~\ref{fig:Wrapping-a} after adding vertex~8 and edge~7}
+\label{fig:Wrapping-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_8$" [> "root", > "7"'] -> {
+   "$\texttt{*}_2$" [> "left", > "9"'] -> {
+    "$\texttt{var(x)}_4$" [> "left", > "3"'],
+    "$\texttt{var(z)}_6$" [> "right", > "5"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Figure~\ref{fig:Wrapping-b} after adding edge~9}
+\label{fig:Wrapping-c}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Wrapping}
@@
 \end{figure}
+
+TODO: dotted lines for deleted edges
+
+TODO: mention that naive add that creates a multi-child is discussed later
 
@@
 \begin{figure}
+% \centering
+% \begin{tikzpicture}
+% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
+% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
+% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
+% \end{tikzpicture}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
 \centering
 \begin{tikzpicture}
-\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
-\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
-\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_8$" [> "root", > "7"']
+ }
+};
+\path (0cm,-3cm) graph [graph style] {
+ "$\texttt{*}_2$" -> {
+  "$\texttt{var(x)}_4$" [> "left", > "3"'],
+  "$\texttt{var(z)}_6$" [> "right", > "5"']
+ }
+};
 \end{tikzpicture}
-\caption{Move}
+\caption{Figure~\ref{fig:Wrapping-c} after deleting edge~9}
+\label{fig:Move-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_8$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{*}_2$" [> "right", > "11"'] -> {
+    "$\texttt{var(x)}_4$" [> "left", > "3"'],
+    "$\texttt{var(z)}_6$" [> "right", > "5"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Figure~\ref{fig:Move-a} after adding edge~11}
+\label{fig:Move-b}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Moving vertex~2 from the left child to the right child of vertex~8}
 \label{fig:Move}

### 2020-08-18 grove-paper.tex
Use \autoref and macros \edge and \vertex

@@
 
+%%%%%%%%
+
 \renewcommand{\topfraction}{1} % Allow floats to take up the page
 \renewcommand{\textfraction}{0}
+
+%%%%%%%%
+% \autoref from hyperref
+\renewcommand{\AMSautorefname}          {Equation}
+\renewcommand{\appendixautorefname}     {Appendix}
+\renewcommand{\chapterautorefname}      {Chapter}
+\renewcommand{\equationautorefname}     {Equation}
+\renewcommand{\FancyVerbLineautorefname}{Line}
+\renewcommand{\figureautorefname}       {Figure}
+\renewcommand{\footnoteautorefname}     {Footnote}
+\renewcommand{\Hfootnoteautorefname}    {Footnote}
+\renewcommand{\itemautorefname}         {Item}
+\renewcommand{\Itemautorefname}         {Item}
+\renewcommand{\pageautorefname}         {Page}
+\renewcommand{\paragraphautorefname}    {Section}
+\renewcommand{\partautorefname}         {Part}
+\renewcommand{\sectionautorefname}      {Section}
+\renewcommand{\subparagraphautorefname} {Section}
+\renewcommand{\subsectionautorefname}   {Section}
+\renewcommand{\subsubsectionautorefname}{Section}
+\renewcommand{\tableautorefname}        {Table}
+\renewcommand{\theoremautorefname}      {Theorem}
 
@@
 \usepackage{graphicx}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
 
@@
 \newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
+
+% Support \includegraphics of .dot files
+\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
+
+\newcommand{\edge}[1]{Edge~#1}
+\newcommand{\vertex}[1]{Vertex~#1}
 
@@
 This section introduces collaborative structure editing in Grove by example. 
-We begin in Section~\ref{sub:Program Representation} by introducing basic edit actions
+We begin in \autoref{sub:Program Representation} by introducing basic edit actions
 and the underlying graph representation of expressions with holes.
 We will then discuss scenarios that arise in the collaborative setting in 
-Section~\ref{sub:Collaborative Structure Editing}.
-
-TODO: gray-scale unchanged parts of diagrams
+\autoref{sub:Collaborative Structure Editing}.
+
+TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts?
 
@@
 The editor state for each replica in Grove consists of an expression with holes.
-For example, Figure~\ref{fig:A simple tree with a hole} shows the simple arithmetic expression \texttt{x * \hole}, which
+For example, \autoref{fig:A simple tree with a hole} shows the simple arithmetic expression~\texttt{x * \hole}, which
 has a hole on the right side of the plus operator.
@@
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In this figure, vertices~2 and~4 have the constructors \texttt{times} and \texttt{var(x)}, respectively.
+In this figure, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.
 Each term constructor has an associated set of child positions.
-The \texttt{*} constructor has positions for \texttt{left} and \texttt{right} children
-and the \texttt{var} constructors have no child positions but takes an argument naming the particular variable referenced
-(e.g., \texttt{var(x)} for the variable \texttt{x}).
+The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
+and the~\texttt{var} constructors have no child positions but takes an argument naming the particular variable referenced
+(e.g.,~\texttt{var(x)} for the variable~\texttt{x}).
 The root vertex has a distinguished identifier and a constructor with a single child position.
@@
 % as leaves and represent their names or values as
-% arguments to those constructors (e.g., \texttt{2} in \texttt{num(2)}).
+% arguments to those constructors (e.g.,~\texttt{2} in~\texttt{num(2)}).
 % (In TODO:REF, we discuss how to allow edits within these.)
@@
 
-(though as we show in Section~\ref{sec:formalism} creating a vertex is implicit in edge creation)
+(though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)
 
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Addition-b} after adding vertex~6 and edge~5}
+\caption{\autoref{fig:Addition-b} after adding \vertex{6} and \edge{5}}
 \label{fig:Addition-b}
@@
 We model edits to the code as sequences of additions and deletions of edges to this graph.
-For example, suppose the user adds \texttt{y} as the right child of vertex~2 in
-Figure~\ref{fig:A simple tree with a hole}
+For example, suppose the user adds~\texttt{y} as the right child of \vertex{2} in
+\autoref{fig:A simple tree with a hole}
 so that the expression becomes \texttt{x * y}.
-This corresponds to creating a vertex for the variable reference \texttt{y} and adding an edge
-to that vertex from vertex~2.
-Figure~\ref{fig:Addition} shows the result, where the new vertex has identifier~6.
+This corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
+to that vertex from \vertex{2}.
+\autoref{fig:Addition} shows the result, where the new vertex has identifier~6.
 
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Addition-b} after deleting edge~5}
+\caption{\autoref{fig:Addition-b} after deleting \edge{5}}
 \label{fig:Change-a}
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Change-a} after adding vertex~8 and edge~7}
+\caption{\autoref{fig:Change-a} after adding \vertex{8} and \edge{7}}
 \label{fig:Change-b}
@@
 
-Next suppose we want to replace \texttt{y} with \texttt{z} so that the code becomes \texttt{x * z}.
-This is accomplished by first deleting edge~5.
-This results in Figure~\ref{fig:Change-a}, which represents the expression \texttt{x * \hole{}}.
+Next suppose we want to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
+This is accomplished by first deleting \edge{5}.
+This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
 After that deletion, we add a new vertex to represent~\texttt{z} and an
-edge from vertex~2 to that vertex.
-This results in the bottom graph in Figure~\ref{fig:Change-b},
-which represents the expression \texttt{x * z}.
+edge from \vertex{2} to that vertex.
+This results in the bottom graph in \autoref{fig:Change-b},
+which represents the expression~\texttt{x * z}.
 
@@
 
-Note that vertex~3 continues to exist, and if it had any children, those children could remain connected to it.
+Note that \vertex{3} continues to exist, and if it had any children, those children could remain connected to it.
 In our system, once a vertex is created it is never deleted.
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Change-b} after deleting edge~1}
+\caption{\autoref{fig:Change-b} after deleting \edge{1}}
 \label{fig:Wrapping-a}
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Wrapping-a} after adding vertex~8 and edge~7}
+\caption{\autoref{fig:Wrapping-a} after adding \vertex{8} and \edge{7}}
 \label{fig:Wrapping-b}
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Wrapping-b} after adding edge~9}
+\caption{\autoref{fig:Wrapping-b} after adding \edge{9}}
 \label{fig:Wrapping-c}
@@
 
-TODO: dotted lines for deleted edges
-
 TODO: mention that naive add that creates a multi-child is discussed later
@@
 we want to point out a common idiom that we call \emph{wrapping}.
-Consider the first graph in Figure~\ref{fig:Wrapping}, which
-represents the code \texttt{$\lambda$ x . x}, and suppose
+Consider the first graph in \autoref{fig:Wrapping}, which
+represents the code~\texttt{$\lambda$ x . x}, and suppose
 the user's cursor is in the body field of the lambda (i.e., around the variable
-reference to \texttt{x} in the body of the lambda).
+reference to~\texttt{x} in the body of the lambda).
 
@@
 This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
-and results in the code \texttt{$\lambda$ x . x + \hole}.
+and results in the code~\texttt{$\lambda$ x . x + \hole}.
 (This is similar to the behavior in other structure editors such as in CITE:TODO.)
-Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
+Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.
 Note that this is a layer above
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Wrapping-c} after deleting edge~9}
+\caption{\autoref{fig:Wrapping-c} after deleting \edge{9}}
 \label{fig:Move-a}
@@
 \end{tikzpicture}
-\caption{Figure~\ref{fig:Move-a} after adding edge~11}
+\caption{\autoref{fig:Move-a} after adding \edge{11}}
 \label{fig:Move-b}
@@
 \hfill{}
-\caption{Moving vertex~2 from the left child to the right child of vertex~8}
+\caption{Moving \vertex{2} from the left child to the right child of \vertex{8}}
 \label{fig:Move}
@@
 
-For example, consider Figure~\ref{fig:Move}, which represents
-the movement of \texttt{1 * 2} from the left child of \texttt{plus}
+For example, consider \autoref{fig:Move}, which represents
+the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
 to its right child.  The top graph is the code before the move.
-First, edge~7 is deleted, which results in the middle graph
-where the rooted code is now \texttt{\hole{} + \hole},
-and vertex~6 is orphaned.
-
-Next edge~16 is added, which results in the bottom graph
-where the rooted code is now \texttt{\hole{} + 1 * 2},
-and vertex~6 is no longer orphaned.
+First, \edge{7} is deleted, which results in the middle graph
+where the rooted code is now~\texttt{\hole{} + \hole},
+and \vertex{6} is orphaned.
+
+Next \edge{16} is added, which results in the bottom graph
+where the rooted code is now~\texttt{\hole{} + 1 * 2},
+and \vertex{6} is no longer orphaned.
 
@@
 that are synchronized with each other and
-have the code \texttt{3 + 4}.
+have the code~\texttt{3 + 4}.
 That is to say, they both have
-the top graphs in Figure~\ref{fig:Editing Different Parts of the Code}.
-
-Next, suppose user A changes the \texttt{3} to a \texttt{13},
-and user B changes the \texttt{4} to a \texttt{14}.
+the top graphs in \autoref{fig:Editing Different Parts of the Code}.
+
+Next, suppose user A changes the~\texttt{3} to a~\texttt{13},
+and user B changes the~\texttt{4} to a~\texttt{14}.
 Before transmitting their edits to each other,
 user A thus has left graph and user B has the right graph.
-For user A, this involves deleting edge~7 and creating edge~14.
-For user B, this involves deleting edge~10 and creating edge~18.
+For user A, this involves deleting \edge{7} and creating \edge{14}.
+For user B, this involves deleting \edge{10} and creating \edge{18}.
 
@@
 For example, suppose that A and B did not transmit their edits to each other
-but to a third user, C, that has the original graph at the top of Figure~\ref{fig:Editing Different Parts of the Code}.
+but to a third user, C, that has the original graph at the top of \autoref{fig:Editing Different Parts of the Code}.
 User C can replay the edits from A and B in any order and will always get
-the graph at the bottom of Figure~\ref{fig:Editing Different Parts of the Code}.
+the graph at the bottom of \autoref{fig:Editing Different Parts of the Code}.
 
 Stated more formally, our edit model obeys commutativity
-as shown in Figure~\ref{fig:Commutativity}.
+as shown in \autoref{fig:Commutativity}.
 For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
@@
 
-Consider Figure~\ref{fig:Editing Nested Parts of the Code}.
-Alice and Bob both start with the top graph, which represents the code \texttt{2 * x + \hole{}}.
-Alice changes the \texttt{2} to a \texttt{3}, which results in the left graph.
-Bob moves the \texttt{2 * x} from the left of the plus to the right of the plus
-by deleting edge~7 and adding edge~16, which results in the right graph.
+Consider \autoref{fig:Editing Nested Parts of the Code}.
+Alice and Bob both start with the top graph, which represents the code~\texttt{2 * x + \hole{}}.
+Alice changes the~\texttt{2} to a~\texttt{3}, which results in the left graph.
+Bob moves the~\texttt{2 * x} from the left of the plus to the right of the plus
+by deleting \edge{7} and adding \edge{16}, which results in the right graph.
 Since vertex and edge identifiers are preserved by these,
@@
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of edge~7 before making her
-edits and before receiving Bob's adding of edge~16.
+Alice might receive Bob's deletion of \edge{7} before making her
+edits and before receiving Bob's adding of \edge{16}.
 Her edits would then be within the code deleted by Bob.
@@
 Finally, we address genuine conflicts between edits.
-Consider the top graph of Figure~\ref{fig:Direct Conflicts}.
-Suppose Alice inserts the number literal \texttt{2} into the body of the lambda
-while Bob inserts the function application \texttt{f x} also into the body of the lambda.
-These results in the left and right graphs of \ref{fig:Direct Conflicts} respectively.
-In particular pay attention to the edges going out of the \texttt{body} index of node~3.
-In the left graph it is edge~10, which points to a number literal,
-and in the right graph it is edge~13, which points to a function application.
+Consider the top graph of \autoref{fig:Direct Conflicts}.
+Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
+while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
+These results in the left and right graphs of \autoref{fig:Direct Conflicts} respectively.
+In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
+In the left graph it is \edge{10}, which points to a number literal,
+and in the right graph it is \edge{13}, which points to a function application.
 
 We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
-of node~3.
-This results in the bottom graph of Figure~\ref{fig:Direct Conflicts}.
+of \vertex{3}.
+This results in the bottom graph of \autoref{fig:Direct Conflicts}.
 The conflict between Alice's and Bob's changes is
 represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
-index of node~3).
+index of \vertex{3}).
 In the graph representation, we flag this conflict by highlighting its edges
 in red.
-In the textual representation, we flag this conflict with the notation \texttt{\{ 2 | f x \}}.
+In the textual representation, we flag this conflict with the notation~\texttt{\{ 2 | f x \}}.
 This is a conflict that much eventually be resolved by the user by deleting
 one of the edges and editing the other to be the merged value.
-For example, one could delete edge~10 and then edit the \texttt{x}
-to be \texttt{x + 2}.
+For example, one could delete \edge{10} and then edit the~\texttt{x}
+to be~\texttt{x + 2}.
 Note that users can continue to edit the code while in this conflicted state.
@@
 line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of \texttt{<}, \texttt{>}, or \texttt{=})
+Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
 when a merge conflict happens.
@@
 
-An example of this is shown in Figure~\ref{fig:Multi-parent conflict},
-which starts with the top graph, which represents the expression \texttt{\hole + (\hole + x)}.
-Suppose Alice moves the variable \texttt{x} to the left-most hole.
-This results in the left graph, which represents the expression \texttt{x + (\hole + \hole)}.
-At the same time, suppose Bob moves the variable \texttt{x} to the right-most hole.
-This results in the right graph, which represents the expression \texttt{\hole + (x + \hole)}.
+An example of this is shown in \autoref{fig:Multi-parent conflict},
+which starts with the top graph, which represents the expression~\texttt{\hole + (\hole + x)}.
+Suppose Alice moves the variable~\texttt{x} to the left-most hole.
+This results in the left graph, which represents the expression~\texttt{x + (\hole + \hole)}.
+At the same time, suppose Bob moves the variable~\texttt{x} to the right-most hole.
+This results in the right graph, which represents the expression~\texttt{\hole + (x + \hole)}.
 In both cases, these edits are achieved with one edge delete and one edge add.
@@
 a vertex as a child of itself.
-An example of this is shown in Figure~\ref{fig:Single-User Cycles},
+An example of this is shown in \autoref{fig:Single-User Cycles},
 where the user has pasted vertex~TODO as a child of vertex~TODO.
@@
 
-For example, consider the situation shown in Figure~\ref{fig:Multi-User Cycles}.
+For example, consider the situation shown in \autoref{fig:Multi-User Cycles}.
 The top graph represents code that is synchronized between Alice and Bob
-and represents the code \texttt{(1 * \hole) * (2 * \hole)}.
+and represents the code~\texttt{(1 * \hole) * (2 * \hole)}.
 In the left graph, Alice moves vertex~TODO to the \texttt{right}
 index of vertex~TODO.
-The resulting graph represents the code \texttt{(1 * (2 * \hole)) * \hole}.
+The resulting graph represents the code~\texttt{(1 * (2 * \hole)) * \hole}.
 In the right graph, Bob moves vertex~TODO to the \texttt{right} index of vertex~TODO.
-The resulting graph represents the code \texttt{\hole * (2 * (1 * \hole))}.
+The resulting graph represents the code~\texttt{\hole * (2 * (1 * \hole))}.
 On their own, neither edit creates a cycle.
@@
 
-The language we are modeling is shown in Figure~\ref{fig:Syntax as a grammar}.
+The language we are modeling is shown in \autoref{fig:Syntax as a grammar}.
 Standard for a simply-typed lambda calculus except that

### 2020-08-18 grove-paper.tex
Light editing pass on sections 2.1 and 2.2.1-2.2.3

@@
  nodes={draw, rounded corners=1mm},
- edges={draw, sloped, font={\small}}
+ edges={draw, sloped, font={\small\ttfamily}}
 }}
@@
 
+TODO: Alice and Bob and narrative continuity
+
 This section introduces collaborative structure editing in Grove by example. 
@@
 
+TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
+
+We present this at two levels.
 One is the underlying graph model, which
@@
 operations.
-
-We present this at two levels.
 The other is the interface a user might see,
@@
 as a graph with.
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
 
@@
 
-The editor state for each replica in Grove consists of an expression with holes.
-For example, \autoref{fig:A simple tree with a hole} shows the simple arithmetic expression~\texttt{x * \hole}, which
-has a hole on the right side of the plus operator.
+The state for each editor instance in Grove consists of an expression containing holes along with a cursor.
+For example, \autoref{fig:A simple tree with a hole} shows the arithmetic expression~\texttt{x * \hole}, which
+has a hole on the right side of the times operator.
 
 Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
+For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In this figure, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.
+In \autoref{fig:A simple tree with a hole}, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.
+
 Each term constructor has an associated set of child positions.
 The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular variable referenced
-(e.g.,~\texttt{var(x)} for the variable~\texttt{x}).
-The root vertex has a distinguished identifier and a constructor with a single child position.
-For simplicity, we visually represent the root with a dot.
-
+and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
+variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).
+The root vertex has a distinguished identifier~0 and the constructor \textbullet with the single child position \texttt{root}.
+
+Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position (e.g., \texttt{left} and \texttt{root} in the figure).
+An edge indicates that the destination vertex is a child of the origin vertex at the given position.
+%Visually, we indicate the position of an edge by the location of its origin.
 For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
 
-Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position.
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-Visually, we indicate the position of an edge by the location of its origin.
-For example, Figure~TODO:1(b) shows the graph corresponding to the term in Figure~TODO:1(a).
-
-TODO: holes
+Finally, holes are represented by the absence of a child.
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vertex{2}
+corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
 
@@
 
+Note that, for the purposes of this paper, we represent identifiers and number literals
+indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.
+
 \subsection{Single User Edits}
@@
 % TODO: remove cursor from Fig 1 and Fig 2
-
-(though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)
 
@@
 
-TODO: Note that adding an edge implicitly creates.
-
-For the purposes of this paper, we represent identifiers and number literals
-indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.
+TODO: (though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)
 
@@
 
-Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
-we want to point out a common idiom that we call \emph{wrapping}.
-Consider the first graph in \autoref{fig:Wrapping}, which
-represents the code~\texttt{$\lambda$ x . x}, and suppose
-the user's cursor is in the body field of the lambda (i.e., around the variable
-reference to~\texttt{x} in the body of the lambda).
+Next suppose a user inserts cod at a position already occupied by some code.
+For example, suppose, the user types~\texttt{+} when the cursor is on \vertex{2}.
+If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
+as a child of \vertex{0}.
 
@@
 
-Traditional: no relation (and indent might change)
+Traditional diff: no relation (and indent might change)
 

### 2020-08-24 grove-paper.tex
Added some TikZ figures showing graphs

@@
  %draw=none,
- as={\textbullet$_0$}
+ as={\textbullet$_{0}$}
 }}
@@
 
+TODO: "run through several senarios though in practice these occur less frequently than portrayed here"
+
 TODO: Alice and Bob and narrative continuity
+
+TODO: group user-level edits (with user visible)
 
@@
  root[root style] -> {
-  "$\texttt{*}_2$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_4$" [> "left", > "3"'],
+  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
    {}
@@
  root[root style] -> {
-  "$\texttt{*}_2$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_4$" [> "left", > "3"'],
-   "$\texttt{var(y)}_6$" [> "right", > "5"']
-  }
- }
-};
-\end{tikzpicture}
-\caption{\autoref{fig:Addition-b} after adding \vertex{6} and \edge{5}}
+  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+   "$\texttt{var(y)}_{6}$" [> "right", > "5"']
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:A simple tree with a hole-b} after adding \vertex{6} and \edge{5}}
 \label{fig:Addition-b}
@@
 
+TODO: z -> foo (or multi-editor)
+
 \subsubsection{Changes}
@@
 \path (2cm,-2cm) graph[graph style] {
- y [as={$\texttt{var(y)}_6$}]
+ y [as={$\texttt{var(y)}_{6}$}]
 };
@@
  root[root style] -> {
-  "$\texttt{*}_2$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_4$" [> "left", > "3"'],
-   "$\texttt{var(z)}_8$" [> "right", > "7"']
+  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
+   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+   "$\texttt{var(z)}_{8}$" [> "right", > "7"']
   }
@@
 \path (2cm,-2cm) graph[graph style] {
- "$\texttt{var(y)}_6$"
+ "$\texttt{var(y)}_{6}$"
 };
@@
 \path (-1cm,-3cm) graph[graph style] {
- "$\texttt{*}_2$" -> {
-  "$\texttt{var(x)}_4$" [> "left", > "3"'],
-  "$\texttt{var(z)}_6$" [> "right", > "5"']
+ "$\texttt{*}_{2}$" -> {
+  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
  }
@@
  root[root style] -> {
-  "$\texttt{+}_8$" [> "root", > "7"']
+  "$\texttt{+}_{8}$" [> "root", > "7"']
  }
@@
  root[root style] -> {
-  "$\texttt{+}_8$" [> "root", > "7"'] -> {
-   "$\texttt{*}_2$" [> "left", > "9"'] -> {
-    "$\texttt{var(x)}_4$" [> "left", > "3"'],
-    "$\texttt{var(z)}_6$" [> "right", > "5"']
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "9"'] -> {
+    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
    },
@@
 \path (0cm,-3cm) graph [graph style] {
- "$\texttt{*}_2$" -> {
-  "$\texttt{var(x)}_4$" [> "left", > "3"'],
-  "$\texttt{var(z)}_6$" [> "right", > "5"']
+ "$\texttt{*}_{2}$" -> {
+  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
  }
@@
  root[root style] -> {
-  "$\texttt{+}_8$" [> "root", > "7"'] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
    {},
-   "$\texttt{*}_2$" [> "right", > "11"'] -> {
-    "$\texttt{var(x)}_4$" [> "left", > "3"'],
-    "$\texttt{var(z)}_6$" [> "right", > "5"']
+   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
+    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
    }
@@
 
+TODO: mention that id numbers are global
+
 \subsubsection{Editing Different Parts of the Code}
@@
 \begin{figure}
-\centering
-TODO: add screenshot of code text to each node
-TODO: remove orphans from diagram
-\begin{tikzpicture}
-\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
-\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
-\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
-\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
-\end{tikzpicture}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Move-b} after Alice replaces \vertex{4} with \vertex{12}}
+\label{fig:Editing Different Parts of the Code-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
+    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Move-b} after Bob replaces \vertex{6} with \vertex{14}}
+\label{fig:Editing Different Parts of the Code-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{The result of merging \autoref{fig:Editing Different Parts of the Code-a} with \autoref{fig:Editing Different Parts of the Code-b}}
+\label{fig:Editing Different Parts of the Code-c}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Editing Different Parts of the Code}
@@
 \end{figure}
+
+% \begin{figure}
+% \centering
+% TODO: add screenshot of code text to each node
+% TODO: remove orphans from diagram
+% \begin{tikzpicture}
+% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
+% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
+% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
+% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
+% \end{tikzpicture}
+% \caption{Editing Different Parts of the Code}
+% \label{fig:Editing Different Parts of the Code}
+% \end{figure}
 
@@
 \begin{figure}
-\centering
-TODO: add screenshot of code text to each node
-TODO: remove orphans from diagram
-TODO: annotate edges with delete and restore
-\begin{tikzpicture}
-\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
-\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
-\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
-\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
-\end{tikzpicture}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Editing Different Parts of the Code-c} after Alice replaces \vertex{14} with \vertex{16}}
+\label{fig:Editing Nested Parts of the Code-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Editing Different Parts of the Code-c} after Bob move \vertex{2} to the left child of \vertex{8}}
+\label{fig:Editing Nested Parts of the Code-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}}
+\label{fig:Editing Nested Parts of the Code-c}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Editing Nested Parts of the Code}
@@
 \end{figure}
+
+% \begin{figure}
+% \centering
+% TODO: add screenshot of code text to each node
+% TODO: remove orphans from diagram
+% TODO: annotate edges with delete and restore
+% \begin{tikzpicture}
+% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
+% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
+% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
+% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
+% \end{tikzpicture}
+% \caption{Editing Nested Parts of the Code}
+% \label{fig:Editing Nested Parts of the Code}
+% \end{figure}
 
@@
 \begin{figure}
-\centering
-\begin{tikzpicture}
-\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
-\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
-\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
-\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
-\end{tikzpicture}
-\caption{Direct Conflicts}
-\label{fig:Direct Conflicts}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   },
+   {
+    "$\texttt{var(x)}_{20}$" [> "right", > "21"']
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO}
+\label{fig:Multi-child-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   },
+   {
+    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
+     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
+     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO}
+\label{fig:Multi-child-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.3\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
+    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   },
+   {
+    "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
+    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
+     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
+     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
+\label{fig:Multi-child-c}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-child conflicts}
+\label{fig:Multi-child}
 \end{figure}
 
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}
+% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
+% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
+% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
+% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
+% \end{tikzpicture}
+% \caption{Direct Conflicts}
+% \label{fig:Direct Conflicts}
+% \end{figure}
+
 Finally, we address genuine conflicts between edits.
-Consider the top graph of \autoref{fig:Direct Conflicts}.
+Consider the top graph of \autoref{fig:Multi-child}.
 Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
 while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
-These results in the left and right graphs of \autoref{fig:Direct Conflicts} respectively.
+These results in the left and right graphs of \autoref{fig:Multi-child} respectively.
 In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
@@
 of \vertex{3}.
-This results in the bottom graph of \autoref{fig:Direct Conflicts}.
+This results in the bottom graph of \autoref{fig:Multi-child}.
 The conflict between Alice's and Bob's changes is
@@
 \begin{figure}
-\centering
-
-TODO TOP: +(\hole, +(\hole, x))
-
-TODO LEFT: +(x, +(\hole, \hole))
-
-TODO RIGHT: +(\hole, +(x, \hole))
-
-TODO BOTTOM: Two parents pointing to x
-
-\caption{Multi-parent conflict}
-\label{fig:Multi-parent conflict}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+    {}
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Multi-parent-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'],
+   "$\texttt{var(x)}_{2}$" [> "left", > "19"']
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Multi-parent-a} after Alice ... TODO}
+\label{fig:Multi-parent-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    {},
+    "$\texttt{var(x)}_{2}$" [> "right", > "19"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Multi-child-a} after Bob ... TODO}
+\label{fig:Multi-parent-c}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  a/{$\texttt{+}_{8}$} [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] ->["right", "19"'] {
+    {},
+    b/"$\texttt{var(x)}_{2}$"
+   },
+   {}
+  }
+ },
+ a ->["right", "27"'] b
+};
+\end{tikzpicture}
+\caption{The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}}
+\label{fig:Multi-parent-d}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-parent conflicts}
+\label{fig:Multi-parent}
 \end{figure}
 
-An example of this is shown in \autoref{fig:Multi-parent conflict},
+An example of this is shown in \autoref{fig:Multi-parent},
 which starts with the top graph, which represents the expression~\texttt{\hole + (\hole + x)}.
@@
 \begin{figure}
-\centering
-TODO: lam x. app(\#1, \hole)
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+    {}
+   },
+   {}
+  }
+ },
+ v2 -> v8
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Single-User Cycles}
@@
 \end{figure}
+
+% \begin{figure}
+% \centering
+% TODO: lam x. app(\#1, \hole)
+% \caption{Single-User Cycles}
+% \label{fig:Single-User Cycles}
+% \end{figure}
 
@@
 
-Can also be detected due to neigther original leaving them disconnected.
+Can also be detected due to neither original leaving them disconnected.
 
@@
 
+TODO: memory usage
+
 \subsection{Variable names, strings, and numbers}

### 2020-08-27 grove-paper.tex
Add labels to all sections.  Update the multi-user sections with Alice and Bob and references to the new figures.

@@
 \section{Introduction}
-\label{sec:introduction}
+\label{sec:Introduction}
 
@@
 \subsection{Contributions and Paper Organization}
-\label{sec:paper-organization}
+\label{sec:Contributions and Paper Organization}
 
 \section{Grove By Example}
-\label{sec:grove-by-example}
-
-TODO: "run through several senarios though in practice these occur less frequently than portrayed here"
-
-TODO: Alice and Bob and narrative continuity
+\label{sec:Grove By Example}
+
+This section introduces collaborative structure editing in Grove.
+We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
+\autoref{sub:Single User Edits} explores edits by a single user, Alice,
+and \autoref{sub:Collaborative Structure Editing} explores edits by multiple users, Alice and Bob,
+each of whom is editing their own instance of a program.
 
@@
 
-This section introduces collaborative structure editing in Grove by example. 
-We begin in \autoref{sub:Program Representation} by introducing basic edit actions
-and the underlying graph representation of expressions with holes.
-We will then discuss scenarios that arise in the collaborative setting in 
-\autoref{sub:Collaborative Structure Editing}.
-
-TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts?
-
-TODO: typewritter font in figure graphs
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-We present this at two levels.
-One is the underlying graph model, which
-involves explicit identifiers, graph edges, indexes, and add and delete
-operations.
-The other is the interface a user might see,
-which omits these details unless needed.
-Throughout these examples the graph model is presented
-as a graph with.
+TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts? describe semantic vs presentational in diagrams
 
@@
 
-The state for each editor instance in Grove consists of an expression containing holes along with a cursor.
-For example, \autoref{fig:A simple tree with a hole} shows the arithmetic expression~\texttt{x * \hole}, which
+%We represent expression contain holes as part of the editor state 
+The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
+This can be presented to the user as an expression containing holes.
+For example, \autoref{fig:A simple tree with a hole-b} shows one such graph,
+and \autoref{fig:A simple tree with a hole-a} shows its corresponding expression,~\texttt{x * \hole}, which
 has a hole on the right side of the times operator.
 
-Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
-For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
+%We present this from two perspectives.
+%One is the underlying graph model, the other is the interface a user might see, which omits these details unless needed.
+
+%which involves explicit graph edges, positions, and add and delete operations.
+
+Program state is a directed graph with a distinguished root vertex.
+%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole}, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.
+In \autoref{fig:A simple tree with a hole-b}, \vertex{2} and \vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
 
@@
 and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).
-The root vertex has a distinguished identifier~0 and the constructor \textbullet with the single child position \texttt{root}.
-
-Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position (e.g., \texttt{left} and \texttt{root} in the figure).
+variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
+for the purposes of this paper, we represent identifiers and number literals
+indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
+The root vertex constructor,~\textbullet, has the single child position \texttt{root}.
+
+Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
+a child position (e.g., \texttt{left} and \texttt{root} in \autoref{fig:A simple tree with a hole-b}).
 An edge indicates that the destination vertex is a child of the origin vertex at the given position.
@@
 
-Note that, for the purposes of this paper, we represent identifiers and number literals
-indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.
+\marginpar{START REV HERE}
 
@@
 \subsubsection{Additions}
+\label{sub:Additions}
 
@@
 
-We model edits to the code as sequences of additions and deletions of edges to this graph.
-For example, suppose the user adds~\texttt{y} as the right child of \vertex{2} in
-\autoref{fig:A simple tree with a hole}
-so that the expression becomes \texttt{x * y}.
-This corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
+TODO: Actions act at a cursor.  Each instance also has a cursor that it controls, but this is not part of the graph representation.
+
+Alice starts with the code in \autoref{fig:A simple tree with a hole},
+She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
+We model edits as sequences of additions and deletions of edges to the graph.
+Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
 to that vertex from \vertex{2}.
-\autoref{fig:Addition} shows the result, where the new vertex has identifier~6.
-
-TODO: (though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)
+\autoref{fig:Addition-b} shows the result, where the new vertex has identifier~6.
+This graph represents the expression \texttt{x * y}.
+
+TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
 
@@
 \subsubsection{Changes}
+\label{sub:Changes}
 
@@
 
-Next suppose we want to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
-This is accomplished by first deleting \edge{5}.
+Next Alice wants to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
+This is modeled by first deleting \edge{5}.
 This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
@@
 Note that this edge has a new identifier distinct from the one before.
-An edge once deleted cannot be restored.  Instead a new edge (with a new id) is created.
+In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
+Instead a new edge with a new identifiers is created.
 
@@
 This allows further manipulations of those vertexes as shown later in this section.
-
-TODO: explain that multiple delete or add are idempotent
-
-TODO: explain that delete overrides add
-
-TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
-
-Explain orphans
 
@@
 
+TODO: explain that multiple delete or add are idempotent
+
+TODO: explain that delete overrides add
+
+TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
+
+TODO: Explain orphans
+
 \subsubsection{Wrapping Code}
+\label{sub:Wrapping Code}
 
@@
 
-TODO: mention that naive add that creates a multi-child is discussed later
-
-Next suppose a user inserts cod at a position already occupied by some code.
-For example, suppose, the user types~\texttt{+} when the cursor is on \vertex{2}.
+Next, Alice types~\texttt{+} when the cursor is on \vertex{2}.
 If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
 as a child of \vertex{0}.
-
-If the user requests that a plus be inserted,
-the editor may notice that there is something already in that hole.
-A naive editor could certainly add the plus by creating a new
+However, this position is already occupied by some code.
+
+A naive editor could add the~\texttt{+} by creating a new
 edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
-This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
+This would result in a conflict between the \texttt{+} and the existing \texttt{var}
 as shown in Figure~REF:TODO.
-A better user experience
+
+How to handle multiple children in the same position is explored in Sec~REF:TODO,
+but for this sort of edit, a better user experience
 would be interpret the request for a plus to mean to
@@
 This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
-and results in the code~\texttt{$\lambda$ x . x + \hole}.
+and results in the code~\texttt{(x * z) + \hole}.
 (This is similar to the behavior in other structure editors such as in CITE:TODO.)
 Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.
-Note that this is a layer above
-
-Traditional diff: no relation (and indent might change)
+
+TODO: "Note that the logic of to insert or wrap is a layer above graph actions."
+
+TODO: Traditional diff: no relation (and indent might change)
 
 \subsubsection{Move}
+\label{sub:Move}
 
@@
 the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
-to its right child.  The top graph is the code before the move.
-First, \edge{7} is deleted, which results in the middle graph
-where the rooted code is now~\texttt{\hole{} + \hole},
-and \vertex{6} is orphaned.
-
-Next \edge{16} is added, which results in the bottom graph
+to its right child.
+\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
+First, \edge{7} is deleted, which results in \autoref{fig:Move-a}
+where the rooted code is now~\texttt{\hole{} + \hole{}},
+and \vertex{2} is orphaned.
+Next \edge{11} is added, which results in the bottom graph
 where the rooted code is now~\texttt{\hole{} + 1 * 2},
-and \vertex{6} is no longer orphaned.
+and \vertex{2} is no longer orphaned.
 
@@
 
-TODO: mention that id numbers are global
+TODO: mention that id numbers are global (or maybe use A1 and B1 for different users?)
+
+We assume multiple instances
 
 \subsubsection{Editing Different Parts of the Code}
+\label{sub:Editing Different Parts of the Code}
 
@@
 
-When editing different parts of the code
-
-For example, suppose there are users, A and B,
-that are synchronized with each other and
-have the code~\texttt{3 + 4}.
-That is to say, they both have
-the top graphs in \autoref{fig:Editing Different Parts of the Code}.
-
-Next, suppose user A changes the~\texttt{3} to a~\texttt{13},
-and user B changes the~\texttt{4} to a~\texttt{14}.
+To start, Alice and Bob are synchronized with each other and
+both have the graph in \autoref{fig:Move-b}, which
+represents the code~\texttt{\hole{} + (x + z)}.
+
+Alice changes the~\texttt{x} to a~\texttt{u},
+and Bob changes the~\texttt{z} to a~\texttt{v}.
 Before transmitting their edits to each other,
-user A thus has left graph and user B has the right graph.
-For user A, this involves deleting \edge{7} and creating \edge{14}.
-For user B, this involves deleting \edge{10} and creating \edge{18}.
-
-To merge these two graphs each user simply replays the edge deletions and additions
-that the other user performed.
-
-Both user then have the bottom graph and are again synchronized with each other.
+Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
+and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
+For Alice, this involves deleting \edge{3} and creating \edge{13}.
+For Bob, this involves deleting \edge{5} and creating \edge{15}.
+
+Alice and Bob then transmit their graph actions (i.e., edge additions and deletions) to each other
+and applies the other's graph actions to their own copy of the graph.
+By design, graph actions commute, so this results in both Alice and Bob having the graph in
+\autoref{fig:Editing Different Parts of the Code-c}.
 
 Note that aside from transmitting the edge deletions and additions,
-no coordination between A and B is needed to coordinate the merge.
-For example, suppose that A and B did not transmit their edits to each other
-but to a third user, C, that has the original graph at the top of \autoref{fig:Editing Different Parts of the Code}.
-User C can replay the edits from A and B in any order and will always get
-the graph at the bottom of \autoref{fig:Editing Different Parts of the Code}.
-
-Stated more formally, our edit model obeys commutativity
-as shown in \autoref{fig:Commutativity}.
+no coordination between Alice and Bob is needed to coordinate the merge.
+For example, suppose that Alice and Bob did not transmit their edits to each other
+but to a third user, Chris, that has the original graph at the top of \autoref{fig:Move-b}.
+Chris can apply the edits from Alice and Bob in any order and will always get
+the graph in \autoref{fig:Editing Different Parts of the Code-c}.
+
+Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
 For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
@@
 \subsubsection{Editing Nested Parts of the Code}
-
-(Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
+\label{sub:Editing Nested Parts of the Code}
 
@@
 
-Consider \autoref{fig:Editing Nested Parts of the Code}.
-Alice and Bob both start with the top graph, which represents the code~\texttt{2 * x + \hole{}}.
-Alice changes the~\texttt{2} to a~\texttt{3}, which results in the left graph.
-Bob moves the~\texttt{2 * x} from the left of the plus to the right of the plus
-by deleting \edge{7} and adding \edge{16}, which results in the right graph.
+After the edits in the previous subsection synchronize,
+Alice changes the~\texttt{v} to a~\texttt{w}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
+On the other hand, Bob moves the~\texttt{u * v} from the right of the plus to the left of the plus
+by deleting \edge{11} and adding \edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
 Since vertex and edge identifiers are preserved by these,
 both Alice's and Bob's edits can be applied to the other's graph,
-which results in both Alice and Bob having the bottom graph.
-
-As with Subsection~SEC:TODO, this merge does not require coordination
+which results in both Alice and Bob having the graph in \autoref{fig:Editing Nested Parts of the Code-c}.
+
+As with \autoref{sub:Editing Different Parts of the Code}, this merge does not require coordination
 between Alice and Bob.  They could independently send their edits
 to a third party, Charley, who could replay their edits in any order
-and will always get the bottom graph.
+and will always get the graph in \autoref{fig:Editing Nested Parts of the Code-c}.
 
@@
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of \edge{7} before making her
-edits and before receiving Bob's adding of \edge{16}.
+Alice might receive Bob's deletion of \edge{11} before making her
+edits and before receiving Bob's addition of \edge{19}.
 Her edits would then be within the code deleted by Bob.
@@
 
+TODO: put this somewhere.
+(Similar to how git merged across moves.
+A notable difference though is that \texttt{git} detects
+file moves by structural similarity rather than
+our nominal equality.
+(We exploit this in Subsection~REF:TODO).
+
 \subsubsection{Multi-child conflicts}
+\label{sub:Multi-child conflicts}
 
@@
    },
-   {
-    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
-     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
-     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
-    }
-   }
+   "$\texttt{var(y)}_{22}$" [> "right", > "23"']
   }
@@
     "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
-    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
-     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
-     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
-    }
+    "$\texttt{var(y)}_{22}$" [> "right", > "23"']
    }
@@
 
-Finally, we address genuine conflicts between edits.
-Consider the top graph of \autoref{fig:Multi-child}.
-Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
-while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
-These results in the left and right graphs of \autoref{fig:Multi-child} respectively.
-In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
-In the left graph it is \edge{10}, which points to a number literal,
-and in the right graph it is \edge{13}, which points to a function application.
-
-We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
-of \vertex{3}.
-This results in the bottom graph of \autoref{fig:Multi-child}.
+Finally, we address our first kind of genuine conflict between edits.
+Alice inserts the variable~\texttt{x} to the right side of the~\texttt{+}
+while Bob add the variable~\texttt{y} also to the right side of the~\texttt{+}
+These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
+In particular pay attention to the edges going out of the \texttt{right} index of \vertex{8}.
+In \autoref{fig:Multi-child-a} it is \edge{21}, which points to the variable~\texttt{x},
+and in \autoref{fig:Multi-child-b} it is \edge{23}, which points to the variable~\texttt{y}.
+
+We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
+of \vertex{8}.
+This results in \autoref{fig:Multi-child-c}.
 The conflict between Alice's and Bob's changes is
-represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
-index of \vertex{3}).
+represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{right}
+index of \vertex{8}).
 In the graph representation, we flag this conflict by highlighting its edges
-in red.
-In the textual representation, we flag this conflict with the notation~\texttt{\{ 2 | f x \}}.
-This is a conflict that much eventually be resolved by the user by deleting
+in red (TODO).
+In the textual representation, we flag this conflict with the notation~\texttt{\{ x | y \}}.
+This is a conflict that must eventually be resolved by the user by deleting
 one of the edges and editing the other to be the merged value.
-For example, one could delete \edge{10} and then edit the~\texttt{x}
-to be~\texttt{x + 2}.
+For example, one could delete \edge{23} and then edit the~\texttt{x}
+to be~\texttt{x * y}.
 Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code can be run.\footnote{
-It may be possible to develop evaluation models that allow these sorts of conflicts,
+It just must be resolved before the the code is run.\footnote{It
+may be possible to develop evaluation models that allow these sorts of conflicts,
 but that is beyond the scope of this paper.}
@@
 \subsubsection{Multi-parent conflicts}
-
-In addition to conflicts due multiple edges coming \emph{out of} an index of a particular vertex,
-conflicts can manifest as multiple edges going \emph{into} a particular vertex.
-We call these \emph{multi-parent conflicts}.
+\label{sub:Multi-parent conflicts}
 
@@
 \end{figure}
+
+TODO: alice and bob
+
+TODO: edge and vertex numbers
+
+In addition to conflicts due multiple edges coming \emph{out of} an index of a particular vertex,
+conflicts can manifest as multiple edges going \emph{into} a particular vertex.
+We call these \emph{multi-parent conflicts}.
 
@@
 We call this situation a multi-parent conflict.
+
+TODO: plus(x, \hole) that both Alice and Bob move the "x" to the right -> This results in two edges from plus to x.
 
@@
 \subsubsection{Cycles}
+\label{sub:Cycles}
 
@@
 % \end{figure}
+
+Although this case is likely to be infrequent.
+
+Finally, note that through we present several complex scenarios, this is merely for presentation.
+In practice, these complex scenarios occur less frequently than portrayed here.
 
@@
 \section{Formalism}
-\label{sec:formalism}
+\label{sec:Formalism}
 
@@
 \subsection{Syntax}
+\label{sub:Syntax}
 
@@
 \subsection{Convergent Graphs}
+\label{sub:Convergent Graphs}
 
 \subsubsection{Graphs}
+\label{sub:Graphs}
 
@@
 \subsubsection{Graph Actions}
+\label{sub:Graph Actions}
+
 A graph action is a +E or -E.
@@
 \subsection{Agda Mechanization}
-\label{sub:agda-mechanization}
+\label{sub:Agda Mechanization}
 
@@
 
-\subsection{Type System}
-\label{sub:type-system}
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
 \subsection{Edit Actions}
+\label{sub:Edit Actions}
+
 Mapping to graph actions
@@
 
+\section{Type System}
+\label{sec:Type System}
+
+TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
+
+editor states:
+
+conflicts (multiple children in the same spot)
+
+multi-parent
+
+cycles
+
+non-empty holes
+
+----------------
+
+User must supply edge selection for multi-parent
+
 \section{TODO}
@@
 \section{Implementation}
-\label{sec:implementation}
+\label{sec:Implementation}
 GRV -- how it is implemented, how it connects to the formalism, describe the graph,
@@
 \section{Related Work}
-\label{sec:related-work}
+\label{sec:Related Work}
 
@@
 \section{Discussion and Conclusion}
-\label{sec:conclusion}
+\label{sec:Discussion and Conclusion}
 
@@
 \subsection{Variable names, strings, and numbers}
+\label{sub:Variable names, strings, and numbers}
 
@@
 \section{Appendix}
+\label{apx:Appendix}
 

### 2020-08-31 grove-paper.tex
Revise figure and diagram for multi-parent conflict section

@@
   "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+   "$\texttt{+}_{24}$" [> "left", > "23"'] -> {
+    "$\texttt{var(x)}_{26}$" [> "left", > "25"'],
     {}
@@
   "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'],
-   "$\texttt{var(x)}_{2}$" [> "left", > "19"']
+   "$\texttt{+}_{24}$" [> "left", > "23"'],
+   "$\texttt{var(x)}_{26}$" [> "right", > "27"']
   }
@@
   "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   "$\texttt{+}_{24}$" [> "left", > "23"'] -> {
     {},
-    "$\texttt{var(x)}_{2}$" [> "right", > "19"']
+    "$\texttt{var(x)}_{26}$" [> "right", > "29"']
    },
@@
   a/{$\texttt{+}_{8}$} [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] ->["right", "19"'] {
+   "$\texttt{+}_{24}$" [> "left", > "23"'] ->["right", "29"'] {
     {},
-    b/"$\texttt{var(x)}_{2}$"
+    b/"$\texttt{var(x)}_{26}$"
    },
@@
 
-TODO: edge and vertex numbers
-
 In addition to conflicts due multiple edges coming \emph{out of} an index of a particular vertex,
@@
 
-An example of this is shown in \autoref{fig:Multi-parent},
-which starts with the top graph, which represents the expression~\texttt{\hole + (\hole + x)}.
-Suppose Alice moves the variable~\texttt{x} to the left-most hole.
-This results in the left graph, which represents the expression~\texttt{x + (\hole + \hole)}.
-At the same time, suppose Bob moves the variable~\texttt{x} to the right-most hole.
-This results in the right graph, which represents the expression~\texttt{\hole + (x + \hole)}.
+An example of this is shown in \autoref{fig:Multi-parent}.
+\autoref{fig:Multi-parent-a} shows the graph after Alice and Bob have made some more edits,
+and Alice and Bob's editors are synchronized with each other.
+This graph represents the expression~\texttt{(x + \hole{}) + \hole{})}.
+Alice moves the variable~\texttt{x} to the left-most hole,
+while Bob moves it to the right-most hole.
+This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
+which represent respectively the expression~\texttt{(\hole{} + \hole{}) + x} for Alice
+and the expression~\texttt{(\hole{} + x) + \hole{}} for Bob.
 In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes edge TODO and adds edge TODO, while Bob deletes edge TODO and adds edge TODO.
-As explained in Subsection~TODO:REF, deletes are idempotent so both Alice and Bob deleting
-edge TODO is not a problem.
-However, both edge TODO (added by Alice) and edge TODO (added by Bob) point to the same vertex.
-This generates the bottom graph, in which vertex TODO as two parents pointing to it.
+Alice deletes \edge{25} and adds \edge{27}, while Bob deletes \edge{25} and adds edge \edge{29}.
+As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \edge{25} is not a problem.
+However, both \edge{27} (added by Alice) and edge \edge{29} (added by Bob) point to the same vertex.
+This generates the bottom graph, in which \vertex{26} as two parents pointing to it.
 We call this situation a multi-parent conflict.
-
-TODO: plus(x, \hole) that both Alice and Bob move the "x" to the right -> This results in two edges from plus to x.
 

### 2020-09-08 grove-paper.tex
Figure out examples to use for multi-user cycles and disconnection
Rough out last figures in section 2

@@
 \end{figure}
-
-TODO: alice and bob
 
@@
 
-\subsubsection{Cycles}
+\subsubsection{Single-User Cycles}
 \label{sub:Cycles}
@@
 edits would create a cycle.
+
+\subsubsection{Multi-User Cycles}
+\label{sub:Multi-User Cycles}
 
@@
 \begin{figure}
-\centering
-
-TOP: *(*(1, \hole), *(2, \hole))
-
-LEFT: *(*(1, *(2, \hole)), \hole)
-
-RIGHT: *(\hole, *(2, *(1, \hole)))
-
-BOTTOM: *(\hole, \hole) *(1, *(2, \#1))
-
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{+}_{4}$" [> "left", > "19"'],
+    "$\texttt{*}_{6}$" [> "left", > "19"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
+     "$\texttt{*}_{6}$" [> "left", > "19"']
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
+     "$\texttt{+}_{4}$" [> "left", > "19"']
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
+    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
+   }
+  }
+ }
+};
+\path[draw] (v1) -> (v2);
+\path[draw] (v2) -> (v1);
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
 \caption{Multi-User Cycles}
@@
 (Maybe not in initial grove by example) The A and B diamond example
+
+\subsection{Disconnection}
+\label{sub:Disconnection}
+
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{1}_{4}$" [> "left", > "19"'],
+   },
+   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
+    "$\texttt{2}_{4}$" [> "left", > "19"'],
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{1}_{4}$" [> "left", > "19"'],
+    "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
+     "$\texttt{2}_{4}$" [> "left", > "19"'],
+    }
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   {},
+   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
+    "$\texttt{2}_{4}$" [> "left", > "19"'],
+    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+     "$\texttt{1}_{4}$" [> "left", > "19"'],
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"']
+ }
+};
+\path (-1cm,-2cm) graph[graph style] {
+ "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
+  "$\texttt{2}_{4}$" [> "left", > "19"'],
+ }
+};
+\path (1cm,-2cm) graph[graph style] {
+ "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+  "$\texttt{1}_{4}$" [> "left", > "19"'],
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Disconnection}
+\label{fig:Disconnection}
+\end{figure}
 

### 2020-09-14 grove-paper.tex
Convert to PLDI format and add bibliography

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 \bibliographystyle{ACM-Reference-Format}
-\citestyle{acmauthoryear}
+%\citestyle{acmauthoryear}
+\citestyle{acmnumeric}
 
@@
 \affiliation{
-  \position{Assistant Research Scientist}
+  %\position{Assistant Research Scientist}
   \department[0]{Computer Science and Engineering}
@@
 \label{sec:Introduction}
+
+TODO: PLDI: two column format
 
@@
 
+TODO: have star or dot dot arrows for skipping steps
+
+Arrow coming from 5b to 6a and 6b
+
+Diamond diagram referencing diagrams
+
 To start, Alice and Bob are synchronized with each other and
@@
 \end{tikzpicture}
+
+TODO: show UI
 \caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
@@
 \end{tikzpicture}
+
+TODO: use * and +
+
+TODO: UI (at least moc up)
 \caption{Initial code}
@@
 
-\subsection{Disconnection}
+\subsubsection{Disconnection}
 \label{sub:Disconnection}
@@
 Hazel
+\citep{Omar:2019:10.1145/3290327}
 
@@
 
-Pidjul
+Operational Transforms
+
+Pijul
 
@@
 \appendix
+
+%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
 \section{Appendix}
@@
 %% Bibliography
-%\bibliography{bibfile}
+\bibliography{grove-paper}
+
+%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
 

### 2020-09-15 grove-paper.tex
Fix LaTeX build errors

@@
 
-TODO: PLDI: two column format
+TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
 
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Multi-User Cycles-a}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Multi-User Cycles-b}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Multi-User Cycles-c}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Multi-User Cycles-d}
 \end{subfigure}
@@
 
-\begin{figure}
+\begin{figure*}
 \hfill
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Disconnection-a}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Disconnection-b}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Disconnection-c}
 \end{subfigure}
@@
 \caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\label{fig:Disconnection-d}
 \end{subfigure}
@@
 \label{fig:Disconnection}
-\end{figure}
+\end{figure*}
 
@@
 %% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
+
+%% Bibliography
+\bibliography{grove-paper}
+
+%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
+
+\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
+
 \section{Appendix}
@@
 
-Text of appendix \ldots
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
+\subsection{Concepts}
+
+\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
+  $\mathcal{A}$ & Actions & $\mathcal{S}$ & Edge States  \\
+  $\mathcal{C}$ & Cursors & $\mathcal{U}$ & UUIDs        \\
+  $\mathcal{E}$ & Edges   & $\mathcal{V}$ & Vertexes     \\
+  $\mathcal{G}$ & Graphs  & $\textbf{A}$ & Graph Actions \\
+  $\mathcal{I}$ & Indices & $\textbf{E}$ & Editors       \\
+  $\mathcal{K}$ & Constructors \\
+\end{tabular}
+
+A \emph{UUID} $u \in \mathcal{U}$ is a universally unique identifier.
+
+\subsubsection{The Target Language}
+
+\paragraph{Constructor} ~
+
+A \emph{constructor} $k \in \mathcal{K}$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.
+
+\paragraph{Index} ~
+
+An \emph{index} $i \in \mathcal{I}$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST.
+Let $i_{root}$ be the hyper-position of all rooted constructors.
+
+\subsubsection{The Graph}
+\label{sec:concepts:the-graph}
+
+\paragraph{Vertex} ~
+
+A \emph{vertex} $v = (k, u) \in \mathcal{V} = \mathcal{K} \times \mathcal{U}$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$.
+Let $v_{root} = (k_{root}, u_{root})$ be the root vertex of a graph.
+The \emph{rooted vertexes} of a graph are the immediate children of its $v_{root}$.
+
+\paragraph{Cursor} ~
+
+A \emph{cursor} $c = (v,i) \in \mathcal{C} = \mathcal{V} \times \mathcal{I}$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
+Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference
+to all of the rooted vertexes, of a graph.
+
+\paragraph{Edge} ~
+
+An \emph{edge} $\e = (c, v, u) \in \mathcal{E} = \mathcal{C} \times \mathcal{V} \times \mathcal{U}$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
+
+\paragraph{Edge State} ~
+
+An \emph{edge state} $s \in \mathcal{S} = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).
+
+\paragraph{Graph Action} ~
+
+A \emph{graph action} $A = (\varepsilon,s,u) \in \mathcal{A} = \mathcal{E} \times \mathcal{S} \times \mathcal{U}$ is an instance of a binding from edge $\varepsilon$ to edge state $s$ that can be identified by UUID $u$.
+Denote by $[\varepsilon :-> s]_{u}$ a binding from edge $\varepsilon$ to edge state $s$ (identified by UUID $u$).
+
+\paragraph{Graph} ~
+
+A \emph{graph} $G = [\varepsilon :-> s]_{u}^{*} \in \mathcal{G} = \mathcal{A}^{*}$ is a function from edges to edge states.
+Denote by $G(\varepsilon) = s$ the state $s$ of edge $\varepsilon$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.
+
+\subsubsection{The Editor}
+
+\paragraph{Move Action} ~
+
+A \emph{move action} $a \in \mathcal{A}^{move} = \{ \text{Left},\text{Right},\text{Up},\text{Down} \} \cup \{ \text{Select} \} \times \mathcal{C}$ re-positions the cursor.
+
+\paragraph{Edit Action} ~
+
+An \emph{edit action} $a \in \mathcal{A}^{edit}= \{ \text{Create} \} \times \mathcal{K} \cup \{ \text{Destroy} \} \cup \{ \text{Restore} \} \times \mathcal{V}$ adds or removes an edge at the cursor.
+
+\paragraph{Communication Action} ~
+
+A \emph{communication action} $a = (\text{Send}, A^{*}, u^{*}) \in \mathcal{A}^{comm} = \{ \text{Send} \} \times \textbf{A}^{*} \times \mathcal{U}^{*}$ applies the graph action sequence $A^{*}$ to the editors identified by UUIDs $u^{*}$.
+
+\paragraph{Action} ~
+
+An \emph{action} $a \in \mathcal{A} = \mathcal{A}^{move} \cup \mathcal{A}^{edit} \cup \mathcal{A}^{comm}$ describes a change to one or more editors.
+
+\paragraph{Editor} ~
+
+An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \textbf{E} = \mathcal{G} \times \mathcal{C} \times \textbf{A}^{*} \times \textbf{A}^{*} \times \mathcal{U}$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
+
+\subsubsection{The Environment}
+
+An \emph{environment} $(E^{*}, A^{*}) \in \textbf{E}^{*} \times \textbf{A}^{*}$ is a set of communicating editors $E^{*}$ that share a global graph action history $A^{*}$.
+
+\subsection{The Target Language}
+
+\paragraph{Formal Syntax}
+
+$e ::= \lambda p : \tau.e \mid e~e \mid e+e \mid n \mid \_ $
+
+$p ::= x $
+
+$\tau ::= \tau \rightarrow \tau \mid \text{Num}$
+
+\subsection{Sets and Relations}
+
+\paragraph{Edges}
+
+\newcommand{\edges}[1]{TODO:#1}
+\newcommand{\vertexes}[1]{TODO:#1}
+\newcommand{\liveEdges}[1]{TODO:#1}
+\newcommand{\vertexChildren}[1]{TODO:#1}
+\newcommand{\cursorChildren}[1]{TODO:#1}
+\newcommand{\parents}[1]{TODO:#1}
+\newcommand{\dom}[1]{TODO:#1}
+\newcommand{\childVertexes}[1]{TODO:#1}
+\newcommand{\liveVertexes}[1]{TODO:#1}
+\newcommand{\multiparents}[1]{TODO:#1}
+\newcommand{\known}[1]{TODO:#1}
+\newcommand{\orphans}[1]{TODO:#1}
+
+$\edges(G) = \dom(G)$
+
+$\liveEdges(G) = \{ \e \in \edges(G) | G(\e)=+ \}              \\
+  \vertexChildren(G,v) = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
+  \cursorChildren(G,c) = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
+         \parents(G,v) = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
+$
+
+\paragraph{Vertexes} ~
+
+\[
+\begin{array}{lr}
+\vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
+    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
+ \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
+         \orphans(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || = 0 \}       \\
+    \multiparents(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || > 1 \}       \\
+        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
+    %    \reachable(G) &= \orphans(G) \cup \multiparents(G) \cup \childVertexes(G,v \in \reachable(G)) \\
+    % \deleted(G) &= (\orphans(G) \setminus \{ v_{root} \})
+    %                \cup (\vertexes(G) \setminus  \\
+\end{array}
+\]
+
+\paragraph{Graph Actions} ~
+
+$\known(E_1 \cdots E_n) = A_{H1}^{*} \cdots A_{Hn}^{*}
+\iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j))$
 

### 2020-09-17 grove-paper.tex
finish adding formalism notes to appendix

--- 2020-09-14 grove-paper.tex
@@
 \usepackage{graphicx}
+\usepackage{semantic}
+\usepackage{colonequals}
+\usepackage{mathpartir}
+
+% math ligatures
+\mathlig{::=}{\coloncolonequals}
+\mathlig{|-}{\vdash}
+\mathlig{:->}{\mapsto}
+\mathlig{-->a}{\overset{a}{\longrightarrow}}
+\mathlig{||}{|}
+\mathlig{|}{:}
+\mathlig{:}{\!:\!}
+
+\def\_{\texttt{\textunderscore}}
 
@@
 % sets
+\def\A{\mathcal{A}}
+\def\C{\mathcal{C}}
 \def\E{\mathcal{E}}
@@
 \def\K{\mathcal{K}}
+\def\S{\mathcal{S}}
 \def\U{\mathcal{U}}
 \def\V{\mathcal{V}}
+\def\e{\varepsilon}
+\def\AA{\textbf{A}}
+\def\EE{\textbf{E}}
 
@@
 
+% actions
+\def\Create{\text{Create}}
+\def\Destroy{\text{Destroy}}
+\def\Down{\text{Down}}
+\def\Enqueue{\text{Enqueue}}
+\def\Left{\text{Left}}
+\def\Move{\text{Move}}
+\def\Num{\text{Num}}
+\def\Restore{\text{Restore}}
+\def\Right{\text{Right}}
+\def\Select{\text{Select}}
+\def\Send{\text{Send}}
+\def\Up{\text{Up}}
+
 % relations
 \DeclareMathOperator{\arity}{\text{arity}}
+\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
+\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
+\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
+\DeclareMathOperator{\deleted}{\text{deleted}}
+\DeclareMathOperator{\dom}{\text{dom}}
+\DeclareMathOperator{\edges}{\text{edges}}
+\DeclareMathOperator{\known}{\text{known}}
+\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
+\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
+\DeclareMathOperator{\multiparents}{\text{multiparents}}
+\DeclareMathOperator{\orphans}{\text{orphans}}
+\DeclareMathOperator{\parents}{\text{parents}}
+\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
+\DeclareMathOperator{\reachable}{\text{reachable}}
+\DeclareMathOperator{\seen}{\text{seen}}
+\DeclareMathOperator{\unseen}{\text{unseen}}
+\DeclareMathOperator{\vertexes}{\text{vertexes}}
+\DeclareMathOperator{\vertex}{\text{vertex}}
+\DeclareMathOperator{\publish}{\text{publish}}
+\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
+\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
+\DeclareMathOperator{\downIndex}{\text{downIndex}}
+\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
 
@@
 
-Text of appendix \ldots
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
+\subsection{Concepts}
+
+\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
+  $\A$ & Actions & $\S$ & Edge States  \\
+  $\C$ & Cursors & $\U$ & UUIDs        \\
+  $\E$ & Edges   & $\V$ & Vertexes     \\
+  $\G$ & Graphs  & $\A$ & Graph Actions \\
+  $\I$ & Indices & $\E$ & Editors       \\
+  $\K$ & Constructors \\
+\end{tabular}
+
+A \emph{UUID} $u \in \U$ is a universally unique identifier.
+
+\subsubsection{The Target Language}
+
+\paragraph{Constructor} ~
+
+A \emph{constructor} $k \in \K$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.
+
+\paragraph{Index} ~
+
+An \emph{index} $i \in \I$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST.
+Let $i_{root}$ be the hyper-position of all rooted constructors.
+
+\subsubsection{The Graph}
+
+\paragraph{Vertex} ~
+
+A \emph{vertex} $v = (k, u) \in \V = \K \times \U$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$.
+Let $v_{root} = (k_{root}, u_{root})$ be the root vertex of a graph.
+The \emph{rooted vertexes} of a graph are the immediate children of its $v_{root}$.
+
+\paragraph{Cursor} ~
+
+A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
+Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.
+
+\paragraph{Edge} ~
+
+An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
+
+\paragraph{Edge State} ~
+
+An \emph{edge state} $s \in \S = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).
+
+\paragraph{Graph Action} ~
+
+A \emph{graph action} $A = (\e,s,u) \in \A = \E \times \S \times \U$ is an instance of a binding from edge $\e$ to edge state $s$ that can be identified by UUID $u$.
+Denote by $[\e :-> s]_{u}$ a binding from edge $\e$ to edge state $s$ (identified by UUID $u$).
+
+\paragraph{Graph} ~
+
+A \emph{graph} $G = [\e :-> s]_{u}^{*} \in \G = \A^{*}$ is a function from edges to edge states.
+Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.
+
+\subsubsection{The Editor}
+
+\paragraph{Move Action} ~
+
+A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
+
+\paragraph{Edit Action} ~
+
+An \emph{edit action} $a \in \A^{edit}= \{ \Create \} \times \K \cup \{ \Destroy \} \cup \{ \Restore \} \times \V$ adds or removes an edge at the cursor.
+
+\paragraph{Communication Action} ~
+
+A \emph{communication action} $a = (\Send, A^{*}, u^{*}) \in \A^{comm} = \{ \Send \} \times \A^{*} \times \U^{*}$ applies the graph action sequence $A^{*}$ to the editors identified by UUIDs $u^{*}$.
+
+\paragraph{Action} ~
+
+An \emph{action} $a \in \A = \A^{move} \cup \A^{edit} \cup \A^{comm}$ describes a change to one or more editors.
+
+\paragraph{Editor} ~
+
+An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
+
+\subsubsection{The Environment}
+
+An \emph{environment} $(E^{*}, A^{*}) \in \EE^{*} \times \AA^{*}$ is a set of communicating editors $E^{*}$ that share a global graph action history $A^{*}$.
+
+\subsection{The Target Language}
+
+\paragraph{Formal Syntax} ~
+
+$e ::= \lambda p : \tau.e \mid e~e \mid e+e \mid n \mid \_ $
+
+$p ::= x $
+
+$\tau ::= \tau -> \tau \mid \Num$
+
+\subsection{Sets and Relations}
+
+\paragraph{Edges} ~
+
+\begin{align*}
+             \edges(G) &= \dom(G) \\
+         \liveEdges(G) &= \{ \e \in \edges(G) | G(\e)=+ \}              \\
+  \vertexChildren(G,v) &= \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
+  \cursorChildren(G,c) &= \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
+         \parents(G,v) &= \{ (c,v',u) \in \liveEdges(G) | v'=v \}
+\end{align*}
+
+\paragraph{Vertexes} ~
+
+\begin{align*}
+        \vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
+    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
+ \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
+         \orphans(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || = 0 \}       \\
+    \multiparents(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || > 1 \}       \\
+        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
+\end{align*}
+
+\paragraph{Graph Actions} ~
+
+$\known(E_1 \cdots E_n) = A_{H1}^{*} \cdots A_{Hn}^{*}
+\iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j))$
+
+\section{Operational Semantics}
+
+\subsection{Movement}
+
+$\boxed{E,A -->a E,A}$
+
+\begin{mathpar}
+  \inferrule{
+    \leftIndex(i) = i'
+  }{
+    (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{*},A_H^{*},u),A
+  }
+
+  \inferrule{
+    \rightIndex(i) = i'
+  }{
+    (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{*},A_H^{*},u),A
+  }
+
+  \inferrule{
+    \parents(v) = \{(c',v',u')\}
+  }{
+    (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Up} (G,c',A_Q^{*},A_H^{*},u),A
+  }
+
+  \inferrule{
+    \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
+    \downIndex(k) = i
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{*},A_H^{*},u),A
+  }
+
+  \inferrule{}{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{*},A_H^{*},u),A
+  }
+\end{mathpar}
+
+\subsection{Editing}
+
+\begin{mathpar}
+  \inferrule{
+    \defaultIndex(c) = \None \\
+    u, u_k \in \U \text{ fresh} \\
+    A_H' = [(c,(k,u_k),u) :-> +]
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{*},A_H^{*}A_H',u),A
+  }
+
+  \inferrule{
+    \defaultIndex(c) = i' \\
+    c' = ((k,u_k),i') \\
+    u, u_k \in \U \text{ fresh} \\
+    \\\\
+    \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
+    A_H'^{*} =
+    [(c  ,(k,u_k),u  ) :-> +]
+    [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
+    [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k}
+    (G A_H'^{*},c',A_Q^{*},A_H^{*}A_H'^{*},u),A
+  }
+
+  \inferrule{
+    \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
+    A_H'^{*} = [\e_j :-> -]_{j=1}^n
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Destroy}
+    (G A_H'^{*},c,A_Q^{*},A_H^{*}A_H'^{*},u),A
+  }
+
+  \inferrule{
+    A_H = [(c,v,u') :-> +] \\
+    u' \text{ fresh}
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{*},A_H^{*}A_H,u),A %
+  }
+\end{mathpar}
+
+\subsection{Communication}
+
+\paragraph{Record} ~
+%
+\begin{mathpar}
+  \inferrule{
+    a \in \A^{edit} \\
+    (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*},A_H'^{*},u),A
+  }{
+    (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*}a,A_H'^{*},u),A
+  }
+\end{mathpar}
+
+\paragraph{Replay} ~
+
+$\boxed{E,A \xrightarrow{a^{*}} E,A}$
+%
+\begin{mathpar}
+  \inferrule{
+    E,A -->a E'',A'' \\
+    E'',A'' \xrightarrow{a^{*}} E',A' \\
+  }{
+    E,A \xrightarrow{aa^{*}} E',A'
+  }
+\end{mathpar}
+
+\paragraph{Publish} ~
+
+$\boxed{(E,A)^{*} -->a (E,A)^{*}}$
+%
+\begin{mathpar}
+  \inferrule{
+    E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j) \\
+    E_j,A_j \xrightarrow{a'^{*}} (G_j',c_j',A_{Qj}'^{*}, A_{Hj}'^{*},u_j),A_j' \\
+    E_j' =
+    ( G_j'
+    , c_j'
+    , A_{Qj}'^{*} \setminus a'^{*}
+    , A_{Hj}'^{*} \setminus \known(E_1 \cdots E_n),u_j) \\
+    j = 1,\ldots,n
+  }{
+    (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{*} } (E_1',A_1') \cdots (E_n',A_n')
+  }
+\end{mathpar}
+
+Missing Action.env members: Record / Report / Stop / Replay, Dump / Load, Clone / Drop
 

### 2020-09-18 grove-paper.tex
Fix more LaTeX errors

@@
 \usepackage{semantic}
+\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
+\let\colonsim\undefined % Avoid redefinition error in `colonequals`
 \usepackage{colonequals}
@@
 
-\newcommand{\edge}[1]{Edge~#1}
-\newcommand{\vertex}[1]{Vertex~#1}
+\newcommand{\Edge}[1]{Edge~#1}
+\newcommand{\Vertex}[1]{Vertex~#1}
 
@@
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \vertex{2} and \vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+In \autoref{fig:A simple tree with a hole-b}, \Vertex{2} and \Vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
 The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vertex{2}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{2}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 \end{tikzpicture}
-\caption{\autoref{fig:A simple tree with a hole-b} after adding \vertex{6} and \edge{5}}
+\caption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
 \label{fig:Addition-b}
@@
 Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
-to that vertex from \vertex{2}.
+to that vertex from \Vertex{2}.
 \autoref{fig:Addition-b} shows the result, where the new vertex has identifier~6.
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Addition-b} after deleting \edge{5}}
+\caption{\autoref{fig:Addition-b} after deleting \Edge{5}}
 \label{fig:Change-a}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Change-a} after adding \vertex{8} and \edge{7}}
+\caption{\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}}
 \label{fig:Change-b}
@@
 Next Alice wants to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
-This is modeled by first deleting \edge{5}.
+This is modeled by first deleting \Edge{5}.
 This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
 After that deletion, we add a new vertex to represent~\texttt{z} and an
-edge from \vertex{2} to that vertex.
+edge from \Vertex{2} to that vertex.
 This results in the bottom graph in \autoref{fig:Change-b},
@@
 
-Note that \vertex{3} continues to exist, and if it had any children, those children could remain connected to it.
+Note that \Vertex{3} continues to exist, and if it had any children, those children could remain connected to it.
 In our system, once a vertex is created it is never deleted.
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Change-b} after deleting \edge{1}}
+\caption{\autoref{fig:Change-b} after deleting \Edge{1}}
 \label{fig:Wrapping-a}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-a} after adding \vertex{8} and \edge{7}}
+\caption{\autoref{fig:Wrapping-a} after adding \Vertex{8} and \Edge{7}}
 \label{fig:Wrapping-b}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-b} after adding \edge{9}}
+\caption{\autoref{fig:Wrapping-b} after adding \Edge{9}}
 \label{fig:Wrapping-c}
@@
 
-Next, Alice types~\texttt{+} when the cursor is on \vertex{2}.
+Next, Alice types~\texttt{+} when the cursor is on \Vertex{2}.
 If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
-as a child of \vertex{0}.
+as a child of \Vertex{0}.
 However, this position is already occupied by some code.
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-c} after deleting \edge{9}}
+\caption{\autoref{fig:Wrapping-c} after deleting \Edge{9}}
 \label{fig:Move-a}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-a} after adding \edge{11}}
+\caption{\autoref{fig:Move-a} after adding \Edge{11}}
 \label{fig:Move-b}
@@
 \hfill{}
-\caption{Moving \vertex{2} from the left child to the right child of \vertex{8}}
+\caption{Moving \Vertex{2} from the left child to the right child of \Vertex{8}}
 \label{fig:Move}
@@
 \autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
-First, \edge{7} is deleted, which results in \autoref{fig:Move-a}
+First, \Edge{7} is deleted, which results in \autoref{fig:Move-a}
 where the rooted code is now~\texttt{\hole{} + \hole{}},
-and \vertex{2} is orphaned.
-Next \edge{11} is added, which results in the bottom graph
+and \Vertex{2} is orphaned.
+Next \Edge{11} is added, which results in the bottom graph
 where the rooted code is now~\texttt{\hole{} + 1 * 2},
-and \vertex{2} is no longer orphaned.
+and \Vertex{2} is no longer orphaned.
 
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-b} after Alice replaces \vertex{4} with \vertex{12}}
+\caption{\autoref{fig:Move-b} after Alice replaces \Vertex{4} with \Vertex{12}}
 \label{fig:Editing Different Parts of the Code-a}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-b} after Bob replaces \vertex{6} with \vertex{14}}
+\caption{\autoref{fig:Move-b} after Bob replaces \Vertex{6} with \Vertex{14}}
 \label{fig:Editing Different Parts of the Code-b}
@@
 and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
-For Alice, this involves deleting \edge{3} and creating \edge{13}.
-For Bob, this involves deleting \edge{5} and creating \edge{15}.
+For Alice, this involves deleting \Edge{3} and creating \Edge{13}.
+For Bob, this involves deleting \Edge{5} and creating \Edge{15}.
 
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Different Parts of the Code-c} after Alice replaces \vertex{14} with \vertex{16}}
+\caption{\autoref{fig:Editing Different Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}}
 \label{fig:Editing Nested Parts of the Code-a}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Different Parts of the Code-c} after Bob move \vertex{2} to the left child of \vertex{8}}
+\caption{\autoref{fig:Editing Different Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}}
 \label{fig:Editing Nested Parts of the Code-b}
@@
 On the other hand, Bob moves the~\texttt{u * v} from the right of the plus to the left of the plus
-by deleting \edge{11} and adding \edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
+by deleting \Edge{11} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
 Since vertex and edge identifiers are preserved by these,
@@
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of \edge{11} before making her
-edits and before receiving Bob's addition of \edge{19}.
+Alice might receive Bob's deletion of \Edge{11} before making her
+edits and before receiving Bob's addition of \Edge{19}.
 Her edits would then be within the code deleted by Bob.
@@
 These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
-In particular pay attention to the edges going out of the \texttt{right} index of \vertex{8}.
-In \autoref{fig:Multi-child-a} it is \edge{21}, which points to the variable~\texttt{x},
-and in \autoref{fig:Multi-child-b} it is \edge{23}, which points to the variable~\texttt{y}.
+In particular pay attention to the edges going out of the \texttt{right} index of \Vertex{8}.
+In \autoref{fig:Multi-child-a} it is \Edge{21}, which points to the variable~\texttt{x},
+and in \autoref{fig:Multi-child-b} it is \Edge{23}, which points to the variable~\texttt{y}.
 
 We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
-of \vertex{8}.
+of \Vertex{8}.
 This results in \autoref{fig:Multi-child-c}.
@@
 represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{right}
-index of \vertex{8}).
+index of \Vertex{8}).
 In the graph representation, we flag this conflict by highlighting its edges
@@
 one of the edges and editing the other to be the merged value.
-For example, one could delete \edge{23} and then edit the~\texttt{x}
+For example, one could delete \Edge{23} and then edit the~\texttt{x}
 to be~\texttt{x * y}.
@@
 In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes \edge{25} and adds \edge{27}, while Bob deletes \edge{25} and adds edge \edge{29}.
-As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \edge{25} is not a problem.
-However, both \edge{27} (added by Alice) and edge \edge{29} (added by Bob) point to the same vertex.
-This generates the bottom graph, in which \vertex{26} as two parents pointing to it.
+Alice deletes \Edge{25} and adds \Edge{27}, while Bob deletes \Edge{25} and adds edge \Edge{29}.
+As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \Edge{25} is not a problem.
+However, both \Edge{27} (added by Alice) and edge \Edge{29} (added by Bob) point to the same vertex.
+This generates the bottom graph, in which \Vertex{26} as two parents pointing to it.
 We call this situation a multi-parent conflict.
@@
 \begin{array}{lrlll}
- \textrm{Patterns:}          & p    & {}\in Pat & {}\mathrel{::=}{} & x \mid \_ \\
- \textrm{Expressions:\qquad} & e    & {}\in Exp & {}\mathrel{::=}{} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
- \textrm{Types:}             & \tau & {}\in Typ & {}\mathrel{::=}{} & \tau \rightarrow \tau \mid Num \mid \_ \\
+ \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \_ \\
+ \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
+ \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \_ \\
 \end{array}
@@
 
-\paragraph{Constructor} ~
+\paragraph{Constructor} \par
 
@@
 
-\paragraph{Index} ~
+\paragraph{Index} \par
 
@@
 
-\paragraph{Vertex} ~
+\paragraph{Vertex} \par
 
@@
 
-\paragraph{Cursor} ~
+\paragraph{Cursor} \par
 
@@
 
-\paragraph{Edge} ~
+\paragraph{Edge} \par
 
@@
 
-\paragraph{Edge State} ~
+\paragraph{Edge State} \par
 
@@
 
-\paragraph{Graph Action} ~
+\paragraph{Graph Action} \par
 
@@
 
-\paragraph{Graph} ~
+\paragraph{Graph} \par
 
@@
 
-\paragraph{Move Action} ~
+\paragraph{Move Action} \par
 
@@
 
-\paragraph{Edit Action} ~
+\paragraph{Edit Action} \par
 
@@
 
-\paragraph{Communication Action} ~
+\paragraph{Communication Action} \par
 
@@
 
-\paragraph{Action} ~
+\paragraph{Action} \par
 
@@
 
-\paragraph{Editor} ~
+\paragraph{Editor} \par
 
@@
 
-\subsubsection{The Environment}
+\subsubsection{The Environment} \par
 
@@
 
-\paragraph{Formal Syntax} ~
+\paragraph{Formal Syntax} \par
 
@@
 
-\paragraph{Edges} ~
+\paragraph{Edges} \par
 
@@
 
-\paragraph{Vertexes} ~
+\paragraph{Vertexes} \par
 
@@
 
-\paragraph{Graph Actions} ~
+\paragraph{Graph Actions} \par
 
@@
   \inferrule{
-    \defaultIndex(c) = \None \\
+    \defaultIndex(c) = TODO(Was None) \\
     u, u_k \in \U \text{ fresh} \\
@@
 
-\paragraph{Record} ~
+\paragraph{Record} \par
 %
@@
 
-\paragraph{Replay} ~
+\paragraph{Replay} \par
 
@@
 
-\paragraph{Publish} ~
+\paragraph{Publish} \par
 

### 2020-09-21 grove-paper.tex
Some figure organization and sketches of the commutativity proof (Michael D. Adams)

@@
 \usepackage{mathpartir}
+\usepackage{stmaryrd}
 
@@
 
-\begin{figure}
+\begin{figure*}
 % TODO: naive add that creates a conflict
@@
 \label{fig:Wrapping}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:Editing Different Parts of the Code}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:Editing Nested Parts of the Code}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:Multi-child}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:Multi-parent}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:Multi-User Cycles}
-\end{figure}
+\end{figure*}
 
@@
 \label{sec:Formalism}
+
+We will now make the intuitions developed in the previous section precise
+by defining a collaborative structure editor calculus called Grove.
+We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
+the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
+...
+
+TODO: point of this section is commutativity theorem
+
+The formalism is rather simple, \autoref{sec:Grove By Example} shows
+that it is sufficient to handle many common situations,
+and this simplicity aids in the predictability.
 
@@
 
-\subsection{Agda Mechanization}
+\subsection{Commutativity}
+
+In order to define commutativity, we first define a lattice over $G$.
+Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
+$G \overset{\alpha}\rightarrow G'$
+iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
+Thus the commutativity of graph actions is established
+by the commutativity of the join operations defining those graph actions.
+
+Discuss uniqueness of uuid
+
+\begin{lemma}[Join Commutativity]
+\label{lem:Join Commutativity}
+\end{lemma}
+
+\begin{lemma}[Joining]
+\label{lem:Joining}
+For all graphs $G$ and $G'$ and all edit actions $\alpha$,
+$G \overset{\alpha}{\longrightarrow} G'$
+iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
+\end{lemma}
+\begin{proof}
+If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
+\end{proof}
+
+\begin{theorem}[Commutativity]
+\label{thm:Commutativity}
+For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
+if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
+then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
+\end{theorem}
+\begin{proof}
+Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
+$G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
+Then by \autoref{lem:Join Commutativity},
+$G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
+Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
+$G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
+\end{proof}
+
+\subsubsection{Agda Mechanization}
 \label{sub:Agda Mechanization}
-
-Discuss uniqueness of uuid
-
-\begin{theorem}[Commutativity]
-
-G  -- a1 a2 --> G'
-
-------------------
-
-G' -- a2 a1 --> G
-
-\end{theorem}
 
@@
 
-\subsection{The Target Language}
-
-\paragraph{Formal Syntax} \par
-
-$e ::= \lambda p : \tau.e \mid e~e \mid e+e \mid n \mid \_ $
-
-$p ::= x $
-
-$\tau ::= \tau -> \tau \mid \Num$
-
 \subsection{Sets and Relations}

### 2020-09-21 grove-paper.tex
Rough text for Section 2.3.5 Single-User Cycles (Michael D. Adams)

@@
 In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
-Instead a new edge with a new identifiers is created.
+Instead a new edge with a new identifier is created.
 
@@
   }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Single-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+    {}
+   },
+   "$\texttt{+}_{18}$" [> "right", > "7"'] -> {
+    "$\texttt{*}_{10}$" [> "left", > "19"'] -> {
+     "$\texttt{var(x)}_{12}$" [> "left", > "19"'],
+     {}
+    },
+    {}
+   }
+  }
+ },
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
+\label{fig:Single-User Cycles-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+    {}
+   },
+   {}
+  }
  },
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by reference as the right child of \Vertex{TODO}}
+\label{fig:Single-User Cycles-c}
 \end{subfigure}
@@
 \label{fig:Single-User Cycles}
-\end{figure}
+\end{figure*}
 
@@
 
-Although this case is likely to be infrequent.
-
-Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-Since we are representing programs as graphs,
-we have to consider cycles.
-
-Cycles can occur in a both single-user and multi-user contexts.
-
-In a single user context, a cycle can appear when a user pastes
-a vertex as a child of itself.
-An example of this is shown in \autoref{fig:Single-User Cycles},
-where the user has pasted vertex~TODO as a child of vertex~TODO.
-
-As a user interface consideration, it might be best to either
+Another case we must consider is when cycles appear in the graph.
+We categorize these into cycles caused by the action of a single user
+and cycles caused by the interaction of the actions of multiple users.
+
+In a single user context,
+normal insertion and deletion of code by the user cannot create cycles.
+However, it is possible with certain kinds of copy and paste.
+For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles-a}
+and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
+There are two ways to interpret the paste action.
+The first interpretation is to create a deep copy of \Vertex{TODO}.
+This results in \autoref{fig:Single-User Cycles-b} and
+does not cause a cycle.
+The second interpretation is to simply add an edge to \Vertex{TODO}.
+This results in \autoref{fig:Single-User Cycles-b}
+and causes a cycle.
+
+Note that not all pastes should be deep copies.
+For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
+done by a cutting from the old position and pasting to the new position.
+Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
+Distinguishing when a paste should be by reference versus by copy
+is ultimately a user interface question.
+Cycles caused by the local user's edits be detected as soon as a user enters them
+by noting either that the graph would contain a cycle or the vertex
+already has a parent somewhere in the graph.
+
+Thus, as a user interface consideration, it might be best to either
 disallow such edits to at least warn users that their
@@
 TODO: cursors are informational
+
+\subsection{Summary}
+
+Finally, note that through we present several complex scenarios, this is merely for presentation.
+In practice, these complex scenarios occur less frequently than portrayed here.
 

### 2020-09-22 grove-paper.tex
pair writing with Cyrus & Michael - Eric

@@
 
-Graph is a function from edges to edge states.
-
-$\G : \E \rightarrow \{ \bot, +, - \}$
-
-An edge is a uuid * source vertex * index * target vertex.  (We are assuming you can generate globally unique IDs.)
-
-\def\e{\varepsilon}
-
-An edge $\e \in \E = \U \times \V \times \I \times \V$ is ...
-
-A vertex is a pair of a uuid and a constructor. (uuid + constructor is the identity)
-
-$\V = \U \times \K$
-
-An edge state is bottom + - . Define the total ordering (forms a lattice).
+TODO: rename indices to positions
+TODO: look into using bold plus and minus (google results)
+
+A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states, 
+where $\E = \U \times \V \times \I \times \V$ 
+and   unique IDs are drawn from some suitable set $\U$
+and   vertices are drawn from $\V$, 
+and   indices are drawn from $\I$, 
+and   edge states are drawn from $\Sigma$, all defined below.
+
+Each edge, $\e = (u, v_s, p, v_t)$, has a unique name, $u \in \U$, and connects the child position $p$ of a source vertex $v_s$ to a target vertex $v_t$.
+The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, +, - \}$. 
+If $\G(\e) = \bot$, then $\e$ does not yet exist.
+If $\G(\e)$ is $\mathbf{+}$, the edge has been created.
+If $\G(\e)$ is -, the edge has been destroyed.
+The total ordering $\bot \sqsubseteq \mathbb{+} \sqsubseteq -$ forms a lattice over $\Sigma$.
+
+A vertex $v = (u, k) \in \U \times \K$ is a pair of a globally unique identifier $u$ and a constructor $k$. We say $u$ identifies vertex $v$ as a distinct instance of constructor $k$.
 

### 2020-09-22 grove-paper.tex
Create \Plus and \Minus commands for outlined plus and minus symbols (Michael D. Adams)

@@
 \usepackage{stmaryrd}
+\usepackage{fontawesome}
 
@@
 \DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
+
+% Define outline versions of + and -
+\def\outlinepad{0.4pt}
+\def\outlinestroke{0.4pt}
+\newcommand{\Plus}{\mathord{
+\begin{tikzpicture}[anchor=base, baseline]
+%\node at (0,0) {+};
+\path[draw, line width=\outlinestroke]
+   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
+ --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
+ --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
+ --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
+ --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --cycle
+ ;
+\end{tikzpicture}
+}}
+\newcommand{\Minus}{\mathord{
+\begin{tikzpicture}[anchor=base, baseline]
+%\node at (0,0) {$-$};
+\path[draw, line width=\outlinestroke]
+   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
+ --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
+ --cycle
+ ;
+\end{tikzpicture}
+}}
 
@@
 Each edge, $\e = (u, v_s, p, v_t)$, has a unique name, $u \in \U$, and connects the child position $p$ of a source vertex $v_s$ to a target vertex $v_t$.
-The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, +, - \}$. 
+The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. 
 If $\G(\e) = \bot$, then $\e$ does not yet exist.
@@
 If $\G(\e)$ is -, the edge has been destroyed.
-The total ordering $\bot \sqsubseteq \mathbb{+} \sqsubseteq -$ forms a lattice over $\Sigma$.
+The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
 

### 2020-09-23 grove-paper.tex
Wrote Section 2.3.6 Multi-User Cycles (Michael D. Adams)

@@
 
-While a single-user edits can always be checked to see
+While single-user edits can be checked to see
 whether they create a cycle and thus the system
@@
 multi-user edits may not manifest cycles until after
-the edits are merged.
-
-For example, consider the situation shown in \autoref{fig:Multi-User Cycles}.
-The top graph represents code that is synchronized between Alice and Bob
-and represents the code~\texttt{(1 * \hole) * (2 * \hole)}.
-In the left graph, Alice moves vertex~TODO to the \texttt{right}
-index of vertex~TODO.
-The resulting graph represents the code~\texttt{(1 * (2 * \hole)) * \hole}.
-In the right graph, Bob moves vertex~TODO to the \texttt{right} index of vertex~TODO.
-The resulting graph represents the code~\texttt{\hole * (2 * (1 * \hole))}.
-On their own, neither edit creates a cycle.
-However, when these edits are merged,
-not only is a cycle created, but
-the vertexes in the cycle
-
-Can also be detected due to neither original leaving them disconnected.
+the edits from multiple users are merged.
+
+For example, consider the situation in \autoref{fig:Multi-User Cycles}.
+\autoref{fig:Multi-User Cycles-a} shows the graph after Alice and Bob have made some more edits,
+and Alice and Bob's editors are synchronized with each other.
+This graph represents the expression~\texttt{((\hole+\hole)+(\hole*\hole))+\hole}.
+Alice moves the leftmost \texttt{+} into the rightmost hold
+and then the \texttt{*} underneath that.
+This results in \autoref{fig:Multi-User Cycles-b},
+which represents the expression~\texttt{(\hole+\hole)+((\hole*\hole)+\hole)}.
+On the other hand, Bob does the same move
+but puts the \texttt{+} and under the \texttt{*}.
+This results in \autoref{fig:Multi-User Cycles-c},
+which represents the expression~\texttt{(\hole+\hole)+((\hole+\hole)*\hole)}.
+
+On their own, neither of these edits creates a cycle.
+However, merging the edit actions of both Alice and Bob results in the graph
+in \autoref{fig:Multi-User Cycles-d}, which has a cycle
+between \Vertex{TODO} and \Vertex{TODO}.
+This situation is akin to a merge conflict in traditional version control systems.
+Resolving it requires user input, but is as simple as
+either deleting \Edge{TODO} and \Edge{TODO} (thus favoring Alice's version)
+or deleting \Edge{TODO} and \Edge{TODO} (thus favoring Bob's version).
+Since a cycle like this that is connected to the root vertex always
+contains vertexes with multiple parents that if removed would break the cycle,
+we use the user interface described in \autoref{sub:Multi-parent conflicts}
+to display these kinds of programs.
+For example, \autoref{fig:Multi-User Cycles-d} would be displayed
+as shown in TODO:FIGURE.
+We leave the user interface considerations of resolving the conflict such a
+cycle represents to future work.
 
@@
     "$\texttt{+}_{4}$" [> "left", > "19"'],
-    "$\texttt{*}_{6}$" [> "left", > "19"']
+    "$\texttt{*}_{6}$" [> "right", > "19"']
    },
@@
     "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
-     "$\texttt{*}_{6}$" [> "left", > "19"']
+     "$\texttt{*}_{6}$" [> "left", > "19"'],
+     {}
     }
@@
     "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
-     "$\texttt{+}_{4}$" [> "left", > "19"']
+     "$\texttt{+}_{4}$" [> "left", > "19"'],
+     {}
     }
@@
 };
-\path[draw] (v1) -> (v2);
-\path[draw] (v2) -> (v1);
+\path[draw,->] (v1) to [out=-45,in=-135] (v2);
+\path[draw,->] (v2) to [out=-90,in=-90] (v1);
 \end{tikzpicture}
@@
 If $\G(\e) = \bot$, then $\e$ does not yet exist.
-If $\G(\e)$ is $\mathbf{+}$, the edge has been created.
-If $\G(\e)$ is -, the edge has been destroyed.
+If $\G(\e)$ is $\Plus$, the edge has been created.
+If $\G(\e)$ is $\Minus$, the edge has been destroyed.
 The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
@@
 
+\section{Supplemental Material: Complete Graph Sequences for All Figures}
+\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
+
+TODO: put all intermediate states for all graphs along with the edge actions
+
 \section{Appendix}

### 2020-09-23 grove-paper.tex
Add vertexes, constructors, and positions to section 3.2.1 graph formalism (Eric)

@@
 \def\K{\mathcal{K}}
+\def\P{\mathcal{P}}
 \def\S{\mathcal{S}}
@@
 A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states, 
-where $\E = \U \times \V \times \I \times \V$ 
-and   unique IDs are drawn from some suitable set $\U$
+where $\E = \U \times \V \times \P \times \V$,
+and   unique IDs are drawn from some suitable set $\U$,
 and   vertices are drawn from $\V$, 
-and   indices are drawn from $\I$, 
+and   positions are drawn from $\P$, 
 and   edge states are drawn from $\Sigma$, all defined below.
 
-Each edge, $\e = (u, v_s, p, v_t)$, has a unique name, $u \in \U$, and connects the child position $p$ of a source vertex $v_s$ to a target vertex $v_t$.
+Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects the child position $p$ of a source vertex $v$ to a target vertex $v^\prime$.
 The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
@@
 
-A vertex $v = (u, k) \in \U \times \K$ is a pair of a globally unique identifier $u$ and a constructor $k$. We say $u$ identifies vertex $v$ as a distinct instance of constructor $k$.
-
-Define $\sqcup_{edge state}$ (i.e. join)
-
-lattice ordering $\bot < + < -$ LUB e.g., join A B = C
+A vertex $v = (u, k)$ identifies a distinct instance of the constructor $k$.
+Each constructor represents a different type of node in the target language's abstract syntax tree.
+
+A position $p$ identifies a location adjacent to some parent vertex.
+
+Every graph contains a distinguished root vertex, root constructor, and root position, respectively named $v_0$, $k_0$, and $p_0$.
+
+% Define $\sqcup_{edge state}$ (i.e. join)
+
+% lattice ordering $\bot < + < -$ LUB e.g., join A B = C
 
@@
   $\E$ & Edges   & $\V$ & Vertexes      \\
-  $\G$ & Graphs  & $\A$ & Graph Actions \\
-  $\I$ & Indices & $\E$ & Editors       \\
+  $\G$ & Graphs  & $\AA$ & Graph Actions \\
+  $\I$ & Indices & $\EE$ & Editors       \\
   $\K$ & Constructors                   \\

### 2020-09-28 grove-paper.tex
Rough draft of Section 2.3.7 Disconnection

@@
 \mathlig{-->a}{\overset{a}{\longrightarrow}}
-\mathlig{||}{|}
-\mathlig{|}{:}
-\mathlig{:}{\!:\!}
+%\mathlig{||}{|}
+%\mathlig{|}{:}
+%\mathlig{:}{\!:\!}
+\newcommand{\abs}[1]{\left|{#1}\right|}
 
@@
 
+\begin{figure*}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+    "$\texttt{+}_{4}$" [> "left", > "19"'],
+    "$\texttt{*}_{6}$" [> "right", > "19"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Multi-User Cycles-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
+     "$\texttt{*}_{6}$" [> "left", > "19"'],
+     {}
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Multi-User Cycles-b}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
+     "$\texttt{+}_{4}$" [> "left", > "19"'],
+     {}
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Multi-User Cycles-c}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+   },
+   {
+    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
+    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
+   }
+  }
+ }
+};
+\path[draw,->] (v1) to [out=-45,in=-135] (v2);
+\path[draw,->] (v2) to [out=-90,in=-90] (v1);
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Multi-User Cycles-d}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-User Cycles}
+\label{fig:Multi-User Cycles}
+\end{figure*}
+
+TODO: NEEDS UI MOCKUP
+
 While single-user edits can be checked to see
@@
 
+\subsubsection{Disconnection}
+\label{sub:Disconnection}
+
 \begin{figure*}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\textwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+   "$\texttt{+}_{2}$" [> "left", > "19"'],
+   "$\texttt{*}_{6}$" [> "left", > "19"']
+  }
+ }
+};
+\end{tikzpicture}
+\caption{Initial code}
+\label{fig:Disconnection-a}
+\end{subfigure}
+%%%%%%%%
 \hfill
@@
    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{+}_{4}$" [> "left", > "19"'],
-    "$\texttt{*}_{6}$" [> "right", > "19"']
+    "$\texttt{*}_{6}$" [> "left", > "19"'],
+    {}
    },
@@
 \caption{Initial code}
-\label{fig:Multi-User Cycles-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-   },
-   {
-    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
-     "$\texttt{*}_{6}$" [> "left", > "19"'],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-   },
-   {
-    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
-     "$\texttt{+}_{4}$" [> "left", > "19"'],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-   },
-   {
-    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
-    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
-   }
-  }
- }
-};
-\path[draw,->] (v1) to [out=-45,in=-135] (v2);
-\path[draw,->] (v2) to [out=-90,in=-90] (v1);
-\end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-d}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Multi-User Cycles}
-\label{fig:Multi-User Cycles}
-\end{figure*}
-
-NEEDS UI MOCKUP
-
-(Maybe not in initial grove by example) The A and B diamond example
-
-\subsubsection{Disconnection}
-\label{sub:Disconnection}
-
-\begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{1}_{4}$" [> "left", > "19"'],
-   },
-   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
-    "$\texttt{2}_{4}$" [> "left", > "19"'],
-   }
-  }
- }
-};
-\end{tikzpicture}
-\caption{Initial code}
-\label{fig:Disconnection-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{1}_{4}$" [> "left", > "19"'],
-    "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
-     "$\texttt{2}_{4}$" [> "left", > "19"'],
-    }
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\caption{Initial code}
 \label{fig:Disconnection-b}
@@
    {},
-   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
-    "$\texttt{2}_{4}$" [> "left", > "19"'],
-    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-     "$\texttt{1}_{4}$" [> "left", > "19"'],
-    }
+   "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
+    "$\texttt{+}_{2}$" [> "left", > "19"'],
+    {}
    }
@@
 };
-\path (-1cm,-2cm) graph[graph style] {
- "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
-  "$\texttt{2}_{4}$" [> "left", > "19"'],
- }
-};
-\path (1cm,-2cm) graph[graph style] {
- "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-  "$\texttt{1}_{4}$" [> "left", > "19"'],
- }
-};
+\path (-1cm,-3cm) graph[graph style] {
+ times/"$\texttt{*}_{6}$" [> "left", > "19"']
+};
+\path (1cm,-3cm) graph[graph style] {
+ plus/"$\texttt{+}_{2}$" [> "left", > "19"']
+};
+\path[draw,->] (times) to [out=45,in=135] (plus);
+\path[draw,->] (plus) to [out=-135,in=-45] (times);
 \end{tikzpicture}
@@
 \end{figure*}
+
+Finally, when multi-user edits are merged,
+parts of the graph can become disconnected
+even though they are connected in each user's
+copy of the graph before the merge.
+
+For example, consider the situation in \autoref{fig:Disconnection}.
+\autoref{fig:Disconnection-a} shows the graph after Alice and Bob have made some more edits,
+and Alice and Bob's editors are synchronized with each other.
+This graph represents the expression~\texttt{(\hole+\hole)+(\hole*\hole)}.
+Alice moves the \texttt{*} under \texttt{+}$_2$.
+This results in \autoref{fig:Disconnection-b},
+which represents the expression~\texttt{((\hole*\hole)+\hole)+\hole}.
+On the other hand, Bob puts \texttt{+}$_2$ and under the \texttt{*}.
+This results in \autoref{fig:Disconnection-c},
+which represents the expression~\texttt{\hole+((\hole+\hole)*\hole)}.
+
+When these edits are merged, the result is the graph in \autoref{fig:Disconnection-d}.
+Since Alice deleted the edge from \Vertex{TODO} to \Vertex{TODO}
+and Bob deleted the edge from \Vertex{TODO} to \Vertex{TODO},
+when these are merged \Vertex{TODO} and \Vertex{TODO}
+are completely disconnected from the root.
+
+This is similar to the type of merge conflict
+that can arise in traditional version control systems
+when two users delete different functions that
+are duplicates of the other.
+
+Fortunately, we can detect this situation and warn the user
+since the disconnection appears after merging two graphs
+that did not have disconnected vertexes.
 
@@
  \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
-         \orphans(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || = 0 \}       \\
-    \multiparents(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || > 1 \}       \\
+         \orphans(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}       \\
+    \multiparents(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}       \\
         \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))

### 2020-09-29 grove-paper.tex
Commit discussion from Monday (Michael D. Adams)

@@
  nodes={draw, rounded corners=1mm},
- edges={draw, sloped, font={\small\ttfamily}}
+ edges={draw, font={\small\ttfamily}}
 }}
@@
 
+do not model cursors
+but model later in \autoref{sub:Cursors} 
+they are not involved in the communication of graph edits
+though their representation is slightly subtly
+
 TODO: group user-level edits (with user visible)
@@
 
+TODO: use mathit instead of var(x)
+
 %We represent expression contain holes as part of the editor state 
@@
   "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+   "$\texttt{x}_{4}$" [> "left", > "3"'],
    {}
@@
 \path (2cm,-2cm) graph[graph style] {
- y [as={$\texttt{var(y)}_{6}$}]
+ y [as={$\texttt{y}_{6}$}]
 };
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-   "$\texttt{var(z)}_{8}$" [> "right", > "7"']
+  "$\texttt{*}_{2}$" [> "RT$_1$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_3$"'],
+   "$\texttt{z}_{8}$" [> "R$_7$"]
   }
@@
 \path (2cm,-2cm) graph[graph style] {
- "$\texttt{var(y)}_{6}$"
+ "$\texttt{y}_{6}$"
 };
@@
 \end{figure}
+
+Arrows between figures
 
@@
 We assume multiple instances
+
+TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
+In practice, these complex scenarios occur less frequently than portrayed here.
+
+Common scenarios first and then consider more unusual scenarios
+that cause cycles or multiple parents
+
+TODO: \autoref{sub:Editing Different Parts of the Code}
 
@@
 
-TODO: cursors are informational
-
-\subsection{Summary}
-
-Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
+Cursor is either on edge or on position
+
+cursors are advisory
+
+Don't model communication but would be separate layer anyway
 

### 2020-09-30 grove-paper.tex
Commit further work on figures

@@
 \usepackage{booktabs}
-\usepackage{subcaption}
+\usepackage[rule=false]{subcaption}
 \usepackage{graphicx}
@@
 \usetikzlibrary{
+  backgrounds,  % Provides "framed" and "gridded"
   graphs,       % Graph *notation*
@@
 
+\tikzset{
+  every picture/.style={framed, background rectangle/.style={draw=gray!50}},
+}
 \tikzset{graphs/graph style/.style={
@@
   nodes={draw, rounded corners=1mm},
- edges={draw, font={\small\ttfamily}}
+  edges={draw, font={\small\ttfamily}},
+  /tikz/every edge quotes/.style={
+    %draw=gray!20,
+    anchor=west,
+    swap/.append code={
+      \ifpgfarrowswap
+        \pgfkeysalso{anchor=west}
+      \else
+        \pgfkeysalso{anchor=east}
+      \fi}},
 }}
@@
 }}
-%\tikzset{graphs/every edge/.style={}
 
@@
 \autoref{sub:Single User Edits} explores edits by a single user, Alice,
-and \autoref{sub:Collaborative Structure Editing} explores edits by multiple users, Alice and Bob,
+and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
 each of whom is editing their own instance of a program.
@@
 
-% TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
-
-% TODO: remove cursor from Fig 1 and Fig 2
-
-\subsubsection{Additions}
-\label{sub:Additions}
+In the remainder of this section,
+we will consider user edits, starting
+in this subsection with single-user
+edits and continuing in \autoref{sub:Multi-user Edits}
+with multi-user edits.
+
+Each user edit that we will consider corresponds to one or more graph edits.
+Graph edits can only add fresh edges and vertices or remove existing edges.
+
+User edits act relative to a cursor. 
+We do not model cursor sharing and the cursor does not appear in the underlying graph.
+We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
+
+\subsubsection{Construction}
+\label{sub:Construction}
 
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.3\textwidth}
+\begin{tabular}{|p{0.49\linewidth}|p{0.49\linewidth}|}
+%\centering
+TODO
+&
+%\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_3$"'],
+   "$\texttt{y}_{6}$" [> "R$_5$"]
+  }
+ }
+};
+\end{tikzpicture}
+\\
+\subcaption{User interface}
+\label{fig:Construction-a}
+ &
+\subcaption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
+\label{fig:Construction-b}
+\\ \hline
+\end{tabular}
+\caption{Construction}
+\label{fig:Construction}
+\end{figure}
+
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.49\columnwidth}
 \centering
@@
 \caption{User interface}
-\label{fig:Addition-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-   "$\texttt{var(y)}_{6}$" [> "right", > "5"']
+\label{fig:Construction-a}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.49\columnwidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_3$"'],
+   "$\texttt{y}_{6}$" [> "R$_5$"]
   }
@@
 \caption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
-\label{fig:Addition-b}
+\label{fig:Construction-b}
 \end{subfigure}
@@
 \hfill{}
-\caption{Addition}
-\label{fig:Addition}
+\caption{Construction}
+\label{fig:Construction}
 \end{figure}
 
-TODO: Actions act at a cursor.  Each instance also has a cursor that it controls, but this is not part of the graph representation.
-
-Alice starts with the code in \autoref{fig:A simple tree with a hole},
+TODO: "user edits" and "graph edits"
+
+To start our examples, Alice starts with the code in \autoref{fig:A simple tree with a hole}.
 She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
@@
 to that vertex from \Vertex{2}.
-\autoref{fig:Addition-b} shows the result, where the new vertex has identifier~6.
+\autoref{fig:Construction-b} shows the result, where the new vertex has identifier~6.
 This graph represents the expression \texttt{x * y}.
@@
 TODO: z -> foo (or multi-editor)
+
+% TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
+
+% TODO: remove cursor from Fig 1 and Fig 2
 
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
+\begin{subfigure}{0.49\linewidth}
 \centering
@@
 };
-\path (2cm,-2cm) graph[graph style] {
+\path (1cm,-2cm) graph[graph style] {
  y [as={$\texttt{y}_{6}$}]
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Addition-b} after deleting \Edge{5}}
+\caption{\autoref{fig:Construction-b} after deleting \Edge{5}}
 \label{fig:Change-a}
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
+\begin{subfigure}{0.49\linewidth}
+\centering
+\begin{tikzpicture}[framed]
 \path (0cm,0cm) graph[graph style] {
@@
 };
-\path (2cm,-2cm) graph[graph style] {
+\path (1cm,-2cm) graph[graph style] {
  "$\texttt{y}_{6}$"
@@
 
-\subsection{Collaborative Structure Editing}
-\label{sub:Collaborative Structure Editing}
+\subsection{Multi-user Edits}
+\label{sub:Multi-user Edits}
 

### 2020-09-30 grove-paper.tex
Rework figures (Michael D. Adams)

@@
 \usepackage{fontawesome}
+\usepackage{array}
 
@@
 }
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level sep=1cm,
-  nodes={draw, rounded corners=1mm},
-  edges={draw, font={\small\ttfamily}},
+\tikzset{edge style/.style={
+  draw,
+  font={\small\ttfamily},
   /tikz/every edge quotes/.style={
@@
       \fi}},
+}}
+\tikzset{graphs/graph style/.style={
+  tree layout,
+  level sep=1cm,
+  nodes={draw, rounded corners=1mm},
+  edges={edge style},
 }}
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
-   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
+\begin{subfigure}{0.49\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_{3}$"],
    {}
@@
 \begin{figure}
-\begin{tabular}{|p{0.49\linewidth}|p{0.49\linewidth}|}
-%\centering
-TODO
-&
-%\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_3$"'],
-   "$\texttt{y}_{6}$" [> "R$_5$"]
-  }
- }
-};
-\end{tikzpicture}
-\\
-\subcaption{User interface}
-\label{fig:Construction-a}
- &
-\subcaption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
-\label{fig:Construction-b}
-\\ \hline
-\end{tabular}
-\caption{Construction}
-\label{fig:Construction}
-\end{figure}
-
-\begin{figure}
 \hfill
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_3$"'],
-   "$\texttt{y}_{6}$" [> "R$_5$"]
+  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+   "$\texttt{y}_{6}$" [> "R$_{5}$"]
   }
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
-   "$\texttt{x}_{4}$" [> "left", > "3"'],
+  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
    {}
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "RT$_1$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_3$"'],
-   "$\texttt{z}_{8}$" [> "R$_7$"]
+  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
+   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+   "$\texttt{z}_{8}$" [> "R$_{7}$"]
   }
@@
 
-\begin{figure*}
+\begin{figure}
 % TODO: naive add that creates a conflict
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
+\begin{subfigure}{0.32\linewidth}
 \centering
@@
 };
-\path (-1cm,-3cm) graph[graph style] {
+\path (-0.5,-3cm) graph[graph style] {
  "$\texttt{*}_{2}$" -> {
-  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+  "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+  "$\texttt{z}_{6}$" [> "R$_{5}$"]
  }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"']
- }
-};
-\path (-1cm,-3cm) graph[graph style] {
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"]
+ }
+};
+\path (-0.5cm,-3cm) graph[graph style] {
  "$\texttt{*}_{2}$" -> {
-  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+  "$\texttt{x}_{4}$" [> "L$_{5}$"'],
+  "$\texttt{z}_{6}$" [> "R$_{5}$"]
  }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "9"'] -> {
-    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{9}$"'] -> {
+    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+    "$\texttt{z}_{6}$" [> "R$_{5}$"]
    },
@@
 \label{fig:Wrapping}
-\end{figure*}
+\end{figure}
 
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"']
+\begin{subfigure}{0.49\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"]
  }
@@
  "$\texttt{*}_{2}$" -> {
-  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+  "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+  "$\texttt{z}_{6}$" [> "R$_{5}$"]
  }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+\begin{subfigure}{0.45\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
-    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+    "$\texttt{z}_{6}$" [> "R$_{5}$"]
    }
@@
 
-\begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
+   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{z}_{6}$" [> "R$_{5}$"]
    }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
-    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
-    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+    "$\texttt{v}_{14}$" [> "R$_{15}$"]
    }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{v}_{14}$" [> "R$_{15}$"]
    }
@@
 \label{fig:Editing Different Parts of the Code}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{w}_{16}$" [> "R$_{17}$"]
    }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{v}_{14}$" [> "R$_{15}$"]
    },
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{w}_{16}$" [> "R$_{17}$"]
    },
@@
 \label{fig:Editing Nested Parts of the Code}
-\end{figure*}
+\end{figure}
 
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{w}_{16}$" [> "R$_{17}$"]
    },
    {
-    "$\texttt{var(x)}_{20}$" [> "right", > "21"']
+    "$\texttt{x}_{20}$" [> "R$_{21}$"]
    }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{w}_{16}$" [> "R$_{17}$"]
    },
-   "$\texttt{var(y)}_{22}$" [> "right", > "23"']
+   "$\texttt{y}_{22}$" [> "R$_{23}$"]
   }
@@
 %%%%%%%%
-\begin{subfigure}{0.3\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
-    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
+    "$\texttt{w}_{16}$" [> "R$_{17}$"]
    },
    {
-    "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
-    "$\texttt{var(y)}_{22}$" [> "right", > "23"']
-   }
-  }
- }
-};
-\end{tikzpicture}
-
-TODO: show UI
+    "$\texttt{x}_{20}$" [> "R$_{21}$"],
+    "$\texttt{y}_{22}$" [> "R$_{23}$"]
+   }
+  }
+ }
+};
+\end{tikzpicture}
 \caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
 \label{fig:Multi-child-c}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.24\linewidth}
+\centering
+TODO: show UI
+\caption{User interface showing the multi-child conflict in \autoref{fig:Multi-child-d}}
+\label{fig:Multi-child-d}
 \end{subfigure}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{24}$" [> "left", > "23"'] -> {
-    "$\texttt{var(x)}_{26}$" [> "left", > "25"'],
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
+    "$\texttt{x}_{26}$" [> "L$_{25}$"'],
     {}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{24}$" [> "left", > "23"'],
-   "$\texttt{var(x)}_{26}$" [> "right", > "27"']
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{24}$" [> "L$_{23}$"'],
+   "$\texttt{x}_{26}$" [> "R$_{27}$"]
   }
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{24}$" [> "left", > "23"'] -> {
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
     {},
-    "$\texttt{var(x)}_{26}$" [> "right", > "29"']
+    "$\texttt{x}_{26}$" [> "R$_{29}$"']
    },
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{$\texttt{+}_{8}$} [> "root", > "7"'] -> {
-   "$\texttt{+}_{24}$" [> "left", > "23"'] ->["right", "29"'] {
+\begin{subfigure}{0.24\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  a/{$\texttt{+}_{8}$} [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
     {},
-    b/"$\texttt{var(x)}_{26}$"
+    b/"$\texttt{x}_{26}$" [> "R$_{29}$"']
    },
@@
  },
- a ->["right", "27"'] b
-};
+};
+\path [edge style,->] (a) to ["R$_{27}$",out=-60,in=60] (b);
 \end{tikzpicture}
@@
 
-\begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
     {}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
     {}
    },
-   "$\texttt{+}_{18}$" [> "right", > "7"'] -> {
-    "$\texttt{*}_{10}$" [> "left", > "19"'] -> {
-     "$\texttt{var(x)}_{12}$" [> "left", > "19"'],
+   "$\texttt{+}_{18}$" [> "R$_{7}$"] -> {
+    "$\texttt{*}_{10}$" [> "L$_{19}$"'] -> {
+     "$\texttt{x}_{12}$" [> "L$_{19}$"'],
      {}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
+\begin{subfigure}{0.32\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
     {}
@@
   }
- },
- v2 -> v8
-};
+ }
+};
+\path[edge style,->] (v2) to ["R$_{19}$",out=-60,in=-60] (v8);
 \end{tikzpicture}
@@
 \label{fig:Single-User Cycles}
-\end{figure*}
+\end{figure}
 
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{+}_{4}$" [> "left", > "19"'],
-    "$\texttt{*}_{6}$" [> "right", > "19"']
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+    "$\texttt{+}_{4}$" [> "L$_{19}$"'],
+    "$\texttt{*}_{6}$" [> "R$_{19}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
    },
    {
-    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
-     "$\texttt{*}_{6}$" [> "left", > "19"'],
+    "$\texttt{+}_{4}$" [> "R$_{19}$"] -> {
+     "$\texttt{*}_{6}$" [> "L$_{19}$"'],
      {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
    },
    {
-    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
-     "$\texttt{+}_{4}$" [> "left", > "19"'],
+    "$\texttt{*}_{6}$" [> "R$_{19}$"] -> {
+     "$\texttt{+}_{4}$" [> "L$_{19}$"'],
      {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
    },
    {
-    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
-    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
+    v1/"$\texttt{*}_{6}$" [> "R$_{19}$"],
+    v2/"$\texttt{+}_{4}$" [> "R$_{19}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'],
-   "$\texttt{*}_{6}$" [> "left", > "19"']
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'],
+   "$\texttt{*}_{6}$" [> "R$_{19}$"]
   }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
-   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
-    "$\texttt{*}_{6}$" [> "left", > "19"'],
-    {}
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+    {},
+    "$\texttt{*}_{6}$" [> "R$_{19}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
    {},
-   "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
-    "$\texttt{+}_{2}$" [> "left", > "19"'],
+   "$\texttt{*}_{6}$" [> "R$_{19}$"] -> {
+    "$\texttt{+}_{2}$" [> "L$_{19}$"'],
     {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "root", > "7"']
+  "$\texttt{+}_{8}$" [> "Root$_{7}$"]
  }
@@
 \path (-1cm,-3cm) graph[graph style] {
- times/"$\texttt{*}_{6}$" [> "left", > "19"']
+ times/"$\texttt{*}_{6}$" [> "L$_{19}$"']
 };
 \path (1cm,-3cm) graph[graph style] {
- plus/"$\texttt{+}_{2}$" [> "left", > "19"']
+ plus/"$\texttt{+}_{2}$" [> "L$_{19}$"']
 };

### 2020-10-03 grove-paper.tex
Fix vertex and edge numbering in (most) figures (Michael D. Adams)

@@
   "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_{3}$"],
+   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
    {}
@@
   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-  "$\texttt{z}_{6}$" [> "R$_{5}$"]
+  "$\texttt{z}_{8}$" [> "R$_{7}$"]
  }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"]
  }
@@
   "$\texttt{x}_{4}$" [> "L$_{5}$"'],
-  "$\texttt{z}_{6}$" [> "R$_{5}$"]
+  "$\texttt{z}_{8}$" [> "R$_{7}$"]
  }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{9}$"'] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{11}$"'] -> {
     "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-    "$\texttt{z}_{6}$" [> "R$_{5}$"]
+    "$\texttt{z}_{8}$" [> "R$_{7}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
     "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-    "$\texttt{z}_{6}$" [> "R$_{5}$"]
+    "$\texttt{z}_{6}$" [> "R$_{7}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{z}_{6}$" [> "R$_{5}$"]
+   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
+    "$\texttt{z}_{6}$" [> "R$_{7}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
+   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
     "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-    "$\texttt{v}_{14}$" [> "R$_{15}$"]
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{v}_{14}$" [> "R$_{15}$"]
+   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{11}$"] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{w}_{16}$" [> "R$_{17}$"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\caption{\autoref{fig:Editing Different Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}}
+   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}}
 \label{fig:Editing Nested Parts of the Code-a}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{v}_{14}$" [> "R$_{15}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    },
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Different Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}}
+\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}}
 \label{fig:Editing Nested Parts of the Code-b}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{w}_{16}$" [> "R$_{17}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{w}_{16}$" [> "R$_{17}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    },
    {
-    "$\texttt{x}_{20}$" [> "R$_{21}$"]
+    "$\texttt{x}_{22}$" [> "R$_{21}$"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{w}_{16}$" [> "R$_{17}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    },
-   "$\texttt{y}_{22}$" [> "R$_{23}$"]
+   "$\texttt{y}_{24}$" [> "R$_{23}$"]
   }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{u}_{12}$" [> "L$_{13}$"'],
-    "$\texttt{w}_{16}$" [> "R$_{17}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+    "$\texttt{v}_{18}$" [> "R$_{17}$"]
    },
    {
-    "$\texttt{x}_{20}$" [> "R$_{21}$"],
-    "$\texttt{y}_{22}$" [> "R$_{23}$"]
+    "$\texttt{x}_{22}$" [> "R$_{21}$"],
+    "$\texttt{y}_{24}$" [> "R$_{23}$"]
    }
@@
 %%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
-    "$\texttt{x}_{26}$" [> "L$_{25}$"'],
+\begin{subfigure}{0.19\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
     {}
@@
 \end{tikzpicture}
-
-TODO: use * and +
-
 TODO: UI (at least moc up)
@@
 %%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{24}$" [> "L$_{23}$"'],
-   "$\texttt{x}_{26}$" [> "R$_{27}$"]
+\begin{subfigure}{0.19\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    {},
+    "$\texttt{w}_{20}$" [> "R$_{25}$"']
+   },
+   {}
   }
@@
 %%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
+\begin{subfigure}{0.19\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'],
+   "$\texttt{w}_{20}$" [> "R$_{27}$"]
+  }
+ }
+};
+\end{tikzpicture}
+\caption{\autoref{fig:Multi-child-a} after Bob ... TODO}
+\label{fig:Multi-parent-c}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.19\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  a/{$\texttt{+}_{10}$} [> "Root$_{9}$"] -> {
+   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
     {},
-    "$\texttt{x}_{26}$" [> "R$_{29}$"']
+    b/"$\texttt{w}_{20}$" [> "R$_{25}$"']
    },
@@
   }
- }
-};
-\end{tikzpicture}
-\caption{\autoref{fig:Multi-child-a} after Bob ... TODO}
-\label{fig:Multi-parent-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{$\texttt{+}_{8}$} [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{24}$" [> "L$_{23}$"'] -> {
+ },
+};
+\path [edge style,->] (a) to ["R$_{27}$",out=-60,in=60] (b);
+\end{tikzpicture}
+\caption{The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}}
+\label{fig:Multi-parent-d}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}{0.19\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
     {},
-    b/"$\texttt{x}_{26}$" [> "R$_{29}$"']
+    "$\texttt{w}_{20}$" [> "R$_{25}$"']
    },
@@
   }
- },
-};
-\path [edge style,->] (a) to ["R$_{27}$",out=-60,in=60] (b);
-\end{tikzpicture}
-\caption{The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}}
-\label{fig:Multi-parent-d}
+ }
+};
+\end{tikzpicture}
+\caption{Initial code TODO: move to previous figure}
+\label{fig:Multi-parent-e}
 \end{subfigure}
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
+  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"'] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    {},
     {}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Single-User Cycles-a}
+\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
+\label{fig:Single-User Cycles-b}
 \end{subfigure}
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
-    {}
-   },
-   "$\texttt{+}_{18}$" [> "R$_{7}$"] -> {
-    "$\texttt{*}_{10}$" [> "L$_{19}$"'] -> {
-     "$\texttt{x}_{12}$" [> "L$_{19}$"'],
+  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"'] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+    {},
+    "$\texttt{+}_{30}$" [> "R$_{29}$"] -> {
+     "$\texttt{*}_{32}$" [> "L$_{31}$"'] -> {
+      {},
       {}
@@
     }
-  }
    },
+  }
+ }
 };
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{x}_{2}$" [> "L$_{19}$"'],
+  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'],
    {}
-   },
-   {}
-  }
- }
-};
-\path[edge style,->] (v2) to ["R$_{19}$",out=-60,in=-60] (v8);
+  }
+ }
+};
+\path[edge style,->] (v2) to ["R$_{29}$",out=-60,in=-60] (v8);
 \end{tikzpicture}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {

### 2020-10-05 grove-paper.tex
Add graph actions to formalism with Michael and Cyrus (Eric)

@@
 
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects the child position $p$ of a source vertex $v$ to a target vertex $v^\prime$.
+Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
+and connects the child position $p$ of a source vertex $v$ to a target vertex $v^\prime$.
 The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
@@
 A vertex $v = (u, k)$ identifies a distinct instance of the constructor $k$.
-Each constructor represents a different type of node in the target language's abstract syntax tree.
-
-A position $p$ identifies a location adjacent to some parent vertex.
+
+A constructor $k \in \K$ represents a different type of node in the target language's abstract syntax tree.
+
+A position $p \in \P$ identifies a location, adjacent to some parent vertex.
 
@@
 
-A graph action is a +E or -E.
-
-Graph action semantics is a transition system
-between graphs. (join with whatever it was before)
-
-G --action--> G'
+We define graph edit actions as a pair of an edge state, excluding $\bot$, and an edge.
+Thus, $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$.
+As a shorthand, we define $+e$ to be $(\Plus, e)$ and $-e$ to be $(\Minus, e)$.
+
+We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
+For example, $\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
+and $\Plus \sqcup \Minus = \Minus$, which means that once an edge is deleted it can never be restored.
+%smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.
+
+We define the semantics of graph actions via the following
+transition relation between graphs.
+\[
+G \overset{\left(s,e\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right\}
+\]
+In other words, applying graph action $\alpha = \left(s, e\right)$ to a graph $G$
+results in $G\left[e \mapsto s \sqcup G(e) \right]$ (i.e., the graph is updated at $e$ to be
+the join of $s$ with whatever was already at $e$).
+
+% Graph action semantics is a transition system between graphs.
+% (join with whatever it was before)
+
+% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
+% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
 
@@
 
+$G \overset{\left(e,s\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right]$
+
+% Edge.Map.add action.value.edge (join old_state new_state) graph
+
 \end{document}

### 2020-10-06 grove-paper.tex
Finish figure renumbering

@@
    {
-    "$\texttt{x}_{22}$" [> "R$_{21}$"]
+    "$\texttt{x}_{24}$" [> "R$_{23}$"]
    }
@@
    },
-   "$\texttt{y}_{24}$" [> "R$_{23}$"]
+   "$\texttt{y}_{26}$" [> "R$_{25}$"]
   }
@@
    {
-    "$\texttt{x}_{22}$" [> "R$_{21}$"],
-    "$\texttt{y}_{24}$" [> "R$_{23}$"]
+    "$\texttt{x}_{24}$" [> "R$_{23}$"],
+    "$\texttt{y}_{26}$" [> "R$_{25}$"]
    }
@@
     {},
-    "$\texttt{w}_{20}$" [> "R$_{25}$"']
+    "$\texttt{w}_{20}$" [> "R$_{27}$"']
    },
@@
    "$\texttt{*}_{2}$" [> "L$_{21}$"'],
-   "$\texttt{w}_{20}$" [> "R$_{27}$"]
+   "$\texttt{w}_{20}$" [> "R$_{29}$"]
   }
@@
     {},
-    b/"$\texttt{w}_{20}$" [> "R$_{25}$"']
+    b/"$\texttt{w}_{20}$" [> "R$_{27}$"']
    },
@@
 };
-\path [edge style,->] (a) to ["R$_{27}$",out=-60,in=60] (b);
+\path [edge style,->] (a) to ["R$_{29}$",out=-60,in=60] (b);
 \end{tikzpicture}
@@
 \caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
-\label{fig:Single-User Cycles-b}
+\label{fig:Single-User Cycles-a}
 \end{subfigure}
@@
     {},
-    "$\texttt{+}_{30}$" [> "R$_{29}$"] -> {
-     "$\texttt{*}_{32}$" [> "L$_{31}$"'] -> {
+    "$\texttt{+}_{32}$" [> "R$_{31}$"] -> {
+     "$\texttt{*}_{34}$" [> "L$_{33}$"'] -> {
       {},
@@
 };
-\path[edge style,->] (v2) to ["R$_{29}$",out=-60,in=-60] (v8);
+\path[edge style,->] (v2) to ["R$_{31}$",out=-60,in=-60] (v8);
 \end{tikzpicture}
@@
   "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
-    "$\texttt{+}_{4}$" [> "L$_{19}$"'],
-    "$\texttt{*}_{6}$" [> "R$_{19}$"]
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
+    "$\texttt{+}_{34}$" [> "L$_{33}$"'],
+    "$\texttt{*}_{36}$" [> "R$_{35}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
    },
    {
-    "$\texttt{+}_{4}$" [> "R$_{19}$"] -> {
-     "$\texttt{*}_{6}$" [> "L$_{19}$"'],
+    "$\texttt{+}_{34}$" [> "R$_{37}$"] -> {
+     "$\texttt{*}_{36}$" [> "L$_{39}$"'],
      {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
    },
    {
-    "$\texttt{*}_{6}$" [> "R$_{19}$"] -> {
-     "$\texttt{+}_{4}$" [> "L$_{19}$"'],
+    "$\texttt{*}_{36}$" [> "R$_{41}$"] -> {
+     "$\texttt{+}_{34}$" [> "L$_{43}$"'],
      {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
    },
    {
-    v1/"$\texttt{*}_{6}$" [> "R$_{19}$"],
-    v2/"$\texttt{+}_{4}$" [> "R$_{19}$"]
-   }
-  }
- }
-};
-\path[draw,->] (v1) to [out=-45,in=-135] (v2);
-\path[draw,->] (v2) to [out=-90,in=-90] (v1);
+    v1/"$\texttt{+}_{34}$" [> "R$_{37}$"],
+    v2/"$\texttt{*}_{36}$" [> "R$_{41}$"]
+   }
+  }
+ }
+};
+\path[edge style,->] (v1) to ["L$_{39}$",out=-45,in=-135] (v2);
+\path[edge style,->] (v2) to ["L$_{43}$",out=-90,in=-90] (v1);
 \end{tikzpicture}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'],
-   "$\texttt{*}_{6}$" [> "R$_{19}$"]
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'],
+   "$\texttt{*}_{36}$" [> "R$_{41}$"]
   }
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{19}$"'] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
     {},
-    "$\texttt{*}_{6}$" [> "R$_{19}$"]
+    "$\texttt{*}_{36}$" [> "R$_{45}$"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"] -> {
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
    {},
-   "$\texttt{*}_{6}$" [> "R$_{19}$"] -> {
-    "$\texttt{+}_{2}$" [> "L$_{19}$"'],
+   "$\texttt{*}_{36}$" [> "R$_{37}$"] -> {
+    "$\texttt{+}_{2}$" [> "L$_{47}$"'],
     {}
@@
  root[root style] -> {
-  "$\texttt{+}_{8}$" [> "Root$_{7}$"]
- }
-};
-\path (-1cm,-3cm) graph[graph style] {
- times/"$\texttt{*}_{6}$" [> "L$_{19}$"']
-};
-\path (1cm,-3cm) graph[graph style] {
- plus/"$\texttt{+}_{2}$" [> "L$_{19}$"']
-};
-\path[draw,->] (times) to [out=45,in=135] (plus);
-\path[draw,->] (plus) to [out=-135,in=-45] (times);
+  "$\texttt{+}_{10}$" [> "Root$_{9}$"]
+ }
+};
+\path (-1cm,-3cm) graph[graph style] { plus/"$\texttt{+}_{2}$" };
+\path ( 1cm,-3cm) graph[graph style] { times/"$\texttt{*}_{32}$" };
+\path[edge style,->] (times) to ["L$_{43}$"anchor=south,out=-135,in=-45] (plus);
+\path[edge style,->] (plus) to ["R$_{41}$"anchor=north,out=45,in=135] (times);
 \end{tikzpicture}

### 2020-10-08 grove-paper.tex
Get macros setup for vertex and edge names (Michael D. Adams)

@@
 
+% Calling \newvertex{Foo}{bar} defines
+%   \vidFoo to be a new id number, and
+%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
+\newcounter{NodeVertexCounter}
+\newcommand{\newvertex}[2]{%
+\ifodd\theNodeVertexCounter
+  \addtocounter{NodeVertexCounter}{1}%
+\else
+  \addtocounter{NodeVertexCounter}{2}%
+\fi
+\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
+\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
+\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
+\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\csname vid#1\endcsname}}}%
+}
+\newcommand{\newedge}[2]{%
+\ifodd\theNodeVertexCounter
+  \addtocounter{NodeVertexCounter}{2}%
+\else
+  \addtocounter{NodeVertexCounter}{1}%
+\fi
+\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
+\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
+\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
+\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\csname eid#1\endcsname}}}%
+}
+\setcounter{NodeVertexCounter}{-1}
+\newvertex{Root}{Root}
+
 % Support \includegraphics of .dot files
@@
 \centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+\newedge{SimpleTimes}{Root}%
+\newvertex{SimpleTimes}{*}%
+\newedge{SimpleX}{L}%
+\newvertex{SimpleX}{x}%
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   "\vSimpleX" [> "\eSimpleX"'],
    {}
@@
 \end{tikzpicture}
+TODO: labels for edge numbers
+
+TODO: sub figure with diagram of relation of subfigures
+
+TODO: move subcaptions into master caption
+
+TODO: move subcaptions above figure
+
+TODO: top align subfigures
 \caption{Graph representation}
@@
 \centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-   "$\texttt{y}_{6}$" [> "R$_{5}$"]
+\newedge{ConstructionY}{R}%
+\newvertex{ConstructionY}{y}%
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   "\vSimpleX" [> "\eSimpleX"'],
+   "\vConstructionY" [> "\eConstructionY"]
   }
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   "\vSimpleX" [> "\eSimpleX"'],
    {}
@@
 \centering
+\newedge{ChangeZ}{R}
+\newvertex{ChangeZ}{z}
 \begin{tikzpicture}[framed]
@@
  root[root style] -> {
-  "$\texttt{*}_{2}$" [> "Root$_{1}$"] -> {
-   "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-   "$\texttt{z}_{8}$" [> "R$_{7}$"]
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   "\vSimpleX" [> "\eSimpleX"'],
+   "\vChangeZ" [> "\eChangeZ"]
   }
@@
 \path (-0.5,-3cm) graph[graph style] {
- "$\texttt{*}_{2}$" -> {
-  "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-  "$\texttt{z}_{8}$" [> "R$_{7}$"]
+ "\vSimpleTimes" -> {
+  "\vSimpleX" [> "\eSimpleX"'],
+  "\vChangeZ" [> "\eChangeZ"]
  }
@@
 \centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"]
+\newedge{WrapPlus}{Root}
+\newvertex{WrapPlus}{+}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"]
  }
@@
 \path (-0.5cm,-3cm) graph[graph style] {
- "$\texttt{*}_{2}$" -> {
-  "$\texttt{x}_{4}$" [> "L$_{5}$"'],
-  "$\texttt{z}_{8}$" [> "R$_{7}$"]
+ "\vSimpleTimes" -> {
+  "\vSimpleX" [> "\eSimpleX"'],
+  "\vChangeZ" [> "\eChangeZ"]
  }
@@
 \centering
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{11}$"'] -> {
-    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-    "$\texttt{z}_{8}$" [> "R$_{7}$"]
+\newedge{WrapTimes}{L}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eWrapTimes"'] -> {
+    "\vSimpleX" [> "\eSimpleX"'],
+    "\vChangeZ" [> "\eChangeZ"]
    },
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"]
+  "\vWrapPlus" [> "\eWrapPlus"]
  }
@@
 \path (0cm,-3cm) graph [graph style] {
- "$\texttt{*}_{2}$" -> {
-  "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-  "$\texttt{z}_{8}$" [> "R$_{7}$"]
+ "\vSimpleTimes" -> {
+  "\vSimpleX" [> "\eSimpleX"'],
+  "\vChangeZ" [> "\eChangeZ"]
  }
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
-    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
-    "$\texttt{z}_{6}$" [> "R$_{7}$"]
+   "\vSimpleTimes" [> "R$_{13}$"] -> {
+    "\vSimpleX" [> "\eSimpleX"'],
+    "$\texttt{z}_{6}$" [> "\eChangeZ"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "R$_{13}$"] -> {
     "$\texttt{u}_{16}$" [> "L$_{15}$"'],
-    "$\texttt{z}_{6}$" [> "R$_{7}$"]
+    "$\texttt{z}_{6}$" [> "\eChangeZ"]
    }
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
-    "$\texttt{x}_{4}$" [> "L$_{3}$"'],
+   "\vSimpleTimes" [> "R$_{13}$"] -> {
+    "\vSimpleX" [> "\eSimpleX"'],
     "$\texttt{v}_{18}$" [> "R$_{17}$"]
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "R$_{13}$"] -> {
     "$\texttt{u}_{16}$" [> "L$_{15}$"'],
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},
-   "$\texttt{*}_{2}$" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "R$_{13}$"] -> {
     "$\texttt{w}_{20}$" [> "L$_{19}$"'],
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "L$_{21}$"'] -> {
     "$\texttt{u}_{16}$" [> "L$_{15}$"'],
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "L$_{21}$"'] -> {
     "$\texttt{w}_{20}$" [> "L$_{19}$"'],
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "L$_{21}$"'] -> {
     {},
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{21}$"'],
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "L$_{21}$"'],
    "$\texttt{w}_{20}$" [> "R$_{29}$"]
@@
  root[root style] -> {
-  a/{$\texttt{+}_{10}$} [> "Root$_{9}$"] -> {
-   "$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "L$_{21}$"'] -> {
     {},
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
+   v2/"\vSimpleTimes" [> "L$_{21}$"'] -> {
     {},
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"'] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'] -> {
+  v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
+   v2/"\vSimpleTimes" [> "L$_{21}$"'] -> {
     {},
@@
  root[root style] -> {
-  v8/"$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
-   v2/"$\texttt{*}_{2}$" [> "L$_{21}$"'],
+  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
+   v2/"\vSimpleTimes" [> "L$_{21}$"'],
    {}
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    "$\texttt{+}_{2}$" [> "L$_{21}$"'],
@@
  root[root style] -> {
-  "$\texttt{+}_{10}$" [> "Root$_{9}$"] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
    {},

### 2020-10-13 grove-paper.tex
start math for edit action syntax and semantics (Eric)

@@
 \mathlig{-->a}{\overset{a}{\longrightarrow}}
-%\mathlig{||}{|}
-%\mathlig{|}{:}
-%\mathlig{:}{\!:\!}
 \newcommand{\abs}[1]{\left|{#1}\right|}
@@
 \def\_{\texttt{\textunderscore}}
+
+\newcommand{\conflict}[2]{\llparenthesis#1|#2\rrparenthesis}
 
@@
 \def\A{\mathcal{A}}
-\def\C{\mathcal{C}}
+%\def\C{\mathcal{C}}
 \def\E{\mathcal{E}}
@@
 \def\P{\mathcal{P}}
-\def\S{\mathcal{S}}
 \def\U{\mathcal{U}}
@@
 % relations
+\DeclareMathOperator{\s}{\text{S}}
+\DeclareMathOperator{\MP}{\text{MP}}
+\DeclareMathOperator{\CC}{\text{C}}
+\DeclareMathOperator{\D}{\text{D}}
+
 \DeclareMathOperator{\arity}{\text{arity}}
@@
 You can map from syntax to graph by selecting
-unique IDs (for both edges and vertexes).
+unique IDs $u$ (for both edges and vertexes).
     - holes are not explicit in the graph
@@
 Mapping to graph actions
+
+\[
+\begin{array}{l}
+  \textrm{Edit Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}} \\
+  \s : \G \to Act \times \V^{*} \times \V^{*} \times \V^{*} \\
+  \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G)) \\
+  \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right. \\
+      \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\} \\
+  \CC(G) = \ldots \\
+  \D(G) = \orphans(G) \setminus \left\{ v_0 \right\} \\
+  \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right. \\
+      \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\} \\
+  \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
+  \edges(G) = \left\{ e \mid G(e) \ne \bot \right\} \\
+  \liveEdges(G) = \left\{ e \mid G(e) = + \right\} \\
+\end{array}
+\]
 
@@
   $\A$ & Actions & $\S$ & Edge States   \\
-  $\C$ & Cursors & $\U$ & UUIDs         \\
+%  $\C$ & Cursors & $\U$ & UUIDs         \\
   $\E$ & Edges   & $\V$ & Vertexes      \\
@@
 
-A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
+%A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
 Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.
@@
 
-An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
+%An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
 
@@
 
-A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
+%A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
 
@@
 
-An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
+%An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
 

### 2020-10-14 grove-paper.tex
Commit yesterdays work (Michael D. Adams)

@@
 \usepackage{array}
+\usepackage{todonotes}
 
@@
 
-\marginpar{START REV HERE}
-
 \subsection{Single User Edits}
@@
 
-Each user edit that we will consider corresponds to one or more graph edits.
+Each user edit that we consider corresponds to one or more graph edits.
 Graph edits can only add fresh edges and vertices or remove existing edges.
@@
 \centering
-TODO
+TODO: remove this figure
 \caption{User interface}
@@
 
-TODO: "user edits" and "graph edits"
-
-To start our examples, Alice starts with the code in \autoref{fig:A simple tree with a hole}.
-She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
-We model edits as sequences of additions and deletions of edges to the graph.
-Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
-to that vertex from \Vertex{2}.
-\autoref{fig:Construction-b} shows the result, where the new vertex has identifier~6.
-This graph represents the expression \texttt{x * y}.
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-TODO: z -> foo (or multi-editor)
-
-% TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)
-
-% TODO: remove cursor from Fig 1 and Fig 2
-
-\subsubsection{Changes}
-\label{sub:Changes}
+\todo{"user edits" and "graph edits"}
+To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole} and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
+This edit corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly created vertex \Vertex{\vidConstructionY} containing the variable reference \texttt{y}.
+The resulting graph represents the expression \texttt{x * y}.
+
+\subsubsection{Deletion}
+\label{sub:Deletion}
 
@@
 \centering
+TODO: delete and leave deleted z8
+
+TODO: delete "x" instead of "y"
 \begin{tikzpicture}
@@
 \centering
+TODO: Delete this figure
 \newedge{ChangeZ}{R}
@@
 \end{figure}
-
-Arrows between figures
 
@@
 
-Next Alice wants to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
-This is modeled by first deleting \Edge{5}.
-This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
-After that deletion, we add a new vertex to represent~\texttt{z} and an
-edge from \Vertex{2} to that vertex.
-This results in the bottom graph in \autoref{fig:Change-b},
-which represents the expression~\texttt{x * z}.
-
-Note that this edge has a new identifier distinct from the one before.
-In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
-Instead a new edge with a new identifier is created.
-
-Note that \Vertex{3} continues to exist, and if it had any children, those children could remain connected to it.
+\todo{Arrows between figures}
+Next Alice decides to delete~\texttt{x} so that the code becomes~\texttt{\hole  * y}.
+This is modeled by deleting \Edge{\eidConstructionY} as shown in \autoref{fig:Change-a}. 
+Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children could remain connected to it.
 In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertexes as shown later in this section.
-
-Note that we will omit orphans from diagrams if they are not relevant
-(e.g., they are deleted and are not added back later)
-
-Note that we do not explicitly delete or add vertexes.
-Our only actions are deleting or adding \textit{edges}.
-Vertexes are implicitly created whenever an edge
-is created that points to it.
-Deleting vertexes is not necessary, as we can simply
-remove the edge pointing to it.
-
-TODO: explain that multiple delete or add are idempotent
-
-TODO: explain that delete overrides add
-
-TODO: once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)
-
-TODO: Explain orphans
+This allows further manipulations of those vertices by other users as shown later in this section.
+Note that we omit such orphans from the remaining diagrams if they are not relevant to the exposition.
+
+Once an edge with a particular identifier is deleted, it cannot be recreated.
+Thus if Alice performed an "undo" on this deletion, Grove would create a fresh edge between \Vertex{TODO} and \Vertex{TODO}.
+
+% Note that we do not explicitly add vertexes.
+% Our only actions are deleting or adding \textit{edges}.
+% Vertexes are implicitly created whenever an edge
+% is created that points to it.
+% Deleting vertexes is not necessary, as we can simply
+% remove the edge pointing to it.
+
+% TODO: explain that multiple delete or add are idempotent
+
+% TODO: explain that delete overrides add
 
@@
 
+\marginpar{START REV HERE}
+
 Next, Alice types~\texttt{+} when the cursor is on \Vertex{2}.
@@
 \centering
+\newedge{MoveTimes}{R}
 \begin{tikzpicture}
@@
    {},
-   "\vSimpleTimes" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
     "\vSimpleX" [> "\eSimpleX"'],
-    "$\texttt{z}_{6}$" [> "\eChangeZ"]
+    "\vChangeZ" [> "\eChangeZ"]
    }
@@
    {},
-   "\vSimpleTimes" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
     "$\texttt{u}_{16}$" [> "L$_{15}$"'],
-    "$\texttt{z}_{6}$" [> "\eChangeZ"]
+    "\vChangeZ" [> "\eChangeZ"]
    }
@@
    {},
-   "\vSimpleTimes" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
     "\vSimpleX" [> "\eSimpleX"'],
@@
    {},
-   "\vSimpleTimes" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
     "$\texttt{u}_{16}$" [> "L$_{15}$"'],
@@
    {},
-   "\vSimpleTimes" [> "R$_{13}$"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
     "$\texttt{w}_{20}$" [> "L$_{19}$"'],
@@
 
+\todo{UI would show orphans only once edits orphans occur}
+
 Finally, when multi-user edits are merged,
@@
 TODO: point of this section is commutativity theorem
+
+TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
 

### 2020-10-15 grove-paper.tex
Partially fix vertex and edge data in figures

@@
 \centering
+\newedge{DifferentAlice}{L}
+\newvertex{DifferentAlice}{u}
 \begin{tikzpicture}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
+    "\vDifferentAlice" [> "\eDifferentAlice"'],
     "\vChangeZ" [> "\eChangeZ"]
@@
 \centering
+\newedge{DifferentBob}{R}
+\newvertex{DifferentBob}{v}
 \begin{tikzpicture}
@@
     "\vSimpleX" [> "\eSimpleX"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+    "\vDifferentBob" [> "\eDifferentBob"]
    }
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+    "\vDifferentAlice" [> "\eDifferentAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    }
@@
 \centering
+\newedge{NestedAlice}{L}
+\newvertex{NestedAlice}{w}
 \begin{tikzpicture}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    }
@@
 \centering
+\newedge{NestedBob}{L}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{u}_{16}$" [> "L$_{15}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vDifferentAlice" [> "\eDifferentAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    },
@@
 \centering
+\newedge{MultiChildAlice}{R}
+\newvertex{MultiChildAlice}{x}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    },
    {
-    "$\texttt{x}_{24}$" [> "R$_{23}$"]
+    "\vMultiChildAlice" [> "\eMultiChildAlice"]
    }
@@
 \centering
+\newedge{MultiChildBob}{R}
+\newvertex{MultiChildBob}{y}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    },
-   "$\texttt{y}_{26}$" [> "R$_{25}$"]
+   "\vMultiChildBob" [> "\eMultiChildBob"]
   }
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
-    "$\texttt{v}_{18}$" [> "R$_{17}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
    },
    {
-    "$\texttt{x}_{24}$" [> "R$_{23}$"],
-    "$\texttt{y}_{26}$" [> "R$_{25}$"]
+    "\vMultiChildAlice" [> "\eMultiChildAlice"],
+    "\vMultiChildBob" [> "\eMultiChildBob"]
    }
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
-    "$\texttt{w}_{20}$" [> "L$_{19}$"'],
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
     {}
@@
 \centering
+\newedge{MultiParentAlice}{R}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
-    "$\texttt{w}_{20}$" [> "R$_{27}$"']
+    "\vNestedAlice" [> "\eMultiParentAlice"']
    },
@@
 \centering
+\newedge{MultiParentBob}{R}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'],
-   "$\texttt{w}_{20}$" [> "R$_{29}$"]
+   "\vSimpleTimes" [> "\eNestedBob"'],
+   "\vNestedAlice" [> "\eMultiParentBob"]
   }
@@
   a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "L$_{21}$"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
-    b/"$\texttt{w}_{20}$" [> "R$_{27}$"']
+    b/"\vNestedAlice" [> "\eMultiParentAlice"']
    },
@@
 };
-\path [edge style,->] (a) to ["R$_{29}$",out=-60,in=60] (b);
+\path [edge style,->] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
 \end{tikzpicture}
@@
   v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "L$_{21}$"'] -> {
+   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
-    "$\texttt{w}_{20}$" [> "R$_{27}$"']
+    "\vNestedAlice" [> "\eMultiParentAlice"']
    },
@@
   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-   v2/"\vSimpleTimes" [> "L$_{21}$"'] -> {
+   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
@@
 \centering
+\newedge{SingleCyclePlus}{R}
+\newvertex{SingleCyclePlus}{+}
+\newedge{SingleCycleTimes}{L}
+\newvertex{SingleCycleTimes}{*}
 \begin{tikzpicture}
@@
   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-   v2/"\vSimpleTimes" [> "L$_{21}$"'] -> {
+   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
-    "$\texttt{+}_{32}$" [> "R$_{31}$"] -> {
-     "$\texttt{*}_{34}$" [> "L$_{33}$"'] -> {
+    "\vSingleCyclePlus" [> "\eSingleCyclePlus"] -> {
+     "\vSingleCycleTimes" [> "\eSingleCycleTimes"'] -> {
       {},
@@
 \centering
+\newedge{SingleCycleCycle}{R}
 \begin{tikzpicture}
@@
   v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "L$_{21}$"'],
+   v2/"\vSimpleTimes" [> "\eNestedBob"'],
    {}
@@
 };
-\path[edge style,->] (v2) to ["R$_{31}$",out=-60,in=-60] (v8);
+\path[edge style,->] (v2) to ["\eSingleCycleCycle",out=-60,in=-60] (v8);
 \end{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
     "$\texttt{+}_{34}$" [> "L$_{33}$"'],
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{21}$"'],
+   "\vSimpleTimes" [> "\eNestedBob"'],
    "$\texttt{*}_{36}$" [> "R$_{41}$"]
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "$\texttt{+}_{2}$" [> "L$_{21}$"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
     {},
@@
    "$\texttt{*}_{36}$" [> "R$_{37}$"] -> {
-    "$\texttt{+}_{2}$" [> "L$_{47}$"'],
+    "\vSimpleTimes" [> "L$_{47}$"'],
     {}
@@
 };
-\path (-1cm,-3cm) graph[graph style] { plus/"$\texttt{+}_{2}$" };
+\path (-1cm,-3cm) graph[graph style] { plus/"\vSimpleTimes" };
 \path ( 1cm,-3cm) graph[graph style] { times/"$\texttt{*}_{32}$" };

### 2020-10-16 grove-paper.tex
Finish renumbering vertexes and edges in figures

@@
 \centering
+\newedge{MultiCyclePlus}{L}
+\newvertex{MultiCyclePlus}{+}
+\newedge{MultiCycleTimes}{R}
+\newvertex{MultiCycleTimes}{*}
 \begin{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "$\texttt{+}_{34}$" [> "L$_{33}$"'],
-    "$\texttt{*}_{36}$" [> "R$_{35}$"]
+    "\vMultiCyclePlus" [> "\eMultiCyclePlus"'],
+    "\vMultiCycleTimes" [> "\eMultiCycleTimes"]
    },
@@
 \centering
+\newedge{MultiCycleAlicePlus}{R}
+\newedge{MultiCycleAliceTimes}{L}
 \begin{tikzpicture}
@@
    {
-    "$\texttt{+}_{34}$" [> "R$_{37}$"] -> {
-     "$\texttt{*}_{36}$" [> "L$_{39}$"'],
+    "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"] -> {
+     "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"'],
      {}
@@
 \centering
+\newedge{MultiCycleBobTimes}{R}
+\newedge{MultiCycleBobPlus}{L}
 \begin{tikzpicture}
@@
    {
-    "$\texttt{*}_{36}$" [> "R$_{41}$"] -> {
-     "$\texttt{+}_{34}$" [> "L$_{43}$"'],
+    "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"] -> {
+     "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"'],
      {}
@@
    {
-    v1/"$\texttt{+}_{34}$" [> "R$_{37}$"],
-    v2/"$\texttt{*}_{36}$" [> "R$_{41}$"]
-   }
-  }
- }
-};
-\path[edge style,->] (v1) to ["L$_{39}$",out=-45,in=-135] (v2);
-\path[edge style,->] (v2) to ["L$_{43}$",out=-90,in=-90] (v1);
+    v1/"\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"],
+    v2/"\vMultiCycleTimes" [> "\eMultiCycleBobTimes"]
+   }
+  }
+ }
+};
+\path[edge style,->] (v1) to ["\eMultiCycleAliceTimes",out=-45,in=-135] (v2);
+\path[edge style,->] (v2) to ["\eMultiCycleBobPlus",out=-90,in=-90] (v1);
 \end{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedBob"'],
-   "$\texttt{*}_{36}$" [> "R$_{41}$"]
+   "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"]
   }
@@
 \centering
+\newedge{DisconnectionAlice}{R}
 \begin{tikzpicture}
@@
     {},
-    "$\texttt{*}_{36}$" [> "R$_{45}$"]
+    "\vMultiCycleTimes" [> "\eDisconnectionAlice"]
    },
@@
 \centering
+\newedge{DisconnectionBob}{L}
 \begin{tikzpicture}
@@
    {},
-   "$\texttt{*}_{36}$" [> "R$_{37}$"] -> {
-    "\vSimpleTimes" [> "L$_{47}$"'],
+   "\vMultiCycleTimes" [> "\eMultiCycleAlicePlus"] -> {
+    "\vSimpleTimes" [> "\eDisconnectionBob"'],
     {}
@@
 \path (-1cm,-3cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 1cm,-3cm) graph[graph style] { times/"$\texttt{*}_{32}$" };
-\path[edge style,->] (times) to ["L$_{43}$"anchor=south,out=-135,in=-45] (plus);
-\path[edge style,->] (plus) to ["R$_{41}$"anchor=north,out=45,in=135] (times);
+\path ( 1cm,-3cm) graph[graph style] { times/"\vMultiCycleTimes" };
+\path[edge style,->] (times) to ["\eMultiCycleBobPlus"anchor=south,out=-135,in=-45] (plus);
+\path[edge style,->] (plus) to ["\eMultiCycleBobTimes"anchor=north,out=45,in=135] (times);
 \end{tikzpicture}

### 2020-10-20 grove-paper.tex
Move subcaptions to top and top align subfigures. (Michael D. Adams)
Move subcaptions into main captions (Michael D. Adams)

@@
 \tikzset{
-  every picture/.style={framed, background rectangle/.style={draw=gray!50}},
+  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
 }
@@
 %% Title information
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
+\title[Grove]{Grove: A Convergent Collaborative Structure-Editor~Calculus}
 %\subtitle{Subtitle}
@@
 %%%%%%%%
-\begin{subfigure}{0.49\linewidth}
-\centering
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:A simple tree with a hole-a}
 TODO
-\caption{User interface}
-\label{fig:A simple tree with a hole-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.49\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:A simple tree with a hole-b}
 \newedge{SimpleTimes}{Root}%
@@
 \end{tikzpicture}
+
 TODO: labels for edge numbers
@@
 TODO: move subcaptions into master caption
-
-TODO: move subcaptions above figure
-
-TODO: top align subfigures
-\caption{Graph representation}
-\label{fig:A simple tree with a hole-b}
 \end{subfigure}
@@
 \hfill{}
-\caption{A simple tree with a hole}
+\caption{A simple tree with a hole.
+\autoref{fig:A simple tree with a hole-a}:~User interface.
+\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
 \label{fig:A simple tree with a hole}
 \end{figure}
-
-TODO: use mathit instead of var(x)
 
@@
 %%%%%%%%
-\begin{subfigure}{0.49\columnwidth}
-\centering
+\begin{subfigure}[t]{0.49\columnwidth}
+\centering
+\caption{}
+\label{fig:Construction-a}
 TODO: remove this figure
-\caption{User interface}
-\label{fig:Construction-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.49\columnwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.49\columnwidth}
+\centering
+\caption{}
+\label{fig:Construction-b}
 \newedge{ConstructionY}{R}%
@@
 \end{tikzpicture}
-\caption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
-\label{fig:Construction-b}
 \end{subfigure}
@@
 \hfill{}
-\caption{Construction}
+\caption{Construction.
+\autoref{fig:Construction-a}:~User interface.
+\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}.}
 \label{fig:Construction}
@@
 %%%%%%%%
-\begin{subfigure}{0.49\linewidth}
-\centering
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:Change-a}
 TODO: delete and leave deleted z8
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Construction-b} after deleting \Edge{5}}
-\label{fig:Change-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.49\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:Change-b}
 TODO: Delete this figure
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}}
-\label{fig:Change-b}
 \end{subfigure}
@@
 \hfill{}
-\caption{Change}
+\caption{Change.
+\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{5}.
+\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}.}
 \label{fig:Change}
@@
 %%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Wrapping-a}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Change-b} after deleting \Edge{1}}
-\label{fig:Wrapping-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Wrapping-b}
 \newedge{WrapPlus}{Root}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-a} after adding \Vertex{8} and \Edge{7}}
-\label{fig:Wrapping-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Wrapping-c}
 \newedge{WrapTimes}{L}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-b} after adding \Edge{9}}
-\label{fig:Wrapping-c}
 \end{subfigure}
@@
 \hfill{}
-\caption{Wrapping}
+\caption{Wrapping.
+\autoref{fig:Wrapping-a}: \autoref{fig:Change-b} after deleting \Edge{1}.
+\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{8} and \Edge{7}.
+\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{9}.
+}
 \label{fig:Wrapping}
@@
 %%%%%%%%
-\begin{subfigure}{0.49\linewidth}
-\centering
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:Move-a}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Wrapping-c} after deleting \Edge{9}}
-\label{fig:Move-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.45\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.45\linewidth}
+\centering
+\caption{}
+\label{fig:Move-b}
 \newedge{MoveTimes}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-a} after adding \Edge{11}}
-\label{fig:Move-b}
 \end{subfigure}
@@
 \hfill{}
-\caption{Moving \Vertex{2} from the left child to the right child of \Vertex{8}}
+\caption{Moving \Vertex{2} from the left child to the right child of \Vertex{8}.
+\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{9}.
+\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{11}.}
 \label{fig:Move}
@@
 %%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Different Parts of the Code-a}
 \newedge{DifferentAlice}{L}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-b} after Alice replaces \Vertex{4} with \Vertex{12}}
-\label{fig:Editing Different Parts of the Code-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Different Parts of the Code-b}
 \newedge{DifferentBob}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Move-b} after Bob replaces \Vertex{6} with \Vertex{14}}
-\label{fig:Editing Different Parts of the Code-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Different Parts of the Code-c}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{The result of merging \autoref{fig:Editing Different Parts of the Code-a} with \autoref{fig:Editing Different Parts of the Code-b}}
-\label{fig:Editing Different Parts of the Code-c}
 \end{subfigure}
@@
 \hfill{}
-\caption{Editing Different Parts of the Code}
+\caption{Editing Different Parts of the Code.
+\autoref{fig:Editing Different Parts of the Code-a}:~\autoref{fig:Move-b} after Alice replaces \Vertex{4} with \Vertex{12}.
+\autoref{fig:Editing Different Parts of the Code-b}:~\autoref{fig:Move-b} after Bob replaces \Vertex{6} with \Vertex{14}.
+\autoref{fig:Editing Different Parts of the Code-c}:~The
+result of merging \autoref{fig:Editing Different Parts of the Code-a}
+with \autoref{fig:Editing Different Parts of the Code-b}.}
 \label{fig:Editing Different Parts of the Code}
@@
 %%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Nested Parts of the Code-a}
 \newedge{NestedAlice}{L}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}}
-\label{fig:Editing Nested Parts of the Code-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Nested Parts of the Code-b}
 \newedge{NestedBob}{L}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}}
-\label{fig:Editing Nested Parts of the Code-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Editing Nested Parts of the Code-c}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}}
-\label{fig:Editing Nested Parts of the Code-c}
 \end{subfigure}
@@
 \hfill{}
-\caption{Editing Nested Parts of the Code}
+\caption{Editing Nested Parts of the Code.
+\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}.
+\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}.
+\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
 \label{fig:Editing Nested Parts of the Code}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
+\begin{subfigure}[t]{0.24\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-child-a}
 \newedge{MultiChildAlice}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO}
-\label{fig:Multi-child-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-child-b}
 \newedge{MultiChildBob}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO}
-\label{fig:Multi-child-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-child-c}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
-\label{fig:Multi-child-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-child-d}
 TODO: show UI
-\caption{User interface showing the multi-child conflict in \autoref{fig:Multi-child-d}}
-\label{fig:Multi-child-d}
 \end{subfigure}
@@
 \hfill{}
-\caption{Multi-child conflicts}
+\caption{Multi-child conflicts.
+\autoref{fig:Multi-child-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO.
+\autoref{fig:Multi-child-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO.
+\autoref{fig:Multi-child-c}:~The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}.
+\autoref{fig:Multi-child-d}:~User interface showing the multi-child conflict in \autoref{fig:Multi-child-c}.}
 \label{fig:Multi-child}
@@
 %%%%%%%%
-\begin{subfigure}{0.19\linewidth}
-\centering
+\begin{subfigure}[t]{0.19\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-a}
 \begin{tikzpicture}
@@
 TODO: UI (at least moc up)
-\caption{Initial code}
-\label{fig:Multi-parent-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.19\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.19\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-b}
 \newedge{MultiParentAlice}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Multi-parent-a} after Alice ... TODO}
-\label{fig:Multi-parent-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.19\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.19\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-c}
 \newedge{MultiParentBob}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Multi-child-a} after Bob ... TODO}
-\label{fig:Multi-parent-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.19\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.19\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-d}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}}
-\label{fig:Multi-parent-d}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.19\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.19\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-e}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{Initial code TODO: move to previous figure}
-\label{fig:Multi-parent-e}
 \end{subfigure}
@@
 \hfill{}
-\caption{Multi-parent conflicts}
+\caption{Multi-parent conflicts.
+\autoref{fig:Multi-parent-a}:~Initial code.
+\autoref{fig:Multi-parent-b}:~\autoref{fig:Multi-parent-a} after Alice ... TODO.
+\autoref{fig:Multi-parent-c}:~\autoref{fig:Multi-child-a} after Bob ... TODO.
+\autoref{fig:Multi-parent-d}:~The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}.
+\autoref{fig:Multi-parent-e}:~Initial code TODO: move to previous figure.}
 \label{fig:Multi-parent}
@@
 %%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Single-User Cycles-a}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
-\label{fig:Single-User Cycles-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Single-User Cycles-b}
 \newedge{SingleCyclePlus}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
-\label{fig:Single-User Cycles-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.32\linewidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{}
+\label{fig:Single-User Cycles-c}
 \newedge{SingleCycleCycle}{R}
@@
 \end{tikzpicture}
-\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by reference as the right child of \Vertex{TODO}}
-\label{fig:Single-User Cycles-c}
 \end{subfigure}
@@
 \hfill{}
-\caption{Single-User Cycles}
+\caption{Single-User Cycles.
+\autoref{fig:Single-User Cycles-a}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}.
+\autoref{fig:Single-User Cycles-b}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}.
+\autoref{fig:Single-User Cycles-c}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by reference as the right child of \Vertex{TODO}.}
 \label{fig:Single-User Cycles}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Multi-User Cycles-a}
 \newedge{MultiCyclePlus}{L}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Multi-User Cycles-b}
 \newedge{MultiCycleAlicePlus}{R}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Multi-User Cycles-c}
 \newedge{MultiCycleBobTimes}{R}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Multi-User Cycles-d}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Multi-User Cycles-d}
 \end{subfigure}
@@
 \hfill{}
-\caption{Multi-User Cycles}
+\caption{Multi-User Cycles
+\autoref{fig:Multi-User Cycles-a}:~Initial code.
+\autoref{fig:Multi-User Cycles-b}:~TODO.
+\autoref{fig:Multi-User Cycles-c}:~TODO.
+\autoref{fig:Multi-User Cycles-d}:~TODO.}
 \label{fig:Multi-User Cycles}
@@
 %%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Disconnection-a}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Disconnection-a}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Disconnection-b}
 \newedge{DisconnectionAlice}{R}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Disconnection-b}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Disconnection-c}
 \newedge{DisconnectionBob}{L}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Disconnection-c}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}{0.24\textwidth}
-\centering
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.24\textwidth}
+\centering
+\caption{}
+\label{fig:Disconnection-d}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\caption{Initial code}
-\label{fig:Disconnection-d}
 \end{subfigure}
@@
 \hfill{}
-\caption{Disconnection}
+\caption{Disconnection.
+\autoref{fig:Disconnection-a}:~Initial code.
+\autoref{fig:Disconnection-b}:~TODO.
+\autoref{fig:Disconnection-c}:~TODO.
+\autoref{fig:Disconnection-d}:~TODO.}
 \label{fig:Disconnection}
@@
 
-Mapping to graph actions
+Mapping to b actions
 

### 2020-10-28 grove-paper.tex
Commit yesterday's changes

@@
 %% Title information
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor~Calculus}
+% \title[Grove]{Convergent Collaborative Structure Editing}
+\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
 %\subtitle{Subtitle}
@@
 We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
-\autoref{sub:Single User Edits} explores edits by a single user, Alice,
+\autoref{sub:Single-User Actions} explores edits by a single user, Alice,
 and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
@@
 TODO: move subcaptions into master caption
+
+TODO: colors for different editors
+
+TODO: label edges of transitions between graphs
+
+TODO: Remove subfigure captions
 \end{subfigure}
@@
 Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \Vertex{2} and \Vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
 The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{2}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{\vidSimpleTimes}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 
-\subsection{Single User Edits}
-\label{sub:Single User Edits}
+\subsection{Single-User Actions}
+\label{sub:Single-User Actions}
 
 In the remainder of this section,
-we will consider user edits, starting
+we will consider user actions, starting
 in this subsection with single-user
-edits and continuing in \autoref{sub:Multi-user Edits}
-with multi-user edits.
-
-Each user edit that we consider corresponds to one or more graph edits.
+actions and continuing in \autoref{sub:Multi-user Edits}
+with multi-user actions.
+
+Each user action that we consider corresponds to one or more graph edits.
 Graph edits can only add fresh edges and vertices or remove existing edges.
 
-User edits act relative to a cursor. 
+User actions act relative to a cursor. 
 We do not model cursor sharing and the cursor does not appear in the underlying graph.
@@
 \autoref{fig:Construction-a}:~User interface.
-\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}.}
+\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
 \label{fig:Construction}
@@
 
-\todo{"user edits" and "graph edits"}
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole} and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
-This edit corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly created vertex \Vertex{\vidConstructionY} containing the variable reference \texttt{y}.
-The resulting graph represents the expression \texttt{x * y}.
+\todo{"user actions" and "graph edits"}
+To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole}
+and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
+This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
+created vertex, \Vertex{\vidConstructionY}, containing the variable reference \texttt{y}.
+The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.
 
@@
 TODO: delete "x" instead of "y"
+
 \begin{tikzpicture}
@@
 \caption{Change.
-\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{5}.
-\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}.}
+\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{\eidConstructionY}.
+\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{\vidChangeZ} and \Edge{\eidChangeZ}.}
 \label{fig:Change}
@@
 \caption{Wrapping.
-\autoref{fig:Wrapping-a}: \autoref{fig:Change-b} after deleting \Edge{1}.
-\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{8} and \Edge{7}.
-\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{9}.
-}
+\autoref{fig:Wrapping-a}:~\autoref{fig:Change-b} after deleting \Edge{\eidSimpleTimes}.
+\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{\vidWrapPlus} and \Edge{\eidWrapPlus}.
+\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{\eidWrapTimes}.}
 \label{fig:Wrapping}
@@
 
-Next, Alice types~\texttt{+} when the cursor is on \Vertex{2}.
-If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
-as a child of \Vertex{0}.
-However, this position is already occupied by some code.
-
-A naive editor could add the~\texttt{+} by creating a new
-edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
+TODO: revise: "user action" "graph edit"
+
+Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
+If this were a hole, this would cause a construction edit to be performed as in Sec.TODO above.
+However, this position is already occupied by the \texttt{*} expression, 
+so instead, typing~\texttt{+} performs a wrapping edit.
+This corresponds to the following sequence of graph actions, shown in \autoref{fig:Wrapping}: 
+(a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
+(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex,
+(c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
+(The choice of wrapping with a bias for the left child position is arbitrary.)
+
+% A naive editor could add the~\texttt{+} as a new child of the root.
+% However,
+
+by creating a new
+edge from the \texttt{L} index of the \texttt{*} to a freshly created \texttt{plus} vertex.
 This would result in a conflict between the \texttt{+} and the existing \texttt{var}
 as shown in Figure~REF:TODO.
+
+You can have a second edge,
+we discuss conflicts in section TODO
+In the single-user setting, we don't want to create conflicts, so instead we perform a wrapping action.
 
@@
 \hfill{}
-\caption{Moving \Vertex{2} from the left child to the right child of \Vertex{8}.
-\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{9}.
-\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{11}.}
+\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.
+\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{\eidWrapTimes}.
+\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{\eidMoveTimes}.}
 \label{fig:Move}
@@
 \autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
-First, \Edge{7} is deleted, which results in \autoref{fig:Move-a}
+First, \Edge{\eidWrapPlus} is deleted, which results in \autoref{fig:Move-a}
 where the rooted code is now~\texttt{\hole{} + \hole{}},
-and \Vertex{2} is orphaned.
-Next \Edge{11} is added, which results in the bottom graph
+and \Vertex{\vidSimpleTimes} is orphaned.
+Next \Edge{\eidMoveTimes} is added, which results in the bottom graph
 where the rooted code is now~\texttt{\hole{} + 1 * 2},
-and \Vertex{2} is no longer orphaned.
+and \Vertex{\vidSimpleTimes} is no longer orphaned.
 
@@
 \caption{Editing Different Parts of the Code.
-\autoref{fig:Editing Different Parts of the Code-a}:~\autoref{fig:Move-b} after Alice replaces \Vertex{4} with \Vertex{12}.
-\autoref{fig:Editing Different Parts of the Code-b}:~\autoref{fig:Move-b} after Bob replaces \Vertex{6} with \Vertex{14}.
-\autoref{fig:Editing Different Parts of the Code-c}:~The
+\textbf{(a)}:~\autoref{fig:Move-b} after Alice replaces \Vertex{\vidSimpleX} with \Vertex{\vidDifferentAlice}.
+\textbf{(b)}:~\autoref{fig:Move-b} after Bob replaces \Vertex{\vidConstructionY} with \Vertex{\vidDifferentBob}.
+\textbf{(c)}:~The
 result of merging \autoref{fig:Editing Different Parts of the Code-a}
@@
 For Alice, this involves deleting \Edge{3} and creating \Edge{13}.
-For Bob, this involves deleting \Edge{5} and creating \Edge{15}.
+For Bob, this involves deleting \Edge{\eidConstructionY} and creating \Edge{15}.
 
@@
 \caption{Editing Nested Parts of the Code.
-\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}.
-\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}.
+\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{\vidDifferentAlice} with \Vertex{\vidNestedAlice}.
+\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{\vidSimpleTimes} to the left child of \Vertex{\vidWrapPlus}.
 \autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
@@
 On the other hand, Bob moves the~\texttt{u * v} from the right of the plus to the left of the plus
-by deleting \Edge{11} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
+by deleting \Edge{\eidMoveTimes} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
 Since vertex and edge identifiers are preserved by these,
@@
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of \Edge{11} before making her
+Alice might receive Bob's deletion of \Edge{\eidMoveTimes} before making her
 edits and before receiving Bob's addition of \Edge{19}.
@@
 These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
-In particular pay attention to the edges going out of the \texttt{right} index of \Vertex{8}.
+In particular pay attention to the edges going out of the \texttt{right} index of \Vertex{\vidWrapPlus}.
 In \autoref{fig:Multi-child-a} it is \Edge{21}, which points to the variable~\texttt{x},
@@
 We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
-of \Vertex{8}.
+of \Vertex{\vidWrapPlus}.
 This results in \autoref{fig:Multi-child-c}.
@@
 represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{right}
-index of \Vertex{8}).
+index of \Vertex{\vidWrapPlus}).
 In the graph representation, we flag this conflict by highlighting its edges

### 2020-10-29 grove-paper.tex
Commit before massive figure fixup

@@
 This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
+We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
 \autoref{sub:Single-User Actions} explores edits by a single user, Alice,
@@
 \end{tikzpicture}
-
-TODO: labels for edge numbers
-
-TODO: sub figure with diagram of relation of subfigures
-
-TODO: move subcaptions into master caption
-
-TODO: colors for different editors
-
-TODO: label edges of transitions between graphs
-
-TODO: Remove subfigure captions
 \end{subfigure}
@@
 \end{figure}
+
+TODO: labels for edge numbers
+
+TODO: sub figure with diagram of relation of subfigures
+
+TODO: move subcaptions into master caption
+
+TODO: colors for different editors
+
+TODO: label edges of transitions between graphs
+
+TODO: Remove subfigure captions
 
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.49\columnwidth}
-\centering
-\caption{}
-\label{fig:Construction-a}
-TODO: remove this figure
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\columnwidth}
-\centering
-\caption{}
-\label{fig:Construction-b}
+\begin{minipage}[t]{0.49\columnwidth}
+\centering
 \newedge{ConstructionY}{R}%
@@
 \end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
 \caption{Construction.
-\autoref{fig:Construction-a}:~User interface.
 \autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
 \label{fig:Construction}
+\label{fig:Construction-b}
+\end{minipage}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{minipage}[t]{0.49\linewidth}
+\centering
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   {},
+   "\vConstructionY" [> "\eConstructionY"]
+  }
+ }
+};
+\path (-1cm,-2.5cm) graph[graph style] {
+ "\vSimpleX" % [as={$\texttt{y}_{6}$}]
+};
+\end{tikzpicture}
+\caption{Change.}
+\label{fig:Change-a}
+\label{fig:Change}
+\end{minipage}
+%%%%%%%%
+\hfill{}
 \end{figure}
@@
 \label{sub:Deletion}
-
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
-\label{fig:Change-a}
-TODO: delete and leave deleted z8
-
-TODO: delete "x" instead of "y"
-
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   "\vSimpleX" [> "\eSimpleX"'],
-   {}
-  }
- }
-};
-\path (1cm,-2cm) graph[graph style] {
- y [as={$\texttt{y}_{6}$}]
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
-\label{fig:Change-b}
-TODO: Delete this figure
-\newedge{ChangeZ}{R}
-\newvertex{ChangeZ}{z}
-\begin{tikzpicture}[framed]
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   "\vSimpleX" [> "\eSimpleX"'],
-   "\vChangeZ" [> "\eChangeZ"]
-  }
- }
-};
-\path (1cm,-2cm) graph[graph style] {
- "$\texttt{y}_{6}$"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Change.
-\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{\eidConstructionY}.
-\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{\vidChangeZ} and \Edge{\eidChangeZ}.}
-\label{fig:Change}
-\end{figure}
 

### 2020-10-29 grove-paper.tex
Daily revision (Michael D. Adams)

@@
  as={\textbullet$_{0}$}
+}}
+\tikzset{alice/.style={
+  color=red,
 }}
@@
 \autoref{sub:Single-User Actions} explores edits by a single user, Alice,
-and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
+and \autoref{sub:Multi-user Actions} explores edits by multiple users, Alice and Bob,
 each of whom is editing their own instance of a program.
@@
 
-TODO: labels for edge numbers
-
 TODO: sub figure with diagram of relation of subfigures
 
-TODO: move subcaptions into master caption
-
 TODO: colors for different editors
@@
 TODO: label edges of transitions between graphs
-
-TODO: Remove subfigure captions
 
@@
 %For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
-Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
+In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)},\todo{var} respectively.
 The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
@@
 Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
+The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
 and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
@@
 indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{root}.
-
-Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
-a child position (e.g., \texttt{left} and \texttt{root} in \autoref{fig:A simple tree with a hole-b}).
+The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
+
+Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
+a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:A simple tree with a hole-b}).
 An edge indicates that the destination vertex is a child of the origin vertex at the given position.
@@
 In the remainder of this section,
-we will consider user actions, starting
+we consider user actions, starting
 in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Edits}
+actions and continuing in \autoref{sub:Multi-user Actions}
 with multi-user actions.
 
-Each user action that we consider corresponds to one or more graph edits.
+Each user action corresponds to one or more graph edits.
 Graph edits can only add fresh edges and vertices or remove existing edges.
@@
    "\vSimpleX" [> "\eSimpleX"'],
-   "\vConstructionY" [> "\eConstructionY"]
+   "\vConstructionY" [> "\eConstructionY", >alice, alice]
   }
@@
 };
-\path (-1cm,-2.5cm) graph[graph style] {
+\path (-0.75cm,-2.75cm) graph[graph style] {
  "\vSimpleX" % [as={$\texttt{y}_{6}$}]
@@
 This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
-created vertex, \Vertex{\vidConstructionY}, containing the variable reference \texttt{y}.
+created vertex, \Vertex{\vidConstructionY}, containing the variable reference~\texttt{y}.
 The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.
@@
 \todo{Arrows between figures}
-Next Alice decides to delete~\texttt{x} so that the code becomes~\texttt{\hole  * y}.
+Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
 This is modeled by deleting \Edge{\eidConstructionY} as shown in \autoref{fig:Change-a}. 
-Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children could remain connected to it.
+Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children would remain connected to it.
 In our system, once a vertex is created it is never deleted.
 This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphans from the remaining diagrams if they are not relevant to the exposition.
+Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
 
 Once an edge with a particular identifier is deleted, it cannot be recreated.
-Thus if Alice performed an "undo" on this deletion, Grove would create a fresh edge between \Vertex{TODO} and \Vertex{TODO}.
+Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX}.
 
@@
 };
-\path (-0.5,-3cm) graph[graph style] {
+\path (-0.5cm,-3cm) graph[graph style] {
  "\vSimpleTimes" -> {
-  "\vSimpleX" [> "\eSimpleX"'],
-  "\vChangeZ" [> "\eChangeZ"]
+  {},
+  "\vConstructionY" [> "\eConstructionY"]
  }
@@
  root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
+  "\vWrapPlus" [> "\eWrapPlus", >alice, alice]
  }
@@
  "\vSimpleTimes" -> {
-  "\vSimpleX" [> "\eSimpleX"'],
-  "\vChangeZ" [> "\eChangeZ"]
+  {},
+  "\vConstructionY" [> "\eConstructionY"]
  }
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"'] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vChangeZ" [> "\eChangeZ"]
+   "\vSimpleTimes" [> "\eWrapTimes"', >alice] -> {
+    {},
+    "\vConstructionY" [> "\eConstructionY"]
    },
@@
 \hfill{}
-\caption{Wrapping.
-\autoref{fig:Wrapping-a}:~\autoref{fig:Change-b} after deleting \Edge{\eidSimpleTimes}.
-\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{\vidWrapPlus} and \Edge{\eidWrapPlus}.
-\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{\eidWrapTimes}.}
+\caption{Wrapping.}
 \label{fig:Wrapping}
@@
 
-TODO: revise: "user action" "graph edit"
-
 Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
-If this were a hole, this would cause a construction edit to be performed as in Sec.TODO above.
+If this were a hole, this would cause a construction edit to be performed as in \autoref{sub:Construction} above.
 However, this position is already occupied by the \texttt{*} expression, 
-so instead, typing~\texttt{+} performs a wrapping edit.
-This corresponds to the following sequence of graph actions, shown in \autoref{fig:Wrapping}: 
+so instead, typing~\texttt{+} is interpreted as a wrapping action.\footnote{%
+Multi-user merges can still create multi-child conflicts,
+but it is better to not interpret user actions ina way that creates them.
+How to handle multiple children in the same position is explored in \autoref{sub:Multi-child conflicts}.}
+This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrapping}: 
 (a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex,
+(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex, and
 (c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
 (The choice of wrapping with a bias for the left child position is arbitrary.)
-
-% A naive editor could add the~\texttt{+} as a new child of the root.
-% However,
-
-by creating a new
-edge from the \texttt{L} index of the \texttt{*} to a freshly created \texttt{plus} vertex.
-This would result in a conflict between the \texttt{+} and the existing \texttt{var}
-as shown in Figure~REF:TODO.
-
-You can have a second edge,
-we discuss conflicts in section TODO
-In the single-user setting, we don't want to create conflicts, so instead we perform a wrapping action.
-
-How to handle multiple children in the same position is explored in Sec~REF:TODO,
-but for this sort of edit, a better user experience
-would be interpret the request for a plus to mean to
-take whatever (if anything) is in the cursor
-and place it as a child of the plus.
-This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
-and results in the code~\texttt{(x * z) + \hole}.
-(This is similar to the behavior in other structure editors such as in CITE:TODO.)
-Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.
-
-TODO: "Note that the logic of to insert or wrap is a layer above graph actions."
-
-TODO: Traditional diff: no relation (and indent might change)
 
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vChangeZ" [> "\eChangeZ"]
+    {},
+    "\vConstructionY" [> "\eConstructionY"]
    }
@@
 
-For example, consider \autoref{fig:Move}, which represents
-the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
-to its right child.
-\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
-First, \Edge{\eidWrapPlus} is deleted, which results in \autoref{fig:Move-a}
-where the rooted code is now~\texttt{\hole{} + \hole{}},
-and \Vertex{\vidSimpleTimes} is orphaned.
-Next \Edge{\eidMoveTimes} is added, which results in the bottom graph
-where the rooted code is now~\texttt{\hole{} + 1 * 2},
-and \Vertex{\vidSimpleTimes} is no longer orphaned.
+For example, \autoref{fig:Move} Alice
+moves \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
+This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
+(a) delete \Edge{\eidWrapPlus} leaving \Vertex{\vidSimpleTimes} temporarily orphaned and 
+(b) add an edge (\Edge{\eidMoveTimes}) to \Vertex{\vidSimpleTimes} in the right child position of \Vertex{\vidWrapPlus}.
 
@@
 
-\subsection{Multi-user Edits}
-\label{sub:Multi-user Edits}
+\subsection{Multi-user Actions}
+\label{sub:Multi-user Actions}
 
@@
     "\vDifferentAlice" [> "\eDifferentAlice"'],
-    "\vChangeZ" [> "\eChangeZ"]
+    "\vConstructionY" [> "\eConstructionY"]
    }
@@
 
+TODO: Traditional diff: no relation (and indent might change)
+
 TODO: memory usage

### 2020-11-02 grove-paper.tex
Revistions to Section 2.3 (Michael D. Adams)

@@
 
+%\selectcolormodel{gray}
+
 % math ligatures
@@
   draw,
+  %color=gray,
   font={\small\ttfamily},
@@
   level sep=1cm,
-  nodes={draw, rounded corners=1mm},
+  nodes={
+    draw,
+    %color=gray,
+    rounded corners=1mm},
   edges={edge style},
@@
 \tikzset{alice/.style={
-  color=red,
+  color=red!80!black,
+  font={\bfseries},
+  thick,
+}}
+\tikzset{bob/.style={
+  color=green!60!black,
+  font={\bfseries},
+  thick,
+}}
+\tikzset{merge/.style={
+  color=blue,
+  font={\bfseries},
+  thick,
 }}
@@
 \autoref{sub:Single-User Actions} explores edits by a single user, Alice,
-and \autoref{sub:Multi-user Actions} explores edits by multiple users, Alice and Bob,
+and \autoref{sub:Multi-user Interactions} explores edits by multiple users, Alice and Bob,
 each of whom is editing their own instance of a program.
@@
 \path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   "\vSimpleX" [> "\eSimpleX"'],
+ root[root style, color=black] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
+   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
    {}
@@
 
-TODO: colors for different editors
+TODO: deleted edges (vertecies?) in gray with X in them
 
 TODO: label edges of transitions between graphs
+
+TODO: explain red, green, blue
+
+TODO: $+_32$ and L$_37$
 
@@
 in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Actions}
+actions and continuing in \autoref{sub:Multi-user Interactions}
 with multi-user actions.
@@
 
-\subsubsection{Construction}
+\subsubsection{Construction Actions}
 \label{sub:Construction}
@@
 \end{tikzpicture}
-\caption{Change.}
+\caption{Deletion.}
 \label{fig:Change-a}
@@
 
-\subsubsection{Deletion}
+\subsubsection{Deletion Actions}
 \label{sub:Deletion}
@@
 
-\subsubsection{Wrapping Code}
-\label{sub:Wrapping Code}
+\subsubsection{Wrapping Actions}
+\label{sub:Wrapping}
 
@@
 \hfill{}
-\caption{Wrapping.}
+\caption{Wrapping.  Note that this figure omits the orphaned \Vertex{\vidSimpleX} since it is no longer relevant to our narrative.}
 \label{fig:Wrapping}
@@
 
-\marginpar{START REV HERE}
-
 Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
-If this were a hole, this would cause a construction edit to be performed as in \autoref{sub:Construction} above.
-However, this position is already occupied by the \texttt{*} expression, 
-so instead, typing~\texttt{+} is interpreted as a wrapping action.\footnote{%
-Multi-user merges can still create multi-child conflicts,
-but it is better to not interpret user actions ina way that creates them.
-How to handle multiple children in the same position is explored in \autoref{sub:Multi-child conflicts}.}
+If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
+However, Alice's editor sees that this position is already occupied by the \texttt{*} expression, 
+so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrapping}: 
@@
 
-\subsubsection{Move}
+\subsubsection{Movement Actions}
 \label{sub:Move}
@@
    {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
+   "\vSimpleTimes" [> "\eMoveTimes", >alice] -> {
     {},
@@
 \hfill{}
-\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.
-\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{\eidWrapTimes}.
-\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{\eidMoveTimes}.}
+\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.}
 \label{fig:Move}
@@
 
-An interesting consequence of our design is that a move operation
+\marginpar{START REV HERE}
+
+A consequence of our design is that a move operation
 does not have to delete code from one position and create a copy
@@
 
-\subsection{Multi-user Actions}
-\label{sub:Multi-user Actions}
+\subsection{Multi-user Interactions}
+\label{sub:Multi-user Interactions}
 
@@
 We assume multiple instances
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
 
@@
 
-TODO: \autoref{sub:Editing Different Parts of the Code}
+\subsubsection{Commutativity}
+\label{sub:Commutativity}
+
+Note that aside from transmitting the edge deletions and additions,
+the merge requires no coordination between Alice and Bob.
+For example, suppose that Alice and Bob did not transmit their edits to each other
+but to a third user, Chris, that has the original graph in \autoref{fig:Move-b}.
+Chris can apply the edits from Alice and Bob in any order and will always get
+the graph in \autoref{fig:Editing Different Parts of the Code-c}.
+
+Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
+For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
+applying them in either order produces the same results.
 
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"'],
+    "\vDifferentAlice" [> "\eDifferentAlice"', >alice, alice],
     "\vConstructionY" [> "\eConstructionY"]
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
+    {},
+    "\vDifferentBob" [> "\eDifferentBob", >bob, bob]
    }
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
+    "\vDifferentAlice" [> "\eDifferentAlice"', >merge, merge],
+    "\vDifferentBob" [> "\eDifferentBob", >merge, merge]
    }
@@
 \hfill{}
-\caption{Editing Different Parts of the Code.
-\textbf{(a)}:~\autoref{fig:Move-b} after Alice replaces \Vertex{\vidSimpleX} with \Vertex{\vidDifferentAlice}.
-\textbf{(b)}:~\autoref{fig:Move-b} after Bob replaces \Vertex{\vidConstructionY} with \Vertex{\vidDifferentBob}.
-\textbf{(c)}:~The
-result of merging \autoref{fig:Editing Different Parts of the Code-a}
-with \autoref{fig:Editing Different Parts of the Code-b}.}
+\caption{Editing Different Parts of the Code.}
 \label{fig:Editing Different Parts of the Code}
@@
 
-TODO: have star or dot dot arrows for skipping steps
-
-Arrow coming from 5b to 6a and 6b
-
-Diamond diagram referencing diagrams
-
 To start, Alice and Bob are synchronized with each other and
 both have the graph in \autoref{fig:Move-b}, which
-represents the code~\texttt{\hole{} + (x + z)}.
-
-Alice changes the~\texttt{x} to a~\texttt{u},
-and Bob changes the~\texttt{z} to a~\texttt{v}.
+represents the code~\texttt{\hole{} + (\hole{} + y)}.
+
+Alice add~\texttt{u} as the left child of \Vertex{TODO}
+and Bob changes the~\texttt{y} to a~\texttt{v}.
 Before transmitting their edits to each other,
@@
 and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
-For Alice, this involves deleting \Edge{3} and creating \Edge{13}.
-For Bob, this involves deleting \Edge{\eidConstructionY} and creating \Edge{15}.
-
-Alice and Bob then transmit their graph actions (i.e., edge additions and deletions) to each other
-and applies the other's graph actions to their own copy of the graph.
-By design, graph actions commute, so this results in both Alice and Bob having the graph in
+
+Alice and Bob then transmit their graph edits to each other
+and apply the other's graph edits to their own copy of the graph.
+This results in both Alice and Bob having the graph in
 \autoref{fig:Editing Different Parts of the Code-c}.
-
-Note that aside from transmitting the edge deletions and additions,
-no coordination between Alice and Bob is needed to coordinate the merge.
-For example, suppose that Alice and Bob did not transmit their edits to each other
-but to a third user, Chris, that has the original graph at the top of \autoref{fig:Move-b}.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:Editing Different Parts of the Code-c}.
-
-Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
-applying them in either order produces the same results.
 
@@
 
+TODO: have star or dot dot arrows for skipping steps
+
+TODO: Arrow coming from 5b to 6a and 6b
+
+TODO: Diamond diagram referencing diagrams
+
 \subsubsection{Editing Nested Parts of the Code}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vNestedAlice" [> "\eNestedAlice"', >alice, alice],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedBob"', >bob] -> {
     "\vDifferentAlice" [> "\eDifferentAlice"'],
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
+   "\vSimpleTimes" [> "\eNestedBob"', >merge] -> {
+    "\vNestedAlice" [> "\eNestedAlice"', >merge, merge],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
 Alice changes the~\texttt{v} to a~\texttt{w}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
-On the other hand, Bob moves the~\texttt{u * v} from the right of the plus to the left of the plus
+On the other hand, Bob moves the~\texttt{u * v} from the left of the plus to the right of the plus
 by deleting \Edge{\eidMoveTimes} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
@@
 which results in both Alice and Bob having the graph in \autoref{fig:Editing Nested Parts of the Code-c}.
-
-As with \autoref{sub:Editing Different Parts of the Code}, this merge does not require coordination
-between Alice and Bob.  They could independently send their edits
-to a third party, Charley, who could replay their edits in any order
-and will always get the graph in \autoref{fig:Editing Nested Parts of the Code-c}.
 
@@
 Her edits would then be within the code deleted by Bob.
+TODO: She can restore that code using a restoration action add a fresh edge to it.
 
@@
 
-TODO: put this somewhere.
-(Similar to how git merged across moves.
+Similar to how git merged across moves.
 A notable difference though is that \texttt{git} detects
@@
    {
-    "\vMultiChildAlice" [> "\eMultiChildAlice"]
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice, alice]
    }
@@
    },
-   "\vMultiChildBob" [> "\eMultiChildBob"]
+   "\vMultiChildBob" [> "\eMultiChildBob", >bob, bob]
   }
@@
    {
-    "\vMultiChildAlice" [> "\eMultiChildAlice"],
-    "\vMultiChildBob" [> "\eMultiChildBob"]
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >merge, merge],
+    "\vMultiChildBob" [> "\eMultiChildBob", >merge, merge]
    }
@@
 
+TODO: Multi-user merges can still create multi-child conflicts,
+but it is better to not interpret user actions in a way that creates them.
+How to handle multiple children in the same position is explored in \autoref{sub:Multi-child conflicts}.
+
 Finally, we address our first kind of genuine conflict between edits.
 Alice inserts the variable~\texttt{x} to the right side of the~\texttt{+}
-while Bob add the variable~\texttt{y} also to the right side of the~\texttt{+}
+while Bob add the variable~\texttt{y} also to the right side of the~\texttt{+}.
 These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
@@
 The conflict between Alice's and Bob's changes is
-represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{right}
+represented by the fact that there are multiple edges going out of one index (i.e., the \texttt{right}
 index of \Vertex{\vidWrapPlus}).
@@
 
+TODO: need example of fully-automatic resolution and semi-automatic resolution
+
+Multiple adds cause conflicts
+(Not usually presented in the UI)
+
 Compare this to what happens when there are conflicts in a traditional,
@@
 when a merge conflict happens.
-These show alternate versions versions of the code, and it is up to the user
+These show alternate versions of the code, and it is up to the user
 to replace those alternates with the desired merged result.
@@
 of the code, and it is up to the user to replace those alternates with the desired merged result.
-Our model is different, in that merge conflicts in the line-based model
+Our model is different in that merge conflicts in the line-based model
 prevent some semantic editor services from running
@@
 
-TODO: need example of fully-automatic resolution and semi-automatic resolution
-
-Multiple adds cause conflicts
-(Not usually presented in the UI)
-
 \subsubsection{Multi-parent conflicts}
@@
     {},
-    "\vNestedAlice" [> "\eMultiParentAlice"']
+    "\vNestedAlice" [> "\eMultiParentAlice"', >alice]
    },
@@
    "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vNestedAlice" [> "\eMultiParentBob"]
+   "\vNestedAlice" [> "\eMultiParentBob", >bob]
   }
@@
     {},
-    b/"\vNestedAlice" [> "\eMultiParentAlice"']
+    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge]
    },
@@
 };
-\path [edge style,->] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
+\path [edge style,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge] (b);
 \end{tikzpicture}
@@
 
-In addition to conflicts due multiple edges coming \emph{out of} an index of a particular vertex,
+In addition to conflicts due to multiple edges coming \emph{out of} child position of a particular vertex,
 conflicts can manifest as multiple edges going \emph{into} a particular vertex.
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(x + \hole{}) + \hole{})}.
-Alice moves the variable~\texttt{x} to the left-most hole,
+This graph represents the expression~\texttt{(w + \hole{}) + \hole{})}.
+Alice moves the variable~\texttt{w} to the left-most hole,
 while Bob moves it to the right-most hole.
 This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
-which represent respectively the expression~\texttt{(\hole{} + \hole{}) + x} for Alice
-and the expression~\texttt{(\hole{} + x) + \hole{}} for Bob.
+which represent respectively the expression~\texttt{(\hole{} + w) + \hole{}} for Alice
+and the expression~\texttt{(\hole{} + \hole{}) + w} for Bob.
 In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes \Edge{25} and adds \Edge{27}, while Bob deletes \Edge{25} and adds edge \Edge{29}.
-As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \Edge{25} is not a problem.
-However, both \Edge{27} (added by Alice) and edge \Edge{29} (added by Bob) point to the same vertex.
-This generates the bottom graph, in which \Vertex{26} as two parents pointing to it.
+Alice deletes \Edge{TODO} and adds \Edge{TODO}, while Bob deletes \Edge{TODO} and adds edge \Edge{TODO}.
+As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \Edge{TODO} is not a problem.
+However, both \Edge{TODO} (added by Alice) and edge \Edge{TODO} (added by Bob) point to the same vertex.
+This generates REF:FIG, in which \Vertex{TODO} as two parents pointing to it.
 We call this situation a multi-parent conflict.
+
+TODO: talk about how to resolve the conflict
 
@@
     {},
-    "\vSingleCyclePlus" [> "\eSingleCyclePlus"] -> {
-     "\vSingleCycleTimes" [> "\eSingleCycleTimes"'] -> {
+    "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice, alice] -> {
+     "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice, alice] -> {
       {},
@@
 };
-\path[edge style,->] (v2) to ["\eSingleCycleCycle",out=-60,in=-60] (v8);
+\path[edge style,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice] (v8);
 \end{tikzpicture}
@@
 For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-done by a cutting from the old position and pasting to the new position.
+by a cutting from the old position and pasting to the new position.
 Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
@@
 is ultimately a user interface question.
-Cycles caused by the local user's edits be detected as soon as a user enters them
+Cycles caused by the local user's edits can be detected as soon as a user enters them
 by noting either that the graph would contain a cycle or the vertex
@@
 Thus, as a user interface consideration, it might be best to either
-disallow such edits to at least warn users that their
+disallow such edits to at least warn users when their
 edits would create a cycle.
@@
    {
-    "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"'],
+    "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus", >alice, alice] -> {
+     "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"', >alice, alice],
      {}
@@
    {
-    "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"'],
+    "\vMultiCycleTimes" [> "\eMultiCycleBobTimes", >bob, bob] -> {
+     "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"', >bob, bob],
      {}
@@
    {
-    v1/"\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"],
-    v2/"\vMultiCycleTimes" [> "\eMultiCycleBobTimes"]
-   }
-  }
- }
-};
-\path[edge style,->] (v1) to ["\eMultiCycleAliceTimes",out=-45,in=-135] (v2);
-\path[edge style,->] (v2) to ["\eMultiCycleBobPlus",out=-90,in=-90] (v1);
+    v1/"\vMultiCyclePlus" [> "\eMultiCycleAlicePlus", >merge, merge],
+    v2/"\vMultiCycleTimes" [> "\eMultiCycleBobTimes", >merge, merge]
+   }
+  }
+ }
+};
+\path[edge style,->,merge] (v1) to ["\eMultiCycleAliceTimes",out=-45,in=-135,merge] (v2);
+\path[edge style,->,merge] (v2) to ["\eMultiCycleBobPlus",out=-90,in=-90,merge] (v1);
 \end{tikzpicture}
@@
 \hfill{}
-\caption{Multi-User Cycles
-\autoref{fig:Multi-User Cycles-a}:~Initial code.
-\autoref{fig:Multi-User Cycles-b}:~TODO.
-\autoref{fig:Multi-User Cycles-c}:~TODO.
-\autoref{fig:Multi-User Cycles-d}:~TODO.}
+\caption{Multi-User Cycles.}
 \label{fig:Multi-User Cycles}
@@
     {},
-    "\vMultiCycleTimes" [> "\eDisconnectionAlice"]
+    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice]
    },
@@
    "\vMultiCycleTimes" [> "\eMultiCycleAlicePlus"] -> {
-    "\vSimpleTimes" [> "\eDisconnectionBob"'],
+    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob],
     {}
@@
 \path ( 1cm,-3cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[edge style,->] (times) to ["\eMultiCycleBobPlus"anchor=south,out=-135,in=-45] (plus);
-\path[edge style,->] (plus) to ["\eMultiCycleBobTimes"anchor=north,out=45,in=135] (times);
+\path[edge style,->,merge] (times) to ["\eMultiCycleBobPlus"anchor=south,out=-135,in=-45, merge] (plus);
+\path[edge style,->,merge] (plus) to ["\eMultiCycleBobTimes"anchor=north,out=45,in=135, merge] (times);
 \end{tikzpicture}
@@
 \hfill{}
-\caption{Disconnection.
-\autoref{fig:Disconnection-a}:~Initial code.
-\autoref{fig:Disconnection-b}:~TODO.
-\autoref{fig:Disconnection-c}:~TODO.
-\autoref{fig:Disconnection-d}:~TODO.}
+\caption{Disconnection.}
 \label{fig:Disconnection}
@@
 
+TODO: cache eviction algorithm
+
+TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
+In practice, these complex scenarios occur less frequently than portrayed here.
+
 \subsection{Variable names, strings, and numbers}

### 2020-11-04 grove-paper.tex
Add some inter-figure arrows (Michael D. Adams)

@@
 \tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
+  every picture/.style={framed, background rectangle/.style={draw=gray!50}},
 }
@@
 
-TODO: deleted edges (vertecies?) in gray with X in them
+TODO: [don't do this] deleted edges (vertecies?) in gray with X in them
 
@@
 TODO: $+_32$ and L$_37$
+
+TODO: extra space in figure 12 d
+
+TODO: reduce vertical tree sep
+
+TODO: list actions in an overview subfigure
+
+TODO: maybe label actions between
 
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Move-b) {};
+\end{tikzpicture}
+}
 \label{fig:Move-b}
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-a) {};
+\path [draw,->,alice] (fig-Move-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-a}
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-b) {};
+\path [draw, bob,->] (fig-Move-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-b}
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-c) {};
+\path [draw, merge,->] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path [draw, merge,->] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-c}
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+%\hfill
+%%%%%%%%
+\hskip3em%\hfill
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {\textbf{Fig.~\ref*{fig:Editing Different Parts of the Code-c}}};
+%%%%
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
+\path [draw,merge,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-1.5em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-a}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
+\path [draw,merge,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-1.5em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-b}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
+\path [draw,merge,->,merge] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-1.5em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path [draw,merge,->,merge] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-1.5em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-c}
@@
 
+Homotopical Patch Theorey: https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7
+
 \section{Discussion and Conclusion}

### 2020-11-05 grove-paper.tex
Finish inter-figure arrows (Michael D. Adams)

--- 2020-11-02 grove-paper.tex
@@
   %color=gray,
-  font={\small\ttfamily},
+  font={\footnotesize\ttfamily},
   /tikz/every edge quotes/.style={
@@
   tree layout,
-  level sep=1cm,
+  level distance=0.5cm,
+  level sep=0.5cm,
+  sibling distance=0.5cm,
+  sibling sep=0.1cm,
+  part distance=0.1cm,
+  part sep=0.1cm,
+  component distance=0.1cm,
+  component sep=0.1cm,
   nodes={
@@
     %color=gray,
+    inner sep=2pt,
     rounded corners=1mm},
@@
   color=red!80!black,
-  font={\bfseries},
+  font={\bfseries\footnotesize},
   thick,
@@
   color=green!60!black,
-  font={\bfseries},
+  font={\bfseries\footnotesize},
   thick,
@@
   color=blue,
-  font={\bfseries},
+  font={\bfseries\footnotesize},
   thick,
 }}
+\tikzset{alice edge/.style={alice, edge style, font={\bfseries\footnotesize}}}
+\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\footnotesize}}
+\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\footnotesize}}
 
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{minipage}[t]{0.49\columnwidth}
-\centering
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-A simple tree with a hole-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:A simple tree with a hole-b}}}};
+%%%%
+\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-a}
+\label{fig:Construction}
+\label{fig:Construction-b}
 \newedge{ConstructionY}{R}%
@@
    "\vSimpleX" [> "\eSimpleX"'],
-   "\vConstructionY" [> "\eConstructionY", >alice, alice]
-  }
- }
-};
-\end{tikzpicture}
-\caption{Construction.
-\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
-\label{fig:Construction}
-\label{fig:Construction-b}
-\end{minipage}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{minipage}[t]{0.49\linewidth}
-\centering
+   "\vConstructionY" [> "\eConstructionY", >alice edge, alice]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Construction) to[alice edge] (-2em, 0cm |- fig-Deletion);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-b}
+\label{fig:Change-a}
+\label{fig:Change}
 \begin{tikzpicture}
@@
 \path (-0.75cm,-2.75cm) graph[graph style] {
- "\vSimpleX" % [as={$\texttt{y}_{6}$}]
-};
-\end{tikzpicture}
-\caption{Deletion.}
-\label{fig:Change-a}
-\label{fig:Change}
-\end{minipage}
+ "\vSimpleX"
+};
+\end{tikzpicture}
+\end{subfigure}
 %%%%%%%%
 \hfill{}
+\caption{Construction and Deletion.}
 \end{figure}
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Deletion) to[alice edge] (-2em, 0cm |- fig-Wrapping-a);
+\end{tikzpicture}
+}
 \label{fig:Wrapping-a}
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-a) to[alice edge] (-2em, 0cm |- fig-Wrapping-b);
+\end{tikzpicture}
+}
 \label{fig:Wrapping-b}
@@
  root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice, alice]
+  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice]
  }
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-c) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-b) to[alice edge] (-2em, 0cm |- fig-Wrapping-c);
+\end{tikzpicture}
+}
 \label{fig:Wrapping-c}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice] -> {
+   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
     {},
@@
 \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
-% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
-% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
-% \end{tikzpicture}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
+\end{tikzpicture}
+}
 \label{fig:Move-a}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.45\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Moving-b) {};
+\path [draw,->,alice] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
+\end{tikzpicture}
+}
 \label{fig:Move-b}
@@
    {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice] -> {
+   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
     {},
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-a) {};
+\path [draw,->,alice] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-a}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >alice, alice],
+    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice],
     "\vConstructionY" [> "\eConstructionY"]
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-b) {};
+\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-b}
@@
     {},
-    "\vDifferentBob" [> "\eDifferentBob", >bob, bob]
+    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob]
    }
@@
 \centering
-\caption{}
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,1ex) node (fig-Editing Different Parts of the Code-c) {};
+\path [draw,->,merge] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path [draw,->,merge] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\end{tikzpicture}
+}
 \label{fig:Editing Different Parts of the Code-c}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >merge, merge],
-    "\vDifferentBob" [> "\eDifferentBob", >merge, merge]
+    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge],
+    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge]
    }
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-a}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >alice, alice],
+    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-b}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >bob] -> {
+   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
     "\vDifferentAlice" [> "\eDifferentAlice"'],
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path [draw,->,merge] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\end{tikzpicture}
+}
 \label{fig:Editing Nested Parts of the Code-c}
+\label{fig:Editing Nested-c}
 \begin{tikzpicture}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >merge] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >merge, merge],
+   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
+    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
 \begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\linewidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
+\end{tikzpicture}
+}
 \label{fig:Multi-child-a}
@@
    {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice, alice]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\linewidth}
-\centering
-\caption{}
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
+\end{tikzpicture}
+}
 \label{fig:Multi-child-b}
@@
    },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob, bob]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\linewidth}
-\centering
-\caption{}
+   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
+\path [draw,->,merge] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
+\end{tikzpicture}
+}
 \label{fig:Multi-child-c}
@@
    {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >merge, merge],
-    "\vMultiChildBob" [> "\eMultiChildBob", >merge, merge]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\linewidth}
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >merge edge, merge],
+    "\vMultiChildBob" [> "\eMultiChildBob", >merge edge, merge]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
 \centering
@@
 \begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.19\linewidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-child-c-x) to (-2em, 0cm |- fig-Multi-parent-a);
+\end{tikzpicture}
+}
 \label{fig:Multi-parent-a}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.19\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-a) to (-2em, 0cm |- fig-Multi-parent-b);
+\end{tikzpicture}
+}
 \label{fig:Multi-parent-b}
@@
     {},
-    "\vNestedAlice" [> "\eMultiParentAlice"', >alice]
+    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
    },
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.19\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-c);
+\end{tikzpicture}
+}
 \label{fig:Multi-parent-c}
@@
    "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vNestedAlice" [> "\eMultiParentBob", >bob]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.19\linewidth}
-\centering
-\caption{}
+   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\path [draw,->,merge] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\end{tikzpicture}
+}
 \label{fig:Multi-parent-d}
@@
     {},
-    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge]
+    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
    },
@@
 };
-\path [edge style,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.19\linewidth}
-\centering
-\caption{}
+\path [draw,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge edge] (b);
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
+\end{tikzpicture}
+}
 \label{fig:Multi-parent-e}
@@
 \hfill{}
-\caption{Multi-parent conflicts.
-\autoref{fig:Multi-parent-a}:~Initial code.
-\autoref{fig:Multi-parent-b}:~\autoref{fig:Multi-parent-a} after Alice ... TODO.
-\autoref{fig:Multi-parent-c}:~\autoref{fig:Multi-child-a} after Bob ... TODO.
-\autoref{fig:Multi-parent-d}:~The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}.
-\autoref{fig:Multi-parent-e}:~Initial code TODO: move to previous figure.}
+\caption{Multi-parent conflicts.}
 \label{fig:Multi-parent}
@@
 
-% \subsubsection{Conflicting wrap operations}
-
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% % TODO: texttt
-% % TODO: position the different graphs
-% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
-% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
-% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
-% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
-% % \path (0cm,0cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     x
-% %   }
-% % };
-% % \path (-3cm,-2cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     + -> {
-% %       1,
-% %       x
-% %     }
-% %   }
-% % };
-% % \path (3cm,-2cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     * -> {
-% %       x,
-% %       2
-% %     }
-% %   }
-% % };
-% % \path (0cm,-4cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     + -> {
-% %       1,
-% %       x
-% %     },
-% %     * -> {
-% %       x,
-% %       2
-% %     }
-% %   }
-% % };
-% % \path (2cm,-8cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     + -> {
-% %       1,
-% %       * -> {
-% %         x,
-% %         2
-% %       }
-% %     }
-% %   }
-% % };
-% % \path (-2cm,-8cm) graph[tree layout] {
-% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
-% %     * -> {
-% %       + -> {
-% %         1,
-% %         x
-% %       },
-% %       2
-% %     }
-% %   }
-% % };
-% \end{tikzpicture}
-% \caption{Conflicting wraps}
-% \label{fig:Conflicting wraps}
-% \end{figure}
-
-% If two different 
-% As seen in REF:SEC:TODO, conflicting wraps gener
-
-% Multi-parent
-
-% TODO: SHOW intermediate Steps (maybe no graph)
-%   Show user's perspective
-%   Accept one side = delete others (select this)
-%   Delete x in *
-%   Move + to x
-
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (0cm,0cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     + -> {
-%       1,
-%       x
-%     },
-%     * -> {
-%       x,
-%       2
-%     }
-%   }
-% };
-% \path (2cm,-4cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     + -> {
-%       1,
-%       * -> {
-%         x,
-%         2
-%       }
-%     }
-%   }
-% };
-% \path (-2cm,-4cm) graph[tree layout] {
-%   lam[as={$\lambda_{\texttt{x}}$}] -> {
-%     * -> {
-%       + -> {
-%         1,
-%         x
-%       },
-%       2
-%     }
-%   }
-% };
-% \end{tikzpicture}
-% \caption{Wrap Resolution}
-% \label{fig:Wrap Resolution}
-% \end{figure}
-
-% NEEDS UI MOCKUP
-
-% Say: "might not be a popup"
-
 \subsubsection{Single-User Cycles}
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.28\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
+%%%%
+\path (0cm,0cm) node (fig-Single-User Cycles-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
+\end{tikzpicture}
+}
 \label{fig:Single-User Cycles-a}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.28\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Single-User Cycles-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
+\end{tikzpicture}
+}
 \label{fig:Single-User Cycles-b}
@@
     {},
-    "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice, alice] -> {
-     "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice, alice] -> {
+    "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice] -> {
+     "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice] -> {
       {},
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.28\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Single-User Cycles-c) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Single-User Cycles-b) to (-2em, 0cm |- fig-Single-User Cycles-c);
+\end{tikzpicture}
+}
 \label{fig:Single-User Cycles-c}
@@
 };
-\path[edge style,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice] (v8);
+\path[draw,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice edge] (v8);
 \end{tikzpicture}
@@
 \label{sub:Multi-User Cycles}
-
+MultiCyclePlus
 \begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Single-User Cycles-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Single-User Cycles-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Single-User Cycles-c-x) to (-2em, 0cm |- fig-Multi-User Cycles-a);
+\end{tikzpicture}
+}
 \label{fig:Multi-User Cycles-a}
-\newedge{MultiCyclePlus}{L}
+\newedge{MultiCycleTimes}{L}
+\newvertex{MultiCycleTimes}{*}
+\newedge{MultiCyclePlus}{R}
 \newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleTimes}{R}
-\newvertex{MultiCycleTimes}{*}
 \begin{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus"'],
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"]
+    "\vMultiCycleTimes" [> "\eMultiCycleTimes"'],
+    "\vMultiCyclePlus" [> "\eMultiCyclePlus"]
    },
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-User Cycles-a) to (-2em, 0cm |- fig-Multi-User Cycles-b);
+\end{tikzpicture}
+}
 \label{fig:Multi-User Cycles-b}
-\newedge{MultiCycleAlicePlus}{R}
-\newedge{MultiCycleAliceTimes}{L}
+\newedge{MultiCycleAliceTimes}{R}
+\newedge{MultiCycleAlicePlus}{L}
 \begin{tikzpicture}
@@
    {
-    "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus", >alice, alice] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"', >alice, alice],
+    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice] -> {
+     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice],
      {}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-c);
+\end{tikzpicture}
+}
 \label{fig:Multi-User Cycles-c}
-\newedge{MultiCycleBobTimes}{R}
-\newedge{MultiCycleBobPlus}{L}
+\newedge{MultiCycleBobPlus}{R}
+\newedge{MultiCycleBobTimes}{L}
 \begin{tikzpicture}
@@
    {
-    "\vMultiCycleTimes" [> "\eMultiCycleBobTimes", >bob, bob] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"', >bob, bob],
+    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob] -> {
+     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob],
      {}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
+\path [draw,->,merge] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
+\end{tikzpicture}
+}
 \label{fig:Multi-User Cycles-d}
@@
    {
-    v1/"\vMultiCyclePlus" [> "\eMultiCycleAlicePlus", >merge, merge],
-    v2/"\vMultiCycleTimes" [> "\eMultiCycleBobTimes", >merge, merge]
-   }
-  }
- }
-};
-\path[edge style,->,merge] (v1) to ["\eMultiCycleAliceTimes",out=-45,in=-135,merge] (v2);
-\path[edge style,->,merge] (v2) to ["\eMultiCycleBobPlus",out=-90,in=-90,merge] (v1);
+    v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >merge edge, merge],
+    v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >merge edge, merge]
+   }
+  }
+ }
+};
+\path[draw,->,merge] (v1) to ["\eMultiCycleAlicePlus",out=-45,in=-135,merge edge] (v2);
+\path[draw,->,merge] (v2) to ["\eMultiCycleBobTimes",out=-90,in=-90,merge edge] (v1);
 \end{tikzpicture}
@@
 \begin{figure*}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
+%%%%
+\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to (-2em, 0cm |- fig-Disconnection-a);
+\end{tikzpicture}
+}
 \label{fig:Disconnection-a}
@@
    "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+   "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Disconnection-a) to (-2em, 0cm |- fig-Disconnection-b);
+\end{tikzpicture}
+}
 \label{fig:Disconnection-b}
@@
     {},
-    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice]
+    "\vMultiCyclePlus" [> "\eDisconnectionAlice", >alice edge]
    },
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-c);
+\end{tikzpicture}
+}
 \label{fig:Disconnection-c}
@@
    {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAlicePlus"] -> {
-    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob],
+   "\vMultiCyclePlus" [> "\eMultiCycleBobTimes"] -> {
+    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
     {}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.24\textwidth}
-\centering
-\caption{}
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
+\path [draw,->,merge] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
+\end{tikzpicture}
+}
 \label{fig:Disconnection-d}
@@
 \path (-1cm,-3cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 1cm,-3cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[edge style,->,merge] (times) to ["\eMultiCycleBobPlus"anchor=south,out=-135,in=-45, merge] (plus);
-\path[edge style,->,merge] (plus) to ["\eMultiCycleBobTimes"anchor=north,out=45,in=135, merge] (times);
+\path ( 1cm,-3cm) graph[graph style] { times/"\vMultiCyclePlus" };
+\path[draw,->,merge] (times) to ["\eMultiCycleBobTimes"anchor=south,out=-135,in=-45, merge edge] (plus);
+\path[draw,->,merge] (plus) to ["\eMultiCycleBobPlus"anchor=north,out=45,in=135, merge edge] (times);
 \end{tikzpicture}

### 2020-11-06 grove-paper.tex
Move figures into commands (Michael D. Adams)
Fix spacing and various refinements to figures

@@
 
-%\selectcolormodel{gray}
+% \selectcolormodel{gray}
 
@@
   backgrounds,  % Provides "framed" and "gridded"
+  bending,      % bending arrow tips
   graphs,       % Graph *notation*
@@
   %color=gray,
-  font={\footnotesize\ttfamily},
+  font={\small\ttfamily},
   /tikz/every edge quotes/.style={
@@
   color=red!80!black,
-  font={\bfseries\footnotesize},
+  font={\bfseries\small},
   thick,
@@
   color=green!60!black,
-  font={\bfseries\footnotesize},
+  font={\bfseries\small},
   thick,
@@
   color=blue,
-  font={\bfseries\footnotesize},
+  font={\bfseries\small},
   thick,
 }}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\footnotesize}}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\footnotesize}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\footnotesize}}
+\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
+\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
+\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureSimple
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{SimpleTimes}{Root}%
+\newvertex{SimpleTimes}{*}%
+\newedge{SimpleX}{L}%
+\newvertex{SimpleX}{x}%
+\newcommand{\figureSimple}{
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:A simple tree with a hole-a}
+TODO
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.49\linewidth}
+\centering
+\caption{}
+\label{fig:A simple tree with a hole-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style, color=black] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
+   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{A simple tree with a hole.
+\autoref{fig:A simple tree with a hole-a}:~User interface.
+\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
+\label{fig:A simple tree with a hole}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureConstructionAndDeletion
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{ConstructionY}{R}%
+\newvertex{ConstructionY}{y}%
+\newcommand{\figureConstructionAndDeletion}{
+\begin{figure}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-A simple tree with a hole-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:A simple tree with a hole-b}}}};
+%%%%
+\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-a}
+\label{fig:Construction}
+\label{fig:Construction-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   "\vSimpleX" [> "\eSimpleX"'],
+   "\vConstructionY" [> "\eConstructionY", >alice edge, alice]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Construction) to[alice edge] (-2em, 0cm |- fig-Deletion);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-b}
+\label{fig:Change-a}
+\label{fig:Change}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+   {},
+   "\vConstructionY" [> "\eConstructionY"]
+  }
+ }
+};
+\path (-0.25cm,-1.7cm) graph[graph style] {
+ "\vSimpleX"
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Construction and Deletion.}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureWrapping
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{WrapPlus}{Root}
+\newvertex{WrapPlus}{+}
+\newedge{WrapTimes}{L}
+\newcommand{\figureWrapping}{
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Deletion) to[alice edge] (-2em, 0cm |- fig-Wrapping-a);
+\end{tikzpicture}
+}
+\label{fig:Wrapping-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style]
+};
+\path (0cm,-1.7cm) graph[graph style] {
+ "\vSimpleTimes" -> {
+  {},
+  "\vConstructionY" [> "\eConstructionY"]
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-a) to[alice edge] (-2em, 0cm |- fig-Wrapping-b);
+\end{tikzpicture}
+}
+\label{fig:Wrapping-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice]
+ }
+};
+\path (0cm,-1.7cm) graph[graph style] {
+ "\vSimpleTimes" -> {
+  {},
+  "\vConstructionY" [> "\eConstructionY"]
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Wrapping-c) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-b) to[alice edge] (-2em, 0cm |- fig-Wrapping-c);
+\end{tikzpicture}
+}
+\label{fig:Wrapping-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
+    {},
+    "\vConstructionY" [> "\eConstructionY"]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Wrapping.  Note that this figure omits the orphaned \Vertex{\vidSimpleX} since it is no longer relevant to our narrative.}
+\label{fig:Wrapping}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureMovement
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{MoveTimes}{R}
+\newcommand{\figureMovement}{
+\begin{figure}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
+\end{tikzpicture}
+}
+\label{fig:Move-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"]
+ }
+};
+\path (0cm,-1.7cm) graph [graph style] {
+ "\vSimpleTimes" -> {
+  {},
+  "\vConstructionY" [> "\eConstructionY"]
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.43\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Moving-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
+\end{tikzpicture}
+}
+\label{fig:Move-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
+    {},
+    "\vConstructionY" [> "\eConstructionY"]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Moving \vSimpleTimes{} to the right child of~\vWrapPlus{}.}
+\label{fig:Move}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureEditingDifferentParts
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{DifferentAlice}{L}
+\newvertex{DifferentAlice}{u}
+\newedge{DifferentBob}{R}
+\newvertex{DifferentBob}{v}
+\newcommand{\figureEditingDifferentParts}{
+\begin{figure}
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Different Parts of the Code-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\end{tikzpicture}
+}
+\label{fig:Editing Different Parts of the Code-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
+    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice],
+    "\vConstructionY" [> "\eConstructionY"]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Different Parts of the Code-b) [anchor=base] {\strut};
+\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\end{tikzpicture}
+}
+\label{fig:Editing Different Parts of the Code-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
+    {},
+    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Different Parts of the Code-c) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path [draw,->,merge] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\end{tikzpicture}
+}
+\label{fig:Editing Different Parts of the Code-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
+    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge],
+    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Editing Different Parts of the Code.}
+\label{fig:Editing Different Parts of the Code}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureCommutativity
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\figureCommutativity}{
+\begin{figure}
+\centering
+
+TODO: commutativity diagram of screenshots with two cursors
+
+TODO: include legend of arrow colors
+
+\caption{TODO:Commutativity}
+\label{fig:Commutativity}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureEditingNestedParts
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{NestedAlice}{L}
+\newvertex{NestedAlice}{w}
+\newedge{NestedBob}{L}
+\newcommand{\figureEditingNestedParts}{
+\begin{figure}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\end{tikzpicture}
+}
+\label{fig:Editing Nested Parts of the Code-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vSimpleTimes" [> "\eMoveTimes"] -> {
+    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\end{tikzpicture}
+}
+\label{fig:Editing Nested Parts of the Code-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
+    "\vDifferentAlice" [> "\eDifferentAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.25\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path [draw,->,merge] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\end{tikzpicture}
+}
+\label{fig:Editing Nested Parts of the Code-c}
+\label{fig:Editing Nested-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
+    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Editing Nested Parts of the Code.
+\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{\vidDifferentAlice} with \Vertex{\vidNestedAlice}.
+\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{\vidSimpleTimes} to the left child of \Vertex{\vidWrapPlus}.
+\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
+\label{fig:Editing Nested Parts of the Code}
+\end{figure}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureMultiChild
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{MultiChildAlice}{R}
+\newvertex{MultiChildAlice}{x}
+\newedge{MultiChildBob}{R}
+\newvertex{MultiChildBob}{y}
+\newcommand{\figureMultiChild}{
+\begin{figure*}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
+\end{tikzpicture}
+}
+\label{fig:Multi-child-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   },
+   {
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice]
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
+\end{tikzpicture}
+}
+\label{fig:Multi-child-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   },
+   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
+\path [draw,->,merge] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
+\end{tikzpicture}
+}
+\label{fig:Multi-child-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    "\vDifferentBob" [> "\eDifferentBob"]
+   },
+   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge],
+   {},
+   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-child-d}
+TODO: show UI
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-child conflicts.}
+\label{fig:Multi-child}
+\end{figure*}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureMultiParent
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{MultiParentAlice}{R}
+\newedge{MultiParentBob}{R}
+\newcommand{\figureMultiParent}{
+\begin{figure*}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-child-c-x) to (-2em, 0cm |- fig-Multi-parent-a);
+\end{tikzpicture}
+}
+\label{fig:Multi-parent-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vNestedAlice" [> "\eNestedAlice"'],
+    {}
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-a) to (-2em, 0cm |- fig-Multi-parent-b);
+\end{tikzpicture}
+}
+\label{fig:Multi-parent-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    {},
+    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-c);
+\end{tikzpicture}
+}
+\label{fig:Multi-parent-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'],
+   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\path [draw,->,merge] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\end{tikzpicture}
+}
+\label{fig:Multi-parent-d}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    {},
+    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
+   },
+   {}
+  }
+ },
+};
+\path [draw,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge edge] (b);
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
+\end{tikzpicture}
+}
+\label{fig:Multi-parent-e}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
+   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
+    {},
+    "\vNestedAlice" [> "\eMultiParentAlice"']
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.14\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-parent-f}
+TODO: UI
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-parent conflicts.}
+\label{fig:Multi-parent}
+\end{figure*}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureSingleUserCycles
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \newedge{SingleCyclePlus}{R}
+% \newvertex{SingleCyclePlus}{+}
+% \newedge{SingleCycleTimes}{L}
+% \newvertex{SingleCycleTimes}{*}
+% \newedge{SingleCycleCycle}{R}
+% \newcommand{\figureSingleUserCycles}{
+% \begin{figure}
+% %\hfill
+% \hskip0.12\linewidth
+% %%%%%%%%
+% \begin{subfigure}[t]{0.28\linewidth}
+% \centering
+% \caption{
+% \begin{tikzpicture}[remember picture, overlay]
+% \path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
+% %%%%
+% \path (0cm,0cm) node (fig-Single-User Cycles-a) [anchor=base] {\strut};
+% \path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
+% \end{tikzpicture}
+% }
+% \label{fig:Single-User Cycles-a}
+% \begin{tikzpicture}
+% \path (0cm,0cm) graph[graph style] {
+%  root[root style] -> {
+%   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
+%   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
+%     {},
+%     {}
+%   },
+%   {}
+%   }
+%  }
+% };
+% \end{tikzpicture}
+% \end{subfigure}
+% %%%%%%%%
+% \hfill
+% %%%%%%%%
+% \begin{subfigure}[t]{0.28\linewidth}
+% \centering
+% \caption{
+% \begin{tikzpicture}[remember picture, overlay]
+% \path (0cm,0cm) node (fig-Single-User Cycles-b) [anchor=base] {\strut};
+% \path [draw,->,alice] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
+% \end{tikzpicture}
+% }
+% \label{fig:Single-User Cycles-b}
+% \begin{tikzpicture}
+% \path (0cm,0cm) graph[graph style] {
+%  root[root style] -> {
+%   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
+%   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
+%     {},
+%     "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice] -> {
+%      "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice] -> {
+%       {},
+%       {}
+%      },
+%      {}
+%     }
+%   },
+%   }
+%  }
+% };
+% \end{tikzpicture}
+% \end{subfigure}
+% %%%%%%%%
+% \hfill
+% %%%%%%%%
+% \begin{subfigure}[t]{0.28\linewidth}
+% \centering
+% \caption{
+% \begin{tikzpicture}[remember picture, overlay]
+% \path (0cm,0cm) node (fig-Single-User Cycles-c) [anchor=base] {\strut};
+% \path [draw,->,alice] (fig-Single-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Single-User Cycles-c);
+% \end{tikzpicture}
+% }
+% \label{fig:Single-User Cycles-c}
+% \begin{tikzpicture}
+% \path (0cm,0cm) graph[graph style] {
+%  root[root style] -> {
+%   v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
+%   v2/"\vSimpleTimes" [> "\eNestedBob"'],
+%   {}
+%   }
+%  }
+% };
+% \path[draw,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice edge] (v8);
+% \end{tikzpicture}
+% \end{subfigure}
+% %%%%%%%%
+% \hfill{}
+% \caption{Single-User Cycles.}
+% \label{fig:Single-User Cycles}
+% \end{figure}
+% }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureMultiUserCycles
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{MultiCycleTimes}{L}
+\newvertex{MultiCycleTimes}{*}
+\newedge{MultiCyclePlus}{R}
+\newvertex{MultiCyclePlus}{+}
+\newedge{MultiCycleAliceTimes}{R}
+\newedge{MultiCycleAlicePlus}{L}
+\newedge{MultiCycleBobPlus}{R}
+\newedge{MultiCycleBobTimes}{L}
+\newcommand{\figureMultiUserCycles}{
+\begin{figure*}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
+%%%%
+\path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Multi-User Cycles-a);
+\end{tikzpicture}
+}
+\label{fig:Multi-User Cycles-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice],
+    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-User Cycles-a) to (-2em, 0cm |- fig-Multi-User Cycles-b);
+\end{tikzpicture}
+}
+\label{fig:Multi-User Cycles-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   },
+   {
+    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice] -> {
+     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice],
+     {}
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-c);
+\end{tikzpicture}
+}
+\label{fig:Multi-User Cycles-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   },
+   {
+    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob] -> {
+     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob],
+     {}
+    }
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Multi-User Cycles-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
+\path [draw,->,merge] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
+\end{tikzpicture}
+}
+\label{fig:Multi-User Cycles-d}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
+   },
+   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge],
+   {},
+   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge]
+  }
+ }
+};
+\path[draw,-{>[bend]},merge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120,merge edge] (v2);
+\path[draw,-{>[bend]},merge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90,merge edge] (v1);
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Multi-User Cycles.}
+\label{fig:Multi-User Cycles}
+\end{figure*}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \figureDisconnection
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newedge{DisconnectionAlice}{R}
+\newedge{DisconnectionBob}{L}
+\newcommand{\figureDisconnection}{
+\begin{figure*}
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
+%%%%
+\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to (-2em, 0cm |- fig-Disconnection-a);
+\end{tikzpicture}
+}
+\label{fig:Disconnection-a}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'],
+   "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"]
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
+\path [draw,->,alice] (fig-Disconnection-a) to (-2em, 0cm |- fig-Disconnection-b);
+\end{tikzpicture}
+}
+\label{fig:Disconnection-b}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+    {},
+    "\vMultiCyclePlus" [> "\eDisconnectionAlice", >alice edge]
+   },
+   {}
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
+\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-c);
+\end{tikzpicture}
+}
+\label{fig:Disconnection-c}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"] -> {
+   {},
+   "\vMultiCyclePlus" [> "\eMultiCycleBobTimes"] -> {
+    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
+    {}
+   }
+  }
+ }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
+\path [draw,->,merge] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
+\path [draw,->,merge] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
+\end{tikzpicture}
+}
+\label{fig:Disconnection-d}
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+ root[root style] -> {
+  "\vWrapPlus" [> "\eWrapPlus"]
+ }
+};
+\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
+\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCyclePlus" };
+\path[draw,->,merge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45, merge edge] (plus);
+\path[draw,->,merge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135, merge edge] (times);
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill{}
+\caption{Disconnection.}
+\label{fig:Disconnection}
+\end{figure*}
+}
+
 \begin{document}
@@
 
-TODO: group user-level edits (with user visible)
-
-TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts? describe semantic vs presentational in diagrams
-
 \subsection{Program Representation}
@@
 
-% \begin{figure}
-% \centering
-% TODO: UI
-% \includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
-% \caption{A simple tree with a hole}
-% \label{fig:A simple tree with a hole}
-% \end{figure}
-
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
-\label{fig:A simple tree with a hole-a}
-TODO
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
-\label{fig:A simple tree with a hole-b}
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{A simple tree with a hole.
-\autoref{fig:A simple tree with a hole-a}:~User interface.
-\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
-\label{fig:A simple tree with a hole}
-\end{figure}
-
-TODO: sub figure with diagram of relation of subfigures
-
-TODO: [don't do this] deleted edges (vertecies?) in gray with X in them
-
-TODO: label edges of transitions between graphs
+\figureSimple
 
@@
 TODO: extra space in figure 12 d
-
-TODO: reduce vertical tree sep
 
@@
 
-% In this paper we represent variables and numeric constants
-% as leaves and represent their names or values as
-% arguments to those constructors (e.g.,~\texttt{2} in~\texttt{num(2)}).
-% (In TODO:REF, we discuss how to allow edits within these.)
-
 \subsection{Single-User Actions}
@@
 
-\begin{figure}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-A simple tree with a hole-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:A simple tree with a hole-b}}}};
-%%%%
-\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
-\end{tikzpicture}
-}
-\label{fig:Construction and Deletion-a}
-\label{fig:Construction}
-\label{fig:Construction-b}
-\newedge{ConstructionY}{R}%
-\newvertex{ConstructionY}{y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   "\vSimpleX" [> "\eSimpleX"'],
-   "\vConstructionY" [> "\eConstructionY", >alice edge, alice]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Construction) to[alice edge] (-2em, 0cm |- fig-Deletion);
-\end{tikzpicture}
-}
-\label{fig:Construction and Deletion-b}
-\label{fig:Change-a}
-\label{fig:Change}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   {},
-   "\vConstructionY" [> "\eConstructionY"]
-  }
- }
-};
-\path (-0.75cm,-2.75cm) graph[graph style] {
- "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Construction and Deletion.}
-\end{figure}
+\figureConstructionAndDeletion
 
@@
 \label{sub:Deletion}
-
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/delete-and-add/1.dot}};
-% \path (0,-4cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/2.dot}};
-% \path (0,-8cm) node {\includegraphics[scale=0.75]{graphviz/delete-and-add/3.dot}};
-% \end{tikzpicture}
-% \caption{Delete and Add}
-% \label{fig:Delete and Add}
-% \end{figure}
 
@@
 
-\begin{figure}
-% TODO: naive add that creates a conflict
-% \begin{tikzpicture}
-% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
-% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
-% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
-% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
-% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
-% \end{tikzpicture}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Deletion) to[alice edge] (-2em, 0cm |- fig-Wrapping-a);
-\end{tikzpicture}
-}
-\label{fig:Wrapping-a}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (-0.5cm,-3cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vConstructionY" [> "\eConstructionY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Wrapping-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-a) to[alice edge] (-2em, 0cm |- fig-Wrapping-b);
-\end{tikzpicture}
-}
-\label{fig:Wrapping-b}
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice]
- }
-};
-\path (-0.5cm,-3cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vConstructionY" [> "\eConstructionY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Wrapping-c) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-b) to[alice edge] (-2em, 0cm |- fig-Wrapping-c);
-\end{tikzpicture}
-}
-\label{fig:Wrapping-c}
-\newedge{WrapTimes}{L}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vConstructionY" [> "\eConstructionY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Wrapping.  Note that this figure omits the orphaned \Vertex{\vidSimpleX} since it is no longer relevant to our narrative.}
-\label{fig:Wrapping}
-\end{figure}
+\figureWrapping
 
@@
 
-\begin{figure}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
-\end{tikzpicture}
-}
-\label{fig:Move-a}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-3cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vConstructionY" [> "\eConstructionY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,1ex) node (fig-Moving-b) {};
-\path [draw,->,alice] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
-\end{tikzpicture}
-}
-\label{fig:Move-b}
-\newedge{MoveTimes}{R}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vConstructionY" [> "\eConstructionY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.}
-\label{fig:Move}
-\end{figure}
-
 \marginpar{START REV HERE}
+
+TODO: in this section, discuss how we have to be careful to avoid creating cycles
 
@@
 
+\figureMovement
+
 \subsection{Multi-user Interactions}
@@
 
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,1ex) node (fig-Editing Different Parts of the Code-a) {};
-\path [draw,->,alice] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
-\end{tikzpicture}
-}
-\label{fig:Editing Different Parts of the Code-a}
-\newedge{DifferentAlice}{L}
-\newvertex{DifferentAlice}{u}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice],
-    "\vConstructionY" [> "\eConstructionY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,1ex) node (fig-Editing Different Parts of the Code-b) {};
-\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
-\end{tikzpicture}
-}
-\label{fig:Editing Different Parts of the Code-b}
-\newedge{DifferentBob}{R}
-\newvertex{DifferentBob}{v}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,1ex) node (fig-Editing Different Parts of the Code-c) {};
-\path [draw,->,merge] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
-\path [draw,->,merge] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
-\end{tikzpicture}
-}
-\label{fig:Editing Different Parts of the Code-c}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge],
-    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Editing Different Parts of the Code.}
-\label{fig:Editing Different Parts of the Code}
-\end{figure}
-
-% \begin{figure}
-% \centering
-% TODO: add screenshot of code text to each node
-% TODO: remove orphans from diagram
-% \begin{tikzpicture}
-% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
-% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
-% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
-% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
-% \end{tikzpicture}
-% \caption{Editing Different Parts of the Code}
-% \label{fig:Editing Different Parts of the Code}
-% \end{figure}
-
 To start, Alice and Bob are synchronized with each other and
@@
 
-\begin{figure}
-\centering
-TODO: commutativity diagram of screenshots with two cursors
-\caption{TODO:Commutativity}
-\label{fig:Commutativity}
-\end{figure}
-
-TODO: have star or dot dot arrows for skipping steps
-
-TODO: Arrow coming from 5b to 6a and 6b
-
-TODO: Diamond diagram referencing diagrams
+\figureEditingDifferentParts
+
+\figureCommutativity
 
@@
 
-\begin{figure}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
-\end{tikzpicture}
-}
-\label{fig:Editing Nested Parts of the Code-a}
-\newedge{NestedAlice}{L}
-\newvertex{NestedAlice}{w}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
-\end{tikzpicture}
-}
-\label{fig:Editing Nested Parts of the Code-b}
-\newedge{NestedBob}{L}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
-\path [draw,->,merge] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
-\end{tikzpicture}
-}
-\label{fig:Editing Nested Parts of the Code-c}
-\label{fig:Editing Nested-c}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Editing Nested Parts of the Code.
-\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{\vidDifferentAlice} with \Vertex{\vidNestedAlice}.
-\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{\vidSimpleTimes} to the left child of \Vertex{\vidWrapPlus}.
-\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
-\label{fig:Editing Nested Parts of the Code}
-\end{figure}
-
-% \begin{figure}
-% \centering
-% TODO: add screenshot of code text to each node
-% TODO: remove orphans from diagram
-% TODO: annotate edges with delete and restore
-% \begin{tikzpicture}
-% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
-% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
-% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
-% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
-% \end{tikzpicture}
-% \caption{Editing Nested Parts of the Code}
-% \label{fig:Editing Nested Parts of the Code}
-% \end{figure}
+\figureEditingNestedParts
 
@@
 
-\begin{figure*}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
-\end{tikzpicture}
-}
-\label{fig:Multi-child-a}
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
-\end{tikzpicture}
-}
-\label{fig:Multi-child-b}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
-\path [draw,->,merge] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
-\end{tikzpicture}
-}
-\label{fig:Multi-child-c}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >merge edge, merge],
-    "\vMultiChildBob" [> "\eMultiChildBob", >merge edge, merge]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{}
-\label{fig:Multi-child-d}
-TODO: show UI
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Multi-child conflicts.
-\autoref{fig:Multi-child-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO.
-\autoref{fig:Multi-child-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO.
-\autoref{fig:Multi-child-c}:~The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}.
-\autoref{fig:Multi-child-d}:~User interface showing the multi-child conflict in \autoref{fig:Multi-child-c}.}
-\label{fig:Multi-child}
-\end{figure*}
-
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
-% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
-% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
-% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
-% \end{tikzpicture}
-% \caption{Direct Conflicts}
-% \label{fig:Direct Conflicts}
-% \end{figure}
+\figureMultiChild
 
@@
 
-\begin{figure*}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-child-c-x) to (-2em, 0cm |- fig-Multi-parent-a);
-\end{tikzpicture}
-}
-\label{fig:Multi-parent-a}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-TODO: UI (at least moc up)
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-a) to (-2em, 0cm |- fig-Multi-parent-b);
-\end{tikzpicture}
-}
-\label{fig:Multi-parent-b}
-\newedge{MultiParentAlice}{R}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-c);
-\end{tikzpicture}
-}
-\label{fig:Multi-parent-c}
-\newedge{MultiParentBob}{R}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
-\path [draw,->,merge] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
-\end{tikzpicture}
-}
-\label{fig:Multi-parent-d}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge edge] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
-\end{tikzpicture}
-}
-\label{fig:Multi-parent-e}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    "\vNestedAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Multi-parent conflicts.}
-\label{fig:Multi-parent}
-\end{figure*}
+\figureMultiParent
 
@@
 
-\subsubsection{Single-User Cycles}
-\label{sub:Cycles}
-
-\begin{figure}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
-%%%%
-\path (0cm,0cm) node (fig-Single-User Cycles-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
-\end{tikzpicture}
-}
-\label{fig:Single-User Cycles-a}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Single-User Cycles-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
-\end{tikzpicture}
-}
-\label{fig:Single-User Cycles-b}
-\newedge{SingleCyclePlus}{R}
-\newvertex{SingleCyclePlus}{+}
-\newedge{SingleCycleTimes}{L}
-\newvertex{SingleCycleTimes}{*}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice] -> {
-     "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice] -> {
-      {},
-      {}
-     },
-     {}
-    }
-   },
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Single-User Cycles-c) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Single-User Cycles-b) to (-2em, 0cm |- fig-Single-User Cycles-c);
-\end{tikzpicture}
-}
-\label{fig:Single-User Cycles-c}
-\newedge{SingleCycleCycle}{R}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedBob"'],
-   {}
-  }
- }
-};
-\path[draw,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice edge] (v8);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Single-User Cycles.
-\autoref{fig:Single-User Cycles-a}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}.
-\autoref{fig:Single-User Cycles-b}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}.
-\autoref{fig:Single-User Cycles-c}:~\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by reference as the right child of \Vertex{TODO}.}
-\label{fig:Single-User Cycles}
-\end{figure}
-
-% \begin{figure}
-% \centering
-% TODO: lam x. app(\#1, \hole)
-% \caption{Single-User Cycles}
-% \label{fig:Single-User Cycles}
-% \end{figure}
-
-Another case we must consider is when cycles appear in the graph.
-We categorize these into cycles caused by the action of a single user
-and cycles caused by the interaction of the actions of multiple users.
-
-In a single user context,
-normal insertion and deletion of code by the user cannot create cycles.
-However, it is possible with certain kinds of copy and paste.
-For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles-a}
-and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-There are two ways to interpret the paste action.
-The first interpretation is to create a deep copy of \Vertex{TODO}.
-This results in \autoref{fig:Single-User Cycles-b} and
-does not cause a cycle.
-The second interpretation is to simply add an edge to \Vertex{TODO}.
-This results in \autoref{fig:Single-User Cycles-b}
-and causes a cycle.
-
-Note that not all pastes should be deep copies.
-For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-by a cutting from the old position and pasting to the new position.
-Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-Distinguishing when a paste should be by reference versus by copy
-is ultimately a user interface question.
-Cycles caused by the local user's edits can be detected as soon as a user enters them
-by noting either that the graph would contain a cycle or the vertex
-already has a parent somewhere in the graph.
-
-Thus, as a user interface consideration, it might be best to either
-disallow such edits to at least warn users when their
-edits would create a cycle.
+% \subsubsection{Single-User Cycles}
+% \label{sub:Cycles}
+
+% TODO: cut this section
+
+% \figureSingleUserCycles
+
+% Another case we must consider is when cycles appear in the graph.
+% We categorize these into cycles caused by the action of a single user
+% and cycles caused by the interaction of the actions of multiple users.
+
+% In a single user context,
+% normal insertion and deletion of code by the user cannot create cycles.
+% However, it is possible with certain kinds of copy and paste.
+% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles-a}
+% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
+% There are two ways to interpret the paste action.
+% The first interpretation is to create a deep copy of \Vertex{TODO}.
+% This results in \autoref{fig:Single-User Cycles-b} and
+% does not cause a cycle.
+% The second interpretation is to simply add an edge to \Vertex{TODO}.
+% This results in \autoref{fig:Single-User Cycles-b}
+% and causes a cycle.
+
+% Note that not all pastes should be deep copies.
+% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
+% by a cutting from the old position and pasting to the new position.
+% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
+% Distinguishing when a paste should be by reference versus by copy
+% is ultimately a user interface question.
+% Cycles caused by the local user's edits can be detected as soon as a user enters them
+% by noting either that the graph would contain a cycle or the vertex
+% already has a parent somewhere in the graph.
+
+% Thus, as a user interface consideration, it might be best to either
+% disallow such edits to at least warn users when their
+% edits would create a cycle.
 
@@
 \label{sub:Multi-User Cycles}
-MultiCyclePlus
-\begin{figure*}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Single-User Cycles-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Single-User Cycles-c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Single-User Cycles-c-x) to (-2em, 0cm |- fig-Multi-User Cycles-a);
-\end{tikzpicture}
-}
-\label{fig:Multi-User Cycles-a}
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"'],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-a) to (-2em, 0cm |- fig-Multi-User Cycles-b);
-\end{tikzpicture}
-}
-\label{fig:Multi-User Cycles-b}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-c);
-\end{tikzpicture}
-}
-\label{fig:Multi-User Cycles-c}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
-\path [draw,->,merge] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
-\end{tikzpicture}
-}
-\label{fig:Multi-User Cycles-d}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-   },
-   {
-    v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >merge edge, merge],
-    v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >merge edge, merge]
-   }
-  }
- }
-};
-\path[draw,->,merge] (v1) to ["\eMultiCycleAlicePlus",out=-45,in=-135,merge edge] (v2);
-\path[draw,->,merge] (v2) to ["\eMultiCycleBobTimes",out=-90,in=-90,merge edge] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Multi-User Cycles.}
-\label{fig:Multi-User Cycles}
-\end{figure*}
+
+\figureMultiUserCycles
 
 TODO: NEEDS UI MOCKUP
+
+TODO: paragraph about why single-user cycles cannot happen
 
@@
 
-\begin{figure*}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
-%%%%
-\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to (-2em, 0cm |- fig-Disconnection-a);
-\end{tikzpicture}
-}
-\label{fig:Disconnection-a}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Disconnection-a) to (-2em, 0cm |- fig-Disconnection-b);
-\end{tikzpicture}
-}
-\label{fig:Disconnection-b}
-\newedge{DisconnectionAlice}{R}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    {},
-    "\vMultiCyclePlus" [> "\eDisconnectionAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-c);
-\end{tikzpicture}
-}
-\label{fig:Disconnection-c}
-\newedge{DisconnectionBob}{L}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCyclePlus" [> "\eMultiCycleBobTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
-\path [draw,->,merge] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
-\end{tikzpicture}
-}
-\label{fig:Disconnection-d}
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-1cm,-3cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 1cm,-3cm) graph[graph style] { times/"\vMultiCyclePlus" };
-\path[draw,->,merge] (times) to ["\eMultiCycleBobTimes"anchor=south,out=-135,in=-45, merge edge] (plus);
-\path[draw,->,merge] (plus) to ["\eMultiCycleBobPlus"anchor=north,out=45,in=135, merge edge] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Disconnection.}
-\label{fig:Disconnection}
-\end{figure*}
+\figureDisconnection
 
@@
 \section{TODO}
+
 TODO: leaves

### 2020-11-06 grove-paper.tex
Edits and commutativity merge (Michael D. Adams)

@@
 \centering
+\begin{tikzpicture}
+\path ( 0cm, 0cm) node (a) {Original\\Version};
+% \path (-1cm,-1cm) node (b) [alice] {Alice's Version};
+% \path ( 1cm,-1cm) node (c) [bob] {Bob's Version};
+% \path ( 0cm,-2cm) node (d) {Combined Version};
+% \path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
+% \path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
+% \path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
+\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
+\end{tikzpicture}
 
@@
 TODO: include legend of arrow colors
+
+TODO: move into section 2.3.1
+
+TODO: star on multi-step arrows
 
@@
 
-TODO: explain red, green, blue
-
 TODO: $+_32$ and L$_37$
@@
 \label{sub:Single-User Actions}
+
+TODO: explain cross-page arrows
 
@@
 
-\subsubsection{Movement Actions}
-\label{sub:Move}
+\subsubsection{Repositioning Actions}
+\label{sub:Repositioning}
+
+If a user want to move code from one place to another we have two options.
+The first is to delete the code from its old position, and
+reconstruct the code in its new position.
+However, edges and vertices have unique identifiers in Grove 
+to support the collaborative editing features we discuss next,
+so reconstruction will not preserve identity of vertices and edges. 
+Consequently, we include explicit repositioning actions that preserve identity.
+These would be triggered by, for example, a drag and drop, or a cut and paste
+(but not copy and paste, or a second paste after a cut, which do not need to
+preserve identity and so would rely on reconstruction as already described).
+
+In terms of the graph, a repositioning action involves simply deleting the edge
+from the original parent then adding an edge from the new parent.
+Thus the code is actually \textit{moved}, not copied.
+
+For example, in \autoref{fig:Move} Alice continues her edits by
+moving \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
+This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
+(a) delete \eWrapPlus leaving \vSimpleTimes temporarily orphaned and 
+(b) add \eMoveTimes to \vSimpleTimes in the right child position of \vWrapPlus.
+
+\figureMovement
+
+\subsection{Multi-user Interactions}
+\label{sub:Multi-user Interactions}
 
@@
 
-TODO: in this section, discuss how we have to be careful to avoid creating cycles
-
-A consequence of our design is that a move operation
-does not have to delete code from one position and create a copy
-in another position.
-Instead, by deleting just the edge to the code to be moved and then
-adding a new edge to that code somewhere else, the actual code is
-moved, not just a copy.
-
-For example, \autoref{fig:Move} Alice
-moves \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \Edge{\eidWrapPlus} leaving \Vertex{\vidSimpleTimes} temporarily orphaned and 
-(b) add an edge (\Edge{\eidMoveTimes}) to \Vertex{\vidSimpleTimes} in the right child position of \Vertex{\vidWrapPlus}.
-
-This move semantics gives us a richer structure than
+TODO: mention that id numbers are global (or maybe use A1 and B1 for different users?)
+
+TODO: (Place somewhere) This move semantics gives us a richer structure than
 when treating code as a list of lines.  We exploit
 this in Section~REF:TODO in order merge edits that involve moving code.
-
-Of course, if the user wants a genuine copy
-and not just a move, that can be achieved
-by creating new vertexes with fresh identifiers.
-(TODO: figure for this?)
-
-This is similar to the paste with formatting versus
-paste without formatting.
-
-If only one restore, then it does what you expect (basically a cut and paste)
-
-If multiple replicas restore to different places, then you get some complexity:
-
-\figureMovement
-
-\subsection{Multi-user Interactions}
-\label{sub:Multi-user Interactions}
-
-TODO: mention that id numbers are global (or maybe use A1 and B1 for different users?)
 
@@
 
-\figureCommutativity
+%\figureCommutativity
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path ( 0cm, 0cm) node (a) [align=center]       {Original \\ Version};
+\path (-1cm,-2cm) node (b) [align=center,alice] {Alice's  \\ Version};
+\path ( 1cm,-2cm) node (c) [align=center,bob]   {Bob's    \\ Version};
+\path ( 0cm,-4cm) node (d) [align=center]       {Combined \\ Version};
+\path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
+\path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
+\path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
+\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
+\end{tikzpicture}
+
+TODO: commutativity diagram of screenshots with two cursors
+
+TODO: include legend of arrow colors
+
+TODO: move into section 2.3.1
+
+TODO: star on multi-step arrows
+
+\caption{TODO:Commutativity}
+\label{fig:Commutativity}
+\end{figure}
 

### 2020-11-07 grove-paper.tex
Removed \Vertex and \Edge (Michael D. Adams)

@@
 \hfill{}
-\caption{Wrapping.  Note that this figure omits the orphaned \Vertex{\vidSimpleX} since it is no longer relevant to our narrative.}
+\caption{Wrapping.  Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to our narrative.}
 \label{fig:Wrapping}
@@
 \caption{Editing Nested Parts of the Code.
-\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{\vidDifferentAlice} with \Vertex{\vidNestedAlice}.
-\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{\vidSimpleTimes} to the left child of \Vertex{\vidWrapPlus}.
+\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \vDifferentAlice{} with \vNestedAlice{}.
+\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \vSimpleTimes{} to the left child of \vWrapPlus{}.
 \autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
@@
 
+TODO: star on multi-step arrows
+
 Motivation:
@@
 
-TODO: $+_32$ and L$_37$
-
-TODO: extra space in figure 12 d
-
-TODO: list actions in an overview subfigure
-
-TODO: maybe label actions between
-
-%We represent expression contain holes as part of the editor state 
 The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
@@
 
-%We present this from two perspectives.
-%One is the underlying graph model, the other is the interface a user might see, which omits these details unless needed.
-
-%which involves explicit graph edges, positions, and add and delete operations.
-
 Program state is a directed graph with a distinguished root vertex.
@@
 Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)},\todo{var} respectively.
+In \autoref{fig:A simple tree with a hole-b}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)},\todo{var} respectively.
 The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{\vidSimpleTimes}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vSimpleTimes{}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
-This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
-created vertex, \Vertex{\vidConstructionY}, containing the variable reference~\texttt{y}.
+This action corresponds to creating \eConstructionY{} from \vSimpleX{} at the \texttt{R} position to a newly
+created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
 The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.
@@
 Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \Edge{\eidConstructionY} as shown in \autoref{fig:Change-a}. 
-Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children would remain connected to it.
+This is modeled by deleting \eConstructionY{} as shown in \autoref{fig:Change-a}. 
+Notice that \vConstructionY{} continues to exist, and if it had any children, those children would remain connected to it.
 In our system, once a vertex is created it is never deleted.
@@
 Once an edge with a particular identifier is deleted, it cannot be recreated.
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX}.
-
-% Note that we do not explicitly add vertexes.
-% Our only actions are deleting or adding \textit{edges}.
-% Vertexes are implicitly created whenever an edge
-% is created that points to it.
-% Deleting vertexes is not necessary, as we can simply
-% remove the edge pointing to it.
-
-% TODO: explain that multiple delete or add are idempotent
-
-% TODO: explain that delete overrides add
+Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
 
@@
 
-Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
+Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
 If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
@@
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrapping}: 
-(a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex, and
-(c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
+(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
+(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
+(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
 (The choice of wrapping with a bias for the left child position is arbitrary.)
@@
 
-TODO: star on multi-step arrows
-
 \caption{TODO:Commutativity}
@@
 On the other hand, Bob moves the~\texttt{u * v} from the left of the plus to the right of the plus
-by deleting \Edge{\eidMoveTimes} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
+by deleting \eMoveTimes{} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
 Since vertex and edge identifiers are preserved by these,
@@
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of \Edge{\eidMoveTimes} before making her
+Alice might receive Bob's deletion of \eMoveTimes{} before making her
 edits and before receiving Bob's addition of \Edge{19}.
@@
 These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
-In particular pay attention to the edges going out of the \texttt{right} index of \Vertex{\vidWrapPlus}.
+In particular pay attention to the edges going out of the \texttt{right} index of \vWrapPlus{}.
 In \autoref{fig:Multi-child-a} it is \Edge{21}, which points to the variable~\texttt{x},
@@
 We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
-of \Vertex{\vidWrapPlus}.
+of \vWrapPlus{}.
 This results in \autoref{fig:Multi-child-c}.
@@
 represented by the fact that there are multiple edges going out of one index (i.e., the \texttt{right}
-index of \Vertex{\vidWrapPlus}).
+index of \vWrapPlus{}).
 In the graph representation, we flag this conflict by highlighting its edges
@@
 TODO: rename indices to positions
+
 TODO: look into using bold plus and minus (google results)

### 2020-11-09 grove-paper.tex
Add stars to figures (Michael D. Adams)

@@
 \tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
+\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\large\textbf{${}^{*}$}}}}}
+
+%%%% From https://tex.stackexchange.com/questions/104119/star-next-to-arrowhead-in-tikz
+\usepgflibrary{plotmarks}
+\tikzset{
+    to*/.style={
+        shorten >=.25em,#1-to,
+        to path={-- node[inner sep=0pt,at end,sloped] {${}^*$} (\tikztotarget) \tikztonodes}
+    },
+    to*/.default=
+}
+
+\makeatletter
+\pgfarrowsdeclare{to*}{to*}
+{
+  \pgfutil@tempdima=-0.84pt%
+  \advance\pgfutil@tempdima by-1.3\pgflinewidth%
+  \pgfutil@tempdimb=0.21pt%
+  \advance\pgfutil@tempdimb by.625\pgflinewidth%
+  \advance\pgfutil@tempdimb by2.5pt%
+  \pgfarrowsleftextend{+\pgfutil@tempdima}
+  \pgfarrowsrightextend{+\pgfutil@tempdimb}
+}
+{
+  \pgfutil@tempdima=0.28pt%
+  \advance\pgfutil@tempdima by.3\pgflinewidth%
+  \pgfsetlinewidth{0.8\pgflinewidth}
+  \pgfsetdash{}{+0pt}
+  \pgfsetroundcap
+  \pgfsetroundjoin
+  \pgfpathmoveto{\pgfqpoint{-3\pgfutil@tempdima}{4\pgfutil@tempdima}}
+  \pgfpathcurveto
+  {\pgfqpoint{-2.75\pgfutil@tempdima}{2.5\pgfutil@tempdima}}
+  {\pgfqpoint{0pt}{0.25\pgfutil@tempdima}}
+  {\pgfqpoint{0.75\pgfutil@tempdima}{0pt}}
+  \pgfpathcurveto
+  {\pgfqpoint{0pt}{-0.25\pgfutil@tempdima}}
+  {\pgfqpoint{-2.75\pgfutil@tempdima}{-2.5\pgfutil@tempdima}}
+  {\pgfqpoint{-3\pgfutil@tempdima}{-4\pgfutil@tempdima}}
+  \pgfusepathqstroke
+  \begingroup
+    \pgftransformxshift{2.5pt}
+    \pgftransformyshift{2pt}
+    \pgftransformscale{.7}
+    \pgfuseplotmark{asterisk}
+  \endgroup
+%  \pgftext[left,y=2pt]{$\scriptstyle *$}
+}
+
+\pgfarrowsdeclare{*to}{*to}
+{
+  \pgfutil@tempdima=-0.84pt%
+  \advance\pgfutil@tempdima by-1.3\pgflinewidth%
+  \pgfutil@tempdimb=0.21pt%
+  \advance\pgfutil@tempdimb by.625\pgflinewidth%
+  \advance\pgfutil@tempdimb by2.5pt%
+  \pgfarrowsleftextend{+\pgfutil@tempdima}
+  \pgfarrowsrightextend{+\pgfutil@tempdimb}
+}
+{
+  \pgfutil@tempdima=0.28pt%
+  \advance\pgfutil@tempdima by.3\pgflinewidth%
+  \pgfsetlinewidth{0.8\pgflinewidth}
+  \pgfsetdash{}{+0pt}
+  \pgfsetroundcap
+  \pgfsetroundjoin
+  \pgfpathmoveto{\pgfqpoint{-3\pgfutil@tempdima}{4\pgfutil@tempdima}}
+  \pgfpathcurveto
+  {\pgfqpoint{-2.75\pgfutil@tempdima}{2.5\pgfutil@tempdima}}
+  {\pgfqpoint{0pt}{0.25\pgfutil@tempdima}}
+  {\pgfqpoint{0.75\pgfutil@tempdima}{0pt}}
+  \pgfpathcurveto
+  {\pgfqpoint{0pt}{-0.25\pgfutil@tempdima}}
+  {\pgfqpoint{-2.75\pgfutil@tempdima}{-2.5\pgfutil@tempdima}}
+  {\pgfqpoint{-3\pgfutil@tempdima}{-4\pgfutil@tempdima}}
+  \pgfusepathqstroke
+  \begingroup
+    \pgftransformxshift{2.5pt}
+    \pgftransformyshift{-2pt}
+    \pgftransformscale{.7}
+    \pgfuseplotmark{asterisk}
+  \endgroup
+%  \pgftext[left,y=-2pt]{$\scriptstyle *$}
+}
+
+\makeatother
 
@@
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-b) [anchor=base] {\strut};
-\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-child-c-x) to (-2em, 0cm |- fig-Multi-parent-a);
+\path [draw,->,alice] (fig-Multi-child-c-x) to[star] (-2em, 0cm |- fig-Multi-parent-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-a) to (-2em, 0cm |- fig-Multi-parent-b);
+\path [draw,->,alice] (fig-Multi-parent-a) to[star] (-2em, 0cm |- fig-Multi-parent-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-c);
+\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-parent-c);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Multi-User Cycles-a);
+\path [draw,->,alice] (fig-Multi-parent-e-x) to[star] (-2em, 0cm |- fig-Multi-User Cycles-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-a) to (-2em, 0cm |- fig-Multi-User Cycles-b);
+\path [draw,->,alice] (fig-Multi-User Cycles-a) to[star] (-2em, 0cm |- fig-Multi-User Cycles-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-c);
+\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-User Cycles-c);
 \end{tikzpicture}
@@
 \hfill{}
-\caption{Multi-User Cycles.}
+\caption{Cycles.}
 \label{fig:Multi-User Cycles}
@@
 \path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to (-2em, 0cm |- fig-Disconnection-a);
+\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to[star] (-2em, 0cm |- fig-Disconnection-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Disconnection-a) to (-2em, 0cm |- fig-Disconnection-b);
+\path [draw,->,alice] (fig-Disconnection-a) to[star] (-2em, 0cm |- fig-Disconnection-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-c);
+\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Disconnection-c);
 \end{tikzpicture}
@@
 \path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
-\path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
-\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
+\path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Sync} (d);
+\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Sync} (d);
 \end{tikzpicture}
@@
 TODO: move into section 2.3.1
+
+TODO: layout horizontally
 
@@
 
-Pijul
+Pijul and Anu
 

### 2020-11-09 grove-paper.tex
Make merge steps wavy (Michael D. Adams)

--- 2020-11-07 grove-paper.tex
@@
   bending,      % bending arrow tips
+  decorations.pathmorphing,   % Provides wavy edges
   graphs,       % Graph *notation*
@@
 \tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
+\tikzset{alice node/.style={alice}}
+\tikzset{alice step/.style={alice}}
 \tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
+\tikzset{bob node/.style={bob}}
+\tikzset{bob step/.style={bob}}
 \tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
+\tikzset{merge node/.style={merge}}
+\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
+\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\large\textbf{${}^{*}$}}}}}
 
@@
 \path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
+\path [draw,->,alice step] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
 \end{tikzpicture}
@@
    "\vSimpleX" [> "\eSimpleX"'],
-   "\vConstructionY" [> "\eConstructionY", >alice edge, alice]
+   "\vConstructionY" [> "\eConstructionY", >alice edge, alice node]
   }
@@
 \path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Construction) to[alice edge] (-2em, 0cm |- fig-Deletion);
+\path [draw,->,alice step] (fig-Construction) to (-2em, 0cm |- fig-Deletion);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Deletion) to[alice edge] (-2em, 0cm |- fig-Wrapping-a);
+\path [draw,->,alice step] (fig-Deletion) to (-2em, 0cm |- fig-Wrapping-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Wrapping-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-a) to[alice edge] (-2em, 0cm |- fig-Wrapping-b);
+\path [draw,->,alice step] (fig-Wrapping-a) to (-2em, 0cm |- fig-Wrapping-b);
 \end{tikzpicture}
@@
  root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice]
+  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
  }
@@
 \path (0cm,0cm) node (fig-Wrapping-c) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-b) to[alice edge] (-2em, 0cm |- fig-Wrapping-c);
+\path [draw,->,alice step] (fig-Wrapping-b) to (-2em, 0cm |- fig-Wrapping-c);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
+\path [draw,->,alice step] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Moving-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
+\path [draw,->,alice step] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\path [draw,->,alice step] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
 \end{tikzpicture}
@@
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice],
+    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice node],
     "\vConstructionY" [> "\eConstructionY"]
@@
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-b) [anchor=base] {\strut};
-\path [draw, bob,->] (fig-Moving-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\path [draw,->,bob step] (fig-Moving-b) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
 \end{tikzpicture}
@@
     {},
-    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob]
+    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob node]
   }
@@
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-c) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
-\path [draw,->,merge] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path [draw,->,merge step] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path [draw,->,merge step] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
 \end{tikzpicture}
@@
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge],
-    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge]
+    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge node],
+    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge node]
   }
@@
 \centering
-\begin{tikzpicture}
-\path ( 0cm, 0cm) node (a) {Original\\Version};
-% \path (-1cm,-1cm) node (b) [alice] {Alice's Version};
-% \path ( 1cm,-1cm) node (c) [bob] {Bob's Version};
-% \path ( 0cm,-2cm) node (d) {Combined Version};
-% \path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
-% \path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
-% \path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
-\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
-\end{tikzpicture}
 
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Editing Different Parts of the Code-c-x) to (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c-x) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
 \end{tikzpicture}
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice],
+    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice node],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
-\path [draw,->,merge] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
 \end{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge],
+    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge node],
     "\vDifferentBob" [> "\eDifferentBob"]
@@
 \path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
+\path [draw,->,alice step] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
 \end{tikzpicture}
@@
    {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice]
+    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
    }
@@
 \path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
+\path [draw,->,bob step] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
 \end{tikzpicture}
@@
    },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob]
+   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
   }
@@
 \path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
-\path [draw,->,merge] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
+\path [draw,->,merge step] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
+\path [draw,->,merge step] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
 \end{tikzpicture}
@@
    },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge],
+   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
    {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge]
+   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
   }
@@
 \path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-child-c-x) to (-2em, 0cm |- fig-Multi-parent-a);
+\path [draw,->,alice step] (fig-Multi-child-c-x) to[star] (-2em, 0cm |- fig-Multi-parent-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-a) to (-2em, 0cm |- fig-Multi-parent-b);
+\path [draw,->,alice step] (fig-Multi-parent-a) to[star] (-2em, 0cm |- fig-Multi-parent-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-parent-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-c);
+\path [draw,->,bob step] (fig-Multi-parent-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-parent-c);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
-\path [draw,->,merge] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\path [draw,->,merge step] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\path [draw,->,merge step] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
 \end{tikzpicture}
@@
 };
-\path [draw,->,merge] (a) to ["\eMultiParentBob",out=-60,in=60,merge edge] (b);
+\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
+\path [draw,->,alice step] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
 \end{tikzpicture}
@@
 % \path (0cm,0cm) node (fig-Single-User Cycles-a) [anchor=base] {\strut};
-% \path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
+% \path [draw,->,alice step] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
 % \end{tikzpicture}
@@
 % \path (0cm,0cm) node (fig-Single-User Cycles-b) [anchor=base] {\strut};
-% \path [draw,->,alice] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
+% \path [draw,->,alice step] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
 % \end{tikzpicture}
@@
 %     {},
-%     "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice] -> {
-%      "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice] -> {
+%     "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice node] -> {
+%      "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice node] -> {
 %       {},
@@
 % \path (0cm,0cm) node (fig-Single-User Cycles-c) [anchor=base] {\strut};
-% \path [draw,->,alice] (fig-Single-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Single-User Cycles-c);
+% \path [draw,->,alice step] (fig-Single-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Single-User Cycles-c);
 % \end{tikzpicture}
@@
 % };
-% \path[draw,->,alice] (v2) to ["\eSingleCycleCycle",out=-60,in=-60,alice edge] (v8);
+% \path[draw,->,alice edge] (v2) to ["\eSingleCycleCycle",out=-60,in=-60] (v8);
 % \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Multi-User Cycles-a);
+\path [draw,->,alice step] (fig-Multi-parent-e-x) to[star] (-2em, 0cm |- fig-Multi-User Cycles-a);
 \end{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice]
+    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
+    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
    },
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-a) to (-2em, 0cm |- fig-Multi-User Cycles-b);
+\path [draw,->,alice step] (fig-Multi-User Cycles-a) to[star] (-2em, 0cm |- fig-Multi-User Cycles-b);
 \end{tikzpicture}
@@
    {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice],
+    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
+     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
      {}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Multi-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-c);
+\path [draw,->,bob step] (fig-Multi-User Cycles-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-User Cycles-c);
 \end{tikzpicture}
@@
    {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob],
+    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
+     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
      {}
@@
 \path (0cm,0cm) node (fig-Multi-User Cycles-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
-\path [draw,->,merge] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
+\path [draw,->,merge step] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
+\path [draw,->,merge step] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
 \end{tikzpicture}
@@
    },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge],
+   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
    {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge]
-  }
- }
-};
-\path[draw,-{>[bend]},merge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120,merge edge] (v2);
-\path[draw,-{>[bend]},merge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90,merge edge] (v1);
+   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
+  }
+ }
+};
+\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
+\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Multi-User Cycles-d-x) to (-2em, 0cm |- fig-Disconnection-a);
+\path [draw,->,alice step] (fig-Multi-User Cycles-d-x) to[star] (-2em, 0cm |- fig-Disconnection-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
-\path [draw,->,alice] (fig-Disconnection-a) to (-2em, 0cm |- fig-Disconnection-b);
+\path [draw,->,alice step] (fig-Disconnection-a) to[star] (-2em, 0cm |- fig-Disconnection-b);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
-\path [draw,->,bob] (fig-Disconnection-a) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-c);
+\path [draw,->,bob step] (fig-Disconnection-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Disconnection-c);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
-\path [draw,->,merge] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
-\path [draw,->,merge] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
+\path [draw,->,merge step] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
+\path [draw,->,merge step] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
 \end{tikzpicture}
@@
 \path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCyclePlus" };
-\path[draw,->,merge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45, merge edge] (plus);
-\path[draw,->,merge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135, merge edge] (times);
+\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
+\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
 \end{tikzpicture}
@@
 \path ( 0cm, 0cm) node (a) [align=center]       {Original \\ Version};
-\path (-1cm,-2cm) node (b) [align=center,alice] {Alice's  \\ Version};
-\path ( 1cm,-2cm) node (c) [align=center,bob]   {Bob's    \\ Version};
+\path (-1cm,-2cm) node (b) [align=center,alice node] {Alice's  \\ Version};
+\path ( 1cm,-2cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
 \path ( 0cm,-4cm) node (d) [align=center]       {Combined \\ Version};
-\path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
-\path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
-\path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
-\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
+\path [draw,->,alice step] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
+\path [draw,->,bob step]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
+\path [draw,->,merge step] (b) -- node [pos=0.3,auto,swap] {Sync} (d);
+\path [draw,->,merge step] (c) -- node [pos=0.3,auto]      {Sync} (d);
 \end{tikzpicture}

### 2020-11-10 grove-paper.tex
Finish figured and captions (Michael D. Adams)

@@
 \hfill{}
-\caption{A simple tree with a hole.
-\autoref{fig:A simple tree with a hole-a}:~User interface.
-\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
+\caption{Example of Code that Includes a Hole.
+\autoref{fig:A simple tree with a hole-a}:~User Interface for the Code.
+\autoref{fig:A simple tree with a hole-b}:~Graph Representation of the Code.}
 \label{fig:A simple tree with a hole}
@@
 \hfill{}
-\caption{Construction and Deletion.}
+\caption{Example of Construction and Deletion.}
 \end{figure}
@@
 \hfill{}
-\caption{Wrapping.  Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to our narrative.}
+\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to our narrative.)}
 \label{fig:Wrapping}
@@
 \hfill{}
-\caption{Moving \vSimpleTimes{} to the right child of~\vWrapPlus{}.}
+\caption{Example of Repositioning Code.}
 \label{fig:Move}
@@
 \hfill{}
-\caption{Editing Different Parts of the Code.}
+\caption{Example of Users Editing Different Parts of the Code.}
 \label{fig:Editing Different Parts of the Code}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\figureCommutativity}{
-\begin{figure}
-\centering
-
-TODO: commutativity diagram of screenshots with two cursors
-
-TODO: include legend of arrow colors
-
-TODO: move into section 2.3.1
-
-TODO: star on multi-step arrows
-
-\caption{TODO:Commutativity}
-\label{fig:Commutativity}
-\end{figure}
-}
+% \newcommand{\figureCommutativity}{
+% \begin{figure}
+% \centering
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}
+% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
+% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
+% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
+% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
+% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
+% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
+% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
+% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
+% \end{tikzpicture}
+% \caption{TODO:Commutativity}
+% \label{fig:Commutativity}
+% \end{figure}
+% }
 
@@
 \hfill{}
-\caption{Editing Nested Parts of the Code.
-\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \vDifferentAlice{} with \vNestedAlice{}.
-\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \vSimpleTimes{} to the left child of \vWrapPlus{}.
-\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
+\caption{Example of Users Editing Nested Parts of the Code.}
 \label{fig:Editing Nested Parts of the Code}
@@
 \hfill{}
-\caption{Multi-child conflicts.}
+\caption{Example of Multi-Child Conflicts.}
 \label{fig:Multi-child}
@@
 \hfill{}
-\caption{Multi-parent conflicts.}
+\caption{Example of Multi-Parent Conflicts.}
 \label{fig:Multi-parent}
@@
 \hfill{}
-\caption{Cycles.}
+\caption{Example of Cycles.}
 \label{fig:Multi-User Cycles}
@@
 \hfill{}
-\caption{Disconnection.}
+\caption{Example of Disconnection.}
 \label{fig:Disconnection}
@@
 
-TODO: star on multi-step arrows
-
 Motivation:
+
 - collaborative editing (both synchronous ala Google Docs and asynchronous version control) 
   is good and important as computing grows
+
 - semantic structure editing is good because it solves the gap problem (semantic editor services 
   are always available) -- cite Hazelnut papers (talk about holes)
+
 - previous approaches to collaborative editing have limitations
+
   - diff/merge based approaches (trying to solve the inverse problem based on final states -- 
@@
     knowledge we have about the edits!)
+
   - operational transforms (complexity, you have to patch previous actions based on new actions)
+
   - CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
@@
     convergent by design
+
 - we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
@@
   this is tricky because:
+
   - some edits might be conflicting -- solve this with "conflict holes"
+
   - adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
+
 - since we are commutative, we solve both synchronus and async collaborative editing
+
   - and this resolves issues around merges and conflicts
+
 - contribution of this paper is to solve these problems from type-theoretic first principles:
+
   - ...
+
 - Hazel
@@
 
-\todo{Arrows between figures}
 Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
@@
 %\figureCommutativity
-
 \begin{figure}
@@
 \begin{tikzpicture}
-\path ( 0cm, 0cm) node (a) [align=center]       {Original \\ Version};
-\path (-1cm,-2cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-\path ( 1cm,-2cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-\path ( 0cm,-4cm) node (d) [align=center]       {Combined \\ Version};
-\path [draw,->,alice step] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
-\path [draw,->,bob step]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
-\path [draw,->,merge step] (b) -- node [pos=0.3,auto,swap] {Sync} (d);
-\path [draw,->,merge step] (c) -- node [pos=0.3,auto]      {Sync} (d);
-\end{tikzpicture}
-
-TODO: commutativity diagram of screenshots with two cursors
-
-TODO: include legend of arrow colors
-
-TODO: move into section 2.3.1
-
-TODO: layout horizontally
-
-\caption{TODO:Commutativity}
+\path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
+\path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
+\path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
+\path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
+\path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
+\path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
+\path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
+\path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
+\end{tikzpicture}
+\caption{Commutativity of Edits.}
 \label{fig:Commutativity}

### 2020-11-10 grove-paper.tex
Fix spacing after vertex and edge names (Michael D. Adams)

@@
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapPlus leaving \vSimpleTimes temporarily orphaned and 
-(b) add \eMoveTimes to \vSimpleTimes in the right child position of \vWrapPlus.
+(a) delete \eWrapPlus{} leaving \vSimpleTimes{} temporarily orphaned and 
+(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
 

### 2020-11-12 grove-paper.tex
Fix placeholder references to vertexes and edges (Michael D. Adams)

@@
 \tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\large\textbf{${}^{*}$}}}}}
+\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{*}$}}}}}
 
@@
    "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vMultiCyclePlus" [> "\eMultiCycleBobPlus"]
+   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
   }
@@
     {},
-    "\vMultiCyclePlus" [> "\eDisconnectionAlice", >alice edge]
+    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice edge]
    },
@@
    {},
-   "\vMultiCyclePlus" [> "\eMultiCycleBobTimes"] -> {
+   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
     "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
@@
 \path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCyclePlus" };
+\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
 \path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
@@
 We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores edits by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores edits by multiple users, Alice and Bob,
+\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
+and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
 each of whom is editing their own instance of a program.
-
-do not model cursors
-but model later in \autoref{sub:Cursors} 
-they are not involved in the communication of graph edits
-though their representation is slightly subtly
 
@@
 Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)},\todo{var} respectively.
+In \autoref{fig:A simple tree with a hole-b}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
 The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
@@
 
-TODO: explain cross-page arrows
-
 In the remainder of this section,
@@
 
-\todo{"user actions" and "graph edits"}
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole}
+To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole-b}
 and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
@@
 
-If a user want to move code from one place to another we have two options.
+If a user wants to reposition code from one place to another, we have two options.
 The first is to delete the code from its old position, and
@@
 so reconstruction will not preserve identity of vertices and edges. 
-Consequently, we include explicit repositioning actions that preserve identity.
-These would be triggered by, for example, a drag and drop, or a cut and paste
+Consequently, we take the second option: explicit repositioning actions that preserve identity.
+These actions would be triggered by, for example, a drag and drop, or a cut and paste
 (but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described).
+preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
 
@@
 from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{moved}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues her edits by
-moving \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
+Thus the code is actually \textit{repositioned}, not copied.
+
+For example, in \autoref{fig:Move} Alice continues by
+repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
@@
 
-\marginpar{START REV HERE}
+Having described single-user actions,
+we turn our attention to how Grove behaves in the presence of multiple users. 
+Everything we discuss in this section generalizes to any number of users,
+but for simplicity we consider only two: Alice and Bob. 
+Alice and Bob each maintain their own editor instance, and perform 
+actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
+We start each user where we left off in the previous section, i.e., in the state shown in \autoref{fig:Move-b}.
+
+\todo{START REV HERE}
 
@@
 that cause cycles or multiple parents
+
+\subsubsection{Editing Different Parts of the Code}
+\label{sub:Editing Different Parts of the Code}
+
+To start, Alice and Bob are synchronized with each other and
+both have the graph in \autoref{fig:Move-b}, which
+represents the code~\texttt{\hole{} + (\hole{} + y)}.
+
+Alice adds~\vDifferentAlice{} as the left child of \vSimpleTimes{}
+and Bob changes \vConstructionY{} to \vDifferentBob{}.
+Before transmitting their edits to each other,
+Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
+and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
+
+Alice and Bob then transmit their graph edits to each other
+and apply the other's graph edits to their own copy of the graph.
+This results in both Alice and Bob having the graph in
+\autoref{fig:Editing Different Parts of the Code-c}.
+
+\figureEditingDifferentParts
+
+%\figureCommutativity
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
+\path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
+\path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
+\path ( 3cm, 0cm) node (d) [align=center]       {Synchronized \\ Version};
+\path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
+\path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
+\path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
+\path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
+\end{tikzpicture}
+\caption{Commutativity of Edits.}
+\label{fig:Commutativity}
+\end{figure}
 
@@
 
-\subsubsection{Editing Different Parts of the Code}
-\label{sub:Editing Different Parts of the Code}
-
-To start, Alice and Bob are synchronized with each other and
-both have the graph in \autoref{fig:Move-b}, which
-represents the code~\texttt{\hole{} + (\hole{} + y)}.
-
-Alice add~\texttt{u} as the left child of \Vertex{TODO}
-and Bob changes the~\texttt{y} to a~\texttt{v}.
-Before transmitting their edits to each other,
-Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
-and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:Editing Different Parts of the Code-c}.
-
-\figureEditingDifferentParts
-
-%\figureCommutativity
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-\path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-\path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-\path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-\path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
-\path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
-\path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-\path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-\end{tikzpicture}
-\caption{Commutativity of Edits.}
-\label{fig:Commutativity}
-\end{figure}
-
 \subsubsection{Editing Nested Parts of the Code}
@@
 After the edits in the previous subsection synchronize,
-Alice changes the~\texttt{v} to a~\texttt{w}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
-On the other hand, Bob moves the~\texttt{u * v} from the left of the plus to the right of the plus
-by deleting \eMoveTimes{} and adding \Edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
+Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
+On the other hand, Bob moves~\vSimpleTimes{} and its children from the left child of~\vWrapPlus{} to its right child
+by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
 Since vertex and edge identifiers are preserved by these,
@@
 
-A consequence of this design is that we allow users to edit
+A consequence of this design is that we allow users to modify
 parts of the code that have been deleted.  For example,
-Alice might receive Bob's deletion of \eMoveTimes{} before making her
-edits and before receiving Bob's addition of \Edge{19}.
+Alice might receive Bob's deletion of~\eMoveTimes{} before making her
+changes and before receiving Bob's addition of~\eNestedBob{}.
 Her edits would then be within the code deleted by Bob.
@@
 
-Finally, we address our first kind of genuine conflict between edits.
-Alice inserts the variable~\texttt{x} to the right side of the~\texttt{+}
-while Bob add the variable~\texttt{y} also to the right side of the~\texttt{+}.
-These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
-In particular pay attention to the edges going out of the \texttt{right} index of \vWrapPlus{}.
-In \autoref{fig:Multi-child-a} it is \Edge{21}, which points to the variable~\texttt{x},
-and in \autoref{fig:Multi-child-b} it is \Edge{23}, which points to the variable~\texttt{y}.
-
-We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
+Finally, we address our first kind of conflict between edits.
+Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
+while Bob adds~\vMultiChildBob{} also as the right child of~\vWrapPlus{}.
+These result in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
+In particular pay attention to the edges going out of the \texttt{R} position of \vWrapPlus{}.
+In \autoref{fig:Multi-child-a} it is~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
+and in \autoref{fig:Multi-child-b} it is~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
+
+We merge these by including \emph{both} edges as edges going out of the \texttt{R} position
 of \vWrapPlus{}.
@@
 The conflict between Alice's and Bob's changes is
-represented by the fact that there are multiple edges going out of one index (i.e., the \texttt{right}
-index of \vWrapPlus{}).
+represented by the fact that there are multiple edges going out of one index (i.e., the \texttt{R}
+position of \vWrapPlus{}).
 In the graph representation, we flag this conflict by highlighting its edges
@@
 one of the edges and editing the other to be the merged value.
-For example, one could delete \Edge{23} and then edit the~\texttt{x}
+For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{}
 to be~\texttt{x * y}.
@@
 In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes \Edge{TODO} and adds \Edge{TODO}, while Bob deletes \Edge{TODO} and adds edge \Edge{TODO}.
-As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting \Edge{TODO} is not a problem.
-However, both \Edge{TODO} (added by Alice) and edge \Edge{TODO} (added by Bob) point to the same vertex.
-This generates REF:FIG, in which \Vertex{TODO} as two parents pointing to it.
+Alice deletes~\eNestedAlice{} and adds~\eMultiParentAlice{}, while
+Bob deletes~\eNestedAlice{} and adds~\eMultiParentBob{}.
+As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting~\eNestedAlice{} is not a problem.
+However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
+This generates \autoref{fig:Multi-parent-d}, in which~\vNestedAlice{} has two parents pointing to it.
 We call this situation a multi-parent conflict.
@@
 
-\subsubsection{Multi-User Cycles}
+\subsubsection{Cycles}
 \label{sub:Multi-User Cycles}
@@
 This graph represents the expression~\texttt{((\hole+\hole)+(\hole*\hole))+\hole}.
-Alice moves the leftmost \texttt{+} into the rightmost hold
-and then the \texttt{*} underneath that.
+Alice moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
+and then~\vMultiCyclePlus{} underneath.
 This results in \autoref{fig:Multi-User Cycles-b},
@@
 On the other hand, Bob does the same move
-but puts the \texttt{+} and under the \texttt{*}.
+but puts~\vMultiCycleTimes{} and under~\vMultiCyclePlus{}.
 This results in \autoref{fig:Multi-User Cycles-c},
@@
 in \autoref{fig:Multi-User Cycles-d}, which has a cycle
-between \Vertex{TODO} and \Vertex{TODO}.
+between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
 This situation is akin to a merge conflict in traditional version control systems.
 Resolving it requires user input, but is as simple as
-either deleting \Edge{TODO} and \Edge{TODO} (thus favoring Alice's version)
-or deleting \Edge{TODO} and \Edge{TODO} (thus favoring Bob's version).
+either deleting~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
+or deleting~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
 Since a cycle like this that is connected to the root vertex always
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(\hole+\hole)+(\hole*\hole)}.
-Alice moves the \texttt{*} under \texttt{+}$_2$.
+This graph represents the expression~\texttt{(\hole*\hole)+(\hole*\hole)}.
+Alice moves~\vMultiCycleTimes{} under~\vSimpleTimes{}.
 This results in \autoref{fig:Disconnection-b},
-which represents the expression~\texttt{((\hole*\hole)+\hole)+\hole}.
-On the other hand, Bob puts \texttt{+}$_2$ and under the \texttt{*}.
+which represents the expression~\texttt{(\hole*(\hole*\hole))+\hole}.
+On the other hand, Bob puts~\vSimpleTimes{} and under~\vMultiCycleTimes{}.
 This results in \autoref{fig:Disconnection-c},
-which represents the expression~\texttt{\hole+((\hole+\hole)*\hole)}.
+which represents the expression~\texttt{\hole+((\hole*\hole)*\hole)}.
 
 When these edits are merged, the result is the graph in \autoref{fig:Disconnection-d}.
-Since Alice deleted the edge from \Vertex{TODO} to \Vertex{TODO}
-and Bob deleted the edge from \Vertex{TODO} to \Vertex{TODO},
-when these are merged \Vertex{TODO} and \Vertex{TODO}
-are completely disconnected from the root.
+Since Alice deleted~\eNestedBob{}
+and Bob deleted~\eMultiCycleAliceTimes{},
+when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
+are disconnected from the root.
 

### 2020-11-13 grove-paper.tex
Revisions to Section 2.3 (Michael D. Adams)

@@
 %%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
+\begin{subfigure}[t]{0.17\textwidth}
 \centering
@@
 \end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{}
+\label{fig:Multi-User Cycles-e}
+TODO: UI
 \end{subfigure}
@@
 \end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.17\linewidth}
+\centering
+\caption{}
+\label{fig:Disconnection-e}
+TODO: UI
 \end{subfigure}
@@
 Having described single-user actions,
-we turn our attention to how Grove behaves in the presence of multiple users. 
-Everything we discuss in this section generalizes to any number of users,
+we turn our attention to how Grove handles multiple users. 
+This section generalizes to any number of users,
 but for simplicity we consider only two: Alice and Bob. 
-Alice and Bob each maintain their own editor instance, and perform 
+Alice and Bob each maintain their own editor instance and perform 
 actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-We start each user where we left off in the previous section, i.e., in the state shown in \autoref{fig:Move-b}.
+We start each user where we left off in the previous section with the state shown in \autoref{fig:Move-b}.
 
@@
 
-We assume multiple instances
-
-Common scenarios first and then consider more unusual scenarios
-that cause cycles or multiple parents
+TODO: Common scenarios first and then consider more unusual scenarios
+that cause cycles or vertices with multiple parents
 
@@
 
-To start, Alice and Bob are synchronized with each other and
-both have the graph in \autoref{fig:Move-b}, which
+To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
 represents the code~\texttt{\hole{} + (\hole{} + y)}.
@@
 Before transmitting their edits to each other,
-Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
-and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
+Alice and Bob have the graphs in \autoref{fig:Editing Different Parts of the Code-a}
+and \autoref{fig:Editing Different Parts of the Code-b}, respectively.
+(Note that thee transition edge from \autoref{fig:Move-b} to
+\autoref{fig:Editing Different Parts of the Code-b} represents
+multiple edges (i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}).
+We thus mark it with a star.)\todo{star size}\todo{"sync"}
 
@@
 
-Note that aside from transmitting the edge deletions and additions,
-the merge requires no coordination between Alice and Bob.
-For example, suppose that Alice and Bob did not transmit their edits to each other
+Note that aside from transmitting edge deletions and additions,
+merging requires no coordination between Alice and Bob.
+For example, suppose Alice and Bob did not transmit their edits to each other
 but to a third user, Chris, that has the original graph in \autoref{fig:Move-b}.
@@
 Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
+For any sequences of edits from Alice and Bob,
 applying them in either order produces the same results.
@@
 
-After the edits in the previous subsection synchronize,
+After Alice and Bob share the edits they made in the previous subsection,
 Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
@@
 by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
-Since vertex and edge identifiers are preserved by these,
+Since edits are based on edge identities,
 both Alice's and Bob's edits can be applied to the other's graph,
@@
 changes and before receiving Bob's addition of~\eNestedBob{}.
-Her edits would then be within the code deleted by Bob.
-TODO: She can restore that code using a restoration action add a fresh edge to it.
+Her edits would then temporarily be within the code deleted by Bob.
+TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
 
@@
 
-TODO: Multi-user merges can still create multi-child conflicts,
-but it is better to not interpret user actions in a way that creates them.
-How to handle multiple children in the same position is explored in \autoref{sub:Multi-child conflicts}.
-
-Finally, we address our first kind of conflict between edits.
-Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
+Next we address what happens whn edits conflict.
+For example, Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
 while Bob adds~\vMultiChildBob{} also as the right child of~\vWrapPlus{}.
-These result in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
-In particular pay attention to the edges going out of the \texttt{R} position of \vWrapPlus{}.
-In \autoref{fig:Multi-child-a} it is~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
-and in \autoref{fig:Multi-child-b} it is~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
-
-We merge these by including \emph{both} edges as edges going out of the \texttt{R} position
+These result in Alice and Bob having the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
+These graphs have two different edges for the~\texttt{R} position of~\vWrapPlus{}.
+\autoref{fig:Multi-child-a} has~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
+and \autoref{fig:Multi-child-b} has~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
+We call this a \emph{multi-child conflict}.
+
+We merge these edits by including \emph{both} edges as edges going out of the \texttt{R} position
 of \vWrapPlus{}.
@@
 The conflict between Alice's and Bob's changes is
-represented by the fact that there are multiple edges going out of one index (i.e., the \texttt{R}
-position of \vWrapPlus{}).
-In the graph representation, we flag this conflict by highlighting its edges
-in red (TODO).
-In the textual representation, we flag this conflict with the notation~\texttt{\{ x | y \}}.
-This is a conflict that must eventually be resolved by the user by deleting
+represented by the fact that there are multiple edges going out of one index.
+In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}
+as shown in \autoref{fig:Multi-child-d}.
+
+This conflict must eventually be resolved by the user by deleting
 one of the edges and editing the other to be the merged value.
@@
 to be~\texttt{x * y}.
-Note that users can continue to edit the code while in this conflicted state.
+(Note that users can continue to edit the code while in this conflicted state.
 It just must be resolved before the the code is run.\footnote{It
 may be possible to develop evaluation models that allow these sorts of conflicts,
-but that is beyond the scope of this paper.}
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution
-
-Multiple adds cause conflicts
-(Not usually presented in the UI)
+but that is beyond the scope of this paper.})
+
+TODO: need example of fully-automatic resolution and semi-automatic resolution.
+Note that as a convenience to the user, certain simple
+conflicts might be automatically resolved,
+but we consider this a higher-level, user-interface consideration.
+
+TODO: Multiple adds cause conflicts (Not usually presented in the UI)
 
@@
 Another place our model is different
-is that changes are explicit rather than being inferred.
+is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
 Finally, in our model, since all changes are in terms of
@@
 This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the structure of the code.
-Finally, our model differs from the traditional model
-in that we have explict add and remove actions
-where the traditional model infers these from a differencing algorithm.
-
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
+that a merge conflict follows the grammatical structure of the code.
 
@@
 
-In addition to conflicts due to multiple edges coming \emph{out of} child position of a particular vertex,
+In addition to conflicts due to multiple edges coming \emph{out of} the same child position of a particular vertex,
 conflicts can manifest as multiple edges going \emph{into} a particular vertex.
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(w + \hole{}) + \hole{})}.
-Alice moves the variable~\texttt{w} to the left-most hole,
-while Bob moves it to the right-most hole.
+This graph represents the expression~\texttt{(w + \hole{}) + \hole{}}.
+Alice moves~\vNestedAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
+while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
 This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
@@
 Bob deletes~\eNestedAlice{} and adds~\eMultiParentBob{}.
-As explained in Subsection~TODO:REF, deletes are idempotent so Alice and Bob both deleting~\eNestedAlice{} is not a problem.
+As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedAlice{} is not a problem.
 However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
@@
 
-TODO: NEEDS UI MOCKUP
-
-TODO: paragraph about why single-user cycles cannot happen
-
+TODO: paragraph about why single-user cycles cannot happen.
 While single-user edits can be checked to see
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{((\hole+\hole)+(\hole*\hole))+\hole}.
+This graph represents the expression~\texttt{((\hole{} + \hole{}) + (\hole{} * \hole{})) + \hole}.
 Alice moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
@@
 This results in \autoref{fig:Multi-User Cycles-b},
-which represents the expression~\texttt{(\hole+\hole)+((\hole*\hole)+\hole)}.
-On the other hand, Bob does the same move
+which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
+On the other hand, Bob does the same
 but puts~\vMultiCycleTimes{} and under~\vMultiCyclePlus{}.
 This results in \autoref{fig:Multi-User Cycles-c},
-which represents the expression~\texttt{(\hole+\hole)+((\hole+\hole)*\hole)}.
+which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
 
@@
 For example, \autoref{fig:Multi-User Cycles-d} would be displayed
-as shown in TODO:FIGURE.
+as shown in \autoref{fig:Multi-User Cycles-e}.
 We leave the user interface considerations of resolving the conflict such a
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(\hole*\hole)+(\hole*\hole)}.
+This graph represents the expression~\texttt{(\hole{} * \hole{}) + (\hole{}  * \hole{})}.
 Alice moves~\vMultiCycleTimes{} under~\vSimpleTimes{}.
 This results in \autoref{fig:Disconnection-b},
-which represents the expression~\texttt{(\hole*(\hole*\hole))+\hole}.
+which represents the expression~\texttt{(\hole{} * (\hole{} * \hole{})) + \hole}.
 On the other hand, Bob puts~\vSimpleTimes{} and under~\vMultiCycleTimes{}.
 This results in \autoref{fig:Disconnection-c},
-which represents the expression~\texttt{\hole+((\hole*\hole)*\hole)}.
+which represents the expression~\texttt{\hole{} + ((\hole{} * \hole{}) * \hole{})}.
 

### 2020-11-13 grove-paper.tex
add informal definition of simple cycles (Eric Griffis)

@@
 
+\figureMultiChild
+
 After Alice and Bob share the edits they made in the previous subsection,
@@
 \label{sub:Multi-child conflicts}
-
-\figureMultiChild
 
@@
 
-\subsubsection{Graph Actions}
-\label{sub:Graph Actions}
+\subsubsection{Graph Edits}
+\label{sub:Graph Edits}
 
@@
 
-\subsection{Edit Actions}
-\label{sub:Edit Actions}
+\subsection{Graph Decomposition}
+\label{sub:Graph Decomposition}
+
+What does a graph decomposition produce? a 4-tuple (reachable, multiparented, deleted, simple cycles).
+
+A simple cycle is a set of vertices connected by a closed path, consisting
+   of:
+
+   1) two or more vertices, all descendants of each other, and
+   2) the vertices of any non-multiparent subtrees reachable from the core, and
+      not reachable from the root vertex.
+
+  To determine the simple cycles of a graph, we:
+
+  1) check if any vertex in the graph is a descendant of itself or some other
+     vertex that is (a descendant of itself),
+  2) find the least vertex in the core of this new(ly discovered) cycle, and
+  3) return a tree connecting every vertex reachable from the least (vertex).
+
+  The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.
+
+<Mathematical definition of above.>
+
+\subsection{User Actions}
+\label{sub:User Actions}
 
@@
 \begin{array}{l}
-  \textrm{Edit Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}} \\
+  \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}} \\
   \s : \G \to Act \times \V^{*} \times \V^{*} \times \V^{*} \\

### 2020-11-15 grove-paper.tex
add code view screenshots
add my orcid
auto-format whitespace
VSCode insists, so I'll wrap it in a pull request, in case the changes
get in the way.
make code in prose match the diagrams
also fix a few small grammar errors and add some whitespace to latex
source

@@
 \label{fig:A simple tree with a hole-a}
-TODO
+\includegraphics{code-views/simple.png}
 \end{subfigure}
@@
 \label{fig:Multi-child-d}
-TODO: show UI
+\includegraphics{code-views/multi-child-conflict.png}
 \end{subfigure}
@@
 \label{fig:Multi-parent-f}
-TODO: UI
+\includegraphics{code-views/multi-parent-conflict.png}
 \end{subfigure}
@@
 \label{fig:Multi-User Cycles-e}
-TODO: UI
+\includegraphics{code-views/cycle.png}
 \end{subfigure}
@@
 \label{fig:Disconnection-e}
-TODO: UI
+\includegraphics{code-views/disconnected-reachable.png}
+\includegraphics{code-views/disconnected-unreachable.png}
 \end{subfigure}
@@
 \author{Eric Griffis}
-\orcid{nnnn-nnnn-nnnn-nnnn}
+\orcid{0000-0003-1693-6172}
 
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vSimpleTimes{}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a right edge coming from \vSimpleTimes{}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
-This action corresponds to creating \eConstructionY{} from \vSimpleX{} at the \texttt{R} position to a newly
+This action corresponds to creating \eConstructionY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
 created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
@@
 Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eConstructionY{} as shown in \autoref{fig:Change-a}. 
-Notice that \vConstructionY{} continues to exist, and if it had any children, those children would remain connected to it.
+This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Change-a}.
+Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
 In our system, once a vertex is created it is never deleted.
@@
 Once an edge with a particular identifier is deleted, it cannot be recreated.
+
 Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
@@
 (The choice of wrapping with a bias for the left child position is arbitrary.)
+
+\newcommand{\eWrapTimesWrapC}{\eWrapTimes}
 
@@
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapPlus{} leaving \vSimpleTimes{} temporarily orphaned and 
+(a) delete \eWrapTimesWrapC{} leaving \vSimpleTimes{} temporarily orphaned and
 (b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
@@
 To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
-represents the code~\texttt{\hole{} + (\hole{} + y)}.
+represents the code~\texttt{\hole{} + (\hole{} * y)}.
 
@@
 and \autoref{fig:Editing Different Parts of the Code-b}, respectively.
-(Note that thee transition edge from \autoref{fig:Move-b} to
+(Note that the transition edge from \autoref{fig:Move-b} to
 \autoref{fig:Editing Different Parts of the Code-b} represents
-multiple edges (i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}).
+multiple edges, i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
 We thus mark it with a star.)\todo{star size}\todo{"sync"}
@@
 Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
-On the other hand, Bob moves~\vSimpleTimes{} and its children from the left child of~\vWrapPlus{} to its right child
+On the other hand, Bob moves~\vSimpleTimes{} and its children from the right child of~\vWrapPlus{} to its left child
 by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
@@
 
-Next we address what happens whn edits conflict.
+Next we address what happens when edits conflict.
 For example, Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(w + \hole{}) + \hole{}}.
+This graph represents the expression~\texttt{(w * \hole{}) + \hole{}}.
 Alice moves~\vNestedAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
@@
 This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
-which represent respectively the expression~\texttt{(\hole{} + w) + \hole{}} for Alice
-and the expression~\texttt{(\hole{} + \hole{}) + w} for Bob.
+which represent respectively the expression~\texttt{(\hole{} * w) + \hole{}} for Alice
+and the expression~\texttt{(\hole{} * \hole{}) + w} for Bob.
 In both cases, these edits are achieved with one edge delete and one edge add.
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{((\hole{} + \hole{}) + (\hole{} * \hole{})) + \hole}.
+This graph represents the expression~\texttt{((\hole{} * \hole{}) * (\hole{} + \hole{})) + \hole}.
 Alice moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
@@
 This results in \autoref{fig:Multi-User Cycles-b},
-which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
+which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
 On the other hand, Bob does the same
-but puts~\vMultiCycleTimes{} and under~\vMultiCyclePlus{}.
+but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
 This results in \autoref{fig:Multi-User Cycles-c},
-which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
+which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
 
@@
 which represents the expression~\texttt{(\hole{} * (\hole{} * \hole{})) + \hole}.
-On the other hand, Bob puts~\vSimpleTimes{} and under~\vMultiCycleTimes{}.
+On the other hand, Bob puts~\vSimpleTimes{} under~\vMultiCycleTimes{}.
 This results in \autoref{fig:Disconnection-c},
-which represents the expression~\texttt{\hole{} + ((\hole{} * \hole{}) * \hole{})}.
+which represents the expression~\texttt{((\hole{} * \hole{}) * \hole{}) + \hole{}}.
 
@@
 
+% XXX deletes are idempotent
+
 We define graph edit actions as a pair of an edge state, excluding $\bot$, and an edge.
@@
 \begin{align*}
-             \edges(G) &= \dom(G) \\
-         \liveEdges(G) &= \{ \e \in \edges(G) | G(\e)=+ \}              \\
-  \vertexChildren(G,v) &= \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
-  \cursorChildren(G,c) &= \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
-         \parents(G,v) &= \{ (c,v',u) \in \liveEdges(G) | v'=v \}
+  \edges(G)            & = \dom(G)                                       \\
+  \liveEdges(G)        & = \{ \e \in \edges(G) | G(\e)=+ \}              \\
+  \vertexChildren(G,v) & = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
+  \cursorChildren(G,c) & = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
+  \parents(G,v)        & = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
 \end{align*}
@@
 \begin{align*}
-        \vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
-    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
- \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
-         \orphans(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}       \\
-    \multiparents(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}       \\
-        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
+  \vertexes(G)        & = \{ v \in \V | \exists (c,v,u) \in \edges(G) \}              \\
+  \liveVertexes(G)    & = \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}          \\
+  \childVertexes(G,v) & = \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \} \\
+  \orphans(G)         & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}      \\
+  \multiparents(G)    & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}      \\
+  \vertex(G,u)        & = v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
 \end{align*}

### 2020-11-15 grove-paper.tex
make code in prose match the diagrams
also fix a few small grammar errors and add some whitespace to latex
source

@@
 
-% $\arraycolsep=2pt\begin{array}{llcl}
-% \mathsf{Typ} & \tau & ::= & t ~\vert~ \aparr{\tau}{\tau} ~\vert~ \aall{t}{\tau} ~\vert~ \arec{t}{\tau} ~\vert~ \aprod{\labelset}{\mapschema{\tau}{i}{\labelset}} ~\vert~ \asum{\labelset}{\mapschema{\tau}{i}{\labelset}}\\
-% \mathsf{Exp} & e & ::= & x ~\vert~ \aelam{\tau}{x}{e} ~\vert~ \aeap{e}{e} ~\vert~ \aetlam{t}{e} ~\vert~ \aetap{e}{\tau} ~\vert~ \aefold{e} ~\vert~ \aetpl{\labelset}{\mapschema{e}{i}{\labelset}} ~\vert~  \aein{\ell}{e} \\
-% & & \vert & \aematchwith{n}{e}{\seqschemaX{r}}\\
-% \mathsf{Rule} & r & ::= & \aematchrule{p}{e}\\
-% \mathsf{Pat} & p & ::= & x  ~\vert~ \aewildp ~\vert~ \aefoldp{p} ~\vert~ \aetplp{\labelset}{\mapschema{p}{i}{\labelset}} ~\vert~ \aeinjp{\ell}{p}
-% \end{array}$
+% XXX "p" is overloaded - could mean pattern or position
 
@@
     \begin{array}{lrlll}
-      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \_                                                            \\
-      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
-      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \_                               \\
+      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \hole                                                            \\
+      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \hole \\
+      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole                               \\
     \end{array}
@@
 
-The language we are modeling is shown in \autoref{fig:Syntax as a grammar}.
-Standard for a simply-typed lambda calculus except that
-all terms have holes (represented by $\_$)
-and the argument to a lambda is its own term.
-The latter of which allows us to express lambdas with a hole for the bound variable.
-
-In order to model this we define
-
-Turn that into a constructor set K + index set I + arity function (has type: K -> $\wp$(I)).
+The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
+standard for a simply-typed lambda calculus except that all types of terms have
+holes (represented by $\hole$). In particular, the argument to a lambda can be
+either a bound variable or a hole.
+
+% In order to model this we define
+
+% Turn that into a constructor set K + position set P + arity function (has type: K -> $\wp$(P)).
 
@@
     \begin{array}{ll}
-      \arity : \K \rightarrow & \wp(\I)                                        \\
+      \arity : \K \rightarrow & \wp(\P)                                        \\
       \hline
@@
 
-TODO: rename indices to positions
-
-TODO: look into using bold plus and minus (google results)
-
 A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states,
@@
 
-A vertex $v = (u, k)$ identifies a distinct instance of the constructor $k$.
-
-A constructor $k \in \K$ represents a different type of node in the target language's abstract syntax tree.
-
-A position $p \in \P$ identifies a location, adjacent to some parent vertex.
-
-Every graph contains a distinguished root vertex, root constructor, and root position, respectively named $v_0$, $k_0$, and $p_0$.
-
-% Define $\sqcup_{edge state}$ (i.e. join)
-
-% lattice ordering $\bot < + < -$ LUB e.g., join A B = C
-
-You can map from syntax to graph by selecting
-unique IDs $u$ (for both edges and vertexes).
-- holes are not explicit in the graph
+A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
+where $k \in \K$ corresponds to a type of expression. Each contructor has an
+arity, which we define as a set of positions. A position $p \in \P$ identifies a
+location, adjacent to some parent vertex, from which edges to child vertices may
+originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
+  Arity} gives the complete mapping from expression constructors to potential edge
+positions. Every graph contains a distinguished root vertex, root constructor,
+and root position, respectively named $v_0$, $k_0$, and $p_0$.
+
+A graph can be derived from an expression in the calculus by starting with an
+empty graph, selecting unique ids $u, u_k$ and a constructor $k$ for the
+outermost expression, and creating the edge $\e = (u, v_0, p_0, (u_k, k))$.
+Then, for each sub-expression at some position $p \in \arity(k)$, select unique
+id $u'$ and constructor $k'$, and create the edge $e' = (u', v, p, v')$, where
+$v = (u_k, k)$ and $v' = (u', k')$. Continuing in this manner, treat each
+sub-expression, starting with the ones at positions in $\arity(k')$, as if it
+were the outermost and create an edge to the vertices representing each of its
+immediate sub-expressions. Note that holes are not modeled explicitly in the
+graph. When an expression has a hole, it means that the graph has no edges
+originating from that position.
+
+% You can map from syntax to graph by selecting
+% unique IDs $u$ (for both edges and vertexes).
+% - holes are not explicit in the graph
 
@@
 We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
-For example, $\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
-and $\Plus \sqcup \Minus = \Minus$, which means that once an edge is deleted it can never be restored.
+For example,
+$\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
+and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$, which means that once an edge is deleted it can never be restored.
+
 %smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.

### 2020-11-16 grove-paper.tex
finish merging last changes

@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a right edge coming from \vSimpleTimes{}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vSimpleTimes{}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
-This action corresponds to creating \eConstructionY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
+This action corresponds to creating \eConstructionY{} from \vSimpleX{} at the \texttt{R} position to a newly
 created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
@@
 Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Change-a}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
+This is modeled by deleting \eConstructionY{} as shown in \autoref{fig:Change-a}.
+Notice that \vConstructionY{} continues to exist, and if it had any children, those children would remain connected to it.
 In our system, once a vertex is created it is never deleted.
@@
 Once an edge with a particular identifier is deleted, it cannot be recreated.
-
 Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
@@
 (The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\newcommand{\eWrapTimesWrapC}{\eWrapTimes}
 
@@
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimesWrapC{} leaving \vSimpleTimes{} temporarily orphaned and
+(a) delete \eWrapPlus{} leaving \vSimpleTimes{} temporarily orphaned and
 (b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
@@
 To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
-represents the code~\texttt{\hole{} + (\hole{} * y)}.
+represents the code~\texttt{\hole{} + (\hole{} + y)}.
 
@@
 and \autoref{fig:Editing Different Parts of the Code-b}, respectively.
-(Note that the transition edge from \autoref{fig:Move-b} to
+(Note that thee transition edge from \autoref{fig:Move-b} to
 \autoref{fig:Editing Different Parts of the Code-b} represents
-multiple edges, i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
+multiple edges (i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}).
 We thus mark it with a star.)\todo{star size}\todo{"sync"}
@@
 Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
-On the other hand, Bob moves~\vSimpleTimes{} and its children from the right child of~\vWrapPlus{} to its left child
+On the other hand, Bob moves~\vSimpleTimes{} and its children from the left child of~\vWrapPlus{} to its right child
 by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
@@
 
-Next we address what happens when edits conflict.
+Next we address what happens whn edits conflict.
 For example, Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(w * \hole{}) + \hole{}}.
+This graph represents the expression~\texttt{(w + \hole{}) + \hole{}}.
 Alice moves~\vNestedAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
@@
 This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
-which represent respectively the expression~\texttt{(\hole{} * w) + \hole{}} for Alice
-and the expression~\texttt{(\hole{} * \hole{}) + w} for Bob.
+which represent respectively the expression~\texttt{(\hole{} + w) + \hole{}} for Alice
+and the expression~\texttt{(\hole{} + \hole{}) + w} for Bob.
 In both cases, these edits are achieved with one edge delete and one edge add.
@@
 and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{((\hole{} * \hole{}) * (\hole{} + \hole{})) + \hole}.
+This graph represents the expression~\texttt{((\hole{} + \hole{}) + (\hole{} * \hole{})) + \hole}.
 Alice moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
@@
 This results in \autoref{fig:Multi-User Cycles-b},
-which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
+which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
 On the other hand, Bob does the same
-but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
+but puts~\vMultiCycleTimes{} and under~\vMultiCyclePlus{}.
 This results in \autoref{fig:Multi-User Cycles-c},
-which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
+which represents the expression~\texttt{(\hole{} + \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
 
@@
 which represents the expression~\texttt{(\hole{} * (\hole{} * \hole{})) + \hole}.
-On the other hand, Bob puts~\vSimpleTimes{} under~\vMultiCycleTimes{}.
+On the other hand, Bob puts~\vSimpleTimes{} and under~\vMultiCycleTimes{}.
 This results in \autoref{fig:Disconnection-c},
-which represents the expression~\texttt{((\hole{} * \hole{}) * \hole{}) + \hole{}}.
+which represents the expression~\texttt{\hole{} + ((\hole{} * \hole{}) * \hole{})}.
 
@@
 
-% XXX deletes are idempotent
-
 We define graph edit actions as a pair of an edge state, excluding $\bot$, and an edge.

### 2020-11-16 grove-paper.tex
Tweaks to paper

--- 2020-11-15 grove-paper.tex
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of a right edge coming from \vSimpleTimes{}
+For example, in \autoref{fig:A simple tree with a hole-b} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
@@
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimesWrapC{} leaving \vSimpleTimes{} temporarily orphaned and
+(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
 (b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.

### 2020-11-17 grove-paper.tex
fill in section 3, through Graph Decomposition

@@
 
-% XXX "p" is overloaded - could mean pattern or position
-
 \begin{figure}
@@
     \begin{array}{lrlll}
-      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \hole                                                            \\
-      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \hole \\
-      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole                               \\
+      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid * \\
+      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
+      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
+                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
+      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
     \end{array}
@@
 standard for a simply-typed lambda calculus except that all types of terms have
-holes (represented by $\hole$). In particular, the argument to a lambda can be
-either a bound variable or a hole.
+unfilled holes (represented by $\hole$), and most types of terms have conflict holes (of the form $\{\cdot | \ldots | \cdot\}$).
+In particular, the argument to a lambda can be either a bound variable or a hole.
 
@@
     \begin{array}{ll}
-      \arity : \K \rightarrow & \wp(\P)                                        \\
+      \arity : \K \hfil \rightarrow & \wp(\P)                                        \\
       \hline
-      \arity(\Root)={}        & \left\{ \Root \right\}                         \\
-      \arity(\PatVar)={}      & \left\{ \right\}                               \\
-      \arity(\ExpVar)={}      & \left\{ \right\}                               \\
-      \arity(\ExpLam)={}      & \left\{ \LamParam, \LamType, \LamBody \right\} \\
-      \arity(\ExpApp)={}      & \left\{ \AppFun, \AppArg \right\}              \\
-      \arity(\ExpPlus)={}     & \left\{ \PlusLeft, \PlusRight \right\}         \\
-      \arity(\TypArrow)={}    & \left\{ \ArrowArg, \ArrowResult \right\}       \\
-      \arity(\TypNum)={}      & \left\{ \right\}                               \\
+      \arity(k_0)={}        & \left\{ \Root \right\}                         \\
+      \arity(x \in Pat)={}      & \left\{ \right\}                               \\
+      \arity(x \in Exp)={}      & \left\{ \right\}                               \\
+      \arity(\lambda q \mathord{:}\tau.e)={}      & \left\{ \mathtt{Param}, \mathtt{Type}, \mathtt{Body} \right\} \\
+      \arity(e~e)={}      & \left\{ \mathtt{Fun}, \mathtt{Arg} \right\}              \\
+      \arity(e~op~e)={}     & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
+      \arity(n)={}      & \left\{ \right\}                               \\
+      \arity(\tau \rightarrow \tau)={}    & \left\{ \mathtt{Arg}, \mathtt{Result} \right\}       \\
+      \arity(Num)={}      & \left\{ \right\}                               \\
     \end{array}
@@
 where $\E = \U \times \V \times \P \times \V$,
-and   unique IDs are drawn from some suitable set $\U$,
-and   vertices are drawn from $\V$,
-and   positions are drawn from $\P$,
+unique IDs are drawn from some suitable set $\U$,
+vertices are drawn from $\V$,
+positions are drawn from $\P$,
 and edge states are drawn from $\Sigma$, all defined below.
@@
 A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
-where $k \in \K$ corresponds to a type of expression. Each contructor has an
-arity, which we define as a set of positions. A position $p \in \P$ identifies a
+where $k \in \K$ corresponds to one of the forms defined in \autoref{fig:Syntax
+as a grammar} that is not an operator or a hole. Each contructor has an arity,
+which we define as a set of positions. A position $p \in \P$ identifies a
 location, adjacent to some parent vertex, from which edges to child vertices may
 originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
-  Arity} gives the complete mapping from expression constructors to potential edge
+Arity} gives the complete mapping from constructors to potential edge
 positions. Every graph contains a distinguished root vertex, root constructor,
@@
 A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and a constructor $k$ for the
-outermost expression, and creating the edge $\e = (u, v_0, p_0, (u_k, k))$.
-Then, for each sub-expression at some position $p \in \arity(k)$, select unique
-id $u'$ and constructor $k'$, and create the edge $e' = (u', v, p, v')$, where
-$v = (u_k, k)$ and $v' = (u', k')$. Continuing in this manner, treat each
-sub-expression, starting with the ones at positions in $\arity(k')$, as if it
-were the outermost and create an edge to the vertices representing each of its
-immediate sub-expressions. Note that holes are not modeled explicitly in the
-graph. When an expression has a hole, it means that the graph has no edges
-originating from that position.
+empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
+expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
+sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
+constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
+= (u', k')$. Continuing in this manner, treat each sub-term, starting with the
+ones at positions in $\arity(k')$, as if it were the outermost and create edges
+to the vertices representing its immediate sub-terms until only operators and
+holes remain. Note that empty holes are not modeled explicitly in the graph.
+Wherever an empty hole term appears, the graph has no edges originating from
+that position.
 
@@
 
-% XXX deletes are idempotent
-
-We define graph edit actions as a pair of an edge state, excluding $\bot$, and an edge.
-Thus, $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$.
-As a shorthand, we define $+e$ to be $(\Plus, e)$ and $-e$ to be $(\Minus, e)$.
-
-We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
-For example,
-$\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
-and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$, which means that once an edge is deleted it can never be restored.
-
-%smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.
-
-We define the semantics of graph actions via the following
-transition relation between graphs.
+We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
+as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
+define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
+
+We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
+$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
+\sqcup \bot = \Plus$ means that edges can only be created if they do not already
+exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
+an edge is deleted it can never be restored.
+
+We define the semantics of graph actions via the following transition relation
+between graphs.
 \[
-  G \overset{\left(s,e\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right\}
+  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
 \]
-In other words, applying graph action $\alpha = \left(s, e\right)$ to a graph $G$
-results in $G\left[e \mapsto s \sqcup G(e) \right]$ (i.e., the graph is updated at $e$ to be
-the join of $s$ with whatever was already at $e$).
+In words, applying graph action $\left(s, \e\right)$ to graph $G$ results in the
+updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
+state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
+$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
+$G$.
 
@@
 
-What does a graph decomposition produce? a 4-tuple (reachable, multiparented, deleted, simple cycles).
+If we translate the graph into expressions in the calculus, interactions with
+the graph can be modeled as a stream of user actions as described in
+\autoref{sub:User Actions}. Our decomposition algorithm partitions the graph
+into disjount connected components. We encode cycles by resolving all but one
+vertex of each multi-edge into a by-id reference to the one vertex.
+
+The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
+\wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
+portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
+expressions that involve vertices with multiple parents. $D$ contains deleted
+expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
+$SC$ contains the deleted expressions that do not involve multiple parents but
+do contain cycles.
 
@@
 
+\section{Type System}
+\label{sec:Type System}
+
 Sensibility -- mapping back from graphs
 to well-typed states that fixes holes
-
-\section{Type System}
-\label{sec:Type System}
 

### 2020-11-17 grove-paper.tex
Add code labels in figure 7 (Editing Nested)
Updates from Overleaf

--- 2020-11-16 grove-paper.tex
@@
 \label{fig:A simple tree with a hole-a}
-\includegraphics{code-views/simple.png}
+TODO
 \end{subfigure}
@@
 \label{fig:Multi-child-d}
-\includegraphics{code-views/multi-child-conflict.png}
+TODO: show UI
 \end{subfigure}
@@
 \label{fig:Multi-parent-f}
-\includegraphics{code-views/multi-parent-conflict.png}
+TODO: UI
 \end{subfigure}
@@
 \label{fig:Multi-User Cycles-e}
-\includegraphics{code-views/cycle.png}
+TODO: UI
 \end{subfigure}
@@
 \label{fig:Disconnection-e}
-\includegraphics{code-views/disconnected-reachable.png}
-\includegraphics{code-views/disconnected-unreachable.png}
+TODO: UI
 \end{subfigure}
@@
 \author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
+\orcid{nnnn-nnnn-nnnn-nnnn}
 
@@
 multiple edges (i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}).
-We thus mark it with a star.)\todo{star size}\todo{"sync"}
+We thus mark it with a star.)
 
@@
 
+Like Hazel, cursor position are a position in the tree, not a particular character.
+They wrap around an expression
+
+In most cases, cursor on position.  This has the advantage of remaining stable
+if the other user deletes the code under the cursor.
+However, when there are multiple children, cursors need to be specific
+to a particular edge.
+
 Cursor is either on edge or on position
@@
 
-% XXX "p" is overloaded - could mean pattern or position
+% $\arraycolsep=2pt\begin{array}{llcl}
+% \mathsf{Typ} & \tau & ::= & t ~\vert~ \aparr{\tau}{\tau} ~\vert~ \aall{t}{\tau} ~\vert~ \arec{t}{\tau} ~\vert~ \aprod{\labelset}{\mapschema{\tau}{i}{\labelset}} ~\vert~ \asum{\labelset}{\mapschema{\tau}{i}{\labelset}}\\
+% \mathsf{Exp} & e & ::= & x ~\vert~ \aelam{\tau}{x}{e} ~\vert~ \aeap{e}{e} ~\vert~ \aetlam{t}{e} ~\vert~ \aetap{e}{\tau} ~\vert~ \aefold{e} ~\vert~ \aetpl{\labelset}{\mapschema{e}{i}{\labelset}} ~\vert~  \aein{\ell}{e} \\
+% & & \vert & \aematchwith{n}{e}{\seqschemaX{r}}\\
+% \mathsf{Rule} & r & ::= & \aematchrule{p}{e}\\
+% \mathsf{Pat} & p & ::= & x  ~\vert~ \aewildp ~\vert~ \aefoldp{p} ~\vert~ \aetplp{\labelset}{\mapschema{p}{i}{\labelset}} ~\vert~ \aeinjp{\ell}{p}
+% \end{array}$
 
@@
 \begin{array}{lrlll}
-      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \hole                                                            \\
-      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \hole \\
-      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole                               \\
+ \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \_ \\
+ \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
+ \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \_ \\
 \end{array}
@@
 
-The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
-standard for a simply-typed lambda calculus except that all types of terms have
-holes (represented by $\hole$). In particular, the argument to a lambda can be
-either a bound variable or a hole.
-
-% In order to model this we define
-
-% Turn that into a constructor set K + position set P + arity function (has type: K -> $\wp$(P)).
+The language we are modeling is shown in \autoref{fig:Syntax as a grammar}.
+Standard for a simply-typed lambda calculus except that
+all terms have holes (represented by $\_$)
+and the argument to a lambda is its own term.
+The latter of which allows us to express lambdas with a hole for the bound variable.
+
+In order to model this we define
+
+Turn that into a constructor set K + index set I + arity function (has type: K -> $\wp$(I)).
 
@@
 \begin{array}{ll}
-      \arity : \K \rightarrow & \wp(\P)                                        \\
+ \arity : \K \rightarrow & \wp(\I) \\
  \hline
@@
 
+TODO: rename indices to positions
+
+TODO: look into using bold plus and minus (google results)
+
 A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states, 
@@
 
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
-where $k \in \K$ corresponds to a type of expression. Each contructor has an
-arity, which we define as a set of positions. A position $p \in \P$ identifies a
-location, adjacent to some parent vertex, from which edges to child vertices may
-originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
-  Arity} gives the complete mapping from expression constructors to potential edge
-positions. Every graph contains a distinguished root vertex, root constructor,
-and root position, respectively named $v_0$, $k_0$, and $p_0$.
-
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and a constructor $k$ for the
-outermost expression, and creating the edge $\e = (u, v_0, p_0, (u_k, k))$.
-Then, for each sub-expression at some position $p \in \arity(k)$, select unique
-id $u'$ and constructor $k'$, and create the edge $e' = (u', v, p, v')$, where
-$v = (u_k, k)$ and $v' = (u', k')$. Continuing in this manner, treat each
-sub-expression, starting with the ones at positions in $\arity(k')$, as if it
-were the outermost and create an edge to the vertices representing each of its
-immediate sub-expressions. Note that holes are not modeled explicitly in the
-graph. When an expression has a hole, it means that the graph has no edges
-originating from that position.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
+A vertex $v = (u, k)$ identifies a distinct instance of the constructor $k$.
+
+A constructor $k \in \K$ represents a different type of node in the target language's abstract syntax tree.
+
+A position $p \in \P$ identifies a location, adjacent to some parent vertex.
+
+Every graph contains a distinguished root vertex, root constructor, and root position, respectively named $v_0$, $k_0$, and $p_0$.
+
+% Define $\sqcup_{edge state}$ (i.e. join)
+
+% lattice ordering $\bot < + < -$ LUB e.g., join A B = C
+
+You can map from syntax to graph by selecting
+unique IDs $u$ (for both edges and vertexes).
+    - holes are not explicit in the graph
 
@@
 We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
-For example,
-$\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
-and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$, which means that once an edge is deleted it can never be restored.
-
+For example, $\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
+and $\Plus \sqcup \Minus = \Minus$, which means that once an edge is deleted it can never be restored.
 %smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.
@@
 \begin{align*}
-  \edges(G)            & = \dom(G)                                       \\
-  \liveEdges(G)        & = \{ \e \in \edges(G) | G(\e)=+ \}              \\
-  \vertexChildren(G,v) & = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
-  \cursorChildren(G,c) & = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
-  \parents(G,v)        & = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
+             \edges(G) &= \dom(G) \\
+         \liveEdges(G) &= \{ \e \in \edges(G) | G(\e)=+ \}              \\
+  \vertexChildren(G,v) &= \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
+  \cursorChildren(G,c) &= \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
+         \parents(G,v) &= \{ (c,v',u) \in \liveEdges(G) | v'=v \}
 \end{align*}
@@
 \begin{align*}
-  \vertexes(G)        & = \{ v \in \V | \exists (c,v,u) \in \edges(G) \}              \\
-  \liveVertexes(G)    & = \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}          \\
-  \childVertexes(G,v) & = \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \} \\
-  \orphans(G)         & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}      \\
-  \multiparents(G)    & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}      \\
-  \vertex(G,u)        & = v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
+        \vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
+    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
+ \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
+         \orphans(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}       \\
+    \multiparents(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}       \\
+        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
 \end{align*}

### 2020-11-17 grove-paper.tex

--- 2020-11-17 grove-paper.tex
@@
 \label{fig:Editing Nested Parts of the Code-a}
+\textbf{\texttt{\hole{} + w * v}}\vskip1em%
 \begin{tikzpicture}
@@
 \label{fig:Editing Nested Parts of the Code-b}
+\textbf{\texttt{u * v + \hole{}}}\vskip1em%
 \begin{tikzpicture}
@@
 \label{fig:Editing Nested-c}
+\textbf{\texttt{w * v + \hole{}}}\vskip1em%
 \begin{tikzpicture}
@@
 \autoref{fig:Editing Different Parts of the Code-b} represents
-multiple edges, i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
-We thus mark it with a star.)\todo{star size}\todo{"sync"}
+multiple edges (i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}).
+We thus mark it with a star.)
 
@@
 \label{sub:Cursors}
+
+Like Hazel, cursor position are a position in the tree, not a particular character.
+They wrap around an expression
+
+In most cases, cursor on position.  This has the advantage of remaining stable
+if the other user deletes the code under the cursor.
+However, when there are multiple children, cursors need to be specific
+to a particular edge.
 

### 2020-11-17 grove-paper.tex
Add a very rough draft of the cursor section

--- 2020-11-17 grove-paper.tex
@@
 \newcommand{\Vertex}[1]{Vertex~#1}
+\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
 
@@
 \label{fig:Editing Nested Parts of the Code-a}
+\figureCode{\hole{} + w * v}%
 \begin{tikzpicture}
@@
 \label{fig:Editing Nested Parts of the Code-b}
+\figureCode{u * v + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Editing Nested-c}
+\figureCode{w * v + \hole{}}%
 \begin{tikzpicture}
@@
 
-Cursor is either on edge or on position
-
-cursors are advisory
+Finally we address the question of cursors.
+
+All edits use edge and vertex identifiers.
+Thus cursor positions do not affect the interpretation of graph edits.
+Cursors are not considered part of the graph representation,
+and need not be communicated to other users.
+
+COMMUNICATION
+
+TODO: use Alice and Bob
+
+However as a convienience to other users, the current cursor position
+may be transmitted to other users.
+In this role, cursors are merely advisory and we can follow the simple
+protocol of the sender marking every cursor announcement with a timestamp
+and the receiver displaying the cursor position With the most recently received position being the one displayed.
+Note that this cursor position may refer to edges or verticies that
+do not yet exist on the receivers machine.
+In these cases, we can either not display the other user's cursor (and perhaps
+have a visual display flagging this fact), or we can display the most recent
+cursor that represents a valid position in the local graph (perhaps
+shown in a fadded color to show that this cursor is known to not be up to date).
+(Or perhaps, all cursors always fade/decay over time like on radar blips.)
 
 Don't model communication but would be separate layer anyway
+
+REPRESENTATION
+
+For most editing a cursor is represented by a position within a vertex.
+For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
+Representing it this way means that if Bob or some other user deletes TODO:vertex
+and replaces it with a different vertex, Bob's cursor doesn't move.
+We specify a position within a vertex instead of a vertex because TODO.
+
+However, there is one case where we need a more precise cursor.
+That is when a vertex has multiple children in a particular position
+and we want the cursor to be on a specific child.
+In that case, we represent the cursor by an edge identifier.
+An example of this is seen in REF:figure.
+If Bob's cursor is on the \texttt{\{ x | y \}} as a whole,
+Bob's cursor would be the pair of the vertex id TODO and the position~\texttt{R}.
+If Bob's cursor is on a particular element in the multi-child,
+say the~\texttt{x}, then we represent the cursor position with edge identifier TODO:edge.
 

### 2020-11-17 grove-paper.tex
Rough cut of code for each figure

@@
 \usepackage{todonotes}
+
+% Load MnSymbol without clobbering \ast
+% See https://tex.stackexchange.com/a/269691
+\usepackage{amsmath}% needed before mathabx
+\let\amsast=\ast
+\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
+\let\abxast=\ast
+\usepackage{MnSymbol}
+\let\mnast=\ast
+\let\ast=\abxast
 
@@
 \tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{*}$}}}}}
+\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
 
@@
 \newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
+\newcommand{\multiChild}[2]{\textcolor{red}{\{} #1 \textcolor{red}{|} #2 \textcolor{red}{\}}}
+\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
+\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\rcirclearrowleft_{#1}}}}}
+\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
+\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
 
@@
 \label{fig:A simple tree with a hole-b}
+\figureCode{x * \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Construction-b}
+\figureCode{x * y}%
 \begin{tikzpicture}
@@
 \label{fig:Change}
+\figureCode{\hole{} * y}%
 \begin{tikzpicture}
@@
 \label{fig:Wrapping-a}
+\figureCode{\hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Wrapping-b}
+\figureCode{\hole{} + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Wrapping-c}
+\figureCode{\hole{} * y + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Move-a}
+\figureCode{\hole{} + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Move-b}
+\figureCode{\hole{} + \hole{} * y}%
 \begin{tikzpicture}
@@
 \label{fig:Editing Different Parts of the Code-a}
+\figureCode{\hole{} + u * y}%
 \begin{tikzpicture}
@@
 \label{fig:Editing Different Parts of the Code-b}
+\figureCode{\hole{} + \hole{} * v}%
 \begin{tikzpicture}
@@
 \label{fig:Editing Different Parts of the Code-c}
+\figureCode{\hole{} + u * v}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-child-a}
+\figureCode{w * v + x}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-child-b}
+\figureCode{w * v + y}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-child-c}
+\figureCode{w * v + \multiChild{x}{y}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-parent-a}
+\figureCode{w * \hole{} + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-parent-b}
+\figureCode{\hole{} * w + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-parent-c}
+\figureCode{\hole{} * \hole{} + w}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-parent-d}
+\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-parent-e}
+\figureCode{\hole{} * w + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-User Cycles-a}
+\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-User Cycles-b}
+\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-User Cycles-c}
+\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}}%
 \begin{tikzpicture}
@@
 \label{fig:Multi-User Cycles-d}
+\figureCode{\parens{\hole{} * \hole{}} + \multiChild{\multiVertex{30}}{\multiVertex{32}}}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnection-a}
+\figureCode{\hole{} * \hole{} + \hole{} * \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnection-b}
+\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnection-c}
+\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnection-d}
+\figureCode{\hole{} + \hole{}}%
 \begin{tikzpicture}

### 2020-11-18 grove-paper.tex
Combine figures 1 and 2

@@
 \newvertex{SimpleX}{x}%
+\newedge{ConstructionY}{R}%
+\newvertex{ConstructionY}{y}%
 \newcommand{\figureSimple}{
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
-\label{fig:A simple tree with a hole-a}
-\includegraphics{code-views/simple.png}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.49\linewidth}
-\centering
-\caption{}
+% \begin{subfigure}[t]{0.49\linewidth}
+% \centering
+% \caption{}
+% \label{fig:A simple tree with a hole-a}
+% \includegraphics{code-views/simple.png}
+% \end{subfigure}
+% %%%%%%%%
+% \hfill
+% %%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Simple-a) [anchor=base] {\strut};
+%\path [draw,->,alice step] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
+\end{tikzpicture}
+}
 \label{fig:A simple tree with a hole-b}
@@
 %%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
+\path [draw,->,alice step] (fig-Simple-a) to (-2em, 0cm |- fig-Construction);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-a}
+\label{fig:Construction}
+\label{fig:Construction-b}
+\figureCode{x * y}%
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+  root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+    "\vSimpleX" [> "\eSimpleX"'],
+    "\vConstructionY" [> "\eConstructionY", >alice edge, alice node]
+  }
+  }
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%%%%%
+\hfill
+%%%%%%%%
+\begin{subfigure}[t]{0.32\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
+\path [draw,->,alice step] (fig-Construction) to (-2em, 0cm |- fig-Deletion);
+\end{tikzpicture}
+}
+\label{fig:Construction and Deletion-b}
+\label{fig:Change-a}
+\label{fig:Change}
+\figureCode{\hole{} * y}%
+\begin{tikzpicture}
+\path (0cm,0cm) graph[graph style] {
+  root[root style] -> {
+  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
+    {},
+    "\vConstructionY" [> "\eConstructionY"]
+  }
+  }
+};
+\path (-0.25cm,-1.7cm) graph[graph style] {
+  "\vSimpleX"
+};
+\end{tikzpicture}
+\end{subfigure}
+%%%%  
 \hfill{}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{ConstructionY}{R}%
-\newvertex{ConstructionY}{y}%
-\newcommand{\figureConstructionAndDeletion}{
-\begin{figure}
-%\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-A simple tree with a hole-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:A simple tree with a hole-b}}}};
-%%%%
-\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
-\end{tikzpicture}
-}
-\label{fig:Construction and Deletion-a}
-\label{fig:Construction}
-\label{fig:Construction-b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   "\vSimpleX" [> "\eSimpleX"'],
-   "\vConstructionY" [> "\eConstructionY", >alice edge, alice node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Construction) to (-2em, 0cm |- fig-Deletion);
-\end{tikzpicture}
-}
-\label{fig:Construction and Deletion-b}
-\label{fig:Change-a}
-\label{fig:Change}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-   {},
-   "\vConstructionY" [> "\eConstructionY"]
-  }
- }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
- "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Construction and Deletion.}
-\end{figure}
-}
+% \newcommand{\figureConstructionAndDeletion}{
+% \begin{figure}
+% %\hfill
+% \hskip0.12\linewidth
+% %%%%%%%%
+% %%%%%%%%
+% \hfill{}
+% \caption{Example of Construction and Deletion.}
+% \end{figure}
+% }
 
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.28\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-A simple tree with a hole-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Construction and Deletion-b}}}};
+%%%%
 \path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Deletion) to (-2em, 0cm |- fig-Wrapping-a);
+\path [draw,->,alice step] (fig-A simple tree with a hole-c-x) to (-2em, 0cm |- fig-Wrapping-a);
 \end{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
+\begin{subfigure}[t]{0.28\linewidth}
 \centering
@@
 \hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to our narrative.)}
+\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}
 \label{fig:Wrapping}
@@
 \begin{figure}
-%\hfill
-\hskip0.12\linewidth
+\hfill
+%\hskip0.12\linewidth
 %%%%%%%%
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
+%\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
 %%%%
 \path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Wrapping-c-x) to (-2em, 0cm |- fig-Moving-a);
+\path [draw,->,alice step] (fig-Wrapping-c) to (-2em, 0cm |- fig-Moving-a);
 \end{tikzpicture}
@@
 \begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
+%\hfill
+\hskip0.12\linewidth
+%%%%%%%%
+\begin{subfigure}[t]{0.28\linewidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-7.5em,0cm) node (fig-Moving-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested Parts of the Code-b}}}};
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Moving-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\path [draw,->,alice step] (fig-Moving-b-x) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Moving-b) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\path [draw,->,bob step] (fig-Moving-b-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
+%\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
 %%%%
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c-x) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
 \end{tikzpicture}
@@
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
 \end{tikzpicture}
@@
 
+\figureWrapping
+
 \subsection{Single-User Actions}
@@
 
-\figureConstructionAndDeletion
+%\figureConstructionAndDeletion
 
@@
 \label{sub:Deletion}
+
+\figureMovement
 
@@
 \label{sub:Wrapping}
-
-\figureWrapping
 
@@
 
-\figureMovement
+\figureEditingDifferentParts
 
@@
 
-\figureEditingDifferentParts
+\figureEditingNestedParts
 
@@
 \label{sub:Editing Nested Parts of the Code}
-
-\figureEditingNestedParts
 

### 2020-11-18 grove-paper.tex
Regularize figure labels

@@
 \newvertex{SimpleX}{x}%
-\newedge{ConstructionY}{R}%
-\newvertex{ConstructionY}{y}%
+\newedge{SimpleY}{R}%
+\newvertex{SimpleY}{y}%
 \newcommand{\figureSimple}{
@@
 %%%%%%%%
-% \begin{subfigure}[t]{0.49\linewidth}
-% \centering
-% \caption{}
-% \label{fig:A simple tree with a hole-a}
-% \includegraphics{code-views/simple.png}
-% \end{subfigure}
-% %%%%%%%%
-% \hfill
-% %%%%%%%%
 \begin{subfigure}[t]{0.32\linewidth}
@@
 \path (0cm,0cm) node (fig-Simple-a) [anchor=base] {\strut};
-%\path [draw,->,alice step] (fig-A simple tree with a hole-b-x) to (-2em, 0cm |- fig-Construction);
-\end{tikzpicture}
-}
-\label{fig:A simple tree with a hole-b}
+\end{tikzpicture}
+}
+\label{fig:Simple:a}
 \figureCode{x * \hole{}}%
@@
 }
-\label{fig:Construction and Deletion-a}
-\label{fig:Construction}
-\label{fig:Construction-b}
+\label{fig:Simple:b}
 \figureCode{x * y}%
@@
     "\vSimpleX" [> "\eSimpleX"'],
-    "\vConstructionY" [> "\eConstructionY", >alice edge, alice node]
+    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
   }
@@
 }
-\label{fig:Construction and Deletion-b}
-\label{fig:Change-a}
-\label{fig:Change}
+\label{fig:Simple:c}
 \figureCode{\hole{} * y}%
@@
     {},
-    "\vConstructionY" [> "\eConstructionY"]
+    "\vSimpleY" [> "\eSimpleY"]
   }
@@
 \end{subfigure}
-%%%%  
+%%%%%%%%
 \hfill{}
-\caption{Example of Code that Includes a Hole.
-\autoref{fig:A simple tree with a hole-a}:~User Interface for the Code.
-\autoref{fig:A simple tree with a hole-b}:~Graph Representation of the Code.}
-\label{fig:A simple tree with a hole}
+\caption{Code that Includes a Hole and Some Simple Edits.}
+\label{fig:Simple}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureConstructionAndDeletion
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureConstructionAndDeletion}{
-% \begin{figure}
-% %\hfill
-% \hskip0.12\linewidth
-% %%%%%%%%
-% %%%%%%%%
-% \hfill{}
-% \caption{Example of Construction and Deletion.}
-% \end{figure}
-% }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrapping
+%% \figureWrap
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newedge{WrapTimes}{L}
-\newcommand{\figureWrapping}{
+\newcommand{\figureWrap}{
 \begin{figure}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-A simple tree with a hole-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Construction and Deletion-b}}}};
+\path (-7.5em,0cm) node (fig-A simple tree with a hole-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
 %%%%
@@
 }
-\label{fig:Wrapping-a}
+\label{fig:Wrap:a}
 \figureCode{\hole{}}%
@@
   {},
-  "\vConstructionY" [> "\eConstructionY"]
+  "\vSimpleY" [> "\eSimpleY"]
  }
@@
 }
-\label{fig:Wrapping-b}
+\label{fig:Wrap:b}
 \figureCode{\hole{} + \hole{}}%
@@
 }
-\label{fig:Wrapping-c}
+\label{fig:Wrap:c}
 \figureCode{\hole{} * y + \hole{}}%
@@
     {},
-    "\vConstructionY" [> "\eConstructionY"]
+    "\vSimpleY" [> "\eSimpleY"]
    },
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMovement
+%% \figureMove
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newedge{MoveTimes}{R}
-\newcommand{\figureMovement}{
+\newcommand{\figureMove}{
 \begin{figure}
@@
 \begin{tikzpicture}[remember picture, overlay]
-%\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrapping-c}}}};
+%\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrap:c}}}};
 %%%%
@@
 }
-\label{fig:Move-a}
+\label{fig:Move:a}
 \figureCode{\hole{} + \hole{}}%
@@
 }
-\label{fig:Move-b}
+\label{fig:Move:b}
 \figureCode{\hole{} + \hole{} * y}%
@@
     {},
-    "\vConstructionY" [> "\eConstructionY"]
+    "\vSimpleY" [> "\eSimpleY"]
    }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureEditingDifferentParts
+%% \figureDifferentParts
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newvertex{DifferentBob}{v}
-\newcommand{\figureEditingDifferentParts}{
+\newcommand{\figureDifferentParts}{
 \begin{figure}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Moving-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested Parts of the Code-b}}}};
+\path (-7.5em,0cm) node (fig-Moving-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:b}}}};
 \path (0cm,0cm) node (fig-Editing Different Parts of the Code-a) [anchor=base] {\strut};
@@
 }
-\label{fig:Editing Different Parts of the Code-a}
+\label{fig:DifferentParts:a}
 \figureCode{\hole{} + u * y}%
@@
     "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice node],
-    "\vConstructionY" [> "\eConstructionY"]
+    "\vSimpleY" [> "\eSimpleY"]
   }
@@
 }
-\label{fig:Editing Different Parts of the Code-b}
+\label{fig:DifferentParts:b}
 \figureCode{\hole{} + \hole{} * v}%
@@
 }
-\label{fig:Editing Different Parts of the Code-c}
+\label{fig:DifferentParts:c}
 \figureCode{\hole{} + u * v}%
@@
 \caption{Example of Users Editing Different Parts of the Code.}
-\label{fig:Editing Different Parts of the Code}
+\label{fig:DifferentParts}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureEditingNestedParts
+%% \figureNestedParts
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newedge{NestedBob}{L}
-\newcommand{\figureEditingNestedParts}{
+\newcommand{\figureNestedParts}{
 \begin{figure}
@@
 \begin{tikzpicture}[remember picture, overlay]
-%\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
-%%%%
 \path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
@@
 }
-\label{fig:Editing Nested Parts of the Code-a}
+\label{fig:NestedParts:a}
 \figureCode{\hole{} + w * v}%
@@
 }
-\label{fig:Editing Nested Parts of the Code-b}
+\label{fig:NestedParts:b}
 \figureCode{u * v + \hole{}}%
@@
 }
-\label{fig:Editing Nested Parts of the Code-c}
-\label{fig:Editing Nested-c}
+\label{fig:NestedParts:c}
 \figureCode{w * v + \hole{}}%
@@
 \caption{Example of Users Editing Nested Parts of the Code.}
-\label{fig:Editing Nested Parts of the Code}
+\label{fig:NestedParts}
 \end{figure}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
+\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
 %%%%
@@
 }
-\label{fig:Multi-child-a}
+\label{fig:MultiChild:a}
 \figureCode{w * v + x}%
@@
 }
-\label{fig:Multi-child-b}
+\label{fig:MultiChild:b}
 \figureCode{w * v + y}%
@@
 }
-\label{fig:Multi-child-c}
+\label{fig:MultiChild:c}
 \figureCode{w * v + \multiChild{x}{y}}%
@@
 \caption{}
-\label{fig:Multi-child-d}
+\label{fig:MultiChild:d}
 \includegraphics{code-views/multi-child-conflict.png}
@@
 \caption{Example of Multi-Child Conflicts.}
-\label{fig:Multi-child}
+\label{fig:MultiChild}
 \end{figure*}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
+\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
 %%%%
@@
 }
-\label{fig:Multi-parent-a}
+\label{fig:MultiParent:a}
 \figureCode{w * \hole{} + \hole{}}%
@@
 }
-\label{fig:Multi-parent-b}
+\label{fig:MultiParent:b}
 \figureCode{\hole{} * w + \hole{}}%
@@
 }
-\label{fig:Multi-parent-c}
+\label{fig:MultiParent:c}
 \figureCode{\hole{} * \hole{} + w}%
@@
 }
-\label{fig:Multi-parent-d}
+\label{fig:MultiParent:d}
 \figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}}%
@@
 }
-\label{fig:Multi-parent-e}
+\label{fig:MultiParent:e}
 \figureCode{\hole{} * w + \hole{}}%
@@
 \caption{}
-\label{fig:Multi-parent-f}
+\label{fig:MultiParent:f}
 \includegraphics{code-views/multi-parent-conflict.png}
@@
 \caption{Example of Multi-Parent Conflicts.}
-\label{fig:Multi-parent}
+\label{fig:MultiParent}
 \end{figure*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSingleUserCycles
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newedge{SingleCyclePlus}{R}
-% \newvertex{SingleCyclePlus}{+}
-% \newedge{SingleCycleTimes}{L}
-% \newvertex{SingleCycleTimes}{*}
-% \newedge{SingleCycleCycle}{R}
-% \newcommand{\figureSingleUserCycles}{
-% \begin{figure}
-% %\hfill
-% \hskip0.12\linewidth
-% %%%%%%%%
-% \begin{subfigure}[t]{0.28\linewidth}
-% \centering
-% \caption{
-% \begin{tikzpicture}[remember picture, overlay]
-% \path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
-% %%%%
-% \path (0cm,0cm) node (fig-Single-User Cycles-a) [anchor=base] {\strut};
-% \path [draw,->,alice step] (fig-Multi-parent-e-x) to (-2em, 0cm |- fig-Single-User Cycles-a);
-% \end{tikzpicture}
-% }
-% \label{fig:Single-User Cycles-a}
-% \begin{tikzpicture}
-% \path (0cm,0cm) graph[graph style] {
-%  root[root style] -> {
-%   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-%   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
-%     {},
-%     {}
-%   },
-%   {}
-%   }
-%  }
-% };
-% \end{tikzpicture}
-% \end{subfigure}
-% %%%%%%%%
-% \hfill
-% %%%%%%%%
-% \begin{subfigure}[t]{0.28\linewidth}
-% \centering
-% \caption{
-% \begin{tikzpicture}[remember picture, overlay]
-% \path (0cm,0cm) node (fig-Single-User Cycles-b) [anchor=base] {\strut};
-% \path [draw,->,alice step] (fig-Single-User Cycles-a) to (-2em, 0cm |- fig-Single-User Cycles-b);
-% \end{tikzpicture}
-% }
-% \label{fig:Single-User Cycles-b}
-% \begin{tikzpicture}
-% \path (0cm,0cm) graph[graph style] {
-%  root[root style] -> {
-%   v8/"\vWrapPlus" [> "\eWrapPlus"'] -> {
-%   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
-%     {},
-%     "\vSingleCyclePlus" [> "\eSingleCyclePlus", >alice edge, alice node] -> {
-%      "\vSingleCycleTimes" [> "\eSingleCycleTimes"', >alice edge, alice node] -> {
-%       {},
-%       {}
-%      },
-%      {}
-%     }
-%   },
-%   }
-%  }
-% };
-% \end{tikzpicture}
-% \end{subfigure}
-% %%%%%%%%
-% \hfill
-% %%%%%%%%
-% \begin{subfigure}[t]{0.28\linewidth}
-% \centering
-% \caption{
-% \begin{tikzpicture}[remember picture, overlay]
-% \path (0cm,0cm) node (fig-Single-User Cycles-c) [anchor=base] {\strut};
-% \path [draw,->,alice step] (fig-Single-User Cycles-a) to[out=15,in=165] (-2em, 0cm |- fig-Single-User Cycles-c);
-% \end{tikzpicture}
-% }
-% \label{fig:Single-User Cycles-c}
-% \begin{tikzpicture}
-% \path (0cm,0cm) graph[graph style] {
-%  root[root style] -> {
-%   v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-%   v2/"\vSimpleTimes" [> "\eNestedBob"'],
-%   {}
-%   }
-%  }
-% };
-% \path[draw,->,alice edge] (v2) to ["\eSingleCycleCycle",out=-60,in=-60] (v8);
-% \end{tikzpicture}
-% \end{subfigure}
-% %%%%%%%%
-% \hfill{}
-% \caption{Single-User Cycles.}
-% \label{fig:Single-User Cycles}
-% \end{figure}
-% }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiUserCycles
+%% \figureCycle
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureMultiUserCycles}{
+\newcommand{\figureCycle}{
 \begin{figure*}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-parent-e}}}};
+\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
 %%%%
@@
 }
-\label{fig:Multi-User Cycles-a}
+\label{fig:Cycle:a}
 \figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}}%
@@
 }
-\label{fig:Multi-User Cycles-b}
+\label{fig:Cycle:b}
 \figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}}%
@@
 }
-\label{fig:Multi-User Cycles-c}
+\label{fig:Cycle:c}
 \figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}}%
@@
 }
-\label{fig:Multi-User Cycles-d}
+\label{fig:Cycle:d}
 \figureCode{\parens{\hole{} * \hole{}} + \multiChild{\multiVertex{30}}{\multiVertex{32}}}%
@@
 \caption{}
-\label{fig:Multi-User Cycles-e}
+\label{fig:Cycle:e}
 \includegraphics{code-views/cycle.png}
@@
 \caption{Example of Cycles.}
-\label{fig:Multi-User Cycles}
+\label{fig:Cycle}
 \end{figure*}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
+\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
 %%%%
@@
 }
-\label{fig:Disconnection-a}
+\label{fig:Disconnect:a}
 \figureCode{\hole{} * \hole{} + \hole{} * \hole{}}%
@@
 }
-\label{fig:Disconnection-b}
+\label{fig:Disconnect:b}
 \figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}}%
@@
 }
-\label{fig:Disconnection-c}
+\label{fig:Disconnect:c}
 \figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}}%
@@
 }
-\label{fig:Disconnection-d}
+\label{fig:Disconnect:d}
 \figureCode{\hole{} + \hole{}}%
@@
 \caption{}
-\label{fig:Disconnection-e}
+\label{fig:Disconnect:e}
 \includegraphics{code-views/disconnected-reachable.png}
@@
 This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:A simple tree with a hole-b} shows one such graph,
-and \autoref{fig:A simple tree with a hole-a} shows its corresponding expression,~\texttt{x * \hole}, which
+For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
 has a hole on the right side of the times operator.
@@
 Program state is a directed graph with a distinguished root vertex.
-%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
 Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:A simple tree with a hole-b}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
+In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
 (For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
@@
 
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:A simple tree with a hole-b}).
+Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
+a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
 An edge indicates that the destination vertex is a child of the origin vertex at the given position.
@@
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:A simple tree with a hole-b} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
+For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
 corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
 
-\figureWrapping
+\figureWrap
 
@@
 
-%\figureConstructionAndDeletion
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole-b}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
-This action corresponds to creating \eConstructionY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.
+To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
+and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
+This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
+created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
+The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
 
@@
 
-\figureMovement
+\figureMove
 
 Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Change-a}.
+This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
 Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
@@
 
-\figureEditingDifferentParts
+\figureDifferentParts
 
@@
 actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move-b}.
+We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
 
@@
 
-To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
+To start, Alice and Bob both have the graph in \autoref{fig:Move:b}, which
 represents the code~\texttt{\hole{} + (\hole{} * y)}.
@@
 Alice adds~\vDifferentAlice{} as the left child of \vSimpleTimes{}
-and Bob changes \vConstructionY{} to \vDifferentBob{}.
+and Bob changes \vSimpleY{} to \vDifferentBob{}.
 Before transmitting their edits to each other,
-Alice and Bob have the graphs in \autoref{fig:Editing Different Parts of the Code-a}
-and \autoref{fig:Editing Different Parts of the Code-b}, respectively.
-(Note that the transition edge from \autoref{fig:Move-b} to
-\autoref{fig:Editing Different Parts of the Code-b} represents
-multiple edges, i.e., deleting~\eConstructionY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
+Alice and Bob have the graphs in \autoref{fig:DifferentParts:a}
+and \autoref{fig:DifferentParts:b}, respectively.
+(Note that the transition edge from \autoref{fig:Move:b} to
+\autoref{fig:DifferentParts:b} represents
+multiple edges, i.e., deleting~\eSimpleY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
 We thus mark it with a star.)\todo{star size}\todo{"sync"}
@@
 This results in both Alice and Bob having the graph in
-\autoref{fig:Editing Different Parts of the Code-c}.
-
-\figureEditingNestedParts
+\autoref{fig:DifferentParts:c}.
+
+\figureNestedParts
 
@@
 For example, suppose Alice and Bob did not transmit their edits to each other
-but to a third user, Chris, that has the original graph in \autoref{fig:Move-b}.
+but to a third user, Chris, that has the original graph in \autoref{fig:Move:b}.
 Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:Editing Different Parts of the Code-c}.
+the graph in \autoref{fig:DifferentParts:c}.
 
@@
 After Alice and Bob share the edits they made in the previous subsection,
-Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
+Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
 On the other hand, Bob moves~\vSimpleTimes{} and its children from the right child of~\vWrapPlus{} to its left child
-by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
+by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
 Since edits are based on edge identities,
 both Alice's and Bob's edits can be applied to the other's graph,
-which results in both Alice and Bob having the graph in \autoref{fig:Editing Nested Parts of the Code-c}.
+which results in both Alice and Bob having the graph in \autoref{fig:NestedParts:c}.
 
@@
 while Bob adds~\vMultiChildBob{} also as the right child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
+These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
 These graphs have two different edges for the~\texttt{R} position of~\vWrapPlus{}.
-\autoref{fig:Multi-child-a} has~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
-and \autoref{fig:Multi-child-b} has~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
+\autoref{fig:MultiChild:a} has~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
+and \autoref{fig:MultiChild:b} has~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
 We call this a \emph{multi-child conflict}.
@@
 of \vWrapPlus{}.
-This results in \autoref{fig:Multi-child-c}.
+This results in \autoref{fig:MultiChild:c}.
 The conflict between Alice's and Bob's changes is
@@
 In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}
-as shown in \autoref{fig:Multi-child-d}.
+as shown in \autoref{fig:MultiChild:d}.
 
@@
 
-An example of this is shown in \autoref{fig:Multi-parent}.
-\autoref{fig:Multi-parent-a} shows the graph after Alice and Bob have made some more edits,
+An example of this is shown in \autoref{fig:MultiParent}.
+\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some more edits,
 and Alice and Bob's editors are synchronized with each other.
@@
 while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:Multi-parent-b} and \autoref{fig:Multi-parent-c},
+This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c},
 which represent respectively the expression~\texttt{(\hole{} * w) + \hole{}} for Alice
@@
 However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-This generates \autoref{fig:Multi-parent-d}, in which~\vNestedAlice{} has two parents pointing to it.
+This generates \autoref{fig:MultiParent:d}, in which~\vNestedAlice{} has two parents pointing to it.
 We call this situation a multi-parent conflict.
@@
 % However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles-a}
+% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
 % and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
@@
 % The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles-b} and
+% This results in \autoref{fig:Single-User Cycles:b} and
 % does not cause a cycle.
 % The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles-b}
+% This results in \autoref{fig:Single-User Cycles:b}
 % and causes a cycle.
@@
 
-\figureMultiUserCycles
+\figureCycle
 
@@
 
-For example, consider the situation in \autoref{fig:Multi-User Cycles}.
-\autoref{fig:Multi-User Cycles-a} shows the graph after Alice and Bob have made some more edits,
+For example, consider the situation in \autoref{fig:Cycle}.
+\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some more edits,
 and Alice and Bob's editors are synchronized with each other.
@@
 and then~\vMultiCyclePlus{} underneath.
-This results in \autoref{fig:Multi-User Cycles-b},
+This results in \autoref{fig:Cycle:b},
 which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
@@
 but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Multi-User Cycles-c},
+This results in \autoref{fig:Cycle:c},
 which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
@@
 However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Multi-User Cycles-d}, which has a cycle
+in \autoref{fig:Cycle:d}, which has a cycle
 between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
@@
 to display these kinds of programs.
-For example, \autoref{fig:Multi-User Cycles-d} would be displayed
-as shown in \autoref{fig:Multi-User Cycles-e}.
+For example, \autoref{fig:Cycle:d} would be displayed
+as shown in \autoref{fig:Cycle:e}.
 We leave the user interface considerations of resolving the conflict such a
@@
 For example, consider the situation in \autoref{fig:Disconnection}.
-\autoref{fig:Disconnection-a} shows the graph after Alice and Bob have made some more edits,
+\autoref{fig:Disconnect:a} shows the graph after Alice and Bob have made some more edits,
 and Alice and Bob's editors are synchronized with each other.
@@
 Alice moves~\vMultiCycleTimes{} under~\vSimpleTimes{}.
-This results in \autoref{fig:Disconnection-b},
+This results in \autoref{fig:Disconnect:b},
 which represents the expression~\texttt{(\hole{} * (\hole{} * \hole{})) + \hole}.
 On the other hand, Bob puts~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnection-c},
+This results in \autoref{fig:Disconnect:c},
 which represents the expression~\texttt{((\hole{} * \hole{}) * \hole{}) + \hole{}}.
 
-When these edits are merged, the result is the graph in \autoref{fig:Disconnection-d}.
+When these edits are merged, the result is the graph in \autoref{fig:Disconnect:d}.
 Since Alice deleted~\eNestedBob{}

### 2020-11-18 grove-paper.tex
Simplify node naming

@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Simple-a) [anchor=base] {\strut};
+\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Construction) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Simple-a) to (-2em, 0cm |- fig-Construction);
+\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Deletion) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Construction) to (-2em, 0cm |- fig-Deletion);
+\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-A simple tree with a hole-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
+\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
 %%%%
-\path (0cm,0cm) node (fig-Wrapping-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-A simple tree with a hole-c-x) to (-2em, 0cm |- fig-Wrapping-a);
+\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Wrapping-b) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Wrapping-a) to (-2em, 0cm |- fig-Wrapping-b);
+\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Wrapping-c) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Wrapping-b) to (-2em, 0cm |- fig-Wrapping-c);
+\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-%\path (-7.5em,0cm) node (fig-Wrapping-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Wrap:c}}}};
-%%%%
-\path (0cm,0cm) node (fig-Moving-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Wrapping-c) to (-2em, 0cm |- fig-Moving-a);
+\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Moving-b) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Moving-a) to (-2em, 0cm |- fig-Moving-b);
+\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Moving-b-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:b}}}};
-\path (0cm,0cm) node (fig-Editing Different Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Moving-b-x) to (-2em, 0cm |- fig-Editing Different Parts of the Code-a);
+\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:b}}}};
+\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Different Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Moving-b-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Different Parts of the Code-b);
+\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Different Parts of the Code-c) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Editing Different Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
-\path [draw,->,merge step] (fig-Editing Different Parts of the Code-b) to (-2em, 0cm |- fig-Editing Different Parts of the Code-c);
+\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
+\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
+\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
+\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
-\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
+\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
+\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
+\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
 %%%%
-\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
+\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
+\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
-\path [draw,->,merge step] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
+\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
+\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
+\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
 %%%%
-\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-child-c-x) to[star] (-2em, 0cm |- fig-Multi-parent-a);
+\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-parent-a) to[star] (-2em, 0cm |- fig-Multi-parent-b);
+\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Multi-parent-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-parent-c);
+\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
-\path [draw,->,merge step] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
+\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
+\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
+\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-parent-e-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
+\path (-7.5em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
 %%%%
-\path (0cm,0cm) node (fig-Multi-User Cycles-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-parent-e-x) to[star] (-2em, 0cm |- fig-Multi-User Cycles-a);
+\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-b) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-User Cycles-a) to[star] (-2em, 0cm |- fig-Multi-User Cycles-b);
+\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-c) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Multi-User Cycles-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-User Cycles-c);
+\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Multi-User Cycles-d) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Multi-User Cycles-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-User Cycles-d);
-\path [draw,->,merge step] (fig-Multi-User Cycles-c) to (-2em, 0cm |- fig-Multi-User Cycles-d);
+\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
+\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
+\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
 %%%%
-\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Multi-User Cycles-d-x) to[star] (-2em, 0cm |- fig-Disconnection-a);
+\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
-\path [draw,->,alice step] (fig-Disconnection-a) to[star] (-2em, 0cm |- fig-Disconnection-b);
+\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
+\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
-\path [draw,->,bob step] (fig-Disconnection-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Disconnection-c);
+\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
+\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
 \end{tikzpicture}
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
-\path [draw,->,merge step] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
-\path [draw,->,merge step] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
+\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
+\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
+\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
 \end{tikzpicture}

### 2020-11-18 grove-paper.tex
Regularize vertex and edge names

@@
 \caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}
-\label{fig:Wrapping}
+\label{fig:Wrap}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentAlice}{L}
-\newvertex{DifferentAlice}{u}
-\newedge{DifferentBob}{R}
-\newvertex{DifferentBob}{v}
+\newedge{DifferentPartsAlice}{L}
+\newvertex{DifferentPartsAlice}{u}
+\newedge{DifferentPartsBob}{R}
+\newvertex{DifferentPartsBob}{v}
 \newcommand{\figureDifferentParts}{
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:b}}}};
+\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
 \path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
@@
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >alice edge, alice node],
+    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
     "\vSimpleY" [> "\eSimpleY"]
@@
     {},
-    "\vDifferentBob" [> "\eDifferentBob", >bob edge, bob node]
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
   }
@@
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"', >merge edge, merge node],
-    "\vDifferentBob" [> "\eDifferentBob", >merge edge, merge node]
+    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedAlice}{L}
-\newvertex{NestedAlice}{w}
-\newedge{NestedBob}{L}
+\newedge{NestedPartsAlice}{L}
+\newvertex{NestedPartsAlice}{w}
+\newedge{NestedPartsBob}{L}
 \newcommand{\figureNestedParts}{
@@
    "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice node],
-    "\vDifferentBob" [> "\eDifferentBob"]
+    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
    }
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
-    "\vDifferentAlice" [> "\eDifferentAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
+   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
+    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
-    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge node],
-    "\vDifferentBob" [> "\eDifferentBob"]
+   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
+    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
+    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
-    "\vDifferentBob" [> "\eDifferentBob"]
+   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
+    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
+    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
-    "\vNestedAlice" [> "\eNestedAlice"'],
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
+    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
     {}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
     {},
-    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
+    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'],
-   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
+   "\vSimpleTimes" [> "\eNestedPartsBob"'],
+   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
   }
@@
   a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
     {},
-    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
+    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
    },
@@
   v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
+   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
     {},
-    "\vNestedAlice" [> "\eMultiParentAlice"']
+    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
     "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
    },
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
    },
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnection
+%% \figureDisconnect
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectionAlice}{R}
-\newedge{DisconnectionBob}{L}
-\newcommand{\figureDisconnection}{
+\newedge{DisconnectAlice}{R}
+\newedge{DisconnectBob}{L}
+\newcommand{\figureDisconnect}{
 \begin{figure*}
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'],
+   "\vSimpleTimes" [> "\eNestedPartsBob"'],
    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
@@
   "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedBob"'] -> {
+   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
     {},
-    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice edge]
+    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
    },
@@
    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
+    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
     {}
@@
 so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrapping}:
+This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
 (a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
@@
 
-Alice adds~\vDifferentAlice{} as the left child of \vSimpleTimes{}
-and Bob changes \vSimpleY{} to \vDifferentBob{}.
+Alice adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{}
+and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
 Before transmitting their edits to each other,
@@
 \autoref{fig:DifferentParts:b} represents
-multiple edges, i.e., deleting~\eSimpleY{} and adding~\eDifferentBob{} along with its child~\vDifferentBob{}.
+multiple edges, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
 We thus mark it with a star.)\todo{star size}\todo{"sync"}
@@
 After Alice and Bob share the edits they made in the previous subsection,
-Alice changes~\vDifferentAlice{} to~\vNestedAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
+Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
 On the other hand, Bob moves~\vSimpleTimes{} and its children from the right child of~\vWrapPlus{} to its left child
-by deleting~\eMoveTimes{} and adding~\eNestedBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
+by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
 Since edits are based on edge identities,
@@
 Alice might receive Bob's deletion of~\eMoveTimes{} before making her
-changes and before receiving Bob's addition of~\eNestedBob{}.
+changes and before receiving Bob's addition of~\eNestedPartsBob{}.
 Her edits would then temporarily be within the code deleted by Bob.
@@
 This graph represents the expression~\texttt{(w * \hole{}) + \hole{}}.
-Alice moves~\vNestedAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
+Alice moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
 while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
@@
 In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedAlice{} is not a problem.
+Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
+Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
+As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
 However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-This generates \autoref{fig:MultiParent:d}, in which~\vNestedAlice{} has two parents pointing to it.
+This generates \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
 We call this situation a multi-parent conflict.
@@
 
-\figureDisconnection
+\figureDisconnect
 
@@
 When these edits are merged, the result is the graph in \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedBob{}
+Since Alice deleted~\eNestedPartsBob{}
 and Bob deleted~\eMultiCycleAliceTimes{},

### 2020-11-18 grove-paper.tex
add incomplete decomposition algorithm

@@
 \usepackage{todonotes}
+\usepackage[ruled]{algorithm2e}
 
@@
 \mathlig{::=}{\coloncolonequals}
+\mathlig{:=}{\colonequals}
 \mathlig{|-}{\vdash}
@@
 \caption{Example of Cycles.}
+% TODO replace screenshots with text
+% TODO make ref code gray
+% TODO move whole figure left
+% TODO remove cursor from screenshots
+% TODO add multiparent box
+% TODO only show parens where needed - figure 10, 11 - by associativity
+% TODO repplace cycle screenshot with cycles box 
 \label{fig:Cycle}
@@
   \caption{Syntax as a grammar for the lambda calculus that we are considering.}
+  % TODO add references into multiparent store and cycles store
   \label{fig:Syntax as a grammar}
@@
 
+% NOTE base grammar has holes. rooted expression grammar has refs, conflicts, and multiparents
+% NOTE ref = expression reference or vertex reference
+
 \subsection{Convergent Graphs}
@@
 
-If we translate the graph into expressions in the calculus, interactions with
-the graph can be modeled as a stream of user actions as described in
-\autoref{sub:User Actions}. Our decomposition algorithm partitions the graph
-into disjount connected components. We encode cycles by resolving all but one
-vertex of each multi-edge into a by-id reference to the one vertex.
+Translating a graph into expressions in the calculus allows us to model
+user interaction as a stream of edits as described in \autoref{sub:User
+Actions}. Our decomposition algorithm partitions the graph into disjount
+connected components. We encode cycles by resolving all but one vertex of each
+multi-edge into a by-id reference to the one vertex.
+
+\begin{algorithm}[t]
+  \caption{Graph Decomposition}
+  \label{fig:Graph Decomposition}
+
+  \DontPrintSemicolon
+  \SetAlgoNoLine
+  \SetKwFunction{reachable}{reachable}
+  \SetKwFunction{simpleCycles}{simpleCycles}
+  \SetKwProg{Fn}{function}{}{end}
+
+  \KwIn{Graph $G$}
+  \KwOut{Decomposition of $G$ into cycle-free expressions}
+
+  $liveEdges := \{\e : G(\e) = \Plus\}$\;
+  $liveVertexes := \bigcup_{(u, v, p, v') \in liveEdges} \{v, v'\}$\;
+  $allVertexes := \bigcup_{(u, v, p, v') \in \dom(G)} \{v, v'\}$\;
+  $multiparented := \{v \in allVertexes : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in \dom(G), v_1 \ne v_2\}$\;
+  $deleted := allVertexes \setminus liveVertexes \setminus \{v_0\}$\;
+  $MP := \{\reachable(v, \varnothing) : v \in multiparented\}$\;
+  $D := \{\reachable(v, \varnothing) : v \in deleted\}$\;
+  \Return{$(\reachable(v_0, \varnothing), MP, D, \simpleCycles(deleted))$}\;
+  \;
+
+  \Fn{\reachable $((u, k), seen)$}{
+    \If{$(u, k) \in multiparented \cup seen$}{
+      \Return{$u$}
+    }
+    $childEdges := \{(u', v, p, v') \in liveEdges : v = (u, k)\}$\;
+    $children := \{(p, \reachable(v, seen \cup \{u\})) : (p, v) \in childEdges\}$\;
+    \Return{$(k, children)$}
+  }
+  \;
+
+  \Fn{\simpleCycles $(remaining)$}{
+    \Return{}
+  }
+  \;
+\end{algorithm}
 

### 2020-11-18 grove-paper.tex
Remove now redundant UI screenshots

--- 2020-11-18 grove-paper.tex
@@
 \let\ast=\abxast
+%\def\*{\ast}
+\DeclareMathSymbol{*}{\mathbin}{symbols}{"03} % \ast
 
@@
 \begin{figure}
-\hskip0.12\linewidth
+\hskip0.12\columnwidth
 %%%%%%%%
@@
 \hfill
-%\hskip0.12\linewidth
+%\hskip0.12\columnwidth
 %%%%%%%%
@@
 %\hfill
-\hskip0.12\linewidth
+\hskip0.12\columnwidth
 %%%%%%%%
@@
 \newcommand{\figureMultiChild}{
-\begin{figure*}
+\begin{figure}
 %\hfill
-\hskip0.12\linewidth
+\hskip0.12\columnwidth
 %%%%%%%%
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
+\begin{subfigure}[t]{0.4\linewidth}
 \centering
@@
 %%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{}
-\label{fig:MultiChild:d}
-\includegraphics{code-views/multi-child-conflict.png}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
+%\hfill{}
 \caption{Example of Multi-Child Conflicts.}
 \label{fig:MultiChild}
-\end{figure*}
+\end{figure}
 }
@@
 \end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{}
-\label{fig:MultiParent:f}
-\includegraphics{code-views/multi-parent-conflict.png}
 \end{subfigure}
@@
 %\hfill
-\hskip0.12\linewidth
-%%%%%%%%
-\begin{subfigure}[t]{0.17\textwidth}
+\hskip0.12\columnwidth
+%%%%%%%%
+\begin{subfigure}[t]{0.21\textwidth}
 \centering
@@
 \label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}}%
+\figureCode{\parens{\hole{} * \hole{}} *\\\parens{\hole{} + \hole{}}+ \hole{}}%
 \begin{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.17\textwidth}
+\begin{subfigure}[t]{0.21\textwidth}
 \centering
@@
 \label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}}%
+\figureCode{\hole{} * \hole{} +\\\parens{\hole{} + \hole{}} * \hole{}}%
 \begin{tikzpicture}
@@
 \label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}}%
+\figureCode{\hole{} * \hole{} +\\\parens{\hole{} * \hole{} + \hole{}}}%
 \begin{tikzpicture}
@@
 \label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \multiChild{\multiVertex{30}}{\multiVertex{32}}}%
+\figureCode{\parens{\hole{} * \hole{}} +\\\multiChild{\multiVertex{30}}{\multiVertex{32}}}%
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{}
-\label{fig:Cycle:e}
-\includegraphics{code-views/cycle.png}
 \end{subfigure}
@@
 \label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}}%
+\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\\x}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}}%
+\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\\}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}}%
+\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\\}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}}%
+\figureCode{\hole{} + \hole{}\\\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.17\linewidth}
-\centering
-\caption{}
-\label{fig:Disconnect:e}
-\includegraphics{code-views/disconnected-reachable.png}
-\includegraphics{code-views/disconnected-unreachable.png}
 \end{subfigure}
@@
 
-\figureMultiChild
-
 After Alice and Bob share the edits they made in the previous subsection,
@@
 \label{sub:Multi-child conflicts}
+
+\figureMultiChild
 

### 2020-11-18 grove-paper.tex
Replace '*' with `\ast`

@@
 \let\ast=\abxast
-%\def\*{\ast}
-\DeclareMathSymbol{*}{\mathbin}{symbols}{"03} % \ast
 
@@
     \begin{array}{lrlll}
-      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid * \\
+      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid \ast{} \\
       \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
@@
     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-    \s : \G \to Act \times \V^{*} \times \V^{*} \times \V^{*}                                                                                          \\
+    \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
@@
 
-A \emph{graph} $G = [\e :-> s]_{u}^{*} \in \G = \A^{*}$ is a function from edges to edge states.
+A \emph{graph} $G = [\e :-> s]_{u}^{\ast{}} \in \G = \A^{\ast{}}$ is a function from edges to edge states.
 Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.
@@
 
-A \emph{communication action} $a = (\Send, A^{*}, u^{*}) \in \A^{comm} = \{ \Send \} \times \A^{*} \times \U^{*}$ applies the graph action sequence $A^{*}$ to the editors identified by UUIDs $u^{*}$.
+A \emph{communication action} $a = (\Send, A^{\ast{}}, u^{\ast{}}) \in \A^{comm} = \{ \Send \} \times \A^{\ast{}} \times \U^{\ast{}}$ applies the graph action sequence $A^{\ast{}}$ to the editors identified by UUIDs $u^{\ast{}}$.
 
@@
 
-%An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
+%An \emph{editor} $E = (G,c,A_{Q}^{\ast{}},A_{H}^{\ast{}},u) \in \EE = \mathcal{G} \times \C \times \AA^{\ast{}} \times \AA^{\ast{}} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{\ast{}}$, and known graph action history $A_{H}^{\ast{}}$, that can be identified by UUID $u$.
 
@@
 
-An \emph{environment} $(E^{*}, A^{*}) \in \EE^{*} \times \AA^{*}$ is a set of communicating editors $E^{*}$ that share a global graph action history $A^{*}$.
+An \emph{environment} $(E^{\ast{}}, A^{\ast{}}) \in \EE^{\ast{}} \times \AA^{\ast{}}$ is a set of communicating editors $E^{\ast{}}$ that share a global graph action history $A^{\ast{}}$.
 
@@
 
-$\known(E_1 \cdots E_n) = A_{H1}^{*} \cdots A_{Hn}^{*}
-  \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j))$
+$\known(E_1 \cdots E_n) = A_{H1}^{\ast{}} \cdots A_{Hn}^{\ast{}}
+  \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j))$
 
@@
   }{
-  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{*},A_H^{*},u),A
+  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
   }
@@
   }{
-  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{*},A_H^{*},u),A
+  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
   }
@@
   }{
-  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Up} (G,c',A_Q^{*},A_H^{*},u),A
+  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Up} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
   }
@@
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{*},A_H^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{\ast{}},A_H^{\ast{}},u),A
   }
@@
   \inferrule{}{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{*},A_H^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
   }
@@
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{*},A_H^{*}A_H',u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{\ast{}},A_H^{\ast{}}A_H',u),A
   }
@@
   \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
-  A_H'^{*} =
+  A_H'^{\ast{}} =
     [(c  ,(k,u_k),u  ) :-> +]
@@
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k}
-  (G A_H'^{*},c',A_Q^{*},A_H^{*}A_H'^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k}
+  (G A_H'^{\ast{}},c',A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
   }
@@
   \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
-  A_H'^{*} = [\e_j :-> -]_{j=1}^n
+  A_H'^{\ast{}} = [\e_j :-> -]_{j=1}^n
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Destroy}
-  (G A_H'^{*},c,A_Q^{*},A_H^{*}A_H'^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Destroy}
+  (G A_H'^{\ast{}},c,A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
   }
@@
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{*},A_H^{*}A_H,u),A %
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{\ast{}},A_H^{\ast{}}A_H,u),A %
   }
@@
   a \in \A^{edit} \\
-  (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*},A_H'^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}},A_H'^{\ast{}},u),A
   }{
-  (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*}a,A_H'^{*},u),A
+  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}}a,A_H'^{\ast{}},u),A
   }
@@
 
-$\boxed{E,A \xrightarrow{a^{*}} E,A}$
+$\boxed{E,A \xrightarrow{a^{\ast{}}} E,A}$
 %
@@
     E,A -->a E'',A'' \\
-    E'',A'' \xrightarrow{a^{*}} E',A' \\
+    E'',A'' \xrightarrow{a^{\ast{}}} E',A' \\
   }{
-    E,A \xrightarrow{aa^{*}} E',A'
+    E,A \xrightarrow{aa^{\ast{}}} E',A'
   }
@@
 
-$\boxed{(E,A)^{*} -->a (E,A)^{*}}$
+$\boxed{(E,A)^{\ast{}} -->a (E,A)^{\ast{}}}$
 %
@@
   \inferrule{
-  E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j) \\
-  E_j,A_j \xrightarrow{a'^{*}} (G_j',c_j',A_{Qj}'^{*}, A_{Hj}'^{*},u_j),A_j' \\
+  E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j) \\
+  E_j,A_j \xrightarrow{a'^{\ast{}}} (G_j',c_j',A_{Qj}'^{\ast{}}, A_{Hj}'^{\ast{}},u_j),A_j' \\
   E_j' =
@@
   , c_j'
-  , A_{Qj}'^{*} \setminus a'^{*}
-  , A_{Hj}'^{*} \setminus \known(E_1 \cdots E_n),u_j) \\
+  , A_{Qj}'^{\ast{}} \setminus a'^{\ast{}}
+  , A_{Hj}'^{\ast{}} \setminus \known(E_1 \cdots E_n),u_j) \\
   j = 1,\ldots,n
   }{
-  (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{*} } (E_1',A_1') \cdots (E_n',A_n')
+  (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{\ast{}} } (E_1',A_1') \cdots (E_n',A_n')
   }

### 2020-11-18 grove-paper.tex
Figure improvements

@@
 \newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
+\newcommand{\otherVertexVskip}{\vskip0.5em}
 
@@
 \label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}}%
+\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}}%
+\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w}%
+\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}}%
+\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}}%
+\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
 \begin{tikzpicture}
@@
 %\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
+\hskip0.07\columnwidth
+%%%%%%%%
+\begin{subfigure}[t]{0.23\textwidth}
+\centering
+\caption{
+\begin{tikzpicture}[remember picture, overlay]
+\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
 %%%%
@@
 \label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} *\\\parens{\hole{} + \hole{}}+ \hole{}}%
+\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
+  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
+  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
 \begin{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
+\begin{subfigure}[t]{0.23\textwidth}
 \centering
@@
 \label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} +\\\parens{\hole{} + \hole{}} * \hole{}}%
+\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
+  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
+  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
 \begin{tikzpicture}
@@
 \label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} +\\\parens{\hole{} * \hole{} + \hole{}}}%
+\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
+  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
+  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
 \begin{tikzpicture}
@@
 \label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} +\\\multiChild{\multiVertex{30}}{\multiVertex{32}}}%
+\figureCode{\parens{\hole{} * \hole{}} + \multiChild{\multiVertex{30}}{\multiVertex{32}}%
+  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
+  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
 \begin{tikzpicture}
@@
 %%%%%%%%
-\hfill{}
+%\hfill{}
 \caption{Example of Cycles.}
@@
 \label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\\x}%
+\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\\}%
+\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\\}%
+\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
 \begin{tikzpicture}
@@
 \label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\\\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
+\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
 \begin{tikzpicture}
@@
 
+%\figureCommutativity
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
+    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
+    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
+    \path ( 3cm, 0cm) node (d) [align=center]       {Synchronized \\ Version};
+    \path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
+    \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
+    \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
+    \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
+  \end{tikzpicture}
+  \caption{Commutativity of Edits.}
+  \label{fig:Commutativity}
+\end{figure}
+
 \subsection{Multi-user Interactions}
@@
 \figureNestedParts
-
-%\figureCommutativity
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]       {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
 

### 2020-11-18 grove-paper.tex
Remove references to no longer existing figures

@@
 represented by the fact that there are multiple edges going out of one index.
-In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}
-as shown in \autoref{fig:MultiChild:d}.
+In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}.
 
@@
 to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed
-as shown in \autoref{fig:Cycle:e}.
+For example, \autoref{fig:Cycle:d} would be displayed.
 We leave the user interface considerations of resolving the conflict such a

### 2020-11-18 grove-paper.tex
comment out Eric's notes appendix
Already folded in useful parts. All that's left are inter-editor
semantics, which we're not covering here.

@@
 
-\section{Appendix}
-\label{apx:Appendix}
-
-\subsection{Concepts}
-
-\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
-  $\A$ & Actions      & $\S$  & Edge States   \\
-  %  $\C$ & Cursors & $\U$ & UUIDs         \\
-  $\E$ & Edges        & $\V$  & Vertexes      \\
-  $\G$ & Graphs       & $\AA$ & Graph Actions \\
-  $\I$ & Indices      & $\EE$ & Editors       \\
-  $\K$ & Constructors                         \\
-\end{tabular}
-
-A \emph{UUID} $u \in \U$ is a universally unique identifier.
-
-\subsubsection{The Target Language}
-
-\paragraph{Constructor} \par
-
-A \emph{constructor} $k \in \K$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.
-
-\paragraph{Index} \par
-
-An \emph{index} $i \in \I$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST.
+% \section{Appendix}
+% \label{apx:Appendix}
+
+% \subsection{Concepts}
+
+% \begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
+%   $\A$ & Actions      & $\S$  & Edge States   \\
+%   %  $\C$ & Cursors & $\U$ & UUIDs         \\
+%   $\E$ & Edges        & $\V$  & Vertexes      \\
+%   $\G$ & Graphs       & $\AA$ & Graph Actions \\
+%   $\I$ & Indices      & $\EE$ & Editors       \\
+%   $\K$ & Constructors                         \\
+% \end{tabular}
+
+% A \emph{UUID} $u \in \U$ is a universally unique identifier.
+
+% \subsubsection{The Target Language}
+
+% \paragraph{Constructor} \par
+
+% A \emph{constructor} $k \in \K$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.
+
+% \paragraph{Index} \par
+
+% An \emph{index} $i \in \I$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST.
 Let $i_{root}$ be the hyper-position of all rooted constructors.
 
-\subsubsection{The Graph}
-
-\paragraph{Vertex} \par
-
-A \emph{vertex} $v = (k, u) \in \V = \K \times \U$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$.
+% \subsubsection{The Graph}
+
+% \paragraph{Vertex} \par
+
+% A \emph{vertex} $v = (k, u) \in \V = \K \times \U$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$.
 Let $v_{root} = (k_{root}, u_{root})$ be the root vertex of a graph.
@@
 
-\paragraph{Cursor} \par
-
-%A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
+% \paragraph{Cursor} \par
+
+% %A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
 Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.
 
-\paragraph{Edge} \par
-
-%An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
-
-\paragraph{Edge State} \par
-
-An \emph{edge state} $s \in \S = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).
-
-\paragraph{Graph Action} \par
-
-A \emph{graph action} $A = (\e,s,u) \in \A = \E \times \S \times \U$ is an instance of a binding from edge $\e$ to edge state $s$ that can be identified by UUID $u$.
+% \paragraph{Edge} \par
+
+% %An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
+
+% \paragraph{Edge State} \par
+
+% An \emph{edge state} $s \in \S = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).
+
+% \paragraph{Graph Action} \par
+
+% A \emph{graph action} $A = (\e,s,u) \in \A = \E \times \S \times \U$ is an instance of a binding from edge $\e$ to edge state $s$ that can be identified by UUID $u$.
 Denote by $[\e :-> s]_{u}$ a binding from edge $\e$ to edge state $s$ (identified by UUID $u$).
 
-\paragraph{Graph} \par
-
-A \emph{graph} $G = [\e :-> s]_{u}^{\ast{}} \in \G = \A^{\ast{}}$ is a function from edges to edge states.
+% \paragraph{Graph} \par
+
+% A \emph{graph} $G = [\e :-> s]_{u}^{\ast{}} \in \G = \A^{\ast{}}$ is a function from edges to edge states.
 Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.
 
-\subsubsection{The Editor}
-
-\paragraph{Move Action} \par
-
-%A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
-
-\paragraph{Edit Action} \par
-
-An \emph{edit action} $a \in \A^{edit}= \{ \Create \} \times \K \cup \{ \Destroy \} \cup \{ \Restore \} \times \V$ adds or removes an edge at the cursor.
-
-\paragraph{Communication Action} \par
-
-A \emph{communication action} $a = (\Send, A^{\ast{}}, u^{\ast{}}) \in \A^{comm} = \{ \Send \} \times \A^{\ast{}} \times \U^{\ast{}}$ applies the graph action sequence $A^{\ast{}}$ to the editors identified by UUIDs $u^{\ast{}}$.
-
-\paragraph{Action} \par
-
-An \emph{action} $a \in \A = \A^{move} \cup \A^{edit} \cup \A^{comm}$ describes a change to one or more editors.
-
-\paragraph{Editor} \par
-
-%An \emph{editor} $E = (G,c,A_{Q}^{\ast{}},A_{H}^{\ast{}},u) \in \EE = \mathcal{G} \times \C \times \AA^{\ast{}} \times \AA^{\ast{}} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{\ast{}}$, and known graph action history $A_{H}^{\ast{}}$, that can be identified by UUID $u$.
-
-\subsubsection{The Environment} \par
-
-An \emph{environment} $(E^{\ast{}}, A^{\ast{}}) \in \EE^{\ast{}} \times \AA^{\ast{}}$ is a set of communicating editors $E^{\ast{}}$ that share a global graph action history $A^{\ast{}}$.
-
-\subsection{Sets and Relations}
-
-\paragraph{Edges} \par
-
-\begin{align*}
-  \edges(G)            & = \dom(G)                                       \\
-  \liveEdges(G)        & = \{ \e \in \edges(G) | G(\e)=+ \}              \\
-  \vertexChildren(G,v) & = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
-  \cursorChildren(G,c) & = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
-  \parents(G,v)        & = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
-\end{align*}
-
-\paragraph{Vertexes} \par
-
-\begin{align*}
-  \vertexes(G)        & = \{ v \in \V | \exists (c,v,u) \in \edges(G) \}              \\
-  \liveVertexes(G)    & = \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}          \\
-  \childVertexes(G,v) & = \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \} \\
-  \orphans(G)         & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}      \\
-  \multiparents(G)    & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}      \\
-  \vertex(G,u)        & = v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
-\end{align*}
-
-\paragraph{Graph Actions} \par
-
-$\known(E_1 \cdots E_n) = A_{H1}^{\ast{}} \cdots A_{Hn}^{\ast{}}
-  \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j))$
-
-\section{Operational Semantics}
-
-\subsection{Movement}
-
-$\boxed{E,A -->a E,A}$
-
-\begin{mathpar}
-  \inferrule{
-  \leftIndex(i) = i'
-  }{
-  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
-  }
-
-  \inferrule{
-  \rightIndex(i) = i'
-  }{
-  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
-  }
-
-  \inferrule{
-  \parents(v) = \{(c',v',u')\}
-  }{
-  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Up} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
-  }
-
-  \inferrule{
-  \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
-  \downIndex(k) = i
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{\ast{}},A_H^{\ast{}},u),A
-  }
-
-  \inferrule{}{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
-  }
-\end{mathpar}
-
-\subsection{Editing}
-
-\begin{mathpar}
-  \inferrule{
-  \defaultIndex(c) = TODO(Was None) \\
-  u, u_k \in \U \text{ fresh} \\
-  A_H' = [(c,(k,u_k),u) :-> +]
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{\ast{}},A_H^{\ast{}}A_H',u),A
-  }
-
-  \inferrule{
-  \defaultIndex(c) = i' \\
-  c' = ((k,u_k),i') \\
-  u, u_k \in \U \text{ fresh} \\
-  \\\\
-  \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
-  A_H'^{\ast{}} =
-    [(c  ,(k,u_k),u  ) :-> +]
-    [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
-  [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k}
-  (G A_H'^{\ast{}},c',A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
-  }
-
-  \inferrule{
-  \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
-  A_H'^{\ast{}} = [\e_j :-> -]_{j=1}^n
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Destroy}
-  (G A_H'^{\ast{}},c,A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
-  }
-
-  \inferrule{
-  A_H = [(c,v,u') :-> +] \\
-  u' \text{ fresh}
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{\ast{}},A_H^{\ast{}}A_H,u),A %
-  }
-\end{mathpar}
-
-\subsection{Communication}
-
-\paragraph{Record} \par
-%
-\begin{mathpar}
-  \inferrule{
-  a \in \A^{edit} \\
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}},A_H'^{\ast{}},u),A
-  }{
-  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}}a,A_H'^{\ast{}},u),A
-  }
-\end{mathpar}
-
-\paragraph{Replay} \par
-
-$\boxed{E,A \xrightarrow{a^{\ast{}}} E,A}$
-%
-\begin{mathpar}
-  \inferrule{
-    E,A -->a E'',A'' \\
-    E'',A'' \xrightarrow{a^{\ast{}}} E',A' \\
-  }{
-    E,A \xrightarrow{aa^{\ast{}}} E',A'
-  }
-\end{mathpar}
-
-\paragraph{Publish} \par
-
-$\boxed{(E,A)^{\ast{}} -->a (E,A)^{\ast{}}}$
-%
-\begin{mathpar}
-  \inferrule{
-  E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j) \\
-  E_j,A_j \xrightarrow{a'^{\ast{}}} (G_j',c_j',A_{Qj}'^{\ast{}}, A_{Hj}'^{\ast{}},u_j),A_j' \\
-  E_j' =
-  ( G_j'
-  , c_j'
-  , A_{Qj}'^{\ast{}} \setminus a'^{\ast{}}
-  , A_{Hj}'^{\ast{}} \setminus \known(E_1 \cdots E_n),u_j) \\
-  j = 1,\ldots,n
-  }{
-  (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{\ast{}} } (E_1',A_1') \cdots (E_n',A_n')
-  }
-\end{mathpar}
-
-Missing Action.env members: Record / Report / Stop / Replay, Dump / Load, Clone / Drop
-
-$G \overset{\left(e,s\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right]$
-
-% Edge.Map.add action.value.edge (join old_state new_state) graph
+% \subsubsection{The Editor}
+
+% \paragraph{Move Action} \par
+
+% %A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
+
+% \paragraph{Edit Action} \par
+
+% An \emph{edit action} $a \in \A^{edit}= \{ \Create \} \times \K \cup \{ \Destroy \} \cup \{ \Restore \} \times \V$ adds or removes an edge at the cursor.
+
+% \paragraph{Communication Action} \par
+
+% A \emph{communication action} $a = (\Send, A^{\ast{}}, u^{\ast{}}) \in \A^{comm} = \{ \Send \} \times \A^{\ast{}} \times \U^{\ast{}}$ applies the graph action sequence $A^{\ast{}}$ to the editors identified by UUIDs $u^{\ast{}}$.
+
+% \paragraph{Action} \par
+
+% An \emph{action} $a \in \A = \A^{move} \cup \A^{edit} \cup \A^{comm}$ describes a change to one or more editors.
+
+% \paragraph{Editor} \par
+
+% %An \emph{editor} $E = (G,c,A_{Q}^{\ast{}},A_{H}^{\ast{}},u) \in \EE = \mathcal{G} \times \C \times \AA^{\ast{}} \times \AA^{\ast{}} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{\ast{}}$, and known graph action history $A_{H}^{\ast{}}$, that can be identified by UUID $u$.
+
+% \subsubsection{The Environment} \par
+
+% An \emph{environment} $(E^{\ast{}}, A^{\ast{}}) \in \EE^{\ast{}} \times \AA^{\ast{}}$ is a set of communicating editors $E^{\ast{}}$ that share a global graph action history $A^{\ast{}}$.
+
+% \subsection{Sets and Relations}
+
+% \paragraph{Edges} \par
+
+% \begin{align*}
+%   \edges(G)            & = \dom(G)                                       \\
+%   \liveEdges(G)        & = \{ \e \in \edges(G) | G(\e)=+ \}              \\
+%   \vertexChildren(G,v) & = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
+%   \cursorChildren(G,c) & = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
+%   \parents(G,v)        & = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
+% \end{align*}
+
+% \paragraph{Vertexes} \par
+
+% \begin{align*}
+%   \vertexes(G)        & = \{ v \in \V | \exists (c,v,u) \in \edges(G) \}              \\
+%   \liveVertexes(G)    & = \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}          \\
+%   \childVertexes(G,v) & = \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \} \\
+%   \orphans(G)         & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}      \\
+%   \multiparents(G)    & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}      \\
+%   \vertex(G,u)        & = v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
+% \end{align*}
+
+% \paragraph{Graph Actions} \par
+
+% $\known(E_1 \cdots E_n) = A_{H1}^{\ast{}} \cdots A_{Hn}^{\ast{}}
+%   \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j))$
+
+% \section{Operational Semantics}
+
+% \subsection{Movement}
+
+% $\boxed{E,A -->a E,A}$
+
+% \begin{mathpar}
+%   \inferrule{
+%   \leftIndex(i) = i'
+%   }{
+%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
+%   }
+
+%   \inferrule{
+%   \rightIndex(i) = i'
+%   }{
+%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
+%   }
+
+%   \inferrule{
+%   \parents(v) = \{(c',v',u')\}
+%   }{
+%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Up} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
+%   }
+
+%   \inferrule{
+%   \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
+%   \downIndex(k) = i
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{\ast{}},A_H^{\ast{}},u),A
+%   }
+
+%   \inferrule{}{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
+%   }
+% \end{mathpar}
+
+% \subsection{Editing}
+
+% \begin{mathpar}
+%   \inferrule{
+%   \defaultIndex(c) = TODO(Was None) \\
+%   u, u_k \in \U \text{ fresh} \\
+%   A_H' = [(c,(k,u_k),u) :-> +]
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{\ast{}},A_H^{\ast{}}A_H',u),A
+%   }
+
+%   \inferrule{
+%   \defaultIndex(c) = i' \\
+%   c' = ((k,u_k),i') \\
+%   u, u_k \in \U \text{ fresh} \\
+%   \\\\
+%   \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
+%   A_H'^{\ast{}} =
+%     [(c  ,(k,u_k),u  ) :-> +]
+%     [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
+%   [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k}
+%   (G A_H'^{\ast{}},c',A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
+%   }
+
+%   \inferrule{
+%   \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
+%   A_H'^{\ast{}} = [\e_j :-> -]_{j=1}^n
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Destroy}
+%   (G A_H'^{\ast{}},c,A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
+%   }
+
+%   \inferrule{
+%   A_H = [(c,v,u') :-> +] \\
+%   u' \text{ fresh}
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{\ast{}},A_H^{\ast{}}A_H,u),A %
+%   }
+% \end{mathpar}
+
+% \subsection{Communication}
+
+% \paragraph{Record} \par
+% %
+% \begin{mathpar}
+%   \inferrule{
+%   a \in \A^{edit} \\
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}},A_H'^{\ast{}},u),A
+%   }{
+%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}}a,A_H'^{\ast{}},u),A
+%   }
+% \end{mathpar}
+
+% \paragraph{Replay} \par
+
+% $\boxed{E,A \xrightarrow{a^{\ast{}}} E,A}$
+% %
+% \begin{mathpar}
+%   \inferrule{
+%     E,A -->a E'',A'' \\
+%     E'',A'' \xrightarrow{a^{\ast{}}} E',A' \\
+%   }{
+%     E,A \xrightarrow{aa^{\ast{}}} E',A'
+%   }
+% \end{mathpar}
+
+% \paragraph{Publish} \par
+
+% $\boxed{(E,A)^{\ast{}} -->a (E,A)^{\ast{}}}$
+% %
+% \begin{mathpar}
+%   \inferrule{
+%   E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j) \\
+%   E_j,A_j \xrightarrow{a'^{\ast{}}} (G_j',c_j',A_{Qj}'^{\ast{}}, A_{Hj}'^{\ast{}},u_j),A_j' \\
+%   E_j' =
+%   ( G_j'
+%   , c_j'
+%   , A_{Qj}'^{\ast{}} \setminus a'^{\ast{}}
+%   , A_{Hj}'^{\ast{}} \setminus \known(E_1 \cdots E_n),u_j) \\
+%   j = 1,\ldots,n
+%   }{
+%   (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{\ast{}} } (E_1',A_1') \cdots (E_n',A_n')
+%   }
+% \end{mathpar}
+
+% Missing Action.env members: Record / Report / Stop / Replay, Dump / Load, Clone / Drop
+
+% $G \overset{\left(e,s\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right]$
+
+% % Edge.Map.add action.value.edge (join old_state new_state) graph
 

### 2020-11-19 grove-paper.tex
finish the decomposition algorithm
also various minor edits

@@
 
-A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states,
+A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
 where $\E = \U \times \V \times \P \times \V$,
@@
 Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
-and connects the child position $p$ of a source vertex $v$ to a target vertex $v^\prime$.
-The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
-If $\G(\e) = \bot$, then $\e$ does not yet exist.
-If $\G(\e)$ is $\Plus$, the edge has been created.
-If $\G(\e)$ is $\Minus$, the edge has been destroyed.
+and connects the child position $p$ of source vertex $v$ to target vertex $v^\prime$.
+The state of each edge, $\e$, is determined by the corresponding edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
+When $G(\e) = \bot$, $\e$ has not been created yet.
+When $G(\e)$ is $\Plus$, the edge has been created.
+When $G(\e)$ is $\Minus$, the edge has been destroyed.
 The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
 
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
-where $k \in \K$ corresponds to one of the forms defined in \autoref{fig:Syntax
-as a grammar} that is not an operator or a hole. Each contructor has an arity,
-which we define as a set of positions. A position $p \in \P$ identifies a
-location, adjacent to some parent vertex, from which edges to child vertices may
-originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
-Arity} gives the complete mapping from constructors to potential edge
-positions. Every graph contains a distinguished root vertex, root constructor,
-and root position, respectively named $v_0$, $k_0$, and $p_0$.
-
+A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
+\K$, corresponding to a form defined in \autoref{fig:Syntax as a grammar} that
+is neither an operator nor a hole. Each contructor has an arity, which we define
+as a set of positions. A position $p \in \P$ identifies a location, adjacent to
+some source vertex, from which edges to target vertices may originate. The
+$\arity$ function in \autoref{fig:Constructors, Indexes and Arity} gives the
+complete mapping from constructors to potential edge positions. Every graph
+contains a distinguished root vertex, root constructor, and root position,
+respectively named $v_0$, $k_0$, and $p_0$.
+
+% TODO: should graph "recomposition" have an algorithm, too?
 A graph can be derived from an expression in the calculus by starting with an
@@
 ones at positions in $\arity(k')$, as if it were the outermost and create edges
-to the vertices representing its immediate sub-terms until only operators and
-holes remain. Note that empty holes are not modeled explicitly in the graph.
-Wherever an empty hole term appears, the graph has no edges originating from
-that position.
+to the vertices representing its immediate sub-terms until only empty holes
+remain. Note that empty holes are not modeled explicitly in the graph. Wherever
+an empty hole appears in an expression, the corresponding graph position has no
+edges originating from it.
 
@@
   \SetKwFunction{reachable}{reachable}
-  \SetKwFunction{simpleCycles}{simpleCycles}
   \SetKwProg{Fn}{function}{}{end}
@@
   \KwIn{Graph $G$}
-  \KwOut{Decomposition of $G$ into cycle-free expressions}
-
-  $liveEdges := \{\e : G(\e) = \Plus\}$\;
-  $liveVertexes := \bigcup_{(u, v, p, v') \in liveEdges} \{v, v'\}$\;
-  $allVertexes := \bigcup_{(u, v, p, v') \in \dom(G)} \{v, v'\}$\;
-  $multiparented := \{v \in allVertexes : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in \dom(G), v_1 \ne v_2\}$\;
-  $deleted := allVertexes \setminus liveVertexes \setminus \{v_0\}$\;
-  $MP := \{\reachable(v, \varnothing) : v \in multiparented\}$\;
-  $D := \{\reachable(v, \varnothing) : v \in deleted\}$\;
-  \Return{$(\reachable(v_0, \varnothing), MP, D, \simpleCycles(deleted))$}\;
+  \KwOut{A partitioning of the edges of $G$}
+
+  $live := \{\e : G(\e) = \Plus\}$\;
+  $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
+  $D := \{\reachable(\e) : \e \notin live\}$\;
+  $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
+  $MP := \{\reachable(\e) : \e \in multiparent\}$\;
+  $SC := \varnothing$\;
+  $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
+  \While{$remaining \ne \varnothing$}{
+    $(root, cycle) := \reachable(\min(remaining))$\;
+    $remaining := remaining \setminus \{root\} \setminus cycle$\;
+    $SC := SC \cup \{(root, cycle)\}$\;
+  }
+  \Return{$(R, D, MP, SC)$}\;
   \;
 
-  \Fn{\reachable $((u, k), seen)$}{
-    \If{$(u, k) \in multiparented \cup seen$}{
-      \Return{$u$}
-    }
-    $childEdges := \{(u', v, p, v') \in liveEdges : v = (u, k)\}$\;
-    $children := \{(p, \reachable(v, seen \cup \{u\})) : (p, v) \in childEdges\}$\;
-    \Return{$(k, children)$}
-  }
-  \;
-
-  \Fn{\simpleCycles $(remaining)$}{
-    \Return{}
-  }
-  \;
+  \Fn{$\reachable\ ((u, v, p, v'))$}{
+    $children := \{(u', v', p', v'') \in live\}$\;
+    $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
+    \Return{$((u, v, p, v'), descendants)$}
+  }
 \end{algorithm}

### 2020-11-19 grove-paper.tex
Revise Section 2.3

@@
 but for simplicity we consider only two: Alice and Bob.
-Alice and Bob each maintain their own editor instance and perform
+Alice and Bob each maintain their own editor instances and perform
 actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
 
@@
 
-To start, Alice and Bob both have the graph in \autoref{fig:Move:b}, which
-represents the code~\texttt{\hole{} + (\hole{} * y)}.
-
+We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
 Alice adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{}
@@
 multiple edges, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)\todo{star size}\todo{"sync"}
+We thus mark it with a star.)
 
@@
 For any sequences of edits from Alice and Bob,
-applying them in either order produces the same results.
+applying them in any order always produces the same results.
+(We show this formally in \autoref{sec:Commutativity}.)
 
@@
 
-After Alice and Bob share the edits they made in the previous subsection,
+After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
 Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-On the other hand, Bob moves~\vSimpleTimes{} and its children from the right child of~\vWrapPlus{} to its left child
+Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
 by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identities,
+Since edits are based on edge identifiers and not a location in the tree,
 both Alice's and Bob's edits can be applied to the other's graph,
-which results in both Alice and Bob having the graph in \autoref{fig:NestedParts:c}.
-
-A consequence of this design is that we allow users to modify
-parts of the code that have been deleted.  For example,
+which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
+
+A consequence of this commutativity is that users can modify
+deleted code.  For example,
 Alice might receive Bob's deletion of~\eMoveTimes{} before making her
@@
 
-TODO:figure
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
+\subsection{Merge Conflicts}
+\label{sub:Merge Conflicts}
+
+While the collaborative edits discussed so far merge cleanly,
+they can also generate graphs that do not map cleanly to a syntax tree.
+We call these conflicts and consider cases for \emph{multi-child},
+\emph{multi-parent}, and \emph{cycle} conflicts.\todo{disconnect}
+
+TODO: as with conflict in line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
+these require user intervention to resolve.
 
@@
 
-Next we address what happens when edits conflict.
-For example, Alice adds~\vMultiChildAlice{} as the right child of~\vWrapPlus{}
-while Bob adds~\vMultiChildBob{} also as the right child of~\vWrapPlus{}.
+As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts-c}.
+Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
+while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
 These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-These graphs have two different edges for the~\texttt{R} position of~\vWrapPlus{}.
-\autoref{fig:MultiChild:a} has~\eMultiChildAlice{}, which points to~\vMultiChildAlice{},
-and \autoref{fig:MultiChild:b} has~\eMultiChildBob{}, which points to~\vMultiChildBob{}.
-We call this a \emph{multi-child conflict}.
+Note that these graphs have two different edges for the~\texttt{R} position of~\vWrapPlus{}.
 
@@
 of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}.
-The conflict between Alice's and Bob's changes is
-represented by the fact that there are multiple edges going out of one index.
+This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
+in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
 In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}.
 
-This conflict must eventually be resolved by the user by deleting
-one of the edges and editing the other to be the merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{}
-to be~\texttt{x * y}.
+This conflict can be resolved by the user by deleting
+one of the edges in conflict and editing expression pointed to the other edge to be the correctly merged value.
+For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
 (Note that users can continue to edit the code while in this conflicted state.
@@
 
-TODO: Multiple adds cause conflicts (Not usually presented in the UI)
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Our tree-based model is similar in that it presents alternate versions
-of the code, and it is up to the user to replace those alternates with the desired merged result.
+This handling of conflicts is similar traditional merge conflicts in that
+it presents alternate versions of the code that the user replaces with the desired merge result.
 Our model is different in that merge conflicts in the line-based model
-prevent some semantic editor services from running
-that could continue to run in the tree-based model because
+prevent semantic editor services from running
+that could continue to run in the graph-based model because
 code inside and outside the conflict is still valid code.
-It is just the point where they meet the poses a potential problem for editor services.
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
+It is just the where the multi-child happens that editor services may break down.
+\autoref{sec:Related Work} discusses more differences.
 
@@
 
-In addition to conflicts due to multiple edges coming \emph{out of} the same child position of a particular vertex,
-conflicts can manifest as multiple edges going \emph{into} a particular vertex.
-We call these \emph{multi-parent conflicts}.
-
-An example of this is shown in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some more edits,
-and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(w * \hole{}) + \hole{}}.
-Alice moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
+While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
+multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
+An example of this is in \autoref{fig:MultiParent}.
+\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
+and Alice and Bob's shared their edits with each other.
+Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
 while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c},
-which represent respectively the expression~\texttt{(\hole{} * w) + \hole{}} for Alice
-and the expression~\texttt{(\hole{} * \hole{}) + w} for Bob.
+This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
 In both cases, these edits are achieved with one edge delete and one edge add.
@@
 However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-This generates \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-We call this situation a multi-parent conflict.
+When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
 
@@
 
+TODO: technically a multi-parent conflict
+
 TODO: paragraph about why single-user cycles cannot happen.
@@
 
+In addition to multi-child and multi-parent conflicts, merges
+can cause cycles in the graph.\footnote{In theory a single user could
+generate cycles, but this could be detected immediately and the user action
+that caused that cycle rejected.
+Thus, we are only concerned with cycles that cannot be immediately detected
+and prevented by the editor.}
 For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some more edits,
-and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{((\hole{} * \hole{}) * (\hole{} + \hole{})) + \hole}.
-Alice moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath.
-This results in \autoref{fig:Cycle:b},
-which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} + \hole{}) * \hole{})}.
-On the other hand, Bob does the same
-but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:c},
-which represents the expression~\texttt{(\hole{} * \hole{}) + ((\hole{} * \hole{}) + \hole{})}.
+\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
+and shared them with each other.
+Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
+and then~\vMultiCyclePlus{} underneath that.
+Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
+This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
 
@@
 between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-This situation is akin to a merge conflict in traditional version control systems.
-Resolving it requires user input, but is as simple as
+As with the other kinds of conflict, resolving it requires user input, but is as simple as
 either deleting~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
@@
 
-\subsubsection{Disconnection}
-\label{sub:Disconnection}
+\paragraph{Disconnected Cycles}
 
@@
 
+TODO: technically a cycle
+
+TODO: in general detect when there are or were edits to something now deleted
+
 \todo{UI would show orphans only once edits orphans occur}
@@
 Finally, when multi-user edits are merged,
-parts of the graph can become disconnected
-even though they are connected in each user's
+parts of the graph can become disconnected from the root
+even though they are connected to the root in each user's
 copy of the graph before the merge.
 
-For example, consider the situation in \autoref{fig:Disconnection}.
-\autoref{fig:Disconnect:a} shows the graph after Alice and Bob have made some more edits,
-and Alice and Bob's editors are synchronized with each other.
-This graph represents the expression~\texttt{(\hole{} * \hole{}) + (\hole{}  * \hole{})}.
-Alice moves~\vMultiCycleTimes{} under~\vSimpleTimes{}.
-This results in \autoref{fig:Disconnect:b},
-which represents the expression~\texttt{(\hole{} * (\hole{} * \hole{})) + \hole}.
-On the other hand, Bob puts~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:c},
-which represents the expression~\texttt{((\hole{} * \hole{}) * \hole{}) + \hole{}}.
-
-When these edits are merged, the result is the graph in \autoref{fig:Disconnect:d}.
+For example, consider the situation in \autoref{fig:Disconnection-a},
+which shows the graph after Alice and Bob have made some edits,
+and shared their edits with each other.
+Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
+while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
+This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
+
+When these edits are merged, the result is \autoref{fig:Disconnect:d}.
 Since Alice deleted~\eNestedPartsBob{}
@@
 since the disconnection appears after merging two graphs
-that did not have disconnected vertexes.
+in which vertexes were not already disconnected.
 
@@
 \subsection{Commutativity}
+\label{sub:Commutativity}
 
@@
 
+Contrast with a traditional diff
+
+Similar to how git merged across moves.
+A notable difference though is that \texttt{git} detects
+file moves by structural similarity rather than
+our nominal equality.
+(We exploit this in Subsection~REF:TODO).
+
+Compare this to what happens when there are conflicts in a traditional,
+line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
+Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
+when a merge conflict happens.
+These show alternate versions of the code, and it is up to the user
+to replace those alternates with the desired merged result.
+
+Another place our model is different
+is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
+Finally, in our model, since all changes are in terms of
+the structure of the code rather than lines,
+all changes respect that structure.
+This is unlike the line-based model where there is no guarantee
+that a merge conflict follows the grammatical structure of the code.
+
 Hazel

### 2020-11-19 grove-paper.tex
Try "Share Edits" instead of "Sync" in diagram

@@
     \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,auto,align=center] {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap,align=center]      {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
+    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
+    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
+    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
+    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
   \end{tikzpicture}

### 2020-11-20 grove-paper.tex
Revise cursor section

@@
 
-Finally we address the question of cursors.
-
-All edits use edge and vertex identifiers.
-Thus cursor positions do not affect the interpretation of graph edits.
-Cursors are not considered part of the graph representation,
+\paragraph{Representation}
+
+For most editing a cursor is represented by a position within a vertex.
+For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
+Representing it this way means that if Bob or some other user deletes TODO:vertex
+and replaces it with a different vertex, Bob's cursor is still valid.
+We specify a position within a vertex instead of a vertex because TODO.
+
+However, there is one case where we need a more precise cursor.
+That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
+We want users to beable to put their cussor on either the conflict as a whole
+(e.g., TODO: \{ x | y \})
+or on an individual element (e.g., x or y).
+In the former case, a pair of a vertex and a child position suffices.
+For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
+
+\paragraph{Communication}
+
+All in our system edits use edge and vertex identifiers, and user
+cursors do not affect the interpretation of graph edits.
+Thus, cursors are not part of the graph representation
 and need not be communicated to other users.
 
-COMMUNICATION
-
-TODO: use Alice and Bob
-
-However as a convienience to other users, the current cursor position
-may be transmitted to other users.
-In this role, cursors are merely advisory and we can follow the simple
-protocol of the sender marking every cursor announcement with a timestamp
-and the receiver displaying the cursor position With the most recently received position being the one displayed.
+However, in a collaborative setting, seeing the cursors of other users
+and be useful.
+Thus, the editor announce cursor position to other editors,
+At any point in time, the announcement for a particular user with
+the most recent timestamp is used.
+
 Note that this cursor position may refer to edges or verticies that
@@
 (Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-Don't model communication but would be separate layer anyway
-
-REPRESENTATION
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor doesn't move.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has multiple children in a particular position
-and we want the cursor to be on a specific child.
-In that case, we represent the cursor by an edge identifier.
-An example of this is seen in REF:figure.
-If Bob's cursor is on the \texttt{\{ x | y \}} as a whole,
-Bob's cursor would be the pair of the vertex id TODO and the position~\texttt{R}.
-If Bob's cursor is on a particular element in the multi-child,
-say the~\texttt{x}, then we represent the cursor position with edge identifier TODO:edge.
 

### 2020-12-01 grove-paper.tex
Revise Section 2.3

@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Simple:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Simple:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Simple:c}
@@
 \hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}
+\caption{Code that Includes a Hole and Some Simple Edits.}%
 \label{fig:Simple}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Wrap:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Wrap:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Wrap:c}
@@
 \hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}
+\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
 \label{fig:Wrap}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Move:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Move:b}
@@
 \hfill{}
-\caption{Example of Repositioning Code.}
+\caption{Example of Repositioning Code.}%
 \label{fig:Move}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:DifferentParts:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:DifferentParts:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:DifferentParts:c}
@@
 \hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}
+\caption{Example of Users Editing Different Parts of the Code.}%
 \label{fig:DifferentParts}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:NestedParts:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:NestedParts:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:NestedParts:c}
@@
 \hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}
+\caption{Example of Users Editing Nested Parts of the Code.}%
 \label{fig:NestedParts}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiChild:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiChild:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiChild:c}
@@
 %\hfill{}
-\caption{Example of Multi-Child Conflicts.}
+\caption{Example of Multi-Child Conflicts.}%
 \label{fig:MultiChild}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiParent:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiParent:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiParent:c}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiParent:d}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:MultiParent:e}
@@
 \hfill{}
-\caption{Example of Multi-Parent Conflicts.}
+\caption{Example of Multi-Parent Conflicts.}%
 \label{fig:MultiParent}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Cycle:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Cycle:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Cycle:c}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Cycle:d}
@@
 % TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box 
+% TODO repplace cycle screenshot with cycles box %
 \label{fig:Cycle}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Disconnect:a}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Disconnect:b}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Disconnect:c}
@@
 \end{tikzpicture}
-}
+}%
 \label{fig:Disconnect:d}
@@
 \hfill{}
-\caption{Example of Disconnection.}
+\caption{Example of Disconnection.}%
 \label{fig:Disconnection}
@@
 
-\section{Introduction}
+\section{Introduction}%
 \label{sec:Introduction}
@@
 
-\subsection{Contributions and Paper Organization}
+\subsection{Contributions and Paper Organization}%
 \label{sec:Contributions and Paper Organization}
 
-\section{Grove By Example}
+\section{Grove By Example}%
 \label{sec:Grove By Example}
@@
 
-\subsection{Program Representation}
+\subsection{Program Representation}%
 \label{sub:Program Representation}
@@
 
-\subsection{Single-User Actions}
+\subsection{Single-User Actions}%
 \label{sub:Single-User Actions}
@@
 
-\subsubsection{Construction Actions}
+\subsubsection{Construction Actions}%
 \label{sub:Construction}
@@
 
-\subsubsection{Deletion Actions}
+\subsubsection{Deletion Actions}%
 \label{sub:Deletion}
@@
 
-\subsubsection{Wrapping Actions}
+\subsubsection{Wrapping Actions}%
 \label{sub:Wrapping}
@@
 
-\subsubsection{Repositioning Actions}
+\subsubsection{Repositioning Actions}%
 \label{sub:Repositioning}
@@
 
-\subsection{Multi-user Interactions}
+\subsection{Multi-user Interactions}%
 \label{sub:Multi-user Interactions}
@@
 
-\todo{START REV HERE}
-
-TODO: mention that id numbers are global (or maybe use A1 and B1 for different users?)
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Common scenarios first and then consider more unusual scenarios
-that cause cycles or vertices with multiple parents
-
-\subsubsection{Editing Different Parts of the Code}
+We also assume Alice and Bob have some way to create globally unique identifiers for verticies and edges
+(e.g., GUIDs).
+
+\subsubsection{Editing Different Parts of the Code}%
 \label{sub:Editing Different Parts of the Code}
@@
 We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{}
+Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
 and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
 Before transmitting their edits to each other,
-Alice and Bob have the graphs in \autoref{fig:DifferentParts:a}
+Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
 and \autoref{fig:DifferentParts:b}, respectively.
@@
 \autoref{fig:DifferentParts:b} represents
-multiple edges, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
+multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
 We thus mark it with a star.)
@@
 
-\subsubsection{Commutativity}
-\label{sub:Commutativity}
+\subsubsection{Commutativity}%
+\label{sub:Commutativity:informal}
 
@@
 For example, suppose Alice and Bob did not transmit their edits to each other
-but to a third user, Chris, that has the original graph in \autoref{fig:Move:b}.
+but instead to a third user, Chris.
 Chris can apply the edits from Alice and Bob in any order and will always get
@@
 
-Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
+In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
 For any sequences of edits from Alice and Bob,
 applying them in any order always produces the same results.
-(We show this formally in \autoref{sec:Commutativity}.)
-
-\subsubsection{Editing Nested Parts of the Code}
+(We show this formally in \autoref{sub:Commutativity:informal}.)
+
+\subsubsection{Editing Nested Parts of the Code}%%
 \label{sub:Editing Nested Parts of the Code}
@@
 by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not a location in the tree,
+Since edits are based on edge identifiers and not locations in a tree,
 both Alice's and Bob's edits can be applied to the other's graph,
@@
 
-A consequence of this commutativity is that users can modify
-deleted code.  For example,
-Alice might receive Bob's deletion of~\eMoveTimes{} before making her
-changes and before receiving Bob's addition of~\eNestedPartsBob{}.
-Her edits would then temporarily be within the code deleted by Bob.
+Commutativity has consequences for this scenario.
+Commutativity require that edits produce the same result regardless of the order in which they are applied.
+So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
+changes before receiving Bob's addition of~\eNestedPartsBob{},
+the result must still be the graph in \autoref{fig:NestedParts:c}.
+In that case, her edits would then temporarily be within the code deleted by Bob.
+Thus, our semantics must allows users to edit deleted code.
+(Whether or how this is presented in the user interface is another question.)
+
 TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
 
-\subsection{Merge Conflicts}
+\subsection{Merge Conflicts}%
 \label{sub:Merge Conflicts}
@@
 We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.\todo{disconnect}
-
-TODO: as with conflict in line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these require user intervention to resolve.
-
-\subsubsection{Multi-child conflicts}
+\emph{multi-parent}, and \emph{cycle} conflicts.
+
+As with conflict in line/diff-based version-control
+system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
+these all require user intervention to resolve.
+
+We consider common merge scenarios first, and then consider more unusual scenarios.
+
+\subsubsection{Multi-child conflicts}%
 \label{sub:Multi-child conflicts}
@@
 
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts-c}.
+As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
 Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
@@
 These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} edges as edges going out of the \texttt{R} position
-of \vWrapPlus{}.
+Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
+
+We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
+as edges going out of the \texttt{R} position of \vWrapPlus{}.
 This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
@@
 
+TODO: "delete both"
+
 This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing expression pointed to the other edge to be the correctly merged value.
+one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
 For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
@@
 
-\subsubsection{Multi-parent conflicts}
+\subsubsection{Multi-parent conflicts}%
 \label{sub:Multi-parent conflicts}
@@
 \autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob's shared their edits with each other.
+and Alice and Bob shared their edits with each other.
 Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
@@
 
-\subsubsection{Cycles}
+\subsubsection{Cycles}%
 \label{sub:Multi-User Cycles}
@@
 
-For example, consider the situation in \autoref{fig:Disconnection-a},
+For example, consider the situation in \autoref{fig:Disconnect:a},
 which shows the graph after Alice and Bob have made some edits,
@@
 
-\subsection{Cursors}
+\subsection{Cursors}%
 \label{sub:Cursors}
@@
 
-\section{Formalism}
+\section{Formalism}%
 \label{sec:Formalism}
@@
 
-\subsection{Syntax}
+\subsection{Syntax}%
 \label{sub:Syntax}
@@
 
-\subsection{Convergent Graphs}
+\subsection{Convergent Graphs}%
 \label{sub:Convergent Graphs}
 
-\subsubsection{Graphs}
+\subsubsection{Graphs}%
 \label{sub:Graphs}
@@
 
-\subsubsection{Graph Edits}
+\subsubsection{Graph Edits}%
 \label{sub:Graph Edits}
@@
 
-\subsection{Commutativity}
-\label{sub:Commutativity}
+\subsection{Commutativity}%
+\label{sub:Commutativity:formal}
 
@@
 
-\subsubsection{Agda Mechanization}
+\subsubsection{Agda Mechanization}%
 \label{sub:Agda Mechanization}
 
-\subsection{Graph Decomposition}
+\subsection{Graph Decomposition}%
 \label{sub:Graph Decomposition}
@@
 
-\subsection{User Actions}
+\subsection{User Actions}%
 \label{sub:User Actions}
@@
 
-\section{Type System}
+\section{Type System}%
 \label{sec:Type System}
@@
 
-\section{Implementation}
+\section{Implementation}%
 \label{sec:Implementation}
@@
 
-\section{Related Work}
+\section{Related Work}%
 \label{sec:Related Work}
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion and Conclusion}%
 \label{sec:Discussion and Conclusion}
+
+TODO: (Place somewhere) This move semantics gives us a richer structure than
+when treating code as a list of lines.  We exploit
+this in Section~REF:TODO in order merge edits that involve moving code.
 
@@
 
-\subsection{Variable names, strings, and numbers}
+\subsection{Variable names, strings, and numbers}%
 \label{sub:Variable names, strings, and numbers}
@@
 
-\section{Supplemental Material: Complete Graph Sequences for All Figures}
+\section{Supplemental Material: Complete Graph Sequences for All Figures}%
 \label{apx:Supplemental Material: Complete Graph Sequences for All Figures}

### 2020-12-03 grove-paper.tex
Revisions to "Cycle" section

@@
 
-\figureSimple
+\figureSimple{}
 
@@
 
-\figureWrap
+\figureWrap{}
 
@@
 
-\figureMove
+\figureMove{}
 
@@
 
-\figureDifferentParts
-
-%\figureCommutativity
+\figureDifferentParts{}
+
+%\figureCommutativity{}
 \begin{figure}
@@
 
-\figureNestedParts
+\figureNestedParts{}
 
@@
 
-\figureMultiChild
+\figureMultiChild{}
 
@@
 
-\figureMultiParent
+\figureMultiParent{}
 
@@
 
-% \figureSingleUserCycles
+% \figureSingleUserCycles{}
 
@@
 
-\figureCycle
-
-TODO: technically a multi-parent conflict
-
-TODO: paragraph about why single-user cycles cannot happen.
-While single-user edits can be checked to see
-whether they create a cycle and thus the system
-can immediately warn the user,
-multi-user edits may not manifest cycles until after
-the edits from multiple users are merged.
+\figureCycle{}
 
@@
 between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input, but is as simple as
-either deleting~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
+As with the other kinds of conflict, resolving it requires user input.
+For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
+or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
+
+\paragraph{Disconnected Cycles}
+
+\figureDisconnect{}
+
+TODO: technically a cycle
+
+TODO: in general detect when there are or were edits to something now deleted
+
+\todo{UI would show orphans only once edits orphans occur}
+
+Finally, there is a particular type of cycle worth discussing.
+
 Since a cycle like this that is connected to the root vertex always
@@
 cycle represents to future work.
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
 

### 2020-12-05 grove-paper.tex
add implementation notes on tree decomp

@@
 
+Related: hypertree decomposition is an established topic in AI/DB. Apparently,
+it's NP-complete in general, so people look for tractable subproblems. Recent
+work focuses on reduction to SMT and parallelizability for large db-like
+querying of large graphs, but some papers also focus on efficient decompopsition
+for "nearly acyclic" graphs. The closest I could find to what we're already
+doing are the k-decomp / opt-k-decomp / det-k-decomp algoriths in this paper:
+https://dl.acm.org/doi/10.1145/1412228.1412229, except we're getting determinism
+for free (i.e., not having to guess-and-check pivot points) by virtue of our
+global root and uuid-based total ordering on basically everything.
+
 \section{Related Work}%

### 2020-12-08 grove-paper.tex
Edits from todays meeting

@@
 This section generalizes to any number of users,
-but for simplicity we consider only two: Alice and Bob.
+but for simplicity we consider only two in our examples: Alice and Bob.
 Alice and Bob each maintain their own editor instances and perform
@@
 
-We also assume Alice and Bob have some way to create globally unique identifiers for verticies and edges
-(e.g., GUIDs).
+Users coordinate by sharing with one another the graph edits corresponding to each
+action~(e.g., via direct network connections or relayed through a server).
+TODO: CITE(how to share network data -- DHT? consensus? ).
+The granularity of communication may vary.
+For real-time collaborative editing, edits may be sent as they occur.
+For version control, edits may be batched together in the form of commits.
+As we will show, these edits are commutative, so editors converge to the same state
+when all edits are shared.
+We assume users have some way to create globally unique identifiers for
+verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
 
@@
 
+\todo{rev}
 Note that aside from transmitting edge deletions and additions,

### 2020-12-08 grove-paper.tex
Revise Section 2.4

@@
 We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \{ x | y \})
-or on an individual element (e.g., x or y).
+(e.g., TODO: \texttt{\{ x | y \}})
+or on an individual element (e.g., \texttt{x} or \texttt{y}).
 In the former case, a pair of a vertex and a child position suffices.
@@
 
-All in our system edits use edge and vertex identifiers, and user
+In our system all edits use edge and vertex identifiers, and user
 cursors do not affect the interpretation of graph edits.
-Thus, cursors are not part of the graph representation
-and need not be communicated to other users.
-
+Thus, cursors need not be communicated to other users.
 However, in a collaborative setting, seeing the cursors of other users
-and be useful.
-Thus, the editor announce cursor position to other editors,
-At any point in time, the announcement for a particular user with
-the most recent timestamp is used.
+can be useful.
+For this purpose, editors can announce their cursor position to other editors.
+At any point in time, the announcement
+with the most recent timestamp
+for a particular user
+is used.
 

### 2020-12-08 grove-paper.tex
Add links to some preliminary related work

@@
 
-Hazel
-\citep{Omar:2019:10.1145/3290327}
+\subsection{Trees}
+
+\subsubsection{Structure editors}
+
+Hazel~\citep{Omar:2019:10.1145/3290327}
+
+\subsubsection{Tree diff}
+
+J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
+
+Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
+
+  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
+
+  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
+
+  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
+
+  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
+
+An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
+
+  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
+
+  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
+
+  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
+
+  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
+
+  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
+
+  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
+
+  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
+
+  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
+
+  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
+
+  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
+
+\subsection{Version Control}
+
+Git \url{https://git-scm.com/}
+
+Darcs \url{https://darcs.net/}
+
+  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
+
+Hg? \url{https://www.mercurial-scm.org/}
+
+SVN \url{https://subversion.apache.org/}
+
+Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
+
+  \url{https://pijul.org/}
+  \url{https://pijul.org/manual/theory.html}
+
+  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
+
+\subsection{Collaborative Editing}
 
@@
 
-Pijul and Anu
-
 Etherpad
@@
 
-Git
-
-Darcs
-
-Unision?
-
-Homotopical Patch Theorey: https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7
+\subsection{Synchronization}
+
+Unision
+\url{https://www.cis.upenn.edu/~bcpierce/unison/}
+\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
+
+\subsection{TODO}
+
+Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
+Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
 

### 2020-12-29 grove-paper.tex
minor clean up, mostly graph decomp

@@
 \usepackage{todonotes}
-\usepackage[ruled]{algorithm2e}
+% \usepackage[ruled]{algorithm2e}
 
@@
 \DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
 \DeclareMathOperator{\dom}{\text{dom}}
@@
 \DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parents}{\text{parents}}
 \DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
@@
 \DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
+
+% graph decomposition operators
+\DeclareMathOperator{\parents}{\text{parents}}
+\DeclareMathOperator{\multiParented}{\text{multiParented}}
+\DeclareMathOperator{\singleParented}{\text{singleParented}}
+\DeclareMathOperator{\deleted}{\text{deleted}}
 
@@
 \]
-In words, applying graph action $\left(s, \e\right)$ to graph $G$ results in the
+Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
 updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
@@
 
-Translating a graph into expressions in the calculus allows us to model
-user interaction as a stream of edits as described in \autoref{sub:User
-Actions}. Our decomposition algorithm partitions the graph into disjount
-connected components. We encode cycles by resolving all but one vertex of each
-multi-edge into a by-id reference to the one vertex.
-
-\begin{algorithm}[t]
-  \caption{Graph Decomposition}
-  \label{fig:Graph Decomposition}
-
-  \DontPrintSemicolon
-  \SetAlgoNoLine
-  \SetKwFunction{reachable}{reachable}
-  \SetKwProg{Fn}{function}{}{end}
-
-  \KwIn{Graph $G$}
-  \KwOut{A partitioning of the edges of $G$}
-
-  $live := \{\e : G(\e) = \Plus\}$\;
-  $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
-  $D := \{\reachable(\e) : \e \notin live\}$\;
-  $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
-  $MP := \{\reachable(\e) : \e \in multiparent\}$\;
-  $SC := \varnothing$\;
-  $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
-  \While{$remaining \ne \varnothing$}{
-    $(root, cycle) := \reachable(\min(remaining))$\;
-    $remaining := remaining \setminus \{root\} \setminus cycle$\;
-    $SC := SC \cup \{(root, cycle)\}$\;
-  }
-  \Return{$(R, D, MP, SC)$}\;
-  \;
-
-  \Fn{$\reachable\ ((u, v, p, v'))$}{
-    $children := \{(u', v', p', v'') \in live\}$\;
-    $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
-    \Return{$((u, v, p, v'), descendants)$}
-  }
-\end{algorithm}
-
-The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
-\wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
-portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
-expressions that involve vertices with multiple parents. $D$ contains deleted
-expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
-$SC$ contains the deleted expressions that do not involve multiple parents but
-do contain cycles.
-
-A simple cycle is a set of vertices connected by a closed path, consisting
-of:
-
-1) two or more vertices, all descendants of each other, and
-2) the vertices of any non-multiparent subtrees reachable from the core, and
-not reachable from the root vertex.
-
-To determine the simple cycles of a graph, we:
-
-1) check if any vertex in the graph is a descendant of itself or some other
-vertex that is (a descendant of itself),
-2) find the least vertex in the core of this new(ly discovered) cycle, and
-3) return a tree connecting every vertex reachable from the least (vertex).
-
-The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.
-
-<Mathematical definition of above.>
+% To render the graph as disjoint fragments of program code, we decompose the
+% graph into an equivalent representation using trees. The result is a 4-tuple,
+% where the first component is a tree, the second and third components are sets of
+% trees, and the fourth component is a set of cycles.
+
+% 1. ($O(n \log n)$ for maps, $O(n)$ for hash table) linear scan to partition vertices according to # of parents: multiple parents, single parents, no parents.
+
+% 2. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from root (and remove from single-parented set)
+
+% 3. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from multi-parented vertices (and remove from single-parented set).
+
+% 4. ($n \log n$?) scan to discover edges leading to single-parented [deleted]
+
+% 5. simple cycles—reverse traversal, then forward traversal. Here's where we want to prove some things (e.g., dangling vertices are not deleted)
+
+% tree ::= Vertex (Constructor) | Ref (Vertex.Id) | Conflict (tree list)
+
+%
+% ... () of sets of
+% (poly)trees) wherein each multi-edge has been resolved into a simple edge with
+% one or more symbolic references to it.
+
+% Our decomposition algorithm first separates the graph into three disjoint sets of
+% vertices according to the number of parents and presence of simple cycles.
+% %
+% The first set is the single-parented vertices that have the root as an ancestor.
+% The second set is the vertices that have a multi-parented vertex as an ancestor.
+% The third
+
+% We then derive equivalent tree representations of connected components by tracing
+% the ancestry, taken as the transitive closure of the childhood relation, of the
+% vertices in each set.
+
+% set contains the multi-parented vertices and
+% their non-multiparented descendants. The second set contains the vertices with
+% no parents and their non-multiparented descendants. We can determine both of
+% these sets with a single pass over all of the live vertices in the graph. The
+% third set contains the single-parented vertices that are descendants of the root
+% vertex.
+
+% The first set contains all of the
+% vertices that have the root as an ancestor.
+
+% : vertices with multiple parents, vertices with no parents,
+% single-parented vertices reachable from the root, and everything else.
+
+% separate it into 
+% components that only touch  detect any vertices that either have multiple parents or no parents
+% at all.
+
+% partition its vertices into three disjoint sets: multi-parented, deleted,
+% and any remaining vertices participating in a live edge. We then partition the
+% graph into four non-overlapping sets of trees: multi-parented, single-parented
+% reachable from root, deleted, and simple cycles.
+
+%  detect any multi-parented
+% vertices and use them to partition the remaining vertices of the graph into .
+
+% Our decomposition algorithm partitions the graph into trees of
+% connected components. Cycles are resolved by replacing all but one vertex of
+% each multi-edge with a by-id reference.
+
+% Translating a graph into expressions in the calculus allows us to model user
+% interaction as a stream of edits as described in \autoref{sub:User Actions}.
+
+% \begin{algorithm}[t]
+%   \caption{Graph Decomposition}
+%   \label{fig:Graph Decomposition}
+
+%   \DontPrintSemicolon
+%   \SetAlgoNoLine
+%   \SetKwFunction{reachable}{reachable}
+%   \SetKwProg{Fn}{function}{}{end}
+
+%   \KwIn{Graph $G$}
+%   \KwOut{A partitioning of the edges of $G$}
+
+%   $live := \{\e : G(\e) = \Plus\}$\;
+%   $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
+%   $D := \{\reachable(\e) : \e \notin live\}$\;
+%   $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
+%   $MP := \{\reachable(\e) : \e \in multiparent\}$\;
+%   $SC := \varnothing$\;
+%   $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
+%   \While{$remaining \ne \varnothing$}{
+%     $(root, cycle) := \reachable(\min(remaining))$\;
+%     $remaining := remaining \setminus \{root\} \setminus cycle$\;
+%     $SC := SC \cup \{(root, cycle)\}$\;
+%   }
+%   \Return{$(R, D, MP, SC)$}\;
+%   \;
+
+%   \Fn{$\reachable\ ((u, v, p, v'))$}{
+%     $children := \{(u', v', p', v'') \in live\}$\;
+%     $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
+%     \Return{$((u, v, p, v'), descendants)$}
+%   }
+% \end{algorithm}
+
+% The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
+% \wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
+% portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
+% expressions that involve vertices with multiple parents. $D$ contains deleted
+% expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
+% $SC$ contains the deleted expressions that do not involve multiple parents but
+% do contain cycles.
+
+% A simple cycle is a set of vertices connected by a closed path, consisting
+% of:
+
+% 1) two or more vertices, all descendants of each other, and
+% 2) the vertices of any non-multiparent subtrees reachable from the core, and
+% not reachable from the root vertex.
+
+% To determine the simple cycles of a graph, we:
+
+% 1) check if any vertex in the graph is a descendant of itself or some other
+% vertex that is (a descendant of itself),
+% 2) find the least vertex in the core of this new(ly discovered) cycle, and
+% 3) return a tree connecting every vertex reachable from the least (vertex).
+
+% The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.
+
+% <Mathematical definition of above.>
 

### 2020-12-29 grove-paper.tex
add some text on graph decomposition

@@
 
+Graph decomposition occurs in three steps.
+
+First, a linear scan is performed over every edge that has been created or
+deleted. The vertices at both ends are partitioned into three sets:
+multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
+child edge set lookups are also created.
+
+Second, the subgraphs rooted at vertices with more or less than one parent are
+traversed and equivalent tree structures are generated over components that are
+connected by single-parented vertices. Cycles are broken with UUID vertex
+references. Single-parented vertices reachable from the root vertex are also
+traversed.
+
+Finally, any remaining single-parented vertices are parts of wreaths. A wreath
+is a cycle of single-parented vertices that each may have single-parented
+descendants.
+
+Wreath traversal occurs in two directions. A vertex is chosen as an arbitrary
+starting point and its parents are traversed and marked as seen. Any vertex that
+has been seen twice must be on the cycle. We take the first one as the root of
+the wreath and traverse its children to produce a tree that covers the wreath.
+Any remaining single-parented vertices must be parts of other wreaths and are
+wreath-traversed until there are no more single-parented vertices left.
+
 % To render the graph as disjoint fragments of program code, we decompose the

### 2021-01-06 grove-paper.tex
add tree syntax as a grammar
commit draft notes from last Grove meeting

@@
 
+% XXX
+
+(illustrative running example: 1 of each: mp, d, r, sc)
+
+declarative section: give a hint what decomposition does (incl. wreaths): take a
+graph, turn it into a grove (i.e., 4-tuple) which is what we actually present to
+users and use for things like type checking. In this section, we present an
+algorithm that declaratively say what a grove is and how it connects to the
+graph.
+
+We can't turn the graph into a single tree, so we turn it into a forest. A
+forest is a collection of trees, in fact it's a 4-tuple consisting of a rooted tree
+and 3 sets of trees: mp, d, sc. Before discussing what these sets are, let's 
+define what a tree is.
+
+here's what a tree is. why do we care what a tree is? because we can map them to
+the expressions in figure 11. Explain what a Ref is wrt forest roots. (Maybe
+mention that only well-sorted trees can be mapped, which we discuss below.)
+
+There's this weird stuff left over; we call them wreaths and put them in the 4th
+tuple component.
+
+here's what a wreath is (= a connected graph where each vertex has exactly one
+parent)
+
+(show what a wreath looks like)
+
+here's what a grove is
+
+here are some properties that show how to get one from the other, and to/from
+graph
+
+algorithmic section: if we want to implement these conversions, here are the
+steps (without introducing new concepts):
+
+- linear scan
+- breaking out roots
+- ...
+
 Graph decomposition occurs in three steps.
@@
 
-Finally, any remaining single-parented vertices are parts of wreaths. A wreath
-is a cycle of single-parented vertices that each may have single-parented
+Finally, once we've already accounted for things with more or less than one
+parent, everything else must be single parented. We call these things wreaths. A
+wreath is a cycle of single-parented vertices that each may have single-parented
 descendants.
+
+lemma: a wreath is a cycle with trees hanging off of it
+
+lemma: wreath traversal always produces a valid tree
 
@@
 wreath-traversed until there are no more single-parented vertices left.
+
+% XXX
+
+decomposition produces sets of trees. As long as everything is well sorted, we
+can use them to produce expressions in the main grammar. (alt: define wrt the
+graph instead, once we establish the graph can be turned into expressions)
+
+*** well-sortedness judgment on trees: the tree only has valid positions for the
+constructors, and the child at each position has the sort corresponding to its
+position, and edges in position maps all start at the given constructor.
+
+``A graph has this property iff these conditions hold''
+
+then we can define a total mapping from well-sorted trees to expressions
+(because the tree structure is very generic, so this ties them back to the
+earlier grammar.)
+
+forest = 4-tuple
+
+forest validity: each of the trees in the forest are well sorted, and all Refs
+point to roots of other trees of the correct sort.
+
+every valid forest corresponds to a well behaved grove (i.e., ``expression
+forest'')
+
+soundness: if you start with a valid graph, you get a well behaved grove that
+connects every vertex (and only those vertices) in the graph. (i.e., if you turn
+it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
+Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
+add Root sort
+
+OR:
+
+use edge IDs instead of edges and build a mapping from vertices to
+sorts---change the $\e$s to $u$s)
+
+\begin{figure}
+  \[
+    \arraycolsep=0pt
+    \begin{array}{lrlll}
+      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Con}(v, m) \mid \textrm{Ref}(v) \\
+      \textrm{Maps:}         & m & {}\in Map & {}::={} & [\,] \mid (p \mapsto [(\e, t), \ldots, (\e, t)]); m \\
+    \end{array}
+  \]
+  \caption{Syntax of graph decomposition trees as a grammar.}
+  \label{fig:Syntax of graph decomposition trees}
+\end{figure}
 

### 2021-01-08 grove-paper.tex
Add notes about related work

@@
 
+We are using action based instead of tree diff
+
+We have to deal with merge
+
+We have to deal with edits from multiple people
+
 J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
@@
 
-CRDTs
+Collaborative Structure Editing
+
+SmallTalk collaboration with images
+
+TouchDevelop papers
+
+\subsection{CRDTs}
+(Are we a known CRDT?)
+
+Bottom = Tombstone
+
+https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
+ - Add-Remove Partial Order data type
+ - 2P2P-Sets
+ - Replicated Growable Array
+
+ commutative replicated data types CmRDT
+ convergent replicated data types, or CvRDTs
+ Delta state CRDTs[12][13] (or simply Delta CRDTs
+
+\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
+ - LWW-Element-Set (Last-Write-Wins-Element-Set)
+ - OR-Set (Observed-Remove Set)
+
+PN-Set
+
+https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
+ removeVertex() has priority, all incident edges are removed
+ addEdge() has priority, all removed vertices are re-added
+ Delay removeVertex() execution till all concurrent removeVertex() are executed.
+First one is 2P2P-Set
+
+textcomp Genealogical Symbols
+\textborn
+\textdied
+
+2P-Sets
+Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
 

### 2021-01-12 grove-paper.tex
Add related work for CRDTs

@@
 
+Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
+
 \subsection{CRDTs}
@@
 
+\url{https://github.com/PsychoLlama/graph-crdt}
+ - Graph CRDT
+ - Uses a LWW-E-Set
+
+\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
+ We don't have interleaving problems because
+   - everything is relative to a specific ID not a position
+   - we don't try to auto resolve
+   - we are tree not list
+   Part three: moving sub tree
+   - Last parent writter wins (prevents cycles)
+     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
+
+...
  commutative replicated data types CmRDT
@@
 
+The Causal Graph CRDT for Complex Document Structure
+
 \url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
@@
 
+G-Set
+2P-Set
 PN-Set
+LWW-element-Register
+OR-Set
+MV-Register: Multi-Value Register
+U-Set (this is what we are for edges, not OR-set due to duplicates)
+Add-Remove Partial Order data type
+  2P-Set for vertices, and a G-Set for edges.
+
+A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
+  \url{https://hal.inria.fr/inria-00555588/document}
+
+CRDTs: Consistency without concurrency control
+  \url{https://arxiv.org/abs/0907.0929}
 
@@
 
-textcomp Genealogical Symbols
-\textborn
-\textdied
+https://crdt.tech/papers.html
+[35]
+    Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
+Co-design and verification of an available file system.
+In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
+Springer LNCS volume 10747, January 2018.
+[ bib | DOI | .pdf ]
+
+        Keywords: trees
+
+[41]
+    Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
+
+        Keywords: trees
+
+[67]
+    Vinh Tao, Marc Shapiro, and Vianney Rancurel.
+Merging semantics for conflict updates in geo-distributed file systems.
+In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
+ACM, May 2015.
+[ bib | DOI | .pdf ]
+
+        Keywords: trees
+
+[95]
+    Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
+
+        Keywords: trees
+
+[103]
+    Stéphane Martin, Pascal Urso, and Stéphane Weiss.
+Scalable XML collaborative editing with undo.
+In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
+Springer LNCS volume 6426, October 2010.
+[ bib | DOI | arXiv ]
+
+        Keywords: trees, undo
 

### 2021-01-13 grove-paper.tex
add note on theorem placement
update grammar and arity table

@@
 \def\ExpPlus{\mathsf{Exp\_plus}}
+\def\ExpTimes{\mathsf{Exp\_times}}
 \def\PatVar{\mathsf{Pat\_var}}
@@
     \begin{array}{lrlll}
-      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid \ast{} \\
-      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
-      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
-                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
-      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
+      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \\
+      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \\
+      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
+                                  &      &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \\
     \end{array}
@@
 The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
-standard for a simply-typed lambda calculus except that all types of terms have
-unfilled holes (represented by $\hole$), and most types of terms have conflict holes (of the form $\{\cdot | \ldots | \cdot\}$).
-In particular, the argument to a lambda can be either a bound variable or a hole.
+standard for a simply-typed lambda calculus except that all sorts have empty
+holes (represented by $\hole$) and conflict holes (of the form $\{q | \ldots |
+q\}$). In particular, the argument to a lambda can be either a bound variable or
+a hole. All constructors carry a superscripted identifier $u \in \U$.
 
@@
       \hline
-      \arity(k_0)={}        & \left\{ \Root \right\}                         \\
-      \arity(x \in Pat)={}      & \left\{ \right\}                               \\
-      \arity(x \in Exp)={}      & \left\{ \right\}                               \\
-      \arity(\lambda q \mathord{:}\tau.e)={}      & \left\{ \mathtt{Param}, \mathtt{Type}, \mathtt{Body} \right\} \\
-      \arity(e~e)={}      & \left\{ \mathtt{Fun}, \mathtt{Arg} \right\}              \\
-      \arity(e~op~e)={}     & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
-      \arity(n)={}      & \left\{ \right\}                               \\
-      \arity(\tau \rightarrow \tau)={}    & \left\{ \mathtt{Arg}, \mathtt{Result} \right\}       \\
-      \arity(Num)={}      & \left\{ \right\}                               \\
+      \arity(\Root)={}        & \left\{ \Root \right\}                         \\
+      \arity(\PatVar)={}      & \left\{ \right\}                               \\
+      \arity(\ExpVar)={}      & \left\{ \right\}                               \\
+      \arity(\ExpLam)={}      & \left\{ \mathtt{Param}, \mathtt{Type}, \mathtt{Body} \right\} \\
+      \arity(\ExpApp)={}      & \left\{ \mathtt{Fun}, \mathtt{Arg} \right\}              \\
+      \arity(\ExpPlus)={}     & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
+      \arity(\ExpTimes)={}    & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
+      \arity(\ExpNum)={}      & \left\{ \right\}                               \\
+      \arity(\TypArrow)={}    & \left\{ \mathtt{Arg}, \mathtt{Result} \right\}       \\
+      \arity(\TypNum)={}      & \left\{ \right\}                               \\
     \end{array}
@@
 position, and edges in position maps all start at the given constructor.
+
+Theorem: well-sorted graphs produce decompositions of well-sorted trees
 

### 2021-01-14 grove-paper.tex
section 3 updates with Cyrus

@@
     \begin{array}{lrlll}
-      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \\
-      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \\
-      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
-                                  &      &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \\
+      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \#u \\
+      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \#u \\
+      e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
+           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \#u \\
     \end{array}
   \]
-  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
-  % TODO add references into multiparent store and cycles store
-  \label{fig:Syntax as a grammar}
+  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
+  \label{fig:Base syntax}
 \end{figure}
@@
 
-The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
-standard for a simply-typed lambda calculus except that all sorts have empty
-holes (represented by $\hole$) and conflict holes (of the form $\{q | \ldots |
-q\}$). In particular, the argument to a lambda can be either a bound variable or
-a hole. All constructors carry a superscripted identifier $u \in \U$.
-
-% In order to model this we define
-
-% Turn that into a constructor set K + position set P + arity function (has type: K -> $\wp$(P)).
+The language we are modeling is shown in \autoref{fig:Base syntax}, standard for
+a simply-typed lambda calculus except that all sorts have empty holes
+(represented by $\hole$) and conflict holes (of the form $\{q | \ldots | q\}$).
+In particular, the argument of a lambda can be either a bound variable or a
+hole. Additionally, all non-hole forms carry a superscripted identifier $u \in
+\U$.
+
+% Turn that into a constructor set K + position set P + arity function (has
+% type: K -> $\wp$(P)).
 
@@
 
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts, and multiparents
-% NOTE ref = expression reference or vertex reference
-
 \subsection{Convergent Graphs}%
@@
 
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
-and connects the child position $p$ of source vertex $v$ to target vertex $v^\prime$.
-The state of each edge, $\e$, is determined by the corresponding edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
-When $G(\e) = \bot$, $\e$ has not been created yet.
-When $G(\e)$ is $\Plus$, the edge has been created.
-When $G(\e)$ is $\Minus$, the edge has been destroyed.
-The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
+Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
+the child position $p$ of originating vertex $v$ to destination vertex
+$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
+edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
+$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
+created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
+ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
 
@@
 
-% XXX
+% TODO: clean up and move rooted expression grammar
+
+% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
+% and multiparents
+
+% TODO add references into multiparent store and cycles store
 
@@
 
-declarative section: give a hint what decomposition does (incl. wreaths): take a
-graph, turn it into a grove (i.e., 4-tuple) which is what we actually present to
-users and use for things like type checking. In this section, we present an
-algorithm that declaratively say what a grove is and how it connects to the
-graph.
-
-We can't turn the graph into a single tree, so we turn it into a forest. A
-forest is a collection of trees, in fact it's a 4-tuple consisting of a rooted tree
-and 3 sets of trees: mp, d, sc. Before discussing what these sets are, let's 
-define what a tree is.
+% NOTE: this whole section is for describing the 4-tuple
+
+Tree-like structures are often easier to work with than arbitrary graphs
+directly, for example to present the graph to users, or for type checking, so we
+decompose it into a grove.
+
+A grove is a 4-tuple of the following form:
+\[
+  (e, MP, NP, W)
+\]
+The first component is an expression, governed by the grammar in
+\autoref{fig:Base syntax}, determined by starting at the root of the graph and
+including every connected vertex that has a single parent. In positions where
+there are no children, we leave an empty hole, $\hole$. In positions where there
+are multiple children, we leave a conflict hole of the form $\{e_1 | \ldots |
+e_n\}$. Whenever we encounter a vertex with multiple parents, we refer to it
+indirectly by reference $\#u$. We collect the expressions corresponding to
+vertices with multiple parents in the set $MP$.
+
+These first two components do not account for vertices with no parents. We
+include the expressions corresponding to these vertices in the set $NP$.
+
+We've accounted for vertices with multiple parents, with no parents, and with
+one parent connected to the root. This leaves only vertices with one parent that
+are not connected to the root. These vertices form structures we call wreaths.
+
+Wreaths:
+
+Three components are  <describe each of the components:
+The first component is ... trees. The fourth component is ... wreaths>.
+
+A forest is a collection of trees, in fact it is a 4-tuple consisting of
+a rooted tree $R$ and 3 sets of trees: $MP, D, W$. Before discussing what these
+sets are, let's define what a tree is.
 

### 2021-01-14 grove-paper.tex
add sorts to arity function results
Added notes and links to CRDT related work

@@
     \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P)                                        \\
+      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
       \hline
@@
       \arity(\ExpVar)={} & \left\{ \right\} \\
-      \arity(\ExpLam)={}      & \left\{ \mathtt{Param}, \mathtt{Type}, \mathtt{Body} \right\} \\
-      \arity(\ExpApp)={}      & \left\{ \mathtt{Fun}, \mathtt{Arg} \right\}              \\
-      \arity(\ExpPlus)={}     & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
-      \arity(\ExpTimes)={}    & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
+      \arity(\ExpLam)={} & \left\{ (\mathtt{Param}, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
+      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
+      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
+      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
       \arity(\ExpNum)={} & \left\{ \right\} \\
-      \arity(\TypArrow)={}    & \left\{ \mathtt{Arg}, \mathtt{Result} \right\}       \\
+      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg} Typ), (\mathtt{Result}, Typ) \right\} \\
       \arity(\TypNum)={} & \left\{ \right\} \\
@@
 
+A commutative replicated data type for cooperative editing
+  \url{https://hal.inria.fr/inria-00445975/document}
+  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
+
+Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
+  \url{}
+
+Specification and Complexity of Collaborative Text Editing
+  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
+
 ...
@@
 The Causal Graph CRDT for Complex Document Structure
+  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
 
 \url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
- - LWW-Element-Set (Last-Write-Wins-Element-Set)
- - OR-Set (Observed-Remove Set)
 
 G-Set
+PN-Set
 2P-Set
-PN-Set
-LWW-element-Register
+LWW-element-Set (Last-Write-Wins)
 OR-Set
@@
 https://crdt.tech/papers.html
-[35]
+
 Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
@@
 [ bib | DOI | .pdf ]
-
-        Keywords: trees
-
-[41]
+  http://dx.doi.org/10.1007/978-3-319-73721-8_17
+  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
+
 Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-
-        Keywords: trees
-
-[67]
+  http://dx.doi.org/10.1109/TPDS.2017.2697382
+  http://arxiv.org/abs/1608.03960
+
 Vinh Tao, Marc Shapiro, and Vianney Rancurel.
@@
 [ bib | DOI | .pdf ]
-
-        Keywords: trees
-
-[95]
+  http://dx.doi.org/10.1145/2757667.2757683
+  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
+
 Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-
-        Keywords: trees
-
-[103]
+  http://arxiv.org/abs/1207.5990
+  https://hal.inria.fr/hal-00720681/
+
 Stéphane Martin, Pascal Urso, and Stéphane Weiss.
@@
 [ bib | DOI | arXiv ]
-
-        Keywords: trees, undo
+  http://dx.doi.org/10.1007/978-3-642-16934-2_37
+  http://arxiv.org/abs/1010.3615
 

### 2021-01-14 grove-paper.tex
Add more CRDT notes to paper

@@
 
+List of CRTD papers: \url{https://crdt.tech/papers.html}
+
 Bottom = Tombstone
@@
 Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{}
+  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
 
@@
   \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
+
+LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
+  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
+
+Data consistency for P2P collaborative editing
+  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
+
+Interleaving anomalies in collaborative text editors
+  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
+
+Moving Elements in List CRDTs
+  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
+
+A highly-available move operation for replicated trees and distributed filesystems
+  \url{https://martin.kleppmann.com/papers/move-op.pdf}
 
@@
 [ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1007/978-3-319-73721-8_17
+  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
   https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
@@
 [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1007/978-3-642-16934-2_37
+  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
   http://arxiv.org/abs/1010.3615

### 2021-01-21 grove-paper.tex
fleshing out decomposition algorithm

--- 2021-01-14 grove-paper.tex
@@
 In particular, the argument of a lambda can be either a bound variable or a
-hole. Additionally, all non-hole forms carry a superscripted identifier $u \in
+hole. Additionally, all non-hole forms have a superscripted identifier $u \in
 \U$.
@@
       \arity(\ExpVar)={} & \left\{ \right\} \\
-      \arity(\ExpLam)={}      & \left\{ \mathtt{Param}, \mathtt{Type}, \mathtt{Body} \right\} \\
-      \arity(\ExpApp)={}      & \left\{ \mathtt{Fun}, \mathtt{Arg} \right\}              \\
-      \arity(\ExpPlus)={}     & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
-      \arity(\ExpTimes)={}    & \left\{ \mathtt{L}, \mathtt{R} \right\}         \\
+      \arity(\ExpLam)={} & \left\{ (\mathtt{Param}, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
+      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
+      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
+      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
       \arity(\ExpNum)={} & \left\{ \right\} \\
-      \arity(\TypArrow)={}    & \left\{ \mathtt{Arg}, \mathtt{Result} \right\}       \\
+      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
       \arity(\TypNum)={} & \left\{ \right\} \\
@@
 A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a form defined in \autoref{fig:Syntax as a grammar} that
-is neither an operator nor a hole. Each contructor has an arity, which we define
-as a set of positions. A position $p \in \P$ identifies a location, adjacent to
-some source vertex, from which edges to target vertices may originate. The
-$\arity$ function in \autoref{fig:Constructors, Indexes and Arity} gives the
-complete mapping from constructors to potential edge positions. Every graph
-contains a distinguished root vertex, root constructor, and root position,
+\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
+contructor has an arity, which we define as a set of positions. A position $p
+\in \P$ identifies a location at some originating vertex from which edges to
+destination vertices may originate. The $\arity$ function in
+\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
+constructors to potential edge positions with their corresponding sorts. Every
+graph contains a distinguished root vertex, root constructor, and root position,
 respectively named $v_0$, $k_0$, and $p_0$.
@@
 constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term, starting with the
-ones at positions in $\arity(k')$, as if it were the outermost and create edges
-to the vertices representing its immediate sub-terms until only empty holes
+= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
+outermost, starting with the ones at positions in $\arity(k')$, and create edges
+to the vertices representing their immediate sub-terms until only empty holes
 remain. Note that empty holes are not modeled explicitly in the graph. Wherever
@@
 
-% TODO: clean up and move rooted expression grammar
-
 % NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
@@
 
-(illustrative running example: 1 of each: mp, d, r, sc)
-
 % NOTE: this whole section is for describing the 4-tuple
 
-Tree-like structures are often easier to work with than arbitrary graphs
-directly, for example to present the graph to users, or for type checking, so we
-decompose it into a grove.
+Trees are often easier to work with than arbitrary graphs directly, for example
+to present the graph to users or for type checking. A graph is not necessarily a
+tree, but we can decompose any graph into a collection of tree-like structure we
+call a grove.
 
@@
 \]
-The first component is an expression, governed by the grammar in
-\autoref{fig:Base syntax}, determined by starting at the root of the graph and
-including every connected vertex that has a single parent. In positions where
-there are no children, we leave an empty hole, $\hole$. In positions where there
-are multiple children, we leave a conflict hole of the form $\{e_1 | \ldots |
-e_n\}$. Whenever we encounter a vertex with multiple parents, we refer to it
-indirectly by reference $\#u$. We collect the expressions corresponding to
-vertices with multiple parents in the set $MP$.
+The first component $e$ is an expression as governed by the grammar in
+\autoref{fig:Base syntax}. We determine $e$ by starting at the root of the graph
+and including every connected vertex that has a single parent. In positions
+where there are no children, we leave an empty hole $\hole$. In positions where
+there are multiple children, we leave a conflict hole of the form $\{e_1 |
+\ldots | e_n\}$. Whenever we encounter a vertex with multiple parents, we leave
+an indirect reference $\#u$, where $u$ is the unique identifier of the
+referenced vertex, and add its corresponding expression to the second component
+$MP$.
 
 These first two components do not account for vertices with no parents. We
-include the expressions corresponding to these vertices in the set $NP$.
-
-We've accounted for vertices with multiple parents, with no parents, and with
-one parent connected to the root. This leaves only vertices with one parent that
-are not connected to the root. These vertices form structures we call wreaths.
-
-Wreaths:
-
-Three components are  <describe each of the components:
-The first component is ... trees. The fourth component is ... wreaths>.
-
-A forest is a collection of trees, in fact it is a 4-tuple consisting of
-a rooted tree $R$ and 3 sets of trees: $MP, D, W$. Before discussing what these
-sets are, let's define what a tree is.
-
-here's what a tree is. why do we care what a tree is? because we can map them to
-the expressions in figure 11. Explain what a Ref is wrt forest roots. (Maybe
-mention that only well-sorted trees can be mapped, which we discuss below.)
-
-There's this weird stuff left over; we call them wreaths and put them in the 4th
-tuple component.
-
-here's what a wreath is (= a connected graph where each vertex has exactly one
-parent)
-
-(show what a wreath looks like)
-
-here's what a grove is
-
-here are some properties that show how to get one from the other, and to/from
-graph
+include the expressions corresponding to these vertices in the third component
+$NP$.
+
+We've accounted for single-parented vertices connected to the root,
+single-parented vertices connected to multi-parented vertices, and vertices with
+no parents. This leaves only single-parented vertices that are not connected to
+the root or any multi-parented vertices. These vertices form cyclic structures
+we call wreaths.
+
+A wreath is a cycle, potentially with acyclic branches. Wreaths are connected
+graphs where each vertex has exactly one incoming edge. The fourth component $W$
+is the set of expressions determined by extracting wreaths from any vertices not
+in the other three components.
+
+% expression = tree
+
+% (show what a wreath looks like)
+
+% here are some properties that show how to get one from the other, and to/from
+% graph
 

### 2021-01-21 grove-paper.tex

@@
 
+A commutative replicated data type for cooperative editing
+  \url{https://hal.inria.fr/inria-00445975/document}
+  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
+
+Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
+  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
+
+Specification and Complexity of Collaborative Text Editing
+  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
+
+LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
+  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
+
+Data consistency for P2P collaborative editing
+  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
+
+Interleaving anomalies in collaborative text editors
+  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
+
+Moving Elements in List CRDTs
+  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
+
+A highly-available move operation for replicated trees and distributed filesystems
+  \url{https://martin.kleppmann.com/papers/move-op.pdf}
+
 ...
@@
 [ bib | DOI | .pdf ]
-
-        Keywords: trees
-
-[41]
+  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
+  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
+
 Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-
-        Keywords: trees
-
-[67]
+  http://dx.doi.org/10.1109/TPDS.2017.2697382
+  http://arxiv.org/abs/1608.03960
+
 Vinh Tao, Marc Shapiro, and Vianney Rancurel.
@@
 [ bib | DOI | arXiv ]
-
-        Keywords: trees, undo
+  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
+  http://arxiv.org/abs/1010.3615
 

### 2021-01-25 grove-paper.tex
updates from Grove meeting

@@
 to present the graph to users or for type checking. A graph is not necessarily a
-tree, but we can decompose any graph into a collection of tree-like structure we
-call a grove.
+tree, but we can decompose any graph into a collection of expressions, governed
+by the grammar in \autoref{fig:Base syntax}, organized into a structure we call
+a grove.
 
@@
 \]
-The first component $e$ is an expression as governed by the grammar in
-\autoref{fig:Base syntax}. We determine $e$ by starting at the root of the graph
-and including every connected vertex that has a single parent. In positions
-where there are no children, we leave an empty hole $\hole$. In positions where
-there are multiple children, we leave a conflict hole of the form $\{e_1 |
-\ldots | e_n\}$. Whenever we encounter a vertex with multiple parents, we leave
-an indirect reference $\#u$, where $u$ is the unique identifier of the
-referenced vertex, and add its corresponding expression to the second component
-$MP$.
+The first component $e$ is the rooted expression.
+The second component $MP$ is the set of expressions rooted at a vertex with multiple parents.
+The third component $NP$ is the set of expressions rooted at a vertex with no parents.
+The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily 
+form cyclic structures we call wreaths, where each vertex has exactly one parent.
+For example, the grove corresponding to the graph in Figure~XXX is ...
+% TODO: add ref to example figure
+
+% TODO: define helper function for finding the expression rooted at some vertex
+
+We determine $e$ by starting at the root of the graph and including every connected
+vertex that has a single parent. In positions where there are no children, we
+leave an empty hole $\hole$. In positions where there are multiple children, we
+leave a conflict hole of the form $\{e_1 | \ldots | e_n\}$. Whenever we
+encounter a vertex with multiple parents, we add its corresponding expression to
+the second component $MP$ and leave an indirect reference $\#u$, where $u$ is
+the unique identifier of the referenced vertex
 
@@
 
-A wreath is a cycle, potentially with acyclic branches. Wreaths are connected
-graphs where each vertex has exactly one incoming edge. The fourth component $W$
-is the set of expressions determined by extracting wreaths from any vertices not
-in the other three components.
+% A wreath is a cycle, potentially with acyclic branches. Wreaths are connected
+% graphs where each vertex has exactly one incoming edge. The fourth component $W$
+% is the set of expressions determined by extracting wreaths from any vertices not
+% in the other three components.
 

### 2021-02-01 grove-paper.tex
updates from last two Grove meetings

@@
 % graph decomposition operators
+\DeclareMathOperator{\expr}{\text{expr}}
+\DeclareMathOperator{\patt}{\text{patt}}
+\DeclareMathOperator{\type}{\text{type}}
 \DeclareMathOperator{\parents}{\text{parents}}
+\DeclareMathOperator{\ancestors}{\text{ancestors}}
+\DeclareMathOperator{\vertices}{\text{vertices}}
+\DeclareMathOperator{\children}{\text{children}}
+
 \DeclareMathOperator{\multiParented}{\text{multiParented}}
@@
 where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$,
+unique IDs are drawn from some suitable set $\U$ equipped with a total ordering,
 vertices are drawn from $\V$,
@@
 tree, but we can decompose any graph into a collection of expressions, governed
-by the grammar in \autoref{fig:Base syntax}, organized into a structure we call
-a grove.
+by the grammar in \autoref{fig:Base syntax}, which we organize into a structure
+we call a grove.
 
@@
 \[
-  (e, MP, NP, W)
+  (e_r, MP, NP, W)
 \]
-The first component $e$ is the rooted expression.
+where
+\begin{align*}
+  e_r &= \expr(v_0) \\
+  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
+  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
+  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
+\end{align*}
+
+The first component $e_r$ is the expression rooted at the root vertex $v_0$.
 The second component $MP$ is the set of expressions rooted at a vertex with multiple parents.
 The third component $NP$ is the set of expressions rooted at a vertex with no parents.
-The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily 
-form cyclic structures we call wreaths, where each vertex has exactly one parent.
+The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent.
+In each of these, $\parents(v)$ is defined as follows.
+\[ \parents(v) = \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\}  \] \\
+
 For example, the grove corresponding to the graph in Figure~XXX is ...
-% TODO: add ref to example figure
-
-% TODO: define helper function for finding the expression rooted at some vertex
+
+\begin{figure}[b]
+  \centering
+  \begin{tikzpicture}
+    \path (0cm,0cm) graph[graph style] {
+      root[root style] -> {
+       a/{\vWrapPlus} [> "\eWrapPlus"] -> {
+        "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
+         {},
+         b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
+        },
+        {}
+       }
+      },
+    };
+    \path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
+    \path (1cm,-1cm) graph[graph style] { plus/"\vSimpleTimes" };
+    \path (2cm,-1cm) graph[graph style] { times/"\vMultiCycleTimes" };
+    \path (2cm,-2cm) graph[graph style] { plus2/"\vSimpleTimes" };
+    \path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
+    \path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
+    \path[draw,->,merge edge] (times) to ["\eMultiCycleBobPlus"anchor=west,out=-90,in=90] (plus2);
+    \path (1.5cm,-2.5cm) graph[graph style] { plus/"\vSimpleTimes" };
+  \end{tikzpicture}
+  \caption{Caption goes here}%
+  \label{fig:Decomposition example}
+\end{figure}
+
+Figure YYY defines $\expr(v)$ wich determines the expression for a given vertex $v$.
+There's one rule for each constructor form, e.g., $\ExpApp$.
+In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).
+
+% TODO: add conflict hole to no-parent vertex
+
+\begin{figure*}
+\begin{align*}
+  e_r &= \expr(v_0) \\
+  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
+  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
+  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
+\end{align*}
+
+\begin{align*}
+  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
+  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
+\end{align*}
+
+\begin{displaymath}
+\begin{array}{l}
+  \expr(v=(u, Root)) = expr'(v, Root) \\
+  \expr(v=(u, \ExpVar(x))) = x^u \\
+  \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
+  \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
+\end{array}
+\end{displaymath}
+
+% \begin{align*}
+%   \expr(v=(u, Root)) &= expr'(v, Root) \\
+%   \expr(v=(u, \ExpVar(x))) &= x^u \\
+%   \expr(v=(u, \ExpApp)) &= (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
+%   % | Exp_var of (string)
+%   % | Exp_lam
+%   % | Exp_app
+%   % | Exp_num of (int)
+%   % | Exp_plus
+%   % | Exp_times
+%   % | Exp_nil
+%   % | Exp_cons
+%   % | Exp_case
+%   % (**** Pat ****)
+%   % | Pat_var of (string)
+%   % (**** Typ ****)
+%   % | Typ_num
+%   % | Typ_arrow
+%   % | Typ_list
+% \end{align*}
+
+% DONE: add ref case to expr'
+% TODO: point out cleverness in W set-builder def -- say that v being the min ancestor of itself means it is part of the cycle
+% TODO: note to take least fixed point of ancestors
+
+% Michael's suggestion:
+%   f : A -> A U parents(v) U ancestors(v)
+%   take lfp of f
+
+% when talking about why W is defined that way: after explaining intuition, say
+% ``we can formalize this intuition by taking the lfp of <the function f>.''
+% fixed-point theory is part of the set of assumptions in this paper.
+
+\begin{align*}
+  \expr'(v,p) = \begin{cases}
+    \hole & \children(v,p) = \varnothing \\
+    \expr(v') & \children(v,p) = \{v'\} \\
+    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| > 1 \\
+    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \land v' = \min(\ancestors(v')) \\
+    \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+  \end{cases}
+\end{align*}
+
+\begin{align*}
+  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
+\end{align*}
+\end{figure*}
+
+% TODO: add text around these definitions -- turn them into paragraphs.
 

### 2021-02-02 grove-paper.tex
clean up stale comments
switch to single-column format

@@
 %\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 \newedge{WrapTimes}{L}
-\newcommand{\figureWrap}{
+\newcommand{\figureWrapMove}{
 \begin{figure}
+\centering
+\begin{minipage}[t]{.45\linewidth}
 \hskip0.12\columnwidth
@@
 \label{fig:Wrap}
-\end{figure}
-}
-
+\end{minipage}
+% \end{figure}
+% }
+%
+\hfil
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newedge{MoveTimes}{R}
-\newcommand{\figureMove}{
-\begin{figure}
+% \newcommand{\figureMove}{
+% \begin{figure}[H]
+\begin{minipage}[t]{.45\linewidth}
 \hfill
@@
 \label{fig:Move}
+\end{minipage}
 \end{figure}
@@
 \newvertex{DifferentPartsBob}{v}
-\newcommand{\figureDifferentParts}{
+% \newcommand{\figureDifferentParts}{
+\newcommand{\figureDifferentPartsNestedParts}{
 \begin{figure}
+\begin{minipage}[t]{0.45\linewidth}
 %\hfill
@@
 \label{fig:DifferentParts}
-\end{figure}
-}
-
+\end{minipage}
+% \end{figure}
+% }
+%
+\hfil
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % }
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newedge{NestedPartsBob}{L}
-\newcommand{\figureNestedParts}{
-\begin{figure}
+% \newcommand{\figureNestedParts}{
+% \begin{figure}
+\begin{minipage}[t]{0.45\linewidth}
 \hfill
@@
 \label{fig:NestedParts}
+\end{minipage}
 \end{figure}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.23\textwidth}
+\begin{subfigure}[t]{0.2\textwidth}
 \centering
@@
 
-\figureWrap{}
+\figureWrapMove{}
+% \figureMove{}
 
@@
 
-\figureMove{}
+% \figureMove{}
 
@@
 
-\figureDifferentParts{}
+% \figureDifferentParts{}
+\figureDifferentPartsNestedParts{}
 
@@
 
-\figureNestedParts{}
+% \figureNestedParts{}
 
@@
 
-\begin{figure}[b]
+\begin{figure}
   \centering
@@
   \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
+  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
 \end{align*}
@@
 
-% DONE: add ref case to expr'
-% TODO: point out cleverness in W set-builder def -- say that v being the min ancestor of itself means it is part of the cycle
-% TODO: note to take least fixed point of ancestors
-
-% Michael's suggestion:
-%   f : A -> A U parents(v) U ancestors(v)
-%   take lfp of f
-
-% when talking about why W is defined that way: after explaining intuition, say
-% ``we can formalize this intuition by taking the lfp of <the function f>.''
-% fixed-point theory is part of the set of assumptions in this paper.
-
 \begin{align*}
@@
 
-\begin{align*}
-  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
-\end{align*}
 \end{figure*}

### 2021-02-02 grove-paper.tex
add constructor args to constructor-arity-sort map
clean up decomp example diagram

@@
       \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar)={} & \left\{ \right\} \\
-      \arity(\ExpVar)={} & \left\{ \right\} \\
+      \arity(\PatVar(x))={} & \left\{ \right\} \\
+      \arity(\ExpVar(x))={} & \left\{ \right\} \\
       \arity(\ExpLam)={} & \left\{ (\mathtt{Param}, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
@@
       \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum)={} & \left\{ \right\} \\
+      \arity(\ExpNum(n))={} & \left\{ \right\} \\
       \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
@@
 
+\newvertex{DecompPlusA}{+}
+\newedge{DecompTimesA}{L}
+\newvertex{DecompTimesA}{*}
+\newedge{DecompPlusAX}{R}
+\newvertex{DecompX}{x}
+\newedge{DecompTimesAX}{R}
+%
+\newedge{DecompPlusB}{L}
+\newvertex{DecompPlusB}{+}
+\newedge{DecompTimesB}{R}
+\newvertex{DecompTimesB}{*}
+\newedge{DecompY}{R}
+\newvertex{DecompY}{y}
+\newedge{DecompZ}{R}
+\newvertex{DecompZ}{z}
+%
+\newedge{DecompTimes}{L}
 \begin{figure}
@@
       root[root style] -> {
-       a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-        "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
+       a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
+        "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
          {},
-         b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
+         b/"\vDecompX" [> "\eDecompTimesAX"']
         },
@@
     };
-    \path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-    \path (1cm,-1cm) graph[graph style] { plus/"\vSimpleTimes" };
-    \path (2cm,-1cm) graph[graph style] { times/"\vMultiCycleTimes" };
-    \path (2cm,-2cm) graph[graph style] { plus2/"\vSimpleTimes" };
-    \path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-    \path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-    \path[draw,->,merge edge] (times) to ["\eMultiCycleBobPlus"anchor=west,out=-90,in=90] (plus2);
-    \path (1.5cm,-2.5cm) graph[graph style] { plus/"\vSimpleTimes" };
+    \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
+    %
+    \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
+    \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
+    \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
+    \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
+    \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
+    \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
+    %
+    \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
   \end{tikzpicture}

### 2021-02-02 grove-paper.tex
updates from call w/ Cyrus

@@
     \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \#u \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \#u \\
+      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \multiVertex{u} \mid \cycleVertex{u} \\
       e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
-           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \#u \\
+           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
     \end{array}
@@
 
-% TODO add references into multiparent store and cycles store
-
 % NOTE: this whole section is for describing the 4-tuple
@@
 we call a grove.
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-
-The first component $e_r$ is the expression rooted at the root vertex $v_0$.
-The second component $MP$ is the set of expressions rooted at a vertex with multiple parents.
-The third component $NP$ is the set of expressions rooted at a vertex with no parents.
-The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent.
-In each of these, $\parents(v)$ is defined as follows.
-\[ \parents(v) = \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\}  \] \\
-
-For example, the grove corresponding to the graph in Figure~XXX is ...
 
@@
 
-Figure YYY defines $\expr(v)$ wich determines the expression for a given vertex $v$.
-There's one rule for each constructor form, e.g., $\ExpApp$.
-In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).
-
-% TODO: add conflict hole to no-parent vertex
-
-\begin{figure*}
+% \newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
+% \newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\rcirclearrowleft_{#1}}}}}
+
+A grove is a 4-tuple of the following form:
+\[
+  (e_r, MP, NP, W)
+\]
+where
 \begin{align*}
@@
 \end{align*}
-
+For example, the grove corresponding to the graph in \autoref{fig:Decomposition example} is
 \begin{align*}
-  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
-  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
+  e_r &= \hole~\texttt{*}^{46}~\#48~\texttt{+}^{44}~\#48 \\
+  MP &= \{x^{48}\} \\
+  NP &= \{z^{58}\} \\
+  W &= \{\hole~\texttt{+}^{52}~\#52~\texttt{*}^{54}~y^{56}\} \\
 \end{align*}
 
+Let us now examine each of the four components in more detail.
+The first component $e_r$ is the expression derived from the root vertex $v_0$.
+\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
+There's one rule for each constructor form, e.g., $\ExpApp$.
+In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).
+
+The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
+Wherever an edge to such a vertex appears, we place a reference to it $\#u$.
+In the example, $x^{48}$ is referenced twice in $e_r$.
+
+The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
+In the example, $z^{58}$ has no parents (because it was deleted), so it appears only in $NP$.
+
+The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent.
+
+In $W$, $\ancestors(v)$ is the set of vertices for which there is a path to $v$.
+Note that $v = \min(\ancestors(v))$ only if $v$ is part of a cycle.
+
+% TODO: replace # with \multiVertex and \cycleVertex
+% TODO: update \multiChild with more than 2 children and use it for conflict holes
+% TODO: gray out ids in the grammar and expressions
+% TODO: add a child to z^58 and rotate 51-52
+% TODO: move example grove to right of its diagram
+% TODO: distinguish between multi-parent and cycle refs
+
+\begin{figure}
   \begin{displaymath}
@@
   \end{displaymath}
+
+  \begin{align*}
+    \expr'(v,p) = \begin{cases}
+      \hole & \children(v,p) = \varnothing \\
+      \expr(v') & \children(v,p) = \{v'\} \land |\parents(v')| = 1 \\
+                & \phantom{\children(v,p)} \land v' \ne \min(\ancestors(v')) \\
+      \#u & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
+      \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
+          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
+      \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+    \end{cases}
+  \end{align*}
+
+  \begin{align*}
+    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
+    \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v)) \\
+    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
+    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
+      \end{align*}
+  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of the equation above.}
+  \label{fig:Decomposition expr}
+\end{figure}
+
+% TODO: add conflict hole to no-parent vertex
+
+\begin{figure}[b]
+\begin{align*}
+  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
+  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
+  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
+\end{align*}
 
@@
 
-\begin{align*}
-  \expr'(v,p) = \begin{cases}
-    \hole & \children(v,p) = \varnothing \\
-    \expr(v') & \children(v,p) = \{v'\} \\
-    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| > 1 \\
-    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \land v' = \min(\ancestors(v')) \\
-    \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-  \end{cases}
-\end{align*}
-
-\end{figure*}
+\end{figure}
 

### 2021-02-10 grove-paper.tex
updates from last Grove meeting
updates with Cyrus

@@
 \def\PlusRight{\mathsf{Plus\_right}}
+\def\TimesLeft{\mathsf{Times\_left}}
+\def\TimesRight{\mathsf{Times\_right}}
 \def\ArrowArg{\mathsf{Arrow\_arg}}
@@
 
+\newvertex{DecompPlusA}{+}
+\newedge{DecompTimesA}{L}
+\newvertex{DecompTimesA}{*}
+\newedge{DecompPlusAX}{R}
+\newvertex{DecompX}{x}
+\newedge{DecompTimesAX}{R}
+%
+\newedge{DecompPlusB}{L}
+\newvertex{DecompPlusB}{+}
+\newedge{DecompTimesB}{R}
+\newvertex{DecompTimesB}{*}
+\newedge{DecompY}{R}
+\newvertex{DecompY}{y}
+\newedge{DecompZ}{R}
+\newvertex{DecompZ}{z}
+%
+\newedge{DecompTimes}{L}
+%
+\newcommand{\figureDecompExample}{
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \path (0cm,0cm) graph[graph style] {
+      root[root style] -> {
+       a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
+        "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
+         {},
+         b/"\vDecompX" [> "\eDecompTimesAX"']
+        },
+        {}
+       }
+      },
+    };
+    \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
+    %
+    \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
+    \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
+    \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
+    \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
+    \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
+    \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
+    %
+    \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
+  \end{tikzpicture}
+  \caption{Caption goes here}%
+  \label{fig:Decomposition example}
+\end{figure}
+}
+
 \begin{document}
@@
 
+% TODO: replace positions with macros
+
 \begin{figure}
@@
       \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\mathtt{Param}, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
+      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
       \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
@@
 
-Trees are often easier to work with than arbitrary graphs directly, for example
-to present the graph to users or for type checking. A graph is not necessarily a
-tree, but we can decompose any graph into a collection of expressions, governed
-by the grammar in \autoref{fig:Base syntax}, which we organize into a structure
-we call a grove.
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (0cm,0cm) graph[graph style] {
-      root[root style] -> {
-       a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-        "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-         {},
-         b/"\vDecompX" [> "\eDecompTimesAX"']
-        },
-        {}
-       }
-      },
-    };
-    \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-    %
-    \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-    \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-    \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-    \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-    \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-    \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-    %
-    \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-  \end{tikzpicture}
-  \caption{Caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
+Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
+A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
+
+\figureDecompExample
 
@@
 \autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There's one rule for each constructor form, e.g., $\ExpApp$.
-In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).
+There is one rule for each constructor form.
+For example, the expression derived from $v_{38} = (38, \ExpPlus)$ is $\expr'(v_{38}, \PlusLeft) \text{+}^{38} \expr'(v_{38}, \PlusRight)$.
+In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
+An empty hole $\hole$ appears if there are no children at the given position.
+A conflict hole $...$  appears if there are multiple children.
+If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
+For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
 
@@
 % TODO: replace # with \multiVertex and \cycleVertex
+% TODO: distinguish between multi-parent and cycle refs in text
 % TODO: update \multiChild with more than 2 children and use it for conflict holes
@@
 % TODO: move example grove to right of its diagram
-% TODO: distinguish between multi-parent and cycle refs
+% TODO: use mathord for spaces around + / *
+% TODO: figure out why + --> "" and * --> "\times"
+% TODO: keep condensed "|"s for conflict holes in grammar, but not other places
+% TODO: make lfp a math function
 
@@
       \expr(v=(u, \ExpVar(x))) = x^u \\
+      \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
       \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
-      \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
+      \expr(v=(u, \ExpNum(n))) = n^u \\
+      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^u~\expr'(v, \PlusRight) \\
+      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^u~\expr'(v, \TimesRight) \\
+      \patt(v=(u, \PatVar(x))) = x^u \\
+      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow \type'(\tau, \ArrowResult) \\
+      \type(\tau=(u, \TypNum) = Num^u \\
     \end{array}
   \end{displaymath}
+
+% TODO: make macros for + and *
+% TODO: comment these out:
+%   % | Exp_nil
+%   % | Exp_cons
+%   % | Exp_case
+%   % | Typ_list
 
@@
       \hole & \children(v,p) = \varnothing \\
-      \expr(v') & \children(v,p) = \{v'\} \land |\parents(v')| = 1 \\
-                & \phantom{\children(v,p)} \land v' \ne \min(\ancestors(v')) \\
-      \#u & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
+      \{\expr(v_1) \mid \ldots \mid \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
+      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
           & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \expr(v') & \text{otherwise} \\
     \end{cases}
@@
   \begin{align*}
+    \patt'(v,p) = \begin{cases}
+      \hole & \children(v,p) = \varnothing \\
+      \{\patt(v_1) \mid \ldots \mid \patt(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
+      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
+          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
+      \patt(v') & \text{otherwise} \\
+    \end{cases}
+  \end{align*}
+
+  \begin{align*}
+    \type'(v,p) = \begin{cases}
+      \hole & \children(v,p) = \varnothing \\
+      \{\type(v_1) \mid \ldots \mid \type(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
+      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
+          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
+      \type(v') & \text{otherwise} \\
+    \end{cases}
+  \end{align*}
+
+  \begin{align*}
     \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v)) \\
+    \ancestors(v) &= \left( \textrm{lfp}(\ancestors') \right)(v) \\
+    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
     \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
@@
       \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of the equation above.}
+  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of $a$.}
   \label{fig:Decomposition expr}
@@
 \end{align*}
-
-% \begin{align*}
-%   \expr(v=(u, Root)) &= expr'(v, Root) \\
-%   \expr(v=(u, \ExpVar(x))) &= x^u \\
-%   \expr(v=(u, \ExpApp)) &= (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
-%   % | Exp_var of (string)
-%   % | Exp_lam
-%   % | Exp_app
-%   % | Exp_num of (int)
-%   % | Exp_plus
-%   % | Exp_times
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % (**** Pat ****)
-%   % | Pat_var of (string)
-%   % (**** Typ ****)
-%   % | Typ_num
-%   % | Typ_arrow
-%   % | Typ_list
-% \end{align*}
 

### 2021-02-11 grove-paper.tex
add conflictHole macro

@@
 \usepackage{graphicx}
-\usepackage{semantic}
+% \usepackage{semantic}
 \let\colonapprox\undefined % Avoid redefinition error in `colonequals`
@@
 \usepackage{colonequals}
-\usepackage{mathpartir}
+% \usepackage{mathpartir}
 \usepackage{stmaryrd}
@@
 % \selectcolormodel{gray}
-
-% math ligatures
-\mathlig{::=}{\coloncolonequals}
-\mathlig{:=}{\colonequals}
-\mathlig{|-}{\vdash}
-\mathlig{:->}{\mapsto}
-\mathlig{-->a}{\overset{a}{\longrightarrow}}
-\newcommand{\abs}[1]{\left|{#1}\right|}
 
@@
 
+% custom syntax
+
+\usepackage{xstring}
+
 \newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
+\newcommand{\conflictHole}[1]{%
+  {\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]
+  {\textcolor{red}{\textbf{\{}} \myargs \textcolor{red}{\textbf{\}}}}}}
 
@@
 in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\texttt{\{ x | y \}}.
+In the user interface, we flag this conflict with the notation~\conflictHole{x, y}.
 
@@
     \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \conflictHole{q, \ldots, q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \conflictHole{\tau, \ldots, \tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
       e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
-           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \conflictHole{e, \ldots, e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
     \end{array}
@@
 
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for
-a simply-typed lambda calculus except that all sorts have empty holes
-(represented by $\hole$) and conflict holes (of the form $\{q | \ldots | q\}$).
-In particular, the argument of a lambda can be either a bound variable or a
-hole. Additionally, all non-hole forms have a superscripted identifier $u \in
-\U$.
+The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q, \ldots, q}$).
+In particular, the argument of a lambda can be either a bound variable or a hole.
+Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
 
@@
 An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $...$  appears if there are multiple children.
+A conflict hole $\conflictHole{e_1, \ldots, e_n}$ appears if there are multiple children.
 If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
@@
       \hole & \children(v,p) = \varnothing \\
-      \{\expr(v_1) \mid \ldots \mid \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\expr(v_1), \ldots, \expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
       \hole & \children(v,p) = \varnothing \\
-      \{\patt(v_1) \mid \ldots \mid \patt(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\patt(v_1), \ldots, \patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
       \hole & \children(v,p) = \varnothing \\
-      \{\type(v_1) \mid \ldots \mid \type(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\type(v_1), \ldots, \type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
 leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\{e_1 | \ldots | e_n\}$. Whenever we
+leave a conflict hole of the form $\conflictHole{e_1, \ldots, e_n}$. Whenever we
 encounter a vertex with multiple parents, we add its corresponding expression to

### 2021-02-16 grove-paper.tex
replace \multiChild with \conflictHole
tighten up spacing in \conflictHole

@@
 \newcommand{\conflictHole}[1]{%
-  {\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]
-  {\textcolor{red}{\textbf{\{}} \myargs \textcolor{red}{\textbf{\}}}}}}
+{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
+{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
 
@@
 \newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiChild}[2]{\textcolor{red}{\{} #1 \textcolor{red}{|} #2 \textcolor{red}{\}}}
 \newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
@@
 \label{fig:MultiChild:c}
-\figureCode{w * v + \multiChild{x}{y}}%
+\figureCode{w * v + \conflictHole{x,y}}%
 \begin{tikzpicture}
@@
 \label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \multiChild{\multiVertex{30}}{\multiVertex{32}}%
+\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
   \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
@@
 in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x, y}.
+In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
 
@@
     \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \conflictHole{q, \ldots, q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \conflictHole{\tau, \ldots, \tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
       e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
-           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \conflictHole{e, \ldots, e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
     \end{array}
@@
 
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q, \ldots, q}$).
+The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
 In particular, the argument of a lambda can be either a bound variable or a hole.
@@
 An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1, \ldots, e_n}$ appears if there are multiple children.
+A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
 If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
@@
       \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1), \ldots, \expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
       \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1), \ldots, \patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
       \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1), \ldots, \type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
+      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
       \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
@@
 leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1, \ldots, e_n}$. Whenever we
+leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
 encounter a vertex with multiple parents, we add its corresponding expression to

### 2021-02-16 grove-paper.tex
drop old TODO comment block
replace \# with \multiVertex and \cycleVertex

@@
 \begin{align*}
-  e_r &= \hole~\texttt{*}^{46}~\#48~\texttt{+}^{44}~\#48 \\
-  MP &= \{x^{48}\} \\
-  NP &= \{z^{58}\} \\
-  W &= \{\hole~\texttt{+}^{52}~\#52~\texttt{*}^{54}~y^{56}\} \\
+  e_r &= \hole~\texttt{*}^{40}~\multiVertex{42}~\texttt{+}^{38}~\multiVertex{42} \\
+  MP &= \{x^{42}\} \\
+  NP &= \{z^{52}\} \\
+  W &= \{\hole~\texttt{+}^{46}~\cycleVertex{46}~\texttt{*}^{48}~y^{50}\} \\
 \end{align*}
@@
 The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\#u$.
+Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
 In the example, $x^{48}$ is referenced twice in $e_r$.
@@
 Note that $v = \min(\ancestors(v))$ only if $v$ is part of a cycle.
-
-% TODO: replace # with \multiVertex and \cycleVertex
-% TODO: distinguish between multi-parent and cycle refs in text
-% TODO: update \multiChild with more than 2 children and use it for conflict holes
-% TODO: gray out ids in the grammar and expressions
-% TODO: add a child to z^58 and rotate 51-52
-% TODO: move example grove to right of its diagram
-% TODO: use mathord for spaces around + / *
-% TODO: figure out why + --> "" and * --> "\times"
-% TODO: keep condensed "|"s for conflict holes in grammar, but not other places
-% TODO: make lfp a math function
 
@@
 encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\#u$, where $u$ is
+the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
 the unique identifier of the referenced vertex

### 2021-02-16 grove-paper.tex
comment out old equations
add \id macro for grayed UUIDs

@@
 
-\newcommand{\conflict}[2]{\llparenthesis#1|#2\rrparenthesis}
+\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
 
@@
  %draw=none,
- as={\textbullet$_{0}$}
+ as={\textbullet$_{\id{0}}$}
 }}
@@
 \expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\csname vid#1\endcsname}}}%
+\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
 }
@@
 \expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\csname eid#1\endcsname}}}%
+\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
 }
@@
 \newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\rcirclearrowleft_{#1}}}}}
+\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{\id{#1}}}}}
+\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{\id{#1}}}}}
 \newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
@@
     \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
-           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
+      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
+           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
     \end{array}
@@
 \begin{align*}
-  e_r &= \hole~\texttt{*}^{40}~\multiVertex{42}~\texttt{+}^{38}~\multiVertex{42} \\
-  MP &= \{x^{42}\} \\
-  NP &= \{z^{52}\} \\
-  W &= \{\hole~\texttt{+}^{46}~\cycleVertex{46}~\texttt{*}^{48}~y^{50}\} \\
+  e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
+  MP &= \{x^{\id{42}}\} \\
+  NP &= \{z^{\id{52}}\} \\
+  W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
 \end{align*}
@@
 There is one rule for each constructor form.
-For example, the expression derived from $v_{38} = (38, \ExpPlus)$ is $\expr'(v_{38}, \PlusLeft) \text{+}^{38} \expr'(v_{38}, \PlusRight)$.
+For example, the expression derived from $v_{\id{u}} = (u, \ExpPlus)$ is $\expr'(v_{\id{u}}, \PlusLeft) \text{+}^{\id{u}} \expr'(v_{\id{u}}, \PlusRight)$.
 In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
@@
 Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x^{48}$ is referenced twice in $e_r$.
+In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
 
 The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{58}$ has no parents (because it was deleted), so it appears only in $NP$.
+In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
 
@@
       \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^u \\
-      \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
-      \expr(v=(u, \ExpNum(n))) = n^u \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^u~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^u~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^u \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^u \\
+      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
+      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
+      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
+      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
+      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
+      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
+      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
+      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
+      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
     \end{array}
@@
 
-\[
-  \begin{array}{l}
-    \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-    \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-    \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-    \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-    \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-    \CC(G) = \ldots                                                                                                                                    \\
-    \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-    \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-    \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-    \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-    \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-    \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-  \end{array}
-\]
+% \[
+%   \begin{array}{l}
+%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
+%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
+%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
+%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
+%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
+%     \CC(G) = \ldots                                                                                                                                    \\
+%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
+%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
+%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
+%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
+%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
+%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
+%   \end{array}
+% \]
 

### 2021-02-16 grove-paper.tex
drop spurious figure
elaborate on the meaning of v = min(ancestors(v))

@@
 
-The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent.
-
+The fourth component $W$ is the set of expressions derived from the remaining vertices, which form cyclic structures we call wreaths, where each vertex has exactly one parent.
 In $W$, $\ancestors(v)$ is the set of vertices for which there is a path to $v$.
-Note that $v = \min(\ancestors(v))$ only if $v$ is part of a cycle.
+Note that $v = \min(\ancestors(v))$, where $\min(\{v_{\id{u_1}}, \ldots, v_{\id{u_n}}\}) = v_{\id{u_k}}$ such that $u_k = \min(\{u_{\id{1}}, \ldots, u_{\id{n}}\})$, holds only if $v$ is part of a cycle.
 
@@
 % TODO: add conflict hole to no-parent vertex
-
-\begin{figure}[b]
-\begin{align*}
-  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
-  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
-\end{align*}
-
-\end{figure}
 

### 2021-02-16 grove-paper.tex
make lfp a math operator
move grove example to right of its graph

@@
 \DeclareMathOperator{\children}{\text{children}}
+\DeclareMathOperator{\lfp}{\text{lfp}}
 
@@
   \centering
+  \begin{subfigure}{0.5\textwidth}
+    \centering
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \caption{Caption goes here}%
+    \caption{subcaption 1 goes here}%
+    \label{fig:Decomposition example graph}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \begin{align*}
+      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
+      MP &= \{x^{\id{42}}\} \\
+      NP &= \{z^{\id{52}}\} \\
+      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
+    \end{align*}
+    \caption{subcaption 2 goes here}%
+    \label{fig:Decomposition example grove}
+  \end{subfigure}
+  \caption{Main caption goes here}%
   \label{fig:Decomposition example}
@@
 \end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example} is
-\begin{align*}
-  e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-  MP &= \{x^{\id{42}}\} \\
-  NP &= \{z^{\id{52}}\} \\
-  W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-\end{align*}
+For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
+given in \autoref{fig:Decomposition example grove}.
 
@@
 
-% TODO: make macros for + and *
 % TODO: comment these out:
@@
     \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \textrm{lfp}(\ancestors') \right)(v) \\
+    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
     \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
@@
       \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of $a$.}
+  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
   \label{fig:Decomposition expr}

### 2021-02-22 grove-paper.tex
edits with Cyrus

@@
 \newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{\id{#1}}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{\id{#1}}}}}
+\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
+\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
 \newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
@@
 where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering,
+unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
 vertices are drawn from $\V$,
@@
 There is one rule for each constructor form.
-For example, the expression derived from $v_{\id{u}} = (u, \ExpPlus)$ is $\expr'(v_{\id{u}}, \PlusLeft) \text{+}^{\id{u}} \expr'(v_{\id{u}}, \PlusRight)$.
+For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
 In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
@@
 The fourth component $W$ is the set of expressions derived from the remaining vertices, which form cyclic structures we call wreaths, where each vertex has exactly one parent.
-In $W$, $\ancestors(v)$ is the set of vertices for which there is a path to $v$.
-Note that $v = \min(\ancestors(v))$, where $\min(\{v_{\id{u_1}}, \ldots, v_{\id{u_n}}\}) = v_{\id{u_k}}$ such that $u_k = \min(\{u_{\id{1}}, \ldots, u_{\id{n}}\})$, holds only if $v$ is part of a cycle.
+Every wreath consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
+We need to break the cycle to be able to turn a wreath into a tree.
+To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id according to the total ordering on $\U$.
+Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
+every vertex on the cycle is its own ancestor.
 
@@
 the unique identifier of the referenced vertex
-
-These first two components do not account for vertices with no parents. We
-include the expressions corresponding to these vertices in the third component
-$NP$.
-
-We've accounted for single-parented vertices connected to the root,
-single-parented vertices connected to multi-parented vertices, and vertices with
-no parents. This leaves only single-parented vertices that are not connected to
-the root or any multi-parented vertices. These vertices form cyclic structures
-we call wreaths.
-
-% A wreath is a cycle, potentially with acyclic branches. Wreaths are connected
-% graphs where each vertex has exactly one incoming edge. The fourth component $W$
-% is the set of expressions determined by extracting wreaths from any vertices not
-% in the other three components.
-
-% expression = tree
-
-% (show what a wreath looks like)
 

### 2021-03-01 grove-paper.tex
various small edits

@@
 
+% TODO: make sure we're not claiming to invent the term "unicycular graph"
+% TODO: move "algorithmic section" to Implementation section
+% TODO: maybe add a forward reference at end here to Implementation section
+% TODO: clean up old comments below
+
 A grove is a 4-tuple of the following form:
@@
 
+% TODO: clean up this figure
+% TODO: move to implementation section
+
 \begin{figure}
@@
     \begin{array}{lrlll}
-      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Con}(v, m) \mid \textrm{Ref}(v) \\
-      \textrm{Maps:}         & m & {}\in Map & {}::={} & [\,] \mid (p \mapsto [(\e, t), \ldots, (\e, t)]); m \\
+      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
+      \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
     \end{array}
@@
 
-Mapping to b actions
+% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
+% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
+
+move = delete + add
+
+other composite edits
+
+remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
+
+given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
+This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
+(we probably want to do it this way in Hazel)
+
+OR
+
+We could go the indirect route and avoid having to prove the theorem.
+(this way is easier to start with, and it's how GRV works)
+
+% Mapping to b actions
 
@@
 
-\section{TODO}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
 \section{Implementation}%
 \label{sec:Implementation}
+
+% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
+
 GRV -- how it is implemented, how it connects to the formalism, describe the graph,
@@
 Optimizations: least fixed point
-
-Related: hypertree decomposition is an established topic in AI/DB. Apparently,
-it's NP-complete in general, so people look for tractable subproblems. Recent
-work focuses on reduction to SMT and parallelizability for large db-like
-querying of large graphs, but some papers also focus on efficient decompopsition
-for "nearly acyclic" graphs. The closest I could find to what we're already
-doing are the k-decomp / opt-k-decomp / det-k-decomp algoriths in this paper:
-https://dl.acm.org/doi/10.1145/1412228.1412229, except we're getting determinism
-for free (i.e., not having to guess-and-check pivot points) by virtue of our
-global root and uuid-based total ordering on basically everything.
 
@@
 \label{sec:Discussion and Conclusion}
+
+TODO: leaves
+
+TODO: our model supports treating these as a cons-list of characters
 
@@
 
-% \section{Appendix}
-% \label{apx:Appendix}
-
-% \subsection{Concepts}
-
-% \begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
-%   $\A$ & Actions      & $\S$  & Edge States   \\
-%   %  $\C$ & Cursors & $\U$ & UUIDs         \\
-%   $\E$ & Edges        & $\V$  & Vertexes      \\
-%   $\G$ & Graphs       & $\AA$ & Graph Actions \\
-%   $\I$ & Indices      & $\EE$ & Editors       \\
-%   $\K$ & Constructors                         \\
-% \end{tabular}
-
-% A \emph{UUID} $u \in \U$ is a universally unique identifier.
-
-% \subsubsection{The Target Language}
-
-% \paragraph{Constructor} \par
-
-% A \emph{constructor} $k \in \K$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.
-
-% \paragraph{Index} \par
-
-% An \emph{index} $i \in \I$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST.
-Let $i_{root}$ be the hyper-position of all rooted constructors.
-
-% \subsubsection{The Graph}
-
-% \paragraph{Vertex} \par
-
-% A \emph{vertex} $v = (k, u) \in \V = \K \times \U$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$.
-Let $v_{root} = (k_{root}, u_{root})$ be the root vertex of a graph.
-The \emph{rooted vertexes} of a graph are the immediate children of its $v_{root}$.
-
-% \paragraph{Cursor} \par
-
-% %A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$.
-Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.
-
-% \paragraph{Edge} \par
-
-% %An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.
-
-% \paragraph{Edge State} \par
-
-% An \emph{edge state} $s \in \S = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).
-
-% \paragraph{Graph Action} \par
-
-% A \emph{graph action} $A = (\e,s,u) \in \A = \E \times \S \times \U$ is an instance of a binding from edge $\e$ to edge state $s$ that can be identified by UUID $u$.
-Denote by $[\e :-> s]_{u}$ a binding from edge $\e$ to edge state $s$ (identified by UUID $u$).
-
-% \paragraph{Graph} \par
-
-% A \emph{graph} $G = [\e :-> s]_{u}^{\ast{}} \in \G = \A^{\ast{}}$ is a function from edges to edge states.
-Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.
-
-% \subsubsection{The Editor}
-
-% \paragraph{Move Action} \par
-
-% %A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.
-
-% \paragraph{Edit Action} \par
-
-% An \emph{edit action} $a \in \A^{edit}= \{ \Create \} \times \K \cup \{ \Destroy \} \cup \{ \Restore \} \times \V$ adds or removes an edge at the cursor.
-
-% \paragraph{Communication Action} \par
-
-% A \emph{communication action} $a = (\Send, A^{\ast{}}, u^{\ast{}}) \in \A^{comm} = \{ \Send \} \times \A^{\ast{}} \times \U^{\ast{}}$ applies the graph action sequence $A^{\ast{}}$ to the editors identified by UUIDs $u^{\ast{}}$.
-
-% \paragraph{Action} \par
-
-% An \emph{action} $a \in \A = \A^{move} \cup \A^{edit} \cup \A^{comm}$ describes a change to one or more editors.
-
-% \paragraph{Editor} \par
-
-% %An \emph{editor} $E = (G,c,A_{Q}^{\ast{}},A_{H}^{\ast{}},u) \in \EE = \mathcal{G} \times \C \times \AA^{\ast{}} \times \AA^{\ast{}} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{\ast{}}$, and known graph action history $A_{H}^{\ast{}}$, that can be identified by UUID $u$.
-
-% \subsubsection{The Environment} \par
-
-% An \emph{environment} $(E^{\ast{}}, A^{\ast{}}) \in \EE^{\ast{}} \times \AA^{\ast{}}$ is a set of communicating editors $E^{\ast{}}$ that share a global graph action history $A^{\ast{}}$.
-
-% \subsection{Sets and Relations}
-
-% \paragraph{Edges} \par
-
-% \begin{align*}
-%   \edges(G)            & = \dom(G)                                       \\
-%   \liveEdges(G)        & = \{ \e \in \edges(G) | G(\e)=+ \}              \\
-%   \vertexChildren(G,v) & = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
-%   \cursorChildren(G,c) & = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
-%   \parents(G,v)        & = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
-% \end{align*}
-
-% \paragraph{Vertexes} \par
-
-% \begin{align*}
-%   \vertexes(G)        & = \{ v \in \V | \exists (c,v,u) \in \edges(G) \}              \\
-%   \liveVertexes(G)    & = \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}          \\
-%   \childVertexes(G,v) & = \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \} \\
-%   \orphans(G)         & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}      \\
-%   \multiparents(G)    & = \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}      \\
-%   \vertex(G,u)        & = v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
-% \end{align*}
-
-% \paragraph{Graph Actions} \par
-
-% $\known(E_1 \cdots E_n) = A_{H1}^{\ast{}} \cdots A_{Hn}^{\ast{}}
-%   \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j))$
-
-% \section{Operational Semantics}
-
-% \subsection{Movement}
-
-% $\boxed{E,A -->a E,A}$
-
-% \begin{mathpar}
-%   \inferrule{
-%   \leftIndex(i) = i'
-%   }{
-%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
-%   }
-
-%   \inferrule{
-%   \rightIndex(i) = i'
-%   }{
-%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
-%   }
-
-%   \inferrule{
-%   \parents(v) = \{(c',v',u')\}
-%   }{
-%   (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Up} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
-%   }
-
-%   \inferrule{
-%   \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
-%   \downIndex(k) = i
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{\ast{}},A_H^{\ast{}},u),A
-%   }
-
-%   \inferrule{}{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
-%   }
-% \end{mathpar}
-
-% \subsection{Editing}
-
-% \begin{mathpar}
-%   \inferrule{
-%   \defaultIndex(c) = TODO(Was None) \\
-%   u, u_k \in \U \text{ fresh} \\
-%   A_H' = [(c,(k,u_k),u) :-> +]
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{\ast{}},A_H^{\ast{}}A_H',u),A
-%   }
-
-%   \inferrule{
-%   \defaultIndex(c) = i' \\
-%   c' = ((k,u_k),i') \\
-%   u, u_k \in \U \text{ fresh} \\
-%   \\\\
-%   \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
-%   A_H'^{\ast{}} =
-%     [(c  ,(k,u_k),u  ) :-> +]
-%     [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
-%   [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k}
-%   (G A_H'^{\ast{}},c',A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
-%   }
-
-%   \inferrule{
-%   \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
-%   A_H'^{\ast{}} = [\e_j :-> -]_{j=1}^n
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Destroy}
-%   (G A_H'^{\ast{}},c,A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
-%   }
-
-%   \inferrule{
-%   A_H = [(c,v,u') :-> +] \\
-%   u' \text{ fresh}
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{\ast{}},A_H^{\ast{}}A_H,u),A %
-%   }
-% \end{mathpar}
-
-% \subsection{Communication}
-
-% \paragraph{Record} \par
-% %
-% \begin{mathpar}
-%   \inferrule{
-%   a \in \A^{edit} \\
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}},A_H'^{\ast{}},u),A
-%   }{
-%   (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}}a,A_H'^{\ast{}},u),A
-%   }
-% \end{mathpar}
-
-% \paragraph{Replay} \par
-
-% $\boxed{E,A \xrightarrow{a^{\ast{}}} E,A}$
-% %
-% \begin{mathpar}
-%   \inferrule{
-%     E,A -->a E'',A'' \\
-%     E'',A'' \xrightarrow{a^{\ast{}}} E',A' \\
-%   }{
-%     E,A \xrightarrow{aa^{\ast{}}} E',A'
-%   }
-% \end{mathpar}
-
-% \paragraph{Publish} \par
-
-% $\boxed{(E,A)^{\ast{}} -->a (E,A)^{\ast{}}}$
-% %
-% \begin{mathpar}
-%   \inferrule{
-%   E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j) \\
-%   E_j,A_j \xrightarrow{a'^{\ast{}}} (G_j',c_j',A_{Qj}'^{\ast{}}, A_{Hj}'^{\ast{}},u_j),A_j' \\
-%   E_j' =
-%   ( G_j'
-%   , c_j'
-%   , A_{Qj}'^{\ast{}} \setminus a'^{\ast{}}
-%   , A_{Hj}'^{\ast{}} \setminus \known(E_1 \cdots E_n),u_j) \\
-%   j = 1,\ldots,n
-%   }{
-%   (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{\ast{}} } (E_1',A_1') \cdots (E_n',A_n')
-%   }
-% \end{mathpar}
-
-% Missing Action.env members: Record / Report / Stop / Replay, Dump / Load, Clone / Drop
-
-% $G \overset{\left(e,s\right)}{\longrightarrow} G\left[ e \mapsto s \sqcup G(e) \right]$
-
-% % Edge.Map.add action.value.edge (join old_state new_state) graph
-
 \end{document}

### 2021-03-01 grove-paper.tex
rename W to U and wreath to unicycular graph

@@
   NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
+  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
 \end{align*}
@@
 A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
+If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
 For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
@@
 
-The fourth component $W$ is the set of expressions derived from the remaining vertices, which form cyclic structures we call wreaths, where each vertex has exactly one parent.
-Every wreath consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-We need to break the cycle to be able to turn a wreath into a tree.
+The fourth component $U$ is the set of expressions derived from the remaining vertices, which form unicycular graphs where each vertex has exactly one parent.
+Every unicycular graph consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
+Breaking the cycle turns a unicyclic graph into a tree.
 To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id according to the total ordering on $\U$.
 Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-every vertex on the cycle is its own ancestor.
+only a vertex on the cycle can be its own ancestor.
 
@@
 
-% TODO: add text around these definitions -- turn them into paragraphs.
-
 We determine $e$ by starting at the root of the graph and including every connected
@@
 Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things wreaths. A
-wreath is a cycle of single-parented vertices that each may have single-parented
+parent, everything else must be single parented. We call these things unicycular graphs. A
+unicycular graph is a cycle of single-parented vertices that each may have single-parented
 descendants.
 
-lemma: a wreath is a cycle with trees hanging off of it
-
-lemma: wreath traversal always produces a valid tree
-
-Wreath traversal occurs in two directions. A vertex is chosen as an arbitrary
+lemma: a unicycular graph is a cycle with trees hanging off of it
+
+lemma: unicycular graph traversal always produces a valid tree
+
+Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
 starting point and its parents are traversed and marked as seen. Any vertex that
 has been seen twice must be on the cycle. We take the first one as the root of
-the wreath and traverse its children to produce a tree that covers the wreath.
-Any remaining single-parented vertices must be parts of other wreaths and are
-wreath-traversed until there are no more single-parented vertices left.
+the unicycular graph and traverse its children to produce a tree that covers the unicycular graph.
+Any remaining single-parented vertices must be parts of other unicycular graphs and are
+traversed until there are no more single-parented vertices left.
 

### 2021-03-01 grove-paper.tex
move decomp algorithm notes to Implementation
edits with Cyrus and Michael

@@
 
-% \newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
-% \newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\rcirclearrowleft_{#1}}}}}
-
 % TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: move "algorithmic section" to Implementation section
 % TODO: maybe add a forward reference at end here to Implementation section
-% TODO: clean up old comments below
+% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
 
@@
 
-The fourth component $U$ is the set of expressions derived from the remaining vertices, which form unicycular graphs where each vertex has exactly one parent.
-Every unicycular graph consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
+The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
+These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
+Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
 Breaking the cycle turns a unicyclic graph into a tree.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id according to the total ordering on $\U$.
+% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
+% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
+To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
 Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
 only a vertex on the cycle can be its own ancestor.
+When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
+In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
 
@@
 
-% TODO: add conflict hole to no-parent vertex
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-lemma: a unicycular graph is a cycle with trees hanging off of it
-
-lemma: unicycular graph traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycular graph and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-traversed until there are no more single-parented vertices left.
-
 % XXX
@@
 
-% TODO: clean up this figure
-% TODO: move to implementation section
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-      \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-    \end{array}
-  \]
-  \caption{Syntax of graph decomposition trees as a grammar.}
-  \label{fig:Syntax of graph decomposition trees}
-\end{figure}
-
-% To render the graph as disjoint fragments of program code, we decompose the
-% graph into an equivalent representation using trees. The result is a 4-tuple,
-% where the first component is a tree, the second and third components are sets of
-% trees, and the fourth component is a set of cycles.
-
-% 1. ($O(n \log n)$ for maps, $O(n)$ for hash table) linear scan to partition vertices according to # of parents: multiple parents, single parents, no parents.
-
-% 2. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from root (and remove from single-parented set)
-
-% 3. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from multi-parented vertices (and remove from single-parented set).
-
-% 4. ($n \log n$?) scan to discover edges leading to single-parented [deleted]
-
-% 5. simple cycles—reverse traversal, then forward traversal. Here's where we want to prove some things (e.g., dangling vertices are not deleted)
-
-% tree ::= Vertex (Constructor) | Ref (Vertex.Id) | Conflict (tree list)
-
-%
-% ... () of sets of
-% (poly)trees) wherein each multi-edge has been resolved into a simple edge with
-% one or more symbolic references to it.
-
-% Our decomposition algorithm first separates the graph into three disjoint sets of
-% vertices according to the number of parents and presence of simple cycles.
-% %
-% The first set is the single-parented vertices that have the root as an ancestor.
-% The second set is the vertices that have a multi-parented vertex as an ancestor.
-% The third
-
-% We then derive equivalent tree representations of connected components by tracing
-% the ancestry, taken as the transitive closure of the childhood relation, of the
-% vertices in each set.
-
-% set contains the multi-parented vertices and
-% their non-multiparented descendants. The second set contains the vertices with
-% no parents and their non-multiparented descendants. We can determine both of
-% these sets with a single pass over all of the live vertices in the graph. The
-% third set contains the single-parented vertices that are descendants of the root
-% vertex.
-
-% The first set contains all of the
-% vertices that have the root as an ancestor.
-
-% : vertices with multiple parents, vertices with no parents,
-% single-parented vertices reachable from the root, and everything else.
-
-% separate it into 
-% components that only touch  detect any vertices that either have multiple parents or no parents
-% at all.
-
-% partition its vertices into three disjoint sets: multi-parented, deleted,
-% and any remaining vertices participating in a live edge. We then partition the
-% graph into four non-overlapping sets of trees: multi-parented, single-parented
-% reachable from root, deleted, and simple cycles.
-
-%  detect any multi-parented
-% vertices and use them to partition the remaining vertices of the graph into .
-
-% Our decomposition algorithm partitions the graph into trees of
-% connected components. Cycles are resolved by replacing all but one vertex of
-% each multi-edge with a by-id reference.
-
-% Translating a graph into expressions in the calculus allows us to model user
-% interaction as a stream of edits as described in \autoref{sub:User Actions}.
-
-% \begin{algorithm}[t]
-%   \caption{Graph Decomposition}
-%   \label{fig:Graph Decomposition}
-
-%   \DontPrintSemicolon
-%   \SetAlgoNoLine
-%   \SetKwFunction{reachable}{reachable}
-%   \SetKwProg{Fn}{function}{}{end}
-
-%   \KwIn{Graph $G$}
-%   \KwOut{A partitioning of the edges of $G$}
-
-%   $live := \{\e : G(\e) = \Plus\}$\;
-%   $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
-%   $D := \{\reachable(\e) : \e \notin live\}$\;
-%   $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
-%   $MP := \{\reachable(\e) : \e \in multiparent\}$\;
-%   $SC := \varnothing$\;
-%   $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
-%   \While{$remaining \ne \varnothing$}{
-%     $(root, cycle) := \reachable(\min(remaining))$\;
-%     $remaining := remaining \setminus \{root\} \setminus cycle$\;
-%     $SC := SC \cup \{(root, cycle)\}$\;
-%   }
-%   \Return{$(R, D, MP, SC)$}\;
-%   \;
-
-%   \Fn{$\reachable\ ((u, v, p, v'))$}{
-%     $children := \{(u', v', p', v'') \in live\}$\;
-%     $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
-%     \Return{$((u, v, p, v'), descendants)$}
-%   }
-% \end{algorithm}
-
-% The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
-% \wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
-% portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
-% expressions that involve vertices with multiple parents. $D$ contains deleted
-% expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
-% $SC$ contains the deleted expressions that do not involve multiple parents but
-% do contain cycles.
-
-% A simple cycle is a set of vertices connected by a closed path, consisting
-% of:
-
-% 1) two or more vertices, all descendants of each other, and
-% 2) the vertices of any non-multiparent subtrees reachable from the core, and
-% not reachable from the root vertex.
-
-% To determine the simple cycles of a graph, we:
-
-% 1) check if any vertex in the graph is a descendant of itself or some other
-% vertex that is (a descendant of itself),
-% 2) find the least vertex in the core of this new(ly discovered) cycle, and
-% 3) return a tree connecting every vertex reachable from the least (vertex).
-
-% The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.
-
-% <Mathematical definition of above.>
+% TODO: maybe add a forward reference here to Implementation section
 
@@
 
+% TODO: talk about code generation somewhere in here
+
+% TODO: talk about:
+% - decomp vertex set generation and unicycle traversal
+% - how to choose which branch of expr' to take
+% - lfp(ancestors') and min of it
+
+% TODO: can we save any of this in a README or code comments?
+
 % TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
@@
 Optimizations: least fixed point
+
+We determine $e$ by starting at the root of the graph and including every connected
+vertex that has a single parent. In positions where there are no children, we
+leave an empty hole $\hole$. In positions where there are multiple children, we
+leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
+encounter a vertex with multiple parents, we add its corresponding expression to
+the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
+the unique identifier of the referenced vertex
+
+% here are some properties that show how to get one from the other, and to/from
+% graph
+
+algorithmic section: if we want to implement these conversions, here are the
+steps (without introducing new concepts):
+
+- linear scan
+- breaking out roots
+- ...
+
+Graph decomposition occurs in three steps.
+
+First, a linear scan is performed over every edge that has been created or
+deleted. The vertices at both ends are partitioned into three sets:
+multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
+child edge set lookups are also created.
+
+Second, the subgraphs rooted at vertices with more or less than one parent are
+traversed and equivalent tree structures are generated over components that are
+connected by single-parented vertices. Cycles are broken with UUID vertex
+references. Single-parented vertices reachable from the root vertex are also
+traversed.
+
+Finally, once we've already accounted for things with more or less than one
+parent, everything else must be single parented. We call these things unicycular graphs. A
+unicycular graph is a cycle of single-parented vertices that each may have single-parented
+descendants.
+
+% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
+% TODO: maybe show this figure and say that it's possible to handle generically without code generation
+
+% \begin{figure}
+%   \[
+%     \arraycolsep=0pt
+%     \begin{array}{lrlll}
+%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
+%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
+%     \end{array}
+%   \]
+%   \caption{Syntax of graph decomposition trees as a grammar.}
+%   \label{fig:Syntax of graph decomposition trees}
+% \end{figure}
+
+lemma: a unicycle is a cycle with trees hanging off of it
+
+lemma: unicycle traversal always produces a valid tree
+
+Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
+starting point and its parents are traversed and marked as seen. Any vertex that
+has been seen twice must be on the cycle. We take the first one as the root of
+the unicycle and traverse its children to produce a tree that covers the unicycular graph.
+Any remaining single-parented vertices must be parts of other unicycular graphs and are
+unicycle-traversed until there are no more single-parented vertices left.
 

### 2021-03-08 grove-paper.tex
edits with Cyrus and Michael

@@
 
+\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
+
 % constructors
@@
 edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
+$\e$ has not been created or deleted. When $G(\e)$ is $\Plus$, the edge has been
+created. When $G(\e)$ is $\Minus$, the edge has been deleted. The total
 ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
@@
 
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
+% GRV -- how it is implemented, how it connects to the formalism, describe the graph,
+
+We implemented the core Grove calculus of \autoref{sec:Formalism} as an OCaml library.
+Our code is parameterized by a syntax specification for expressions that we generate automatically, which makes the implemented language of expressions easier to modify or replace.
+We also built a Web browser-based Grove editor as a proof of concept.
+
+Since we cannot implement an infinite mapping directly, the graph is a data structure, \texttt{Graph.t}, that maps from edges to the edge states $\left\{\Plus, \Minus\right\}$.
+We do not represent edges that map to $\bot$.
+
+% TODO: rename Decomposition.ml to Grove.ml
+
+There's a function \verb!decompose! of type \verb!Graph.t) : t * Edge.Set.t Vertex.Map.t! that
+
+From this graph, we compute the set of all edges and the set of live edges with a linear scan.
+
+% TODO: make this into a listing
+
+\begin{verbatim}
+  Graph.edges : Graph.t -> Edge.Set.t
+  Graph.live_edges : Graph.t -> Edge.Set.t
+\end{verbatim}
+
+% TODO: keep connecting the code to the formalism like this
+
+These vertices are partitioned into three sets: multi-parented, single-parented, or orphaned.
+A vertex $v$ is considered a parent of another vertex $v'$ if there is a live edge from $v$ to $v'$.
+
+For each vertex, we take the parent and child edge sets.
+
+% TODO: fix the formalism for parents (to work with edges instead of vertices) and then check that it matches the implementation. Try renaming parents to vparents and adding parents for edges.
+
+% TODO: Talk about how the graph is implemented (not as an actual function, but as a map from edges to edge states) so that we can denote the set of created + deleted edges clearly, as well as all of the vertices that have ever been created.
+
+For a graph $G : \E \to \Sigma$ where $\E = \U \times \V \times \P \times \V$, our graph decomposition algorithm runs in $O(\abs{\V} \log \abs{\V} + \abs{\E} \log \abs{\V})$.
+It begins with a scan of all edges that have been created or deleted $O(\abs{\E})$.
+Their vertices at both ends are partitioned into three sets: multi-parented, single-parented, or orphaned.
+Their relationships are recorded in maps for $O(\log \abs{\V})$ lookups of parent and child edge sets.
+After the vertices have been partitioned, we traverse the various single-parented components and produce equivalent expressions $O(\abs{\V})$.
+For unicycles, we traverse backwards until a vertex is seen twice $O(\abs{\V})$, then proceed forward to find the least vertex on the cycle $O(\abs{\V})$.
+
+% TODO: is this still a thing?
+% Compared to the core language, our implementation lacks support for lists and case expressions.
 
@@
 Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
 

### 2021-03-15 grove-paper.tex
sketching formal properties with Cyrus and Michael

@@
 
+\DeclareMathOperator{\sort}{\text{sort}}
 \DeclareMathOperator{\arity}{\text{arity}}
@@
       \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
+      \arity(\Root)={} & \left\{ (\Root, Exp) \right\} \\
       \arity(\PatVar(x))={} & \left\{ \right\} \\
@@
 
+\vspace*{\baselineskip}
+
+% TODO: add a lead-in sentence here
+
+A graph $G$ is well sorted if
+for all $G((u, (u_1, k_1), p, (u_2, k_2))) \neq \bot$,
+we have $(p, \sort(k_2)) \in \arity(k_1)$.
+
+% In decomp section:
+% - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
+%   - want to show that expr is a total function on Exp's
+% - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
+%   - every vertex should be part of the decomposition
+% - correctness of grove components:
+%   - e_r: rootVertex(e_r) = v_r
+%   - MP: for every expression in MP, |parents(the root vertex)| > 1
+%   - NP: for every expression in NP, |parents(the root vertex)| = 0
+%   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
+%     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
+%   - we've accounted for every edge:
+%     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
+%     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
+
+% correctness covers:
+% - invertibility (w/ some identity assignment)
+% - parentage (things are multiparented correctly)
+
+\vspace*{\baselineskip}
+
 % TODO: should graph "recomposition" have an algorithm, too?
+% TODO: update this to match our current understanding
 A graph can be derived from an expression in the calculus by starting with an
@@
 \[
-  (e_r, MP, NP, W)
+  (e_r, MP, NP, U)
 \]
@@
 The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
+\autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
+There is one rule for each term constructor.
 For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
+In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
 An empty hole $\hole$ appears if there are no children at the given position.
 A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
+If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
 For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.

### 2021-04-09 figure-disconnect.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 }
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
-
-\begin{document}
-
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
-
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
-
-\end{document}

### 2021-04-09 figure-multi-parent.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
-
-\begin{document}
-
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
-
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
-
-\end{document}

### 2021-04-09 formalism.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 \documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
 
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
 \acmJournal{PACMPL}
@@
 
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
 \setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
 
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
+\input{config-tex}
+\input{config-tikz}
+\input{config-glyphs}
+\input{config-definitions}
+\input{config-graphs}
 
@@
 
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
+\title[Grove]{Excerpt: Recomposition}
 
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
 \maketitle
 
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
+\input{appendix-formalism}
 

### 2021-04-09 grove-paper-discussion-and-conclusion.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
-
-\begin{document}
-
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
-
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
 \section{Discussion and Conclusion}%
@@
 GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
-
-\end{document}

### 2021-04-09 grove-paper-formalism.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
-
-\begin{document}
-
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
-
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
 \section{Formalism}%
@@
 
+\vspace*{\baselineskip}
+
+% TODO: add a lead-in sentence here
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: put this in a definition box
+
+A graph $G$ is well sorted if
+for all $\e = (u, (u_1, k_1), p, (u_2, k_2))$
+such that $G(\e) \neq \bot$,
+we have $(p, \sort(k_2)) \in \arity(k_1)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% In decomp section:
+% - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
+%   - want to show that expr is a total function on Exp's
+% - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
+%   - every vertex should be part of the decomposition
+% - correctness of grove components:
+%   - e_r: rootVertex(e_r) = v_r
+%   - MP: for every expression in MP, |parents(the root vertex)| > 1
+%   - NP: for every expression in NP, |parents(the root vertex)| = 0
+%   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
+%     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
+%   - we've accounted for every edge:
+%     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
+%     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
+
+% correctness covers:
+% - invertibility (w/ some identity assignment)
+% - parentage (things are multiparented correctly)
+
+\vspace*{\baselineskip}
+
 % TODO: should graph "recomposition" have an algorithm, too?
+% TODO: update this to match our current understanding
 A graph can be derived from an expression in the calculus by starting with an
@@
 
-A grove is a 4-tuple of the following form:
+A grove is a 3-tuple of the following form:
 \[
-  (e_r, MP, NP, W)
+  (NP, MP, U)
 \]
@@
 \begin{align*}
-  e_r &= \expr(v_0) \\
+  % e_r &= \expr(v_0) \\
+  NP &= \{\expr(v) : |\parents(v)| = 0\} \\
   MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
   U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
@@
 Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
+%
+The first component $NP$ is the set of expressions rooted at a vertex with no parents, including the main root vertex.
+% TODO: change vRoot to dot w/ subscript
+In the example, $e_r$ is the expression corresponding to the main root, $\vRoot$.
+% TODO: change z_... to macro use w/ same appearance as in the graph
+In addition, $z_{\id{52}}$ has no parents because it was deleted, so it also appears in $NP$.
+
+% The first component $e_r$ is the expression derived from the root vertex $v_0$.
+% \autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
+% There is one rule for each term constructor.
+% For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
+% In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
+% An empty hole $\hole$ appears if there are no children at the given position.
+% A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
+% If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
+% For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
 
@@
 
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
+The third component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
 These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
@@
 In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% THEOREMS GO HERE
+
+% - [ ]  State the "exhaustiveness check" theorem formally
+
+% - [ ]  State the "every vertex is part of decomposition" theorem formally
+
+% i.e., {univ, multiv, av} is a partitioning of the vertices of G
+% where, e.g., multiv = vertices(MP)
+
+% - [ ]  Define "recomposition" formally
+
+% not defined yet
+
+% - [ ]  State the relationship between decomp(G) and recomp(grove)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
-
-\end{document}

### 2021-04-09 grove-paper-related-work.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-\bibliographystyle{ACM-Reference-Format}
-%\citestyle{acmauthoryear}
-\citestyle{acmnumeric}
-
-%%%%%%%%
-
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
-
-\begin{document}
-
-%% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
-%\subtitle{Subtitle}
-
-%% Author information
-\author{Michael D. Adams}
-\orcid{0000-0003-3160-6972}
-
-\author{Eric Griffis}
-\orcid{0000-0003-1693-6172}
-
-\author{Cyrus Omar}
-\orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
-  \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
-  \city{Ann Arbor}
-  \state{MI}
-  \postcode{48109-2121}
-  \country{USA}
-}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
 \section{Related Work}%
@@
 Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-  %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Appendix
-\appendix
-
-%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
-
-%% Bibliography
-\bibliography{grove-paper}
-
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
-
-\end{document}

### 2021-04-09 grove-paper.tex
move configs, figures, and sections to separate files

--- 2021-03-01 grove-paper.tex
@@
 
-%%%%%%%%
+\input{config-tex}
+\input{config-tikz}
+\input{config-glyphs}
+\input{config-definitions}
+\input{config-graphs}
 
-\renewcommand{\topfraction}{1} % Allow floats to take up the page
-\renewcommand{\textfraction}{0}
-
-%%%%%%%%
-% \autoref from hyperref
-\renewcommand{\AMSautorefname}          {Equation}
-\renewcommand{\appendixautorefname}     {Appendix}
-\renewcommand{\chapterautorefname}      {Chapter}
-\renewcommand{\equationautorefname}     {Equation}
-\renewcommand{\FancyVerbLineautorefname}{Line}
-\renewcommand{\figureautorefname}       {Figure}
-\renewcommand{\footnoteautorefname}     {Footnote}
-\renewcommand{\Hfootnoteautorefname}    {Footnote}
-\renewcommand{\itemautorefname}         {Item}
-\renewcommand{\Itemautorefname}         {Item}
-\renewcommand{\pageautorefname}         {Page}
-\renewcommand{\paragraphautorefname}    {Section}
-\renewcommand{\partautorefname}         {Part}
-\renewcommand{\sectionautorefname}      {Section}
-\renewcommand{\subparagraphautorefname} {Section}
-\renewcommand{\subsectionautorefname}   {Section}
-\renewcommand{\subsubsectionautorefname}{Section}
-\renewcommand{\tableautorefname}        {Table}
-\renewcommand{\theoremautorefname}      {Theorem}
-
-%% Packages
-\usepackage{booktabs}
-\usepackage[rule=false]{subcaption}
-\usepackage{graphicx}
-% \usepackage{semantic}
-\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
-\let\colonsim\undefined % Avoid redefinition error in `colonequals`
-\usepackage{colonequals}
-% \usepackage{mathpartir}
-\usepackage{stmaryrd}
-\usepackage{fontawesome}
-\usepackage{array}
-\usepackage{todonotes}
-% \usepackage[ruled]{algorithm2e}
-
-% Load MnSymbol without clobbering \ast
-% See https://tex.stackexchange.com/a/269691
-\usepackage{amsmath}% needed before mathabx
-\let\amsast=\ast
-\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
-\let\abxast=\ast
-\usepackage{MnSymbol}
-\let\mnast=\ast
-\let\ast=\abxast
-
-% \selectcolormodel{gray}
-
-\def\_{\texttt{\textunderscore}}
-
-\newcommand{\id}[1]{\textcolor{gray}{\ensuremath{#1}}}
-
-%%%%%%%%
-% TikZ Stuff
-%\usepackage{etex} % Fix "No room for new \dimen" error
-\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
-\usepackage{tikz}
-\usetikzlibrary{babel} % Ensure compatibility the 'babel' package
-
-\usetikzlibrary{external} % Needs to be separately enabled
-%\tikzexternalize % Enable externalization
-%\usepackage{lua-visual-debug}
-
-\usetikzlibrary{arrows.meta} % Arrow Tips
-\tikzset{>=Stealth}
-%\tikzset{<=stealth}
-%\tikzset{arrows={-Stealth[scale=50]}}
-%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
-%\tikzset{wideline/.style={line width=0.7pt}}
-%\tikzset{boldline/.style={color=black,line width=1.0pt}}
-
-\usetikzlibrary{
-  backgrounds,  % Provides "framed" and "gridded"
-  bending,      % bending arrow tips
-  decorations.pathmorphing,   % Provides wavy edges
-  graphs,       % Graph *notation*
-  graphdrawing, % Graph *layout*
-  quotes,       % Quote syntax (e.g., "foo")
-}
-
-\usegdlibrary{
-  trees,
-}
-
-\tikzset{
-  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
-}
-\tikzset{edge style/.style={
-  draw,
-  %color=gray,
-  font={\small\ttfamily},
-  /tikz/every edge quotes/.style={
-    %draw=gray!20,
-    anchor=west,
-    swap/.append code={
-      \ifpgfarrowswap
-        \pgfkeysalso{anchor=west}
-      \else
-        \pgfkeysalso{anchor=east}
-      \fi}},
-}}
-\tikzset{graphs/graph style/.style={
-  tree layout,
-  level distance=0.5cm,
-  level sep=0.5cm,
-  sibling distance=0.5cm,
-  sibling sep=0.1cm,
-  part distance=0.1cm,
-  part sep=0.1cm,
-  component distance=0.1cm,
-  component sep=0.1cm,
-  nodes={
-    draw,
-    %color=gray,
-    inner sep=2pt,
-    rounded corners=1mm},
-  edges={edge style},
-}}
-\tikzset{graphs/root style/.style={
- %draw=none,
- as={\textbullet$_{\id{0}}$}
-}}
-\tikzset{alice/.style={
-  color=red!80!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{bob/.style={
-  color=green!60!black,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{merge/.style={
-  color=blue,
-  font={\bfseries\small},
-  thick,
-}}
-\tikzset{alice edge/.style={alice, edge style, font={\bfseries\small}}}
-\tikzset{alice node/.style={alice}}
-\tikzset{alice step/.style={alice}}
-\tikzset{bob edge/.style={bob, edge style}, font={\bfseries\small}}
-\tikzset{bob node/.style={bob}}
-\tikzset{bob step/.style={bob}}
-\tikzset{merge edge/.style={merge, edge style}, font={\bfseries\small}}
-\tikzset{merge node/.style={merge}}
-\tikzset{merge step/.style={merge,decorate,decoration={coil,amplitude=1.0pt,segment length=7.0pt,aspect=0}}}
-\tikzset{star/.style={edge node={node[inner sep=0pt,at end,sloped] {\textbf{\huge${}^{\ast}$}}}}}
-
-%%%%%%%%
-% Lang components
-
-% sets
-\def\A{\mathcal{A}}
-%\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
-\def\I{\mathcal{I}}
-\def\K{\mathcal{K}}
-\def\P{\mathcal{P}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\e{\varepsilon}
-\def\AA{\textbf{A}}
-\def\EE{\textbf{E}}
-
-% constructors
-\def\ExpVar{\mathsf{Exp\_var}}
-\def\ExpLam{\mathsf{Exp\_lam}}
-\def\ExpApp{\mathsf{Exp\_app}}
-\def\ExpNum{\mathsf{Exp\_num}}
-\def\ExpPlus{\mathsf{Exp\_plus}}
-\def\ExpTimes{\mathsf{Exp\_times}}
-\def\PatVar{\mathsf{Pat\_var}}
-\def\TypNum{\mathsf{Typ\_num}}
-\def\TypArrow{\mathsf{Typ\_arrow}}
-
-% indexes
-\def\Root{\mathsf{Root}}
-\def\LamParam{\mathsf{Lam\_param}}
-\def\LamType{\mathsf{Lam\_type}}
-\def\LamBody{\mathsf{Lam\_body}}
-\def\AppFun{\mathsf{App\_fun}}
-\def\AppArg{\mathsf{App\_arg}}
-\def\PlusLeft{\mathsf{Plus\_left}}
-\def\PlusRight{\mathsf{Plus\_right}}
-\def\TimesLeft{\mathsf{Times\_left}}
-\def\TimesRight{\mathsf{Times\_right}}
-\def\ArrowArg{\mathsf{Arrow\_arg}}
-\def\ArrowResult{\mathsf{Arrow\_result}}
-
-% actions
-\def\Create{\text{Create}}
-\def\Destroy{\text{Destroy}}
-\def\Down{\text{Down}}
-\def\Enqueue{\text{Enqueue}}
-\def\Left{\text{Left}}
-\def\Move{\text{Move}}
-\def\Num{\text{Num}}
-\def\Restore{\text{Restore}}
-\def\Right{\text{Right}}
-\def\Select{\text{Select}}
-\def\Send{\text{Send}}
-\def\Up{\text{Up}}
-
-% relations
-\DeclareMathOperator{\s}{\text{S}}
-\DeclareMathOperator{\MP}{\text{MP}}
-\DeclareMathOperator{\CC}{\text{C}}
-\DeclareMathOperator{\D}{\text{D}}
-
-\DeclareMathOperator{\arity}{\text{arity}}
-\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
-\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
-\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
-\DeclareMathOperator{\dom}{\text{dom}}
-\DeclareMathOperator{\edges}{\text{edges}}
-\DeclareMathOperator{\known}{\text{known}}
-\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
-\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
-\DeclareMathOperator{\multiparents}{\text{multiparents}}
-\DeclareMathOperator{\orphans}{\text{orphans}}
-\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
-\DeclareMathOperator{\reachable}{\text{reachable}}
-\DeclareMathOperator{\seen}{\text{seen}}
-\DeclareMathOperator{\unseen}{\text{unseen}}
-\DeclareMathOperator{\vertexes}{\text{vertexes}}
-\DeclareMathOperator{\vertex}{\text{vertex}}
-\DeclareMathOperator{\publish}{\text{publish}}
-\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
-\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
-\DeclareMathOperator{\downIndex}{\text{downIndex}}
-\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}
-
-% graph decomposition operators
-\DeclareMathOperator{\expr}{\text{expr}}
-\DeclareMathOperator{\patt}{\text{patt}}
-\DeclareMathOperator{\type}{\text{type}}
-\DeclareMathOperator{\parents}{\text{parents}}
-\DeclareMathOperator{\ancestors}{\text{ancestors}}
-\DeclareMathOperator{\vertices}{\text{vertices}}
-\DeclareMathOperator{\children}{\text{children}}
-\DeclareMathOperator{\lfp}{\text{lfp}}
-
-\DeclareMathOperator{\multiParented}{\text{multiParented}}
-\DeclareMathOperator{\singleParented}{\text{singleParented}}
-\DeclareMathOperator{\deleted}{\text{deleted}}
-
-% Define outline versions of + and -
-\def\outlinepad{0.4pt}
-\def\outlinestroke{0.4pt}
-\newcommand{\Plus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {+};
-\path[draw, line width=\outlinestroke]
-   ( 0.333em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.333em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad, -0.084em-\outlinestroke/2-\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.333em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --(-0.020em-\outlinestroke/2-\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.583em+\outlinestroke/2+\outlinepad)
- --( 0.021em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-\newcommand{\Minus}{\mathord{
-\begin{tikzpicture}[anchor=base, baseline]
-%\node at (0,0) {$-$};
-\path[draw, line width=\outlinestroke]
-   ( 0.306em+\outlinestroke/2+\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --( 0.306em+\outlinestroke/2+\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.229em-\outlinestroke/2-\outlinepad)
- --(-0.306em-\outlinestroke/2-\outlinepad,  0.270em+\outlinestroke/2+\outlinepad)
- --cycle
- ;
-\end{tikzpicture}
-}}
-
-% custom syntax
-
-\usepackage{xstring}
-
-\newcommand{\hole}{\ensuremath{\square}} %\textcolor{violet}{\llparenthesis}}\textcolor{violet}{\rrparenthesis}}
-\newcommand{\conflictHole}[1]{%
-{\noexpandarg\StrSubstitute{#1}{,}{\textcolor{red}{\;\textbf{|}\;}}[\myargs]%
-{\textcolor{red}{\textbf{\{}}\myargs\textcolor{red}{\textbf{\}}}}}}%
-
-% Calling \newvertex{Foo}{bar} defines
-%   \vidFoo to be a new id number, and
-%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
-\newcounter{NodeVertexCounter}
-\newcommand{\newvertex}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{1}%
-\else
-  \addtocounter{NodeVertexCounter}{2}%
-\fi
-\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
-}
-\newcommand{\newedge}[2]{%
-\ifodd\theNodeVertexCounter
-  \addtocounter{NodeVertexCounter}{2}%
-\else
-  \addtocounter{NodeVertexCounter}{1}%
-\fi
-\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
-\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
-\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
-}
-\setcounter{NodeVertexCounter}{-1}
-\newvertex{Root}{Root}
-
-% Support \includegraphics of .dot files
-\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}
-
-\newcommand{\Edge}[1]{Edge~#1}
-\newcommand{\Vertex}[1]{Vertex~#1}
-\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
-\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
-\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
-\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
-\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
-\newcommand{\otherVertexVskip}{\vskip0.5em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureSimple
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{SimpleTimes}{Root}%
-\newvertex{SimpleTimes}{*}%
-\newedge{SimpleX}{L}%
-\newvertex{SimpleX}{x}%
-\newedge{SimpleY}{R}%
-\newvertex{SimpleY}{y}%
-\newcommand{\figureSimple}{
-\begin{figure}
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
-\end{tikzpicture}
-}%
-\label{fig:Simple:a}
-\figureCode{x * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style, color=black] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
-   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
-\end{tikzpicture}
-}%
-\label{fig:Simple:b}
-\figureCode{x * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    "\vSimpleX" [> "\eSimpleX"'],
-    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
-  }
-  }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.32\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
-\end{tikzpicture}
-}%
-\label{fig:Simple:c}
-\figureCode{\hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
-  root[root style] -> {
-  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
-};
-\path (-0.25cm,-1.7cm) graph[graph style] {
-  "\vSimpleX"
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Code that Includes a Hole and Some Simple Edits.}%
-\label{fig:Simple}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureWrap
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{WrapPlus}{Root}
-\newvertex{WrapPlus}{+}
-\newedge{WrapTimes}{L}
-\newcommand{\figureWrapMove}{
-\begin{figure}
-\centering
-\begin{minipage}[t]{.45\linewidth}
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
-%%%%
-\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:a}
-\figureCode{\hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style]
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:b}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
- }
-};
-\path (0cm,-1.7cm) graph[graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
-\end{tikzpicture}
-}%
-\label{fig:Wrap:c}
-\figureCode{\hole{} * y + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Wrapping.  (Note that this figure omits the orphaned \vSimpleX{} since it is no longer relevant to the narrative of this paper.)}%
-\label{fig:Wrap}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMove
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MoveTimes}{R}
-% \newcommand{\figureMove}{
-% \begin{figure}[H]
-\begin{minipage}[t]{.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
-\end{tikzpicture}
-}%
-\label{fig:Move:a}
-\figureCode{\hole{} + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (0cm,-1.7cm) graph [graph style] {
- "\vSimpleTimes" -> {
-  {},
-  "\vSimpleY" [> "\eSimpleY"]
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.43\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
-\end{tikzpicture}
-}%
-\label{fig:Move:b}
-\figureCode{\hole{} + \hole{} * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
-    {},
-    "\vSimpleY" [> "\eSimpleY"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Repositioning Code.}%
-\label{fig:Move}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDifferentParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DifferentPartsAlice}{L}
-\newvertex{DifferentPartsAlice}{u}
-\newedge{DifferentPartsBob}{R}
-\newvertex{DifferentPartsBob}{v}
-% \newcommand{\figureDifferentParts}{
-\newcommand{\figureDifferentPartsNestedParts}{
-\begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
-\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:a}
-\figureCode{\hole{} + u * y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
-    "\vSimpleY" [> "\eSimpleY"]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:b}
-\figureCode{\hole{} + \hole{} * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    {},
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.28\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
-\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
-\end{tikzpicture}
-}%
-\label{fig:DifferentParts:c}
-\figureCode{\hole{} + u * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-  {},
-  "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
-  }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Different Parts of the Code.}%
-\label{fig:DifferentParts}
-\end{minipage}
-% \end{figure}
-% }
-%
-\hfil
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCommutativity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\figureCommutativity}{
-% \begin{figure}
-% \centering
-% \begin{figure}
-% \centering
-% \begin{tikzpicture}
-% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
-% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
-% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
-% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
-% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
-% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
-% \end{tikzpicture}
-% \caption{TODO:Commutativity}
-% \label{fig:Commutativity}
-% \end{figure}
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureNestedParts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{NestedPartsAlice}{L}
-\newvertex{NestedPartsAlice}{w}
-\newedge{NestedPartsBob}{L}
-% \newcommand{\figureNestedParts}{
-% \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
-\hfill
-%\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:a}
-\figureCode{\hole{} + w * v}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vSimpleTimes" [> "\eMoveTimes"] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:b}
-\figureCode{u * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
-    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
-\end{tikzpicture}
-}%
-\label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Users Editing Nested Parts of the Code.}%
-\label{fig:NestedParts}
-\end{minipage}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiChild
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiChildAlice}{R}
-\newvertex{MultiChildAlice}{x}
-\newedge{MultiChildBob}{R}
-\newvertex{MultiChildBob}{y}
-\newcommand{\figureMultiChild}{
-\begin{figure}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:a}
-\figureCode{w * v + x}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   {
-    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:b}
-\figureCode{w * v + y}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.4\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
-\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiChild:c}
-\figureCode{w * v + \conflictHole{x,y}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
-   },
-   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
-   {},
-   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
-\label{fig:MultiChild}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureMultiParent
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiParentAlice}{R}
-\newedge{MultiParentBob}{R}
-\newcommand{\figureMultiParent}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
-%%%%
-\path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
-    {}
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/a) to[star] (-2em, 0cm |- node/MultiParent/b);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"', >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/MultiParent/a) to[out=15,in=165,star] (-2em, 0cm |- node/MultiParent/c);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vNestedPartsAlice" [> "\eMultiParentBob", >bob edge]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/MultiParent/b) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\path [draw,->,merge step] (node/MultiParent/c) to[out=15,in=165] (-2em, 0cm |- node/MultiParent/d);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    b/"\vNestedPartsAlice" [> "\eMultiParentAlice"', >merge edge]
-   },
-   {}
-  }
- },
-};
-\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/MultiParent/e) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/d) to (-2em, 0cm |- node/MultiParent/e);
-\end{tikzpicture}
-}%
-\label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
-   v2/"\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vNestedPartsAlice" [> "\eMultiParentAlice"']
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
-\label{fig:MultiParent}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureCycle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{MultiCycleTimes}{L}
-\newvertex{MultiCycleTimes}{*}
-\newedge{MultiCyclePlus}{R}
-\newvertex{MultiCyclePlus}{+}
-\newedge{MultiCycleAliceTimes}{R}
-\newedge{MultiCycleAlicePlus}{L}
-\newedge{MultiCycleBobPlus}{R}
-\newedge{MultiCycleBobTimes}{L}
-\newcommand{\figureCycle}{
-\begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
-%%%%
-\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:a}
-\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:b}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
-     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:c}
-\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
-  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
-  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-   },
-   {
-    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
-     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
-     {}
-    }
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
-\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
-\end{tikzpicture}
-}%
-\label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + \conflictHole{\multiVertex{30},\multiVertex{32}}%
-  \otherVertexVskip\multiVertex{30} = \multiVertex{32} * \hole{}%
-  \\\multiVertex{32} = \multiVertex{30} * \hole{}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
-   },
-   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
-   {},
-   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
-  }
- }
-};
-\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
-\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-%\hfill{}
-\caption{Example of Cycles.}
-% TODO replace screenshots with text
-% TODO make ref code gray
-% TODO move whole figure left
-% TODO remove cursor from screenshots
-% TODO add multiparent box
-% TODO only show parens where needed - figure 10, 11 - by associativity
-% TODO repplace cycle screenshot with cycles box %
-\label{fig:Cycle}
-\end{figure*}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \figureDisconnect
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newedge{DisconnectAlice}{R}
-\newedge{DisconnectBob}{L}
-\newcommand{\figureDisconnect}{
-\begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
-%%%%
-\path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:a}
-\figureCode{\hole{} * \hole{} + \hole{} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'],
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/b) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Disconnect/a) to[star] (-2em, 0cm |- node/Disconnect/b);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:b}
-\figureCode{\hole{} * \parens{\hole{} * \hole{}} + \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    {},
-    "\vMultiCycleTimes" [> "\eDisconnectAlice", >alice edge]
-   },
-   {}
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/c) [anchor=base] {\strut};
-\path [draw,->,bob step] (node/Disconnect/a) to[out=15,in=165,star] (-2em, 0cm |- node/Disconnect/c);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:c}
-\figureCode{\hole{} + \parens{\hole{} * \hole{}} * \hole{}\otherVertexVskip\strut}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"] -> {
-   {},
-   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
-    "\vSimpleTimes" [> "\eDisconnectBob"', >bob edge],
-    {}
-   }
-  }
- }
-};
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill
-%%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
-\centering
-\caption{
-\begin{tikzpicture}[remember picture, overlay]
-\path (0cm,0cm) node (node/Disconnect/d) [anchor=base] {\strut};
-\path [draw,->,merge step] (node/Disconnect/b) to[out=15,in=165] (-2em, 0cm |- node/Disconnect/d);
-\path [draw,->,merge step] (node/Disconnect/c) to (-2em, 0cm |- node/Disconnect/d);
-\end{tikzpicture}
-}%
-\label{fig:Disconnect:d}
-\figureCode{\hole{} + \hole{}\otherVertexVskip\cycleVertex{2} = \hole{} * \parens{\cycleVertex{2} * \hole{}}}%
-\begin{tikzpicture}
-\path (0cm,0cm) graph[graph style] {
- root[root style] -> {
-  "\vWrapPlus" [> "\eWrapPlus"]
- }
-};
-\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
-\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
-\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
-\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
-\end{tikzpicture}
-\end{subfigure}
-%%%%%%%%
-\hfill{}
-\caption{Example of Disconnection.}%
-\label{fig:Disconnection}
-\end{figure*}
-}
-
-\newvertex{DecompPlusA}{+}
-\newedge{DecompTimesA}{L}
-\newvertex{DecompTimesA}{*}
-\newedge{DecompPlusAX}{R}
-\newvertex{DecompX}{x}
-\newedge{DecompTimesAX}{R}
-%
-\newedge{DecompPlusB}{L}
-\newvertex{DecompPlusB}{+}
-\newedge{DecompTimesB}{R}
-\newvertex{DecompTimesB}{*}
-\newedge{DecompY}{R}
-\newvertex{DecompY}{y}
-\newedge{DecompZ}{R}
-\newvertex{DecompZ}{z}
-%
-\newedge{DecompTimes}{L}
-%
-\newcommand{\figureDecompExample}{
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \path (0cm,0cm) graph[graph style] {
-        root[root style] -> {
-        a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
-          "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
-          {},
-          b/"\vDecompX" [> "\eDecompTimesAX"']
-          },
-          {}
-        }
-        },
-      };
-      \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
-      %
-      \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
-      \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
-      \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
-      \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
-      \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
-      \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
-      %
-      \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
-    \end{tikzpicture}
-    \caption{subcaption 1 goes here}%
-    \label{fig:Decomposition example graph}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{align*}
-      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
-      MP &= \{x^{\id{42}}\} \\
-      NP &= \{z^{\id{52}}\} \\
-      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
-    \end{align*}
-    \caption{subcaption 2 goes here}%
-    \label{fig:Decomposition example grove}
-  \end{subfigure}
-  \caption{Main caption goes here}%
-  \label{fig:Decomposition example}
-\end{figure}
-}
+\input{figure-simple}
+\input{figure-wrap-move}
+\input{figure-different-parts-nested-parts}
+\input{figure-multi-child}
+\input{figure-multi-parent}
+\input{figure-cycle}
+\input{figure-disconnect}
+\input{figure-decomposition}
 
@@
 
-\section{Introduction}%
-\label{sec:Introduction}
-
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
-
-Motivation:
-
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
-
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
-
-- previous approaches to collaborative editing have limitations
-
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
-
-- operational transforms (complexity, you have to patch previous actions based on new actions)
-
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
-
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
-
-this is tricky because:
-
-- some edits might be conflicting -- solve this with "conflict holes"
-
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
-
-- since we are commutative, we solve both synchronus and async collaborative editing
-
-- and this resolves issues around merges and conflicts
-
-- contribution of this paper is to solve these problems from type-theoretic first principles:
-
-- ...
-
-- Hazel
-
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
-
-\section{Grove By Example}%
-\label{sec:Grove By Example}
-
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
-
-\subsection{Program Representation}%
-\label{sub:Program Representation}
-
-\figureSimple{}
-
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
-
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
-
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
-
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
-
-Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
-
-\figureWrapMove{}
-% \figureMove{}
-
-\subsection{Single-User Actions}%
-\label{sub:Single-User Actions}
-
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
-
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
-
-\subsubsection{Construction Actions}%
-\label{sub:Construction}
-
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
-
-\subsubsection{Deletion Actions}%
-\label{sub:Deletion}
-
-% \figureMove{}
-
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
-
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
-
-\subsubsection{Wrapping Actions}%
-\label{sub:Wrapping}
-
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
-
-\subsubsection{Repositioning Actions}%
-\label{sub:Repositioning}
-
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
-
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
-
-% \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
-
-%\figureCommutativity{}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
-    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
-    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
-  \end{tikzpicture}
-  \caption{Commutativity of Edits.}
-  \label{fig:Commutativity}
-\end{figure}
-
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
-
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
-
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
-
-\subsubsection{Editing Different Parts of the Code}%
-\label{sub:Editing Different Parts of the Code}
-
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
-
-% \figureNestedParts{}
-
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
-\label{sub:Editing Nested Parts of the Code}
-
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
-
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
-
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
-
-\subsection{Merge Conflicts}%
-\label{sub:Merge Conflicts}
-
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
-
-\subsubsection{Multi-child conflicts}%
-\label{sub:Multi-child conflicts}
-
-\figureMultiChild{}
-
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
-
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
-
-TODO: "delete both"
-
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
-
-\subsubsection{Multi-parent conflicts}%
-\label{sub:Multi-parent conflicts}
-
-\figureMultiParent{}
-
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
-
-TODO: talk about how to resolve the conflict
-
-% \subsubsection{Single-User Cycles}
-% \label{sub:Cycles}
-
-% TODO: cut this section
-
-% \figureSingleUserCycles{}
-
-% Another case we must consider is when cycles appear in the graph.
-% We categorize these into cycles caused by the action of a single user
-% and cycles caused by the interaction of the actions of multiple users.
-
-% In a single user context,
-% normal insertion and deletion of code by the user cannot create cycles.
-% However, it is possible with certain kinds of copy and paste.
-% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
-% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
-% There are two ways to interpret the paste action.
-% The first interpretation is to create a deep copy of \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b} and
-% does not cause a cycle.
-% The second interpretation is to simply add an edge to \Vertex{TODO}.
-% This results in \autoref{fig:Single-User Cycles:b}
-% and causes a cycle.
-
-% Note that not all pastes should be deep copies.
-% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
-% by a cutting from the old position and pasting to the new position.
-% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
-% Distinguishing when a paste should be by reference versus by copy
-% is ultimately a user interface question.
-% Cycles caused by the local user's edits can be detected as soon as a user enters them
-% by noting either that the graph would contain a cycle or the vertex
-% already has a parent somewhere in the graph.
-
-% Thus, as a user interface consideration, it might be best to either
-% disallow such edits to at least warn users when their
-% edits would create a cycle.
-
-\subsubsection{Cycles}%
-\label{sub:Multi-User Cycles}
-
-\figureCycle{}
-
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
-For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
-in \autoref{fig:Cycle:d}, which has a cycle
-between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
-
-\paragraph{Disconnected Cycles}
-
-\figureDisconnect{}
-
-TODO: technically a cycle
-
-TODO: in general detect when there are or were edits to something now deleted
-
-\todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
-
-\subsection{Cursors}%
-\label{sub:Cursors}
-
-\paragraph{Representation}
-
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
-
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
-
-\paragraph{Communication}
-
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
-
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
-
-\section{Formalism}%
-\label{sec:Formalism}
-
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
-
-TODO: point of this section is commutativity theorem
-
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
-
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
-
-\subsection{Syntax}%
-\label{sub:Syntax}
-
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
-
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
-
-% TODO: replace positions with macros
-
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ \Root \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
-
-\subsection{Convergent Graphs}%
-\label{sub:Convergent Graphs}
-
-\subsubsection{Graphs}%
-\label{sub:Graphs}
-
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
-
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
-
-% TODO: should graph "recomposition" have an algorithm, too?
-A graph can be derived from an expression in the calculus by starting with an
-empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
-expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
-constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
-= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
-to the vertices representing their immediate sub-terms until only empty holes
-remain. Note that empty holes are not modeled explicitly in the graph. Wherever
-an empty hole appears in an expression, the corresponding graph position has no
-edges originating from it.
-
-% You can map from syntax to graph by selecting
-% unique IDs $u$ (for both edges and vertexes).
-% - holes are not explicit in the graph
-
-\subsubsection{Graph Edits}%
-\label{sub:Graph Edits}
-
-We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
-define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
-
-We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
-$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
-\sqcup \bot = \Plus$ means that edges can only be created if they do not already
-exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
-an edge is deleted it can never be restored.
-
-We define the semantics of graph actions via the following transition relation
-between graphs.
-\[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
-\]
-Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
-updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
-state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
-$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
-$G$.
-
-% Graph action semantics is a transition system between graphs.
-% (join with whatever it was before)
-
-% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
-% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
-
-\subsection{Commutativity}%
-\label{sub:Commutativity:formal}
-
-In order to define commutativity, we first define a lattice over $G$.
-Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
-$G \overset{\alpha}\rightarrow G'$
-iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
-Thus the commutativity of graph actions is established
-by the commutativity of the join operations defining those graph actions.
-
-Discuss uniqueness of uuid
-
-\begin{lemma}[Join Commutativity]
-  \label{lem:Join Commutativity}
-\end{lemma}
-
-\begin{lemma}[Joining]
-  \label{lem:Joining}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
-  $G \overset{\alpha}{\longrightarrow} G'$
-  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
-\end{lemma}
-\begin{proof}
-  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
-\end{proof}
-
-\begin{theorem}[Commutativity]
-  \label{thm:Commutativity}
-  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
-  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
-  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{theorem}
-\begin{proof}
-  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
-  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
-  Then by \autoref{lem:Join Commutativity},
-  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
-  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
-  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
-\end{proof}
-
-\subsubsection{Agda Mechanization}%
-\label{sub:Agda Mechanization}
-
-\subsection{Graph Decomposition}%
-\label{sub:Graph Decomposition}
-
-% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
-% and multiparents
-
-% NOTE: this whole section is for describing the 4-tuple
-
-Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
-
-\figureDecompExample
-
-% TODO: make sure we're not claiming to invent the term "unicycular graph"
-% TODO: maybe add a forward reference at end here to Implementation section
-% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
-
-A grove is a 4-tuple of the following form:
-\[
-  (e_r, MP, NP, W)
-\]
-where
-\begin{align*}
-  e_r &= \expr(v_0) \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
-\end{align*}
-For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
-given in \autoref{fig:Decomposition example grove}.
-
-Let us now examine each of the four components in more detail.
-The first component $e_r$ is the expression derived from the root vertex $v_0$.
-\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
-There is one rule for each constructor form.
-For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
-An empty hole $\hole$ appears if there are no children at the given position.
-A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
-For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
-
-The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
-Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
-In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
-
-The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
-In the example, $z^{\id{58}}$ has no parents (because it was deleted), so it appears only in $NP$.
-
-The fourth component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
-These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
-Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
-Breaking the cycle turns a unicyclic graph into a tree.
-% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
-% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
-To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
-only a vertex on the cycle can be its own ancestor.
-When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
-In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
-
-% XXX
-
-decomposition produces sets of trees. As long as everything is well sorted, we
-can use them to produce expressions in the main grammar. (alt: define wrt the
-graph instead, once we establish the graph can be turned into expressions)
-
-*** well-sortedness judgment on trees: the tree only has valid positions for the
-constructors, and the child at each position has the sort corresponding to its
-position, and edges in position maps all start at the given constructor.
-
-Theorem: well-sorted graphs produce decompositions of well-sorted trees
-
-``A graph has this property iff these conditions hold''
-
-then we can define a total mapping from well-sorted trees to expressions
-(because the tree structure is very generic, so this ties them back to the
-earlier grammar.)
-
-forest = 4-tuple
-
-forest validity: each of the trees in the forest are well sorted, and all Refs
-point to roots of other trees of the correct sort.
-
-every valid forest corresponds to a well behaved grove (i.e., ``expression
-forest'')
-
-soundness: if you start with a valid graph, you get a well behaved grove that
-connects every vertex (and only those vertices) in the graph. (i.e., if you turn
-it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
-Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
-add Root sort
-
-OR:
-
-use edge IDs instead of edges and build a mapping from vertices to
-sorts---change the $\e$s to $u$s)
-
-% TODO: maybe add a forward reference here to Implementation section
-
-\subsection{User Actions}%
-\label{sub:User Actions}
-
-% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
-% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
-
-move = delete + add
-
-other composite edits
-
-remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
-
-given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
-This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
-(we probably want to do it this way in Hazel)
-
-OR
-
-We could go the indirect route and avoid having to prove the theorem.
-(this way is easier to start with, and it's how GRV works)
-
-% Mapping to b actions
-
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent
-
-\section{Implementation}%
-\label{sec:Implementation}
-
-% TODO: talk about code generation somewhere in here
-
-% TODO: talk about:
-% - decomp vertex set generation and unicycle traversal
-% - how to choose which branch of expr' to take
-% - lfp(ancestors') and min of it
-
-% TODO: can we save any of this in a README or code comments?
-
-% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)
-
-GRV -- how it is implemented, how it connects to the formalism, describe the graph,
-
-We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.
-
-Compared to the core language, our implementation lacks support for lists and case expressions.
-
-Optimizations: least fixed point
-
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
-
-% here are some properties that show how to get one from the other, and to/from
-% graph
-
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
-
-- linear scan
-- breaking out roots
-- ...
-
-Graph decomposition occurs in three steps.
-
-First, a linear scan is performed over every edge that has been created or
-deleted. The vertices at both ends are partitioned into three sets:
-multi-parented, single-parented, or orphaned. Maps for O($\log n$) parent and
-child edge set lookups are also created.
-
-Second, the subgraphs rooted at vertices with more or less than one parent are
-traversed and equivalent tree structures are generated over components that are
-connected by single-parented vertices. Cycles are broken with UUID vertex
-references. Single-parented vertices reachable from the root vertex are also
-traversed.
-
-Finally, once we've already accounted for things with more or less than one
-parent, everything else must be single parented. We call these things unicycular graphs. A
-unicycular graph is a cycle of single-parented vertices that each may have single-parented
-descendants.
-
-% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
-% TODO: maybe show this figure and say that it's possible to handle generically without code generation
-
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
-
-lemma: a unicycle is a cycle with trees hanging off of it
-
-lemma: unicycle traversal always produces a valid tree
-
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
-
-\section{Related Work}%
-\label{sec:Related Work}
-
-Contrast with a traditional diff
-
-Similar to how git merged across moves.
-A notable difference though is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
-
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
-
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
-
-\subsection{Trees}
-
-\subsubsection{Structure editors}
-
-Hazel~\citep{Omar:2019:10.1145/3290327}
-
-\subsubsection{Tree diff}
-
-We are using action based instead of tree diff
-
-We have to deal with merge
-
-We have to deal with edits from multiple people
-
-J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.
-
-Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}
-
-  Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}
-
-  An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}
-
-  Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}
-
-An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}
-
-  Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}
-
-  A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}
-
-  Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}
-
-  Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}
-
-  Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}
-
-  A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}
-
-  Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}
-
-  The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}
-
-  The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}
-
-  Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}
-
-\subsection{Version Control}
-
-Git \url{https://git-scm.com/}
-
-Darcs \url{https://darcs.net/}
-
-  Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}
-
-Hg? \url{https://www.mercurial-scm.org/}
-
-SVN \url{https://subversion.apache.org/}
-
-Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)
-
-  \url{https://pijul.org/}
-  \url{https://pijul.org/manual/theory.html}
-
-  \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}
-
-\subsection{Collaborative Editing}
-
-Collaborative Structure Editing
-
-SmallTalk collaboration with images
-
-TouchDevelop papers
-
-Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)
-
-\subsection{CRDTs}
-(Are we a known CRDT?)
-
-List of CRTD papers: \url{https://crdt.tech/papers.html}
-
-Bottom = Tombstone
-
-https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read
- - Add-Remove Partial Order data type
- - 2P2P-Sets
- - Replicated Growable Array
-
-\url{https://github.com/PsychoLlama/graph-crdt}
- - Graph CRDT
- - Uses a LWW-E-Set
-
-\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
- We don't have interleaving problems because
-   - everything is relative to a specific ID not a position
-   - we don't try to auto resolve
-   - we are tree not list
-   Part three: moving sub tree
-   - Last parent writter wins (prevents cycles)
-     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?
-
-A commutative replicated data type for cooperative editing
-  \url{https://hal.inria.fr/inria-00445975/document}
-  Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)
-
-Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
-  \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}
-
-Specification and Complexity of Collaborative Text Editing
-  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}
-
-LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}
-
-Data consistency for P2P collaborative editing
-  \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}
-
-Interleaving anomalies in collaborative text editors
-  \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}
-
-Moving Elements in List CRDTs
-  \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}
-
-A highly-available move operation for replicated trees and distributed filesystems
-  \url{https://martin.kleppmann.com/papers/move-op.pdf}
-
-...
- commutative replicated data types CmRDT
- convergent replicated data types, or CvRDTs
- Delta state CRDTs[12][13] (or simply Delta CRDTs
-
-The Causal Graph CRDT for Complex Document Structure
-  \url{https://dl.acm.org/doi/10.1145/3209280.3229110}
-
-\url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}
-
-G-Set
-PN-Set
-2P-Set
-LWW-element-Set (Last-Write-Wins)
-OR-Set
-MV-Register: Multi-Value Register
-U-Set (this is what we are for edges, not OR-set due to duplicates)
-Add-Remove Partial Order data type
-  2P-Set for vertices, and a G-Set for edges.
-
-A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
-  \url{https://hal.inria.fr/inria-00555588/document}
-
-CRDTs: Consistency without concurrency control
-  \url{https://arxiv.org/abs/0907.0929}
-
-https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
- removeVertex() has priority, all incident edges are removed
- addEdge() has priority, all removed vertices are re-added
- Delay removeVertex() execution till all concurrent removeVertex() are executed.
-First one is 2P2P-Set
-
-https://crdt.tech/papers.html
-
-Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster.
-Co-design and verification of an available file system.
-In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381.
-Springer LNCS volume 10747, January 2018.
-[ bib | DOI | .pdf ]
-  \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
-  https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf
-
-Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
-  http://dx.doi.org/10.1109/TPDS.2017.2697382
-  http://arxiv.org/abs/1608.03960
-
-Vinh Tao, Marc Shapiro, and Vianney Rancurel.
-Merging semantics for conflict updates in geo-distributed file systems.
-In 8th ACM International Systems and Storage Conference, SYSTOR 2015.
-ACM, May 2015.
-[ bib | DOI | .pdf ]
-  http://dx.doi.org/10.1145/2757667.2757683
-  https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf
-
-Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
-  http://arxiv.org/abs/1207.5990
-  https://hal.inria.fr/hal-00720681/
-
-Stéphane Martin, Pascal Urso, and Stéphane Weiss.
-Scalable XML collaborative editing with undo.
-In On the Move to Meaningful Internet Systems (OTM), pages 507--514.
-Springer LNCS volume 6426, October 2010.
-[ bib | DOI | arXiv ]
-  \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
-  http://arxiv.org/abs/1010.3615
-
-2P-Sets
-Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge
-
-Operational Transforms
-
-Etherpad
-
-Live Share
-
-\subsection{Synchronization}
-
-Unision
-\url{https://www.cis.upenn.edu/~bcpierce/unison/}
-\url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}
-
-\subsection{TODO}
-
-Homotopical Patch Theorey: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
-Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
-
-\section{Discussion and Conclusion}%
-\label{sec:Discussion and Conclusion}
-
-TODO: leaves
-
-TODO: our model supports treating these as a cons-list of characters
-
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
-
-TODO: Traditional diff: no relation (and indent might change)
-
-TODO: memory usage
-
-TODO: cache eviction algorithm
-
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
-
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
-
-NOTE: we could implement each digit as a separate characters
-
-GUI for string conflicts: use popups
+\input{grove-paper-introduction}
+\input{grove-paper-collaborative-structure-editing}
+\input{grove-paper-formalism}
+\input{grove-paper-implementation}
+\input{grove-paper-related-work}
+\input{grove-paper-discussion-and-conclusion}
 
@@
 
-\section{Supplemental Material: Complete Graph Sequences for All Figures}%
-\label{apx:Supplemental Material: Complete Graph Sequences for All Figures}
-
-TODO: put all intermediate states for all graphs along with the edge actions
+\input{appendix-graph-sequences}
+\input{appendix-formalism}
 

### 2021-05-03 grove-paper-1-introduction.tex
make file structure match paper structure

\section{Introduction}%
\label{sec:Introduction}

TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.

Motivation:

- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
is good and important as computing grows

- semantic structure editing is good because it solves the gap problem (semantic editor services
are always available) -- cite Hazelnut papers (talk about holes)

- previous approaches to collaborative editing have limitations

- diff/merge based approaches (trying to solve the inverse problem based on final states --
you lose the actual actions that were performed, and have to reconstruct them or an approx.
of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
some papers have started to look at that, but fundamentally we don't want to throw away the
knowledge we have about the edits!)

- operational transforms (complexity, you have to patch previous actions based on new actions)

- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
because it is relatively simple: you just send all the edits to all the replicas and they are
convergent by design

- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
the semantic editor services will be disabled because the program is going to be broken in multiple places
transiently)

this is tricky because:

- some edits might be conflicting -- solve this with "conflict holes"

- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)

- since we are commutative, we solve both synchronus and async collaborative editing

- and this resolves issues around merges and conflicts

- contribution of this paper is to solve these problems from type-theoretic first principles:

- ...

- Hazel

\subsection{Contributions and Paper Organization}%
\label{sec:Contributions and Paper Organization}

### 2021-05-03 grove-paper-2-grove-by-example.tex
make file structure match paper structure

\section{Grove By Example}%
\label{sec:Grove By Example}

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

\subsection{Program Representation}%
\label{sub:Program Representation}

\figureSimple{}

The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
This can be presented to the user as an expression containing holes.
For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
has a hole on the right side of the times operator.

Program state is a directed graph with a distinguished root vertex.
Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively. (For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
  for the purposes of this paper, we represent identifiers and number literals
  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.

Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

\figureWrapMove{}
% \figureMove{}

\subsection{Single-User Actions}%
\label{sub:Single-User Actions}

In the remainder of this section,
we consider user actions, starting
in this subsection with single-user
actions and continuing in \autoref{sub:Multi-user Interactions}
with multi-user actions.

Each user action corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

User actions act relative to a cursor.
We do not model cursor sharing and the cursor does not appear in the underlying graph.
We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.

\subsubsection{Construction Actions}%
\label{sub:Construction}

To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.

\subsubsection{Deletion Actions}%
\label{sub:Deletion}

% \figureMove{}

Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
In our system, once a vertex is created it is never deleted.
This allows further manipulations of those vertices by other users as shown later in this section.
Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.

Once an edge with a particular identifier is deleted, it cannot be recreated.

Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.

\subsubsection{Wrapping Actions}%
\label{sub:Wrapping}

Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
(The choice of wrapping with a bias for the left child position is arbitrary.)

\subsubsection{Repositioning Actions}%
\label{sub:Repositioning}

If a user wants to reposition code from one place to another, we have two options.
The first is to delete the code from its old position, and
reconstruct the code in its new position.
However, edges and vertices have unique identifiers in Grove
to support the collaborative editing features we discuss next,
so reconstruction will not preserve identity of vertices and edges.
Consequently, we take the second option: explicit repositioning actions that preserve identity.
These actions would be triggered by, for example, a drag and drop, or a cut and paste
(but not copy and paste, or a second paste after a cut, which do not need to
preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).

In terms of the graph, a repositioning action involves simply deleting the edge
from the original parent then adding an edge from the new parent.
Thus the code is actually \textit{repositioned}, not copied.

For example, in \autoref{fig:Move} Alice continues by
repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.

% \figureDifferentParts{}
\figureDifferentPartsNestedParts{}

%\figureCommutativity{}
\begin{figure}
  \centering
  \begin{tikzpicture}
    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
    \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
    \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
  \end{tikzpicture}
  \caption{Commutativity of Edits.}
  \label{fig:Commutativity}
\end{figure}

\subsection{Multi-user Interactions}%
\label{sub:Multi-user Interactions}

Having described single-user actions,
we turn our attention to how Grove handles multiple users.
This section generalizes to any number of users,
but for simplicity we consider only two in our examples: Alice and Bob.
Alice and Bob each maintain their own editor instances and perform
actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).

Users coordinate by sharing with one another the graph edits corresponding to each
action~(e.g., via direct network connections or relayed through a server).
TODO: CITE(how to share network data -- DHT? consensus? ).
The granularity of communication may vary.
For real-time collaborative editing, edits may be sent as they occur.
For version control, edits may be batched together in the form of commits.
As we will show, these edits are commutative, so editors converge to the same state
when all edits are shared.
We assume users have some way to create globally unique identifiers for
verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).

\subsubsection{Editing Different Parts of the Code}%
\label{sub:Editing Different Parts of the Code}

We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
Before transmitting their edits to each other,
Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
and \autoref{fig:DifferentParts:b}, respectively.
(Note that the transition edge from \autoref{fig:Move:b} to
\autoref{fig:DifferentParts:b} represents
multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
We thus mark it with a star.)

Alice and Bob then transmit their graph edits to each other
and apply the other's graph edits to their own copy of the graph.
This results in both Alice and Bob having the graph in
\autoref{fig:DifferentParts:c}.

% \figureNestedParts{}

\subsubsection{Commutativity}%
\label{sub:Commutativity:informal}

\todo{rev}
Note that aside from transmitting edge deletions and additions,
merging requires no coordination between Alice and Bob.
For example, suppose Alice and Bob did not transmit their edits to each other
but instead to a third user, Chris.
Chris can apply the edits from Alice and Bob in any order and will always get
the graph in \autoref{fig:DifferentParts:c}.

In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
For any sequences of edits from Alice and Bob,
applying them in any order always produces the same results.
(We show this formally in \autoref{sub:Commutativity:informal}.)

\subsubsection{Editing Nested Parts of the Code}%%
\label{sub:Editing Nested Parts of the Code}

After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
Since edits are based on edge identifiers and not locations in a tree,
both Alice's and Bob's edits can be applied to the other's graph,
which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.

Commutativity has consequences for this scenario.
Commutativity require that edits produce the same result regardless of the order in which they are applied.
So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
changes before receiving Bob's addition of~\eNestedPartsBob{},
the result must still be the graph in \autoref{fig:NestedParts:c}.
In that case, her edits would then temporarily be within the code deleted by Bob.
Thus, our semantics must allows users to edit deleted code.
(Whether or how this is presented in the user interface is another question.)

TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.

\subsection{Merge Conflicts}%
\label{sub:Merge Conflicts}

While the collaborative edits discussed so far merge cleanly,
they can also generate graphs that do not map cleanly to a syntax tree.
We call these conflicts and consider cases for \emph{multi-child},
\emph{multi-parent}, and \emph{cycle} conflicts.

As with conflict in line/diff-based version-control
system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
these all require user intervention to resolve.

We consider common merge scenarios first, and then consider more unusual scenarios.

\subsubsection{Multi-child conflicts}%
\label{sub:Multi-child conflicts}

\figureMultiChild{}

As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.

We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
as edges going out of the \texttt{R} position of \vWrapPlus{}.
This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.

TODO: "delete both"

This conflict can be resolved by the user by deleting
one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
(Note that users can continue to edit the code while in this conflicted state.
It just must be resolved before the the code is run.\footnote{It
  may be possible to develop evaluation models that allow these sorts of conflicts,
  but that is beyond the scope of this paper.})

TODO: need example of fully-automatic resolution and semi-automatic resolution.
Note that as a convenience to the user, certain simple
conflicts might be automatically resolved,
but we consider this a higher-level, user-interface consideration.

This handling of conflicts is similar traditional merge conflicts in that
it presents alternate versions of the code that the user replaces with the desired merge result.
Our model is different in that merge conflicts in the line-based model
prevent semantic editor services from running
that could continue to run in the graph-based model because
code inside and outside the conflict is still valid code.
It is just the where the multi-child happens that editor services may break down.
\autoref{sec:Related Work} discusses more differences.

\subsubsection{Multi-parent conflicts}%
\label{sub:Multi-parent conflicts}

\figureMultiParent{}

While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
An example of this is in \autoref{fig:MultiParent}.
\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
and Alice and Bob shared their edits with each other.
Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
In both cases, these edits are achieved with one edge delete and one edge add.
Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.

TODO: talk about how to resolve the conflict

% \subsubsection{Single-User Cycles}
% \label{sub:Cycles}

% TODO: cut this section

% \figureSingleUserCycles{}

% Another case we must consider is when cycles appear in the graph.
% We categorize these into cycles caused by the action of a single user
% and cycles caused by the interaction of the actions of multiple users.

% In a single user context,
% normal insertion and deletion of code by the user cannot create cycles.
% However, it is possible with certain kinds of copy and paste.
% For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles:a}
% and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
% There are two ways to interpret the paste action.
% The first interpretation is to create a deep copy of \Vertex{TODO}.
% This results in \autoref{fig:Single-User Cycles:b} and
% does not cause a cycle.
% The second interpretation is to simply add an edge to \Vertex{TODO}.
% This results in \autoref{fig:Single-User Cycles:b}
% and causes a cycle.

% Note that not all pastes should be deep copies.
% For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
% by a cutting from the old position and pasting to the new position.
% Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
% Distinguishing when a paste should be by reference versus by copy
% is ultimately a user interface question.
% Cycles caused by the local user's edits can be detected as soon as a user enters them
% by noting either that the graph would contain a cycle or the vertex
% already has a parent somewhere in the graph.

% Thus, as a user interface consideration, it might be best to either
% disallow such edits to at least warn users when their
% edits would create a cycle.

\subsubsection{Cycles}%
\label{sub:Multi-User Cycles}

\figureCycle{}

In addition to multi-child and multi-parent conflicts, merges
can cause cycles in the graph.\footnote{In theory a single user could
generate cycles, but this could be detected immediately and the user action
that caused that cycle rejected.
Thus, we are only concerned with cycles that cannot be immediately detected
and prevented by the editor.}
For example, consider the situation in \autoref{fig:Cycle}.
\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
and shared them with each other.
Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
and then~\vMultiCyclePlus{} underneath that.
Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.

On their own, neither of these edits creates a cycle.
However, merging the edit actions of both Alice and Bob results in the graph
in \autoref{fig:Cycle:d}, which has a cycle
between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
As with the other kinds of conflict, resolving it requires user input.
For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).

\paragraph{Disconnected Cycles}

\figureDisconnect{}

TODO: technically a cycle

TODO: in general detect when there are or were edits to something now deleted

\todo{UI would show orphans only once edits orphans occur}

Finally, there is a particular type of cycle worth discussing.

Since a cycle like this that is connected to the root vertex always
contains vertexes with multiple parents that if removed would break the cycle,
we use the user interface described in \autoref{sub:Multi-parent conflicts}
to display these kinds of programs.
For example, \autoref{fig:Cycle:d} would be displayed.
We leave the user interface considerations of resolving the conflict such a
cycle represents to future work.

Finally, when multi-user edits are merged,
parts of the graph can become disconnected from the root
even though they are connected to the root in each user's
copy of the graph before the merge.

For example, consider the situation in \autoref{fig:Disconnect:a},
which shows the graph after Alice and Bob have made some edits,
and shared their edits with each other.
Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.

When these edits are merged, the result is \autoref{fig:Disconnect:d}.
Since Alice deleted~\eNestedPartsBob{}
and Bob deleted~\eMultiCycleAliceTimes{},
when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
are disconnected from the root.

This is similar to the type of merge conflict
that can arise in traditional version control systems
when two users delete different functions that
are duplicates of the other.

Fortunately, we can detect this situation and warn the user
since the disconnection appears after merging two graphs
in which vertexes were not already disconnected.

\subsection{Cursors}%
\label{sub:Cursors}

\paragraph{Representation}

For most editing a cursor is represented by a position within a vertex.
For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
Representing it this way means that if Bob or some other user deletes TODO:vertex
and replaces it with a different vertex, Bob's cursor is still valid.
We specify a position within a vertex instead of a vertex because TODO.

However, there is one case where we need a more precise cursor.
That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
We want users to beable to put their cussor on either the conflict as a whole
(e.g., TODO: \texttt{\{ x | y \}})
or on an individual element (e.g., \texttt{x} or \texttt{y}).
In the former case, a pair of a vertex and a child position suffices.
For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).

\paragraph{Communication}

In our system all edits use edge and vertex identifiers, and user
cursors do not affect the interpretation of graph edits.
Thus, cursors need not be communicated to other users.
However, in a collaborative setting, seeing the cursors of other users
can be useful.
For this purpose, editors can announce their cursor position to other editors.
At any point in time, the announcement
with the most recent timestamp
for a particular user
is used.

Note that this cursor position may refer to edges or verticies that
do not yet exist on the receivers machine.
In these cases, we can either not display the other user's cursor (and perhaps
have a visual display flagging this fact), or we can display the most recent
cursor that represents a valid position in the local graph (perhaps
shown in a fadded color to show that this cursor is known to not be up to date).
(Or perhaps, all cursors always fade/decay over time like on radar blips.)

### 2021-05-03 grove-paper-3-formalism.tex
make file structure match paper structure

--- 2021-04-09 grove-paper-formalism.tex
@@
 \[
-  (NP, MP, U)
+  (NP, MP, U, D)
 \]
@@
   MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
+  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\} \\
+  D  &= \{\e : G(\e) = \Minus\}
 \end{align*}
@@
 In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% THEOREMS GO HERE
-
-% - [ ]  State the "exhaustiveness check" theorem formally
-
-% - [ ]  State the "every vertex is part of decomposition" theorem formally
-
-% i.e., {univ, multiv, av} is a partitioning of the vertices of G
-% where, e.g., multiv = vertices(MP)
-
-% - [ ]  Define "recomposition" formally
-
-% not defined yet
-
-% - [ ]  State the relationship between decomp(G) and recomp(grove)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-  \begin{displaymath}
-    \begin{array}{l}
-      \expr(v=(u, Root)) = expr'(v, Root) \\
-      \expr(v=(u, \ExpVar(x))) = x^{\id{u}} \\
-      \expr(v=(u, \ExpLam)) = \lambda^{\id{u}} \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
-      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^{\id{u}} \\
-      \expr(v=(u, \ExpNum(n))) = n^{\id{u}} \\
-      \expr(v=(u, \ExpPlus)) = \expr'(v, \PlusLeft)~\texttt{+}^{\id{u}}~\expr'(v, \PlusRight) \\
-      \expr(v=(u, \ExpTimes)) = \expr'(v, \TimesLeft)~\texttt{*}^{\id{u}}~\expr'(v, \TimesRight) \\
-      \patt(v=(u, \PatVar(x))) = x^{\id{u}} \\
-      \type(\tau=(u, \TypArrow)) = \type'(\tau, \ArrowArg) \rightarrow^{\id{u}} \type'(\tau, \ArrowResult) \\
-      \type(\tau=(u, \TypNum) = Num^{\id{u}} \\
-    \end{array}
-  \end{displaymath}
-
-% TODO: comment these out:
-%   % | Exp_nil
-%   % | Exp_cons
-%   % | Exp_case
-%   % | Typ_list
-
-  \begin{align*}
-    \expr'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\expr(v_1),\ldots,\expr(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \expr(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \patt'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\patt(v_1),\ldots,\patt(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \patt(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \type'(v,p) = \begin{cases}
-      \hole & \children(v,p) = \varnothing \\
-      \conflictHole{\type(v_1),\ldots,\type(v_n)} & \children(v,p) = \{v_1, \ldots, v_n\} \\
-      \multiVertex{u} & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
-      \cycleVertex{u} & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
-          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
-      \type(v') & \text{otherwise} \\
-    \end{cases}
-  \end{align*}
-
-  \begin{align*}
-    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
-    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
-    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
-    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
-    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
-      \end{align*}
-  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
-  \label{fig:Decomposition expr}
-\end{figure}
 
@@
 % \]
-
-\section{Type System}%
-\label{sec:Type System}
-
-Sensibility -- mapping back from graphs
-to well-typed states that fixes holes
-
-TODO: somewhere in here, maybe multiple places, we want to talk about typing the intermediate states that arise (background for structured editing)
-
-editor states:
-
-conflicts (multiple children in the same spot)
-
-multi-parent
-
-cycles
-
-non-empty holes
-
-----------------
-
-User must supply edge selection for multi-parent

### 2021-05-11 grove-paper-3-formalism.tex
move term syntax and arity figures into new files and deduplicate
rename expr, patt, type to erecomp, precomp, trecomp
add decomp back to main formalism section (deduped)
make decomp return a pair and recomp accept a pair
finish decomp / recomp edge tracking

@@
 
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{lrlll}
-      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
-           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
-    \end{array}
-  \]
-  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
-  \label{fig:Base syntax}
-\end{figure}
+\figureTermSyntax
 
@@
 
-\begin{figure}
-  \[
-    \arraycolsep=0pt
-    \begin{array}{ll}
-      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
-      \hline
-      \arity(\Root)={} & \left\{ (\Root, Exp) \right\} \\
-      \arity(\PatVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpVar(x))={} & \left\{ \right\} \\
-      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
-      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
-      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
-      \arity(\ExpNum(n))={} & \left\{ \right\} \\
-      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
-      \arity(\TypNum)={} & \left\{ \right\} \\
-    \end{array}
-  \]
-  \caption{Constructors, Indexes and Arity}
-  \label{fig:Constructors, Indexes and Arity}
-\end{figure}
+\figureArity
 
@@
 \[
-  (NP, MP, U, D)
+  (NP, MP, U)
 \]
@@
 \begin{align*}
-  % e_r &= \expr(v_0) \\
-  NP &= \{\expr(v) : |\parents(v)| = 0\} \\
-  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
-  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\} \\
-  D  &= \{\e : G(\e) = \Minus\}
+  \figureDecompositionDefDecompComponents
 \end{align*}
@@
 given in \autoref{fig:Decomposition example grove}.
+
+\figureDecompositionDef
+
+\figureDecompositionDefHelpers
 
@@
 To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
-Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
+Formally, we choose the vertex $v$ such that $v = \min{\ancestors{v}}$ because
 only a vertex on the cycle can be its own ancestor.

### 2021-07-27 grove-paper-3-formalism.tex
fix well sortedness defs
add source annotation to \conflictHole

@@
 
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}$).
+The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}{v}{p}$).
 In particular, the argument of a lambda can be either a bound variable or a hole.
@@
 \in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arity$ function in
+destination vertices may originate. The $\arityOp$ function in
 \autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
@@
 \[
-  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
+  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right]
 \]

### 2021-07-27 grove-paper-5-implementation.tex
add source annotation to \conflictHole

\section{Implementation}%
\label{sec:Implementation}

% TODO: talk about code generation somewhere in here

% TODO: talk about:
% - decomp vertex set generation and unicycle traversal
% - how to choose which branch of expr' to take
% - lfp(ancestors') and min of it

% TODO: can we save any of this in a README or code comments?

% TODO: talk about performance: O(decomp) given "size of the graph" (probably linear in # edges)

% GRV -- how it is implemented, how it connects to the formalism, describe the graph,

We implemented the core Grove calculus of \autoref{sec:Formalism} as an OCaml library.
Our code is parameterized by a syntax specification for expressions that we generate automatically, which makes the implemented language of expressions easier to modify or replace.
We also built a Web browser-based Grove editor as a proof of concept.

Since we cannot implement an infinite mapping directly, the graph is a data structure, \texttt{Graph.t}, that maps from edges to the edge states $\left\{\Plus, \Minus\right\}$.
We do not represent edges that map to $\bot$.

% TODO: rename Decomposition.ml to Grove.ml

There's a function \verb!decompose! of type \verb!Graph.t) : t * Edge.Set.t Vertex.Map.t! that

From this graph, we compute the set of all edges and the set of live edges with a linear scan.

% TODO: make this into a listing

\begin{verbatim}
  Graph.edges : Graph.t -> Edge.Set.t
  Graph.live_edges : Graph.t -> Edge.Set.t
\end{verbatim}

% TODO: keep connecting the code to the formalism like this

These vertices are partitioned into three sets: multi-parented, single-parented, or orphaned.
A vertex $v$ is considered a parent of another vertex $v'$ if there is a live edge from $v$ to $v'$.

For each vertex, we take the parent and child edge sets.

% TODO: fix the formalism for parents (to work with edges instead of vertices) and then check that it matches the implementation. Try renaming parents to vparents and adding parents for edges.

% TODO: Talk about how the graph is implemented (not as an actual function, but as a map from edges to edge states) so that we can denote the set of created + deleted edges clearly, as well as all of the vertices that have ever been created.

For a graph $G : \E \to \Sigma$ where $\E = \U \times \V \times \P \times \V$, our graph decomposition algorithm runs in $O(\abs{\V} \log \abs{\V} + \abs{\E} \log \abs{\V})$.
It begins with a scan of all edges that have been created or deleted $O(\abs{\E})$.
Their vertices at both ends are partitioned into three sets: multi-parented, single-parented, or orphaned.
Their relationships are recorded in maps for $O(\log \abs{\V})$ lookups of parent and child edge sets.
After the vertices have been partitioned, we traverse the various single-parented components and produce equivalent expressions $O(\abs{\V})$.
For unicycles, we traverse backwards until a vertex is seen twice $O(\abs{\V})$, then proceed forward to find the least vertex on the cycle $O(\abs{\V})$.

% TODO: is this still a thing?
% Compared to the core language, our implementation lacks support for lists and case expressions.

Optimizations: least fixed point

We determine $e$ by starting at the root of the graph and including every connected
vertex that has a single parent. In positions where there are no children, we
leave an empty hole $\hole$. In positions where there are multiple children, we
leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}{v}{p}$. Whenever we
encounter a vertex with multiple parents, we add its corresponding expression to
the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
the unique identifier of the referenced vertex

% here are some properties that show how to get one from the other, and to/from
% graph

algorithmic section: if we want to implement these conversions, here are the
steps (without introducing new concepts):

- linear scan
- breaking out roots
- ...

Graph decomposition occurs in three steps.

% TODO: use code generation instead: have Lang.ml generate expr / expr' / ... for us; then we don't need this figure.
% TODO: maybe show this figure and say that it's possible to handle generically without code generation

% \begin{figure}
%   \[
%     \arraycolsep=0pt
%     \begin{array}{lrlll}
%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
%     \end{array}
%   \]
%   \caption{Syntax of graph decomposition trees as a grammar.}
%   \label{fig:Syntax of graph decomposition trees}
% \end{figure}

lemma: a unicycle is a cycle with trees hanging off of it

lemma: unicycle traversal always produces a valid tree

Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
starting point and its parents are traversed and marked as seen. Any vertex that
has been seen twice must be on the cycle. We take the first one as the root of
the unicycle and traverse its children to produce a tree that covers the unicycular graph.
Any remaining single-parented vertices must be parts of other unicycular graphs and are
unicycle-traversed until there are no more single-parented vertices left.

### 2021-08-19 grove-paper-3-formalism.tex
edits with Cyrus
fiill in details of definitions
start merging appendix defs into section 3

@@
 
-We will now make the intuitions developed in the previous section precise
-by defining a collaborative structure editor calculus called Grove.
-We begin in Sec.~\ref{sub:Syntax} with the syntax, then discuss
-the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
-...
+We will now make the intuitions developed in the previous section precise by defining a collaborative structure editor calculus called Grove.
+This section
+  gives the syntax of our source language,
+  discusses the internal representations of graphs and terms,
+  presents a commutative graph edit semantics,
+  and introduces the user action semantics.
 
-TODO: point of this section is commutativity theorem
+% We begin in Sec.~\ref{sub:Syntax} with the syntax,
+% then discuss the internal graph representation in Sec.~\ref{sub:Convergent Graphs},
+% ...
 
-TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
+% TODO: point of this section is commutativity theorem
 
-The formalism is rather simple, \autoref{sec:Grove By Example} shows
-that it is sufficient to handle many common situations,
-and this simplicity aids in the predictability.
+% TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
+
+% The formalism is rather simple, \autoref{sec:Grove By Example} shows
+% that it is sufficient to handle many common situations,
+% and this simplicity aids in the predictability.
 
@@
 
-The language we are modeling is shown in \autoref{fig:Base syntax}, standard for a simply-typed lambda calculus except that all sorts have empty holes (represented by $\hole$) and conflict holes (of the form $\conflictHole{q,\ldots,q}{v}{p}$).
-In particular, the argument of a lambda can be either a bound variable or a hole.
-Additionally, all non-hole forms have a superscripted identifier $u \in \U$.
+Our source language (\autoref{fig:Syntax}) is a simply typed lambda calculus with references, empty holes, and conflict holes.
+The syntax of terms is sorted into expressions, types, and patterns.
+Expressions include
+  variables $\eVar{G}{x}$,
+  lambda abstractions with pattern arguments $\eFun{G}{q}{\tau}{e}$,
+  applications $\eApp{G}{e_1}{e_2}$,
+  and numbers $\eNum{G}{n}$ with addition $\ePlus{G}{e_1}{e_2}$ and multiplication $\eTimes{G}{e_1}{e_2}$.
+Types include
+  arrows $\tArrow{G}{\tau_1}{\tau_2}$
+  and the type of numbers $\tNum{G}$.
+Patterns include variables $\pVar{G}{x}$.
+Each sort of term also contains
+multiparent references $\multiVertex{\e}$,
+unicycle references $\cycleVertex{\e}$,
+conflict holes $\conflictHole{t_i}_{i < n}$,
+and empty holes $\hole$.
+Note that for conflict holes, it must be the case that $n \geq 2$.
+We identify conflict holes up to reordering.
 
-% Turn that into a constructor set K + position set P + arity function (has
-% type: K -> $\wp$(P)).
+Each term that is not a conflict hole carries an annotation.
+For terms that are not holes, the annotation corresponds to the graph or edge it recomposes to.
+Multiparent references correspond to edges that target
+  \emph{multiparent roots}---vertices targeted by multiple edges.
+Unicycle references correspond to edges that target
+  \emph{unicycle roots}---vertices that are their own least ancestor.
+Empty holes, which denote the absence of any corresponding structure,
+are annotated with the source needed to construct terms in their place.
 
-% TODO: replace positions with macros
+Conflict holes correspond to multiple edges with a common source.
+Since conflict holes cannot be nested or contain empty holes,
+each term in a conflict hole is annotated with the graph or edge it recomposes to,
+and the combined annotations correspond to the graph the entire conflict hole recomposes to.
 
-\figureArity
+% \figureArity
 
@@
 
-\subsubsection{Graphs}%
-\label{sub:Graphs}
+Let $\U$ be a set of \emph{unique identifiers} equipped with a total ordering $\leq$,
+and $\K$ the \emph{constructors} of terms in the source language,
+and $\P$ the \emph{positions} of subterms in the source language,
+and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
+  \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$
+  that forms a lattice over $\Sigma$.
+Now, let $\V = \U \times \K$ all possible \emph{vertices},
+and $\S = \V \times \P$ all possible \emph{sources},
+and $\E = \U \times \S \times \V$ all possible \emph{edges}.
 
-A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
-where $\E = \U \times \V \times \P \times \V$,
-unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
-vertices are drawn from $\V$,
-positions are drawn from $\P$,
-and edge states are drawn from $\Sigma$, all defined below.
+A \emph{graph} $G : \E \rightarrow \Sigma$ is a function from edges to edge states.
+Graphs that represent a complete edit state have a distinguished
+  root vertex $\rootVertex$
+  and root position $\rootPosition$.
+The \emph{inedges} of $G$ are the edges on which $G$ returns $\Plus$.
+The \emph{outedges} of a source are the inedges that have the source.
 
-Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
-the child position $p$ of originating vertex $v$ to destination vertex
-$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
-edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
-$\e$ has not been created or deleted. When $G(\e)$ is $\Plus$, the edge has been
-created. When $G(\e)$ is $\Minus$, the edge has been deleted. The total
-ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
+An \emph{edge} $\e{=}(u, (v, p), v^\prime) \in \E$ represents a directed multi-edge
+identified by $u \in \U$, originating from source $(v, p) \in \S$, with destination vertex $v' \in \V$.
 
-A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
-\K$, corresponding to a non-hole form defined in \autoref{fig:Base syntax}. Each
-contructor has an arity, which we define as a set of positions. A position $p
-\in \P$ identifies a location at some originating vertex from which edges to
-destination vertices may originate. The $\arityOp$ function in
-\autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
-constructors to potential edge positions with their corresponding sorts. Every
-graph contains a distinguished root vertex, root constructor, and root position,
-respectively named $v_0$, $k_0$, and $p_0$.
+A vertex $v = (u, k) \in \V$ is an instance of constructor $k \in \K$ identified by $u \in \U$.
+Vertices correspond to \emph{constructable terms}---terms that are not references or holes.
+The \emph{inedges} of $v$ are the inedges that target $v$.
+The \emph{ingraph} of $v$ is the subgraph on inedges that target $v$.
+We write $G_v$ to denote the ingraph of $v$.
+The \emph{parents} of $v$ are the source vertices of the inedges of $v$.
+The \emph{ancestors} of $v$ is defined inductively as the parent vertices of $v$ plus their ancestors.
+The \emph{least ancestor} of $v$ is the ancestor with the smallest unique identifier.
 
-\vspace*{\baselineskip}
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
-% TODO: add a lead-in sentence here
+\subsubsection{Well sorted graphs}
+\label{sub:Well sorted graphs}
+
+% TODO: where does this go?
+% 
+% Each contructor has an arity, which we define as a set of position-sort pairs. A position $p
+% \in \P$ identifies a location at some originating vertex from which edges to
+% destination vertices may originate. The $\arityOp$ function in
+% \autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
+% constructors to potential edge positions with their corresponding sorts.
 
@@
 Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
-A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Base syntax}, which we organize into a structure called a grove.
+A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Syntax}, which we organize into a structure called a grove.
 
@@
 \figureDecompositionDef
-
-\figureDecompositionDefHelpers
 

### 2022-03-07 appendix/language-definitions.tex
simplify section file names and start separate appendix doc

\section{Language Definitions}
\label{sec:language-definitions}

### 2022-03-07 formalism.tex
move decomp to appendix and add more detail
simplify section file names and start separate appendix doc

@@
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\section{Formalism}%
+\label{sec:Formalism}
 
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+% TODO: try to introduce groves in section 2?
+% TODO: use key phrases in section 2, e.g.: "unresolved parentage"
 
-\setcopyright{none}
+We will now make the intuitions developed in the previous section precise by defining a collaborative structure editor calculus called Grove.
+This section introduces
+  graphs (\autoref{sub:Convergent Graphs})
+  and commutative graph edits (\autoref{sub:Graph Edits}).
+Users do not edit graphs directly.
+Instead, they work with the decomposition of graphs into groves (\autoref{sub:Groves})%
+---collections of terms with empty holes, conflict holes, and references to terms with unresolved parentage.
+User actions on groves translate to corresponding graph edits (\autoref{sub:User Actions}).
 
-\input{config-tex}
-\input{config-tikz}
-\input{config-glyphs}
-\input{config-definitions}
-\input{config-graphs}
+% TODO: point of this section is commutativity theorem
 
-\begin{document}
+% TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
 
-\title[Grove]{Excerpt: Recomposition}
+% The formalism is rather simple, \autoref{sec:Grove By Example} shows
+% that it is sufficient to handle many common situations,
+% and this simplicity aids in the predictability.
 
-\maketitle
+% \figureArity
 
-\input{appendix-formalism}
+\subsection{Convergent Graphs}%
+\label{sub:Convergent Graphs}
 
-\end{document}
+% Let $\U$ be a set of \emph{unique identifiers} equipped with a total ordering $\leq$,
+% and $\K$ the \emph{constructors} of terms in the source language,
+% and $\P$ the \emph{positions} of subterms in the source language,
+% and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
+%   \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ that forms a lattice over $\Sigma$.
+% Now, let $\V = \U \times \K$ all possible \emph{vertices},
+% and $\S = \V \times \P$ all possible \emph{sources},
+% and $\E = \U \times \S \times \V$ all possible \emph{edges}.
+
+A \emph{graph} $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
+  where $\E = \U \times \S \times \V$,
+  unique identifiers are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
+  sources are drawn from $\S$,
+  vertices are drawn from $\V$,
+  and edge states are drawn from $\Sigma$,
+  all defined below.
+Graphs that represent a complete edit state have a distinguished
+  root vertex $\rootVertex$
+  and root position $\rootPosition$.
+The \emph{inedges} of $G$ are the edges on which $G$ returns $\Plus$.
+The \emph{outedges} of a source are the inedges that have the source.
+
+An \emph{edge} $\e{=}(u, (v, p), v^\prime) \in \E$ represents a directed multi-edge
+identified by $u \in \U$, originating from source $(v, p) \in \S$, with destination vertex $v' \in \V$.
+
+The \emph{edge states} $\Sigma = \SetOf{\bot, \Plus, \Minus}$ determine which edges are live in a graph.
+For example, if $G(\e) = \Plus$ for some $G : \E \rightarrow \Sigma, \e \in \E$,
+then we say $\e$ is \emph{live}, i.e., it corresponds to a term in the graph decomposition, as explained in \autoref{sub:Graph Decomposition}.
+The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
+
+A \emph{source} is a pair $(v, p) \in \S$,
+  where $\S = \V \times \P$,
+  vertices are drawn from $\V$,
+  and positions are drawn from $\P$.
+
+A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is a constructor instance,
+  where unique identifiers are drawn from $\U$
+  and constructors are drawn from $k \in \K$.
+Vertices represent \emph{constructable terms}, i.e., terms that are not references or holes.
+The \emph{inedges} of $v$ are the inedges that target $v$.
+The \emph{ingraph} of $v$ is the subgraph on inedges that target $v$.
+We write $G_v$ to denote the ingraph of $v$.
+The \emph{parents} of $v$ are the source vertices of the inedges of $v$.
+The \emph{ancestors} of $v$ is defined inductively as the parent vertices of $v$ plus their ancestors.
+The \emph{least ancestor} of $v$ is the ancestor with the smallest unique identifier.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+\subsubsection{Well sorted graphs}
+\label{sub:Well sorted graphs}
+
+% TODO: where does this go?
+% 
+% Each contructor has an arity, which we define as a set of position-sort pairs. A position $p
+% \in \P$ identifies a location at some originating vertex from which edges to
+% destination vertices may originate. The $\arityOp$ function in
+% \autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
+% constructors to potential edge positions with their corresponding sorts.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: put this in a definition box
+
+A graph $G$ is well sorted if
+for all $\e = (u, (u_1, k_1), p, (u_2, k_2))$
+such that $G(\e) \neq \bot$,
+we have $(p, \sort(k_2)) \in \arity(k_1)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% In decomp section:
+% - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
+%   - want to show that expr is a total function on Exp's
+% - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
+%   - every vertex should be part of the decomposition
+% - correctness of grove components:
+%   - e_r: rootVertex(e_r) = v_r
+%   - MP: for every expression in MP, |parents(the root vertex)| > 1
+%   - NP: for every expression in NP, |parents(the root vertex)| = 0
+%   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
+%     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
+%   - we've accounted for every edge:
+%     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
+%     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
+
+% correctness covers:
+% - invertibility (w/ some identity assignment)
+% - parentage (things are multiparented correctly)
+
+\vspace*{\baselineskip}
+
+% TODO: should graph "recomposition" have an algorithm, too?
+% TODO: update this to match our current understanding
+A graph can be derived from an expression in the calculus by starting with an
+empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
+expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
+sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
+constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
+= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
+outermost, starting with the ones at positions in $\arity(k')$, and create edges
+to the vertices representing their immediate sub-terms until only empty holes
+remain. Note that empty holes are not modeled explicitly in the graph. Wherever
+an empty hole appears in an expression, the corresponding graph position has no
+edges originating from it.
+
+% You can map from syntax to graph by selecting
+% unique IDs $u$ (for both edges and vertexes).
+% - holes are not explicit in the graph
+
+\subsection{Graph Edits}%
+\label{sub:Graph Edits}
+
+We define a graph edit $\alpha \in \A = \left\{\Plus, \Minus\right\} \times \E$
+as a pair of an edge state, excluding $\bot$, and an edge. As a shorthand, we
+define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
+
+We define the join operation $s_1 \sqcup s_2$ to be the least upper bound of
+$s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
+\sqcup \bot = \Plus$ means that edges can only be created if they do not already
+exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
+an edge is deleted it can never be restored.
+
+We define the semantics of graph actions via the following transition relation
+between graphs.
+\[
+  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right]
+\]
+Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
+updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
+state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
+$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
+$G$.
+
+% Graph action semantics is a transition system between graphs.
+% (join with whatever it was before)
+
+% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
+% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
+
+\subsection{Commutativity}%
+\label{sub:Commutativity:formal}
+
+In order to define commutativity, we first define a lattice over $G$.
+Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
+$G \overset{\alpha}\rightarrow G'$
+iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
+Thus the commutativity of graph actions is established
+by the commutativity of the join operations defining those graph actions.
+
+Discuss uniqueness of uuid
+
+\begin{lemma}[Join Commutativity]
+  \label{lem:Join Commutativity}
+\end{lemma}
+
+\begin{lemma}[Joining]
+  \label{lem:Joining}
+  For all graphs $G$ and $G'$ and all edit actions $\alpha$,
+  $G \overset{\alpha}{\longrightarrow} G'$
+  iff $G \sqcup \llbracket\alpha_1\rrbracket = G'$.
+\end{lemma}
+\begin{proof}
+  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
+\end{proof}
+
+\begin{theorem}[Commutativity]
+  \label{thm:Commutativity}
+  For all graphs $G$ and $G'$ and all edit actions $\alpha_1$ and $\alpha_2$,
+  if $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$,
+  then $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
+\end{theorem}
+\begin{proof}
+  Using \autoref{lem:Joining} to unfold $\rightarrow$ into $\sqcup$,
+  $G \sqcup \llbracket\alpha_1 \sqcup \llbracket \sqcup \alpha_2 \rrbracket = G'$.
+  Then by \autoref{lem:Join Commutativity},
+  $G \sqcup \llbracket\alpha_2 \sqcup \llbracket \sqcup \alpha_1 \rrbracket = G'$.
+  Finally using \autoref{lem:Joining} to fold $\sqcup$ into $\rightarrow$,
+  $G \overset{\alpha_2\alpha_1}{\longrightarrow} G'$.
+\end{proof}
+
+\subsubsection{Agda Mechanization}%
+\label{sub:Agda Mechanization}
+
+\subsection{Groves}%
+\label{sub:Groves}
+
+\figureTermSyntax
+
+% TODO: check for \hole --> \emptyHole
+
+Our source language (\autoref{fig:Syntax}) is a simply typed lambda calculus with references, empty holes, and conflict holes.
+The syntax of terms is sorted into
+  expressions $e$,
+  types $\tau$,
+  and patterns $q$.
+These can include
+  graph annotations $G$,
+  edge annotations $\e$,
+  or source annotations $(v, p)$,
+  which we introduce in \autoref{sub:Convergent Graphs}.
+Expressions include
+  variables $\eVar{G}{x}$,
+  lambda abstractions with pattern arguments $\eFun{G}{q}{\tau}{e}$,
+  applications $\eApp{G}{e_1}{e_2}$,
+  and numbers $\eNum{G}{n}$ with addition $\ePlus{G}{e_1}{e_2}$ and multiplication $\eTimes{G}{e_1}{e_2}$.
+Types include
+  arrows $\tArrow{G}{\tau_1}{\tau_2}$
+  and the type of numbers $\tNum{G}$.
+Patterns include variables $\pVar{G}{x}$.
+Each sort of term also contains
+and empty holes $\hole_{(v, p)}$.
+multiparent references $\multiVertex{\e}$,
+unicycle references $\cycleVertex{\e}$,
+and conflict holes $\conflictHole{t_i}_{i < n}$.
+For conflict holes, it must be the case that $n \geq 2$.
+We identify conflict holes up to reordering.
+
+Each term that is not a conflict hole carries an annotation.
+For terms that are not holes, the annotation corresponds to the graph or edge it recomposes to.
+Multiparent references correspond to edges that target
+  \emph{multiparent roots}---vertices targeted by multiple edges.
+Unicycle references correspond to edges that target
+  \emph{unicycle roots}---vertices that are their own least ancestor.
+Empty holes, which denote the absence of any corresponding structure,
+are annotated with the source needed to construct terms in their place.
+
+Conflict holes correspond to multiple edges with a common source.
+Since conflict holes cannot be nested or contain empty holes,
+each term in a conflict hole is annotated with the graph or edge it recomposes to,
+and the combined annotations correspond to the graph the entire conflict hole recomposes to.
+
+\subsection{Graph Decomposition}%
+\label{sub:Graph Decomposition}
+
+% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
+% and multiparents
+
+% NOTE: this whole section is for describing the 4-tuple
+
+Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
+A graph can be decomposed into a collection of expressions governed by the grammar in \autoref{fig:Syntax}, which we organize into a structure called a grove.
+
+\figureDecompExample
+
+% TODO: make sure we're not claiming to invent the term "unicycular graph"
+% TODO: maybe add a forward reference at end here to Implementation section
+% TODO: make ID of y lower than the unicycle root, to show what we're minimizing
+
+A grove is a 3-tuple of the following form:
+\[
+  (NP, MP, U)
+\]
+where
+\begin{align*}
+  \figureDecompositionDefDecompComponents
+\end{align*}
+For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
+given in \autoref{fig:Decomposition example grove}.
+
+\figureDecompositionDef
+
+Let us now examine each of the four components in more detail.
+%
+The first component $NP$ is the set of expressions rooted at a vertex with no parents, including the main root vertex.
+% TODO: change vRoot to dot w/ subscript
+In the example, $e_r$ is the expression corresponding to the main root, $\vRoot$.
+% TODO: change z_... to macro use w/ same appearance as in the graph
+In addition, $z_{\id{52}}$ has no parents because it was deleted, so it also appears in $NP$.
+
+% The first component $e_r$ is the expression derived from the root vertex $v_0$.
+% \autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
+% There is one rule for each term constructor.
+% For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
+% In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
+% An empty hole $\hole$ appears if there are no children at the given position.
+% A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
+% If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
+% For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
+
+The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
+Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
+In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
+
+The third component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
+These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
+Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
+Breaking the cycle turns a unicyclic graph into a tree.
+% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
+% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
+To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
+Formally, we choose the vertex $v$ such that $v = \min{\ancestors{v}}$ because
+only a vertex on the cycle can be its own ancestor.
+When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
+In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
+
+% XXX
+
+decomposition produces sets of trees. As long as everything is well sorted, we
+can use them to produce expressions in the main grammar. (alt: define wrt the
+graph instead, once we establish the graph can be turned into expressions)
+
+*** well-sortedness judgment on trees: the tree only has valid positions for the
+constructors, and the child at each position has the sort corresponding to its
+position, and edges in position maps all start at the given constructor.
+
+Theorem: well-sorted graphs produce decompositions of well-sorted trees
+
+``A graph has this property iff these conditions hold''
+
+then we can define a total mapping from well-sorted trees to expressions
+(because the tree structure is very generic, so this ties them back to the
+earlier grammar.)
+
+forest = 4-tuple
+
+forest validity: each of the trees in the forest are well sorted, and all Refs
+point to roots of other trees of the correct sort.
+
+every valid forest corresponds to a well behaved grove (i.e., ``expression
+forest'')
+
+soundness: if you start with a valid graph, you get a well behaved grove that
+connects every vertex (and only those vertices) in the graph. (i.e., if you turn
+it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
+Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
+add Root sort
+
+OR:
+
+use edge IDs instead of edges and build a mapping from vertices to
+sorts---change the $\e$s to $u$s)
+
+% TODO: maybe add a forward reference here to Implementation section
+
+\subsection{User Actions}%
+\label{sub:User Actions}
+
+% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
+% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
+
+move = delete + add
+
+other composite edits
+
+remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
+
+given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
+This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
+(we probably want to do it this way in Hazel)
+
+OR
+
+We could go the indirect route and avoid having to prove the theorem.
+(this way is easier to start with, and it's how GRV works)
+
+% Mapping to b actions
+
+% \[
+%   \begin{array}{l}
+%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
+%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
+%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
+%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
+%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
+%     \CC(G) = \ldots                                                                                                                                    \\
+%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
+%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
+%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
+%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
+%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
+%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
+%   \end{array}
+% \]

### 2022-03-08 appendix/language-definitions.tex
define language representations with ports and arity maps

@@
 \label{sec:language-definitions}
+
+\begin{definition}
+  A \emph{language representation}
+  $\Delta = (\Sorts, \Constructors, \Positions, \ConstructorSorts, \PositionSorts, \PositionConstructors)$
+  is a six-tuple consisting of three finite ordered sets with three binary relations:
+
+  \begin{minipage}{.48\linewidth}
+    \[
+      \arraycolsep=4pt
+      \begin{array}{lrcr}
+        \text{Sorts}        & \Sorts        & = & \Setof{S_i}_{i < |\Sorts| }        \\
+        \text{Constructors} & \Constructors & = & \Setof{C_i}_{i < |\Constructors| } \\
+        \text{Positions}    & \Positions    & = & \Setof{P_i}_{i < |\Positions| }    \\
+      \end{array}
+    \]
+  \end{minipage}
+  \hfil
+  \begin{minipage}{.48\linewidth}
+    \[
+      \arraycolsep=4pt
+      \begin{array}{lrcr}
+        \text{Position Constructors} & \PositionConstructors & \subseteq & \Positions \times \Constructors \\
+        \text{Position Sorts}        & \PositionSorts        & \subseteq & \Positions \times \Sorts        \\
+        \text{Constructor Sorts}     & \ConstructorSorts     & \subseteq & \Constructors \times \Sorts     \\
+      \end{array}
+    \]
+  \end{minipage}
+  %
+  such that the relations satisfy the following properties:
+  \begin{enumerate}
+    \item For each position $P \in \Positions$,
+          there is a unique constructor $C \in \Constructors$ such that $(P, C) \in \PositionConstructors$.
+    \item For each position $P \in \Positions$,
+          there is a unique sort $S \in \Sorts$ such that $(P, S) \in \PositionSorts$.
+    \item For each constructor $C \in \Constructors$,
+          there is a unique sort $S \in \Sorts$ such that $(C, S) \in \ConstructorSorts$.
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  In any language representation $\Delta$,
+  the \emph{ports} of a constructor $C \in \Constructors$, denoted $\Ports{C}$,
+  is the largest subset of $\Positions$ such that $(P, C) \in \PositionConstructors$ for all $P \in \Ports{C}$.
+\end{definition}
+
+\begin{definition}
+  The \emph{arity map} of a language representation $\Delta$
+  is a binary relation $\Arity \subseteq \Constructors \times \Pow{\Positions}$
+  defined by $\Arity = \Setof{(C, [C]_\Positions)}_{C \in \Constructors}$.
+\end{definition}
+
+\begin{theorem}
+  \label{thm:langrep-rel-funs}
+  The relations of a language representation are total functions.
+\end{theorem}
+
+\begin{theorem}
+  \label{thm:langrep-arity-fun}
+  In any language representation, arity maps are total functions.
+\end{theorem}
+
+Theorems \ref{thm:langrep-rel-funs} and \ref{thm:langrep-arity-fun}
+justify the use of function notation when working with language representations and arity maps.
+For example, for any $P \in \Positions$, we may write $\PositionConstructors(P)$
+to denote the unique constructor $C \in \Constructors$ such that $(P, C) \in \PositionConstructors$.
+
+\begin{theorem}
+  Any language representation $\Delta$ with arity map $\Arity$
+  satisfies the property that $\PositionConstructors(P) = C$ if and only if $P \in \Arity(C)$.
+\end{theorem}
+
+\begin{proof}
+  Let $\Delta$ be a language representation with arity map $\Arity$.
+  For any $P \in \Positions$, if $\PositionConstructors(P) = C$ then $P \in \Ports{C}$ implies $P \in \Arity(C)$.
+  Conversely, for any $C \in \Constructors$, if $P \in \Arity(C)$ then $P \in \Ports{C}$ implies $\PositionConstructors(P) = C$.
+\end{proof}

### 2022-08-09 appendix/language-construction.tex
snapshot of current working directory

\section{Language Construction}
\label{sec:language-construction}

\begin{definition}
  A \emph{signature}
  $\Sig = (\Sorts, \Constructors, \Positions, \Sp, \Sc, \ArityInv)$ consists of
  \begin{enumerate}
    \item a set $\Sorts$ of \emph{sorts},
    \item a set $\Constructors$ of \emph{constructors}, or vertex types,
    \item a set $\Positions$ of \emph{positions}, or edge types,
    \item a function $\Sc : \Constructors \to \Sorts$ assigning each constructor a sort,
    \item a function $\Sp : \Positions \to \Sorts$ assigning each position a sort,
    \item a function $\ArityInv : \Positions \to \Constructors$ assigning each position a constructor.
  \end{enumerate}
\end{definition}

\begin{definition}
  The \emph{ports} $\Ports{C}$ over constructor $C$ is the set
  \[ \Setof{P \in \Positions \suchthat \ArityInv(P) = C} \]
  of positions assigned to constructor $C$.
\end{definition}

\begin{definition}
  The \emph{arity map} $\Arity : \Constructors \to \Pow{\Positions}$ of signature $\Sig$ is
  \[ \Arity = (\ArityInv)^{-1} \]
  the inverse of its constructor assignment function.
\end{definition}

### 2022-08-09 introduction.tex
working on intro

--- 2021-05-03 grove-paper-1-introduction.tex
@@
 
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
+A \emph{collaborative structure editor} is an editor that
+(1) allows multiple concurrent users to work on a shared document, while also
+(2) providing structure-aware editor services such as projectional editing, syntax highlighting, or automatic code folding.
+%
+Collaborative editing research focuses on the design and implementation of real-time, multi-user, character-based communication systems,
+whereas structure editors typically presume a more complex document schema and then focus on some other aspect of the user experience.
+In both settings, preservation of user intent is a core technical challenge.
+%
+Although collaborative editors and structure editors have overlapping goals (optimal user experience)
+and complementary design challenges (subject-subject versus subject-object harmony),
+to our knowledge, there is no comprehensive, principled account of their combined use.
+
+Since collaborative editors are essentially distributed systems, existing work tends to focus on extensions to distribution protocols.
+Lots of examples using OT. (Google Docs)
+OT is complex and largely textual.
+OT can make sense for real time systems: users typically change one character at a time, and instant feedback can help to prevent conflicts.
+On the other hand, OT system designs can be difficult to extend.
+
+Alternatively, there's CRDT. (Peritext?)
+CRDT is easier to implement, but harder to design for.
+There's an All-CRDT editor---it turned out to be not so realistic. (what's it called again?)
+
+Structure editing has been a recurring theme in the computer science literature since at least Engelbart's ``Mother of All Demos.''
+Provides automation for domain experts and reduces the barrier to entry for everyone else.
+Popular for editing programs, i.e., for programming language-specific editors.
+
+However, modern program editors typically disable editor services (like what?) when the document is not in a consistent state,
+a phenomenon called the ``gap problem.''
+Of course, in the presence of multiple concurrent users, the problem gets worse.
+
+In a collaborative setting,
+Hazel is a structure editor with support for advanced editing services, (e.g., semantic actions?).
+
+\newpage
 

### 2022-08-10 introduction.tex
add some intro notes w/ cyrus

@@
 \label{sec:Introduction}
+
+start with something that everyone can accept: ``programming is collaborative.''
+
+briefly introduce 3 types: diff, OT, CRDT
+
+tease out some common problems (tree-native, collaborative)
+
+propose a solution
+
+general motivation: programming is a collaborative activity
+
+today usually takes the form of coarse-grained solutions, e.g, git
+
+if you try to use eg google docs for this, there are several problems
+(gap problem, diffs don't handle repositioning well)
+
+diff is not really an ideal way of communicating change: e.g., how repositioning gets mishandled
+
+CRDT: tree-native vs text. think about obvious limitation of existing tools
+
+algorithms for collaborative text editing are well developed
+natural fit for text
+not a natural fit for trees
+so if you want to build a grove, a more natural approach is needed
+
+a lot of collaborative editing of other sorts (not necessarily programming) requires a tree model,
+but that problem doesn't seem to have been study directly (maybe except JSON CRDT).
 

### 2022-08-12 introduction.tex
still working on the intro

--- 2021-05-03 grove-paper-1-introduction.tex
@@
 
-TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Programming is an increasingly collaborative activity.
+Today, collaborative editing services are either provided directly by an asynchronous,
+diff-and-merge-based version control system like git, darcs, or pijul,
+or they are integrated into a real-time editor built on top of a distributed consistency protocol such as OT or CRDT.
+
+Diff-and-merge is not an ideal collaborative editing solution.
+Since diffs do not provide information on the original scope or ordering of edits,
+the merging process must reconstruct an approximation, e.g., as a sequence of line additions and deletions.
+Clearly, in the context of editing programs,
+a line-based merge is not only less likely to succeed than one which retains finer grained information about changes over time,
+it is also less likely to produce useful information about conflicting edit states, when they arise.
+
+Operational Transformation (OT), on the other hand, goes perhaps too far in the opposite direction.
+OT is a distributed system architecture that coordinates the proliferation and maintainance of edit states and actions.
+Industrial-scale collaborative editors such as Google Docs tend to employ OT
+as it provides strong theoretical guarantees on the eventual consistency of causally-ordered events
+while remaining suitable for real-time applications.
+It does so by maintaining a history of recent edits, or operations, while transforming individual edits to preserve a causal ordering.
+But fully featured distributed systems that maintain causal orderings of arbitrary events are hard to design well,
+so in practice OT-based systems tend to be restricted to linear text editing or have fewer features than comparable single-user alternatives.
+
+Convergent Replicated Data Types (CRDTs) were proposed as a reaction to the complexity of Operational Transformation,
+though the claim is currently under strutiny by the research community.
+A CRDT is essentially a distributed data structure with an API for building and interacting with it.
+CRDTs avoid the need for explicit state synchronization by requiring that allowable edit actions inhabit a lattice.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% start with something that everyone can accept: ``programming is collaborative.''
+
+% briefly introduce 3 types: diff, OT, CRDT
+
+% tease out some common problems (tree-native, collaborative)
+
+% propose a solution
+
+% general motivation: programming is a collaborative activity
+
+% today usually takes the form of coarse-grained solutions, e.g, git
+
+% if you try to use eg google docs for this, there are several problems
+% (gap problem, diffs don't handle repositioning well)
+
+% diff is not really an ideal way of communicating change: e.g., how repositioning gets mishandled
+
+% CRDT: tree-native vs text. think about obvious limitation of existing tools
+
+% algorithms for collaborative text editing are well developed
+% natural fit for text
+% not a natural fit for trees
+% so if you want to build a grove, a more natural approach is needed
+
+% a lot of collaborative editing of other sorts (not necessarily programming) requires a tree model,
+% but that problem doesn't seem to have been study directly (maybe except JSON CRDT).
+
+% A \emph{collaborative structure editor} is an editor that
+% (1) allows multiple concurrent users to work on a shared document, while also
+% (2) providing structure-aware editor services such as projectional editing, syntax highlighting, or automatic code folding.
+% %
+% Collaborative editing research focuses on the design and implementation of real-time, multi-user, character-based communication systems,
+% whereas structure editors typically presume a more complex document schema and then focus on some other aspect of the user experience.
+% In both settings, preservation of user intent is a core technical challenge.
+% %
+% Although collaborative editors and structure editors have overlapping goals (optimal user experience)
+% and complementary design challenges (subject-subject versus subject-object harmony),
+% to our knowledge, there is no comprehensive, principled account of their combined use.
+
+% Since collaborative editors are essentially distributed systems, existing work tends to focus on extensions to distribution protocols.
+% Lots of examples using OT. (Google Docs)
+% OT is complex and largely textual.
+% OT can make sense for real time systems: users typically change one character at a time, and instant feedback can help to prevent conflicts.
+% On the other hand, OT system designs can be difficult to extend.
+
+% Alternatively, there's CRDT. (Peritext?)
+% CRDT is easier to implement, but harder to design for.
+% There's an All-CRDT editor---it turned out to be not so realistic. (what's it called again?)
+
+% Structure editing has been a recurring theme in the computer science literature since at least Engelbart's ``Mother of All Demos.''
+% Provides automation for domain experts and reduces the barrier to entry for everyone else.
+% Popular for editing programs, i.e., for programming language-specific editors.
+
+% However, modern program editors typically disable editor services (like what?) when the document is not in a consistent state,
+% a phenomenon called the ``gap problem.''
+% Of course, in the presence of multiple concurrent users, the problem gets worse.
+
+% In a collaborative setting,
+% Hazel is a structure editor with support for advanced editing services, (e.g., semantic actions?).
+
+\newpage
 

### 2022-08-18 introduction.tex
intro

@@
 or they are integrated into a real-time editor built on top of a distributed consistency protocol such as OT or CRDT.
+Whereas algorithms for collaborative text editing have been well developed,
+we know far less about collaborative editing of more structured documents---specifically, of program source files---%
+hence we are motivated to ask the following questions:
+\begin{itemize}
+\item Is there a more natural model for collaborative editing of tree structures?
+\item How do we preserve user intent in the presence of concurrent, non-linear edits?
+\end{itemize}
 
-Diff-and-merge is not an ideal collaborative editing solution.
+Diff-and-merge is clearly not an ideal solution.
 Since diffs do not provide information on the original scope or ordering of edits,
 the merging process must reconstruct an approximation, e.g., as a sequence of line additions and deletions.
-Clearly, in the context of editing programs,
-a line-based merge is not only less likely to succeed than one which retains finer grained information about changes over time,
-it is also less likely to produce useful information about conflicting edit states, when they arise.
+In a tree-native setting,
+line-based merging is not only less likely to succeed than one which retains finer grained information,
+it is also less able to provide useful information about conflicting edit states as they arise.
+Moreover, because diffs cannot guarantee preservation of user intent,
+critical features such as identity-preserving repositioning are impossible to implement on top of them.
 
 Operational Transformation (OT), on the other hand, goes perhaps too far in the opposite direction.
-OT is a distributed system architecture that coordinates the proliferation and maintainance of edit states and actions.
-Industrial-scale collaborative editors such as Google Docs tend to employ OT
-as it provides strong theoretical guarantees on the eventual consistency of causally-ordered events
-while remaining suitable for real-time applications.
-It does so by maintaining a history of recent edits, or operations, while transforming individual edits to preserve a causal ordering.
-But fully featured distributed systems that maintain causal orderings of arbitrary events are hard to design well,
-so in practice OT-based systems tend to be restricted to linear text editing or have fewer features than comparable single-user alternatives.
+It is a distributed system architecture that coordinates the proliferation and maintainance of edits and edit states.
+Industrial-scale collaborative editors, such as Google Docs,
+employ OT for real-time applications requiring strong eventual consistency of causally ordered events.
+It does so by maintaining an event log it transforms upon each new entry in order to preserve the causal ordering.
 
-Convergent Replicated Data Types (CRDTs) were proposed as a reaction to the complexity of Operational Transformation,
-though the claim is currently under strutiny by the research community.
-A CRDT is essentially a distributed data structure with an API for building and interacting with it.
-CRDTs avoid the need for explicit state synchronization by requiring that allowable edit actions inhabit a lattice.
+The Convergent Replicated Data Type (CRDT) was proposed as a simpler alternative to the OT architecture.
+CRDTs do not require causal ordering maintenance because the language of edits must form a join semilattice,
+thereby ensuring the edit state update function increases monotoncally.
+(Foreshadow solution by warning about a shortcoming of CRDT.)
+
+(Propose Grove as a solution)
+
+% OT is a natural fit for character-based editing, but the architecture is hard to extend, so it is rarely used for more structured documents.
+% CRDT imposes less restrictions on the structure of the underlying document,
+% providing a more natural substrate for collaborative structure editing.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% start with something that everyone can accept: ``programming is collaborative.''
-
-% briefly introduce 3 types: diff, OT, CRDT
 
@@
 % propose a solution
-
-% general motivation: programming is a collaborative activity
-
-% today usually takes the form of coarse-grained solutions, e.g, git
-
-% if you try to use eg google docs for this, there are several problems
-% (gap problem, diffs don't handle repositioning well)
-
-% diff is not really an ideal way of communicating change: e.g., how repositioning gets mishandled
-
-% CRDT: tree-native vs text. think about obvious limitation of existing tools
 

### 2022-08-23 introduction.tex
intro

@@
 Programming is an increasingly collaborative activity.
-Today, collaborative editing services are either provided directly by an asynchronous,
-diff-and-merge-based version control system like git, darcs, or pijul,
-or they are integrated into a real-time editor built on top of a distributed consistency protocol such as OT or CRDT.
-Whereas algorithms for collaborative text editing have been well developed,
-we know far less about collaborative editing of more structured documents---specifically, of program source files---%
-hence we are motivated to ask the following questions:
-\begin{itemize}
-\item Is there a more natural model for collaborative editing of tree structures?
-\item How do we preserve user intent in the presence of concurrent, non-linear edits?
-\end{itemize}
+Today, collaborative editing services are either
+provided directly by an asynchronous diff-and-merge version control system like git, darcs, or pijul,
+or they are integrated into a real-time editor backed by a distributed collaboration protocol such as OT or CRDT.
+Whereas algorithms for collaborative text editing have been well developed in the OT setting,
+we know far less about collaborative editing of tree-structured documents.
 
-Diff-and-merge is clearly not an ideal solution.
+Diff-and-merge is clearly not an ideal solution for collaborative tree editing.
 Since diffs do not provide information on the original scope or ordering of edits,
 the merging process must reconstruct an approximation, e.g., as a sequence of line additions and deletions.
-In a tree-native setting,
-line-based merging is not only less likely to succeed than one which retains finer grained information,
-it is also less able to provide useful information about conflicting edit states as they arise.
-Moreover, because diffs cannot guarantee preservation of user intent,
-critical features such as identity-preserving repositioning are impossible to implement on top of them.
+In a tree-native setting, line-based merging can fail in ways that methods which retain finer grained information would not.
+Moreover, user intent can be impossible to preserve,
+so critical features such as identity-preserving repositioning are impossible to implement reliably on top of them.
 
 Operational Transformation (OT), on the other hand, goes perhaps too far in the opposite direction.
-It is a distributed system architecture that coordinates the proliferation and maintainance of edits and edit states.
+OT is a distributed system architecture that coordinates the proliferation and maintainance of edits and edit states.
 Industrial-scale collaborative editors, such as Google Docs,
@@
 CRDTs do not require causal ordering maintenance because the language of edits must form a join semilattice,
-thereby ensuring the edit state update function increases monotoncally.
-(Foreshadow solution by warning about a shortcoming of CRDT.)
+thereby ensuring the edit state update function increases monotonically.
+CRDT places fewer constraints than OT on the language of edits,
+so calculi for editing tree-native documents can be simpler to model for CRDT than OT.
+Similarly, by eliminating all but one design constraint, edit calculi for CRDT can be easier to implement and extend.
 
-(Propose Grove as a solution)
+Unfortunately, monotonicity merely changes the nature of the complexity rather than eliminating it entirely.
+Although tree-native document editing calculi may ultimately admit simpler designs under CRDT than OT,
+there are no generic recipes for transforming a sequential tree editing calculus into one suitable for CRDT.
+Consequently, progress of CRDT-backed tree editors has slowed as the research community debates whether CRDT is living up to its promise.
+Thus, we are motivated to ask:
+Is there a more natural model for collaborative editing of trees?
+And how do we preserve user intent in the presence of concurrent edits in a tree-native editor?
 
-% OT is a natural fit for character-based editing, but the architecture is hard to extend, so it is rarely used for more structured documents.
-% CRDT imposes less restrictions on the structure of the underlying document,
-% providing a more natural substrate for collaborative structure editing.
+To address these gaps in the literature, we propose \emph{Grove}, a CRDT-backed calculus for collaborative structure editing.
+At its heart is a simple partially ordered set of operations over a labeled multidigraph with a distinguished root vertex.
+In this setting, conflicting edit states may be reified as multi-edges,
+allowing conflicts to be exposed as they arise and enabling novel forms of collaboration.
+We establish an isomorphism between labeled multidigraphs and a collection of forests called a \emph{grove},
+enabling more familiar interactions with sub-trees of the graph in more familiar settings, e.g., as a term in a programming language editor.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% tease out some common problems (tree-native, collaborative)
-
-% propose a solution
+We are further motivated by whether there exists a more natural medium for collaborative extension of sophisticated structure editors like Hazel.
+Hazel maintains a sensibility invariant, i.e., every edit state is semantically valid.
+Scaling editor services in the presence of many collaborators is difficult but necessary in a world of increasing connectivity.
+By providing a first step in that direction,
+we believe Grove has the potential to spark a new wave of inquiry into collaborative structure editing.
 

### 2022-08-31 grove-by-example.tex
simplify section file names and start separate appendix doc
redo most of section 2

--- 2021-05-03 grove-paper-2-grove-by-example.tex
@@
 This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
+We begin in \autoref{sub:Program Representation} with how we use graphs to represent programs with holes.
 \autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores actions performed by multiple users, Alice and Bob,
-each of whom is editing their own instance of a program.
+and \autoref{sub:Multi-user Interactions} explores interactions between two users, Alice and Bob,
+as they edit their own instances of a shared program.
 
@@
 
-The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
-This can be presented to the user as an expression containing holes.
-For example, \autoref{fig:Simple:a} shows one such graph and its corresponding expression,~\texttt{x * \hole}, which
-has a hole on the right side of the times operator.
+The \textit{edit state} of a Grove instance is a graph representing a program
+in a language with concrete syntax for holes,
+or terms that have not been written yet or have already been deleted.
+For example, \autoref{fig:Simple:a} gives one such graph and its corresponding program,~\texttt{x * \hole},
+whose missing right operand is denoted by $\hole$.
 
-Program state is a directed graph with a distinguished root vertex.
-Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
-In \autoref{fig:Simple:a}, \vSimpleTimes{} and \vSimpleX{} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
-(For the sake of compactness, in our figures, we abbreviate \texttt{var(x)} as simply \texttt{x}.)
-The root vertex has the distinguished identifier~0 and the constructor~\textbullet.
+The edit state is a directed graph with a distinguished root vertex.
+Each vertex represents a term in a hypothetical programming language
+and is labeled with a globally unique identifier and a term constructor.     % TODO: should we informally define `constructor' here?
+In \autoref{fig:Simple:a}, the root vertex has identifier 0 and constructor~\textbullet.
+Terms \vSimpleTimes{} and \vSimpleX{} have constructors~\texttt{*} and \texttt{var(x)}, respectively.
+For clarity, we abbreviate \texttt{var(x)} as simply \texttt{x}.
 
-Each term constructor has an associated set of child positions.
-The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
-and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
-variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
+Each constructor has an associated set of \emph{child positions} for naming subterms.
+For instance, the~\texttt{*} constructor has positions
+for~\texttt{L}~(i.e., left child)
+and~\texttt{R}~(i.e., right child).
+The~\texttt{var} constructor has no child positions
+and takes a variable name as an argument, e.g.,~\texttt{x} in \texttt{var(x)}.\footnote{Note that,
   for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
-The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.
+  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}      % TODO: should we keep this footnote?
+The root vertex constructor~\textbullet~has a single child position \texttt{Root}.
 
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a}) and
-a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:Simple:b}).
+Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a})
+and a child position (e.g., \texttt{Root} and \texttt{L} in \autoref{fig:Simple:a}).
 An edge indicates that the destination vertex is a child of the origin vertex at the given position.
 %Visually, we indicate the position of an edge by the location of its origin.
-For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.
+For clarity, we reserve odd numbers for vertex identifiers and even numbers for edge identifiers.
 
 Finally, holes are represented by the absence of a child.
-For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} edge coming from \vSimpleTimes{}
-corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.
+For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} child under \vSimpleTimes{}
+corresponds to the hole in the right operand of~\texttt{x * \hole}.
 
@@
 
-In the remainder of this section,
-we consider user actions, starting
-in this subsection with single-user
-actions and continuing in \autoref{sub:Multi-user Interactions}
-with multi-user actions.
+In the remainder of this section, we consider user actions and interactions,
+starting in this subsection with a single user performing one edit at a time
+and continuing in \autoref{sub:Multi-user Interactions} with multiple users performing single edits concurrently.
 
-Each user action corresponds to one or more graph edits.
-Graph edits can only add fresh edges and vertices or remove existing edges.
-
-User actions act relative to a cursor.
-We do not model cursor sharing and the cursor does not appear in the underlying graph.
-We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.
+Each action by the user corresponds to one or more graph update operations.
+Graph updates can only add fresh edges or vertices or remove existing edges.
+All user actions are relative to an instance-local cursor and cursors are not replicated across instances.
+We discuss cursor representations in detail in \autoref{sub:Cursors}.
 
@@
 
-To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}
-and constructs the variable~\texttt{y} as shown in \autoref{fig:Simple:b}.
-This action corresponds to creating \eSimpleY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
-created vertex, \vSimpleY{}, containing the variable reference~\texttt{y}.
-The resulting graph, shown in \autoref{fig:Simple:b}, represents the expression \texttt{x * y}.
+To start our examples, say Alice's cursor is on the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}.
+Alice constructs the variable~\texttt{y}, which produces the edit state in \autoref{fig:Simple:b}.
+This fills the hole by creating a new vertex \vSimpleY{} containing the variable reference
+as a child of \vSimpleTimes{} at position \texttt{R} connected via edge \eSimpleY{}.
+The resulting graph represents the expression \texttt{x * y}.
 
@@
 
-Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
-This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
-In our system, once a vertex is created it is never deleted.
-This allows further manipulations of those vertices by other users as shown later in this section.
-Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.
+Next, Alice moves the cursor to~\texttt{x} and deletes it, leavingx~\texttt{\hole{} * y}.
+This deletes edge~\eSimpleX{} as shown in \autoref{fig:Simple:c}.
+Notice that vertex \vSimpleX{} continues to exist,
+and if it had any children, those children would remain connected to it.
+
+Once a vertex is created, it is never deleted.
+This allows further interaction with those vertices by other users without risk of spontaneous cursor motion,
+as shown later in this section.
+In remaining diagrams, we omit such orphaned vertices if they are not relevant to the exposition.
 
 Once an edge with a particular identifier is deleted, it cannot be recreated.
-
-Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.
+For instance, if Alice performed an ``undo'' on this deletion,
+a fresh edge between \vSimpleTimes{} and \vSimpleX{} would be created.
 
@@
 
-Next, Alice types~\texttt{+} when the cursor is on \vSimpleTimes{}.
-If this were a hole, this would cause a construction action to be performed as in \autoref{sub:Construction} above.
-However, Alice's editor sees that this position is already occupied by the \texttt{*} expression,
-so typing~\texttt{+} is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrap}:
-(a) delete \eSimpleTimes{}, leaving \vSimpleTimes{} temporarily orphaned,
-(b) add a fresh~\texttt{+} vertex (\vWrapPlus{}) as a child of the root vertex, and
-(c) add an edge to \vSimpleTimes{} in the left child position of this new~\texttt{+} vertex.
-(The choice of wrapping with a bias for the left child position is arbitrary.)
+Next, Alice moves the cursor to parent vertex \vSimpleTimes{}
+and constructs a binary addition expression with constructor~\texttt{+}.
+If the cursor were on a hole,
+a fresh term would be constructed in its place, as in \autoref{sub:Construction} above.
+However, since Alice's cursor is on an existing term \texttt{*},
+it is interpreted as a wrapping action instead of a construction action.
+This corresponds to the following sequence of graph updates, shown in \autoref{fig:Wrap}:
+(a) delete edge \eSimpleTimes{} leaving \vSimpleTimes{} temporarily orphaned,
+(b) add fresh vertex \vWrapPlus{} as a child of the root vertex, and
+(c) add fresh edge to \vSimpleTimes{} in left (\texttt{L}) child position of the new vertex.
+Our choice here of wrapping with a bias for the left child position is arbitrary.
 
@@
 
-If a user wants to reposition code from one place to another, we have two options.
-The first is to delete the code from its old position, and
-reconstruct the code in its new position.
-However, edges and vertices have unique identifiers in Grove
-to support the collaborative editing features we discuss next,
-so reconstruction will not preserve identity of vertices and edges.
-Consequently, we take the second option: explicit repositioning actions that preserve identity.
-These actions would be triggered by, for example, a drag and drop, or a cut and paste
-(but not copy and paste, or a second paste after a cut, which do not need to
-preserve identity and so would rely on reconstruction as already described in \autoref{sub:Construction}).
+We have two options for moving a term from one place to another.
+The first is to delete the term from its original position and then reconstruct it in its new position.
+Since edges are uniquely identifiable,
+this option precludes the use of edge identities in any cursor representation,
+as doing so would lead directly to the repositioning problem.
 
-In terms of the graph, a repositioning action involves simply deleting the edge
-from the original parent then adding an edge from the new parent.
-Thus the code is actually \textit{repositioned}, not copied.
+Consequently, we take the second option: explicit repositioning actions that preserve identities.
+These could be triggered by, for example, a drag and drop or a cut and paste---but not a copy and paste,
+or a second paste after a cut, which do not need to preserve identity
+and so could rely on reconstruction as already described in \autoref{sub:Construction}.
+
+In the graph, explicit repositioning involves simply deleting the edge from the original parent
+and then adding a new edge to it from the new parent.
+Thus the term is indeed \textit{repositioned}, not copied.
 
 For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
+repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to the right child.
 This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
@@
 
-Having described single-user actions,
-we turn our attention to how Grove handles multiple users.
-This section generalizes to any number of users,
-but for simplicity we consider only two in our examples: Alice and Bob.
-Alice and Bob each maintain their own editor instances and perform
-actions relative to its state (using their own cursors, which we discuss in \autoref{sub:Cursors}).
+We now turn our attention to how Grove handles multiple users.
+The examples in this section generalize to any number of users,
+but for simplicity we consider only two: Alice and Bob.
+Alice and Bob each maintain their own editor instances independently
+and perform actions relative to its state using their own cursors, as we discuss in \autoref{sub:Cursors}.
 
-Users coordinate by sharing with one another the graph edits corresponding to each
-action~(e.g., via direct network connections or relayed through a server).
-TODO: CITE(how to share network data -- DHT? consensus? ).
+Users collaborate by sharing graph updates corresponding to user edits,
+e.g., via direct network connections or relayed through a server.
+% TODO: CITE(how to share network data -- DHT? consensus? ).
 The granularity of communication may vary.
-For real-time collaborative editing, edits may be sent as they occur.
-For version control, edits may be batched together in the form of commits.
-As we will show, these edits are commutative, so editors converge to the same state
-when all edits are shared.
-We assume users have some way to create globally unique identifiers for
-verticies and edges without coordination~(e.g.,~GUIDs TODO:CITE:GUIDS).
+For real-time collaboration, graph updates may be sent as they occur.
+For asynchronous version control, updates may be batched together in the form of commits.
+As we will show, since these edits are commutative,
+all instances converge to the same state once all updates are shared.
+We assume users have some way to create globally unique identifiers
+% for verticies and edges
+without coordination~(e.g.,~GUIDs).
+% TODO:CITE:GUIDS).
 
@@
 
-We start each user where we left off in the previous section with the state shown in \autoref{fig:Move:b}.
-Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{},
-and Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before transmitting their edits to each other,
-Alice and Bob thus have the graphs in \autoref{fig:DifferentParts:a}
-and \autoref{fig:DifferentParts:b}, respectively.
-(Note that the transition edge from \autoref{fig:Move:b} to
-\autoref{fig:DifferentParts:b} represents
-multiple graph-edit steps, i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
-We thus mark it with a star.)
-
-Alice and Bob then transmit their graph edits to each other
-and apply the other's graph edits to their own copy of the graph.
-This results in both Alice and Bob having the graph in
-\autoref{fig:DifferentParts:c}.
+Alice and Bob start where Alice left off in the previous section
+with the state \texttt{\hole{} + \hole{} * y} shown in \autoref{fig:Move:b}.
+Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{}.
+At the same time, Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
+Before either shares their graph updates with the other,
+Alice and Bob have the graphs in \autoref{fig:DifferentParts:a} and \autoref{fig:DifferentParts:b}, respectively.
+Note that the transition from \autoref{fig:Move:b} to \autoref{fig:DifferentParts:b}
+represents multiple graph updates,
+i.e., deleting~\eSimpleY{} and adding~\eDifferentPartsBob{} along with its child~\vDifferentPartsBob{}.
+We thus mark the transition with a star.
+Once Alice and Bob share their graph edits and apply each other's updates to their own copy of the graph,
+they both have the graph in \autoref{fig:DifferentParts:c}.
 
@@
 
-\todo{rev}
-Note that aside from transmitting edge deletions and additions,
+% \todo{rev}
+Note that aside from sharing edge deletions and additions,
 merging requires no coordination between Alice and Bob.
@@
 but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order and will always get
-the graph in \autoref{fig:DifferentParts:c}.
+Chris can apply the edits from Alice and Bob in any order
+and will always get the graph in \autoref{fig:DifferentParts:c}.
 
-In other words, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
+In other words, our update model obeys the commutativity shown in \autoref{fig:Commutativity}.
 For any sequences of edits from Alice and Bob,
 applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:informal}.)
+(We show this formally in \autoref{sub:Commutativity:formal}.)
 
@@
 
-After Alice and Bob share their edits to get the graph in \autoref{fig:DifferentParts:c},
-Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{}, which results in the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob moves~\vSimpleTimes{} and its children from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}, which results in the graph in \autoref{fig:NestedParts:b}.
-Since edits are based on edge identifiers and not locations in a tree,
-both Alice's and Bob's edits can be applied to the other's graph,
-which results in Alice and Bob both having the graph in \autoref{fig:NestedParts:c}.
+After all of Alice's and Bob's updates have been shared,
+they both have the edit state \texttt{\hole{} + u * y} in \autoref{fig:DifferentParts:c}.
+Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{},
+producing the graph in \autoref{fig:NestedParts:a}.
+Meanwhile, Bob repositions~\vSimpleTimes{} (along with its children)
+from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
+by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}.
+The resulting graph is shown in \autoref{fig:NestedParts:b}.
+Because updates are located by edge identifiers instead of by paths relative to the root vertex,
+each of Alice's and Bob's updates can be applied to the other's graph
+and afterwards they will both have the graph in \autoref{fig:NestedParts:c}.
 
-Commutativity has consequences for this scenario.
-Commutativity require that edits produce the same result regardless of the order in which they are applied.
-So, if Alice receives Bob's deletion of~\eMoveTimes{} then makes her
-changes before receiving Bob's addition of~\eNestedPartsBob{},
+Note that commutativity has consequences in this scenario.
+Since all updates must eventually lead to the same edit state regardles of the order in which they are applied,
+if Alice, say, receives Bob's deletion of~\eMoveTimes{}, makes some local changes,
+and then receives Bob's addition of~\eNestedPartsBob{},
 the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, her edits would then temporarily be within the code deleted by Bob.
-Thus, our semantics must allows users to edit deleted code.
-(Whether or how this is presented in the user interface is another question.)
+In that case, Alice's edits would then temporarily be within the term deleted by Bob.
+Thus, our semantics must allows users to edit deleted terms.
+Whether or how this is presented in the user interface is another question.
 
-TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
+% TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
 
@@
 
-While the collaborative edits discussed so far merge cleanly,
-they can also generate graphs that do not map cleanly to a syntax tree.
-We call these conflicts and consider cases for \emph{multi-child},
-\emph{multi-parent}, and \emph{cycle} conflicts.
-
-As with conflict in line/diff-based version-control
-system~(e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.),
-these all require user intervention to resolve.
-
-We consider common merge scenarios first, and then consider more unusual scenarios.
+Although the collaborative edits discussed so far merge cleanly,
+they can also lead to graphs that do not map cleanly to a syntax tree.
+We call these states \emph{conflicts}
+and consider separate cases for \emph{multi-child}, \emph{multi-parent}, and \emph{unicycle} conflicts.
+As with merge conflicts in version-control systems such as git, these all require user intervention to resolve.
+We consider common merge scenarios first, then move onto more unusual scenarios.
 
@@
 
-As an example of a multi-chid conflict, suppose Alice and Bob both have the graph in \autoref{fig:NestedParts:c}.
-Alice then adds~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}
-while Bob also adds~\vMultiChildBob{} as the~\texttt{R} child of~\vWrapPlus{}.
-These result in Alice and Bob having the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges, \eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
+Suppose Alice and Bob both have the edit state \texttt{w * v + \hole} in \autoref{fig:NestedParts:c}.
+Alice moves the cursor to the hole and constructs~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}.
+At the same time, Bob constructs~\vMultiChildBob{} in the same position.
+Now Alice and Bob have the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
+Note that these graphs have two different edges,
+\eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
 
 We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges going out of the \texttt{R} position of \vWrapPlus{}.
-This results in \autoref{fig:MultiChild:c}, where there is a multi-child conflict
-in the~\texttt{R} position of~\vWrapPlus~(i.e., there are multiple edges for the same position in a vertex).
-In the user interface, we flag this conflict with the notation~\conflictHole{x,y}.
+as edges originating from the \texttt{R} position of \vWrapPlus{}.
+This results in a multi-child conflict in the~\texttt{R} position of~\vWrapPlus,
+shown in \autoref{fig:MultiChild:c},
+where there are multiple edges originating from the same position of a vertex.
+In the user interface, we flag this conflict with the notation~$\conflictHole{x,y}$.
 
-TODO: "delete both"
+% TODO: "delete both"
 
-This conflict can be resolved by the user by deleting
-one of the edges in conflict and editing the expression pointed to by the other edge to be the correctly merged value.
-For example, one could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} to be~\texttt{x * y}.
-(Note that users can continue to edit the code while in this conflicted state.
-It just must be resolved before the the code is run.\footnote{It
-  may be possible to develop evaluation models that allow these sorts of conflicts,
-  but that is beyond the scope of this paper.})
-
-TODO: need example of fully-automatic resolution and semi-automatic resolution.
-Note that as a convenience to the user, certain simple
-conflicts might be automatically resolved,
-but we consider this a higher-level, user-interface consideration.
-
-This handling of conflicts is similar traditional merge conflicts in that
-it presents alternate versions of the code that the user replaces with the desired merge result.
-Our model is different in that merge conflicts in the line-based model
-prevent semantic editor services from running
-that could continue to run in the graph-based model because
-code inside and outside the conflict is still valid code.
-It is just the where the multi-child happens that editor services may break down.
+Such conflicts can be resolved by deleting one of the conflicting edges
+and editing the remaining term into the correctly merged value.
+For example, Alice could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} into~\texttt{x * y}.
+% To ensure users can continue editing while in a conflicted state,
+% all conflicts must be resolved before a program can be run.\footnote{It
+%   may be possible to develop evaluation models that allow these sorts of conflicts,
+%   but that is beyond the scope of this paper.})
+% % TODO: need example of fully-automatic resolution and semi-automatic resolution.
+% Note that as a convenience to the user, certain simple conflicts might be automatically resolved,
+% but we consider this a higher-level, user-interface consideration.
+Handling conflicts in this way is similar to handling traditional merge conflicts
+in that it presents alternate versions of the conflicted state
+that must be transformed by users into the desired result.
+Our approach is different in that merge conflicts in a line-based model
+may break semantic editor services, such as semantic search or type-preserving refactorings,
+that could continue to work inside and outside the conflict in a graph-based model.
+% It is just the where the multi-child happens that editor services may break down.
 \autoref{sec:Related Work} discusses more differences.
@@
 
-While multi-child conflicts are when multiple edges are in the same \emph{child} position of a vertex,
-multi-parent conflicts are when there are multiple \emph{parents} of a vertex.
-An example of this is in \autoref{fig:MultiParent}.
-\autoref{fig:MultiParent:a} shows the graph after Alice and Bob have made some edits,
-and Alice and Bob shared their edits with each other.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{},
-while Bob moves it to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the graphs in \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}.
-In both cases, these edits are achieved with one edge delete and one edge add.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}, while
-Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As explained in \autoref{sec:Formalism}, deletes are idempotent so Alice and Bob both deleting~\eNestedPartsAlice{} is not a problem.
-However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob) point to the same vertex.
-When their edits are merged, the result is \autoref{fig:MultiParent:d}, in which~\vNestedPartsAlice{} has two parents pointing to it.
+Multi-parent conflicts occur when multiple edges share a destination vertex,
+as opposed to multi-child conflicts which share a common origin.
+An example of this is in \autoref{fig:MultiParent},
+which shows the edit state Alice and Bob converge on after deleting the conflict and the \texttt{v}.
+Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{}.
+At the same time, Bob also moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vWrapPlus{}.
+This results in the edit states of \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}
+In both cases, the edits are modeled as one edge deletion followed by one edge addition.
+Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}.
+At the same times, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
+As we explain in \autoref{sec:Formalism}, deletions are idempotent,
+so the fact that Alice and Bob both deleted~\eNestedPartsAlice{} will not lead to a conflict.
+However, both~\eMultiParentAlice{} (added by Alice) and~\eMultiParentBob{} (added by Bob)
+point to the same vertex.
+Once all updates are merged, the resulting edit state is given in \autoref{fig:MultiParent:d}.
+Notice~\vNestedPartsAlice{} has two edges pointing to it.
 

### 2024-03-25 formalism.tex
feat: latest changes - uncomment sections

@@
 
-% TODO: point of this section is commutativity theorem
+TODO: point of this section is commutativity theorem
 
-% TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
+TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
 
-% The formalism is rather simple, \autoref{sec:Grove By Example} shows
-% that it is sufficient to handle many common situations,
-% and this simplicity aids in the predictability.
+The formalism is rather simple, \autoref{sec:Grove By Example} shows
+that it is sufficient to handle many common situations,
+and this simplicity aids in the predictability.
 
-% \figureArity
+\figureArity{}
 
@@
 
-% Let $\U$ be a set of \emph{unique identifiers} equipped with a total ordering $\leq$,
-% and $\K$ the \emph{constructors} of terms in the source language,
-% and $\P$ the \emph{positions} of subterms in the source language,
-% and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
-%   \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ that forms a lattice over $\Sigma$.
-% Now, let $\V = \U \times \K$ all possible \emph{vertices},
-% and $\S = \V \times \P$ all possible \emph{sources},
-% and $\E = \U \times \S \times \V$ all possible \emph{edges}.
+Let $\U$ be a set of \emph{unique identifiers} equipped with a total ordering $\leq$,
+and $\K$ the \emph{constructors} of terms in the source language,
+and $\P$ the \emph{positions} of subterms in the source language,
+and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
+  \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ that forms a lattice over $\Sigma$.
+Now, let $\V = \U \times \K$ all possible \emph{vertices},
+and $\S = \V \times \P$ all possible \emph{sources},
+and $\E = \U \times \S \times \V$ all possible \emph{edges}.
 
@@
 % TODO: put this in a definition box
-
 A graph $G$ is well sorted if
@@
 such that $G(\e) \neq \bot$,
-we have $(p, \sort(k_2)) \in \arity(k_1)$.
+we have $(p, \sort (k_2)) \in \arity k_1$.
 
@@
 
-% In decomp section:
-% - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
-%   - want to show that expr is a total function on Exp's
-% - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
-%   - every vertex should be part of the decomposition
-% - correctness of grove components:
-%   - e_r: rootVertex(e_r) = v_r
-%   - MP: for every expression in MP, |parents(the root vertex)| > 1
-%   - NP: for every expression in NP, |parents(the root vertex)| = 0
-%   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
-%     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
-%   - we've accounted for every edge:
-%     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
-%     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
+In decomp section:
+- If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
+  - want to show that expr is a total function on Exps
+- If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
+  - every vertex should be part of the decomposition
+- correctness of grove components:
+  - $e_r$: rootVertex($e_r) = v_r$
+  - MP: for every expression in MP, |parents(the root vertex)| > 1
+  - NP: for every expression in NP, |parents(the root vertex)| = 0
+  - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
+    - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
+  - we've accounted for every edge:
+    - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
+    - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
 
-% correctness covers:
-% - invertibility (w/ some identity assignment)
-% - parentage (things are multiparented correctly)
+correctness covers:
+- invertibility (w/ some identity assignment)
+- parentage (things are multiparented correctly)
 
@@
 expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
-sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
+sub-term at some position $p \in \arity k$, select unique id $u'$ and
 constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
 = (u', k')$. Continuing in this manner, treat each sub-term as if it were the
-outermost, starting with the ones at positions in $\arity(k')$, and create edges
+outermost, starting with the ones at positions in $\arity k'$, and create edges
 to the vertices representing their immediate sub-terms until only empty holes
@@
 
-\figureTermSyntax
+\figureTermSyntax{}
 
@@
 
-% The first component $e_r$ is the expression derived from the root vertex $v_0$.
-% \autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
-% There is one rule for each term constructor.
-% For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
-% In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
-% An empty hole $\hole$ appears if there are no children at the given position.
-% A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
-% If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
-% For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
+The first component $e_r$ is the expression derived from the root vertex $v_0$.
+\autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
+There is one rule for each term constructor.
+For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
+In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
+An empty hole $\hole$ appears if there are no children at the given position.
+A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
+If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
+For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
 
@@
 
-% Mapping to b actions
+Mapping to b actions
 
-% \[
-%   \begin{array}{l}
-%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
-%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
-%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
-%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
-%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
-%     \CC(G) = \ldots                                                                                                                                    \\
-%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
-%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
-%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
-%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
-%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
-%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
-%   \end{array}
-% \]
+\[
+  \begin{array}{l}
+    \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
+    \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
+    \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
+    \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
+    \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
+    \CC(G) = \ldots                                                                                                                                    \\
+    \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
+    \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
+    \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
+    \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
+    \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
+    \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
+  \end{array}
+\]

### 2024-03-25 implementation.tex
feat: latest changes - uncomment sections

--- 2021-07-27 grove-paper-5-implementation.tex
@@
 leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}{v}{p}$. Whenever we
+leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
 encounter a vertex with multiple parents, we add its corresponding expression to
@@
 
-% \begin{figure}
-%   \[
-%     \arraycolsep=0pt
-%     \begin{array}{lrlll}
-%       \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
-%       \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
-%     \end{array}
-%   \]
-%   \caption{Syntax of graph decomposition trees as a grammar.}
-%   \label{fig:Syntax of graph decomposition trees}
-% \end{figure}
+\begin{figure}
+  \[
+    \arraycolsep=0pt
+    \begin{array}{lrlll}
+      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Vertex}(v, m) \mid \textrm{Ref}(v) \\
+      \textrm{Maps:}         & m & {}\in Map & {}::={} & \varnothing \mid p \mapsto [(\e, t), \ldots, (\e, t)]; m \\
+    \end{array}
+  \]
+  \caption{Syntax of graph decomposition trees as a grammar.}
+  \label{fig:Syntax of graph decomposition trees}
+\end{figure}
 

### 2024-07-11 abstract.tex
Updates from Overleaf

\begin{abstract}
Version control systems typically rely on a \emph{patch language}, heuristic \emph{patch synthesis algorithms} like \li{diff}, and \emph{three-way merge algorithms}. Standard patch languages and merge algorithms often fail to identify conflicts correctly when there are multiple edits to one line of code or code is relocated. This paper introduces Grove, a collaborative structure editor calculus  that eliminates patch synthesis and three-way merge algorithms entirely. 
Instead, patches are derived directly from the log of the developer's edit actions 
and all edits commute, i.e. the repository state forms a commutative replicated data type (CmRDT). To handle conflicts that can arise due to code relocation, 
the core datatype in Grove is a labeled directed multi-graph with uniquely identified 
vertices and edges. All edits amount to edge insertion and deletion, with deletion being permanent. 
To support tree-based editing, we define a decomposition from graphs into \emph{groves}, which are a set of syntax trees with conflicts---including multi-child, multi-parent, and cycle conflicts---represented explicitly using holes and references between trees. Finally, we define a type error localization system for groves that enjoys a \emph{totality} property, i.e. all editor states in Grove are statically meaningful, so developers can use standard editor services while 
working to resolve these explicitly represented conflicts. The static semantics is defined as a bidirectional marking system in line with recent work, with gradual typing employed to handle situations where errors and conflicts 
prevent type determination. We then layer on a unification-based type inference system  to opportunistically fill type holes and fail gracefully when no solution exists. We mechanize the metatheory of Grove using the Agda theorem prover. We  implement these ideas as the \emph{Grove Workbench}, which generates the necessary data structures and algorithms in OCaml given a syntax tree specification.
\end{abstract}

### 2024-07-11 config/config-graphs.tex
Updates from Overleaf

% Calling \newvertex{Foo}{bar} defines
%   \vidFoo to be a new id number, and
%   \vFoo to be \texttt{bar}\ensuremath{_{\vidFoo}}
\newcounter{NodeVertexCounter}
\newcommand{\newvertex}[2]{%
\ifodd\theNodeVertexCounter
  \addtocounter{NodeVertexCounter}{1}%
\else
  \addtocounter{NodeVertexCounter}{2}%
\fi
\expandafter\newcommand\csname vid#1\endcsname{}% This is just to check if this is a redefinition
\expandafter\global\expandafter\edef\csname vid#1\endcsname{\theNodeVertexCounter}%
\expandafter\newcommand\csname v#1\endcsname{}% This is just to check if this is a redefinition
\expandafter\gdef\csname v#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname vid#1\endcsname}}}}%
}
\newcommand{\newedge}[2]{%
\ifodd\theNodeVertexCounter
  \addtocounter{NodeVertexCounter}{2}%
\else
  \addtocounter{NodeVertexCounter}{1}%
\fi
\expandafter\newcommand\csname eid#1\endcsname{}% This is just to check if this is a redefinition
\expandafter\global\expandafter\edef\csname eid#1\endcsname{\theNodeVertexCounter}%
\expandafter\newcommand\csname e#1\endcsname{}% This is just to check if this is a redefinition
\expandafter\gdef\csname e#1\endcsname{\texttt{#2}\ensuremath{_{\id{\csname eid#1\endcsname}}}}%
}
\setcounter{NodeVertexCounter}{-1}
\newvertex{Root}{Root}

% Support \includegraphics of .dot files
\DeclareGraphicsRule{.dot}{pdf}{.pdf}{`dot -Tpdf #1 -o \noexpand\OutputFile}

### 2024-07-11 decomp-recomp.tex
Updates from Overleaf

\documentclass[acmsmall,dvipsnames,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}

% \title{Decomp-Recomp}

\newcommand{\edgevar}[0]{\ensuremath{\epsilon}}
\newcommand{\set}[1]{\{#1\}}

\begin{document}
% \maketitle

\begin{definition}
    U-top(v) holds if and only if there exists a sequence w(i), $1 \leq i \leq n$, where $n > 1$, such that: 
    \begin{itemize}
        \item w(i+1) is the unique parent of w(i) for all $1\leq i < n$
        \item w(1) = v = w(n)
        \item the id of v is the min of the ids of w(i)
    \end{itemize}
\end{definition}

\begin{definition}
    X-inner(v, w) holds if and only if not U-top(v), v has unique parent v’, and either w = v’ and X-top(w), or X-inner(v’ ,w).
\end{definition}

\begin{lemma}
\label{lem:u-top}
    If U-top(v), then v has unique parent w such that either w = v or U-inner(w, v).
\end{lemma}
\begin{proof}

    $v=w(1)$ has unique parent $w=w(2)$. If $n=2$, then $w=w(2)=v$, and we have $w=v$.

    Otherwise, consider the minimum $1<m\leq n$ such that $v=w(m)$.

    Helper fact: not U-top(w(i)) for all $1 < i < m$.

    \

    Prove X-inner(w(m-i), v) for all $1\leq i < m-1$, by induction on $i$.

    Base case: w(m-1). X-inner(w(m-1), v) because not U-top(w(m-1)) by the helper, w(m-1) has unique parent w(m)=v, and U-top(v).

    Inductive step: w(m-i-1). X-inner(w(m-i-1), v) because not U-top(w(m-i-1)) by the helper, w(m-i-1) has unique parent w(m-i), and by the induction hypothesis, X-inner(w(m-i), v). 
\end{proof}

\begin{theorem}
\label{thm:vertex-class}
    For each vertex v, one and only one of the following holds, and this classification is decidable:
\begin{itemize}
    \item NP-top(v)
    \item MP-top(v) 
    \item U-top(v)
    \item NP-inner(v,w)
    \item MP-inner(v,w)
    \item U-inner(v,w)
\end{itemize}
\end{theorem}
% \begin{proof}
%     (uses lemma \ref{lem:u-top})
% \end{proof}

\begin{corollary}
\label{cor:edge-class}
    For each edge \edgevar, one and only one of the following holds, and this classification is decidable:
\begin{itemize}
    \item NP-edge(\edgevar)
    \item MP-edge(\edgevar) 
    \item U-edge(\edgevar)
\end{itemize}
\end{corollary}

\begin{lemma}
    If X-inner(v, w), then X-top(w).
\end{lemma}

\begin{lemma}
    If X-edge(e, v), then X-top(v).
\end{lemma}

\begin{lemma}
    For each vertex v, v-of(decomp’(v)) = v.
\end{lemma}

\begin{lemma}
\label{lem:redecomp-eq}
    For each vertex v, recomp(decomp(v)) = $\bigcup_{(E\ u`\ v\ p\ v`) \in G}$ \set{(E u’ v p v’)} $\cup$ recomp(decomp’(v’)).
\end{lemma}

\begin{lemma}
\label{lem:parent-top-inner}
    If (E u’ v p v’) and X-top(v), and not MP-top(v’), and not U-top(v’), then X-inner(v’, v).
\end{lemma}

\begin{lemma}
\label{lem:parent-inner-inner}
    If (E u’ v p v’) and X-top(v), and not MP-top(v’), and not U-top(v’), then X-inner(v’, v).
\end{lemma}

\begin{lemma}
\label{lem:redecomp-side1}
    For each w s.t. X-inner(w, v), for each e in recomp(decomp(w)), X-edge(e, v).
\end{lemma}
% uses {lem:parent-inner-inner}

\begin{lemma}
\label{lem:redecomp-source}
    e is in recomp(decomp(source of e))
\end{lemma}

\begin{lemma}
\label{lem:redecomp-side2}
    If X-inner(w, v), then recomp(decomp(w)) $\subseteq$ recomp(decomp(v)). 
\end{lemma}

\begin{lemma}
\label{lem:redecomp-top}
    For each vertex v such that X-top(v), recomp(decomp(v)) = \set{\edgevar s.t. X-edge(\edgevar,v)}.
\end{lemma}
% \begin{proof}
%     (uses lemmas \ref{lem:redecomp-eq}, \ref{lem:redecomp-side1}, \ref{lem:parent-top-inner} for first direction, uses {lem:redecomp-source} and {lem:redecomp-side2} for second direction)
% \end{proof}

\begin{theorem}
\label{thm:redecomp}
    For each graph G, recomp(decomp(G)) = G. 
\end{theorem}
% \begin{proof}
%     (uses lemma \ref{lem:redecomp-top})
% \end{proof}

\end{document}

### 2024-07-11 discussion-and-conclusion.tex
Updates from Overleaf

--- 2021-04-09 grove-paper-discussion-and-conclusion.tex
@@
 \label{sec:Discussion and Conclusion}
+
+Contribution 1 and Contribution 2 together result in a typed collaborative structure calculus
+where, uniquely, all edits, including code relocations that stymie existing approaches, commute and where there are no semantic gaps: all possible editor states, including editor states with various kinds of unresolved conflicts, are semantically meaningful.
+
+In practice, it may be helpful to garbage collect orphaned vertices once there is consensus across collaborators that the deletion is permanent, but we do not consider this consensus protocol formally in this paper.
+In an open-ended collaboration scenario (where the set of collaborators is not known, e.g.
+on GitHub), we simply retain all vertices.
 

### 2024-07-11 figures/figure-arity.tex
Updates from Overleaf

\newcommand{\figureArityContent}{%
\[
  \arraycolsep=0pt
  \begin{array}{ll}
    \multicolumn{2}{l}{\arityOp : \K \rightarrow \wp(\P \times \{Exp, Pat, Typ\})} \\
    \hline
    \arity{\Root}={} & \left\{ (\Root, Exp) \right\} \\
    \arity{\PatVar(x)}={} & \left\{ \right\} \\
    \arity{\ExpVar(x)}={} & \left\{ \right\} \\
    \arity{\ExpLam}={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
    \arity{\ExpApp}={} & \left\{ (\AppFun, Exp), (\AppArg, Exp) \right\} \\
    \arity{\ExpPlus}={} & \left\{ (\PlusLeft, Exp), (\PlusRight, Exp) \right\} \\
    \arity{\ExpTimes}={} & \left\{ (\PlusLeft, Exp), (\PlusRight, Exp) \right\} \\
    \arity{\ExpNum(n)}={} & \left\{ \right\} \\
    \arity{\TypArrow}={} & \left\{ (\ArrowArg, Typ), (\ArrowResult, Typ) \right\} \\
    \arity{\TypNum}={} & \left\{ \right\} \\
  \end{array}
\]%
}

\newcommand{\figureArity}{%
\begin{figure}
\figureArityContent
\caption{Constructors, Indexes and Arity}
\Description{This figure shows the mapping from constructors to their corresponding indexes and arity. For each constructor, such as \ExpLam or \ExpApp, the arity is defined in terms of the parts that make up the constructor. For example, \ExpLam has three parts: a pattern (\LamParam), a type (\mathtt{Type}), and a body (\mathtt{Body}), each with their corresponding expression, pattern, or type.}
\label{fig:Constructors, Indexes and Arity}
\end{figure}%
}

### 2024-07-11 figures/figure-cycle.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureCycle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiCycleTimes}{L}
\newvertex{MultiCycleTimes}{*}
\newedge{MultiCyclePlus}{R}
\newvertex{MultiCyclePlus}{+}
\newedge{MultiCycleAliceTimes}{R}
\newedge{MultiCycleAlicePlus}{L}
\newedge{MultiCycleBobPlus}{R}
\newedge{MultiCycleBobTimes}{L}
\newcommand{\figureCycle}{
\begin{figure*}
%\hfill
\hskip0.07\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.2\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
%%%%
\path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
\end{tikzpicture}
}%
\label{fig:Cycle:a}
\figureCode{\parens{\hole{} * \hole{}} * \parens{\hole{} + \hole{}} + \hole{}%
  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.2\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Cycle/b) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Cycle/a) to[star] (-2em, 0cm |- node/Cycle/b);
\end{tikzpicture}
}%
\label{fig:Cycle:b}
\figureCode{\hole{} * \hole{} + \parens{\hole{} + \hole{}} * \hole{}%
  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
   },
   {
    "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes", >alice edge, alice node] -> {
     "\vMultiCyclePlus" [> "\eMultiCycleAlicePlus"', >alice edge, alice node],
     {}
    }
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.2\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Cycle/c) [anchor=base] {\strut};
\path [draw,->,bob step] (node/Cycle/a) to[out=15,in=165,star] (-2em, 0cm |- node/Cycle/c);
\end{tikzpicture}
}%
\label{fig:Cycle:c}
\figureCode{\hole{} * \hole{} + \parens{\hole{} * \hole{} + \hole{}}%
  \otherVertexVskip\phantom{\multiVertex{30} = \multiVertex{32} * \hole{}}%
  \\\phantom{\multiVertex{32} = \multiVertex{30} * \hole{}}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
   },
   {
    "\vMultiCyclePlus" [> "\eMultiCycleBobPlus", >bob edge, bob node] -> {
     "\vMultiCycleTimes" [> "\eMultiCycleBobTimes"', >bob edge, bob node],
     {}
    }
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.2\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Cycle/d) [anchor=base] {\strut};
\path [draw,->,merge step] (node/Cycle/b) to[out=15,in=165] (-2em, 0cm |- node/Cycle/d);
\path [draw,->,merge step] (node/Cycle/c) to (-2em, 0cm |- node/Cycle/d);
\end{tikzpicture}
}%
\label{fig:Cycle:d}
\figureCode{\parens{\hole{} * \hole{}} + $\conflictHole{\multiVertex{\vidMultiCycleTimes},\multiVertex{\vidMultiCyclePlus}}$%
  \otherVertexVskip\multiVertex{\vidMultiCycleTimes} = \multiVertex{\vidMultiCyclePlus} * \hole{}%
  \\\multiVertex{\vidMultiCyclePlus} = \multiVertex{\vidMultiCycleTimes} * \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
   },
   v1/"\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"anchor=155, >merge edge, merge node],
   {},
   v2/"\vMultiCyclePlus" [> "\eMultiCycleBobPlus"anchor=-165, >merge edge, merge node]
  }
 }
};
\path[draw,-{>[bend]},merge edge] (v1) to ["\eMultiCycleAlicePlus"anchor=south,out=-60,in=-120] (v2);
\path[draw,-{>[bend]},merge edge] (v2) to ["\eMultiCycleBobTimes"anchor=north,out=-90,in=-90] (v1);
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
%\hfill{}
\caption{Example of Cycles.}
% TODO replace screenshots with text
% TODO make ref code gray
% TODO move whole figure left
% TODO remove cursor from screenshots
% TODO add multiparent box
% TODO only show parens where needed - figure 10, 11 - by associativity
% TODO repplace cycle screenshot with cycles box %
\Description{This figures shows an example of a cycle}
\label{fig:Cycle}
\end{figure*}
}

### 2024-07-11 figures/figure-different-parts-nested-parts.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureDifferentParts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{DifferentPartsAlice}{L}
\newvertex{DifferentPartsAlice}{u}
\newedge{DifferentPartsBob}{R}
\newvertex{DifferentPartsBob}{v}
% \newcommand{\figureDifferentParts}{
\newcommand{\figureDifferentPartsNestedParts}{
\begin{figure}
\begin{minipage}[t]{0.45\linewidth}
%\hfill
\hskip0.12\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
\path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
\end{tikzpicture}
}%
\label{fig:DifferentParts:a}
\figureCode{\hole{} + u * y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
  {},
  "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >alice edge, alice node],
    "\vSimpleY" [> "\eSimpleY"]
  }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/DifferentParts/b) [anchor=base] {\strut};
\path [draw,->,bob step] (node/Move/bx) to [out=15,in=165,star] (-2em, 0cm |- node/DifferentParts/b);
\end{tikzpicture}
}%
\label{fig:DifferentParts:b}
\figureCode{\hole{} + \hole{} * v}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
  {},
  "\vSimpleTimes" [> "\eMoveTimes"] -> {
    {},
    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >bob edge, bob node]
  }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/DifferentParts/c) [anchor=base] {\strut};
\path [draw,->,merge step] (node/DifferentParts/a) to [out=15,in=165] (-2em, 0cm |- node/DifferentParts/c);
\path [draw,->,merge step] (node/DifferentParts/b) to (-2em, 0cm |- node/DifferentParts/c);
\end{tikzpicture}
}%
\label{fig:DifferentParts:c}
\figureCode{\hole{} + u * v}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
  {},
  "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"', >merge edge, merge node],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob", >merge edge, merge node]
  }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{(a) Alice fills the hole in the multiplication with \li{u}. (b) Bob renames \li{y} to \li{v}. (c) These  edits to the same expression commute, addressing the \textbf{granularity problem}.}%
\label{fig:DifferentParts}
\end{minipage}
% \end{figure}
% }
%
\hfil
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureCommutativity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newcommand{\figureCommutativity}{
% \begin{figure}
% \centering
% \begin{figure}
% \centering
% \begin{tikzpicture}
% \path (-3cm, 0cm) node (a) [align=center]       {Original \\ Version};
% \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
% \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
% \path ( 3cm, 0cm) node (d) [align=center]       {Combined \\ Version};
% \path [draw,->,alice step] (a) -- node [pos=0.7,auto] {Alice's Edits} (b);
% \path [draw,->,bob step]   (a) -- node [pos=0.7,auto,swap]      {Bob's Edits}   (c);
% \path [draw,->,merge step] (b) -- node [pos=0.3,auto] {Sync} (d);
% \path [draw,->,merge step] (c) -- node [pos=0.3,auto,swap]      {Sync} (d);
% \end{tikzpicture}
% \caption{TODO:Commutativity}
% \label{fig:Commutativity}
% \end{figure}
% }
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureNestedParts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{NestedPartsAlice}{L}
\newvertex{NestedPartsAlice}{w}
\newedge{NestedPartsBob}{L}
% \newcommand{\figureNestedParts}{
% \begin{figure}
\begin{minipage}[t]{0.45\linewidth}
\hfill
%\hskip0.12\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/NestedParts/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/DifferentParts/c) to[star] (-2em, 0cm |- node/NestedParts/a);
\end{tikzpicture}
}%
\label{fig:NestedParts:a}
\figureCode{\hole{} + w * v}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >alice edge, alice node],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/NestedParts/b) [anchor=base] {\strut};
\path [draw,->,bob step] (node/DifferentParts/c) to [out=15,in=165,star] (-2em, 0cm |- node/NestedParts/b);
\end{tikzpicture}
}%
\label{fig:NestedParts:b}
\figureCode{u * v + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"', >bob edge] -> {
    "\vDifferentPartsAlice" [> "\eDifferentPartsAlice"'],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/NestedParts/c) [anchor=base] {\strut};
\path [draw,->,merge step] (node/NestedParts/a) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
\path [draw,->,merge step] (node/NestedParts/b) to [out=15,in=165] (-2em, 0cm |- node/NestedParts/c);
\end{tikzpicture}
}%
\label{fig:NestedParts:c}
\figureCode{w * v + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"', >merge edge] -> {
    "\vNestedPartsAlice" [> "\eNestedPartsAlice"', >merge edge, merge node],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{(a) Alice renames \li{u} to \li{w}. (b) Bob relocates the multiplication to the left side. (c) These edits commute without conflict or duplication, addressing the \textbf{relocation modification problem}.}%
\Description {This figure describes users editing nested parts of code}
\label{fig:NestedParts}
\end{minipage}

\end{figure}
}

### 2024-07-11 figures/figure-multi-child.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMultiChild
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiChildAlice}{R}
\newvertex{MultiChildAlice}{x}
\newedge{MultiChildBob}{R}
\newvertex{MultiChildBob}{y}
\newcommand{\figureMultiChild}{
\begin{figure}
%\hfill
\hskip0.12\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
%%%%
\path (0cm,0cm) node (node/MultiChild/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/NestedParts/cx) to (-2em, 0cm |- node/MultiChild/a);
\end{tikzpicture}
}%
\label{fig:MultiChild:a}
\figureCode{w * v + x}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   },
   {
    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/MultiChild/b) [anchor=base] {\strut};
\path [draw,->,bob step] (node/NestedParts/cx) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/b);
\end{tikzpicture}
}%
\label{fig:MultiChild:b}
\figureCode{w * v + y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   },
   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.4\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/MultiChild/c) [anchor=base] {\strut};
\path [draw,->,merge step] (node/MultiChild/a) to[out=15,in=165] (-2em, 0cm |- node/MultiChild/c);
\path [draw,->,merge step] (node/MultiChild/b) to (-2em, 0cm |- node/MultiChild/c);
\end{tikzpicture}
}%
\label{fig:MultiChild:c}
\figureCode{w * v + $\conflictHole{x,y}$}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedPartsBob"'anchor=-15] -> {
    "\vNestedPartsAlice" [> "\eNestedPartsAlice"'],
    "\vDifferentPartsBob" [> "\eDifferentPartsBob"]
   },
   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
   {},
   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
%\hfill{}
\caption{Example of Multi-Child Conflicts.}%
\Description{This figure shows an example of multi-child conflicts}
\label{fig:MultiChild}
\end{figure}
}

### 2024-07-11 figures/figure-simple.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureSimple
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{SimpleTimes}{Root}%
\newvertex{SimpleTimes}{*}%
\newedge{SimpleX}{L}%
\newvertex{SimpleX}{x}%
\newedge{SimpleY}{R}%
\newvertex{SimpleY}{y}%
\newcommand{\figureSimple}{
\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Simple/a) [anchor=base] {\strut};
\end{tikzpicture}
}%
\label{fig:Simple:a}
\figureCode{x * \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style, color=black] -> {
  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Simple/b) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Simple/a) to (-2em, 0cm |- node/Simple/b);
\end{tikzpicture}
}%
\label{fig:Simple:b}
\figureCode{x * y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
  root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vSimpleY" [> "\eSimpleY", >alice edge, alice node]
  }
  }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Simple/c) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Simple/b) to (-2em, 0cm |- node/Simple/c);
\end{tikzpicture}
}%
\label{fig:Simple:c}
\figureCode{\hole{} * y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
  root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
    {},
    "\vSimpleY" [> "\eSimpleY"]
  }
  }
};
\path (-0.25cm,-1.7cm) graph[graph style] {
  "\vSimpleX"
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{(a) We represent collaborative program sketches as graphs. (b) Hole filling translates to edge insertion. (c) Term deletion (or cut) deletes an edge, but the vertex persists. (We omit it in subsequent figures.)}%
\Description{This figure shows some simple edits with holes}
\label{fig:Simple}
\vspace{-10px}
\end{figure}
}

### 2024-07-11 figures/figure-wrap-move.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureWrap
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{WrapPlus}{Root}
\newvertex{WrapPlus}{+}
\newedge{WrapTimes}{L}
\newcommand{\figureWrapMove}{
\begin{figure}
\centering
\begin{minipage}[t]{.45\linewidth}
\hskip0.12\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (node/Simple/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Simple:c}}}};
%%%%
\path (0cm,0cm) node (node/Wrap/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Simple/cx) to (-2em, 0cm |- node/Wrap/a)
\end{tikzpicture}
}%
\label{fig:Wrap:a}
\figureCode{\hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (0cm,-1.7cm) graph[graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vSimpleY" [> "\eSimpleY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Wrap/b) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Wrap/a) to (-2em, 0cm |- node/Wrap/b);
\end{tikzpicture}
}%
\label{fig:Wrap:b}
\figureCode{\hole{} + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus", >alice edge, alice node]
 }
};
\path (0cm,-1.7cm) graph[graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vSimpleY" [> "\eSimpleY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.28\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Wrap/c) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Wrap/b) to (-2em, 0cm |- node/Wrap/c);
\end{tikzpicture}
}%
\label{fig:Wrap:c}
\figureCode{\hole{} * y + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eWrapTimes"', >alice edge] -> {
    {},
    "\vSimpleY" [> "\eSimpleY"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping results in a patch which (a) cuts the original term (edge deletion), (b) creates the outer term at the same location (edge insertion), then (c) pastes the original term (edge insertion).}%
\label{fig:Wrap}
\end{minipage}
% \end{figure}
% }
%
\hfil
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMove
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MoveTimes}{R}
% \newcommand{\figureMove}{
% \begin{figure}[H]
\begin{minipage}[t]{.45\linewidth}
\hfill
%\hskip0.12\columnwidth
%%%%%%%%
\begin{subfigure}[t]{0.43\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Move/a) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Wrap/c) to (-2em, 0cm |- node/Move/a);
\end{tikzpicture}
}%
\label{fig:Move:a}
\figureCode{\hole{} + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (0cm,-1.7cm) graph [graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vSimpleY" [> "\eSimpleY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.43\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (node/Move/b) [anchor=base] {\strut};
\path [draw,->,alice step] (node/Move/a) to (-2em, 0cm |- node/Move/b);
\end{tikzpicture}
}%
\label{fig:Move:b}
\figureCode{\hole{} + \hole{} * y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes", >alice edge] -> {
    {},
    "\vSimpleY" [> "\eSimpleY"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Term relocation results in a patch which (a) cuts the term (edge deletion) then (b) pastes it in its new location (edge insertion). Vertex identities and downstream edges are conserved.}%
\Description{This figures shows an example of the repositioning action}
\label{fig:Move}
\end{minipage}
\vspace{-10px}
\end{figure}
}

### 2024-07-11 formalism.tex
Updates from Overleaf

--- 2021-04-09 formalism.tex
@@
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\section{Formalism}%
+\label{sec:Formalism}
 
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+% % TODO: try to introduce groves in section 2?
+% % TODO: use key phrases in section 2, e.g.: "unresolved parentage"
 
-\setcopyright{none}
+% We will now make the intuitions developed in the previous section precise by defining a collaborative structure editor calculus called Grove.
+% This section introduces
+%   graphs (\autoref{sub:Convergent Graphs})
+%   and commutative graph patches (\autoref{sub:Graph Updates}).
+% Users do not edit graphs directly.
+% Instead, they work with the decomposition of graphs into groves (\autoref{sub:Groves})%
+% ---collections of terms with empty holes, conflict holes, and references to terms with unresolved parentage.
+% User actions on groves translate to corresponding graph edits (\autoref{sub:User Actions}).
 
-\input{config-tex}
-\input{config-tikz}
-\input{config-glyphs}
-\input{config-definitions}
-\input{config-graphs}
+% % TODO: point of this section is commutativity theorem
 
-\begin{document}
+% % TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)
 
-\title[Grove]{Excerpt: Recomposition}
+% The formalism is rather simple, \autoref{sec:Grove By Example} shows
+% that it is sufficient to handle many common situations,
+% and this simplicity aids in the predictability.
 
-\maketitle
+% \figureArity
 
-\input{appendix-formalism}
+\subsection{Convergent Graphs}%
+\label{sub:Convergent Graphs}
 
-\end{document}
+Let $\U$ and $\W$ be separate sets of \emph{unique identifiers} (UIDs) for vertices and edges, respectively. Assume each is equipped with a total ordering $\leq$. 
+Let $\K$ be the set of \emph{constructors} of terms in the source language, and let $\P$ be the set of \emph{positions} of subterms in the source language.
+% and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
+%   \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ that forms a lattice over $\Sigma$.
+% Now, let $\V = \U \times \K$ all possible \emph{vertices},
+% and $\S = \V \times \P$ all possible \emph{sources},
+% and $\E = \U \times \S \times \V$ all possible \emph{edges}.
+
+% A \emph{graph} $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
+%   where $\E = \U \times \S \times \V$,
+%   unique identifiers are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
+%   sources are drawn from $\S$,
+%   vertices are drawn from $\V$,
+%   and edge states are drawn from $\Sigma$,
+%   all defined below.
+% Graphs that represent a complete edit state have a distinguished
+%   root vertex $\rootVertex$
+%   and root position $\rootPosition$.
+
+A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is a constructor instance, where 
+    unique identifiers are drawn from $\U$ and 
+    constructors are drawn from $k \in \K$. 
+The \emph{inedges} of $v$ are the edges that target $v$.
+
+% Vertices represent \emph{constructable terms}, i.e., terms that are not references or holes.
+% The \emph{inedges} of $v$ are the edges that target $v$.
+% The \emph{parents} of $v$ are the source vertices of the inedges of $v$.
+% Not used; need unique parents and ancestors.
+% The \emph{ancestors} of $v$ is defined inductively as the parent vertices of $v$ plus their ancestors.
+% The \emph{least ancestor} of $v$ is the ancestor with the smallest unique identifier.
+
+A \emph{source} is a pair $(v, p) \in \S = \V \times \P$, where
+  vertices are drawn from $\V$,
+  and positions are drawn from $\P$.
+The \emph{outedges} of a source are the edges that have the source.
+
+An \emph{edge} $\e{=}(w, (v, p), v^\prime) \in \E = \W \times \S \times \V$ represents a directed multi-edge 
+    identified by $w \in \W$, 
+    originating from source $(v, p) \in \S$, 
+    with destination vertex $v' \in \V$.
+
+A \emph{graph} $G : \E \rightarrow \Sigma$ is a function from edges to \emph{edge states} $\Sigma = \SetOf{\bot, \Plus, \Minus}$.
+Edge states determine which edges are live in a graph.
+For example, if $G(\e) = \Plus$ for some $G : \E \rightarrow \Sigma$, $\e \in \E$,
+then we say $\e$ is \emph{live}, i.e., it is part of the live subgraph that is decomposed into a grove, as explained in \autoref{sub:Graph Decomposition}.
+The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+% \subsubsection{Constructors, Positions, Arities}
+% \label{sub:Constructors, Positions, and Arities}
+
+% A constructor $k \in \K$ represents a distinct syntactic form belonging to the term language.
+For each $k \in \K$ we define its arity \autoref{fig:Constructors, Indexes and Arity} to be its set of valid positions—labels representing potential subterms of terms constructed by $k$—each paired with a corresponding sort indicating which terms may be constructed at that position.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \subsubsection{Well sorted graphs}
+% \label{sub:Well sorted graphs}
+
+% % TODO: where does this go?
+% % 
+% % Each contructor has an arity, which we define as a set of position-sort pairs. A position $p
+% % \in \P$ identifies a location at some originating vertex from which edges to
+% % destination vertices may originate. The $\arityOp$ function in
+% % \autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
+% % constructors to potential edge positions with their corresponding sorts.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % TODO: put this in a definition box
+
+% A graph $G$ is well sorted if
+% for all $\e = (u, (u_1, k_1), p, (u_2, k_2))$
+% such that $G(\e) \neq \bot$,
+% we have $(p, \sort{k_2}) \in \arity{k_1}$.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% % In decomp section:
+% % - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
+% %   - want to show that expr is a total function on Exps
+% % - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
+% %   - every vertex should be part of the decomposition
+% % - correctness of grove components:
+% %   - $e_r$: rootVertex($e_r) = v_r$
+% %   - MP: for every expression in MP, |parents(the root vertex)| > 1
+% %   - NP: for every expression in NP, |parents(the root vertex)| = 0
+% %   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
+% %     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
+% %   - we've accounted for every edge:
+% %     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
+% %     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)
+
+% % correctness covers:
+% % - invertibility (w/ some identity assignment)
+% % - parentage (things are multiparented correctly)
+
+% \vspace*{\baselineskip}
+
+% % TODO: should graph "recomposition" have an algorithm, too?
+% % TODO: update this to match our current understanding
+% A graph can be derived from an expression in the calculus by starting with an
+% empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
+% expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
+% sub-term at some position $p \in \arity k$, select unique id $u'$ and
+% constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
+% = (u', k')$. Continuing in this manner, treat each sub-term as if it were the
+% outermost, starting with the ones at positions in $\arity k'$, and create edges
+% to the vertices representing their immediate sub-terms until only empty holes
+% remain. Note that empty holes are not modeled explicitly in the graph. Wherever
+% an empty hole appears in an expression, the corresponding graph position has no
+% edges originating from it.
+
+% % You can map from syntax to graph by selecting
+% % unique IDs $u$ (for both edges and vertexes).
+% % - holes are not explicit in the graph
+
+\subsubsection{Graph Patch Commands}
+\label{subsub:Graph Patches}
+
+We define a graph patch command $\pi = (\sigma,\e) \in \left\{\Plus, \Minus\right\} \times \E$
+as a pair of an edge state (excluding $\bot$) and an edge. As a shorthand, we define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
+
+For $\sigma_1$, $\sigma_2 \in \Sigma$, we define the join operation $\sigma_1 \sqcup \sigma_2$ to be the least upper bound of $\sigma_1$ and $\sigma_2$ with respect to the $\sqsubset$ ordering.
+Accordingly, $\Plus \sqcup \bot = \Plus$ means that edges can only be created if they do not already exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once an edge is deleted it can never be restored, and hence deletion is idempotent.
+
+We define the semantics of patch commands via the following transition relation
+between graphs.
+\[
+  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right]
+\]
+Applying patch command $\left(\sigma, \e\right)$ to graph $G$ results in the updated graph $G' = G\left[\e \mapsto \sigma \sqcup G(\e) \right]$, wherein the edge
+state associated with edge $\e$ in $G$ becomes the join of $\sigma$ with the state of
+$\e$ in $G$, and $G'(\e') = G(\e')$ for any $\e' \ne \e$.
+Essentially, these patch command semantics act as a transition system between graphs, \emph{joining} the new edge state with the corresponding edge state of the graph to which the patch command is being applied.
+A \textit{graph patch} is a sequence of graph patch commands.
+The transition relation is extended to patches as the composition of the relations of the constituent patch commands.
+
+% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
+% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).
+
+\subsubsection{Commutativity}%
+\label{subsub:Commutativity:formal}
+
+In order to define commutativity, we first define a lattice over $G$.
+Then we show that for any graphs $G$ and $G'$ and graph patch command $\pi$,
+$G \overset{\pi}\rightarrow G'$
+iff $G \sqcup \llbracket\pi\rrbracket = G'$.
+This establishes the commutativity of patch commands
+by the commutativity of the join operations defining those patch commands.
+
+% Discuss uniqueness of uuid
+
+\begin{lemma}[Join Commutativity]
+  \label{lem:Join Commutativity}
+  For all edge states $\sigma_1$ and $\sigma_2$, we have $\sigma_1 \sqcup \sigma_2 = \sigma_2 \sqcup \sigma_1$
+\end{lemma}
+\begin{proof}
+    Follows directly from the total ordering $(\Sigma, \sqsubset)$
+\end{proof}
+
+\begin{lemma}[Joining]
+  \label{lem:Joining}
+  For all graphs $G$ and $G'$ and all graph patch commands $\pi_1 = (\sigma_1, \e_1)$ and  $\pi_2 = (\sigma_2, \e_2)$, we have $G \overset{\pi_1 \pi_2}{\longrightarrow} G'$ iff $G[\e_1 \mapsto \sigma_1 \sqcup G(\e_1)][\e_2 \mapsto \sigma_2 \sqcup G(\e_2)] = G'$
+\end{lemma}
+\begin{proof}
+  If $G \overset{\pi_1\pi_2}{\longrightarrow} G'$
+  there is a graph $G''$ such that $G[\e_1 \mapsto \sigma_1 \sqcup G(\e_1)]= G''$ and $G''[\e_2 \mapsto \sigma_2 \sqcup G(\e_2)] = G'$.
+For any edge \e, if $\e = \e_1$, then $G'(\e)$ = $\sigma_1 \sqcup G''(\e_1)$, and if $\e = \e_2$, then $G'(\e)$ = $\sigma_2 \sqcup G''(\e_2)$ = $\sigma_2 \sqcup G(\e_2)$, and otherwise $G'(\e) = G(\e)$.
+\end{proof}
+
+\begin{theorem}[Commutativity]
+  \label{thm:Commutativity}
+  For all graphs $G$ and $G'$ and all graph patch commands $\pi_1 = (\sigma_1, \e_1)$ and  $\pi_2 = (\sigma_2, \e_2)$, 
+  if $G \overset{\pi_1 \pi_2}{\longrightarrow} G'$,
+  then $G \overset{\pi_2 \pi_1}{\longrightarrow} G'$
+\end{theorem}
+\begin{proof}
+  By \autoref{lem:Joining}, if $\e_1 \ne \e_2$, we have $G'(\e_1) = \sigma_1 \sqcup G'(\e_1)$ and $G'(\e_2) = \sigma_2 \sqcup G(\e_2)$, and if $\e_1 = \e_2 = \e$, we have $G'(\e) = \sigma_1 \sqcup \sigma_2 \sqcup G(\e)$.
+Hence, for any edge \e, $\e_1$ and $\e_2$ are not equal we have $G'(\e) = G[\e_1 \mapsto \sigma_1 \sqcup G(\e_1)][\e_2 \mapsto \sigma_2 \sqcup G(\e_2)](\e) = G[\e_2 \mapsto \sigma_2 \sqcup G(\e_2)][\e_1 \mapsto \sigma_1 \sqcup G(\e_1)](\e)$ and otherwise by \autoref{lem:Join Commutativity}, we have $G'(\e) = \sigma_1 \sqcup \sigma_2 \sqcup G(\e) = \sigma_2 \sqcup \sigma_1 \sqcup G(\e)$
+\end{proof}
+
+The commutativity of patch commands generalizes directly to the commutativity of patches.
+
+% \subsubsection{Agda Mechanization}%
+% \label{sub:Agda Mechanization}
+
+% \textbf{TODO}
+
+\subsection{Groves}%
+\label{sub:Groves}
+
+It is often easier to work with trees than to work directly with arbitrary graphs, such as when the program is presented to the user or type checked.
+When performing these operations on a graph $G$, only the set of edges $\e$ such that $G(\e) = \Plus$ should be considered.
+We call this the \textit{live subgraph} of $G$.
+
+We present \textit{groves}, a tree-based representation of a live subgraph.
+A grove is a set of mutually referential terms governed by the grammar in \autoref{fig:Syntax}.
+Groves correspond precisely to live subgraphs in the sense that a live subgraph can be \emph{decomposed} into a grove, and the grove can be \emph{recomposed} into the original live graph.
+
+\subsubsection{Terms}%
+\label{subsub:Terms}
+
+\figureTermSyntax
+
+% TODO: check for \hole --> \emptyHole
+
+Our source language (\autoref{fig:Syntax}) is based on a simply typed lambda calculus\todo{is there anything here that has to do with the STLC?
+we just have constructors} with an expression sort and a type sort.
+Each sort is extended with cases for multi-location references and unicycle references.
+Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define secondary sorts \textit{ExpList} and \textit{TypList} to represent these possibilities.
+A source with no outgoing edges corresponds to an empty hole, a source with one outgoing edge corresponds to an ordinary subterm, and a source with multiple outgoing edges corresponds to a local conflict between multiple terms.
+
+% TODO: Should we be more formal and say "is annotated with" instead of "carries"?
+These terms also carry information to support additional operations.
+To enable the reconstruction of the original graph, each constructor carries the UID of its original vertex, each reference carries the vertex that it refers to, and each subterm carries the UID of the edge that leads to it.
+To support type hole inference (discussion in Sec.
+TODO), each empty hole and local conflict carries its source.
+Finally, to support user navigation of the grove, each references also carries the UID of the edge that leads to it.
+
+\todo{example of an instantiation? this could be the same one we use in sec 4}
+
+% Multiparent references correspond to edges that target
+%   \emph{multiparent roots}---vertices targeted by multiple edges.
+% Unicycle references correspond to edges that target
+%   \emph{unicycle roots}---vertices that are their own least ancestor.
+
+\subsubsection{Graph Decomposition}%
+\label{subsub:Graph Decomposition}
+
+\figureDecompExample
+
+A live subgraph (or simply graph), which is just a finite set of edges, decomposes into a grove, which is a set of terms.
+In a tree, each node except the root as a unique parent, and there are no cycles.
+There is no guarantee of these properties in a graph.
+Decomposition operates by allowing each `problematic' vertex, in the sense of having multiple parents or forming a cycle, to be the root of its own entry in the grove, and to use references at special term leaves to encode the original graph structure.
+
+Formalizing this reasoning, each vertex in a graph can be classified as either an NP root (if it has no parents), an MP root (if it has multiple parents), a U root (if there is a sequence of unique parents from it to itself, and it has the minimal UID in that sequence), or not a root.
+
+\begin{theorem}
+    (Classification) TODO
+\end{theorem}
+
+The decomposition of a vertex is specified intuitively as the term obtained by traversing the descendants of the vertex until a leaf or a root vertex of some kind is reached, which not further decomposed, but is denoted in the term as the appropriate reference.
+The decomposition of a graph is simply the set obtained by decomposing each vertex classified as a root.
+
+The recomposition of a term is a set of edges, and is specified intuitively as the set of parent-child edges in the term.
+Terms store vertices on constructors and edge UIDs for each child to enable the completeness of this reconstruction.
+
+\begin{theorem}
+    (Recomposability) TODO
+\end{theorem}
+
+% % TODO: make sure we're not claiming to invent the term "unicycular graph"
+% % TODO: maybe add a forward reference at end here to Implementation section
+% % TODO: make ID of y lower than the unicycle root, to show what we're minimizing
+
+% A grove is a 3-tuple of the following form:
+%     \figureDecompositionDefSets
+% where
+% \begin{align*}
+%   \figureDecompositionDefDecompComponents
+% \end{align*}
+% For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
+% given in \autoref{fig:Decomposition example grove}.
+
+% Let us now examine each of the three components in more detail.
+% %
+% The first component $\Set[NP]$ is the set of expressions rooted at a vertex with no parents, including the main root vertex.
+% % TODO: change vRoot to dot w/ subscript
+% In the example, $e_r$ is the expression corresponding to the graph root, $\vRoot$ / $._{\id{0}}$.
+% % TODO: change z_... to macro use w/ same appearance as in the graph
+% Expression, $z_{\id{52}}$ has been deleted and has no parents, so it also appears in $\Set[NP]$.
+
+% The first component $e_r$ is the expression derived from the root vertex $v_0$.
+% % \autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
+% There is one rule for each term constructor.
+% For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr({+}_{\id{38}}, \PlusLeft) \text{+}^ {\id{38}} \expr({+}_{\id{38}}, \PlusRight)$.
+% In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
+% An empty hole $\hole$ appears if there are no children at the given position.
+% A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
+% If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
+% For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.
+
+% The second component, $\Set[MP]$, is the set of expressions derived from a vertex with multiple parents. Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
+% In the example, $x_{\id{42}}$ is referenced twice in $e_r$.
+
+% The third component $\Set[U]$ is the set of expressions derived from the remaining vertices, which necessarily have exactly one parent.
+% These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
+% Every unicycle consists of a single cycle, and each vertex may have additional children that are not part of the cycle. Breaking the cycle turns a unicyclic graph into a tree.
+% % TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
+% % TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
+% To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
+% Formally, we choose the vertex $v$ such that $v = \min{\ancestors{v}}$ because
+% only a vertex on the cycle can be its own ancestor.
+% An edge to this unicycle root decomposes into a unicycle reference $\cycleVertex{u}$.
+% In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
+
+% Finally, we discover the unicycles of $\Set[U]$ by traversing any remaining edges, which necessarily constitute the elements of unicycular graph or another.
+Unicycular graph traversal occurs in two directions.
+We begin by choosing a a vertex as an arbitrary starting point, traverse its parents, and marking all of its vertices as seen.
+Since a vertex that has been seen twice must be on the cycle, we take the first such vertex as the root of the unicycle.
+We then traverse its children to produce a tree that covers the entire unicycular graph.
+Any remaining single-parented vertices must be parts of other unicycular graphs and are subsequently \emph{unicycle}-traversed until there are no more single-parented vertices left.
+
+% \begin{lemma}
+%     Unicycular graph traversal of a well-sorted graph produces a well-sorted tree
+% \end{lemma}
+
+% We say a term is \emph{well-sorted} when it recomposes into a well-sorted graph. The following theorem makes this intuition precise.
+
+% \begin{theorem}
+%     If $G$ is a well-sorted graph which decomposes into a grove $\Grove = (\Set[NP], \Set[MP], \Set[U])$, then any term $t \in \Set[NP] \cup \Set[MP] \cup \Set[U]$ recomposes into a well-sorted graph, and the composite of all such recompositions $\bigcup_{t\in \Set[NP] \cup \Set[MP] \cup \Set[U]}$ $\recomp{t}$ (viewed as sets of mappings from edges to edge states) is a graph $G'$ isomorphic to $G$.
+% \end{theorem}
+
+% \begin{proof}
+%     By rule induction on the derivations of $\Grove$ and $G'$
+% \end{proof}
+
+% That is, a well-sorted graph decomposes into a collection of terms that, when recomposed, produce the original graph.
+We give detailed information on decomposition and recomposition in \autoref{sec:Grove Formalism}.
+
+% % XXX
+
+% % decomposition produces sets of trees. As long as everything is well sorted, we
+% % can use them to produce expressions in the main grammar. (alt: define wrt the
+% % graph instead, once we establish the graph can be turned into expressions)
+
+% % *** well-sortedness judgment on trees: the tree only has valid positions for the
+% % constructors, and the child at each position has the sort corresponding to its
+% % position, and edges in position maps all start at the given constructor.
+
+% % Theorem: well-sorted graphs produce decompositions of well-sorted trees
+
+% % ``A graph has this property iff these conditions hold''
+
+% % then we can define a total mapping from well-sorted trees to expressions
+% % (because the tree structure is very generic, so this ties them back to the
+% % earlier grammar.)
+
+% % forest = 4-tuple
+
+% % forest validity: each of the trees in the forest are well sorted, and all Refs
+% % point to roots of other trees of the correct sort.
+
+% % every valid forest corresponds to a well behaved grove (i.e., ``expression
+% % forest'')
+
+% % soundness: if you start with a valid graph, you get a well behaved grove that
+% % connects every vertex (and only those vertices) in the graph. (i.e., if you turn
+% % it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
+% % Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
+% % add Root sort
+
+% % OR:
+
+% % use edge IDs instead of edges and build a mapping from vertices to
+% % sorts---change the $\e$s to $u$s)
+
+% % % TODO: maybe add a forward reference here to Implementation section
+
+\figureUserActionSyntax
+
+\subsection{User Actions}%
+\label{sub:User Actions}
+
+Given the grammar for user actions in \autoref{fig:User Action Syntax}, we want to show that user actions on terms correspond to equivalent graph patches.
+This can be done in either of two directions: from graph to grove or vice versa.
+Showing that operating on a grove is equivalent to operating on the underlying graph enables a more efficient implementation than the other way around, as then the editor is free to leverage term annotations to avoid a full decomposition pass after each user action, but ensuring the soundness of the resulting graph requires proving an additional theorem.
+For simplicity, we chose to translate all user edits directly to graph patches and then perform a decomposition pass over the entire graph after each user action.
+
+Here, we give intuitive explanations for the user actions of \autoref{fig:User Action Syntax}. For a more formal treatment, see \autoref{sec:Grove Formalism}.
+
+$\Construct{k}$ constructs a new edge whose destination is determined by the cursor's location.
+If the cursor is on an empty hole, the hole is filled by constructing a fresh vertex with constructor $k$.
+If the cursor is on an existing term and the constructor $k$ has a child position of the same sort, we \emph{wrap} the existing term by deleting it from its parent, constructing the new term in its place, then re-connecting the original term as a child of the new term.
+Otherwise, the editor may choose whether to apply patches that would produce conflicts.
+
+Depending on the cursor's location, $\Delete$ deletes one or more edges.
+If the cursor is on a constructable term of reference, its corresponding edge is deleted.
+If the cursor is on a conflict hole, all the edges of the conflict are deleted.
+Otherwise, the cursor is on an empty hole in which case the the action is a no-op.
+
+$\Reposition{v}{p}$ combines edge construction and deletion into a single \emph{atomic} operation.
+If the cursor is not an empty hole, the corresponding edge(s), as indicated above, are simultaneously deleted and then reconstructed as child(ren) with edge source $(v,p)$.
+
+% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
+% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition
+
+% move = delete + add
+
+% other composite edits
+
+% remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs
+
+% given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
+% This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
+% (we probably want to do it this way in Hazel)
+
+% OR
+
+% We could go the indirect route and avoid having to prove the theorem.
+% (this way is easier to start with, and it's how GRV works)
+
+% Mapping to b actions
+
+% \[
+%   \begin{array}{l}
+%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
+%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
+%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
+%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
+%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
+%     \CC(G) = \ldots                                                                                                                                    \\
+%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
+%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
+%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
+%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
+%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
+%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
+%   \end{array}
+% \]

### 2024-07-11 grove-by-example.tex
Updates from Overleaf

@@
 
-This section introduces collaborative structure editing in Grove.
-We begin in \autoref{sub:Program Representation} with how we use graphs to represent programs with holes.
-\autoref{sub:Single-User Actions} explores actions performed by a single user, Alice,
-and \autoref{sub:Multi-user Interactions} explores interactions between two users, Alice and Bob,
-as they edit their own instances of a shared program.
+This section introduces collaborative structure editing in Grove by example.
+\autoref{sub:Program Representation} describes how we use graphs to represent collaborative program sketches. 
+\autoref{sub:Single-User Actions} then shows examples of edits being performed by a single user, Alice. 
+\autoref{sub:Multi-user Interactions}-\ref{sub:Merge Conflicts} then describes a collaboration between two users, Alice and Bob,
+as they edit their own branches of a program and periodically merge in each other's edits, starting with examples without conflicts, then considering the various kinds of conflicts that might arise.
 
-\subsection{Program Representation}%
+For simplicity and concision, all of the examples in this section will be for a 
+language of standard arithmetic operations, 
+but our formalism in \autoref{sec:Formalism} and our implementation in \autoref{sec:Implementation} are parameterized by an arbitrary abstract syntax.
+
+\figureSimple
+
+Grove can form the basis for both a conventional version control workflow,
+where edits are batched into commits, or real-time collaborative editing, 
+where edits are communicated as they occur.
+This paper makes no assumptions about which batching mode is in use, nor do we consider the well-studied problem of reliably and efficiently communicating patches over a network\todo{cite something, dist sys textbook?}.\todo{maybe include a brief note in the discussion about cursors being represented by locations?}
+%\autoref{sub:Cursors} discusses representing cursor locations when in real-time mode.
+
+\subsection{Representing Collaborative Program Sketches as Graphs}%
 \label{sub:Program Representation}
 
-\figureSimple{}
+The \textit{edit state} of a Grove branch is a directed multi-graph representing a \emph{collaborative program sketch}, meaning an incomplete program, i.e.
+one that may have \emph{holes} and (as we will return to) conflicts.
+For example, \autoref{fig:Simple:a} gives one such graph and its corresponding \emph{decomposition} into, in this case, a single syntax tree,~\texttt{x * \hole},
+whose missing right operand is a hole, denoted $\hole$.
 
-The \textit{edit state} of a Grove instance is a graph representing a program
-in a language with concrete syntax for holes,
-or terms that have not been written yet or have already been deleted.
-For example, \autoref{fig:Simple:a} gives one such graph and its corresponding program,~\texttt{x * \hole},
-whose missing right operand is denoted by $\hole$.
+Each vertex represents a term in the specified language, except for a distinguished root vertex, 
+and is labeled with a unique identifier (UID) and a \emph{constructor}. 
+In \autoref{fig:Simple:a}, the root vertex has UID 0 and constructor~\textbullet. 
+Vertices \vSimpleTimes{} and \vSimpleX{} have UIDs 2 and 4 and constructors~\texttt{*} and \texttt{var(}$x$\texttt{)}, respectively.
+For clarity, we abbreviate \texttt{var(}$x$\texttt{)} as simply \texttt{x}; here, $x$ is a constructor parameter.
+We treat identifiers and literals as indivisible, but we discuss character-level editing in Section~TODO:REF\todo{ref}.
 
-The edit state is a directed graph with a distinguished root vertex.
-Each vertex represents a term in a hypothetical programming language
-and is labeled with a globally unique identifier and a term constructor.     % TODO: should we informally define `constructor' here?
-In \autoref{fig:Simple:a}, the root vertex has identifier 0 and constructor~\textbullet.
-Terms \vSimpleTimes{} and \vSimpleX{} have constructors~\texttt{*} and \texttt{var(x)}, respectively.
-For clarity, we abbreviate \texttt{var(x)} as simply \texttt{x}.
+An edge indicates that the destination vertex is a child of the origin vertex. 
+Each edge is labeled with a UID~(e.g.,~1 and~3 in \autoref{fig:Simple:a})
+and a \emph{position} (e.g., \texttt{Root} and \texttt{L} in \autoref{fig:Simple:a}). 
+The parent vertex's constructor determines a set of valid {positions}. 
+For instance, the~\texttt{*} constructor defines positions
+~\texttt{L}~(for the left operand)
+and~\texttt{R}~(for the right operand).
+The~\texttt{var} constructor is a leaf so it defines no positions.
+The root vertex constructor~\textbullet~has a single child position, \texttt{Root}.
 
-Each constructor has an associated set of \emph{child positions} for naming subterms.
-For instance, the~\texttt{*} constructor has positions
-for~\texttt{L}~(i.e., left child)
-and~\texttt{R}~(i.e., right child).
-The~\texttt{var} constructor has no child positions
-and takes a variable name as an argument, e.g.,~\texttt{x} in \texttt{var(x)}.\footnote{Note that,
-  for the purposes of this paper, we represent identifiers and number literals
-  indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}      % TODO: should we keep this footnote?
-The root vertex constructor~\textbullet~has a single child position \texttt{Root}.
+%Visually, we indicate the position of an edge by the location of its origin.
 
-Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:Simple:a})
-and a child position (e.g., \texttt{Root} and \texttt{L} in \autoref{fig:Simple:a}).
-An edge indicates that the destination vertex is a child of the origin vertex at the given position.
-%Visually, we indicate the position of an edge by the location of its origin.
-For clarity, we reserve odd numbers for vertex identifiers and even numbers for edge identifiers.
-
-Finally, holes are represented by the absence of a child.
+Holes arise in the decomposition by the absence of a child at a valid position.
 For example, in \autoref{fig:Simple:a} the absence of an \texttt{R} child under \vSimpleTimes{}
@@
 
-\figureWrapMove{}
+For clarity, we use odd numbers for vertex UIDs and even numbers for edge UIDs. In practice, UIDs would be generated by a mechanism
+that effectively ensures that collaborators always generate distinct UIDs, e.g. by generating
+universally unique IDs (UUIDs)~\cite{IETF RFC 4122}.
+
 % \figureMove{}
 
-\subsection{Single-User Actions}%
+\subsection{Edits}%
 \label{sub:Single-User Actions}
+\figureWrapMove
 
-In the remainder of this section, we consider user actions and interactions,
-starting in this subsection with a single user performing one edit at a time
-and continuing in \autoref{sub:Multi-user Interactions} with multiple users performing single edits concurrently.
+Individual users perform \emph{edits} to evolve the edit state.
+This paper abstracts over the user interface aspects of structure editors and makes no usability-related claims; edits could be performed through, for example, drag-and-drop interactions (as in block-based editors like Scratch) or keyboard interactions (as in MPS and Hazel).
+% User interface design for structure editors remains an active research area.
+% The common thread is that structure editors maintain a program sketch throughout the editing process.
 
-Each action by the user corresponds to one or more graph update operations.
-Graph updates can only add fresh edges or vertices or remove existing edges.
-All user actions are relative to an instance-local cursor and cursors are not replicated across instances.
-We discuss cursor representations in detail in \autoref{sub:Cursors}.
+Each edit translates directly to a \emph{graph patch}, which consists of a sequence of \emph{patch commands}.
+The Grove patch language requires only two patch commands:  \emph{edge insertion} and \emph{edge deletion}.
+A vertex is inserted when it is included in an edge insertion command.
 
-\subsubsection{Construction Actions}%
+To illustrate the Grove patch language, let us consider a sequence of standard edits, found across structure editors, performed by a single user, Alice.
+
+\subsubsection{Hole Filling}%
 \label{sub:Construction}
 
-To start our examples, say Alice's cursor is on the hole in~\texttt{x * \hole} in \autoref{fig:Simple:a}.
-Alice constructs the variable~\texttt{y}, which produces the edit state in \autoref{fig:Simple:b}.
-This fills the hole by creating a new vertex \vSimpleY{} containing the variable reference
-as a child of \vSimpleTimes{} at position \texttt{R} connected via edge \eSimpleY{}.
-The resulting graph represents the expression \texttt{x * y}.
+First, Alice fills the hole in the right position of \texttt{x * \hole} from \autoref{fig:Simple:a} with the variable~\texttt{y}. The resulting edit state is shown in \autoref{fig:Simple:b}.
+The patch corresponding to this hole filling action inserts an edge, labeled \eSimpleY{}, from the vertex corresponding to the parent term, \vSimpleTimes{}, to the newly constructed variable's vertex, \vSimpleY{}.
+The resulting graph decomposes to the term \texttt{x * y}.
 
-\subsubsection{Deletion Actions}%
+\subsubsection{Deletion}%
 \label{sub:Deletion}
@@
 
-Next, Alice moves the cursor to~\texttt{x} and deletes it, leavingx~\texttt{\hole{} * y}.
-This deletes edge~\eSimpleX{} as shown in \autoref{fig:Simple:c}.
-Notice that vertex \vSimpleX{} continues to exist,
-and if it had any children, those children would remain connected to it.
+Next, Alice moves the cursor to~\texttt{x} in \autoref{fig:Simple:b} and deletes it, 
+causing the deletion of edge~\eSimpleX{} and  resulting in the decomposition~\texttt{\hole{} * y} as shown in \autoref{fig:Simple:c}.
 
-Once a vertex is created, it is never deleted.
-This allows further interaction with those vertices by other users without risk of spontaneous cursor motion,
-as shown later in this section.
-In remaining diagrams, we omit such orphaned vertices if they are not relevant to the exposition.
+Once an edge with a particular identifier is deleted, it cannot be re-inserted.
+For instance, if Alice performed an ``undo'' on this deletion,
+a fresh edge between \vSimpleTimes{} and \vSimpleX{} would be created. (We can allow simple undo only if the patch has not yet been communicated to a collaborator).
 
-Once an edge with a particular identifier is deleted, it cannot be recreated.
-For instance, if Alice performed an ``undo'' on this deletion,
-a fresh edge between \vSimpleTimes{} and \vSimpleX{} would be created.
+Notice that vertex \vSimpleX{} continues to exist (and if it had any children, they would remain connected to it; see below for the implications in the collaborative setting).
+In the remaining figures, we omit such orphaned vertices if they are not relevant to the exposition.
 
-\subsubsection{Wrapping Actions}%
+% The same state would arise if Alice cut \li{x}: we assume each user has an individually managed clipboard, so until a corresponding {paste} edit is performed the vertex is deleted but it can be restored by edge insertion.
+
+\subsubsection{Wrapping}%
 \label{sub:Wrapping}
 
-Next, Alice moves the cursor to parent vertex \vSimpleTimes{}
-and constructs a binary addition expression with constructor~\texttt{+}.
-If the cursor were on a hole,
-a fresh term would be constructed in its place, as in \autoref{sub:Construction} above.
-However, since Alice's cursor is on an existing term \texttt{*},
-it is interpreted as a wrapping action instead of a construction action.
-This corresponds to the following sequence of graph updates, shown in \autoref{fig:Wrap}:
-(a) delete edge \eSimpleTimes{} leaving \vSimpleTimes{} temporarily orphaned,
-(b) add fresh vertex \vWrapPlus{} as a child of the root vertex, and
-(c) add fresh edge to \vSimpleTimes{} in left (\texttt{L}) child position of the new vertex.
-Our choice here of wrapping with a bias for the left child position is arbitrary.
+Next, Alice moves the cursor to the parent term \vSimpleTimes{} in \autoref{fig:Simple:c}
+with the intention of wrapping it in a binary addition expression with constructor~\texttt{+}.
 
-\subsubsection{Repositioning Actions}%
+Many structure editors define a primitive wrapping edit, choosing a position heuristically (e.g. favoring the left). Others 
+require the user to cut the original term, construct the new outer term, then paste
+the original term in the intended position.
+
+In either case, the corresponding sequence of patch commands would produce the edit states shown in \autoref{fig:Wrap}: the edge connecting the root to the original term is deleted (effectively cutting the original term), leaving \vSimpleTimes{} temporarily orphaned, then an edge to the new outer term is inserted, followed by an edge reconnecting the original term (effectively pasting the original term).
+
+\subsubsection{Relocation}%
 \label{sub:Repositioning}
 
-We have two options for moving a term from one place to another.
-The first is to delete the term from its original position and then reconstruct it in its new position.
-Since edges are uniquely identifiable,
-this option precludes the use of edge identities in any cursor representation,
-as doing so would lead directly to the repositioning problem.
+Alice changes her mind and decides to relocate the multiplication from the left to the right position of the addition.
+A structure editor might support this using drag-and-drop or cut-and-paste affordances.
+In either case, the resulting patch commands will proceed through the two states shown in \autoref{fig:Move}: deleting the original incoming edge and then inserting an edge at the new location.
+Notice that the sub-graph corresponding to the relocated term itself is never deleted nor re-inserted, in contrast to conventional line-based patch languages.
+See below for the implications in the collaborative setting.
 
-Consequently, we take the second option: explicit repositioning actions that preserve identities.
-These could be triggered by, for example, a drag and drop or a cut and paste---but not a copy and paste,
-or a second paste after a cut, which do not need to preserve identity
-and so could rely on reconstruction as already described in \autoref{sub:Construction}.
-
-In the graph, explicit repositioning involves simply deleting the edge from the original parent
-and then adding a new edge to it from the new parent.
-Thus the term is indeed \textit{repositioned}, not copied.
-
-For example, in \autoref{fig:Move} Alice continues by
-repositioning \texttt{\hole{} * y} from the left child of \texttt{+} to the right child.
-This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
-(a) delete \eWrapTimes{} leaving \vSimpleTimes{} temporarily orphaned and
-(b) add \eMoveTimes{} to \vSimpleTimes{} in the right child position of \vWrapPlus{}.
+\subsubsection{Copying} 
+\label{sub:Copy}
+A copy-and-paste, or a cut followed by multiple pastes, would of course involve copying the graph structure of the original term but generating fresh UIDs (not shown).
 
 % \figureDifferentParts{}
-\figureDifferentPartsNestedParts{}
 
 %\figureCommutativity{}
-\begin{figure}
+\subsection{Collaboration}%
+\label{sub:Multi-user Interactions}
+
+We now turn our attention to how Grove handles collaboration.
+The examples in this section generalize to collaborations between any number of users,
+but for simplicity we consider only two: Alice and Bob.
+Alice and Bob are each concurrently editing their own branches of the repository (or their own instance of a real-time collaborative editor), 
+performing edits that translate to patches as described above. 
+They periodically communicate these patches to one another. \autoref{fig:Commutativity} diagrams the Grove collaboration model.\todo{stretch: include a comparable diagram for }
+
+\begin{figure}[h]
   \centering
   \begin{tikzpicture}
-    \path (-3cm, 0cm) node (a) [align=center]            {Original \\ Version};
-    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Version};
-    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Version};
-    \path ( 3cm, 0cm) node (d) [align=center]            {Synchronized \\ Version};
+    \path (-3cm, 0cm) node (a) [align=center]            {Base \\ Branch};
+    \path ( 0cm, 1cm) node (b) [align=center,alice node] {Alice's  \\ Branch};
+    \path ( 0cm,-1cm) node (c) [align=center,bob node]   {Bob's    \\ Branch};
+    \path ( 3cm, 0cm) node (d) [align=center]            {Merged \\ Branch};
     \path [draw,->,alice step] (a) -- node [pos=0.7,align=center,auto]      {Alice's \\ Edits} (b);
     \path [draw,->,bob step]   (a) -- node [pos=0.7,align=center,auto,swap] {Bob's \\ Edits}   (c);
-    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Share \\ Edits} (d);
-    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Share \\ Edits} (d);
+    \path [draw,->,merge step] (b) -- node [pos=0.3,align=center,auto]      {Apply \\Bob's Patch} (d);
+    \path [draw,->,merge step] (c) -- node [pos=0.3,align=center,auto,swap] {Apply \\Alice's Patch} (d);
   \end{tikzpicture}
-  \caption{Commutativity of Edits.}
+  \caption{Collaboration in Grove is simple due to the commutativity of Grove's patch language.}
+  \Description{This figures describes the commutativity of edits}
   \label{fig:Commutativity}
@@
 
-\subsection{Multi-user Interactions}%
-\label{sub:Multi-user Interactions}
+\subsubsection{Commutativity}%
+\figureDifferentPartsNestedParts
 
-We now turn our attention to how Grove handles multiple users.
-The examples in this section generalize to any number of users,
-but for simplicity we consider only two: Alice and Bob.
-Alice and Bob each maintain their own editor instances independently
-and perform actions relative to its state using their own cursors, as we discuss in \autoref{sub:Cursors}.
+\label{sub:Commutativity:informal}
+The Grove patch language is commutative, meaning that there is no need for a 
+complex three-way merge algorithm (i.e. operational transform). Instead, each 
+user can simply apply incoming patches to their own edit state as they arrive, no matter the order in 
+which they arrive. If two users have received the same set of patches, their edit state will converge.
 
-Users collaborate by sharing graph updates corresponding to user edits,
-e.g., via direct network connections or relayed through a server.
-% TODO: CITE(how to share network data -- DHT? consensus? ).
-The granularity of communication may vary.
-For real-time collaboration, graph updates may be sent as they occur.
-For asynchronous version control, updates may be batched together in the form of commits.
-As we will show, since these edits are commutative,
-all instances converge to the same state once all updates are shared.
-We assume users have some way to create globally unique identifiers
-% for verticies and edges
-without coordination~(e.g.,~GUIDs).
-% TODO:CITE:GUIDS).
+The key properties that make the Grove patch language commutative is that 
+edge deletion is permanent and vertex insertion is permanent.
+We establish commutativity formally in \autoref{sec:Formalism}.
+For now, let us consider several example scenarios that demonstrate 
+how Grove handles different collaborative editing scenarios, particularly 
+the problematic situations outlined in \autoref{sec:Introduction}.
 
-\subsubsection{Editing Different Parts of the Code}%
+\subsubsection{Solving the Granularity Problem}%
 \label{sub:Editing Different Parts of the Code}
 
-Alice and Bob start where Alice left off in the previous section
-with the state \texttt{\hole{} + \hole{} * y} shown in \autoref{fig:Move:b}.
+Alice and Bob start where Alice left off in \autoref{fig:Move:b} 
+with the term \texttt{\hole{} + \hole{} * y}.
 Alice then adds~\vDifferentPartsAlice{} as the left child of \vSimpleTimes{}.
-At the same time, Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
-Before either shares their graph updates with the other,
-Alice and Bob have the graphs in \autoref{fig:DifferentParts:a} and \autoref{fig:DifferentParts:b}, respectively.
+Concurrently, Bob changes \vSimpleY{} to \vDifferentPartsBob{}.
+Before sharing their patches,
+Alice and Bob have the edit states \autoref{fig:DifferentParts:a} and \autoref{fig:DifferentParts:b}, respectively.
 Note that the transition from \autoref{fig:Move:b} to \autoref{fig:DifferentParts:b}
@@
 We thus mark the transition with a star.
-Once Alice and Bob share their graph edits and apply each other's updates to their own copy of the graph,
-they both have the graph in \autoref{fig:DifferentParts:c}.
+Once Alice and Bob share their patches and apply each other's patch to their own edit state,
+both edit states converge to the graph in \autoref{fig:DifferentParts:c}.\todo{add more spacing around terms in Fig 5} 
+Because Grove's patch language is structural rather than line-based, the fact that these edits happened to be close to one another (i.e.
+in the same arithmetic expression) does not run afoul of the \textbf{granularity problem} described in \autoref{sec:Introduction}.
 
@@
 
-\subsubsection{Commutativity}%
-\label{sub:Commutativity:informal}
-
-% \todo{rev}
-Note that aside from sharing edge deletions and additions,
-merging requires no coordination between Alice and Bob.
-For example, suppose Alice and Bob did not transmit their edits to each other
-but instead to a third user, Chris.
-Chris can apply the edits from Alice and Bob in any order
-and will always get the graph in \autoref{fig:DifferentParts:c}.
-
-In other words, our update model obeys the commutativity shown in \autoref{fig:Commutativity}.
-For any sequences of edits from Alice and Bob,
-applying them in any order always produces the same results.
-(We show this formally in \autoref{sub:Commutativity:formal}.)
-
-\subsubsection{Editing Nested Parts of the Code}%%
+\subsubsection{Solving the Relocation Modification Problem}%%
 \label{sub:Editing Nested Parts of the Code}
 
-After all of Alice's and Bob's updates have been shared,
-they both have the edit state \texttt{\hole{} + u * y} in \autoref{fig:DifferentParts:c}.
+After converging on \texttt{\hole{} + u * y} in \autoref{fig:DifferentParts:c}, 
 Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{},
-producing the graph in \autoref{fig:NestedParts:a}.
-Meanwhile, Bob repositions~\vSimpleTimes{} (along with its children)
-from the~\texttt{R} child of~\vWrapPlus{} to its~\texttt{L} child
-by deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}.
-The resulting graph is shown in \autoref{fig:NestedParts:b}.
-Because updates are located by edge identifiers instead of by paths relative to the root vertex,
-each of Alice's and Bob's updates can be applied to the other's graph
-and afterwards they will both have the graph in \autoref{fig:NestedParts:c}.
+producing the edit state in \autoref{fig:NestedParts:a}.
+Meanwhile, Bob relocates~\vSimpleTimes{} (bringing along its children)
+from the~\texttt{R} position of~\vWrapPlus{} to the~\texttt{L} position. As discussed above, this involves deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}.
+Bob's resulting edit state is shown in \autoref{fig:NestedParts:b}.\todo{}
 
-Note that commutativity has consequences in this scenario.
-Since all updates must eventually lead to the same edit state regardles of the order in which they are applied,
-if Alice, say, receives Bob's deletion of~\eMoveTimes{}, makes some local changes,
-and then receives Bob's addition of~\eNestedPartsBob{},
-the result must still be the graph in \autoref{fig:NestedParts:c}.
-In that case, Alice's edits would then temporarily be within the term deleted by Bob.
-Thus, our semantics must allows users to edit deleted terms.
-Whether or how this is presented in the user interface is another question.
+Although Alice has modified a term that Bob has concurrently relocated, the edits commute: Alice's modifications are relocated to the new location chosen by Bob.
+This addresses the \textbf{relocation modification problem} described in \autoref{sec:Introduction}.
+In a line-based setting, this kind of edit can lead to silent code duplication or spurious conflicts (the threat of which, in the author's experience, can inhibit development teams from performing useful code reorganizations).
 
-% TODO: \todo{TODO}She can restore that code using a restoration action add a fresh edge to it.
+% TODO: \todo{TODO}She can restore that code using a restoration action and add a fresh edge to it.
 
-\subsection{Merge Conflicts}%
+\subsubsection{Warning of Edits under Disconnected Terms}
+
+If instead of relocating the multiplication in \autoref{fig:NestedParts}, Bob had deleted it (i.e.
+disconnected it from the root), then Alice's edits would still commute, but her edits would be applied under a deleted term.
+
+This situation could also arise in a real-time collaborative editor, where each individual edit might arrive at any time (rather than in atomic commits).
+If Alice, say, receives Bob's deletion of~\eMoveTimes{}, then makes her edits
+before receiving Bob's subsequent insertion of~\eNestedPartsBob{} to complete the relocation, 
+Alice's edits would then temporarily be under a disconnected term.
+
+This does not present a formal problem or conflict. A subsequent edit might reconnect a disconnected term, so it is sensible for edits to these terms to be recorded. 
+However, heuristically, a system might warn users, perhaps after a period of quiescence in a real-time setting, that Alice's edits were effectively deleted and provide affordances for interacting with disconnected terms.
+
+\subsection{Conflicts}%
 \label{sub:Merge Conflicts}
@@
 
-\subsubsection{Multi-child conflicts}%
+\subsubsection{Local Conflicts}%
 \label{sub:Multi-child conflicts}
 
-\figureMultiChild{}
+\figureMultiChild
 
@@
 
-\subsubsection{Multi-parent conflicts}%
+\subsubsection{Multi-Location Conflicts}%
 \label{sub:Multi-parent conflicts}
 
-\figureMultiParent{}
+\figureMultiParent
 
 Multi-parent conflicts occur when multiple edges share a destination vertex,
-as opposed to multi-child conflicts which share a common origin.
+as opposed to multi-child conflicts, which share a common origin.
 An example of this is in \autoref{fig:MultiParent},
@@
 Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}.
-At the same times, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
+At the same time, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
 As we explain in \autoref{sec:Formalism}, deletions are idempotent,
@@
 
-\figureCycle{}
+\figureCycle
 
@@
 
-\figureDisconnect{}
+\figureDisconnect
 
-TODO: technically a cycle
+% TODO: technically a cycle
 
-TODO: in general detect when there are or were edits to something now deleted
+% TODO: in general detect when there are or were edits to something now deleted
 
-\todo{UI would show orphans only once edits orphans occur}
+% \todo{UI would show orphans only once edits orphans occur}
 
@@
 
-\subsection{Cursors}%
-\label{sub:Cursors}
+% \subsection{Cursors}%
+% \label{sub:Cursors}
 
-\paragraph{Representation}
+% \paragraph{Representation}
 
-For most editing a cursor is represented by a position within a vertex.
-For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
-Representing it this way means that if Bob or some other user deletes TODO:vertex
-and replaces it with a different vertex, Bob's cursor is still valid.
-We specify a position within a vertex instead of a vertex because TODO.
+% For most editing a cursor is represented by a position within a vertex.
+% For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
+% Representing it this way means that if Bob or some other user deletes TODO:vertex
+% and replaces it with a different vertex, Bob's cursor is still valid.
+% We specify a position within a vertex instead of a vertex because TODO.
 
-However, there is one case where we need a more precise cursor.
-That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
-We want users to beable to put their cussor on either the conflict as a whole
-(e.g., TODO: \texttt{\{ x | y \}})
-or on an individual element (e.g., \texttt{x} or \texttt{y}).
-In the former case, a pair of a vertex and a child position suffices.
-For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
+% However, there is one case where we need a more precise cursor.
+% That is when a vertex has a multi-child conflict (e.g. TODO in FIG:TODO).
+% We want users to be able to put their cursor on either the conflict as a whole
+% (e.g., TODO: \texttt{\{ x | y \}})
+% or on an individual element (e.g., \texttt{x} or \texttt{y}).
+% In the former case, a pair of a vertex and a child position suffices.
+% For the latter case, we represent cursors by an edge identifier (e.g., TODO in TODO:FIG).
 
-\paragraph{Communication}
+% \paragraph{Communication}
 
-In our system all edits use edge and vertex identifiers, and user
-cursors do not affect the interpretation of graph edits.
-Thus, cursors need not be communicated to other users.
-However, in a collaborative setting, seeing the cursors of other users
-can be useful.
-For this purpose, editors can announce their cursor position to other editors.
-At any point in time, the announcement
-with the most recent timestamp
-for a particular user
-is used.
+% In our system all edits use edge and vertex identifiers, and user
+% cursors do not affect the interpretation of graph edits.
+% Thus, cursors need not be communicated to other users.
+% However, in a collaborative setting, seeing the cursors of other users
+% can be useful.
+% For this purpose, editors can announce their cursor position to other editors.
+% At any point in time, the announcement
+% with the most recent timestamp
+% for a particular user
+% is used.
 
-Note that this cursor position may refer to edges or verticies that
-do not yet exist on the receivers machine.
-In these cases, we can either not display the other user's cursor (and perhaps
-have a visual display flagging this fact), or we can display the most recent
-cursor that represents a valid position in the local graph (perhaps
-shown in a fadded color to show that this cursor is known to not be up to date).
-(Or perhaps, all cursors always fade/decay over time like on radar blips.)
+% Note that this cursor position may refer to edges or verticies that
+% do not yet exist on the receivers machine.
+% In these cases, we can either not display the other user's cursor (and perhaps
+% have a visual display flagging this fact), or we can display the most recent
+% cursor that represents a valid position in the local graph (perhaps
+% shown in a fadded color to show that this cursor is known to not be up to date).
+% (Or perhaps, all cursors always fade/decay over time like on radar blips.)

### 2024-07-11 grove-paper.tex
make file structure match paper structure
move term syntax and arity figures into new files and deduplicate
add decomp back to main formalism section (deduped)
simplify section file names and start separate appendix doc
more intro - commented out other sections to speed up compilation
redo most of section 2
feat: latest changes - uncomment sections
Updates from Overleaf

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+% !TEX program = lualatex
 
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
+\documentclass[acmsmall,dvipsnames,10pt,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true}
+
 \acmJournal{PACMPL}
 \acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{POPL}
 \acmArticle{1}
-\acmYear{2021}
+\acmYear{2025}
+
 \acmMonth{1}
@@
 %% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
 \setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
 
@@
 
-\input{config-tex}
-\input{config-tikz}
-\input{config-glyphs}
-\input{config-definitions}
-\input{config-graphs}
+%%% Spacing hacks
+\usepackage{titlesec}
+\titlespacing*{\section} {0pt}{6pt}{2pt}
+% \titlespacing*{\subsection} {0pt}{6pt}{2pt}
+\titlespacing*{\subsubsection}{0pt}{6pt}{10pt}
+\titlespacing*{\paragraph} {0pt}{4pt}{10pt}
+% \titlespacing*{\subparagraph} {12pt}{5pt}{10pt}
 
-\input{figure-simple}
-\input{figure-wrap-move}
-\input{figure-different-parts-nested-parts}
-\input{figure-multi-child}
-\input{figure-multi-parent}
-\input{figure-cycle}
-\input{figure-disconnect}
-\input{figure-decomposition}
+\usepackage{enumitem}
+\usepackage{listings}%
+\lstloadlanguages{ML}
+\lstset{tabsize=2, 
+basicstyle=\small\ttfamily, 
+% keywordstyle=\sffamily,
+commentstyle=\itshape\ttfamily\color{gray}, 
+stringstyle=\ttfamily\color{purple},
+mathescape=false,escapechar=\#,
+numbers=none, numberstyle=\scriptsize\color{gray}\ttfamily, language=ML, showspaces=false,showstringspaces=false,xleftmargin=15pt, 
+morekeywords={string, float, int},
+classoffset=0,belowskip=\smallskipamount, aboveskip=\smallskipamount
+}
+\let\li\lstinline
+\usepackage{todonotes}
+
+% TODO: consolidate all the macros into here
+\input{config/macros}
+
+\input{config/config-tex}
+\input{config/config-tikz}
+\input{config/config-glyphs}
+\input{config/config-definitions}
+\input{config/config-graphs}
+
+% figures
+\input{figures/figure-simple}
+\input{figures/figure-wrap-move}
+\input{figures/figure-different-parts-nested-parts}
+\input{figures/figure-multi-child}
+\input{figures/figure-multi-parent}
+\input{figures/figure-cycle}
+\input{figures/figure-disconnect}
+\input{figures/figure-decomposition}
+\input{figures/figure-term-syntax}
+\input{figures/figure-arity}
+\input{figures/figure-decomposition-def}
+\input{figures/figure-user-actions}
+
+% marking
+\input{symbols}
 
@@
 %% Title information
-% \title[Grove]{Convergent Collaborative Structure Editing}
-\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
+\title[Grove]{Grove: A Bidirectionally Typed Collaborative Structure Editor Calculus}
 %\subtitle{Subtitle}
@@
 \orcid{0000-0003-1693-6172}
+
+\author{Sundara Vishnu Satish}
+\orcid{TODO}
+
+\author{Thomas J. Porter}
+\orcid{TODO}
 
@@
 
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  Text of abstract \ldots.
-\end{abstract}
+% Abstract
+\input{abstract}
 
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10011007.10011006.10011008</concept_id>
-%<concept_desc>Software and its engineering~General programming languages</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%<concept>
-%<concept_id>10003456.10003457.10003521.10003525</concept_id>
-%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%<concept_significance>300</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
+% 2012 ACM Computing Classification System (CSS) concepts
+% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+% \begin{CCSXML}
+% <ccs2012>
+% <concept>
+% <concept_id>10011007.10011006.10011008</concept_id>
+% <concept_desc>Software and its engineering~General programming languages</concept_desc>
+% <concept_significance>500</concept_significance>
+% </concept>
+% <concept>
+% <concept_id>10003456.10003457.10003521.10003525</concept_id>
+% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
+% <concept_significance>300</concept_significance>
+% </concept>
+% </ccs2012>
+% \end{CCSXML}
+% \ccsdesc[500]{Software and its engineering~General programming languages}
+% \ccsdesc[300]{Social and professional topics~History of programming languages}
+% End of generated code
 
 %% Keywords
-%% comma separated list
 % \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
 
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
 \maketitle
 
-\input{grove-paper-introduction}
-\input{grove-paper-collaborative-structure-editing}
-\input{grove-paper-formalism}
-\input{grove-paper-implementation}
-\input{grove-paper-related-work}
-\input{grove-paper-discussion-and-conclusion}
+\input{introduction}
+\input{grove-by-example}
+\input{formalism}
+\input{marking-system}
+\input{implementation}
+\input{related-work}
+\input{discussion-and-conclusion}
 
@@
 
-\input{appendix-graph-sequences}
-\input{appendix-formalism}
+% \input{appendix-A-graph-sequences}
+\input{appendix/appendix-A-formalism}  
+% \input{appendix/appendix-B-marking-system}
+\input{appendix/appendix-B-typechecking}
+% \input{appendix/appendix-C-type-inference}
 

### 2024-07-11 implementation.tex
Updates from Overleaf

@@
 
-We implemented the core Grove calculus of \autoref{sec:Formalism} as an OCaml library.
+We implemented the core Grove calculus of \autoref{sec:Grove Formalism} as an OCaml library.
 Our code is parameterized by a syntax specification for expressions that we generate automatically, which makes the implemented language of expressions easier to modify or replace.
@@
 
-Since we cannot implement an infinite mapping directly, the graph is a data structure, \texttt{Graph.t}, that maps from edges to the edge states $\left\{\Plus, \Minus\right\}$.
-We do not represent edges that map to $\bot$.
+\subsection{Graph Implementation}
+\label{sec: Graph Implementation}
+The graph data structure \texttt{Graph.t} is implemented as \texttt{OCaml Map} data structure with insertion and selection operations whose asymptotic worst-case complexity is logarithmic with respect to the size of the map.
+Since we cannot implement an infinite mapping directly, the graph only maps
+live or deleted edges to edge states $\left\{\Plus, \Minus\right\}$. We do not represent edges that map to $\bot$.
 
-% TODO: rename Decomposition.ml to Grove.ml
-
-There's a function \verb!decompose! of type \verb!Graph.t) : t * Edge.Set.t Vertex.Map.t! that
-
-From this graph, we compute the set of all edges and the set of live edges with a linear scan.
-
-% TODO: make this into a listing
-
-\begin{verbatim}
-  Graph.edges : Graph.t -> Edge.Set.t
-  Graph.live_edges : Graph.t -> Edge.Set.t
-\end{verbatim}
-
-% TODO: keep connecting the code to the formalism like this
-
-These vertices are partitioned into three sets: multi-parented, single-parented, or orphaned.
-A vertex $v$ is considered a parent of another vertex $v'$ if there is a live edge from $v$ to $v'$.
-
-For each vertex, we take the parent and child edge sets.
-
-% TODO: fix the formalism for parents (to work with edges instead of vertices) and then check that it matches the implementation. Try renaming parents to vparents and adding parents for edges.
-
-% TODO: Talk about how the graph is implemented (not as an actual function, but as a map from edges to edge states) so that we can denote the set of created + deleted edges clearly, as well as all of the vertices that have ever been created.
-
-For a graph $G : \E \to \Sigma$ where $\E = \U \times \V \times \P \times \V$, our graph decomposition algorithm runs in $O(\abs{\V} \log \abs{\V} + \abs{\E} \log \abs{\V})$.
+For a graph $G : (\E = \U \times \V \times \P \times \V) \to \Sigma$, our graph decomposition algorithm runs in $O(\abs{\V} \log \abs{\V} + \abs{\E} \log \abs{\V})$.
 It begins with a scan of all edges that have been created or deleted $O(\abs{\E})$.
@@
 For unicycles, we traverse backwards until a vertex is seen twice $O(\abs{\V})$, then proceed forward to find the least vertex on the cycle $O(\abs{\V})$.
+Once the least vertex is found, decomposition of unicycle begins with it, thus ensuring any edges to it become references to a unicycle root.
+
+\subsection{Grove Implementation}
+\label{sec: Grove Implementation}
+
+% TODO: Talk about Uuid Generation
+First, we set up \texttt{Uuid.ml}, a module to generate well-founded monotonically increasing unique IDs for all the parts of our code that require unique identifiers.This module has features that allow it to both generate and check for the well-foundedness of a given ID..
+Our programmatically generated target language syntax, explained further in \autoref{sec: Parameterized Code Generation}, provides a \texttt{Lang.Constructor.t} corresponding to $k \in \K$ from \autoref{sec:Grove Formalism}.Further, we have a \texttt{Vertex.ml} module that \emph{wraps} a given \texttt{Lang.Constructor.t} with a UUID, such that, $\V = \K \times \U$.
+This also allows for the vertices to be ordered by their UUIDs.
+Internally, we represent edge-sources ($s = (v,p)$) as a record containing \texttt{Vertex.t}\textit{s} and \texttt{Lang.Position.t}\textit{s}, where the latter encodes the child positions and arities of the various constructors, in \texttt{Cursor.ml}.
+That is, a cursor (or a source) serves as a point of focus for graph actions and it references a position relative to some vertex.
+
+Then we have edges corresponding to $\E = \U \times \S \times \V$, implemented in \texttt{Egde.ml} UUID wrapped \emph{source} from a source to a target.
+Edge states ($\Sigma$), are expressed as variant type with two variants \texttt{Created} and \texttt{Deleted} representing $\Plus$ and $\Minus$ respectively.Since we do not directly implement infinite mappings in \texttt{Graph}, $\bot$ does not have a corresponding variant.
+Further, trees are defined in \texttt{Tree.ml} to represent disjointed graph components.
+A tree is a vertex with a (possibly empty) set of children or is a reference to one.
+The children are a list of \emph{child}s, where a child is a tree paired with an edge UUID.
+We also have a helper \texttt{child\_map : (Lang.Position.t * (Uuid.Id.t $*$ t) list) list $->$ children Position\_map.t} that takes in pairs of target language positions and lists of pairs Uuid and trees to return a map of the children.
+Here, a position\_map is a structure that \emph{maps} \texttt{Lang.position.t}s to their appropriate child representations.
+
+Finally, the logic and infrastructure for a grove that a given graph decomposes into using the decomposition algorithm is implemented in \texttt{Grove.ml}.
+
+% TODO: Tree.ml, Position_map.mk, Grove.ml, Graph_action.ml
+
+\subsection{Parameterized Code Generation}
+\label{sec: Parameterized Code Generation}
+
+% How do you give it the corresponding language definition to generate child positions, arities
+
+% % TODO: rename Decomposition.ml to Grove.ml
+
+% There's a function \verb!decompose! of type \verb!Graph.t) : t * Edge.Set.t Vertex.Map.t! that
+
+% From this graph, we compute the set of all edges and the set of live edges with a linear scan.
+
+% % TODO: make this into a listing
+
+% \begin{verbatim}
+%   Graph.edges : Graph.t -> Edge.Set.t
+%   Graph.live_edges : Graph.t -> Edge.Set.t
+% \end{verbatim}
+
+% % TODO: keep connecting the code to the formalism like this
+
+% These vertices are partitioned into three sets: multi-parented, single-parented, or orphaned.
+% A vertex $v$ is considered a parent of another vertex $v'$ if there is a live edge from $v$ to $v'$.
+
+% For each vertex, we take the parent and child edge sets.
+
+% % TODO: fix the formalism for parents (to work with edges instead of vertices) and then check that it matches the implementation. Try renaming parents to vparents and adding parents for edges.
+
+% % TODO: Talk about how the graph is implemented (not as an actual function, but as a map from edges to edge states) so that we can denote the set of created + deleted edges clearly, as well as all of the vertices that have ever been created.
 
@@
 
-Optimizations: least fixed point
+% Optimizations: least fixed point
 
-We determine $e$ by starting at the root of the graph and including every connected
-vertex that has a single parent. In positions where there are no children, we
-leave an empty hole $\hole$. In positions where there are multiple children, we
-leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
-encounter a vertex with multiple parents, we add its corresponding expression to
-the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
-the unique identifier of the referenced vertex
+% We determine $e$ by starting at the root of the graph and including every connected
+% vertex that has a single parent. In positions where there are no children, we
+% leave an empty hole $\hole$. In positions where there are multiple children, we
+% leave a conflict hole of the form $\conflictHole{e_1,\ldots,e_n}$. Whenever we
+% encounter a vertex with multiple parents, we add its corresponding expression to
+% the second component $MP$ and leave an indirect reference $\multiVertex{u}$, where $u$ is
+% the unique identifier of the referenced vertex
 
@@
 
-algorithmic section: if we want to implement these conversions, here are the
-steps (without introducing new concepts):
+% algorithmic section: if we want to implement these conversions, here are the
+% steps (without introducing new concepts):
 
-- linear scan
-- breaking out roots
-- ...
+% - linear scan
+% - breaking out roots
+% - ...
 
-Graph decomposition occurs in three steps.
+% Graph decomposition occurs in three steps.
 
@@
   \caption{Syntax of graph decomposition trees as a grammar.}
+  \Description{This figure describes the graph decomposition grammar}
   \label{fig:Syntax of graph decomposition trees}
@@
 
-lemma: a unicycle is a cycle with trees hanging off of it
+% lemma: a unicycle is a cycle with trees hanging off of it
 
-lemma: unicycle traversal always produces a valid tree
+% lemma: unicycle traversal always produces a valid tree
 
-Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
-starting point and its parents are traversed and marked as seen. Any vertex that
-has been seen twice must be on the cycle. We take the first one as the root of
-the unicycle and traverse its children to produce a tree that covers the unicycular graph.
-Any remaining single-parented vertices must be parts of other unicycular graphs and are
-unicycle-traversed until there are no more single-parented vertices left.
+% Unicycular graph traversal occurs in two directions. A vertex is chosen as an arbitrary
+% starting point and its parents are traversed and marked as seen. Any vertex that
+% has been seen twice must be on the cycle. We take the first one as the root of
+% the unicycle and traverse its children to produce a tree that covers the unicycular graph.
+% Any remaining single-parented vertices must be parts of other unicycular graphs and are
+% unicycle-traversed until there are no more single-parented vertices left.

### 2024-07-11 introduction.tex
Updates from Overleaf

--- 2022-08-18 introduction.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Software development is a highly collaborative activity. 
+Development teams typically collaborate with the aid of a version control system (VCS) like Git~\cite{git}, Subversion~\cite{subversion}, or Darcs~\cite{darcs}. 
+These systems maintain a branching history of \emph{commits} to a source code repository, each consisting of a \emph{patch} together with various metadata, e.g.
+a human-readable commit message.
+Patches are imperative programs written in a
+\emph{patch language} defining a set of primitive editing commands.
+The standard POSIX \li{patch} language, for example, specifies commands for inserting and deleting specified lines of text at specified line numbers within a file.
 
-Programming is an increasingly collaborative activity.
-Today, collaborative editing services are either provided directly by an asynchronous,
-diff-and-merge-based version control system like git, darcs, or pijul,
-or they are integrated into a real-time editor built on top of a distributed consistency protocol such as OT or CRDT.
-Whereas algorithms for collaborative text editing have been well developed,
-we know far less about collaborative editing of more structured documents---specifically, of program source files---%
-hence we are motivated to ask the following questions:
-\begin{itemize}
-\item Is there a more natural model for collaborative editing of tree structures?
-\item How do we preserve user intent in the presence of concurrent, non-linear edits?
-\end{itemize}
+Developers do not typically express program edits using the patch language directly, nor in any case do version control systems typically have access to a log of the developer's edits.
+Instead, version control systems must \emph{synthesize} patches from the file system state using heuristic algorithms, such as the classic \li{diff} algorithm that synthesizes a patch that minimizes the edit distance between two file system states.\todo{cite}
 
-Diff-and-merge is clearly not an ideal solution.
-Since diffs do not provide information on the original scope or ordering of edits,
-the merging process must reconstruct an approximation, e.g., as a sequence of line additions and deletions.
-In a tree-native setting,
-line-based merging is not only less likely to succeed than one which retains finer grained information,
-it is also less able to provide useful information about conflicting edit states as they arise.
-Moreover, because diffs cannot guarantee preservation of user intent,
-critical features such as identity-preserving repositioning are impossible to implement on top of them.
+When two patches, developed concurrently in branches based on a common ancestral commit, must be merged, version control systems deploy a \emph{three-way merge algorithm}.
+The standard approach is to apply the \emph{local patch} first, then modify the \emph{remote patch} by shifting its line numbers to account for the local patch's line insertions and deletions.
+This algorithm is an \emph{operational transform}~\cite{ot}.
+Character-level operational transforms are similarly used for real-time collaborative editing, e.g. in tools like Google Docs and in Visual Studio Code's Live Share feature~\cite{crdts-for-text}.
 
-Operational Transformation (OT), on the other hand, goes perhaps too far in the opposite direction.
-It is a distributed system architecture that coordinates the proliferation and maintainance of edits and edit states.
-Industrial-scale collaborative editors, such as Google Docs,
-employ OT for real-time applications requiring strong eventual consistency of causally ordered events.
-It does so by maintaining an event log it transforms upon each new entry in order to preserve the causal ordering.
+\paragraph{Problem 1: Merging Sensibly} 
+When merging patches, conflicts (e.g. due to different modifications to the same location) are unavoidable. However, 
+standard three-way merge algorithms for line-based patch languages commonly identify spurious conflicts, fail to identify legitimate conflicts, or silently duplicate or misplace code.
+Let us briefly review some classic problems with these systems.
+Appendix~\ref{appendix:merge-problems} shows examples of the problematic default behavior of Git in each of these scenarios.\todo{write this, can move to supplement if needed}
 
-The Convergent Replicated Data Type (CRDT) was proposed as a simpler alternative to the OT architecture.
-CRDTs do not require causal ordering maintenance because the language of edits must form a join semilattice,
-thereby ensuring the edit state update function increases monotoncally.
-(Foreshadow solution by warning about a shortcoming of CRDT.)
+The \textbf{granularity problem} 
+arises when both commits make edits to different locations within a single line of code. For example, one patch might rename a function argument while the other adds a new argument. 
+Similarly, if one patch renames an type while another 
+makes unrelated changes to code that references that type,
+there will be conflicts at every line of code shared between the two patches. 
+% A notable special case is the \textbf{wrapping problem}, which arises when one patch changes the nesting of code structures, e.g.
+by wrapping a code block within a new control flow construct, thereby changing the indentation of every line in the block.
+These changes can cause conflicts.
 
-(Propose Grove as a solution)
+The \textbf{relocation conflict problem} arises when two patches relocate a code block to different locations. 
+Standard patch languages operationalize code relocation as simply a deletion paired with an insertion.
+The merge will therefore fail to identify this legitimate location conflict and instead silently duplicate the code block at both locations.
 
-% OT is a natural fit for character-based editing, but the architecture is hard to extend, so it is rarely used for more structured documents.
-% CRDT imposes less restrictions on the structure of the underlying document,
-% providing a more natural substrate for collaborative structure editing.
+The \textbf{relocation modification problem} arises when one commit relocates a code block that another commit modifies. 
+A naive approach would silently leave the modifications at the original location.
+A more sophisticated block-based approach, like that deployed by \li{git}, might indicate a conflict when an insertion occurs within the bounds of a code block that has been deleted.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+These classic problems have motivated research into richer patch languages, more sophisticated patch synthesis algorithms, and corresponding improvements to three-way merge algorithms. 
+For example, systems like Git address the nesting problem by allowing indentation changes from one patch to be merged with other changes that do not modify indentation. 
+More sophisticated systems like GumTree, SmartDifferencer\todo{cites, others} deploy parsers and tree differencing algorithms to better
+address the granularity problem.
 
-% tease out some common problems (tree-native, collaborative)
+Addressing the relocation-related problems is more difficult.
+A common approach is to enrich the patch language to make code relocation a primitive command.
+However, correctly synthesizing code relocation commands given only the initial and final state of the repository requires heuristics. 
+The typical approach is to assume that a matching deletion and insertion is due to code relocation. However, relocated code is often also modified. 
+In these cases, it is difficult to determine whether a deletion and a similar but not identical insertion are related by relocation, rather than coincidental code similarity. 
+Developers often intentionally copy and then modify code, so there may be multiple partially matching insertions for a given deletion and there is no clear way to decide which, if any, are related by relocation.
+The developer's actual actions, e.g.
+cuts, copies, and pastes, are not persisted into the file system, nor are there persistent identifiers associated with code structures represented as text, so text-file-based systems have no choice but to deploy imperfect heuristics.
 
-% propose a solution
+\paragraph{Contribution 1: Grove: A Collaborative Structure Editor Calculus} 
+This paper considers the problem of collaborative editing for \emph{structure editors},
+which eschew text editing.
+Instead, developers code by applying tree edit actions directly to a continuously evolving \emph{program sketch}, i.e.
+a syntax tree with holes, shown projected visually in various ways to the developer.
+Structure editing has been studied since the 1980s with the Cornell Program Synthesizer 
+and research continues to this day, with numerous active projects including Scratch, Snap!, and other block-based editors (which are widely used in educational and end-user programming settings), Jetbrains MPS (which has been deployed in industry), and Hazel (a live functional programming environment rooted in a structure editor calculus called Hazelnut, which serves as
+an active research platform).
 
-% algorithms for collaborative text editing are well developed
-% natural fit for text
-% not a natural fit for trees
-% so if you want to build a grove, a more natural approach is needed
+Inspired by Hazelnut, this paper introduces Grove, a \emph{collaborative structure editor calculus} for arbitrary syntax trees that does not suffer from the problems just outlined. 
+This is in large part due to a substantial simplification of the overall collaborative
+editing architecture. In particular, we \emph{eliminate patch synthesis (i.e. diff algorithms) entirely}, 
+instead deriving patches directly from the log of edits performed by the developer. 
+We also \emph{eliminate the need for three-way merge algorithms (i.e. operational transforms) entirely}.
+Instead, we define the patch language such that all edits commute, 
+so remote patches can be applied without transformation. 
+We prove a \emph{convergence theorem} that ensures that replicas (a.k.a. clones) of a repository will converge to the same state 
+when the same set of patches are applied, regardless of the order in which they are applied.
+The patch language forms what is known as a \emph{commutative replicated data type (CmRDT)}.
 
-% a lot of collaborative editing of other sorts (not necessarily programming) requires a tree model,
-% but that problem doesn't seem to have been study directly (maybe except JSON CRDT).
+Defining a structure editor calculus that supports code insertion, deletion, and relocation using only commutative edits, and avoiding the problems outlined above, is not trivial.
+The Hazelnut action language is neither commutative nor does it support relocation.
+Relocation is particularly challenging because of the potential for relocation conflicts,
+including the potential for cyclic relocation (when one commit relocates node $A$ beneath node $B$, and the other \emph{vice versa}).
+
+This need to represent conflicting states means that we cannot use a single syntax tree as Grove's 
+core data structure.
+Instead, we use a directed labeled multi-graph.
+A vertex corresponds to a syntax tree node and is labeled with a \emph{unique identifier (UID)} and a \emph{constructor}, e.g.
+\li{Plus}.
+An edge is also labeled with a UID and establishes a parent-child relationships at a labeled \emph{position} for the parent vertex's constructor, e.g.
+at the \li{L} or \li{R} position of the \li{Plus} constructor.
+We refer to a parent vertex and position collectively as a \emph{location}.
+
+The \emph{Grove patch language} is quite simple: a patch can insert or delete an edge (which might cause the creation of a mentioned vertex if it did not already exist). 
+To ensure commutativity, deletion of an edge is permanent, i.e.
+the edges form a two-phase set (2P-Set) CmRDT. The edit actions that users perform are given meaning by a straightforward translation to a graph patch. 
+Relocation simply translates to 
+edge deletion and insertion. Critically, vertices are \emph{not} deleted during relocation,
+so we do not need to deploy heuristics to identify relocated nodes.
+
+Structure editors are designed to provide editing affordances for trees, not graphs where there may be any number of children at a given location, vertices may have any number of parents, and where there may be cycles.
+To support conventional tree-based structure editing, 
+we define a \emph{decomposition} of our program graph into a \emph{grove}, which is a set of 
+programs with holes, local conflicts, and conflict references between them to account for motifs that arise when editing collaboratively. 
+In particular, a location with no out-edges decomposes to a \emph{hole}. 
+More than one out-edge at a given location decomposes to an explicitly represented \emph{local conflict}. 
+More than one in-edge indicates that a vertex has a \emph{location conflict}, so we leave a \emph{location conflict reference} at each of the conflicting locations. 
+Finally, cycles are broken during decomposition by leaving a \emph{cycle conflict reference} at an arbitrary (but deterministically chosen) edge. 
+Resolving these various conflicts simply requires 
+manipulating these constructs like any other syntactic construct, e.g. deleting all but one location conflict reference to determine a unique location for a node.
+
+\paragraph{Problem 2: Semantic Gaps During Conflict Resolution}
+When working with traditional version control systems,
+resolving conflicts can take time and require reasoning about 
+syntax, types, and program behavior.
+Traditionally, however, conflicts are indicated by inserting extra-linguistic markers
+into files. 
+These markers are not typically understood by the parser, and 
+because they include conflicting alternatives, they cannot generally be removed 
+or concatenated to result in a sensible program. 
+Consequently, language services that require a well-formed, meaningful program
+(e.g. type error localization, go-to-definition, live evaluation and so on) 
+either fail to operate or exhibit gaps in service, e.g. because they are relying on 
+data from a compile prior to the merge attempt.
+Developers are left to reason without the aid of much of their tooling during conflict resolution.
+This is an instance of the more general \textbf{semantic gap problem} when programming tools encounter incomplete programs~\cite{HazelSNAPL}.
+
+The previous work on the Hazelnut structure editor calculus addresses the semantic gap problem in the single-user setting by 
+defining a type system and type error localization system (the \emph{marked lambda calculus}) for incomplete programs, i.e. programs with holes. 
+Notably, type error localization is proven to be \emph{total}, i.e.
+the system is able to assign static meaning to \emph{every} syntactically well-formed expression by inserting \emph{marks} to localize errors.
+Marking employs local type inference as codified by a \emph{bidirectional type system} as well as \emph{gradual typing}, i.e.
+the theory of \emph{type holes}, to recover from situations where type errors make it impossible to
+determine a known type. 
+A separate type hole filling phase deploys unification-based (i.e. non-local) type inference to fill type holes when possible, 
+or allows the user to interactively select from hole fillings that partially satisfy generated constraints when there are type conflicts.
+
+\paragraph{Contribution 2: Total Type Error Localization and Recovery for Groves}
+
+This paper extends this work to develop a total type error localization system for groves, introduced in Contribution 1 above as the result of decomposing a commutatively edited graph into a set of terms with empty holes, local conflicts, location conflict references, and cycle conflict references.
+This paper develops a type (and type error localization) discipline for handling these novel constructs.
+We follow the marked lambda calculus in rooting our \emph{marked grove calculus} in bidirectional type checking, deploying gradual typing when conflicts do not allow a single type to be inferred, and then layering on a unification-based type inference system to opportunistically fill holes or suggest partial solutions when there are conflicting types due to conflicting syntax.
+
+\paragraph{Paper Outline}
+
+\autoref{sec:Grove By Example} introduces Grove by example, demonstrating its behavior in each of the problematic scenarios named above. 
+\autoref{sec:Formalism} then formally defines Grove's graph structure, commutative patch language, grove decomposition procedure, and edit action language.
+We establish key metatheoretic properties using the Agda proof assistant.
+\autoref{sec:Type System} instantiates Grove with a simply typed lambda calculus, then defines a bidirectional type and type localization system for groves and proves totality and other key metatheoretic properties using the Agda proof assistant.
+\autoref{sec:Implementation} describes our OCaml implementation of Grove, which is defined modularly to allow it to be instantiated with arbitrary syntax trees. 
+\autoref{sec:Related Work} reviews related work in more detail. 
+\autoref{sec:Discussion and Conclusion} concludes with a discussion and directions for future work on collaborative editing.
+
+% A lot of collaborative editing of other sorts (not necessarily programming) requires a tree model,
+% But that problem doesn't seem to have been studied directly (maybe except JSON CRDT).
 
 % A \emph{collaborative structure editor} is an editor that
-% (1) allows multiple concurrent users to work on a shared document, while also
+% (1) allows multiple concurrent users to work on a shared document while also
 % (2) providing structure-aware editor services such as projectional editing, syntax highlighting, or automatic code folding.
@@
 % and complementary design challenges (subject-subject versus subject-object harmony),
-% to our knowledge, there is no comprehensive, principled account of their combined use.
+% To our knowledge, there is no comprehensive, principled account of their combined use.
 
@@
 % OT is complex and largely textual.
-% OT can make sense for real time systems: users typically change one character at a time, and instant feedback can help to prevent conflicts.
+% OT can make sense for real-time systems: users typically change one character at a time, and instant feedback can help to prevent conflicts.
 % On the other hand, OT system designs can be difficult to extend.
@@
 % Alternatively, there's CRDT. (Peritext?)
-% CRDT is easier to implement, but harder to design for.
+% CRDT is easier to implement but harder to design for.
 % There's an All-CRDT editor---it turned out to be not so realistic. (what's it called again?)
 
-% Structure editing has been a recurring theme in the computer science literature since at least Engelbart's ``Mother of All Demos.''
+% Structure editing has been a recurring theme in computer science literature since at least Engelbart's ``Mother of All Demos.''
 % Provides automation for domain experts and reduces the barrier to entry for everyone else.
@@
 
-\newpage
+% \newpage
 
-Motivation:
+% Motivation:
 
-- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
-is good and important as computing grows
+% - collaborative editing (both synchronous ala Google Docs and asynchronous version control)
+% is good and important as computing grows
 
-- semantic structure editing is good because it solves the gap problem (semantic editor services
-are always available) -- cite Hazelnut papers (talk about holes)
+% - semantic structure editing is good because it solves the gap problem (semantic editor services
+% are always available) -- cite Hazelnut papers (talk about holes)
 
-- previous approaches to collaborative editing have limitations
+% - previous approaches to collaborative editing have limitations
 
-- diff/merge based approaches (trying to solve the inverse problem based on final states --
-you lose the actual actions that were performed, and have to reconstruct them or an approx.
-of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
-some papers have started to look at that, but fundamentally we don't want to throw away the
-knowledge we have about the edits!)
+% - diff/merge-based approaches (trying to solve the inverse problem based on final states --
+% you lose the actual actions that were performed and have to reconstruct them or an approx.
+% of them i.e., add line/delete line actions -- would need to adapt this to structure editing,
+% Some papers have started to look at that, but fundamentally, we don't want to throw away the
+% knowledge we have about the edits!)
 
-- operational transforms (complexity, you have to patch previous actions based on new actions)
+% - operational transforms (complexity, you have to patch previous actions based on new actions)
 
-- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
-because it is relatively simple: you just send all the edits to all the replicas and they are
-convergent by design
+% - CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
+% because it is relatively simple: you just send all the edits to all the replicas and they are
+% convergent by design
 
-- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
-the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
-allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
-using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
-the semantic editor services will be disabled because the program is going to be broken in multiple places
-transiently)
+% - we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
+% the sensibility invariant of Hazelnut, i.e., every editor state has meaning. mention that maintaining sensibility
+% allows scaling of semantic editor services in the presence of a large number of collaborators (in contrast,
+% using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
+% The semantic editor services will be disabled because the program is going to be broken in multiple places
+% transiently)
 
-this is tricky because:
+% This is tricky because:
 
-- some edits might be conflicting -- solve this with "conflict holes"
+% - some edits might be conflicting -- solve this with "conflict holes"
 
-- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
+% - adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)
 
-- since we are commutative, we solve both synchronus and async collaborative editing
+% - since we are commutative, we solve both synchronous and async collaborative editing
 
-- and this resolves issues around merges and conflicts
+% - and this resolves issues around merges and conflicts
 
-- contribution of this paper is to solve these problems from type-theoretic first principles:
+% - contribution of this paper is to solve these problems from type-theoretic first principles:
 
-- ...
+% - ...
 
-- Hazel
+% - Hazel
 
-\subsection{Contributions and Paper Organization}%
-\label{sec:Contributions and Paper Organization}
+% \subsection{Contributions and Paper Organization}%
+% \label{sec: Contributions and Paper Organization}

### 2024-07-11 marking-system.tex
Updates from Overleaf

\section{Marking System}%
\label{sec:Marking System}

[Motivation about how groves may be ill-typed because of arbitrary actions and conflicts, causing gaps in semantic services (rehash some of intro)]

[POPL 24] introduce the \emph{marked lambda calculus} to ...

We enrich groves and their underlying syntax of terms by \emph{marking} them, that is, systematically adding error marks during the type-checking process to enable \emph{total error localization and recovery}, refining the work from [popl 24]\todo{cite} to fit grove's collaborative needs. This allows the editor to optimistically recover from errors by formalizing the cases in which they can occur, and also provide more helpful error messages and editor services.

In addition to type-checking and marking expressions, we also gather type-inference constraints and \emph{local-info} sets, further augmenting the editor with the ability to inspect and mark \emph{multi-location} and \emph{cycle} conflicts against the different contexts their occurrences can appear in, providing valuable information to the user while resolving these conflicts.

In \autoref{sub:marking-recap}, we revisit the marked lambda calculus and establish all the required machinery to type-check and mark expressions using a conventional bidirectional system \todo{cite TAPL?}. Next, in \autoref{sub:marking-groves}, we introduce the constructs required to handle the various conflicts that can arise from the graph-based setting when type-checking a tree-native surface language. And finally, we describe the mechanization for the marking system in \autoref{sub:marking-agda} done using the Agda \todo{cite} proof assistant.

[Recap marking judgments using old lam and ap rules]

% Further, we will also develop the notion of type consistency and matched arrows from gradual typing to setup the necessary infrastructure required to bidirectionally type-check expressions. This gives rise to typing judgments for both marked and unmarked expressions along with a \emph{marking judgment}, the heart of the [POPL 24] work. Then, we adopt a constraint based type inference system to systematically gather constraints from the child nodes to infer and construct types for the top level program allowing the user to omit type annotations.

% Although, most of the machinery remains the same from [POPL 24], it is interesting to note the case of multiparent references where such terms can be typed according to multiple contexts. Additionally, the marking system along with the base grammar for groves has been mechanized using the Agda proof assistant.

\subsection{Marking Groves}
\label{sub:marking-groves}
We extend the marked lambda calculus to groves. Most of the machinery remains largely the same as in [POPL 24]; we focus our attention on conflicts.

[Discuss the unmarked language as an instantiation of the generic term language described in Sec 3.2; probably can't do this properly until all the language for section 3 is figured out]

First, we introduce here the syntax of the \emph{unmarked} language which comprises two mutually defined expression sorts: $Exp$ \& $ExpList$, denoted by $\EMV$ and $\ELV$ respectively as displayed in \autoref{fig:unmarked-language}. In addition, this unmarked language is an instantiation of the generic term syntax introduced in \autoref{sub:Groves}, with $\EMV$ \& $\TMV$ being the expression and type sorts along with their mutually defined list counterparts $\ELV$ \& $\TLMV$ that allow us to model holes (empty list), \emph{regular} singleton expressions, and local conflicts. Type-checking this base unmarked language is straightforward using conventional bidirectional type-checking rules and we introduce the sort of semantic types $STyp$, denoted by $\sigma$, which builds on ideas from \todo{cite GTLC} Gradually Typed Lambda Calculus (GTLC). This is achieved using 4 mutual type-checking judgments as displayed in \autoref{fig:unmarked-judgments}. The rules are not very different from most simply typed lambda calculi except for the cases where we treat \emph{conflicts} as holes. For the sake of brevity, we have chosen to omit the rules here, but they are expressed in detail in \autoref{sec:marked-unmarked-expressions}.

\begin{figure}
    \judgbox{\ctxSynTypeU{\ctx}{\EMV}{\sigma}} 
    \judgbox{\ctxSynTypeU{\ctx}{\ELV}{\sigma}}
    \judgbox{\ctxAnaTypeU{\ctx}{\EMV}{\sigma}} 
    \judgbox{\ctxAnaTypeU{\ctx}{\ELV}{\sigma}}
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Type-checking Unmarked Expressions}
    \label{fig:unmarked-judgments}
\end{figure}

\begin{figure}
    \[
    \begin{array}{lclll}
     \EMV \in & Exp & \coloneqq & 
        \var{x}
        \mid  \num 
        \mid \plus{\ELV_1}{\ELV_2} 
        \mid \mult{\ELV_1}{\ELV_2}
        \mid \lam{x}{\TLMV}{\ELV}
        \mid \app{\ELV_1}{\ELV_2}
        \mid \multiref
        \mid \uniref \\
     \ELV \in & ExpList & \coloneqq & 
        \ehole
        \mid \lexp{\EMV} 
        \mid \conflict{\EMV} \\
    \TMV \in & \Typ & \coloneqq & 
        \tnum 
        \mid \tarrow 
        \mid \multiref 
        \mid \uniref \\ 
    \TLMV \in & TypList & \coloneqq &
        \ehole
        \mid \lexp{\TMV}
        \mid \conflict{\TMV} \\
    \end{array}
    \]
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Unmarked Language}
    \label{fig:unmarked-language}
\end{figure}

\begin{figure}
    \[
    \begin{array}{lclll}
    \ECMV \in & MExp & \coloneqq &
        \var{x}
        \mid \num
        \mid \plus{\ELMV_1}{\ELMV_2}
        \mid \mult{\ELMV_1}{\ELMV_2}
        \mid \lam{x}{\TLMV}{\ELV}
        \mid \app{\ELMV_1}{\ELMV_2}
        \mid \multiref
        \mid \uniref \\ 
        & & & \ECFree{\var{x}} 
        \mid \incontype
        \mid \lamasc
        \mid \lamanamarr
        \mid \apsynmarr \\ 
        % \mid \lamge \\
     \ELMV \in & MExpList & \coloneqq &
        \ehole
        \mid \lexp{\ECMV}
        \mid \conflict{\ECMV} \\
    \end{array}
    \]
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Marked Language}
    \label{fig:marked-language}
\end{figure}

\begin{figure}
    \[
    \begin{array}{lclll}
    \sigma \in & STyp & \coloneqq & 
        \TUnknown
        \mid \tNum{}
        \mid \starrow \\
    \end{array}
    \]
    \judgbox{\graphErase{\TLMV} = \sigma}
    \[\begin{array}{rrcl}
        \Prov & \Provp & \Coloneqq & u \mid exp(u) \mid \rightarrow_L(\Provp) \mid \rightarrow_R(\Provp)\\
        \TMName  & \TMV  & \Coloneqq & \dots \mid \TUnknown^{p} \mid \TUnknown^{(v,p)}\\
    \end{array}\]
    \judgbox{\ensuremath{\matchedArrowConstraint{\sigma}{\sigma_1}{\sigma_2}{C}}}
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Other machinery}
    \label{fig:marking-machinery}
\end{figure}

\begin{figure}
    \judgbox{\synMarkConstraint{\ctx}{\EMV}{\ECMV}{\sigma}{\cConstraint}{\mConstraint}} 
    \judgbox{\synMarkConstraint{\ctx}{\ELV}{\ELMV}{\sigma}{\cConstraint}{\mConstraint}} 
    \judgbox{\anaMarkConstraint{\ctx}{\EMV}{\ECMV}{\sigma}{\cConstraint}{\mConstraint}} 
    \judgbox{\anaMarkConstraint{\ctx}{\ELV}{\ELMV}{\sigma}{\cConstraint}{\mConstraint}} 
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Marking Expressions}
    \label{fig:marking-judgments}
\end{figure}

Next, we introduce the heart of the marking machinery in \autoref{fig:marking-judgments} \todo{add fig}, again in the form of 4 mutually defined judgments along with the marked language denoted by $\ECMV$ \& $\ELMV$. However, here, we add error marks to the expression when one of the premises fails, and type-checking cannot continue further, thereby adding first-class support for error localization. In addition to marking the expressions and type-checking them, at this stage, we also collect inference constraints and \emph{local-info} sets, which come into play when the user has to deal with \emph{multi-location} and \emph{cycle} conflicts.

Now, we examine the marking rules for lambda abstractions and function applications.

\begin{mathpar}
\label{inf:lam-ana}
\inferrule[MKACLam1]{
        \graphErase{\TLMV} = \sigma \\
        \matchedArrowConstraint{\sigma_3}{\sigma_1}{\sigma_2}{\cConstraint_1} \\
        \consistent{\sigma}{\sigma_1} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\sigma_2}{\cConstraint_2}{\mConstraint_1}
    }{
        \anaMarkConstraint{\ctx}{\lam{x}{\TLMV}{\EMV}}{\lam{x}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\sigma}{\sigma_1}}}{\mConstraint_1}
    }

     \inferrule[MKACLam2]{
        \graphErase{\TLMV} = \sigma \\
        \notMatchedArrow{\sigma_3} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\TUnknown^{anon}}{\cConstraint}{\mConstraint}
    }{
        \anaMarkConstraint{\ctx}{\reglam}{\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint \cup \Setof{\constrain{\TUnknown^{exp(u)}}{\sigma_3}}}{\mConstraint}
    }

    \inferrule[MKACLam3]{
        \erase{\TLMV} = \sigma \\
        \matchedArrowConstraint{\sigma_3}{\sigma_1}{\sigma_2}{\cConstraint_1} \\
        \inconsistent{\sigma}{\sigma_1} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma_1}}{\ELV}{\ELMV}{\sigma_2}{\cConstraint_2}{\mConstraint}
    }{
        \anaMarkConstraint{\ctx}{\reglam}{\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\TUnknown^{exp(u)}}{\sigma_3}}}{\mConstraint}
    }
\end{mathpar}

- core unmarked language
- marked language
- semantic types
- matched arrows from GTLC
- local-info stuff
- provenance and inference stuff

% \subsubsection{Holes} TODO holes aren't really new?

\subsubsection{Local Conflicts}
To mark local conflicts, we recursively mark all subterms, gather the constraint and local info sets, and synthesize the unknown type; such conflicts are treated essentially as holes.

\[
\inferrule[MKSCConflict]{ 
    \Setof{\synMarkConstraint{\ctx}{\EMV}{\ECMV}{\sigma_i}{\cConstraint_i}{\mConstraint_i}}_{i<n}
}{
    \synMarkConstraint{\ctx}{\conflict{\EMV}}{\conflict{\ECMV}}{\TUnknown^{(v,p)}}{\Setof{\constrain{\TUnknown^{(exp(u_i))}}{\sigma_i}}_{i<n} \cup \bigcup_{i<n} \cConstraint_i}{\bigcup_{i<n} \mConstraint_1}
}
\]

\subsubsection{Multi-Location Conflicts}

\[
\inferrule[MKSCMultiParent]{ }{
    \synMarkConstraint{\ctx}{\multiref}{\multiref}{\TUnknown^{exp(id-of(v))}}{\constraintNil{}}{(id-of(v), w , \ctx, syn)}
}
\]

\subsubsection{Cycle Conflicts}

\[
\inferrule[MKSCUnicycle]{ }{
    \synMarkConstraint{\ctx}{\uniref}{\uniref}{\TUnknown^{exp(id-of(v)}}{\constraintNil{}}{(id-of(v),w , \ctx, syn)}
}  
\]

\subsection{Agda Mechanization}
\label{sub:marking-agda}
Our extension of the marked lambda calculus has been mechanized in the Agda proof assistant, taking the standard approach of modeling judgments as inductive datatypes and inference rules as constructors. Marked terms are implicitly typed, allowing local info mappings to be computed as a function of these terms.

% TODO discuss how no type hole inference/constraint generation so 
% TODO discuss how it is separate from the grove agda but a correspondence can be proven?

% Questions while porting over marking system
% - new judgments for marking expressions
% - need to prove totality and unicity of marking
% - then type checking both regular and marked expressions
% - need to check if we need to introduce new marked errors
% - typeset all of this
% - get all of this working in Agda?

### 2024-07-11 related-work.tex
Updates from Overleaf

\section{Related Work}%
\label{sec:Related Work}

Contrast with a traditional diff

Similar to how git merged across moves.
A notable difference though, is that \texttt{git} detects
file moves by structural similarity rather than
our nominal equality.
(We exploit this in Subsection~REF:TODO).

Compare this to what happens when there are conflicts in a traditional,
line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
when a merge conflict happens.
These show alternate versions of the code, and it is up to the user
to replace those alternates with the desired merged result.

Another place our model is different
is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
Finally, in our model, since all changes are in terms of
the structure of the code rather than lines,
all changes respect that structure.
This is unlike the line-based model where there is no guarantee
that a merge conflict follows the grammatical structure of the code.

\subsection{Trees}

\subsubsection{Structure editors}

Hazel~\citep{Omar:2019:10.1145/3290327}

\subsubsection{Tree diff}

% We are using action based instead of tree diff

% We have to deal with merge

% We have to deal with edits from multiple people

% J. W. Hunt and M. D. McIlroy. 1976. An Algorithm for Differential File Comparison. Technical Report CSTR 41. Bell Laboratories, Murray Hill, NJ.

% Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}

%   Approximating Tree Edit Distance through String Edit Distance \url{https://dl.acm.org/doi/10.5555/3118232.3118518}

%   Meaningful change detection in structured data \url{https://dl.acm.org/doi/10.1145/253260.253266}

%   An optimal decomposition algorithm for tree edit distance \url{https://dl.acm.org/doi/10.5555/2394539.2394560}

%   Diff/TS: A Tool for Fine-Grained Structural Change Analysis \url{https://dl.acm.org/doi/10.1109/WCRE.2008.44}

% An efficient algorithm for type-safe structural diffing \url{https://dl.acm.org/doi/10.1145/3341717}

%   Precise Version Control of Trees with Line-Based Version Control Systems \url{https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9}

%   A survey on tree edit distance and related problems \url{https://dl.acm.org/doi/10.1016/j.tcs.2004.12.030}

%   Cycle-aware minimization of acyclic deterministic finite-state automata \url{https://dl.acm.org/doi/10.1016/j.dam.2013.08.003}

%   Computing the Edit-Distance between Unrooted Ordered Trees \url{https://dl.acm.org/doi/10.5555/647908.740125}

%   Type-safe diff for families of datatypes \url{https://dl.acm.org/doi/10.1145/1596614.1596624}

%   A Categorical Theory of Patches \url{https://dl.acm.org/doi/10.1016/j.entcs.2013.09.018}

%   Type-directed diffing of structured data \url{https://dl.acm.org/doi/10.1145/3122975.3122976}

%   The Semantics of Version Control \url{https://dl.acm.org/doi/10.1145/2661136.2661137}

%   The Tree-to-Tree Correction Problem \url{https://dl.acm.org/doi/10.1145/322139.322143}

%   Generic Diff3 for algebraic datatypes \url{https://dl.acm.org/doi/10.1145/2976022.2976026}

\subsection{Version Control}

% Git \url{https://git-scm.com/}

% Darcs \url{https://darcs.net/}

%   Darcs: distributed version management in haskell \url{https://dl.acm.org/doi/10.1145/1088348.1088349}

% Hg? \url{https://www.mercurial-scm.org/}

% SVN \url{https://subversion.apache.org/}

% Pijul and (Anu is a rewrite of Pijul and seems to have been subsumed into Pijul)

%   \url{https://pijul.org/}
%   \url{https://pijul.org/manual/theory.html}

%   \url{https://tahoe-lafs.org/~zooko/badmerge/simple.html}

\subsection{Collaborative Editing}

% Collaborative Structure Editing

% SmallTalk collaboration with images

% TouchDevelop papers

% Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)

% \subsection{CRDTs}
% (Are we a known CRDT?)

% List of CRTD papers: \url{https://crdt.tech/papers.html}

% Bottom = Tombstone

% \url{https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read}
%  - Add-Remove Partial Order data type
%  - 2P2P-Sets
%  - Replicated Growable Array

% \url{https://github.com/PsychoLlama/graph-crdt}
%  - Graph CRDT
%  - Uses a LWW-E-Set

% \url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
%  We don't have interleaving problems because
%    - everything is relative to a specific ID not a position
%    - we don't try to auto resolve
%    - we are tree not list
%    Part three: moving sub tree
%    - Last parent writter wins (prevents cycles)
%      Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?

% A commutative replicated data type for cooperative editing
%   \url{https://hal.inria.fr/inria-00445975/document}
%   Describes TreeDoc but this uses a document model that is a list (tree is just how it is implemented)

% Logoot : a Scalable Optimistic Replication Algorithm for Collaborative Editing on P2P Networks
%   \url{http://pagesperso.lina.univ-nantes.fr/~molli-p/pmwiki/uploads/Main/weiss09.pdf}

% Specification and Complexity of Collaborative Text Editing
%   \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/07/podc16-complete.pdf}

% LSEQ: an Adaptive Structure for Sequences in Distributed Collaborative Editing,
%   \url{https://hal.archives-ouvertes.fr/file/index/docid/921633/filename/fp025-nedelec.pdf}

% Data consistency for P2P collaborative editing
%   \url{https://hal.archives-ouvertes.fr/file/index/docid/108523/filename/OsterCSCW06.pdf}

% Interleaving anomalies in collaborative text editors
%   \url{https://martin.kleppmann.com/papers/interleaving-papoc19.pdf}

% Moving Elements in List CRDTs
%   \url{https://martin.kleppmann.com/papers/list-move-papoc20.pdf}

% A highly-available move operation for replicated trees and distributed filesystems
%   \url{https://martin.kleppmann.com/papers/move-op.pdf}

% ...
%  commutative replicated data types CmRDT
%  convergent replicated data types, or CvRDTs
%  Delta state CRDTs[12][13] (or simply Delta CRDTs

% The Causal Graph CRDT for Complex Document Structure
%   \url{https://dl.acm.org/doi/10.1145/3209280.3229110}

% \url{https://en.wikipedia.org/wiki/Conflict-free_replicated_data_type}

% G-Set
% PN-Set
% 2P-Set
% LWW-element-Set (Last-Write-Wins)
% OR-Set
% MV-Register: Multi-Value Register
% U-Set (this is what we are for edges, not OR-set due to duplicates)
% Add-Remove Partial Order data type
%   2P-Set for vertices, and a G-Set for edges.

% A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
%   \url{https://hal.inria.fr/inria-00555588/document}

% CRDTs: Consistency without concurrency control
%   \url{https://arxiv.org/abs/0907.0929}

% https://medium.com/@amberovsky/crdt-conflict-free-replicated-data-types-b4bfc8459d26
%  removeVertex() has priority, all incident edges are removed
%  addEdge() has priority, all removed vertices are re-added
%  Delay removeVertex() execution till all concurrent removeVertex() are executed.
% First one is 2P2P-Set

% https://crdt.tech/papers.html

% Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster. Co-design and verification of an available file system. In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381. Springer LNCS volume 10747, January 2018. [ bib | DOI | .pdf ]
%   \url{http://dx.doi.org/10.1007/978-3-319-73721-8_17}
%   https://pages.lip6.fr/Marc.Shapiro/papers/VMCAI-2018-filesys.pdf

% Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]
%   http://dx.doi.org/10.1109/TPDS.2017.2697382
%   http://arxiv.org/abs/1608.03960

% Vinh Tao, Marc Shapiro, and Vianney Rancurel. Merging semantics for conflict updates in geo-distributed file systems. In 8th ACM International Systems and Storage Conference, SYSTOR 2015. ACM, May 2015. [ bib | DOI | .pdf ]
%   http://dx.doi.org/10.1145/2757667.2757683
%   https://pages.lip6.fr/Marc.Shapiro/papers/geodistr-FS-Systor-2015.pdf

% Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]
%   http://arxiv.org/abs/1207.5990
%   https://hal.inria.fr/hal-00720681/

% Stéphane Martin, Pascal Urso, and Stéphane Weiss. Scalable XML collaborative editing with undo. In On the Move to Meaningful Internet Systems (OTM), pages 507--514. Springer LNCS volume 6426, October 2010. [ bib | DOI | arXiv ]
%   \url{http://dx.doi.org/10.1007/978-3-642-16934-2_37}
%   http://arxiv.org/abs/1010.3615

% 2P-Sets
% Anomaly: Creating a lone deleted vertex requires create and delete of otherwise unneeded edge

% Operational Transforms

% Etherpad

% Live Share

\subsection{Synchronization}

% Unison
% \url{https://www.cis.upenn.edu/~bcpierce/unison/}
% \url{https://www.cis.upenn.edu/%7Ebcpierce/papers/index.shtml#File%20Synchronization}

\subsection{Homotopical Patch Theory}

% Homotopical Patch Theory: \url{https://www.cambridge.org/core/journals/journal-of-functional-programming/article/homotopical-patch-theory/42AD8BB8A91688BCAC16FD4D6A2C3FE7}
% Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}

### 2024-07-11 formalism.tex
Updates from Overleaf

@@
 Edge states determine which edges are live in a graph.
-For example, if $G(\e) = \Plus$ for some $G : \E \rightarrow \Sigma$, $\e \in \E$,
-then we say $\e$ is \emph{live}, i.e., it is part of the live subgraph that is decomposed into a grove, as explained in \autoref{sub:Graph Decomposition}.
+For some $G : \E \rightarrow \Sigma$, $\e \in \E$, $G(\e) = \bot$ if $\e$ has not yet been constructed in the graph.
+$G(\e) = \Plus$ if $\e$ is \textit{live} in the graph, meaning it is relevant to the user and the type system.
+$G(\e) = \Minus$ if $\e$ has been deleted.
 The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
+We assume a graph assigns only finitely many edges to $\Plus$ to ensure the finiteness of the data structure on which further computation is done.
 
@@
 Essentially, these patch command semantics act as a transition system between graphs, \emph{joining} the new edge state with the corresponding edge state of the graph to which the patch command is being applied.
+Since a patch command can only change the value of $G$ on one input, applying the patch command preserves the property that $G$ maps only a finite set of edges to $\Plus$.
 A \textit{graph patch} is a sequence of graph patch commands.
@@
 
-Our source language (\autoref{fig:Syntax}) is based on a simply typed lambda calculus\todo{is there anything here that has to do with the STLC?
-we just have constructors} with an expression sort and a type sort.
-Each sort is extended with cases for multi-location references and unicycle references.
-Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define secondary sorts \textit{ExpList} and \textit{TypList} to represent these possibilities.
+Given a source language parameterized by a set of constructors $k \in \K$ and an arity function $\K \rightarrow \mathbb{N}$, we define an augmented term language as specified in \autoref{fig:Syntax}.
+Terms are extended with cases for multi-location references and unicycle references.
+Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define a secondary sort \textit{SubTerm} to represent these possibilities.
 A source with no outgoing edges corresponds to an empty hole, a source with one outgoing edge corresponds to an ordinary subterm, and a source with multiple outgoing edges corresponds to a local conflict between multiple terms.
@@
 TODO), each empty hole and local conflict carries its source.
-Finally, to support user navigation of the grove, each references also carries the UID of the edge that leads to it.
+Finally, to support user navigation of the grove, each reference also carries the UID of the edge that leads to it.
 
@@
 
-A live subgraph (or simply graph), which is just a finite set of edges, decomposes into a grove, which is a set of terms.
+A live subgraph, which is just a finite set of edges, decomposes into a set of terms.
 In a tree, each node except the root as a unique parent, and there are no cycles.
@@
 
-\begin{theorem}
-    (Classification) TODO
-\end{theorem}
+\begin{align}
+    \text{parents}_{G}(v) &= \SetOf{v' \in \V \SuchThat{(w, (v', p), v) \in G \land w \in \W \land p \in \P}}
+\end{align}
+
+\begin{definition}
+    A vertex is an \textit{NP root} in $G$ if $\text{parents}_{G}(v)=\emptyset$
+\end{definition}
+
+\begin{definition}
+    A vertex is an \textit{MP root} in $G$ if $\abs{\text{parents}_{G}(v)} > 1$
+\end{definition}
+
+\begin{definition}
+    A vertex is a \textit{U root} in $G$ if $\exists \{v_i\}_{1 \leq i \leq n}$ such that $v_1 = v_n = v$, $\text{parents}_{G}(v_i)=\{v_{i+1}\}$ for all $1 \leq i < n$, and the UID of $v$ is the minimum of the UIDs of $\{v_i\}_{1 \leq i \leq n}$.
+\end{definition}
 
@@
 The decomposition of a graph is simply the set obtained by decomposing each vertex classified as a root.
+
+\begin{definition}
+    $\text{decomp}_{G}(v) = $
+\end{definition}
 

### 2024-07-11 marking-system.tex
Updates from Overleaf

@@
 
-In \autoref{sub:marking-recap}, we revisit the marked lambda calculus and establish all the required machinery to type-check and mark expressions using a conventional bidirectional system \todo{cite TAPL?}.
+Finally, we revisit the marked lambda calculus and establish all the required machinery to type-check and mark expressions using a conventional bidirectional system \todo{cite TAPL?}.
 Next, in \autoref{sub:marking-groves}, we introduce the constructs required to handle the various conflicts that can arise from the graph-based setting when type-checking a tree-native surface language.
@@
 
+\subsubsection{The unmarked language}
+\label{sub:unmarked-lang}
 First, we introduce here the syntax of the \emph{unmarked} language which comprises two mutually defined expression sorts: $Exp$ \& $ExpList$, denoted by $\EMV$ and $\ELV$ respectively as displayed in \autoref{fig:unmarked-language}.
@@
 
+\subsubsection{The marked language}
+\label{sub:marked-lang}
+Next, we introduce the heart of the marking machinery in \autoref{fig:marking-judgments} and \autoref{fig:marked-language}, again in the form of 4 mutually defined judgments along with the marked language denoted by $\ECMV$ \& $\ELMV$.
+However, here, we add error marks to the expression when one of the premises fails, and type-checking cannot continue further, thereby adding first-class support for error localization.
+In addition to marking the expressions and type-checking them, at this stage, we also collect inference constraints and \emph{local-info} sets, which come into play when the user has to deal with \emph{multi-location} and \emph{cycle} conflicts.
+
+\begin{figure}
+    \[
+    \begin{array}{lclll}
+    \ECMV \in & MExp & \coloneqq &
+        \var{x}
+        \mid \num
+        \mid \plus{\ELMV_1}{\ELMV_2}
+        \mid \mult{\ELMV_1}{\ELMV_2}
+        \mid \lam{x}{\TLMV}{\ELV}
+        \mid \app{\ELMV_1}{\ELMV_2}
+        \mid \multiref
+        \mid \uniref \\ 
+        & & & \ECFree{\var{x}} 
+        \mid \incontype
+        \mid \lamasc
+        \mid \lamanamarr
+        \mid \apsynmarr \\ 
+        % \mid \lamge \\
+     \ELMV \in & MExpList & \coloneqq &
+        \ehole
+        \mid \lexp{\ECMV}
+        \mid \conflict{\ECMV} \\
+    \end{array}
+    \]
+    \centering
+    \includegraphics[width=0.5\linewidth]{}
+    \caption{Marked Language}
+    \label{fig:marked-language}
+\end{figure}
+
 \begin{figure}
@@
 
-Next, we introduce the heart of the marking machinery in \autoref{fig:marking-judgments} \todo{add fig}, again in the form of 4 mutually defined judgments along with the marked language denoted by $\ECMV$ \& $\ELMV$.
-However, here, we add error marks to the expression when one of the premises fails, and type-checking cannot continue further, thereby adding first-class support for error localization.
-In addition to marking the expressions and type-checking them, at this stage, we also collect inference constraints and \emph{local-info} sets, which come into play when the user has to deal with \emph{multi-location} and \emph{cycle} conflicts.
-
+\subsubsection{Marking Lambda Abstractions}
+\label{sub:mark-lam}
 Now, we examine the marking rules for lambda abstractions and function applications.
+In \autoref{inf:lam-ana}, in the first case, we successfully mark the lambda abstraction by first \emph{graph-erasing} its type and then using the matched arrow judgment and consistency relations to recursively mark its body.
+While performing this marking, we gather inference constraints \& local-info maps from the matched arrow judgment and the recursive analysis of the lambda body.
+Here, all the premises hold, and the abstraction is analyzed with no issues.
+What happens when the analyzed type is not a matched arrow type or the erased type is \emph{not} consistent with the expected input type?
+Conventional type-checkers fail here and cannot proceed.
+This is where the next rules help in localizing these scenarios to the node where the error occurs.
+In $MKACLam2$, the expected type is not a matched arrow type.
+So, we mark the abstraction as such using $\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}$ which captures the premise that failed while still analyzing the body against the unknown type and gathering constraints.
+Similarly, in $MKACLam3$, the consistency check between the erased ascription and input of the expected type fails, and again, we mark the abstraction as such, depicting the inconsistency in the ascription of the lambda in $\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}$.
 
 \begin{mathpar}
-\label{inf:lam-ana}
+\label{inf:mark-lam-ana}
+\caption{Marking and Analyzing Lambda Abstractions}
 \inferrule[MKACLam1]{
@@
 \end{mathpar}
+
+\subsubsection{Marking Function Applications}
+\label{sub:mark-ap-syn}
+\begin{mathpar}
+\label{inf:ap-syn}
+\label{Marking and Synthesizing Function Applications}
+\inferrule[MKSCAp1]{
+    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma}{\cConstraint_1}{\mConstraint_1} \\
+    \matchedArrowConstraint{\sigma}{\sigma_1}{\sigma_2}{\cConstraint_2} \\
+    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\sigma_1}{\cConstraint_3}{\mConstraint_2} \\
+  }{
+    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\eApp{u}{\ELMV_1}{\ELMV_2}}{\sigma_2}{\cConstraint_1 \cup \cConstraint_2 \cup \cConstraint_3}{\mConstraint_1 \cup \mConstraint_2}
+  }
+
+  \inferrule[MKSCAp2]{
+    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma}{\cConstraint_1}{\mConstraint_1} \\
+    \notMatchedArrow{\sigma} \\
+    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\TUnknown^{\rightarrow_{L}{(exp(u))}}}{\cConstraint_2}{\mConstraint_2}
+  }{
+    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\app{\ECApSynNonMatchedArrow{\ELMV_1}}{\ELMV_2}}{\TUnknown^{\rightarrow_{R}{(exp(u))}}}{\cConstraint_1 \cup \cConstraint_2 \cup \SetOf{\constrain{\TUnknown^{exp(u)}}{\tarr{\TUnknown^{\rightarrow_L(exp(u))}}{\TUnknown^{\rightarrow_R(exp(u))}}}}
+    }{\mConstraint_1 \cup \mConstraint_2}
+  }
+\end{mathpar}
+
+Next, in \autoref{inf:ap-syn}, we look at the case when marking an application succeeds.
+For this to happen, we first synthesize a type for the first sub-term and then match it with an arrow type before analyzing the argument against the expected input type.
+What if the synthesized is not a matched arrow type?
+We insert an error mark again, indicating the failed premise in the synthetic mode in $\app{\ECApSynNonMatchedArrow{\ELMV_1}}{\ELMV_2}$.
 
@@
 
+We observe that \
+
 \subsection{Agda Mechanization}

### 2024-07-11 figures/figure-multi-child.tex
Updates from Overleaf

@@
 \label{fig:MultiChild:c}
-\figureCode{w * v + $\conflictHole{x,y}$}%
+\figureCode{w * v + $\binaryConflictHole{x}{y}$}%
 \begin{tikzpicture}
@@
 %\hfill{}
-\caption{Example of Multi-Child Conflicts.}%
+\caption{(a) Alice fills the hole at the \li{R} position of the addition with \li{x}.
+(b) Bob fills the same hole with \li{y}.
+(c) When the corresponding patches are merged, there are two edges in the \li{R} position of the addition.
+Decomposition turns these into a \emph{local conflict}, leaving it to the users to resolve the problem by performing normal edits that eliminate .}%
 \Description{This figure shows an example of multi-child conflicts}

### 2024-07-11 marking-system.tex
Updates from Overleaf

@@
 
-We enrich groves and their underlying syntax of terms by \emph{marking} them, that is, systematically adding error marks during the type-checking process to enable \emph{total error localization and recovery}, refining the work from [popl 24]\todo{cite} to fit grove's collaborative needs.
+We enrich groves and their underlying syntax of terms by \emph{marking} them, that is, systematically adding error marks during the type-checking process to enable \emph{total error localization and recovery}, refining the work from [popl 24]\todo{cite} to fit Grove's collaborative needs.
 This allows the editor to optimistically recover from errors by formalizing the cases in which they can occur, and also provide more helpful error messages and editor services.
 
-In addition to type-checking and marking expressions, we also gather type-inference constraints and \emph{local-info} sets, further augmenting the editor with the ability to inspect and mark \emph{multi-location} and \emph{cycle} conflicts against the different contexts their occurrences can appear in, providing valuable information to the user while resolving these conflicts.
+In addition to type-checking and marking expressions, we also gather type-inference constraints and \emph{location conflict contexts} , further augmenting the editor with the ability to inspect and mark \emph{multi-location} and \emph{cycle} conflicts against the different contexts their occurrences can appear in, providing valuable information to the user while resolving these conflicts.
 
-Finally, we revisit the marked lambda calculus and establish all the required machinery to type-check and mark expressions using a conventional bidirectional system \todo{cite TAPL?}.
+Next, we revisit the marked lambda calculus and establish all the required machinery to type-check and mark expressions using a conventional bidirectional system \todo{cite TAPL?}.
 Next, in \autoref{sub:marking-groves}, we introduce the constructs required to handle the various conflicts that can arise from the graph-based setting when type-checking a tree-native surface language.
@@
 
-% Further, we will also develop the notion of type consistency and matched arrows from gradual typing to setup the necessary infrastructure required to bidirectionally type-check expressions.
-This gives rise to typing judgments for both marked and unmarked expressions along with a \emph{marking judgment}, the heart of the [POPL 24] work.
-Then, we adopt a constraint based type inference system to systematically gather constraints from the child nodes to infer and construct types for the top level program allowing the user to omit type annotations.
+\subsection{Background: The Marked Lambda Calculus}
 
-% Although, most of the machinery remains the same from [POPL 24], it is interesting to note the case of multiparent references where such terms can be typed according to multiple contexts.
-Additionally, the marking system along with the base grammar for groves has been mechanized using the Agda proof assistant.
+The marked lambda calculus is a bidirectionally typed rewriting system that takes an arbitrary syntactically well-formed expression and \emph{marks} it by localizing type errors, producing a \emph{marked expression}.
+The key judgments are synthetic and analytic marking judgments that mark and type-check expressions.
+For example, the rules relevant to functions and function applications are shown in \autoref{inf:marking-rules}.
+
+\begin{mathpar}
+\label{inf:marking-rules}
+\inferrule[MKALam1]{
+        \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
+        \consistent{\TMV}{\TMV_1} \\
+        \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TMV_2}
+    }{
+        \ctxAnaFixedInto{\ctx}{\eFun{}{x}{\TMV}{\EMV}}{\eFun{}{x}{\TMV}{\ECMV}}{\TMV_3}
+    }
+
+     \inferrule[MKALam2]{
+        \notMatchedArrow{\TMV_3} \\
+        \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TUnknown}
+    }{
+        \ctxAnaFixedInto{\ctx}{\eFun{}{x}{\TMV}{\EMV}}{\ECLamAnaNonMatchedArrow{x}{\TMV}{\ECMV}}{\TMV_3}
+    }
+\end{mathpar}
+
+The first rule for analyzing functions against their expected type requires the matched arrow judgment and a type consistency check between the ascription and the input of the expected type, after which we can analyze the body against the output while recursively marking it.
+What happens when the expected type is not a matched arrow type?
+Conventional type checkers cannot proceed in this case, and type-checking halts.
+However, whenever we encounter an error during type-checking, that is, one of the premises fails, we mark the term as such, indicating that the expected type is not a matched arrow type in the analytic position.
+Yet we are still able to recursively mark the body and analyze it against the unknown type without stopping type-checking on failure.
+...
+
+In some situations, a unique type cannot be determined.
+In these cases, the system recovers by using the unknown type, a.k.a.
+the type hole, from gradual typing.
+For example, the rules for marking variables are given below:
+
+...
+
+Marking is total, meaning every well-formed expression can be marked, producing a well-typed marked expression.
+Once marked, the system layers on unification-based type inference to opportunistically fill type holes.
+The typing judgment for marked expressions generates constraints: <judgement format>
+<Say something about users interactively choosing from partial solutions when the constraints for a type hole are in conflict.>
 
@@
 \label{sub:unmarked-lang}
-First, we introduce here the syntax of the \emph{unmarked} language which comprises two mutually defined expression sorts: $Exp$ \& $ExpList$, denoted by $\EMV$ and $\ELV$ respectively as displayed in \autoref{fig:unmarked-language}.
+First, we introduce the syntax of the \emph{unmarked} language which comprises two mutually defined expression sorts: $Exp$ \& $ExpList$, denoted by $\EMV$ and $\ELV$ respectively as displayed in \autoref{fig:unmarked-language}.
 In addition, this unmarked language is an instantiation of the generic term syntax introduced in \autoref{sub:Groves}, with $\EMV$ \& $\TMV$ being the expression and type sorts along with their mutually defined list counterparts $\ELV$ \& $\TLMV$ that allow us to model holes (empty list), \emph{regular} singleton expressions, and local conflicts.
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
-    \caption{Type-checking Unmarked Expressions}
     \label{fig:unmarked-judgments}
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
     \caption{Unmarked Language}
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
     \caption{Marked Language}
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
     \caption{Other machinery}
@@
 Next, we introduce the heart of the marking machinery in \autoref{fig:marking-judgments} and \autoref{fig:marked-language}, again in the form of 4 mutually defined judgments along with the marked language denoted by $\ECMV$ \& $\ELMV$.
-However, here, we add error marks to the expression when one of the premises fails, and type-checking cannot continue further, thereby adding first-class support for error localization.
-In addition to marking the expressions and type-checking them, at this stage, we also collect inference constraints and \emph{local-info} sets, which come into play when the user has to deal with \emph{multi-location} and \emph{cycle} conflicts.
 
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
     \caption{Marking Expressions}
@@
 In \autoref{inf:lam-ana}, in the first case, we successfully mark the lambda abstraction by first \emph{graph-erasing} its type and then using the matched arrow judgment and consistency relations to recursively mark its body.
-While performing this marking, we gather inference constraints \& local-info maps from the matched arrow judgment and the recursive analysis of the lambda body.
-Here, all the premises hold, and the abstraction is analyzed with no issues.
+When all the premises hold, the abstraction is analyzed with no issues.
 What happens when the analyzed type is not a matched arrow type or the erased type is \emph{not} consistent with the expected input type?
-Conventional type-checkers fail here and cannot proceed.
-This is where the next rules help in localizing these scenarios to the node where the error occurs.
+This is where the next rules help localize these errors to the node where the failure occurs.
 In $MKACLam2$, the expected type is not a matched arrow type.
-So, we mark the abstraction as such using $\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}$ which captures the premise that failed while still analyzing the body against the unknown type and gathering constraints.
 Similarly, in $MKACLam3$, the consistency check between the erased ascription and input of the expected type fails, and again, we mark the abstraction as such, depicting the inconsistency in the ascription of the lambda in $\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}$.
@@
 \caption{Marking and Analyzing Lambda Abstractions}
-\inferrule[MKACLam1]{
+\inferrule[MKALam1]{
         \graphErase{\TLMV} = \sigma \\
-        \matchedArrowConstraint{\sigma_3}{\sigma_1}{\sigma_2}{\cConstraint_1} \\
+        \matchedArrow{\sigma_3}{\sigma_1}{\sigma_2} \\
         \consistent{\sigma}{\sigma_1} \\
-        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\sigma_2}{\cConstraint_2}{\mConstraint_1}
+        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\sigma_2}
     }{
-        \anaMarkConstraint{\ctx}{\lam{x}{\TLMV}{\EMV}}{\lam{x}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\sigma}{\sigma_1}}}{\mConstraint_1}
+        \anaMarkConstraint{\ctx}{\lam{x}{\TLMV}{\EMV}}{\lam{x}{\TLMV}{\ELMV}}{\sigma_3}
     }
 
-     \inferrule[MKACLam2]{
+     \inferrule[MKALam2]{
         \graphErase{\TLMV} = \sigma \\
         \notMatchedArrow{\sigma_3} \\
-        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\TUnknown^{anon}}{\cConstraint}{\mConstraint}
+        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\TUnknown}
     }{
-        \anaMarkConstraint{\ctx}{\reglam}{\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint \cup \Setof{\constrain{\TUnknown^{exp(u)}}{\sigma_3}}}{\mConstraint}
+        \anaMarkConstraint{\ctx}{\reglam}{\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}
     }
 
-    \inferrule[MKACLam3]{
+    \inferrule[MKALam3]{
         \erase{\TLMV} = \sigma \\
-        \matchedArrowConstraint{\sigma_3}{\sigma_1}{\sigma_2}{\cConstraint_1} \\
+        \matchedArrow{\sigma_3}{\sigma_1}{\sigma_2} \\
         \inconsistent{\sigma}{\sigma_1} \\
-        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma_1}}{\ELV}{\ELMV}{\sigma_2}{\cConstraint_2}{\mConstraint}
+        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma_1}}{\ELV}{\ELMV}{\sigma_2}
     }{
-        \anaMarkConstraint{\ctx}{\reglam}{\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\TUnknown^{exp(u)}}{\sigma_3}}}{\mConstraint}
+        \anaMarkConstraint{\ctx}{\reglam}{\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}
     }
@@
 \label{Marking and Synthesizing Function Applications}
-\inferrule[MKSCAp1]{
-    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma}{\cConstraint_1}{\mConstraint_1} \\
-    \matchedArrowConstraint{\sigma}{\sigma_1}{\sigma_2}{\cConstraint_2} \\
-    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\sigma_1}{\cConstraint_3}{\mConstraint_2} \\
+\inferrule[MKSAp1]{
+    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma} \\
+    \matchedArrow{\sigma}{\sigma_1}{\sigma_2}\\
+    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\sigma_1} \\
   }{
-    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\eApp{u}{\ELMV_1}{\ELMV_2}}{\sigma_2}{\cConstraint_1 \cup \cConstraint_2 \cup \cConstraint_3}{\mConstraint_1 \cup \mConstraint_2}
+    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\eApp{u}{\ELMV_1}{\ELMV_2}}{\sigma_2}
   }
 
-  \inferrule[MKSCAp2]{
-    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma}{\cConstraint_1}{\mConstraint_1} \\
+  \inferrule[MKSAp2]{
+    \synMarkConstraint{\ctx}{\ELV_1}{\ELMV_1}{\sigma} \\
     \notMatchedArrow{\sigma} \\
-    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\TUnknown^{\rightarrow_{L}{(exp(u))}}}{\cConstraint_2}{\mConstraint_2}
+    \anaMarkConstraint{\ctx}{\ELV_2}{\ELMV_2}{\TUnknown}
   }{
-    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\app{\ECApSynNonMatchedArrow{\ELMV_1}}{\ELMV_2}}{\TUnknown^{\rightarrow_{R}{(exp(u))}}}{\cConstraint_1 \cup \cConstraint_2 \cup \SetOf{\constrain{\TUnknown^{exp(u)}}{\tarr{\TUnknown^{\rightarrow_L(exp(u))}}{\TUnknown^{\rightarrow_R(exp(u))}}}}
-    }{\mConstraint_1 \cup \mConstraint_2}
+    \synMarkConstraint{\ctx}{\eApp{u}{\ELV_1}{\ELV_2}}{\app{\ECApSynNonMatchedArrow{\ELMV_1}}{\ELMV_2}}{\TUnknown^{\rightarrow_{R}{(exp(u))}}}
   }
@@
 \subsubsection{Local Conflicts}
-To mark local conflicts, we recursively mark all subterms, gather the constraint and local info sets, and synthesize the unknown type; such conflicts are treated essentially as holes.
+To mark local conflicts, we recursively mark all subterms, gather the constraint and location conflict contexts, and synthesize the unknown type; such conflicts are treated essentially as holes.
 
@@
 \subsubsection{Multi-Location Conflicts}
-
+\todo{fix rule names here}
 \[
@@
 
-\subsubsection{Cycle Conflicts}
+\subsubsection{Location Cycle Conflicts}
 
@@
 
-We observe that \
+We observe that both  \uniref and \multiref are instances of \emph{location} conflicts; they are terms that can be marked and type-checked in the various locations where they occur.
+At first, this can confuse the user when they encounter these new conflict constructs while programming collaboratively.
+As a result, we want them to be able to access these different instances \emph{and} their contexts to resolve them in an informed manner.
+This is precisely the motivation behind \emph{location conflict contexts} $(\L)$ which are sets of 4-tuples of vertex-IDs, edge-IDs, contexts, and mode denoted by $v$, $w$, \ctx, and $m$ respectively.
+The goal is to have an inspection window in the editor that gathers the relevant location when a user clicks on one of these location conflicts to display their other instances starting from the clicked context, allowing the user to explore them interactively to resolve the conflicts.
+However, we have left that to the UI considerations of future work while formalizing the machinery required.
+Otherwise, these conflicts are treated as holes during type-checking, with their provenance determined by the vertex-ID $(v)$ they carry.
 
@@
 Marked terms are implicitly typed, allowing local info mappings to be computed as a function of these terms.
+
+\subsection{Inference}
+\label{sub:inference}
+
+I don't know what to write here lol \ldots
 

### 2024-07-11 formalism.tex

@@
 
-A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is a constructor instance, where 
-    unique identifiers are drawn from $\U$ and 
-    constructors are drawn from $k \in \K$. 
+A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is a constructor instance, where unique identifiers are drawn from $\U$ and constructors are drawn from $k \in \K$. 
 The \emph{inedges} of $v$ are the edges that target $v$.
@@
 
-A \emph{source} is a pair $(v, p) \in \S = \V \times \P$, where
-  vertices are drawn from $\V$,
-  and positions are drawn from $\P$.
-The \emph{outedges} of a source are the edges that have the source.
+A \emph{location} $\ell = (v, p) \in \L = \V \times \P$ represents the origin of an edge, where vertices are drawn from $\V$ and positions are drawn from $\P$.
 
-An \emph{edge} $\e{=}(w, (v, p), v^\prime) \in \E = \W \times \S \times \V$ represents a directed multi-edge 
+An \emph{edge} $\e{=}(w, \ell, v^\prime) \in \E = \W \times \L \times \V$ represents a directed multi-edge 
     identified by $w \in \W$, 
-    originating from source $(v, p) \in \S$, 
+    originating from location $\ell \in \L$, 
     with destination vertex $v' \in \V$.
@@
 
-Given a source language parameterized by a set of constructors $k \in \K$ and an arity function $\K \rightarrow \mathbb{N}$, we define an augmented term language as specified in \autoref{fig:Syntax}.
+Given a source language parameterized by a set of constructors $k \in \K$, each associated with a fintie set of positions $arity(k)\subseteq \P$, we define an augmented term language as specified in \autoref{fig:Syntax}.
 Terms are extended with cases for multi-location references and unicycle references.
-Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define a secondary sort \textit{SubTerm} to represent these possibilities.
-A source with no outgoing edges corresponds to an empty hole, a source with one outgoing edge corresponds to an ordinary subterm, and a source with multiple outgoing edges corresponds to a local conflict between multiple terms.
+Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define a secondary sort \textit{ChildTerm} to represent these possibilities.
+A location with no outgoing edges corresponds to an empty hole, a location with one outgoing edge corresponds to an ordinary subterm, and a location with multiple outgoing edges corresponds to a local conflict between multiple terms.
 
@@
 To support type hole inference (discussion in Sec.
-TODO), each empty hole and local conflict carries its source.
+TODO), each empty hole and local conflict carries its location.
 Finally, to support user navigation of the grove, each reference also carries the UID of the edge that leads to it.
@@
 
-\begin{align}
-    \text{parents}_{G}(v) &= \SetOf{v' \in \V \SuchThat{(w, (v', p), v) \in G \land w \in \W \land p \in \P}}
-\end{align}
+\begin{align*}
+    \text{parents}_{G}(v) &= \SetOf{v' \in \V \SuchThat{(w, (v', p), v) \in G \land w \in \W \land p \in \P}}\\
+    \text{children}_{G}(\ell) &= \SetOf{(w,v) \in \W\times\V \SuchThat{(w, \ell, v) \in G}}
+\end{align*}
 
@@
 
+\begin{definition}
+    $\text{root}_G(v)$ if $v$ is an NP root, an MP root, or a U root in $G$.
+\end{definition}
+
 The decomposition of a vertex is specified intuitively as the term obtained by traversing the descendants of the vertex until a leaf or a root vertex of some kind is reached, which not further decomposed, but is denoted in the term as the appropriate reference.
-The decomposition of a graph is simply the set obtained by decomposing each vertex classified as a root.
+The decomposition of a graph is simply the set obtained by decomposing each vertex classified as a root, provided the vertex actually appears as the origin of an edge in $G$.
 
 \begin{definition}
-    $\text{decomp}_{G}(v) = $
+\begin{align*}
+    \text{decompChild}_G(w, v)  &= \begin{cases}
+        \multiref  &\quad v\text{ is an MP root in } G\\
+        \uniref  &\quad v\text{ is a U root in} G\\
+        \text{decompVertex}_{G}(v) &\quad v\text{ is not a root in} G
+    \end{cases} 
+\end{align*}
+\end{definition}
+
+\begin{definition}
+\begin{align*}
+    \text{decompLocation}_{G}(\ell) &= \begin{cases}
+        \ehole &\quad \text{children}_{G}(\ell) = \emptyset \\ \eid{w}{}{\text{decompChild}_G(w, v')} &\quad \text{children}_{G}(\ell) = \SetOf{(w,v')}\\ \conflictHoleTerm{\ell}{\eid{w}{}{\text{decompChild}_G(w, v')} \SuchThat{(w,v') \in \text{children}_{G}(\ell)}} &\quad \abs{\text{children}_{G}(\ell)} > 1 \\
+    \end{cases}
+\end{align*}
+\end{definition}
+
+\begin{definition}
+    $\text{decompVertex}_{G}(v = (u,k)) = \genericTerm{u}{\text{decompLocation}_{G}((v,p))}$
+\end{definition}
+
+\begin{definition}
+    \text{vertices}(G) = \SetOf{v \in \V \SuchThat{(w, (v,p), v') \in G \land w \in \W \land v' \in \V \land p \in \P}}
+\end{definition}
+
+\begin{definition}
+    $\text{decomp}(G) = \SetOf{\text{decompVertex}_{G}(v) \SuchThat{\text{root}_G(v)} \land v \in \text{vertices}(G)}$
 \end{definition}
@@
 Terms store vertices on constructors and edge UIDs for each child to enable the completeness of this reconstruction.
+The recomposition of a set of terms is the union of the recompositions of the terms.
+
+\begin{definition}
+    \begin{align*}
+    \text{vertexOfTerm}(\genericTerm{u}{\subtermMV_p}) &= (u,k)\\
+    \text{vertexOfTerm}(\multiref) &= v\\
+    \text{vertexOfTerm}(\uniref) &= v
+    \end{align*}
+\end{definition}
+
+\begin{definition}
+    \begin{align*}
+    \text{recompChildTerm}(\ell, \ehole) &= \emptyset\\
+    \text{recompChildTerm}(\ell, \eid{w}{}\termMV) &= \SetOf{(w, \ell, \text{vertexOfTerm}(\termMV))}\cup \text{recompTerm}(t)\\
+    \text{recompChildTerm}(\ell, \conflictHoleTerm{\ell}{\eid{w_i}{}\termMV_i}_{i<n}) &= \bigcup_{i<n} (\SetOf{(w_i, \ell, \text{vertexOfTerm}(\termMV_i))}\cup \text{recompTerm}(t_i))
+    \end{align*}
+\end{definition}
+
+\begin{definition}
+    \begin{align*}
+        \text{recompTerm}(\genericTerm{u}{\subtermMV_p}) &= \bigcup_{p\in arity(k)}\text{recompChildTerm}(((u,k),p), \subtermMV_p)\\
+        \text{recompTerm}(\multiref) &= \emptyset\\
+        \text{recompTerm}(\uniref) &= \emptyset
+    \end{align*}
+\end{definition}
+
+\begin{definition}
+    $\text{recomp}(T) = \SetOf{\text{recompTerm}(\termMV) \SuchThat{\termMV \in T}}$
+\end{definition}
 
 \begin{theorem}
-    (Recomposability) TODO
+    (Recomposability) For all subgraphs $G$, $\text{recomp}(\text{decomp}(G))=G$.
 \end{theorem}
+
+\subsubsection{Grove Construction}
+
+In order to facilitate user interaction with the term set that results from decomposition, it is refined into a \textit{grove}.
+A grove is a partition of the term set into three classes, corresponding to the the three kinds of roots: NP, MP, and U.
+A grove also designates a distinguished child term as the primary component of the program.
+
+Formally, a grove $\gamma$ is a quadruple $(\subtermMV_r, NP, MP, U)$ where $\subtermMV \in ChildTerm$ and $NP, MP, U \subseteq Term$.
+The construction of a grove from the term set of $G$ requires the designation of a location $(v_r,p_r)$ as the \textit{distinguished root} such that $v_r$ is an NP root in $G$.
+In that case the grove is defined as follows:
+
+\begin{definition}
+\begin{align*}
+    \text{distinguishedTerm}(\ell_r=((u_r,k_r),p_r), \Theta) &= \subtermMV_r &\quad \text{ if }
+    \genericConstructorTerm{u_r}{k_r}{\subtermMV_r}\in \Theta \\
+    \text{distinguishedTerm}(\ell_r=((u_r,k_r),p_r), \Theta) &= \emptyHole{\ell_r} &\quad \text{ otherwise}
+\end{align*}
+\end{definition}
 
@@
 $\Reposition{v}{p}$ combines edge construction and deletion into a single \emph{atomic} operation.
-If the cursor is not an empty hole, the corresponding edge(s), as indicated above, are simultaneously deleted and then reconstructed as child(ren) with edge source $(v,p)$.
+If the cursor is not an empty hole, the corresponding edge(s), as indicated above, are simultaneously deleted and then reconstructed as child(ren) with edge location $(v,p)$.
 

### 2024-07-11 grove-by-example.tex

@@
 
-Although the collaborative edits discussed so far merge cleanly,
-they can also lead to graphs that do not map cleanly to a syntax tree.
-We call these states \emph{conflicts}
-and consider separate cases for \emph{multi-child}, \emph{multi-parent}, and \emph{unicycle} conflicts.
+The collaborative edits discussed so far merge cleanly,
+but in general, merging patches can lead to graphs that do not map cleanly to a conventional syntax tree.
+We identify several different motifs that might arise, all of which give rise to different kinds of conflicts in the graph decomposition: \emph{local conflicts}, \emph{multi-location conflicts}, and \emph{cyclic location conflicts}.
 As with merge conflicts in version-control systems such as git, these all require user intervention to resolve.
-We consider common merge scenarios first, then move onto more unusual scenarios.
 
@@
 
-Suppose Alice and Bob both have the edit state \texttt{w * v + \hole} in \autoref{fig:NestedParts:c}.
+Suppose Alice and Bob both start with the edit state \texttt{w * v + \hole} from \autoref{fig:NestedParts:c}.
 Alice moves the cursor to the hole and constructs~\vMultiChildAlice{} as the~\texttt{R} child of~\vWrapPlus{}.
-At the same time, Bob constructs~\vMultiChildBob{} in the same position.
+At the same time, Bob constructs~\vMultiChildBob{} at the same location.
 Now Alice and Bob have the graphs in \autoref{fig:MultiChild:a} and \autoref{fig:MultiChild:b}, respectively.
-Note that these graphs have two different edges,
-\eMultiChildAlice{} and \eMultiChildBob{}, for the~\texttt{R} position of~\vWrapPlus{}.
 
-We merge these edits by including \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
-as edges originating from the \texttt{R} position of \vWrapPlus{}.
-This results in a multi-child conflict in the~\texttt{R} position of~\vWrapPlus,
-shown in \autoref{fig:MultiChild:c},
-where there are multiple edges originating from the same position of a vertex.
-In the user interface, we flag this conflict with the notation~$\conflictHole{x,y}$.
+When these patches are merged, \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
+appear in the merged graph. 
+When decomposing this graph to a syntax tree, 
+we resolve this conflict in the~\texttt{R} position of~\vWrapPlus~by decomposing to a \emph{local conflict}, \figureCode{$\binaryConflictHole{x}{y}$}, as 
+shown in \autoref{fig:MultiChild:c}.
 
@@
 
-Such conflicts can be resolved by deleting one of the conflicting edges
-and editing the remaining term into the correctly merged value.
-For example, Alice could delete \eMultiChildBob{} and then change~\vMultiChildAlice{} into~\texttt{x * y}.
+Such conflicts can be resolved simply by deleting relocating one of the conflicting terms (and editing the remaining term into the correctly merged value, if needed), which would remove the corresponding edge.
+For example, Alice could resolve the problem by wrapping \li{x} and \li{y} with a multiplication, effectively moving them to non-conflicting locations.
+No special edit actions are needed for conflict resolution.
+
 % To ensure users can continue editing while in a conflicted state,
@@
 % but we consider this a higher-level, user-interface consideration.
-Handling conflicts in this way is similar to handling traditional merge conflicts
-in that it presents alternate versions of the conflicted state
-that must be transformed by users into the desired result.
-Our approach is different in that merge conflicts in a line-based model
-may break semantic editor services, such as semantic search or type-preserving refactorings,
-that could continue to work inside and outside the conflict in a graph-based model.
-% It is just the where the multi-child happens that editor services may break down.
-\autoref{sec:Related Work} discusses more differences.
+
+It is worth noting one special situation: if Alice and Bob both independently filled the hole with structurally identical terms, e.g.
+\li{x}, Grove would \emph{still} formally identify a conflict, because each \li{x} has a distinct UID.
+In this situation, it would be reasonable for the system to resolve the conflict without further coordination by deterministically choosing one of the two UIDs, e.g.
+the smallest one.
+
+When the conflicted terms are similar up to UID differences but not identical, it might be helpful  to give the developer the option to ``push down'' the conflicts as deeply as possible, using a tree differencing algorithm.
+However, this would increase the number of conflicts overall, so it may not always be preferable.
+Tree differencing is not fundamental to the collaboration model of Grove.
+
+As we will return to in \autoref{sec:Type System}\todo{sec}, 
+a local conflict need not disable type-based editor services (the \textbf{semantic gap problem}).
+Instead, the system can type check around the conflict (and if the conflicted terms have a common type, the system can infer that type for the location conflict itself).
 

### 2024-10-25 review/response.md
Updates from Overleaf

Thank you for your careful readings and thoughtful comments! We are heartened by the encouraging comments and strong overall assessment from RA and RB.

We do not believe RC has identified any significant technical deficiencies in the work. We hope the clarifications we provide below will help illuminate some subtleties regarding our claimed contributions and the technical developments in the paper. We hope that this will lead to a more positive final assessment from RC as well.

# Review A

We agree with the detailed comments and will address them in the final version of the paper.

# Review B

> My only (very minor) complaint is that as somewhat of an outsider to this space, I would have appreciated a longer and more detailed related works section, just to understand more about how the core ideas fit into the broader space. This is not a demand for a longer Related Works–I think the section is clearly already sufficient for acceptance. I only mention this because I think the paper could make a stronger case for its core ideas by offering a deeper analysis of the research space, especially for readers who are not already familiar with the relevant areas.

We agree, and we will fortify our discussion of related works in the final version of the paper. We intend to add further discussion of the literature on CRDTs and on patch synthesis approaches in version control systems in particular.

# Review C

> For the first 18 pages, the paper develops a graph representation for representing
edits and various kinds of conflicts.

It is indeed the case that the central technical idea of the paper is a novel graph representation of collaboratively edited programs (L104), a corresponding patch calculus with commutative edits (L110), and a corresponding invertible decomposition into a grove (L116).

> The graph representation is essentially a different way of representing textual information by using special nodes and edge/node relations to represent conflicts.

Note that the graph representation captures information that cannot be represented in ordinary program text - namely, a subterm occurring in multiple places, multiple subterms in the same location, and a term containing itself as a subterm. To this end, the graph includes unique identifiers of vertices and edges, which are not contained in a typical textual representation. Our decomposition of the graph into groves, which may be presented "textually" to the user as in our implementation, was carefully structured to retain this information, as we prove in Agda.

> Compared to Hazelnut and Hazel, the only difference is that this paper assigns all conflicts to have the Dyn type from gradual typing. ... The idea of using Dyn types were already introduced in Hazel, so there is very little novelty here.

Section 5 on type checking was intended to be a "cherry on top" of the main contributions listed above. It is indeed the case that the type system is based directly on the marked lambda calculus, which is the core calculus of Hazel. However, the contribution of Section 5 is a bit more nuanced than is stated in this review:

- Local conflicts in synthetic position synthesize the dynamic type and generate constraints between this type and the synthesized types of the contents.
- Relocation and unicycle conflicts in synthetic position synthesize the dynamic type with provenance determined by the identifier they refer to. 
- Local conflicts in analytic positions analyze their contents against the same analyzed type.
- Relocation and unicycle conflicts in analytic position generate a constraint between the analyzed type and the dynamic type with provenance determined by the identifier they refer to.

These constraints allow the integration of typing information from all references to the same identifier.

> The paper keeps on saying that "Grove does not suffer from problems due to merge conflicts". However, the only reason this is true is that Grove represents/encapsulates these conflicts. It does not solve any of the merging problems.

The paper does not claim that "Grove does not suffer from problems due to merge conflicts." On line 87, it claims that Grove "does not suffer from the problems just outlined," those problems being the granularity problem (L46), the relocation conflict problem (L52), and the relocation modification problem (L56). Each problem is addressed in Grove as described in Sec. 2.

The paper also states that "The Grove patch language is commutative, meaning that there is no need for a complex three-way merge algorithm (i.e. operational transform)" (L288). Again, this is not a claim that there are no problems due to merge conflicts, but rather that we do not need to rely on that particularly complex kind of algorithm to discover the merge conflicts.

We can see how these statements might not be as clear as possible in emphasizing that Grove does indeed represent/encapsulate merge conflicts without resolving them on the user's behalf fully automatically, and will expand on this in both the introduction and conclusion of the final version of the paper so that readers are less likely to misunderstand the claim.

We believe that automatically resolving merge conflicts would require ad-hoc heuristics and inhibit user agency. Instead, Grove provides a setting in which those unresolved merge conflicts are syntactically and semantically meaningful. In particular, conflicts do not block edits from being made and shared with collaborators, and conflicts do not block the program's semantics. Previous version control systems require merge conflicts to be resolved before editor services may be available again, and Grove eliminates this gap problem by supporting editor services at all times.

Automatic heuristics could be layered on top of Grove as a convenience, and indeed we mention a few possible situations where this might be helpful in the paper (e.g. in the paragraph on L390 and the next paragraph on L415).

Details:

> Section 2.3.2 It is well known that tree-based editors handle the granularity problem better than line-based. But the tree-based editor has a much higher complexity, building the tree and maintaining the correct tree structure during edits such as cut and paste.

Hazel is a structured editor, so the tree structure of the program is maintained natively in the editor. It's true that this is more complex than just maintaining a string, but this complexity is accepted and handled in other aspects of Hazel's theory and implementation (e.g. Omar et al., 2017, Moon et al., 2023). There are a number of successful structure editors in popular usage, e.g. Scratch.

> Section 2.3.3 Grove chooses to silently merge changes from relocation modification, which current versioning control systems often yield a merge conflict. However, it is hard to say that this behavior is always wanted or preferred. In many scenarios, the involved programmers may want to review before such merges are committed. Also, this example seems to work well because the operations involved, + and *, are commutative. What if we have operations like - and /?

The key point here is that Alice's and Bob's edits are not in conflict with each other - one is modifying an expression, and the other is modifying the location of that expression. We posit that in practice, such actions are morally independent, even though they affect some of the same parts of the tree. We acknowledge, however, that in some cases like these it may be helpful to notify the users of the proximity of the edits. Such a notification system would not affect the underlying theory. We will note this possibility in the next version of the submission.

When two edits are truly in conflict, meaning there is no way to apply them both and maintain a correct syntax tree, then Grove 'merges' them, but not silently - it results in a syntactically marked conflict.

The commutativity of the operator here does not matter -- Grove operates purely syntactically.

> Figure 10(d), why not add an L edge from +8 to *2 and a R edge from +8 to *24? Also, why R31 and not R35 and L33 not L37?

The R edge from +8 to *24 is not kept in 10(d) for the same reason that it is not kept in 10(b) - because Alice's action is to relocate *24, which involves both adding an edge and deleting an edge. When this same action is played back on 10(c), the R27 edge is deleted. Playing back all of the merged actions ensures commutativity. The edge names are different for a similar reason - the actions depicted delete and insert edges, so the final edges are new and have new identifiers.

> What is the connection between Line 367 and Line 531? Earlier it says a disconnected term may be connected back after a deletion but later it says deleted edges can't connect again. Also, since an edge includes a term, the deletion of the edge also causes the deletion of the term if it is connected to by only one edge.

Deleting an edge does not delete any vertices. A term may be disconnected by deleting an edge and may be reconnected by inserting a new edge with the same origin and destination but a different identifier. The old edge remains deleted, but the term is reconnected. An edge does not include a term - it refers to an origin location and destination vertex.

> In Rule MKSLocalConflict, it is possible to synthesize a more precise type than ?. A more meaningful choice is the least upper bound of all σi's in terms of some subtyping relation. For example, if all branches of a conflict have Int, then it is fine to say that the conflict has the type Int.

This is true, and we considered taking this approach. We decided not to because it is more complete and modular to defer to the type hole inference system by emitting constraints that each synthesized type from the contents must be equal to the synthesized gradual type (and hence to each other). This allows the type hole inference system to indeed infer Int if they all have type Int, as well as to infer more complete information than the least upper bound would.

> In rules MALocalConflict, MARelationConflict, and MAUnicycleConflict, why do you generate a constraint between a ? and σ? In gradual typing, such constraints are simply dropped. Probably you want to use the location for provenance. However, it is unclear how the location associated with each ? is used.

Indeed, such constraints are not dropped, and indeed we use the location for provenance. This section extends the work on the marked lambda calculus and type hole inference presented in Zhao et al., 2024. This prior work describes the details of how the provenances and constraints are solved and used. To summarize, constraints are solved with each instance of the gradual type from a specific provenance as a metavariable. This allows constraints on the same provenance originating at different points in the bidirectional flow to be combined to form a solution. If solved, this solution is returned to the bidirectional typing flow, and if conflicts are found, the possible solutions for type holes in the surface syntax are offered to the user as an editing affordance.

As it happens, the constraint generated in the rule MALocalConflict is actually redundant because the hole with provenance $\exp(\ell)$ cannot appear in any other constraints generated by the program. This unnecessary constraint will be removed in the final version of the paper. We will also streamline the rules by altering MSLocalConflict to analyze the contents against the synthesized gradual type rather than generating constraints directly. Neither of these is an issue of correctness, but each represents an improvement to the presentation.

Typos:

> Line 163, what does "this work" refer to? Probably it refers to Hazelnut and Hazel, but it is good to be explicit

"This work" refers to "the previous work on the Hazelnut structure editor calculus" (L139), and more specifically, "a type system and type error localization system (the marked lambda calculus) for incomplete programs" (L140) mentioned in the previous paragraph. We will make this reference more explicit in the next version of the submission.

> Lemma 3.8, I am not sure if this is actually correct. For example \*24 in Figure 9(d) is both an MP and a U root.

Lemma 3.8 is correct. \*24 in Figure 9(d) is an MP root, but it is not a U root because there is no nontrivial sequence of vertices that begins and ends at \*24 such that each vertex is the *unique* parent of the previous. \*24 does not have a unique parent - it has multiple parents.

### 2024-10-25 OriginalSubmissionSource/abstract.tex
Updates from Overleaf

--- 2024-07-11 abstract.tex
@@
 vertices and edges. All edits amount to edge insertion and deletion, with deletion being permanent. 
-To support tree-based editing, we define a decomposition from graphs into \emph{groves}, which are a set of syntax trees with conflicts---including multi-child, multi-parent, and cycle conflicts---represented explicitly using holes and references between trees.
+To support tree-based editing, we define a decomposition from graphs into \emph{groves}, which are a set of syntax trees with conflicts---including local, relocation, and unicyclic relocation conflicts--- represented explicitly using holes and references between trees.
 Finally, we define a type error localization system for groves that enjoys a \emph{totality} property, i.e.

### 2024-10-25 OriginalSubmissionSource/discussion-and-conclusion.tex
Updates from Overleaf

\section{Discussion and Conclusion}%
\label{sec:Discussion and Conclusion}

This paper proposes a radically simpler, albeit practically ambitious, rearchitecture of collaborative editing. Our contributions together result in a typed collaborative structure calculus called Grove 
where, uniquely, all edits, including code relocations that stymie existing approaches, commute and where there are no semantic gaps: all possible editor states, including editor states with various kinds of unresolved conflicts, are semantically meaningful.

This paper focuses on the core theoretical underpinnings of this approach,
developing mechanized metatheory for both the patch language and the type system. 
A number of research problems on algorithmic, networking, and user interface aspects of the problem open up given these foundations. For example, we point out several situations where presenting non-conflicted but heuristically attention-worthy merges may be worthwhile, and we leave to future work the user experience design of this process. We intend to use the Grove workbench to integrate these efforts into the Hazel programming environment, though its editor component has been evolving so rapidly as to prevent experimentation in this direction so far.

Although our focus was on tree editing, some aspects of a program are more naturally linearly structured, e.g. string literals. We also leave to future work the problem of combining existing work on sequence CRDTs~\cite{ahmed2011evaluating,kleppmann2020moving} with our work on tree/graph CRDTs.

% In practice, it may be helpful to garbage collect orphaned vertices once there is consensus across collaborators that the deletion is permanent, but we do not consider this consensus protocol formally in this paper. In an open-ended collaboration scenario (where the set of collaborators is not known, e.g. on GitHub), we simply retain all vertices.

% TODO: talk about linear sequences in code

% TODO: leaves

% TODO: our model supports treating these as a cons-list of characters

% TODO: (Place somewhere) This move semantics gives us a richer structure than
% when treating code as a list of lines.  We exploit
% this in Section~REF:TODO in order merge edits that involve moving code.

% TODO: Traditional diff: no relation (and indent might change)

% TODO: memory usage

% TODO: cache eviction algorithm

% TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
% In practice, these complex scenarios occur less frequently than portrayed here.

% \subsection{Variable names, strings, and numbers}%
% \label{sub:Variable names, strings, and numbers}

% NOTE: we could implement each digit as a separate characters

% GUI for string conflicts: use popups

### 2024-10-25 OriginalSubmissionSource/formalism.tex
Updates from Overleaf

\section{The Grove Calculus}%
\label{sec:Formalism}

We now formally consider the ideas presented informally above by defining the Grove Calculus. The calculus consists of a commutative patch language on graphs, defined in \autoref{sub:Convergent Graphs}, a decomposition procedure to go from graphs to groves (i.e. sets of terms with conflicts) in \autoref{sub:Groves}, and a user action language that operates on groves and is defined by translation to the patch language in \autoref{sub:User Actions}.

\subsection{A Commutative Graph Patch Language}%
\label{sub:Convergent Graphs}

\subsubsection{Graphs}
Let $\U$ and $\W$ be separate sets of \emph{unique identifiers} (UIDs) for vertices and edges, respectively. Assume each is equipped with a total ordering $\leq$.

Let $\K$ be the set of \emph{constructors} of terms in the source language, and let $\P$ be the set of \emph{positions} of subterms in the source language. Assume each constructor $k\in\K$ is associated with a finite set of positions $\mathsf{arity}(k)\subseteq \P$, representing the set of positions that children of $k$ can inhabit.

A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is labeled with a UID, $u$, and a constructor, $k$.

A \emph{location} $\ell = (v, p) \in \L$, represents the origin of an edge, where $v=(u,k)$ is drawn from $\V$ and position $p$ is drawn from $\mathsf{arity}(k)$.

An \emph{edge} $\e{=}(w, \ell, v) \in \E = \W \times \L \times \V$ represents a directed multi-edge 
    identified by $w$, 
    originating from location $\ell$, 
    with destination vertex $v$.

A \emph{graph} $g : \E \rightarrow \Sigma$ is a function from edges to \emph{edge states} $\Sigma = \SetOf{\bot, \Plus, \Minus}$. If $g(\e) = \bot$, then $\e$ has not yet been constructed in the graph. If $g(\e) = \Plus$ then $\e$ is \textit{live} in the graph, meaning it is relevant to the user and the type system. If $g(\e) = \Minus$, then $\e$ has been deleted and cannot be constructed again. The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$. We assume a graph assigns only finitely many edges to $\Plus$ to ensure the finiteness of the data structure on which further computation is done.

\subsubsection{Graph Patch Commands}
\label{subsub:Graph Patches}

We define a graph patch command $\pi = (s,\e) \in \left\{\Plus, \Minus\right\} \times \E$
as a pair of an edge state (excluding $\bot$) and an edge. As a shorthand, we define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.

For $s_1$, $s_2 \in \Sigma$, we define the join operation $s_1 \sqcup s_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus \sqcup \bot = \Plus$ means that edges can only be created if they do not already exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once an edge is deleted it can never be restored, and hence deletion is idempotent.

We define the semantics of patch commands via the following transition relation
between graphs.
\[
  g \overset{(s,\e)}{\longrightarrow} g\left[ \e \mapsto s \sqcup g(\e) \right]
\]
Applying patch command $\pi = \left(s, \e\right)$ to graph $g$ results in the updated graph $g' = g\left[\e \mapsto s \sqcup g(\e) \right]$, wherein the edge
state associated with edge $\e$ in $g$ becomes the join of $s$ with the state of $\e$ in $g$, and $g'(\e') = g(\e')$ for any $\e' \ne \e$. Essentially, these patch command semantics act as a transition system between graphs, \emph{joining} the new edge state with the corresponding edge state of the graph to which the patch command is being applied. Since a patch command can only change the value of $g$ on one input, applying the patch command preserves the property that $g$ maps only a finite set of edges to $\Plus$.

A \textit{graph patch} $\Q{\pi}$ is a sequence of graph patch commands, and the transition relation is extended to patches as the composition of the transitions of the constituent patch commands.

\subsubsection{Commutativity}%
\label{subsub:Commutativity:formal}

Using these definitions, commutativity of patch commands can be established using the commutativity and associativity of the join operation on edge states.

\begin{lemma}[Join Semilattice]
  \label{lem:Join Semilattice}
  $(\Sigma, \sqsubset)$ with $\sqcup$ forms a join semilattice. That is, $\sqcup$ is associative, commutative, and idempotent. 
\end{lemma}
\begin{proof}
    Follows directly from the total ordering $(\Sigma, \sqsubset)$
\end{proof}

\begin{theorem}[Commutativity]
  \label{thm:Commutativity}
  For all graphs $g$ and $g'$ and all graph patch commands $\pi_1 = (s_1, \e_1)$ and  $\pi_2 = (s_2, \e_2)$, 
  if $g \overset{\pi_1 \pi_2}{\longrightarrow} g'$,
  then $g \overset{\pi_2 \pi_1}{\longrightarrow} g'$
\end{theorem}
\begin{proof}
  If $g \overset{\pi_1 \pi_2}{\longrightarrow} g'$, then $g' = g[\e_1 \mapsto s_1 \sqcup g(\e_1)][\e_2 \mapsto s_2 \sqcup g(\e_2)]$. We want to show that this equals $g'' = g[\e_1 \mapsto s_1 \sqcup g(\e_1)][\e_2 \mapsto s_2 \sqcup g(\e_2)]$. For any $\e$, if $\e \neq \e_1$ and $\e \neq \e_2$, then $g'(\e) = g''(\e) = g(\e)$. If $\e = \e_1$ but $\e\neq \e_2$, then $g'(\e) = g''(\e) = s_1 \sqcup g(\e)$. Similarly if $\e=\e_2$ but not $\e_1$. Finally, if $\e=\e_1=\e_2$, then $g'(\e)=s_2\sqcup(s_1\sqcup g(\e))$, and $g''(\e)=s_1\sqcup(s_2\sqcup g(\e))$. These are equal by \autoref{lem:Join Semilattice}. Since $g'(\e) = g''(\e)$ for all $\e$, $g' = g''$ and  $g \overset{\pi_2 \pi_1}{\longrightarrow} g'$.
  % $ $\e_1 \ne \e_2$, we have $g'(\e_1) = s_1 \sqcup g(\e_1)$ and $g'(\e_2) = s_2 \sqcup g(\e_2)$, and if $\e_1 = \e_2 = \e$, we have $g'(\e) = s_1 \sqcup s_2 \sqcup g(\e)$. Hence, for any edge \e, if $\e_1$ and $\e_2$ are not equal we have $g'(\e) = g[\e_1 \mapsto s_1 \sqcup g(\e_1)][\e_2 \mapsto s_2 \sqcup g(\e_2)](\e) = g[\e_2 \mapsto s_2 \sqcup g(\e_2)][\e_1 \mapsto s_1 \sqcup g(\e_1)](\e)$, and otherwise by \autoref{lem:Join Commutativity}, we have $g'(\e) = s_1 \sqcup s_2 \sqcup g(\e) = s_2 \sqcup s_1 \sqcup g(\e)$
\end{proof}

The commutativity of patch commands generalizes directly to the commutativity of patches.

\subsubsection{Interpretation as a CmRDT}

The definition above can be understood operationally as a CmRDT. In particular, the vertices can be understood as forming a Grow-Only Set (G-Set). The edges can be understood as forming a two-phase set (2P-Set), because deletion is permanent~\cite{shapiro2011conflict}.

\subsection{Groves}%
\label{sub:Groves}

Structure editors and other tools operate on trees, not on arbitrary graphs. When performing these operations on a graph $g$, only the set of edges $\e$ such that $g(\e) = \Plus$ should be considered. We call this the \textit{live subgraph} of $g$.

We present \textit{groves}, a tree-based representation of a live subgraph. A grove is a collection of mutually referential terms annotated with vertex and edge UIDs governed by the grammar in \autoref{fig:Syntax}. Groves correspond precisely to live subgraphs in the sense that a live subgraph can be \emph{decomposed} into a grove, and the grove can be \emph{recomposed} into the original live graph.

\subsubsection{Terms}%
\label{subsub:Terms}

\figureTermSyntax

Given a source language parameterized by a set of constructors $k \in \K$, we define an augmented term language as specified in \autoref{fig:Syntax}. Terms are extended with cases for relocation conflict references and unicycle conflict references. Additionally, since any position of any constructor may have zero, one, or multiple outgoing edges, we define a secondary sort \textit{ChildTerm} to represent these possibilities. A location with no outgoing edges corresponds to an empty hole, a location with one outgoing edge corresponds to an ordinary subterm, and a location with multiple outgoing edges corresponds to a local conflict between multiple terms.

These terms also carry information to support additional operations. To enable the reconstruction of the original graph, each constructor carries the UID of its original vertex, each reference carries the vertex that it refers to, and each subterm carries the UID of the edge that leads to it. To support type hole inference (as discussed in \autoref{sub:marking-groves}), each empty hole and local conflict carries its location. Finally, to support user navigation of the grove, each reference also carries the UID of the edge that leads to it.

\subsubsection{Graph Decomposition}%
\label{subsub:Graph Decomposition}

\figureDecompExample

A live subgraph $G$, which is just a finite set of edges, decomposes into a grove $\Theta$, which is a finite set of terms. In a tree, each node except the root has a unique parent, and there are no cycles. There is no guarantee of these properties in a graph. Decomposition operates by allowing each `problematic' vertex, in the sense of having multiple parents or forming a cycle, to be the root of its own entry in the grove, and to use references at special term leaves to encode the original graph structure.

Formally, the parents of a vertex $v$ in a graph is defined as the set of origin vertices $v'$ such that the graph includes an edge from $v'$ to $v$. Conversely, the children of a location $\ell$ is the set of $v$ such that the graph includes an edge from $\ell$ to $v$. It is convenient to consider these children to be paired with its edge's UID .

\begin{definition}
\begin{align*}
    \mathsf{parents}_{G}(v) &= \SetOf{v' \in \V \SuchThat{(w, (v', p), v) \in G \land w \in \W \land p \in \P}}\\
    \mathsf{children}_{G}(\ell) &= \SetOf{(w,v) \in \W\times\V \SuchThat{(w, \ell, v) \in G}}
\end{align*}
\end{definition}

To help define decomposittion, each vertex in a graph can be classified as either an NP root (if it has no parents), an MP root (if it has multiple parents), a U root (if there is a sequence of unique parents from it to itself, and it has the minimal UID in that sequence), or not a root. These classes are mutually exclusive.

\begin{definition}
    A vertex is an \textit{NP root} in $G$ if $\mathsf{parents}_{G}(v)=\emptyset$
\end{definition}

\begin{definition}
    A vertex is an \textit{MP root} in $G$ if $\abs{\mathsf{parents}_{G}(v)} > 1$
\end{definition}

\begin{definition}
    A vertex is a \textit{U root} in $G$ if $\exists \{v_i\}_{1 \leq i \leq n}$ such that $v_1 = v_n = v$, $\mathsf{parents}_{G}(v_i)=\{v_{i+1}\}$ for all $1 \leq i < n$, and the UID of $v$ is the minimum of the UIDs of $\{v_i\}_{1 \leq i \leq n}$.
\end{definition}

\begin{definition}
    $\mathsf{root}_G(v)$ if $v$ is an NP root, an MP root, or a U root in $G$.
\end{definition}

\begin{lemma}
\label{lem:uniq-classification}
    (Unique Classification) For any live subgraph $G\subseteq \E$ and vertex $v\in V$, exactly one of the following holds: 
    \begin{itemize}
        \item $v$ is an NP root in $G$
        \item $v$ is an MP root in $G$
        \item $v$ is an U root in $G$
        \item $v$ is not a root in $G$
    \end{itemize}
\end{lemma}

The decomposition of a vertex is specified intuitively as the term obtained by traversing the descendants of the vertex until a leaf or a root vertex of some kind is reached, which is not further decomposed, but is denoted in the term as the appropriate reference. The decomposition of a graph is simply the set obtained by decomposing each vertex classified as a root, provided the vertex actually appears as the origin of an edge in $G$.

Formally, the decomposition of a vertex is given by the function $\mathsf{decompVertex}$ a term with the same UID and constructor as the vertex. At each position in the arity of the constructor, we obtain a child term from the function $\mathsf{decompLoc}$, which decomposes a location in the graph to a corresponding child term. Consider the multiplication constructor with id 40 in \autoref{fig:Decomposition example}. It decomposes to a multiplication term with left and right child terms, the result of $\mathsf{decompLoc}$ applied to its left and right locations.

\begin{definition}
    $\mathsf{decompVertex}_{G}(v = (u,k)) = \genericTerm{u}{\mathsf{decompLoc}_{G}((v,p))}$
\end{definition}

The decomposition $\mathsf{decompLoc}(\ell)$ proceeds by inspecting the children of location $\ell$. If there are no children, the resulting child term is an empty hole. This hole is annotated with $\ell$ to act as a unique identifier. If $\ell$ has children, each is annotated with its edge UID and decomposed using the function $\mathsf{decompChild}$. If there are multiple children, they are placed in a local conflict, also annotated with $\ell$. To continue our example, the left position of vertex 40 has no edges, and therefore the result is a hole annotated with the source $(40, \texttt{*}, L)$. The right position has one edge, so the right argument to the multiplication term is the result of applying $\mathsf{decompChild}$ to the destination vertex of this edge.

\begin{definition}
\begin{align*}
    \mathsf{decompLoc}_{G}(\ell) &= \begin{cases}
        \ehole &\quad \mathsf{children}_{G}(\ell) = \emptyset \\ \eid{w}{}{\mathsf{decompChild}_G(w, v')} &\quad \mathsf{children}_{G}(\ell) = \SetOf{(w,v')}\\ \conflictHoleTerm{\ell}{\eid{w}{}{\mathsf{decompChild}_G(w, v')} \SuchThat{(w,v') \in \mathsf{children}_{G}(\ell)}} &\quad \abs{\mathsf{children}_{G}(\ell)} > 1 \\
    \end{cases}
\end{align*}
\end{definition}

A call to $\mathsf{decompChild}$ is like a call to $\mathsf{decompVertex}$, except that roots are not decomposed further, and a reference is returned instead. Without this case, decomposition could traverse cycles in the graph and never terminate. Instead, decomposition results in well-behaved trees that nevertheless store enough information to enable rich exploration and analysis. In our example, vertex 42 has multiple parents, and therefore is an MP root. Its decomposition is a reference storing the edge UID 41 and the destination vertex.

\begin{definition}
\begin{align*}
    \mathsf{decompChild}_G(w, v)  &= \begin{cases}
        \multiref  &\quad v\text{ is an MP root in } G\\
        \uniref  &\quad v\text{ is a U root in} G\\
        \mathsf{decompVertex}_{G}(v) &\quad v\text{ is not a root in} G
    \end{cases} 
\end{align*}
\end{definition}

Be decomposing each root vertex that appears in $G$, we obtain the whole grove.

\begin{definition}
    \mathsf{vertices}(G) = \SetOf{v \in \V \SuchThat{(w, (v,p), v') \in G \land w \in \W \land v' \in \V \land p \in \P}}
\end{definition}

\begin{definition}
    $\mathsf{decomp}(G) = \SetOf{\mathsf{decompVertex}_{G}(v) \SuchThat{\mathsf{root}_G(v)} \land v \in \mathsf{vertices}(G)}$
\end{definition}

The process can now be reversed and the same live subgraph can be obtained from the grove. At a high level, a term may be considered a tree, and therefore a graph, and the recomposition of a grove may be considered the simple union of each tree in the grove. Formally, the recomposition of a constructor applied to children is a set of edges that includes an edge from the parent to each child, as well as the recomposition of each child. Terms store vertices on constructors and edge UIDs for each child to enable the completeness of this reconstruction.

The recomposition of a constructor term is the set obtained by adding edges from each child of the term, according to the function $\mathsf{recompChildTerm}$, whereas references contain no graph structure of their own and recompose to the empty set.

\begin{definition}
    \begin{align*}
        \mathsf{recompTerm}(\genericTerm{u}{\subtermMV_p}) &= \bigcup_{p\in \mathsf{arity}(k)}\mathsf{recompChildTerm}(((u,k),p), \subtermMV_p)\\
        \mathsf{recompTerm}(\multiref) &= \emptyset\\
        \mathsf{recompTerm}(\uniref) &= \emptyset
    \end{align*}
\end{definition}

The recomposition of a child term at location $\ell$ includes, for each term in $\ell$, an edge from the $\ell$ to the term's corresponding vertex, as well as the term's recomposition. In the case of an empty hole, there are no terms, and this the recomposition is empty. If there are one or more terms, each term contributes its own edge and recomposition.

\begin{definition}
    \begin{align*}
    \mathsf{recompChildTerm}(\ell, \ehole) &= \emptyset\\
    \mathsf{recompChildTerm}(\ell, \eid{w}{}\termMV) &= \SetOf{(w, \ell, \mathsf{vertexOfTerm}(\termMV))}\cup \mathsf{recompTerm}(t)\\
    \mathsf{recompChildTerm}(\ell, \conflictHoleTerm{\ell}{\eid{w_i}{}\termMV_i}_{i<n}) &= \bigcup_{i<n} (\SetOf{(w_i, \ell, \mathsf{vertexOfTerm}(\termMV_i))}\cup \mathsf{recompTerm}(t_i))
    \end{align*}
\end{definition}

The vertex of a term is used as the destination of edges from the term's parent location. Thus the vertex of a constructor term is the vertex with the same UID and constructor, and the vertex of a reference is the vertex it refers to.

\begin{definition}
    \begin{align*}
    \mathsf{vertexOfTerm}(\genericTerm{u}{\subtermMV_p}) &= (u,k)\\
    \mathsf{vertexOfTerm}(\multiref) &= v\\
    \mathsf{vertexOfTerm}(\uniref) &= v
    \end{align*}
\end{definition}

The recomposition of a grove is simply the union of the recompositions of its terms.

\begin{definition}
    $\mathsf{recomp}(T) = \bigup{\termMV \in T}\mathsf{recompTerm}(\termMV)$
\end{definition}

The \autoref{thm:Recomposability} states that recomposition recovers the entirety of the original live subgraph, and therefore that groves faithfully represent the underlying data structure.

\begin{theorem}
\label{thm:Recomposability}
    (Recomposability) For all subgraphs $G$, $\mathsf{recomp}(\mathsf{decomp}(G))=G$.
\end{theorem}

\subsubsection{Partitioned Groves}

In order to facilitate user interaction with groves, we defined a presentation of a grove called a \textit{partitioned grove}. A partitioned grove contains three classes, corresponding to the the three kinds of roots: NP, MP, and U. A partitioned grove also designates a distinguished child term as the primary component of the program.

Formally, a partitioned grove $\gamma$ is a quadruple $(\subtermMV_r, NP, MP, U)$ where $\subtermMV_r \in ChildTerm$ and $NP, MP, U$ are finite sets of terms. The construction of a partitioned grove from the grove $\mathsf{decomp}(G)$ requires the designation of a \textit{distinguished root} location $\ell_r=(v_r=(u_r,k_r),p_r)$ such that $v_r$ is an NP root in $G$ and $\mathsf{arity}(k_r)=\SetOf{p_r}$. The distinguished child term can be thought of as the decomposition of the distinguished root location. The NP, MP, and U partitions are delineated by the class of the root vertex of each term. The root vertex of a term is like the vertex of a term, except that the root vertex of a reference is the \textit{source} of the corresponding edge, rather than the destination. A term rooted at the distinguished root is excluded from the NP class, since its contents are already identified as the distinguished child term. \autoref{fig:Decomposition example partitioned grove} provides an example of a partitioned grove, where the position $\mathsf{Root}$ of vertex 0 is the distinguished root location.

\begin{definition}
\begin{align*}
    \mathsf{distinguishedChildTerm}(\ell_r=((u_r,k_r),p_r), \Theta) &= \subtermMV_r &\quad \text{ if }
    \eid{u_r}{k_r}\ {\subtermMV_r}\in \Theta \\
    \mathsf{distinguishedChildTerm}(\ell_r=((u_r,k_r),p_r), \Theta) &= \emptyHole{\ell_r} &\quad \text{ otherwise}
\end{align*}
\end{definition}

\begin{definition}
    \begin{align*}
    \mathsf{rootVertexOfTerm}_{G}(\genericTerm{u}{\subtermMV_p}) &= (u,k)\\
    \mathsf{rootVertexOfTerm}_{G}(\multiref) &= v'\mathsf{ where }(w, (v',p), v)\in G\\
    \mathsf{rootVertexOfTerm}_{G}(\uniref) &=  v'\mathsf{ where }(w, (v',p), v)\in G
    \end{align*}
\end{definition}

\begin{definition}
    \begin{align*}
    \mathsf{grove}(\ell_r=(v_r,p_r), \Theta, G) = (&\mathsf{distinguishedChildTerm}(\ell_r, \Theta),\\ &\SetOf{t\in \Theta \SuchThat{ \mathsf{rootVertexOfTerm}_{G}(t)\neq v_r\text{ and is an NP root in }G}},\\
    &\SetOf{t\in \Theta \SuchThat{ \mathsf{rootVertexOfTerm}_{G}(t)\text{ is an MP root in }G}},\\ 
    &\SetOf{t\in \Theta \SuchThat{ \mathsf{rootVertexOfTerm}_{G}(t)\text{ is a U root in }G}}) 
    \end{align*}
\end{definition}

\figureUserActionSyntax

\subsection{User Edit Actions}%
\label{sub:User Actions}

We have defined a patch language directly in terms of graph edges and edge states. Since the user interacts with the more intuitive, tree-based grove representation of the program, we require an equally intuitive system of \textit{user edit actions} (edits) that can be translated into the patch language. The new sorts are defined in \autoref{fig:User Action Syntax}.

Edit actions operate on zippered terms~\cite{DBLP:journals/jfp/Huet97,DBLP:conf/popl/OmarVHAH17}. The definitions of zipper terms and child terms are standard, except that the cursor cannot select a child term of the form $\eid{w}{}\termMV$, but is forced to select the term $\termMV$ directly. While inspecting a grove, the user may select an edit action $\alpha$ while the cursor is represented by $\Z{\termMV}$ or $\Z{\subtermMV}$, and this will result in a graph patch $\Q{\pi}$ to be applied directly to the  graph.

$\Construct{k}$ constructs a new edge whose destination is determined by the cursor's location. If the cursor is on an empty hole, the hole is filled by constructing a fresh vertex with constructor $k$. If the cursor is on an existing term or child term, we \emph{wrap} the existing term or child term by deleting it from its parent, constructing the new term in its place, then re-connecting the original term as a child of the new term.

$\Delete$ deletes any edges that pass from constructors above the cursor to constructors below the cursor. If an empty hole is selected, deletion is a no-op. If a local conflict is selected, each edge from that location to a conflicting child is deleted. If a constructor is selected, every edge targeting that constructor is deleted. If a reference is selected, the corresponding edge is deleted.

$\Relocate{\ell}$ combines edge construction and deletion into a single \emph{atomic} operation. Provided the location $\ell$ is empty, the edges through the cursor are deleted, as described above, and new edges are simultaneously constructed with the same destinations as the old, but originating at location $\ell$. This is not equivalent to the composition of a sequence of deletions and constructions, because the UIDs of the relocated vertices remain the same.

Formally, these patches are defined in terms of the function $\mathsf{edges}_G$ (\autoref{def:edges}) which produces the set of edges in the live subgraph $G$ passing through a cursor at the given term or child term. The live subgraph $G$ is therefore an additional input to the patch construction judgment.

\begin{definition}
\label{def:edges}
    \[
    \begin{array}{ccc}
    \mathsf{edges}_G(\ehole) & = & \emptyset\\
    \mathsf{edges}_G(\conflict{\termMV}) & = & \SetOf{(w_i,\ell,v)\SuchThat{(w_i,\ell,v)\in G\land i<n\land \ell\in\L\land v\in\V}}\\
    \mathsf{edges}_G(\genericTerm{u}{\subtermMV_p}) & = & \SetOf{(w,\ell,(u,k))\SuchThat{(w,\ell,(u,k))\in G\land w\in\W\land \ell\in\L}}\\
    \mathsf{edges}_G(\multiref) & = & \SetOf{(w,\ell,v)\SuchThat{(w,\ell,v)\in G\land\ell\in\L}}\\
    \mathsf{edges}_G(\uniref) & = & \SetOf{(w,\ell,v)\SuchThat{(w,\ell,v)\in G\land\ell\in\L}}\\
    \end{array}
\]
\end{definition}

For the purpose of the construction action, it is assumed that each constructor $k\in\K$ in the language is equipped with a designated position $\mathsf{defaultPos}(k)\in \mathsf{arity}(k)$. The formal patch generation judgment is defined in terms of $\mathsf{edges}$ and $\mathsf{defaultPos}$ in \autoref{fig:User Action Transitions}.

\figureUserActionTransitions

\subsection{Mechanized Metatheory}

The definitions and theorems in \autoref{sub:Convergent Graphs} or \autoref{subsub:Terms}, and \autoref{subsub:Graph Decomposition} have been mechanized in the Agda proof assistant, with the exception that termination is justified separately for some definitions and proofs. In particular, \autoref{thm:Commutativity} and \autoref{thm:Recomposability} are formalized and proven. \autoref{lem:uniq-classification} follows from the more powerful theorem of classification correctness and completeness, which is mechanized and used to prove \autoref{thm:Recomposability}.

### 2024-10-25 OriginalSubmissionSource/grove-by-example.tex
Updates from Overleaf

--- 2024-07-11 grove-by-example.tex
@@
 where edits are communicated as they occur.
-This paper makes no assumptions about which batching mode is in use, nor do we consider the well-studied problem of reliably and efficiently communicating patches over a network\todo{cite something, dist sys textbook?}.\todo{maybe include a brief note in the discussion about cursors being represented by locations?}
+This paper makes no assumptions about which batching mode is in use, nor do we consider the well-studied problem of reliably and efficiently communicating patches over a network.
 %\autoref{sub:Cursors} discusses representing cursor locations when in real-time mode.
@@
 For clarity, we abbreviate \texttt{var(}$x$\texttt{)} as simply \texttt{x}; here, $x$ is a constructor parameter.
-We treat identifiers and literals as indivisible, but we discuss character-level editing in Section~TODO:REF\todo{ref}.
+We treat identifiers and literals as indivisible, but we discuss character-level editing in \autoref{sec:Discussion and Conclusion}.
 
@@
 that effectively ensures that collaborators always generate distinct UIDs, e.g. by generating
-universally unique IDs (UUIDs)~\cite{IETF RFC 4122}.
+universally unique IDs (UUIDs)~\cite{paskin1999toward}.
 
@@
 
-\subsection{Edits}%
+\subsection{Structure Editing}%
 \label{sub:Single-User Actions}
@@
 Individual users perform \emph{edits} to evolve the edit state.
-This paper abstracts over the user interface aspects of structure editors and makes no usability-related claims; edits could be performed through, for example, drag-and-drop interactions (as in block-based editors like Scratch) or keyboard interactions (as in MPS and Hazel).
+We consider several standard edits, including insertion, deletion, cut-and-paste (relocation), copy-and-paste, and undo/redo.
+This paper abstracts over the user interface aspects of structure editors and makes no usability-related claims; these edits could be performed through, for example, drag-and-drop interactions (as in block-based editors like Scratch) or keyboard interactions (as in MPS and Hazel).
 % User interface design for structure editors remains an active research area.
@@
 performing edits that translate to patches as described above. 
-They periodically communicate these patches to one another. \autoref{fig:Commutativity} diagrams the Grove collaboration model.\todo{stretch: include a comparable diagram for }
+They periodically communicate these patches to one another. \autoref{fig:Commutativity} diagrams the Grove collaboration model.
 
@@
 Once Alice and Bob share their patches and apply each other's patch to their own edit state,
-both edit states converge to the graph in \autoref{fig:DifferentParts:c}.\todo{add more spacing around terms in Fig 5} 
+both edit states converge to the graph in \autoref{fig:DifferentParts:c}. 
 Because Grove's patch language is structural rather than line-based, the fact that these edits happened to be close to one another (i.e.
@@
 from the~\texttt{R} position of~\vWrapPlus{} to the~\texttt{L} position. As discussed above, this involves deleting~\eMoveTimes{} and adding~\eNestedPartsBob{}.
-Bob's resulting edit state is shown in \autoref{fig:NestedParts:b}.\todo{}
+Bob's resulting edit state is shown in \autoref{fig:NestedParts:b}.
 
@@
 but in general, merging patches can lead to graphs that do not map cleanly to a conventional syntax tree.
-We identify several different motifs that might arise, all of which give rise to different kinds of conflicts in the graph decomposition: \emph{local conflicts}, \emph{multi-location conflicts}, and \emph{cyclic location conflicts}.
+We identify several different motifs that might arise, all of which give rise to different kinds of conflicts in the graph decomposition: \emph{local conflicts}, \emph{relocation conflicts}, and \emph{unicyclic relocation conflicts}.
 As with merge conflicts in version-control systems such as git, these all require user intervention to resolve. 
+%As we will return to in \autoref{sec:Type System}\todo{sec}, conflicts need not disable type-based editor services (the \textbf{semantic gap problem}).
+Instead, the system can type check around these conflict (and in some cases, infer a type for the conflict itself).
 
@@
 
-When these patches are merged, \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
+When these patches are merged in \autoref{fig:MultiChild:c}, \emph{both} \eMultiChildAlice{} and \eMultiChildBob{}
 appear in the merged graph. 
 When decomposing this graph to a syntax tree, 
-we resolve this conflict in the~\texttt{R} position of~\vWrapPlus~by decomposing to a \emph{local conflict}, \figureCode{$\binaryConflictHole{x}{y}$}, as 
-shown in \autoref{fig:MultiChild:c}.
+we resolve this conflict in the~\texttt{R} position of~\vWrapPlus~by decomposing to a \emph{local conflict}, {\binaryConflictHole{x}{y}}.
 
@@
 
-Such conflicts can be resolved simply by deleting relocating one of the conflicting terms (and editing the remaining term into the correctly merged value, if needed), which would remove the corresponding edge.
+Local conflicts can be resolved simply by deleting or relocating all but one of the conflicting terms (and editing the remaining term into the correctly merged value, if needed), which would remove the corresponding edge.
 For example, Alice could resolve the problem by wrapping \li{x} and \li{y} with a multiplication, effectively moving them to non-conflicting locations.
@@
 
-It is worth noting one special situation: if Alice and Bob both independently filled the hole with structurally identical terms, e.g.
-\li{x}, Grove would \emph{still} formally identify a conflict, because each \li{x} has a distinct UID.
-In this situation, it would be reasonable for the system to resolve the conflict without further coordination by deterministically choosing one of the two UIDs, e.g.
-the smallest one.
+It is worth noting one special situation: if Alice and Bob independently filled the hole with structurally identical terms, e.g.
+\li{x}, Grove would \emph{still} formally identify a conflict, because the terms have distinct UIDs.
+In this situation, it would be reasonable for the system to resolve the conflict without further coordination by deterministically choosing one of the two terms, e.g.
+the smallest.
 
@@
 
-As we will return to in \autoref{sec:Type System}\todo{sec}, 
-a local conflict need not disable type-based editor services (the \textbf{semantic gap problem}).
-Instead, the system can type check around the conflict (and if the conflicted terms have a common type, the system can infer that type for the location conflict itself).
-
-\subsubsection{Multi-Location Conflicts}%
+\subsubsection{Relocation Conflicts}%
 \label{sub:Multi-parent conflicts}
@@
 
-Multi-parent conflicts occur when multiple edges share a destination vertex,
-as opposed to multi-child conflicts, which share a common origin.
-An example of this is in \autoref{fig:MultiParent},
-which shows the edit state Alice and Bob converge on after deleting the conflict and the \texttt{v}.
-Alice then moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vSimpleTimes{}.
-At the same time, Bob also moves~\vNestedPartsAlice{} to the~\texttt{R} position of~\vWrapPlus{}.
-This results in the edit states of \autoref{fig:MultiParent:b} and \autoref{fig:MultiParent:c}
+Grove's support for repositioning creates the possibility for \emph{relocation conflicts}.
+These occur when a merge causes a vertex to have multiple incoming edges, indicating that it does not have a uniquely determined location (as opposed to local conflicts, which occur when there are multiple outgoing edges at a specified location).
+
+For example, \autoref{fig:MultiParent} shows an example where Alice and Bob relocate a term, \li{w}, to two different locations (the two holes in \autoref{fig:MultiParent:a}).
 In both cases, the edits are modeled as one edge deletion followed by one edge addition.
-Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{}.
-At the same time, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{}.
-As we explain in \autoref{sec:Formalism}, deletions are idempotent,
+Alice deletes~\eNestedPartsAlice{} and adds~\eMultiParentAlice{} in \autoref{fig:MultiParent:b}.
+At the same time, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{} in \autoref{fig:MultiParent:c}.
+Edge deletions is idempotent,
 so the fact that Alice and Bob both deleted~\eNestedPartsAlice{} will not lead to a conflict. 
@@
 point to the same vertex.
-Once all updates are merged, the resulting edit state is given in \autoref{fig:MultiParent:d}.
+Once these patches are merged, the resulting edit state is given in \autoref{fig:MultiParent:d}.
 Notice~\vNestedPartsAlice{} has two edges pointing to it. 
+When decomposing the graph, we leave a \emph{relocation conflict reference}, $\multiVertex{18}$, at each conflicting location.
+The conflicted term is separately tracked in the result of decomposition, which, due to conflicts like these, is formally a set of terms with references like these between them.
+We call this set a \emph{grove}.
+This approach partially addresses the \textbf{relocation conflict problem} from \autoref{sec:Introduction} (we also need to handle unicycles, see below, to fully address the problem).
 
-TODO: talk about how to resolve the conflict
+To resolve relocation conflicts, a user can simply delete all but one of the relocation conflict references. This will cause the corresponding edges to be deleted,
+and when only one edge remains, there will no longer be a conflict.
+An editor might provide a more convenient way of deleting all but a selected relocation conflict reference, and provide affordances for displaying these terms, e.g.
+by transcluding them inline at each location or showing them in a separate sidebar.
 
@@
 \figureCycle
+\figureDisconnect
 
-In addition to multi-child and multi-parent conflicts, merges
-can cause cycles in the graph.\footnote{In theory a single user could
-generate cycles, but this could be detected immediately and the user action
-that caused that cycle rejected.
-Thus, we are only concerned with cycles that cannot be immediately detected
-and prevented by the editor.}
+Relocation in a collaborative setting can also lead to cycles in the graph. 
 For example, consider the situation in \autoref{fig:Cycle}.
-\autoref{fig:Cycle:a} shows the graph after Alice and Bob have made some edits,
-and shared them with each other.
-Alice then moves~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
-and then~\vMultiCyclePlus{} underneath that.
-Bob does the same but puts~\vMultiCycleTimes{} under~\vMultiCyclePlus{}.
-This results in \autoref{fig:Cycle:b} and \autoref{fig:Cycle:c}, respectively.
-
-On their own, neither of these edits creates a cycle.
-However, merging the edit actions of both Alice and Bob results in the graph
+Starting in \autoref{fig:Cycle:a}, 
+Alice relocates~\vMultiCycleTimes{} to the \texttt{R} child of~\vWrapPlus{}
+and then~\vMultiCyclePlus{} underneath that in \autoref{fig:Cycle:b}.
+Bob does the opposite, putting~\vMultiCycleTimes{} under~\vMultiCyclePlus{} in \autoref{fig:Cycle:c}. On their own, neither of these edits creates a cycle.
+However, merging the two patches results in the graph
 in \autoref{fig:Cycle:d}, which has a cycle
 between~\vMultiCycleTimes{} and~\vMultiCyclePlus{}.
-As with the other kinds of conflict, resolving it requires user input.
-For example, by either deleting both~\eMultiCycleBobPlus{} and~\eMultiCycleBobTimes{}~(thus favoring Alice's version)
-or deleting both~\eMultiCycleAliceTimes{} and~\eMultiCycleAlicePlus{}~(thus favoring Bob's version).
 
-\paragraph{Disconnected Cycles}
+The main difficulty with cycles has to do with decomposition back to a term, i.e.
+a syntax tree.
+We do not want decomposition to traverse endlessly attempting to create an infinite tree, so we need to break the cycle somewhere.
 
-\figureDisconnect
+If the cycle is connected to a larger term, then there will necessarily be at least one vertex along the cycle that has multiple incoming edges.
+In \autoref{fig:Cycle}, both \vMultiCycleTimes{} and \vMultiCyclePlus{} have multiple incoming edges.
+As described above, decomposition will leave relocation conflict references in these positions, thereby breaking the cycle.
+In this example, these references appear within a local conflict as well, because both vertices were relocated under a common parent vertex.
+As before, this cycle can be broken by deleting or otherwise modifying the terms until there are no longer any such conflicts.
+
+It is also possible to merge patches such that a \emph{disconnected unicycle} emerges in the graph, even when neither patch disconnects any vertex from the root.
+\autoref{fig:Disconnection} shows a simple example of when this could occur: Alice relocates~\vMultiCycleTimes{} under~\vSimpleTimes{},
+while Bob relocates~\vSimpleTimes{} under~\vMultiCycleTimes{}. 
+This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
+Merging these causes both vertices to become disconnected, because they were both relocated.
+The inserted edges form a \emph{unicycle}, meaning a cycle where every vertex has in-degree $1$.
+In this case, we cannot rely on relocation references to break the cycle.
+Instead, we break the cycle by arbitrarily but deterministically choosing an edge along the unicycle, e.g.
+the edge with the smallest UID, and leaving a \emph{unicycle conflict reference} at that location, as shown in \autoref{fig:Disconnect:d}.
+A user can be notified of this situation when merging and resolve these conflicts again by relocating or deleting terms until the cycle no longer exists in the graph.
+
+Relocation conflict references and unicycle conflict references together fully address the \textbf{relocation conflict problem} from \autoref{sec:Introduction}.
 
@@
 % \todo{UI would show orphans only once edits orphans occur}
-
-Finally, there is a particular type of cycle worth discussing.
-
-Since a cycle like this that is connected to the root vertex always
-contains vertexes with multiple parents that if removed would break the cycle,
-we use the user interface described in \autoref{sub:Multi-parent conflicts}
-to display these kinds of programs.
-For example, \autoref{fig:Cycle:d} would be displayed.
-We leave the user interface considerations of resolving the conflict such a
-cycle represents to future work.
-
-Finally, when multi-user edits are merged,
-parts of the graph can become disconnected from the root
-even though they are connected to the root in each user's
-copy of the graph before the merge.
-
-For example, consider the situation in \autoref{fig:Disconnect:a},
-which shows the graph after Alice and Bob have made some edits,
-and shared their edits with each other.
-Alice then moves~\vMultiCycleTimes{} under~\vSimpleTimes{},
-while Bob moves~\vSimpleTimes{} under~\vMultiCycleTimes{}.
-This results in \autoref{fig:Disconnect:b} and \autoref{fig:Disconnect:c}, respectively.
-
-When these edits are merged, the result is \autoref{fig:Disconnect:d}.
-Since Alice deleted~\eNestedPartsBob{}
-and Bob deleted~\eMultiCycleAliceTimes{},
-when these are merged~\vMultiCycleTimes{} and~\vSimpleTimes{}
-are disconnected from the root.
-
-This is similar to the type of merge conflict
-that can arise in traditional version control systems
-when two users delete different functions that
-are duplicates of the other.
-
-Fortunately, we can detect this situation and warn the user
-since the disconnection appears after merging two graphs
-in which vertexes were not already disconnected.
 

### 2024-10-25 OriginalSubmissionSource/introduction.tex
Updates from Overleaf

--- 2024-07-11 introduction.tex
@@
 % Software development is a highly collaborative activity. 
-Development teams typically collaborate with the aid of a version control system (VCS) like Git~\cite{git}, Subversion~\cite{subversion}, or Darcs~\cite{darcs}. 
+Development teams typically collaborate with the aid of a version control system (VCS) like Git, Subversion, or Darcs~\cite{DBLP:conf/haskell/Roundy05}. 
 These systems maintain a branching history of \emph{commits} to a source code repository, each consisting of a \emph{patch} together with various metadata, e.g.
@@
 Developers do not typically express program edits using the patch language directly, nor in any case do version control systems typically have access to a log of the developer's edits.
-Instead, version control systems must \emph{synthesize} patches from the file system state using heuristic algorithms, such as the classic \li{diff} algorithm that synthesizes a patch that minimizes the edit distance between two file system states.\todo{cite}
+Instead, version control systems must \emph{synthesize} patches from the file system state using heuristic algorithms, such as the classic \li{diff} algorithm that synthesizes a patch that minimizes the edit distance between two file system states \cite{DiffAlgorithm}.
 
@@
 The standard approach is to apply the \emph{local patch} first, then modify the \emph{remote patch} by shifting its line numbers to account for the local patch's line insertions and deletions.
-This algorithm is an \emph{operational transform}~\cite{ot}.
-Character-level operational transforms are similarly used for real-time collaborative editing, e.g. in tools like Google Docs and in Visual Studio Code's Live Share feature~\cite{crdts-for-text}.
+This algorithm is an \emph{operational transform}~\cite{DBLP:conf/sigmod/EllisG89}.
+Character-level operational transforms are similarly used for real-time collaborative editing, e.g. in tools like Google Docs and in Visual Studio Code's Live Share feature.
 
@@
 Let us briefly review some classic problems with these systems.
-Appendix~\ref{appendix:merge-problems} shows examples of the problematic default behavior of Git in each of these scenarios.\todo{write this, can move to supplement if needed}
+%Appendix~\ref{appendix:merge-problems} shows examples of the problematic default behavior of Git in each of these scenarios.\todo{write this, can move to supplement if needed}
 
@@
 For example, systems like Git address the nesting problem by allowing indentation changes from one patch to be merged with other changes that do not modify indentation. 
-More sophisticated systems like GumTree, SmartDifferencer\todo{cites, others} deploy parsers and tree differencing algorithms to better
+More sophisticated systems deploy parsers and tree differencing algorithms~\cite{DBLP:conf/sigmod/ChawatheG97, DBLP:journals/tse/FluriWPG07,DBLP:conf/kbse/FalleriMBMM14,DBLP:conf/doceng/Lindholm04,DBLP:conf/fase/NguyenNPN10,DBLP:journals/scp/SchwagerlUW15} to better
 address the granularity problem.
@@
 a syntax tree with holes, shown projected visually in various ways to the developer.
-Structure editing has been studied since the 1980s with the Cornell Program Synthesizer 
-and research continues to this day, with numerous active projects including Scratch, Snap!, and other block-based editors (which are widely used in educational and end-user programming settings), Jetbrains MPS (which has been deployed in industry), and Hazel (a live functional programming environment rooted in a structure editor calculus called Hazelnut, which serves as
-an active research platform).
+Structure editing has been studied since the 1980s with the Cornell Program Synthesizer~\cite{DBLP:journals/cacm/TeitelbaumR81}
+and research continues to this day, with numerous active projects including Scratch~\cite{maloney2010scratch}, Snap!, and other block-based editors (which are widely used in educational and end-user programming settings), Jetbrains MPS (which has been deployed in industry)~\cite{voelter2011language}, and Hazel (a live functional programming environment rooted in a structure editor calculus called Hazelnut, which serves as
+an active research platform)~\cite{DBLP:conf/popl/OmarVHAH17}.
 
@@
 so remote patches can be applied without transformation. 
-We prove a \emph{convergence theorem} that ensures that replicas (a.k.a. clones) of a repository will converge to the same state 
+We prove a \emph{convergence theorem} that ensures that branches of a repository will converge to the same state 
 when the same set of patches are applied, regardless of the order in which they are applied.
-The patch language forms what is known as a \emph{commutative replicated data type (CmRDT)}.
+The patch language forms what is known as a \emph{commutative replicated data type (CmRDT)}~\cite{preguicca2018conflict,shapiro2011conflict}.
 
@@
 More than one out-edge at a given location decomposes to an explicitly represented \emph{local conflict}. 
-More than one in-edge indicates that a vertex has a \emph{location conflict}, so we leave a \emph{location conflict reference} at each of the conflicting locations. 
-Finally, cycles are broken during decomposition by leaving a \emph{cycle conflict reference} at an arbitrary (but deterministically chosen) edge. 
+More than one in-edge indicates that a vertex has a \emph{relocation conflict}, so we leave a \emph{relocation conflict reference} at each of the conflicting locations. 
+Finally, cycles are broken during decomposition by leaving a \emph{unicycle conflict reference} at an arbitrary (but deterministically chosen) edge. 
 Resolving these various conflicts simply requires 
-manipulating these constructs like any other syntactic construct, e.g. deleting all but one location conflict reference to determine a unique location for a node.
+manipulating these constructs like any other syntactic construct, e.g. deleting all but one relocation conflict reference to determine a unique location for a node.
 
@@
 Developers are left to reason without the aid of much of their tooling during conflict resolution.
-This is an instance of the more general \textbf{semantic gap problem} when programming tools encounter incomplete programs~\cite{HazelSNAPL}.
+This is an instance of the more general \textbf{semantic gap problem} when programming tools encounter incomplete programs~\cite{DBLP:conf/snapl/OmarVHSGAH17}.
 
 The previous work on the Hazelnut structure editor calculus addresses the semantic gap problem in the single-user setting by 
-defining a type system and type error localization system (the \emph{marked lambda calculus}) for incomplete programs, i.e. programs with holes. 
+defining a type system and type error localization system (the \emph{marked lambda calculus}) for incomplete programs, i.e. programs with holes~\cite{DBLP:journals/pacmpl/ZhaoMDBPO24}. 
 Notably, type error localization is proven to be \emph{total}, i.e.
 the system is able to assign static meaning to \emph{every} syntactically well-formed expression by inserting \emph{marks} to localize errors.
-Marking employs local type inference as codified by a \emph{bidirectional type system} as well as \emph{gradual typing}, i.e.
+Marking employs local type inference as codified by a \emph{bidirectional type system}~\cite{dunfield2019} as well as \emph{gradual typing}~\cite{siek2015}, i.e.
 the theory of \emph{type holes}, to recover from situations where type errors make it impossible to
@@
 
-This paper extends this work to develop a total type error localization system for groves, introduced in Contribution 1 above as the result of decomposing a commutatively edited graph into a set of terms with empty holes, local conflicts, location conflict references, and cycle conflict references.
+This paper extends this work to develop a total type error localization system for groves, introduced in Contribution 1 above as the result of decomposing a commutatively edited graph into a set of terms with empty holes, local conflicts, relocation conflict references, and unicycle conflict references.
 This paper develops a type (and type error localization) discipline for handling these novel constructs.
@@
 We establish key metatheoretic properties using the Agda proof assistant.
-\autoref{sec:Type System} instantiates Grove with a simply typed lambda calculus, then defines a bidirectional type and type localization system for groves and proves totality and other key metatheoretic properties using the Agda proof assistant.
-\autoref{sec:Implementation} describes our OCaml implementation of Grove, which is defined modularly to allow it to be instantiated with arbitrary syntax trees. 
+\autoref{sec:Implementation} describes our implementation of the Grove Workbench, which is defined modularly to allow it to be instantiated with arbitrary syntax trees. 
+\autoref{sec:Marking System} instantiates Grove with a simply typed lambda calculus, then defines a bidirectional type and type localization system for groves and proves totality and other key metatheoretic properties using the Agda proof assistant.
 \autoref{sec:Related Work} reviews related work in more detail. 

### 2024-10-25 OriginalSubmissionSource/marking-system.tex
Updates from Overleaf

\section{Total Type Error Localization and Recovery for Groves}%
\label{sec:Marking System}

Resolving conflicts is often a time-consuming process 
and requires understanding the semantics of the program
even while it still has conflicts.

Traditionally, conflicts are indicated by inserting extra-linguistic conflict markers into files. 
These markers can limit the operation of language services that need a well-formed or well-typed term, 
e.g., type error reporting.

Ignoring the markers is not sufficient because 
multiple conflicting versions of the code may appear
between the markers. This can complicate variable resolution and type checking.

In the previous sections, we developed a system for explicitly representing various forms of syntactic conflict directly in the grove resulting from graph decomposition. We now consider the static semantics of 
groves.

Groves are sets of terms with holes, local conflicts, relocation conflict references, and unicycle conflict references between them. These terms may not yet be well-typed during the course of a collaboration. 
As such, we need to consider the problem of type error localization and recovery.

We build on recent work on the marked lambda calculus~\cite{DBLP:journals/pacmpl/ZhaoMDBPO24}, which we briefly review in \autoref{sec:Background: The Marked Lambda Calculus}. 
The marked lambda calculus specifies a \emph{total} 
type error localization system, i.e. one where every syntactically well-formed term can be \emph{marked} with errors to produce a statically meaningful term.

In this section, we extend the marked lambda calculus to allow reasoning about conflicted programs, i.e. groves (instantiated with the syntax of the simply typed lambda calculus). 
The \emph{marked grove calculus} maintains the key totality property, so \emph{every edit state that can arise in the course of a collaboration is statically meaningful}. Downstream language services can, therefore, continue to provide support while users resolve conflicts. This addresses the \textbf{semantic gap problem} described in \autoref{sec:Introduction}.

We follow the marked lambda calculus in basing our \emph{marked grove calculus} in bidirectional type checking~\cite{dunfield2019} and deploying gradual typing~\cite{siek2015} when conflicts do not allow a single type to be inferred, then layer on a unification-based type inference system to opportunistically fill holes or suggest partial solutions when there are conflicting types due to conflicting syntax. We give an overview of the key judgments and rules in \autoref{sub:marking-groves} but due to space considerations and because much of the development is based directly on the prior work, we leave the full details to the supplemental material and the mechanized metatheory, which is briefly outlined in \ref{sub:marking-agda}.

\subsection{Background: The Marked Lambda Calculus}
\label{sec:Background: The Marked Lambda Calculus}
The marked lambda calculus is a gradual bidirectionally typed rewriting system that takes an arbitrary syntactically well-formed expression and \emph{marks} it by localizing type errors, producing a \emph{marked expression}.

The key judgments are synthetic and analytic marking judgments that mark and type-check expressions: $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$ and $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$.  
The synthetic judgment is used when a type is to be locally inferred from $\EMV$, while the analytic judgment is used when surrounding type annotations determine an expected type for $\EMV$. A subsumption rule, which is not shown, allows analysis at any type consistent with the synthesized type.

For example, the rules for marking variables are reproduced below:
\begin{mathpar}
  \inferrule[MKSVar]{
    \inCtx{\ctx}{x}{\TMV}
  }{
     \synMarkConstraint{\ctx}{\eVar{}{x}}{\eVar{}{x}}{\TMV}
  }

  \inferrule[MKSFree]{
    \notInCtx{\ctx}{x}
  }{
    \synMarkConstraint{\ctx}{\eVar{}{x}}{\ECFree{\eVar{}{x}}}{\TUnknown}
  }
\end{mathpar}
The first rule handles bound variables. The second rule handles free variables, which would normally be statically meaningless. Here, we instead mark it explicitly as a free variable and synthesize the unknown type, $\TUnknown$.

The rules for marking lambda abstractions when an expected type is provided similarly handle both well-typed and ill-typed cases (which is key to proving totality):
\begin{mathpar}
\label{inf:marking-rules}
\inferrule[MKALam1]{
        \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
        \consistent{\TMV}{\TMV_1} \\
        \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TMV_2}
    }{
        \ctxAnaFixedInto{\ctx}{\eFun{}{x}{\TMV}{\EMV}}{\eFun{}{x}{\TMV}{\ECMV}}{\TMV_3}
    }
\end{mathpar}
\begin{mathpar}
     \inferrule[MKALam2]{
        \notMatchedArrow{\TMV_3} \\
        \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TUnknown}
    }{
        \ctxAnaFixedInto{\ctx}{\eFun{}{x}{\TMV}{\EMV}}{\ECLamAnaNonMatchedArrow{x}{\TMV}{\ECMV}}{\TMV_3}
    }

    \inferrule[MKALam3]{
        \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
        \inconsistent{\TMV}{\TMV_1} \\
        \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV_1}}{\EMV}{\ECMV}{\TMV_2}
    }{
        \ctxAnaFixedInto{\ctx}{\eFun{}{x}{\TMV}{\EMV}}{\ECLamInconAsc{x}{\TMV}{\ECMV}}{\TMV_3}
    }

\end{mathpar}

The first rule invokes the matched arrow judgment (which handles the situation where the expected type is unknown \cite{snapl15}) and performs a type consistency check with the argument type annotation, after which we can analyze the body against the expected output type while recursively marking it.

If there is not a matched arrow type, the second rule marks the lambda with an error (while recovering recursively into the body at an unknown type).
If the argument type annotation is inconsistent, the third rule similarly marks the lambda with a different error. 
The subscript and superscript are formal analogs of type error messages.

Marking is total, meaning every well-formed expression can be marked in this way, producing a well-typed marked expression as governed by the bidirectional typing judgments for marked expressions shown here: $\ctxSynTypeM{\ctx}{\ECMV}{\TMV}$, $\ctxAnaTypeM{\ctx}{\ECMV}{\TMV}$.

% \todo{TODO: maybe put the statement of totality here?}

These typing rules emit constraints because once marked, the system layers on unification-based type inference to opportunistically fill the unknown types, i.e., the type holes, that arise. In case the constraints governing a type hole cannot be solved, the system suggests hole fillings that satisfy a subset of the constraints, asking the user to resolve the issue (rather than attempting to heuristically decide which constraints were intended). When the user hovers over a choice, the system returns control to the bidirectional marking system.

\subsection{Marking Groves}
\label{sub:marking-groves}

We extend the marked lambda calculus to groves. The machinery related to type errors remains largely unchanged. We focus our attention on type error localization in the presence of conflicts.

\subsubsection{Syntax}
\label{sub:unmarked-lang}
\autoref{fig:marking-syntax} introduces the syntax of the marked grove calculus. We define an \emph{unmarked} language, which comprises unmarked expressions, $\EMV$,  child expressions, $\ELV$, types, $\TMV$, and  child types $\TLV$.
We also define a corresponding \emph{marked} language,  which comprises marked expressions, $\ECMV$, child expressions, $\ELMV$, types, $\TCMV$, and child types, $\TCLV$. This syntax is an instantiation of the generic term syntax introduced in \autoref{sub:Groves}.
The marked language differs from the unmarked language only by the inclusion of the marks that arise due to type errors taken directly from the prior work.

\subsubsection{Graph Erasure} The presence of conflicts and UIDs requires us to distinguish syntactic types from semantic types, $\STMV$, which are related by a graph erasure function, $\graphErase{\TMV} = \STMV$ and $\graphErase{\TLMV} = \STMV$, that replaces holes and conflicts with the unknown type and erases vertex and edge UIDs (not shown, see supplement for the full definition).

\subsubsection{Marking}
\label{sub:marking}
Marking is performed bidirectionally using four mutually defined judgments:
\synMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV},
\synMarkConstraint{\ctx}{\ELV}{\ELMV}{\STMV},
\anaMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV}, and
\synMarkConstraint{\ctx}{\ELV}{\ELMV}{\STMV}.

\begin{figure}
    \[\begin{array}{lcllll}
    \TMV & \in & \TMName & \Coloneqq & 
        \tnum 
        \mid \tarrow 
        \mid \multiref 
        \mid \uniref \\ 
    \TLMV & \in & \TLMName & \Coloneqq &
        \ehole
        \mid \lexp{\TMV}
        \mid \conflict{\TMV} \\
     \EMV & \in & \EMName & \Coloneqq & 
        \var{x}
        \mid \num 
        \mid \plus{\ELV_1}{\ELV_2} 
        \mid \mult{\ELV_1}{\ELV_2}
        \mid \lam{x}{\TLMV}{\ELV}
        \mid \app{\ELV_1}{\ELV_2}
        \mid \multiref
        \mid \uniref \\
     \ELV & \in & \ELMName & \Coloneqq & 
        \ehole
        \mid \lexp{\EMV} 
        \mid \conflict{\EMV} \\
        & & & \\
     \ECMV & \in & \ECMName & \Coloneqq &
        \var{x}
        \mid \num
        \mid \plus{\ELMV_1}{\ELMV_2}
        \mid \mult{\ELMV_1}{\ELMV_2}
        \mid \lam{x}{\TLMV}{\ELV}
        \mid \app{\ELMV_1}{\ELMV_2}
        \mid \multiref
        \mid \uniref \\ 
        & & & &
          \ECFree{\var{x}} 
        \mid \lamasc
        \mid \lamanamarr
        \mid \apsynmarr
        \mid \incontype \\
    \ELMV & \in & \ELMVName & \Coloneqq &
        \ehole
        \mid \lexp{\ECMV}
        \mid \conflict{\ECMV} \\
        & & & \\
    \STMV & \in & \STMName & \Coloneqq & 
        \TUnknown
        \mid \tNum{}
        \mid \TArrow{\STMV_1}{\STMV_2} \\
        & & & \\
    \PTMV & \in & \PTMName & \Coloneqq & 
        \TUnknown^{\Provp}
        \mid \tNum{}
        \mid \TArrow{\PTMV_1}{\PTMV_2} \\
    \Provp & \in & \Prov & \Coloneqq &
        \mathsf{typ}(\ell)
        \mid \mathsf{exp}(\ell)
        \mid \mathsf{ref}(w)
        \mid \mathsf{mark}(u)
        \mid ~\rightarrow_L(\Provp) \mid ~\rightarrow_R(\Provp)
        \mid \mathsf{anon} \\
    m & \in & \mathsf{Mode} & \Coloneqq & \mathsf{syn} \mid \mathsf{ana}(\STMV)
\end{array}\]
    \vspace{-10px}
    \caption{Syntax}
    \label{fig:marking-syntax}
\end{figure}

For standard forms in the simply typed lambda calculus, the rules correspond directly to those from the marked lambda calculus. For the sake of brevity, we include only the rules related to lambda expressions below; the full set of rules are in the supplemental material.

\begin{mathpar}
\label{inf:mark-lam-ana}
\inferrule[MKALam1]{
        \graphErase{\TLMV} = \sigma \\
        \matchedArrow{\sigma_3}{\sigma_1}{\sigma_2} \\
        \consistent{\sigma}{\sigma_1} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\sigma_2}
    }{
        \anaMarkConstraint{\ctx}{\lam{x}{\TLMV}{\EMV}}{\lam{x}{\TLMV}{\ELMV}}{\sigma_3}
    }
\end{mathpar}
\begin{mathpar}
     \inferrule[MKALam2]{
        \graphErase{\TLMV} = \sigma \\\\
        \notMatchedArrow{\sigma_3} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma}}{\ELV}{\ELMV}{\TUnknown}
    }{
        \anaMarkConstraint{\ctx}{\reglam}{\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}
    }
\end{mathpar}
\begin{mathpar}
     \inferrule[MKALam3]{
        \erase{\TLMV} = \sigma \\\\
        \matchedArrow{\sigma_3}{\sigma_1}{\sigma_2} \\
        \inconsistent{\sigma}{\sigma_1} \\
        \anaMarkConstraint{\extendCtx{\ctx}{x}{\sigma_1}}{\ELV}{\ELMV}{\sigma_2}
    }{
        \anaMarkConstraint{\ctx}{\reglam}{\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}}{\sigma_3}
    }
\end{mathpar}

These rules differ from the original rules, reproduced in \autoref{sec:Background: The Marked Lambda Calculus}, only in that they graph-erase the type annotation and recurse using the marking rules for child expressions. For empty holes and singleton expressions, these are straightforward. We return to local conflicts below.%
% \todo{synthetic rules for child expressions?}
\begin{mathpar}
    \inferrule[MKSHole]{ }{
        \synMarkConstraint{\ctx}{\emptyHole{\ell}}{\emptyHole{\ell}}{\TUnknown}
    }

    \inferrule[MKSOnly]{ 
        \synMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV}
    }{ 
        \synMarkConstraint{\ctx}{\lexp{\EMV}}{\lexp{\ECMV}}{\STMV}
    }

    \inferrule[MKAHole]{ }{
        \anaMarkConstraint{\ctx}{\emptyHole{\ell}}{\emptyHole{\ell}}{\STMV}
    }

    \inferrule[MKAOnly]{ 
        \anaMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV}
    }{ 
        \anaMarkConstraint{\ctx}{\lexp{\EMV}}{\lexp{\ECMV}}{\STMV}
    }
\end{mathpar}

Once marked, the corresponding marked expression typing judgments, $\synConstraint{\ctx}{\ECMV}{\PTMV}{\cConstraint}{\mConstraint}$, $\synConstraint{\ctx}{\ELMV}{\PTMV}{\cConstraint}{\mConstraint}$, 
$\anaConstraint{\ctx}{\ECMV}{\PTMV}{\cConstraint}{\mConstraint}$, and $\anaConstraint{\ctx}{\ELMV}{\PTMV}{\cConstraint}{\mConstraint}$ emit type inference constraints, $\cConstraint$, as in prior work, and, recursively collect location conflict contexts, $\mConstraint$, discussed further below. When the system generates inference constraints and tries to unify them, it encounters unknown types that need to be distinguished. This motivates the need for $\PTMV$ in \autoref{fig:marking-syntax}, which extends the sort of semantic types to enable linking generated unknown types to their associated provenance denoted by $q$. Provenances help locate the origin of the unknown type by associating them with their location $\ell$, edge-ID(w), or vertex-ID (u).

% \todo{mention why we need provenances}
% \todo{marked expression typing rules for lambdas and marked lambdas}

\begin{mathpar}
    \inferrule[MALam1]{
        \matchedArrowConstraint{\PTMV_3}{\PTMV_1}{\PTMV_2}{\cConstraint_1} \\
        \graphErase{\TLMV} = \PTMV \\
        \consistent{\PTMV}{\PTMV_1} \\
        \anaConstraint{\extendCtx{\ctx}{x}{\PTMV}}{\ELMV}{\PTMV_2}{\cConstraint_2}{\mConstraint_1}
    }{
        \anaConstraint{\ctx}{\lam{x}{\TLMV}{\ELMV}}{\PTMV_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\PTMV}{\PTMV_1}}}{\mConstraint_1}
    }
\end{mathpar}
\begin{mathpar}
    \inferrule[MALam2]{
        \notMatchedArrow{\PTMV_3} \\
        \graphErase{\TLMV} = \PTMV \\
        \anaConstraint{\extendCtx{\ctx}{x}{\PTMV}}{\ELMV}{\TUnknown^{\mathsf{anon}}}{\cConstraint}{\mConstraint}
    }{
        \anaConstraint{\ctx}{\ECLamAnaNonMatchedArrow{\var{x}}{\TLMV}{\ELMV}}{\PTMV_3}{\cConstraint \cup \Setof{\constrain{\TUnknown^{\mathsf{mark}(u)}}{\PTMV_3}}}{\mConstraint}
  }
\end{mathpar}
\begin{mathpar}
    \inferrule[MALam3]{
        \matchedArrowConstraint{\PTMV_3}{\PTMV_1}{\PTMV_2}{\cConstraint_1} \\
        \graphErase{\TLMV} = \PTMV \\
        \inconsistent{\PTMV}{\PTMV_1} \\
        \anaConstraint{\extendCtx{\ctx}{x}{\PTMV_1}}{\ELMV}{\PTMV_2}{\cConstraint_2}{\mConstraint}
    }{
        \anaConstraint{\ctx}{\ECLamInconAsc{\var{x}}{\TLMV}{\ELMV}}{\PTMV_3}{\cConstraint_1 \cup \cConstraint_2 \cup \Setof{\constrain{\TUnknown^{\mathsf{mark}(u)}}{\PTMV_3}}}{\mConstraint}
    }
\end{mathpar}

\subsubsection{Local Conflicts}
To mark local conflicts, we recursively mark all conflicted terms under the given context and with the same expected type when available. The local conflict as a whole synthesizes the unknown type, i.e. it is essentially a ``conflict hole'':
\begin{mathpar}
    \inferrule[MKSLocalConflict]{ 
        \Setof{\synMarkConstraint{\ctx}{\EMV}{\ECMV}{\sigma_i}}_{i<n}
    }{
        \synMarkConstraint{\ctx}{\conflict{\EMV}}{\conflict{\ECMV}}{\TUnknown}
    }

    \inferrule[MKALocalConflict]{ 
      \Setof{\anaMarkConstraint{\ctx}{\EMV_i}{\ECMV_i}{\STMV}}_{i<n}
    }{
        \anaMarkConstraint{\ctx}{\conflict{\EMV}}{\conflict{\ECMV}}{\STMV}
    }
\end{mathpar}

After marking, we generate type inference constraints that constrain this hole (identified using a provenance based on the location of conflict, $\ell$) using all of the conflicted terms.
\begin{mathpar}
    \inferrule[MSLocalConflict]{ 
      \Setof{\synConstraint{\ctx}{\ECMV}{\PTMV_i}{\cConstraint_i}{\mConstraint_i}}_{i<n}
    }{
      \synConstraint{\ctx}{\conflict{\ECMV}}{\TUnknown^{\mathsf{exp}(\ell)}}{\bigcup_{i<n} \cConstraint_i \cup \Setof{\constrain{\TUnknown^{\mathsf{exp}(\ell)}}{\PTMV_i}}_{i<n}}{\bigcup_{i<n} \mConstraint_i}
    }

    \inferrule[MALocalConflict]{ 
        \Setof{\anaConstraint{\ctx}{\ECMV_i}{\PTMV}{\cConstraint_i}{\mConstraint_i}}_{i<n}
    }{ 
      \anaConstraint{\ctx}{\conflict{\ECMV}}{\PTMV}{\bigcup_{i<n} \cConstraint_i \cup
      \Setof{\constrain{\TUnknown^{\mathsf{exp}(\ell)}}{\PTMV}}}{\bigcup_{i<n} \mConstraint_i}
    }
\end{mathpar}

 When the conflicted terms have a consistent type, inference will succeed. When it fails, we can fall back to the unknown type and rely on the type hole inference system developed in prior work to allow users to choose from partial hole solutions. When a partial solution is selected (effectively annotating the conflict), the mode becomes analytic, and the conflicted terms are remarked, allowing users to identify where errors would arise if the conflict were resolved at a particular type.

% \todo{check rule names}

\subsubsection{Relocation and Unicycle Conflict References}
%
Relocation and unicycle conflict references also synthesize the unknown type and operate similarly: 
\begin{mathpar}
    \inferrule[MKSRelocationConflict]{ }{
        \synMarkConstraint{\ctx}{\multiref}{\multiref}{\TUnknown}
    }

    \inferrule[MSRelocationConflict]{ }{
        \synConstraint{\ctx}{\multiref}{\TUnknown^{\mathsf{ref}(w)}}{\constraintNil{}}{(v, w, \ctx, \mathsf{syn})}
    }

    \inferrule[MARelocationConflict]{
    }{
      \anaConstraint{\ctx}{\multiref}{\PTMV}{\Setof{\constrain{\TUnknown^{\mathsf{ref}(v)}}{\PTMV}}}{(v, w ,\ctx, \mathsf{ana}(\PTMV))}
    }
\end{mathpar}
\begin{mathpar}
   \inferrule[MKSCycleLocationConflict]{ }{
    \synMarkConstraint{\ctx}{\uniref}{\uniref}{\TUnknown}
  }

  \inferrule[MSUnicycleConflict]{ }{
    \synConstraint{\ctx}{\uniref}{\TUnknown^{\mathsf{ref}(w)}}{\constraintNil{}}{(v ,w, \ctx, \mathsf{syn})}
  }

  \inferrule[MAUnicycleConflict]{ 
    }{
      \anaConstraint{\ctx}{\uniref}{\PTMV}{\Setof{\constrain{\TUnknown^{\mathsf{ref}(v)}}{\PTMV}}}{(v, w ,\ctx, \mathsf{ana}(\PTMV))}
    }
\end{mathpar}

Because the referenced term has multiple possible locations, it also has multiple possible typing contexts and typing modes, so we cannot immediately mark it.
Instead, we gather this information in the location conflict context, $L$, as shown above. This maps a location (identified by an edge UID, $w$, and also for the vertex, $v$, for operational simplicity) to a pair of a typing context, $\Gamma$, and a typing mode, $m$. 
This information can be used by the system to provisionally mark the conflicted term under an explicitly selected context and mode (e.g. in response to which location the user's cursor is on or using some other user interface affordance).

The emitted constraints assume an unknown type for the referenced vertex and constrain it with any expected types that appear at any of the locations where the corresponding term appears. Again, solutions to this unknown type can be presented to the user to help them decide which location might be most sensible.

% The user interface design space for these kinds of affordances is large, so we do not aim to contribute a particular design here. Instead, our intention here was to demonstrate that conflicts can give rise to an interesting static semantics by incorporating ideas from bidirectional typing, gradual typing, and type inference.

\subsection{Mechanized Metatheory}
\label{sub:marking-agda}
The key meta-theoretic property that tells us that we did not miss any cases is \emph{totality}:
% \todo{add marked typing to right side of each clause below}

% TODO: totality statement\todo{totality statement}

\begin{theorem}[name=Marking Totality] \
  \begin{enumerate}
    \item For all $\ctx$ and $\EMV$, there exist $\ECMV$ and $\STMV$ such that
      $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\STMV}$ and $\synConstraint{\ctx}{\ECMV}{\PTMV}{\cConstraint}{\mConstraint}$ and $\provErase{\PTMV} = \STMV$.
    % \item For all $\ctx$ and $\ELV$, there exist $\ELMV$ and $\STMV$ such that
      % $\ctxSynFixedInto{\ctx}{\ELV}{\ELMV}{\STMV}$ and $\synConstraint{\ctx}{\ELMV}{\PTMV}{\cConstraint}{\mConstraint}$ and $\provErase{\PTMV} = \STMV$.
    \item For all $\ctx$, $\EMV$, and $\STMV$, there exists $\ECMV$ such that
      $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\STMV}$ and $\anaConstraint{\ctx}{\ECMV}{\PTMV}{\cConstraint}{\mConstraint}$ and $\provErase{\PTMV} = \STMV$.
    % \item For all $\ctx$, $\ELV$, and $\STMV$, there exists $\ELMV$ such that
      % $\ctxAnaFixedInto{\ctx}{\ELV}{\ELMV}{\STMV}$ and $\anaConstraint{\ctx}{\ELMV}{\PTMV}{\cConstraint}{\mConstraint}$ and $\provErase{\PTMV} = \STMV$.
  \end{enumerate}
\end{theorem}

In addition, the prior work defined a number of other auxiliary metatheorems that help sanity check these definitions: well-formedness (marking preserves syntactic structure), and marking unicity (marking is deterministic).

We have mechanized our extension of the marked lambda calculus and these metatheorems using the Agda proof assistant, taking the standard approach of modeling judgments as inductive datatypes and inference rules as constructors. Marked terms are intrinsically typed, allowing $L$ to be computed as a function of the term.

% TODO discuss how no type hole inference/constraint generation so 
% TODO discuss how it is separate from the grove agda but a correspondence can be proven?
% \begin{figure}
% \label{fig:unmarked-judgments}
% \caption{Typing Unmarked Expressions}
%     \judgbox{\ctxSynTypeU{\ctx}{\EMV}{\sigma}} 
%     \judgbox{\ctxSynTypeU{\ctx}{\ELV}{\sigma}}
%     \judgbox{\ctxAnaTypeU{\ctx}{\EMV}{\sigma}} 
%     \judgbox{\ctxAnaTypeU{\ctx}{\ELV}{\sigma}}
%     \centering   
% \end{figure}

% \begin{figure}
%     \[
%     \begin{array}{lclll}
%      \EMV \in & Exp & \coloneqq & 
%         \var{x}
%         \mid  \num 
%         \mid \plus{\ELV_1}{\ELV_2} 
%         \mid \mult{\ELV_1}{\ELV_2}
%         \mid \lam{x}{\TLMV}{\ELV}
%         \mid \app{\ELV_1}{\ELV_2}
%         \mid \multiref
%         \mid \uniref \\
%      \ELV \in & ExpList & \coloneqq & 
%         \ehole
%         \mid \lexp{\EMV} 
%         \mid \conflict{\EMV} \\
%     \TMV \in & \Typ & \coloneqq & 
%         \tnum 
%         \mid \tarrow 
%         \mid \multiref 
%         \mid \uniref \\ 
%     \TLMV \in & TypList & \coloneqq &
%         \ehole
%         \mid \lexp{\TMV}
%         \mid \conflict{\TMV} \\
%     \end{array}
%     \]
%     \centering
%     \caption{Unmarked Language}
%     \label{fig:unmarked-language}
% \end{figure}

% \begin{figure}
%     \[
%     \begin{array}{lclll}
%     \ECMV \in & MExp & \coloneqq &
%         \var{x}
%         \mid \num
%         \mid \plus{\ELMV_1}{\ELMV_2}
%         \mid \mult{\ELMV_1}{\ELMV_2}
%         \mid \lam{x}{\TLMV}{\ELV}
%         \mid \app{\ELMV_1}{\ELMV_2}
%         \mid \multiref
%         \mid \uniref \\ 
%         & & & \ECFree{\var{x}} 
%         \mid \incontype
%         \mid \lamasc
%         \mid \lamanamarr
%         \mid \apsynmarr \\ 
%         % \mid \lamge \\
%      \ELMV \in & MExpList & \coloneqq &
%         \ehole
%         \mid \lexp{\ECMV}
%         \mid \conflict{\ECMV} \\
%     \end{array}
%     \]
%     \centering
%     \caption{Marked Language}
%     \label{fig:marked-language}
% \end{figure}

% \begin{figure}
%     \[
%     \begin{array}{lclll}
%     \sigma \in & STyp & \coloneqq & 
%         \TUnknown
%         \mid \tNum{}
%         \mid \starrow \\
%     \end{array}
%     \]
%     \judgbox{\graphErase{\TLMV} = \sigma}
%     \[\begin{array}{rrcl}
%         \Prov & \Provp & \Coloneqq & u \mid exp(u) \mid \rightarrow_L(\Provp) \mid \rightarrow_R(\Provp)\\
%         \TMName  & \TMV  & \Coloneqq & \dots \mid \TUnknown^{p} \mid \TUnknown^{(v,p)}\\
%     \end{array}\]
%     \judgbox{\ensuremath{\matchedArrowConstraint{\sigma}{\sigma_1}{\sigma_2}{C}}}
%     \centering
%     \caption{Other machinery}
%     \label{fig:marking-machinery}
% \end{figure}

% \begin{figure}
%     \[
%     \begin{array}{lclll}
%     \ECMV \in & MExp & \coloneqq &
%         \var{x}
%         \mid \num
%         \mid \plus{\ELMV_1}{\ELMV_2}
%         \mid \mult{\ELMV_1}{\ELMV_2}
%         \mid \lam{x}{\TLMV}{\ELV}
%         \mid \app{\ELMV_1}{\ELMV_2}
%         \mid \multiref
%         \mid \uniref \\ 
%         & & & \ECFree{\var{x}} 
%         \mid \incontype
%         \mid \lamasc
%         \mid \lamanamarr
%         \mid \apsynmarr \\ 
%         % \mid \lamge \\
%      \ELMV \in & MExpList & \coloneqq &
%         \ehole
%         \mid \lexp{\ECMV}
%         \mid \conflict{\ECMV} \\
%     \end{array}
%     \]
%     \centering
%     \caption{Marked Language}
%     \label{fig:marked-language}
% \end{figure}

% \begin{figure}
%     \judgbox{\synMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV}}
%     \judgbox{\synMarkConstraint{\ctx}{\ELV}{\ELMV}{\STMV}}
%     \judgbox{\anaMarkConstraint{\ctx}{\EMV}{\ECMV}{\STMV}}
%     \judgbox{\anaMarkConstraint{\ctx}{\ELV}{\ELMV}{\STMV}}
%     \centering
%     \caption{Marking Expressions}
%     \label{fig:marking-judgments}
% \end{figure}

### 2024-10-25 OriginalSubmissionSource/related-work.tex
Updates from Overleaf

--- 2024-07-11 related-work.tex
@@
 
-Contrast with a traditional diff
+Collaboration requires defining a patch language, a method for synthesizing patches from user actions, and an approach for merging patches.
 
-Similar to how git merged across moves.
-A notable difference though, is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
+\subsection{Patch Languages}
+There have been many of different designs for patch languages and indeed many imperative data structures and their associated operations can be construed as patch languages in the most general sense.
+In the context of collaborative coding, patch languages can differ in the data structures they operate on (e.g.
+line-based text, character-based text, tree-structured data).
+They can also differ in how they identify locations within the data (e.g. by using numeric offsets, unique identifiers, or paths through a tree).
+Finally, patch languages differ in which specific actions are supported explicitly.
+Insertion and deletion are common, while code relocation, copying, undo, and other operations are variously also included.
 
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
+In this paper, our focus was on syntax trees with holes (i.e.
+program sketches~\cite{DBLP:conf/aplas/Solar-Lezama09, DBLP:conf/popl/OmarVHAH17}) and explicit conflicts, which we represented as directed graphs.
+We identify locations using unique IDs.
+We have a two-level patch language, with a low level graph patch language supporting only edge insertion and deletion and a higher-level user edit action language focused on insertion, deletion, and relocation, with some additional narrative consideration of copying and undo (a fuller account of which we leave to future work).
+Our user edit language therefore forms a structure editor calculus, inspired closely by recent work on the Hazelnut structure editor calculus (which did not support relocation)~\cite{DBLP:conf/popl/OmarVHAH17} and patch languages for other tree-based data structures~\cite{DBLP:conf/sigmod/ChawatheG97, DBLP:journals/tse/FluriWPG07,DBLP:conf/kbse/FalleriMBMM14,DBLP:conf/doceng/Lindholm04,DBLP:conf/fase/NguyenNPN10,DBLP:journals/scp/SchwagerlUW15}.
 
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
+\subsection{Patch Synthesis}
+The most common approach to patch synthesis is to use a differencing algorithm to compare two states, e.g. from the file system, to generate a patch. 
+The classic \li{diff} algorithm~\cite{DiffAlgorithm}, for example, minimizes edit distance for a patch language involving line insertions and deletions.
 
-\subsection{Trees}
+In the setting of tree-based editing, there have been a number of tree differencing algorithms described in the literature \cite{DBLP:conf/esa/Klein98,DBLP:journals/tcs/Bille05,DBLP:journals/talg/DemaineMRW09,DBLP:journals/fuin/AratsuHK10,DBLP:conf/sigmod/ChawatheG97, DBLP:journals/tse/FluriWPG07,DBLP:conf/kbse/FalleriMBMM14,DBLP:conf/doceng/Lindholm04,DBLP:conf/fase/NguyenNPN10,DBLP:journals/scp/SchwagerlUW15}.
+As described in \autoref{sec:Introduction}, synthesizing insertions and deletions is well-understood, but synthesizing relocations is more complex and requires heuristics.
 
-\subsubsection{Structure editors}
+The approach we explore in this paper is far simpler: we directly translate from the log of user edit actions to graph patches, without needing a differencing algorithm at all.
+This is only possible with a structure editor integrated with the version control system, but the benefit of direct visibility into the edit action log is that we do not need heuristics to synthesize relocations.
 
-Hazel~\citep{Omar:2019:10.1145/3290327}
+\subsection{Merging}
+\subsubsection{Operational Transforms}
+The most common approach to merging concurrently developed patches is to deploy an operational transform~\cite{DBLP:conf/sigmod/EllisG89} whereby locations in a remote patch are modified based on the action of a local patch.
+Standard three-way merge algorithms in version control systems deploy this approach, as do real-time collaborative editors.
+There have been a number of papers studying the algebraic properties of merging patches in this style.
 
-\subsubsection{Tree diff}
+For example, the Darcs~\cite{DBLP:conf/haskell/Roundy05} and Pijul version control systems, like Grove, represent repositories using sets of patches.
+However, unlike Grove, not all orderings of patches are algebraically sensible (e.g.
+commutative and associative), so the theory of these systems define and algebraically characterize when operational transforms do satisfy these properties.
+Recent work on homotopical patch theory \cite{DBLP:journals/jfp/AngiuliMLH16} has similarly developed an abstract algebraic framework for distinguishing sensible merges.
+
+\subsubsection{CRDTs}
+
+The observation that commutativity is a particularly helpful property when dealing with concurrent systems, including version control systems, has led to the development of a number of data structures centered on commutativity.
+These are known as CRDTs, which stands variously for \emph{conflict-free}, \emph{convergent} (CvRDT), or \emph{commutative replicated datatypes} (CmRDT) depending on particular details about the operations and the state representation~\cite{shapiro2011conflict}.
+Our approach draws directly from this line of work: the Hazel patch language forms a CmRDT on directed graphs, from which we observe that we can derive a CmRDT for trees with explicit conflicts.
+
+There have been other recent efforts to develop CmRDTs for tree data structures~\cite{DBLP:conf/icdcs/PreguicaMSL09, Highly-Available}.
+However, the focus in this work has been on avoiding conflicts and cycles entirely by applying \emph{ad hoc} heuristics for conflict resolution at merge-time, e.g.
+using reported timestamps or favoring particular directions in the tree.
+Our approach instead embraces manual conflict resolution, as is common practice in software projects where arbitrarily losing code is not acceptable.
+
+% \url{https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read}
 
@@
 
-\subsection{Version Control}
+% \subsection{Version Control}
 
@@
 
-\subsection{Collaborative Editing}
+% \subsection{Collaborative Editing}
 
@@
 
-\subsection{Synchronization}
+% \subsection{Synchronization}
 
@@
 
-\subsection{Homotopical Patch Theory}
+% \subsection{Homotopical Patch Theory}
 
@@
 % Homotopical patch theory: \url{https://dl.acm.org/doi/10.1145/2628136.2628158}
+
+% - diff
+%     - A file comparison program 1985
+%     - myers an O(ND) difference algorithm and its variations
+% - merging
+%     - a state-of-the-art survey on software merging
+% - homotopy patch theory
+%     - 7 10 15 16 28 31
+%     - general algebraic framework for discussing patch languages
+% - darcs
+%     - unordered series of patches (but dependencies between them)
+%     - we don't focus on inversion in this paper
+%     - some properties of darcs patch theory
+%         - http://urchin.earth.li/darcs/ganesh/darcs-patch-theory/theory/formal.pdf
+%     - A formalization of darcs patch theory using inverse
+% semigroups
+%         - 2, 3, 5, 11, 18
+%         - 15 = selective undo https://ww3.math.ucla.edu/camreport/cam09-83.pdf
+%         - patch commutation requires an operational transform! https://ww3.math.ucla.edu/camreport/cam09-83.pdf
+%         - (we also define it as a semigroup)
+% - dagit - type correct changes
+%     - Type-correct changes—a safe approach to version control
+% implementation
+%     - MS thesis
+%     - Darcs: Distributed version management in haskell.
+%     - "patches that modify different parts of the code are considered, by default, independent"
+%     - patience diff
+% - A categorical theory of patches
+% - OT = "Concurrency control in groupware systems"
+%     - Formal design and
+% verification of operational transformation algorithms for copies convergence
+% - The Semantics of Version Control
+%     - https://dl.acm.org/doi/abs/10.1145/2661136.2661137?casa_token=hKqrcG9g7_UAAAAA:vN0l9AJhdnjc_2wnAIBl3TumxJBgy_4JYYSsG7522fYnbdfzh7Mpc0PIwUKcsX2VbwA0M67Cy_Y
+%     - cite for this is a version control system
+%     - CACM article by o'sullivan
+%         - https://dl.acm.org/doi/pdf/10.1145/1562164.1562183
+%         - "We are not by any means near the end
+% of the road in the evolution of revisioncontrol systems. The field has received
+% only fitful attention from academia.
+% Much work could be done on its formal foundations, which could lead to
+% more powerful and safer ways for developers to work together. Alas, I know
+% of only one notable publication on the
+% topic in the past decade.1
+%  As a simple
+% example, when merging potentially
+% conflicting changes, almost everybody
+% uses either three-way merging, which
+% is decades old, or unpublished ad hoc
+% approaches in which there is little reason to be confident."
+% - touchdevelop
+%     - https://dl.acm.org/doi/pdf/10.1145/2846661.2846672
+% - tree CRDTs
+%     - A Highly-Available Move Operation for Replicated Trees
+%         - "last writer wins" conflict resolution not explicit conflict representation
+%         - using timestamps
+%         - Martin et al 28, 29 CRDT for XML data + Kleppman and Beresford CRDT for JSON but no moves
+%         - [31] outline approaches to handling conflicts on trees but no algor
+%         - treedoc 34 also no move
+%         - 37, 38, 39 Ts for trees
+%         - 40 defines OT with a move operation but need a total order broadcast (not commutative)
+%     - Maram
+%         - The price to pay is that some move operations “lose”, i.e., have no
+% effect; achieving the same end result as previous correct approaches but at a lower cost
+%         - https://arxiv.org/pdf/2103.04828
+%     - UDR tree -- ordering
+%     - Najafzadeh et al Subtree_1 requires locks
+%     - Tao et al [17]
+%         - . Merging semantics for conflict updates in
+% geo-distributed file systems
+%         - Can lead to duplication
+%         - 
+% - tree diff based approaches
+%     - https://dl.acm.org/doi/10.1007/978-3-662-54494-5_9
+%         - 14, 15, 21, 24, 25
+%     - Change distilling: tree differencing
+% for fine-grained source code change extraction.
+%     - Fine-grained
+% and accurate source code differencing.
+%     - Tree-based Version Control in Envision
+%     - A three-way merge for XML documents
+%     - Operation-Based, Fine-Grained Version Control Model for Tree-Based Representation
+%     - https://www.sciencedirect.com/science/article/pii/S0167642315000532
+% - do not guess!
+%     - https://dl.acm.org/doi/pdf/10.1145/2508075.2508092 (brief note)
+%         - Cedalion: a language for language-oriented programming
+% - variability aware execution
+%     - https://dl.acm.org/doi/abs/10.1145/2786805.2803208
+% - text CRDTs?

### 2024-10-25 figures/figure-cycle.tex
Updates from Overleaf

@@
 \begin{figure*}
-%\hfill
-\hskip0.07\columnwidth
+\hfill
+% \hskip0.07\columnwidth
 %%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
+\begin{subfigure}[t]{0.25\textwidth}
 \centering
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
+% \path (-9em,0cm) node (node/MultiParent/ex) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiParent:e}}}};
 %%%%
 \path (0cm,0cm) node (node/Cycle/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
+% \path [draw,->,alice step] (node/MultiParent/ex) to[star] (-2em, 0cm |- node/Cycle/a);
 \end{tikzpicture}
@@
    "\vSimpleTimes" [> "\eNestedPartsBob"'] -> {
-    "\vMultiCycleTimes" [> "\eMultiCycleTimes"', >alice edge, alice node],
-    "\vMultiCyclePlus" [> "\eMultiCyclePlus", >alice edge, alice node]
+    "\vMultiCycleTimes" [> "\eMultiCycleTimes"'],
+    "\vMultiCyclePlus" [> "\eMultiCyclePlus"]
    },
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
+\begin{subfigure}[t]{0.25\textwidth}
 \centering
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.2\textwidth}
+\begin{subfigure}[t]{0.18\textwidth}
 \centering
@@
 \label{fig:Cycle:d}
-\figureCode{\parens{\hole{} * \hole{}} + $\conflictHole{\multiVertex{\vidMultiCycleTimes},\multiVertex{\vidMultiCyclePlus}}$%
+\figureCode{\parens{\hole{} * \hole{}} + $\binaryConflictHole{\multiVertex{\vidMultiCycleTimes}}{\multiVertex{\vidMultiCyclePlus}}$%
   \otherVertexVskip\multiVertex{\vidMultiCycleTimes} = \multiVertex{\vidMultiCyclePlus} * \hole{}%
@@
 %\hfill{}
-\caption{Example of Cycles.}
+\caption{(a) We start with a tree with a multiplication,  \vMultiCycleTimes{}, and addition, \vMultiCyclePlus{}, at the leaves.
+(b) Alice relocates them both, such that \vMultiCycleTimes{} is the parent of \vMultiCyclePlus{}.
+(c) Bob relocates them both, such that \vMultiCyclePlus{} is the parent of the \vMultiCycleTimes{}.
+(d) In the merged state, there is a cycle in the graph.
+Because the terms have a common parent, there is a local conflict.
+Because the cycle is connected to the rest of the graph, the cycle is broken during decomposition by relocation conflict references as shown.}
 % TODO replace screenshots with text

### 2024-10-25 figures/figure-disconnect.tex
Updates from Overleaf

--- 2021-04-09 figure-disconnect.tex
@@
 \begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
+\hfill
+% \hskip0.1\columnwidth
 %%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
+\begin{subfigure}[t]{0.20\textwidth}
 \centering
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
+% \path (-7.5em,0cm) node (node/Cycle/dx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Cycle:d}}}};
 %%%%
 \path (0cm,0cm) node (node/Disconnect/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
+% \path [draw,->,alice step] (node/Cycle/dx) to[star] (-2em, 0cm |- node/Disconnect/a);
 \end{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
+\begin{subfigure}[t]{0.22\textwidth}
 \centering
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.21\textwidth}
+\begin{subfigure}[t]{0.24\textwidth}
 \centering
@@
 \hfill{}
-\caption{Example of Disconnection.}%
+\caption{(a) We start with a tree containing \vSimpleTimes{} and \vMultiCycleTimes{}.
+(b) Alice relocates \vMultiCycleTimes{} to a child position of \vSimpleTimes{}.
+(c) Bob relocates \vSimpleTimes{} to a child position of \vMultiCycleTimes{}.
+(d) When merged, the resulting cycle becomes disconnected from the rest of the graph, forming a \emph{unicycle}.
+During decomposition, we break unicycles by leaving a \emph{unicycle conflict reference} at an arbitrarily chosen edge.}%
+\Description{This figure shows an example of a disconnected component}
 \label{fig:Disconnection}

### 2024-10-25 figures/figure-multi-parent.tex
Updates from Overleaf

--- 2021-04-09 figure-multi-parent.tex
@@
 \begin{figure*}
-%\hfill
-\hskip0.12\columnwidth
+\hfill
+% \hskip0.1\columnwidth
 %%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
+\begin{subfigure}[t]{0.12\linewidth}
 \centering
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
+% \path (-7.5em,0cm) node (node/MultiChild/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:MultiChild:c}}}};
 %%%%
 \path (0cm,0cm) node (node/MultiParent/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
+% \path [draw,->,alice step] (node/MultiChild/cx) to[star] (-2em, 0cm |- node/MultiParent/a);
 \end{tikzpicture}
@@
 \label{fig:MultiParent:a}
-\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
+\figureCode{w * \hole{} + \hole{}\otherVertexVskip\phantom{\multiVertex{58} = w}}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:b}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
+\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{58} = w}}%
 \begin{tikzpicture}
@@
 \label{fig:MultiParent:c}
-\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{18} = w}}%
+\figureCode{\hole{} * \hole{} + w\otherVertexVskip\phantom{\multiVertex{58} = w}}%
 \begin{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
+\begin{subfigure}[t]{0.18\linewidth}
 \centering
@@
 \label{fig:MultiParent:d}
-\figureCode{\hole{} * \multiVertex{18} + \multiVertex{18}\otherVertexVskip\multiVertex{18} = w}%
+\figureCode{\hole{} * \multiVertex{58} + \multiVertex{58}\otherVertexVskip\multiVertex{58} = w}%
 \begin{tikzpicture}
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.14\linewidth}
+\begin{subfigure}[t]{0.13\linewidth}
 \centering
@@
 \label{fig:MultiParent:e}
-\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{18} = w}}%
+\figureCode{\hole{} * w + \hole{}\otherVertexVskip\phantom{\multiVertex{58} = w}}%
 \begin{tikzpicture}
@@
 \hfill{}
-\caption{Example of Multi-Parent Conflicts.}%
+\caption{(a) We start in a state with a variable, \li{w}, and two holes.
+(b) Alice relocates \li{w} to the left hole.
+(c) Bob relocates \li{w} to the right hole.
+(d) After merging, vertex \vNestedPartsAlice{} has two incoming edges, i.e.
+it has a \emph{relocation conflict}.
+The corresponding decomposition leaves a \emph{relocation conflict reference} at both locations, partially addressing the \textbf{relocation conflict problem}.
+Terms that have a relocation conflict are tracked separately by decomposition.
+(e) The relocation conflict can be resolved by deleting all but one reference.}%
+\Description{This figure depicts an example of relocation conflicts}
 \label{fig:MultiParent}

### 2024-10-25 discussion-and-conclusion.tex
updates for final submission

--- 2021-04-09 grove-paper-discussion-and-conclusion.tex
@@
 \label{sec:Discussion and Conclusion}
+\begin{quote}
+    \textit{``The fact that commutation can fail [in Darcs] makes a huge difference in the whole patch formalism.
+It may be possible to create a formalism in which commutation always succeeds, with the result of what would otherwise be a commutation that fails being something like a virtual particle ...
+and it may be that such a formalism would allow strict mathematical proofs ...
+However, I’m not sure how you’d deal with a request to delete a file that has not yet been created, for example.
+Obviously you’d need to create some kind of antifile, which would annihilate with the file when that file finally got created ...''}
 
-TODO: leaves
+    -- David Roundy, Theory of patches~\cite{old-darcs-manual}
+\end{quote}
 
-TODO: our model supports treating these as a cons-list of characters
+This paper proposes a radically simpler, albeit practically ambitious, rearchitecture of collaborative editing.
+Our contributions together result in a typed collaborative structure calculus called Grove
+where, uniquely, all edits, including code relocations that stymie existing approaches, commute and where there are no semantic gaps: all possible editor states, including editor states with various kinds of unresolved conflicts, are semantically meaningful.
 
-TODO: (Place somewhere) This move semantics gives us a richer structure than
-when treating code as a list of lines.  We exploit
-this in Section~REF:TODO in order merge edits that involve moving code.
+This paper focuses on the core theoretical underpinnings of this approach,
+developing mechanized metatheory for both the patch language and the type system. 
+A number of research problems on algorithmic, networking, and user interface aspects of the problem open up given these foundations.
+For example, we point out several situations where presenting non-conflicted but heuristically attention-worthy merges may be worthwhile, and we leave to future work the user experience design of this process.
+We intend to use the Grove workbench to integrate these efforts into the Hazel programming environment, though its editor component has been evolving so rapidly as to prevent experimentation in this direction so far.
 
-TODO: Traditional diff: no relation (and indent might change)
+Although our focus was on tree editing, some aspects of a program are more naturally linearly structured, e.g.
+string literals.
+We also leave to future work the problem of combining existing work on sequence CRDTs~\cite{ahmed2011evaluating,kleppmann2020moving} with our work on tree/graph CRDTs.
 
-TODO: memory usage
+% In practice, it may be helpful to garbage collect orphaned vertices once there is consensus across collaborators that the deletion is permanent, but we do not consider this consensus protocol formally in this paper.
+In an open-ended collaboration scenario (where the set of collaborators is not known, e.g.
+on GitHub), we simply retain all vertices.
 
-TODO: cache eviction algorithm
+% TODO: talk about linear sequences in code
 
-TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
-In practice, these complex scenarios occur less frequently than portrayed here.
+% TODO: leaves
 
-\subsection{Variable names, strings, and numbers}%
-\label{sub:Variable names, strings, and numbers}
+% TODO: our model supports treating these as a cons-list of characters
 
-NOTE: we could implement each digit as a separate characters
+% TODO: (Place somewhere) This move semantics gives us a richer structure than
+% when treating code as a list of lines.  We exploit
+% this in Section~REF:TODO in order merge edits that involve moving code.
 
-GUI for string conflicts: use popups
+% TODO: Traditional diff: no relation (and indent might change)
+
+% TODO: memory usage
+
+% TODO: cache eviction algorithm
+
+% TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
+% In practice, these complex scenarios occur less frequently than portrayed here.
+
+% \subsection{Variable names, strings, and numbers}%
+% \label{sub:Variable names, strings, and numbers}
+
+% NOTE: we could implement each digit as a separate characters
+
+% GUI for string conflicts: use popups

### 2024-10-25 grove-paper.tex
updates for final submission

@@
 
-\documentclass[acmsmall,dvipsnames,10pt,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmsmall,dvipsnames,10pt,review]{acmart}\settopmatter{printfolios=true}
+% Need to tweak for camera-ready submission
+% \documentclass[acmsmall,dvipsnames,10pt]{acmart}\settopmatter{printfolios=true}
 
@@
 \input{figures/figure-user-actions}
+\input{figures/figure-user-action-transitions}
 
@@
 \orcid{0000-0003-3160-6972}
+\affiliation{
+\institution{National University of Singapore}
+\country{Singapore}
+}
 
@@
 \orcid{0000-0003-1693-6172}
+\affiliation{
+\institution{University of Michigan}
+\city{Ann Arbor}
+\state{MI}
+\country{USA}
+}
+
+\author{Thomas J. Porter}
+\orcid{0009-0000-1809-8382}
+\affiliation{
+\institution{University of Michigan}
+\city{Ann Arbor}
+\state{MI}
+\country{USA}
+}
 
 \author{Sundara Vishnu Satish}
-\orcid{TODO}
+\orcid{0009-0002-8685-1342}
+\affiliation{
+\institution{University of Michigan}
+\city{Ann Arbor}
+\state{MI}
+\country{USA}
+}
 
-\author{Thomas J. Porter}
-\orcid{TODO}
+\author{Eric Zhao}
+\orcid{0009-0000-4969-2376}
+\affiliation{
+\institution{University of Michigan}
+\city{Ann Arbor}
+\state{MI}
+\country{USA}
+}
 
@@
   %\position{Assistant Research Scientist}
-  \department[0]{Computer Science and Engineering}
-  \department[1]{Electrical Engineering and Computer Science}
-  \department[2]{College of Engineering}
+  % \department[0]{Computer Science and Engineering}
+  % \department[1]{Electrical Engineering and Computer Science}
+  % \department[2]{College of Engineering}
   \institution{University of Michigan}
-  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
+  % \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
   \city{Ann Arbor}
   \state{MI}
-  \postcode{48109-2121}
+  % \postcode{48109-2121}
   \country{USA}
@@
 \input{formalism}
+\input{implementation}
 \input{marking-system}
-\input{implementation}
 \input{related-work}
@@
   %% extraction tools.
-  This material is based upon work supported by the
+
+  This work was partially funded by the
   \grantsponsor{GS100000001}{National Science
     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+  No.~\grantnum{GS100000001}{2238744}. Any opinions, findings, and
   conclusions or recommendations expressed in this material are those
@@
 %% Appendix
-\appendix
+% \appendix
 
@@
 
-%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count
-
-\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.
-
-% \input{appendix-A-graph-sequences}
-\input{appendix/appendix-A-formalism}  
-% \input{appendix/appendix-B-marking-system}
-\input{appendix/appendix-B-typechecking}
-% \input{appendix/appendix-C-type-inference}
-
 \end{document}

### 2024-10-25 related-work.tex
updates for final submission

@@
 
-Contrast with a traditional diff
+The core components of a version control system are a patch language, a method for synthesizing patches from user actions, and an approach for merging patches. 
+% Existing patch based systems such as Darcs~\cite{DBLP:conf/haskell/Roundy05} and Pijul achieve commutativity between patches in many cases, but Grove achieves total commutativity.
 
-Similar to how git merged across moves.
-A notable difference though, is that \texttt{git} detects
-file moves by structural similarity rather than
-our nominal equality.
-(We exploit this in Subsection~REF:TODO).
+% Grove stands out from existing systems in the precision of its patch language and the simplicity of its patch synthesis and merge processes, as well as its intrinsic guarantee of commutativity by employing a graph CmRDT as its repository data structure.
 
-Compare this to what happens when there are conflicts in a traditional,
-line/diff-based version control system (e.g., \texttt{git}, \texttt{mercurial}, \texttt{svn}, etc.).
-Those systems annotate files with difference markers (e.g., sequences of~\texttt{<},~\texttt{>}, or~\texttt{=})
-when a merge conflict happens.
-These show alternate versions of the code, and it is up to the user
-to replace those alternates with the desired merged result.
+% Patch languages vary in their native data structure and level of precision, with Grove's patch language being exceptionally fine-grained and molded to the language.
+Traditional patch synthesis algorithms reconstruct patches from version differences, while Grove uniquely synthesizes patches directly from edit actions.
+Traditional merge algorithms, such as those used by Git, employ operational transforms to
 
-Another place our model is different
-is that changes are explicit rather than being inferred by a \texttt{diff} algorithm.
-Finally, in our model, since all changes are in terms of
-the structure of the code rather than lines,
-all changes respect that structure.
-This is unlike the line-based model where there is no guarantee
-that a merge conflict follows the grammatical structure of the code.
+% Traditional patch synthesis and merge algorithms, such as those used by Git, are complex and indirect, while Grove joins a family of new systems employing
 
-\subsection{Trees}
+\subsection{Patch Languages}
+There have been many different designs for patch languages and indeed many imperative data structures and their associated operations can be construed as patch languages in the most general sense.
+In the context of collaborative coding, patch languages can differ in the data structures they operate on (e.g.
+line-based text~\cite{DBLP:conf/haskell/Roundy05}, character-based text~\cite{DBLP:conf/sigmod/EllisG89,DBLP:journals/pacmhci/LittLKH22}, tree-structured data).
+They can also differ in how they identify locations within the data (e.g. by using numeric offsets~\cite{DBLP:conf/sigmod/EllisG89}, unique identifiers, or paths through a tree).
+Finally, patch languages differ in which specific actions are supported explicitly.
+Insertion and deletion are common, while code relocation, copying, undo, and other operations are variously also included.
 
-\subsubsection{Structure editors}
+In this paper, our focus was on syntax trees with holes (i.e.
+program sketches~\cite{DBLP:conf/aplas/Solar-Lezama09, DBLP:conf/popl/OmarVHAH17}) and explicit conflicts, which we represented as directed graphs.
+We identify locations using unique IDs.
+We have a two-level patch language, with a low-level graph patch language supporting only edge insertion and deletion and a higher-level user edit action language focused on insertion, deletion, and relocation, with some additional narrative consideration of copying and undo (a fuller account of which we leave to future work).
+Our user edit language therefore forms a structure editor calculus, inspired closely by recent work on the Hazelnut structure editor calculus (which did not support relocation)~\cite{DBLP:conf/popl/OmarVHAH17} and patch languages for other tree-based data structures~\cite{DBLP:conf/sigmod/ChawatheG97, DBLP:journals/tse/FluriWPG07,DBLP:conf/kbse/FalleriMBMM14,DBLP:conf/doceng/Lindholm04,DBLP:conf/fase/NguyenNPN10,DBLP:journals/scp/SchwagerlUW15}.
 
-Hazel~\citep{Omar:2019:10.1145/3290327}
+\subsection{Patch Synthesis}
+The most common approach to patch synthesis is to use a differencing algorithm to compare two states, e.g. from the file system, to generate a patch. 
+The classic \li{diff} algorithm~\cite{DiffAlgorithm}, for example, minimizes edit distance for a patch language involving line insertions and deletions.
 
-\subsubsection{Tree diff}
+In the setting of tree-based editing, there have been a number of tree differencing algorithms described in the literature \cite{DBLP:conf/esa/Klein98,DBLP:journals/tcs/Bille05,DBLP:journals/talg/DemaineMRW09,DBLP:journals/fuin/AratsuHK10,DBLP:conf/sigmod/ChawatheG97, DBLP:journals/tse/FluriWPG07,DBLP:conf/kbse/FalleriMBMM14,DBLP:conf/doceng/Lindholm04,DBLP:conf/fase/NguyenNPN10,DBLP:journals/scp/SchwagerlUW15}.
+As described in \autoref{sec:Introduction}, synthesizing insertions and deletions is well-understood, but synthesizing relocations is more complex and requires heuristics.
+
+The approach we explore in this paper is far simpler: we directly translate from the log of user edit actions to graph patches, without needing a differencing algorithm at all.
+This is only possible with a structure editor integrated with the version control system, but the benefit of direct visibility into the edit action log is that we do not need heuristics to synthesize relocations.
+
+\subsection{Merging}
+\subsubsection{Operational Transforms}
+The most common approach to merging concurrently developed patches is to deploy an operational transform~\cite{DBLP:conf/sigmod/EllisG89} whereby locations in a remote patch are modified based on the action of a local patch.
+Standard three-way merge algorithms in version control systems deploy this approach, as do real-time collaborative editors.
+There have been a number of papers studying the algebraic properties of merging patches in this style.
+
+For example, the Darcs~\cite{DBLP:conf/haskell/Roundy05} version control system, like Grove, represents repositories using sets of patches.
+Using operational transforms, Darcs achieves commutativity in many cases, but not between conflicting patches.
+The theory of Darcs defines and algebraically characterizes when operational transforms do satisfy the properties of associativity and commutativity.
+Recent work on homotopical patch theory \cite{DBLP:journals/jfp/AngiuliMLH16} has similarly developed an abstract algebraic framework for distinguishing sensible merges.
+
+\subsubsection{CRDTs}
+
+The observation that commutativity is a particularly helpful property when dealing with concurrent systems, including version control systems, has led to the development of a number of data structures centered on commutativity.
+These are known as CRDTs, which stands variously for \emph{conflict-free}, \emph{convergent} (CvRDT), or \emph{commutative replicated datatypes} (CmRDT) depending on particular details about the operations and the state representation~\cite{shapiro2011conflict}.
+Our approach draws directly from this line of work: the Grove patch language forms a CmRDT on directed graphs, from which we observe that we can derive a CmRDT for trees with explicit conflicts.
+
+Other examples of CRDT-based editors include and Peritext~\cite{DBLP:journals/pacmhci/LittLKH22} and Zed~\cite{zed-blog}.
+These systems differ from Grove by targeting synchronous editing rather than version control and by operating on sequential text rather than tree structures.
+
+There have been other recent efforts to develop CmRDTs for tree data structures~\cite{DBLP:conf/icdcs/PreguicaMSL09, Highly-Available}.
+However, the focus in this work has been on avoiding conflicts and cycles entirely by applying \emph{ad hoc} heuristics for conflict resolution at merge-time, e.g.
+using reported timestamps or favoring particular directions in the tree.
+Our approach instead embraces manual conflict resolution, as is common practice in software projects where arbitrarily losing code is not acceptable.
+
+% Zed is a Integrated Development Environment (IDE) that reasons text edits using CRDTs and differs in approach by using Lamport timestamps to order events across replicas.
+Peritext, on the other hand, is a character based CRDT editor but is limited to the realm of rich-text editing.
+
+Pijul iterates on the patch based system of Darcs, obtaining commutativity between conflicting patches using a CRDT graph data structure~\cite{pijul-talk}.
+Pijul's graph data structure is very similar to that of Grove in its treatment of edges, but, unlike Grove, requires vertices to be created before they can be referred to.
+This imposes a dependency relation between patches, causing Pijul to fall short of the full commutativity enjoyed by Grove.
+Pijul is also language agnostic and models only the linear structure of text.
+However, Pijul goes beyond this version of Grove by extending the data structure to support history and branches, and Grove may be similarly extended in future work.
+
+% \url{https://www.waitingforcode.com/big-data-algorithms/conflict-free-replicated-data-types-flags-graphs-maps/read}
 

### 2024-11-22 grove-paper.tex
camera ready

@@
 
-\documentclass[acmsmall,dvipsnames,10pt,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmsmall,dvipsnames,10pt]{acmart}\settopmatter{printfolios=true}
 % Need to tweak for camera-ready submission
@@
 
+%%% The following is specific to POPL '25 and the paper
+%%% 'Grove: A Bidirectionally Typed Collaborative Structure Editor Calculus'
+%%% by Michael D. Adams, Eric Griffis, Thomas J. Porter, Sundara Vishnu Satish, Eric Zhao, and Cyrus Omar.
+%%%
+\setcopyright{rightsretained}
+\acmDOI{10.1145/3704909}
+\acmYear{2025}
 \acmJournal{PACMPL}
-\acmVolume{1}
+\acmVolume{9}
 \acmNumber{POPL}
-\acmArticle{1}
-\acmYear{2025}
-
+\acmArticle{73}
 \acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-\setcopyright{none}
+\received{2024-07-11}
+\received[accepted]{2024-11-07}
 
@@
 
-%% Title information
-\title[Grove]{Grove: A Bidirectionally Typed Collaborative Structure Editor Calculus}
-%\subtitle{Subtitle}
+\title{Grove: A Bidirectionally Typed Collaborative Structure Editor Calculus}
 
-%% Author information
 \author{Michael D. Adams}
 \orcid{0000-0003-3160-6972}
-\affiliation{
+\affiliation{%
   \institution{National University of Singapore}
+  \city{Singapore}
   \country{Singapore}
 }
+\email{adamsmd@nus.edu.sg}
 
@@
 \orcid{0000-0003-1693-6172}
-\affiliation{
+\affiliation{%
   \institution{University of Michigan}
   \city{Ann Arbor}
-\state{MI}
   \country{USA}
 }
+\email{egriffis@umich.edu}
 
@@
 \orcid{0009-0000-1809-8382}
-\affiliation{
+\affiliation{%
   \institution{University of Michigan}
   \city{Ann Arbor}
-\state{MI}
   \country{USA}
 }
+\email{thomasjp@umich.edu}
 
@@
 \orcid{0009-0002-8685-1342}
-\affiliation{
+\affiliation{%
   \institution{University of Michigan}
   \city{Ann Arbor}
-\state{MI}
   \country{USA}
 }
+\email{svishnus@umich.edu}
 
@@
 \orcid{0009-0000-4969-2376}
-\affiliation{
+\affiliation{%
   \institution{University of Michigan}
   \city{Ann Arbor}
-\state{MI}
   \country{USA}
 }
+\email{zzhaoe@umich.edu}
 
@@
 \orcid{0000-0003-4502-7971}
-\affiliation{
-  %\position{Assistant Research Scientist}
-  % \department[0]{Computer Science and Engineering}
-  % \department[1]{Electrical Engineering and Computer Science}
-  % \department[2]{College of Engineering}
+\affiliation{%
   \institution{University of Michigan}
-  % \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
   \city{Ann Arbor}
-  \state{MI}
-  % \postcode{48109-2121}
   \country{USA}
 }
+\email{comar@umich.edu}
 
@@
 % Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-% \begin{CCSXML}
-% <ccs2012>
-% <concept>
-% <concept_id>10011007.10011006.10011008</concept_id>
-% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-% <concept_significance>500</concept_significance>
-% </concept>
-% <concept>
-% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-% <concept_significance>300</concept_significance>
-% </concept>
-% </ccs2012>
-% \end{CCSXML}
-% \ccsdesc[500]{Software and its engineering~General programming languages}
-% \ccsdesc[300]{Social and professional topics~History of programming languages}
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10011007.10011006.10011008</concept_id>
+       <concept_desc>Software and its engineering~General programming languages</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011039.10011040</concept_id>
+       <concept_desc>Software and its engineering~Syntax</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011071</concept_id>
+       <concept_desc>Software and its engineering~Software configuration management and version control systems</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Software and its engineering~General programming languages}
+\ccsdesc[500]{Software and its engineering~Syntax}
+\ccsdesc[500]{Software and its engineering~Software configuration management and version control systems}
 % End of generated code
@@
 %% Keywords
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+\keywords{version control systems, collaborative editing, CRDTs, gradual typing}
 

### 2024-11-22 introduction.tex
camera ready

@@
 Let us briefly review some classic problems with these systems.
-Appendix~\ref{appendix:merge-problems} shows examples of the problematic default behavior of Git in each of these scenarios.\todo{write this, can move to supplement if needed}
+The supplemental material includes Git repositories that demonstrate each of these problems.
+%Appendix~\ref{appendix:merge-problems} shows examples of the problematic default behavior of Git in each of these scenarios.\todo{write this, can move to supplement if needed}
 
@@
 arises when both commits make edits to different locations within a single line of code. For example, one patch might rename a function argument while the other adds a new argument. 
-Similarly, if one patch renames an type while another 
+Similarly, if one patch renames a type while another 
 makes unrelated changes to code that references that type,
 there will be conflicts at every line of code shared between the two patches. 
-% A notable special case is the \textbf{wrapping problem}, which arises when one patch changes the nesting of code structures, e.g.
+A notable special case is the \textbf{nesting problem}, which arises when one patch changes the nesting of code structures, e.g.
 by wrapping a code block within a new control flow construct, thereby changing the indentation of every line in the block.
@@
 A common approach is to enrich the patch language to make code relocation a primitive command.
-However, correctly synthesizing code relocation commands given only the initial and final state of the repository requires heuristics. 
+However, correctly synthesizing code relocation commands given only the initial and final states of the repository requires heuristics. 
 The typical approach is to assume that a matching deletion and insertion is due to code relocation. However, relocated code is often also modified. 
@@
 a syntax tree with holes, shown projected visually in various ways to the developer.
-Structure editing has been studied since the 1980s with the Cornell Program Synthesizer 
-and research continues to this day, with numerous active projects including Scratch, Snap!, and other block-based editors (which are widely used in educational and end-user programming settings), Jetbrains MPS (which has been deployed in industry), and Hazel (a live functional programming environment rooted in a structure editor calculus called Hazelnut, which serves as
-an active research platform).
+Structure editing has been studied since the 1980s with the Cornell Program Synthesizer~\cite{DBLP:journals/cacm/TeitelbaumR81}
+and research continues to this day, with numerous active projects including Scratch~\cite{maloney2010scratch} and other block-based editors (which are widely used in educational and end-user programming settings), Jetbrains MPS (which has been deployed in industry)~\cite{voelter2011language}, and Hazel (a live functional programming environment rooted in a structure editor calculus called Hazelnut, which serves as
+an active research platform)~\cite{DBLP:conf/popl/OmarVHAH17}.
 
@@
 \li{Plus}.
-An edge is also labeled with a UID and establishes a parent-child relationships at a labeled \emph{position} for the parent vertex's constructor, e.g.
+An edge is also labeled with a UID and establishes a parent-child relationship at a labeled \emph{position} for the parent vertex's constructor, e.g.
 at the \li{L} or \li{R} position of the \li{Plus} constructor.
@@
 To ensure commutativity, deletion of an edge is permanent, i.e.
-the edges form a two-phase set (2P-Set) CmRDT. The edit actions that users perform are given meaning by a straightforward translation to a graph patch. 
+the edges form a two-phase set (2P-Set) CmRDT~\cite{shapiro2011conflict}. The edit actions that users perform are given meaning by a straightforward translation to a graph patch. 
 Relocation simply translates to 
@@
 Developers are left to reason without the aid of much of their tooling during conflict resolution.
-This is an instance of the more general \textbf{semantic gap problem} when programming tools encounter incomplete programs~\cite{HazelSNAPL}.
+This is an instance of the more general \textbf{semantic gap problem} when programming tools encounter incomplete programs~\cite{DBLP:conf/snapl/OmarVHSGAH17}.
 
 The previous work on the Hazelnut structure editor calculus addresses the semantic gap problem in the single-user setting by 
-defining a type system and type error localization system (the \emph{marked lambda calculus}) for incomplete programs, i.e. programs with holes. 
+defining a type system and type error localization system (the \emph{marked lambda calculus}) for incomplete programs, i.e. programs with holes~\cite{DBLP:journals/pacmpl/ZhaoMDBPO24}. 
 Notably, type error localization is proven to be \emph{total}, i.e.
 the system is able to assign static meaning to \emph{every} syntactically well-formed expression by inserting \emph{marks} to localize errors.
-Marking employs local type inference as codified by a \emph{bidirectional type system} as well as \emph{gradual typing}, i.e.
+Marking employs local type inference as codified by a \emph{bidirectional type system}~\cite{10.1145/3450952} as well as \emph{gradual typing}~\cite{siek2015}, i.e.
 the theory of \emph{type holes}, to recover from situations where type errors make it impossible to
@@
 
-This paper extends this work to develop a total type error localization system for groves, introduced in Contribution 1 above as the result of decomposing a commutatively edited graph into a set of terms with empty holes, local conflicts, location conflict references, and cycle conflict references.
+This paper extends this prior work on the marked lambda calculus to develop a total type error localization system for groves, introduced in Contribution 1 above as the result of decomposing a commutatively edited graph into a set of terms with empty holes, local conflicts, relocation conflict references, and unicycle conflict references.
 This paper develops a type (and type error localization) discipline for handling these novel constructs.

### 2024-11-22 sections/abstract.tex
directory restructuring and hiding appendices

--- 2024-07-11 abstract.tex
@@
 vertices and edges. All edits amount to edge insertion and deletion, with deletion being permanent. 
-To support tree-based editing, we define a decomposition from graphs into \emph{groves}, which are a set of syntax trees with conflicts---including multi-child, multi-parent, and cycle conflicts---represented explicitly using holes and references between trees.
+To support tree-based editing, we define a decomposition from graphs into \emph{groves}, which are a set of syntax trees with conflicts---including local, relocation, and unicyclic relocation conflicts---represented explicitly using holes and references between trees.
 Finally, we define a type error localization system for groves that enjoys a \emph{totality} property, i.e.

### 2024-11-22 sections/formalism.tex
directory restructuring and hiding appendices

--- 2024-10-25 OriginalSubmissionSource/formalism.tex
@@
 
-We now formally consider the ideas presented informally above by defining the Grove Calculus.
+We now formalize the ideas presented informally above by defining the Grove Calculus.
 The calculus consists of a commutative patch language on graphs, defined in \autoref{sub:Convergent Graphs}, a decomposition procedure to go from graphs to groves (i.e.
@@
 
+For example, if the language is the simply typed lambda calculus, $\K$ will include constructors for function abstractions ($\textsc{Lambda}_x$), function applications ($\textsc{Ap}$), variables ($\textsc{Var}_x$), and constants ($\textsc{Const}_c$), as well as function type constructors ($\textsc{Arrow}$) and base types ($\textsc{Base}_b$).
+The arity of each constructor will reflect the child positions of these constructors, e.g.
+$\mathsf{arity}(\textsc{Ap}) = \{\textsc{Function} , \textsc{Argument}\}$, $\mathsf{arity}(\textsc{Lambda}_x) = \{\textsc{Annotation} , \textsc{Body}\}$, and $\mathsf{arity}(\textsc{Var}_x) = \emptyset$.
+
 A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is labeled with a UID, $u$, and a constructor, $k$.
 
-A \emph{location} $\ell = (v, p) \in \L$, represents the origin of an edge, where $v=(u,k)$ is drawn from $\V$ and position $p$ is drawn from $\mathsf{arity}(k)$.
+A \emph{location} $\ell = (v, p) \in \L$ represents the origin of an edge, where $v=(u,k)$ is drawn from $\V$ and position $p$ is drawn from $\mathsf{arity}(k)$.
 
-An \emph{edge} $\e{=}(w, \ell, v) \in \E = \W \times \L \times \V$ represents a directed multi-edge 
+An \emph{edge} $\e = (w, \ell, v) \in \E = \W \times \L \times \V$ represents a directed multi-edge 
     identified by $w$, 
@@
 
-A \emph{graph} $g : \E \rightarrow \Sigma$ is a function from edges to \emph{edge states} $\Sigma = \SetOf{\bot, \Plus, \Minus}$.
+A \emph{graph} $g : \E \rightarrow \Sigma$ is a function from edges to \emph{edge states} $s \in \Sigma = \SetOf{\bot, \Plus, \Minus}$.
 If $g(\e) = \bot$, then $\e$ has not yet been constructed in the graph.
-If $g(\e) = \Plus$ then $\e$ is \textit{live} in the graph, meaning it is relevant to the user and the type system.
+If $g(\e) = \Plus$ then $\e$ is live in the graph.
 If $g(\e) = \Minus$, then $\e$ has been deleted and cannot be constructed again.
 The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.
-We assume a graph assigns only finitely many edges to $\Plus$ to ensure the finiteness of the data structure on which further computation is done.
+The state of each edge can only progress along this ordering over time, from not yet constructed, to live, to deleted.
+The \textit{realized} edges in a graph are those assigned to $\Plus$ or $\Minus$.
+We assume that there are finitely many realized edges in a graph.
 
@@
 We define a graph patch command $\pi = (s,\e) \in \left\{\Plus, \Minus\right\} \times \E$
-as a pair of an edge state (excluding $\bot$) and an edge. As a shorthand, we define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.
+as a pair of an edge state (excluding $\bot$) and an edge. As a shorthand, we use $+\e$ to denote the construction command $(\Plus, \e)$ and $-\e$ to denote the destruction command $(\Minus, \e)$.
 
 For $s_1$, $s_2 \in \Sigma$, we define the join operation $s_1 \sqcup s_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
-Accordingly, $\Plus \sqcup \bot = \Plus$ means that edges can only be created if they do not already exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once an edge is deleted it can never be restored, and hence deletion is idempotent.
+Accordingly, $\Plus \sqcup \bot = \Plus$ means that an edge can be created if it has never existed, and $\Plus \sqcup \Minus = \Minus$ means that once an edge is deleted it can never be restored.
 
@@
 Essentially, these patch command semantics act as a transition system between graphs, \emph{joining} the new edge state with the corresponding edge state of the graph to which the patch command is being applied.
-Since a patch command can only change the value of $g$ on one input, applying the patch command preserves the property that $g$ maps only a finite set of edges to $\Plus$.
+Since a patch command can only change the value of $g$ on one input, applying the patch command preserves the property that $g$ maps all but finitely many edges to $\bot$.
 
@@
 \begin{proof}
-    Follows directly from the total ordering $(\Sigma, \sqsubset)$
+    % Since $\Sigma$ is finite, 
+    Follows directly from the total ordering $(\Sigma, \sqsubset)$. 
 \end{proof}
@@
 The definition above can be understood operationally as a CmRDT.
-In particular, the vertices can be understood as forming a Grow-Only Set (G-Set).
-The edges can be understood as forming a two-phase set (2P-Set), because deletion is permanent~\cite{shapiro2011conflict}.
+In particular, the edges can be understood as forming a two-phase set (2P-Set), because deletion is permanent~\cite{shapiro2011conflict}.
+The set of realized edges forms a grow-only set, as does the set of realized vertices (those included in a realized edge).
 
@@
 
-To help define decomposittion, each vertex in a graph can be classified as either an NP root (if it has no parents), an MP root (if it has multiple parents), a U root (if there is a sequence of unique parents from it to itself, and it has the minimal UID in that sequence), or not a root.
+To help define decomposition, each vertex in a graph can be classified as either an NP root (if it has no parents), an MP root (if it has multiple parents), a U root (if there is a sequence of unique parents from it to itself, and it has the minimal UID in that sequence), or not a root.
 These classes are mutually exclusive.
@@
 \begin{definition}
-    A vertex is an \textit{NP root} in $G$ if $\mathsf{parents}_{G}(v)=\emptyset$
+    A vertex $v$ is an \textit{NP root} in $G$ if $\mathsf{parents}_{G}(v)=\emptyset$
 \end{definition}
@@
 \begin{definition}
-    A vertex is an \textit{MP root} in $G$ if $\abs{\mathsf{parents}_{G}(v)} > 1$
+    A vertex $v$ is an \textit{MP root} in $G$ if $\abs{\mathsf{parents}_{G}(v)} > 1$
 \end{definition}
@@
 \begin{definition}
-    A vertex is a \textit{U root} in $G$ if $\exists \{v_i\}_{1 \leq i \leq n}$ such that $v_1 = v_n = v$, $\mathsf{parents}_{G}(v_i)=\{v_{i+1}\}$ for all $1 \leq i < n$, and the UID of $v$ is the minimum of the UIDs of $\{v_i\}_{1 \leq i \leq n}$.
+    A vertex $v$ is a \textit{U root} in $G$ if $\exists \{v_i\}_{1 \leq i \leq n}$ such that $v_1 = v_n = v$, $\mathsf{parents}_{G}(v_i)=\{v_{i+1}\}$ for all $1 \leq i < n$, and the UID of $v$ is the minimum of the UIDs of $\{v_i\}_{1 \leq i \leq n}$.
 \end{definition}
@@
 
-\begin{lemma}
+\begin{lemma}[Unique Classification]
 \label{lem:uniq-classification}
-    (Unique Classification) For any live subgraph $G\subseteq \E$ and vertex $v\in V$, exactly one of the following holds: 
+    For any live subgraph $G\subseteq \E$ and vertex $v\in V$, exactly one of the following holds: 
     \begin{itemize}
@@
 
-Formally, the decomposition of a vertex is given by the function $\mathsf{decompVertex}$ a term with the same UID and constructor as the vertex.
+Formally, the decomposition of a vertex is given by the function $\mathsf{decompVertex}$, which returns a term with the same UID and constructor as the vertex.
 At each position in the arity of the constructor, we obtain a child term from the function $\mathsf{decompLoc}$, which decomposes a location in the graph to a corresponding child term.
@@
     \mathsf{decompLoc}_{G}(\ell) &= \begin{cases}
-        \ehole &\quad \mathsf{children}_{G}(\ell) = \emptyset \\ \eid{w}{}{\mathsf{decompChild}_G(w, v')} &\quad \mathsf{children}_{G}(\ell) = \SetOf{(w,v')}\\ \conflictHoleTerm{\ell}{\eid{w}{}{\mathsf{decompChild}_G(w, v')} \SuchThat{(w,v') \in \mathsf{children}_{G}(\ell)}} &\quad \abs{\mathsf{children}_{G}(\ell)} > 1 \\
+        \ehole &\quad \mathsf{children}_{G}(\ell) = \emptyset \\ \eid{w}{}{\mathsf{decompChild}_G(w, v')} &\quad abs{\mathsf{children}_{G}(\ell)} = 1\\ \conflictHoleTerm{\ell}{\eid{w}{}{\mathsf{decompChild}_G(w, v')} \SuchThat{(w,v') \in \mathsf{children}_{G}(\ell)}} &\quad \abs{\mathsf{children}_{G}(\ell)} > 1 \\
     \end{cases}
@@
 \begin{definition}
-    \mathsf{vertices}(G) = \SetOf{v \in \V \SuchThat{(w, (v,p), v') \in G \land w \in \W \land v' \in \V \land p \in \P}}
+    $\mathsf{vertices}(G) = \SetOf{v \in \V \SuchThat{(w, (v,p), v') \in G \land w \in \W \land v' \in \V \land p \in \P}}$
 \end{definition}
@@
 \begin{definition}
-    $\mathsf{recomp}(T) = \bigup{\termMV \in T}\mathsf{recompTerm}(\termMV)$
+    $\mathsf{recomp}(T) = \bigcup_{\termMV \in T}\mathsf{recompTerm}(\termMV)$
 \end{definition}
@@
 
-\begin{theorem}
+\begin{theorem}[Recomposability]
 \label{thm:Recomposability}
-    (Recomposability) For all subgraphs $G$, $\mathsf{recomp}(\mathsf{decomp}(G))=G$.
+    For all subgraphs $G$, $\mathsf{recomp}(\mathsf{decomp}(G))=G$.
 \end{theorem}
@@
 
-Formally, a partitioned grove $\gamma$ is a quadruple $(\subtermMV_r, NP, MP, U)$ where $\subtermMV_r \in ChildTerm$ and $NP, MP, U$ are finite sets of terms.
+Formally, a partitioned grove $\gamma$ is a quadruple $(\subtermMV_r, NP, MP, U)$ where $\subtermMV_r \in \textit{ChildTerm}$ and $NP, MP, U$ are finite sets of terms.
 The construction of a partitioned grove from the grove $\mathsf{decomp}(G)$ requires the designation of a \textit{distinguished root} location $\ell_r=(v_r=(u_r,k_r),p_r)$ such that $v_r$ is an NP root in $G$ and $\mathsf{arity}(k_r)=\SetOf{p_r}$.
@@
 
-Edit actions operate on zippered terms~\cite{DBLP:journals/jfp/Huet97,DBLP:conf/popl/OmarVHAH17}.
+Edit actions operate on zipper terms~\cite{DBLP:journals/jfp/Huet97,DBLP:conf/popl/OmarVHAH17}.
 The definitions of zipper terms and child terms are standard, except that the cursor cannot select a child term of the form $\eid{w}{}\termMV$, but is forced to select the term $\termMV$ directly.
@@
 
-The definitions and theorems in \autoref{sub:Convergent Graphs} or \autoref{subsub:Terms}, and \autoref{subsub:Graph Decomposition} have been mechanized in the Agda proof assistant, with the exception that termination is justified separately for some definitions and proofs.
+The definitions and theorems in \autoref{sub:Convergent Graphs} and \autoref{subsub:Graph Decomposition} have been mechanized in the Agda proof assistant, with the exception that termination is justified separately for some definitions and proofs.
 In particular, \autoref{thm:Commutativity} and \autoref{thm:Recomposability} are formalized and proven.

### 2024-11-22 sections/grove-by-example.tex
directory restructuring and hiding appendices

--- 2024-10-25 OriginalSubmissionSource/grove-by-example.tex
@@
 \autoref{sub:Single-User Actions} then shows examples of edits being performed by a single user, Alice. 
-\autoref{sub:Multi-user Interactions}-\ref{sub:Merge Conflicts} then describes a collaboration between two users, Alice and Bob,
+Sections \ref{sub:Multi-user Interactions}--\ref{sub:Merge Conflicts} then describe a collaboration between two users, Alice and Bob,
 as they edit their own branches of a program and periodically merge in each other's edits, starting with examples without conflicts, then considering the various kinds of conflicts that might arise.
@@
 
-For clarity, we use odd numbers for vertex UIDs and even numbers for edge UIDs. In practice, UIDs would be generated by a mechanism
+For clarity, we use even numbers for vertex UIDs and odd numbers for edge UIDs. In practice, UIDs would be generated by a mechanism
 that effectively ensures that collaborators always generate distinct UIDs, e.g. by generating
@@
 
-After converging on \texttt{\hole{} + u * y} in \autoref{fig:DifferentParts:c}, 
+After converging on \texttt{\hole{} + u * v} in \autoref{fig:DifferentParts:c}, 
 Alice changes~\vDifferentPartsAlice{} to~\vNestedPartsAlice{},
@@
 
-Grove's support for repositioning creates the possibility for \emph{relocation conflicts}.
+Grove's support for code relocation creates the possibility for \emph{relocation conflicts}.
 These occur when a merge causes a vertex to have multiple incoming edges, indicating that it does not have a uniquely determined location (as opposed to local conflicts, which occur when there are multiple outgoing edges at a specified location).
@@
 At the same time, Bob deletes~\eNestedPartsAlice{} and adds~\eMultiParentBob{} in \autoref{fig:MultiParent:c}.
-Edge deletions is idempotent,
+Edge deletion is idempotent,
 so the fact that Alice and Bob both deleted~\eNestedPartsAlice{} will not lead to a conflict. 
@@
 When decomposing the graph, we leave a \emph{relocation conflict reference}, $\multiVertex{18}$, at each conflicting location.
-The conflicted term is separately tracked in the result of decomposition, which, due to conflicts like these, is formally a set of terms with references like these between them.
+The conflicted term is separately tracked in the result of decomposition, which, due to conflicts like these, is formally a set of terms with references between them.
 We call this set a \emph{grove}.
@@
 The inserted edges form a \emph{unicycle}, meaning a cycle where every vertex has in-degree $1$.
-In this case, we cannot rely on relocation references to break the cycle.
+In this case, we cannot rely on relocation conflict references to break the cycle.
 Instead, we break the cycle by arbitrarily but deterministically choosing an edge along the unicycle, e.g.

### 2024-11-22 sections/marking-system.tex
directory restructuring and hiding appendices

--- 2024-10-25 OriginalSubmissionSource/marking-system.tex
@@
 and requires understanding the semantics of the program
-even while it still has conflicts.
+even while it still has conflicts. Traditionally, conflicts are indicated by inserting extra-linguistic conflict markers into files.
 
-Traditionally, conflicts are indicated by inserting extra-linguistic conflict markers into files. 
 These markers can limit the operation of language services that need a well-formed or well-typed term, 
-e.g., type error reporting.
-
-Ignoring the markers is not sufficient because 
-multiple conflicting versions of the code may appear
-between the markers. This can complicate variable resolution and type checking.
+e.g., type error reporting. Ignoring the markers is not sufficient because 
+multiple conflicting versions of the code may appear between the markers. This can complicate variable resolution and type checking.
 
@@
 
-We follow the marked lambda calculus in basing our \emph{marked grove calculus} in bidirectional type checking~\cite{dunfield2019} and deploying gradual typing~\cite{siek2015} when conflicts do not allow a single type to be inferred, then layer on a unification-based type inference system to opportunistically fill holes or suggest partial solutions when there are conflicting types due to conflicting syntax.
+We follow the marked lambda calculus in basing our \emph{marked grove calculus} in bidirectional type checking~\cite{10.1145/3450952} and deploying gradual typing~\cite{siek2015} when conflicts do not allow a single type to be inferred, then layer on a unification-based type inference system to opportunistically fill holes or suggest partial solutions when there are conflicting types due to conflicting syntax.
 We give an overview of the key judgments and rules in \autoref{sub:marking-groves} but due to space considerations and because much of the development is based directly on the prior work, we leave the full details to the supplemental material and the mechanized metatheory, which is briefly outlined in \ref{sub:marking-agda}.
@@
 
-The first rule invokes the matched arrow judgment (which handles the situation where the expected type is unknown \cite{snapl15}) and performs a type consistency check with the argument type annotation, after which we can analyze the body against the expected output type while recursively marking it.
+The first rule invokes the matched arrow judgment (which handles the situation where the expected type is unknown \cite{siek2015}) and performs a type consistency check with the argument type annotation, after which we can analyze the body against the expected output type while recursively marking it.
 
@@
 \autoref{fig:marking-syntax} introduces the syntax of the marked grove calculus.
-We define an \emph{unmarked} language, which comprises unmarked expressions, $\EMV$,  child expressions, $\ELV$, types, $\TMV$, and  child types $\TLV$.
+We define an \emph{unmarked} language, which comprises unmarked expressions, $\EMV$,  child expressions, $\ELV$, types, $\TMV$, and  child types $\TLMV$.
 We also define a corresponding \emph{marked} language,  which comprises marked expressions, $\ECMV$, child expressions, $\ELMV$, types, $\TCMV$, and child types, $\TCLV$.
@@
 This motivates the need for $\PTMV$ in \autoref{fig:marking-syntax}, which extends the sort of semantic types to enable linking generated unknown types to their associated provenance denoted by $q$.
-Provenances help locate the origin of the unknown type by associating them with their location $\ell$, edge-ID(w), or vertex-ID (u).
+Provenances help locate the origin of the unknown type by associating them with their location ($\ell$), edge-ID ($w$), or vertex-ID ($u$).
 
@@
     }{ 
-      \anaConstraint{\ctx}{\conflict{\ECMV}}{\PTMV}{\bigcup_{i<n} \cConstraint_i \cup
-      \Setof{\constrain{\TUnknown^{\mathsf{exp}(\ell)}}{\PTMV}}}{\bigcup_{i<n} \mConstraint_i}
+      \anaConstraint{\ctx}{\conflict{\ECMV}}{\PTMV}{\bigcup_{i<n} \cConstraint_i}{\bigcup_{i<n} \mConstraint_i}
     }

### 2024-11-22 sections/old-formalism.tex
directory restructuring and hiding appendices

\section{Formalism}%
\label{sec:Formalism}

% TODO: try to introduce groves in section 2?
% TODO: use key phrases in section 2, e.g.: "unresolved parentage"

We will now make the intuitions developed in the previous section precise by defining a collaborative structure editor calculus called Grove.
This section introduces
  graphs (\autoref{sub:Convergent Graphs})
  and commutative graph updates (\autoref{sub:Graph Updates}).
Users do not edit graphs directly.
Instead, they work with the decomposition of graphs into groves (\autoref{sub:Groves})%
---collections of terms with empty holes, conflict holes, and references to terms with unresolved parentage.
User actions on groves translate to corresponding graph edits (\autoref{sub:User Actions}).

% TODO: point of this section is commutativity theorem

% TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)

The formalism is rather simple, \autoref{sec:Grove By Example} shows
that it is sufficient to handle many common situations,
and this simplicity aids in the predictability.

\figureArity

\subsection{Convergent Graphs}%
\label{sub:Convergent Graphs}

Let $\U$ be a set of \emph{unique identifiers} equipped with a total ordering $\leq$,
and $\K$ the \emph{constructors} of terms in the source language,
and $\P$ the \emph{positions} of subterms in the source language,
and $\Sigma = \SetOf{\bot, \Plus, \Minus}$ the
  \emph{edge states} equipped with a total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ that forms a lattice over $\Sigma$.
Now, let $\V = \U \times \K$ all possible \emph{vertices},
and $\S = \V \times \P$ all possible \emph{sources},
and $\E = \U \times \S \times \V$ all possible \emph{edges}.

A \emph{graph} $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
  where $\E = \U \times \S \times \V$,
  unique identifiers are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
  sources are drawn from $\S$,
  vertices are drawn from $\V$,
  and edge states are drawn from $\Sigma$,
  all defined below.
Graphs that represent a complete edit state have a distinguished
  root vertex $\rootVertex$
  and root position $\rootPosition$.
The \emph{inedges} of $G$ are the edges on which $G$ returns $\Plus$.
The \emph{outedges} of a source are the inedges that have the source.

An \emph{edge} $\e{=}(u, (v, p), v^\prime) \in \E$ represents a directed multi-edge
identified by $u \in \U$, originating from source $(v, p) \in \S$, with destination vertex $v' \in \V$.

The \emph{edge states} $\Sigma = \SetOf{\bot, \Plus, \Minus}$ determine which edges are live in a graph.
For example, if $G(\e) = \Plus$ for some $G : \E \rightarrow \Sigma, \e \in \E$,
then we say $\e$ is \emph{live}, i.e., it corresponds to a term in the graph decomposition, as explained in \autoref{sub:Graph Decomposition}.
The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.

A \emph{source} is a pair $(v, p) \in \S$,
  where $\S = \V \times \P$,
  vertices are drawn from $\V$,
  and positions are drawn from $\P$.

A \emph{vertex} $v = (u, k) \in \V = \U \times \K$ is a constructor instance,
  where unique identifiers are drawn from $\U$
  and constructors are drawn from $k \in \K$.
Vertices represent \emph{constructable terms}, i.e., terms that are not references or holes.
The \emph{inedges} of $v$ are the inedges that target $v$.
The \emph{ingraph} of $v$ is the subgraph on inedges that target $v$.
We write $G_v$ to denote the ingraph of $v$.
The \emph{parents} of $v$ are the source vertices of the inedges of $v$.
The \emph{ancestors} of $v$ is defined inductively as the parent vertices of $v$ plus their ancestors.
The \emph{least ancestor} of $v$ is the ancestor with the smallest unique identifier.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsubsection{Constructors, Positions, Arities}
\label{sub:Constructors, Positions, and Arities}

A constructor $k \in \K$ represents a distinct syntactic form belonging to the term language. For each $k \in \K$ we define its arity \autoref{fig:Constructors, Indexes and Arity} to be its set of valid positions—labels representing potential subterms of terms constructed by $k$—each paired with a corresponding sort indicating which terms may be constructed at that position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Well sorted graphs}
\label{sub:Well sorted graphs}

% TODO: where does this go?
% 
% Each contructor has an arity, which we define as a set of position-sort pairs. A position $p
% \in \P$ identifies a location at some originating vertex from which edges to
% destination vertices may originate. The $\arityOp$ function in
% \autoref{fig:Constructors, Indexes and Arity} gives the complete mapping from
% constructors to potential edge positions with their corresponding sorts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: put this in a definition box

A graph $G$ is well sorted if
for all $\e = (u, (u_1, k_1), p, (u_2, k_2))$
such that $G(\e) \neq \bot$,
we have $(p, \sort{k_2}) \in \arity{k_1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% In decomp section:
% - If a graph is well sorted, then there exists a grove. (i.e., exhaustiveness check for expr)
%   - want to show that expr is a total function on Exps
% - If decomp(G) = Grove, then vertices(Grove) = vertices(edges(G))
%   - every vertex should be part of the decomposition
% - correctness of grove components:
%   - $e_r$: rootVertex($e_r) = v_r$
%   - MP: for every expression in MP, |parents(the root vertex)| > 1
%   - NP: for every expression in NP, |parents(the root vertex)| = 0
%   - U: for every expression in U, |parents(the root vertex)| = 1 and the root vertex = min(ancestors(the root vertex))
%     - every other vertex notin U has exactly 1 parent and is not its own min-ancestor
%   - we've accounted for every edge:
%     - want to say edges(Grove) = edges(G), but we don't have a clean way to say that
%     - there is a choice from (an infinite family of graphs) recomposition(Grove) (which are all structurally equivalent to G, modulo edge ids)

% correctness covers:
% - invertibility (w/ some identity assignment)
% - parentage (things are multiparented correctly)

\vspace*{\baselineskip}

% TODO: should graph "recomposition" have an algorithm, too?
% TODO: update this to match our current understanding
A graph can be derived from an expression in the calculus by starting with an
empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
sub-term at some position $p \in \arity k$, select unique id $u'$ and
constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
= (u', k')$. Continuing in this manner, treat each sub-term as if it were the
outermost, starting with the ones at positions in $\arity k'$, and create edges
to the vertices representing their immediate sub-terms until only empty holes
remain. Note that empty holes are not modeled explicitly in the graph. Wherever
an empty hole appears in an expression, the corresponding graph position has no
edges originating from it.

% You can map from syntax to graph by selecting
% unique IDs $u$ (for both edges and vertexes).
% - holes are not explicit in the graph

\subsection{Graph Updates}
\label{sub:Graph Updates}

We define a graph update $s = (\alpha,\e) \in \left\{\Plus, \Minus\right\} \times \E$
as a pair of an edge state (excluding $\bot$) and an edge. As a shorthand, we
define $+\e$ to be $(\Plus, \e)$ and $-\e$ to be $(\Minus, \e)$.

We define the join operation $\alpha_1 \sqcup \alpha_2$ to be the least upper bound of
$\alpha_1$ and $\alpha_2$ with respect to the $\sqsubset$ ordering. Accordingly, $\Plus
\sqcup \bot = \Plus$ means that edges can only be created if they do not already
exist, and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$ means that once
an edge is deleted it can never be restored, and hence deletion is idempotent.

We define the semantics of graph actions via the following transition relation
between graphs.
\[
  G \overset{(\alpha,\e)}{\longrightarrow} G\left[ \e \mapsto \alpha \sqcup G(\e) \right]
\]
Applying graph action $\left(\alpha, \e\right)$ to graph $G$ results in the
updated graph $G' = G\left[\e \mapsto \alpha \sqcup G(\e) \right]$, wherein the edge
state associated with edge $\e$ in $G$ becomes the join of $\alpha$ with the state of
$\e$ in $G$, and $G'(\e') = G(\e')$ for any $\e' \ne \e$. Essentially, these graph update semantics act as a transition system between graphs, \emph{joining} the new edge state with the corresponding edge state of the graph to which the update is being applied.

% Intuitively, joining any edge state with $\bot$ is always the non-$\bot$ edge state (once an edge exists, it always exists).
% Joining anything with $\Minus$ is always $\Minus$ (once an edge is marked as deleted, it is permanently deleted).

\subsection{Commutativity}%
\label{sub:Commutativity:formal}

In order to define commutativity, we first define a lattice over $G$.
Then we show that for any graphs $G$ and $G'$ and graph action $\alpha$,
$G \overset{\alpha}\rightarrow G'$
iff $G \sqcup \llbracket\alpha\rrbracket = G'$.
This established the commutativity of graph actions
by the commutativity of the join operations defining those graph actions.

% Discuss uniqueness of uuid

\begin{lemma}[Join Commutativity]
  \label{lem:Join Commutativity}
  For all graphs G and G' and all graph updates $s_1$ and $s_2$, we have $s_1 \sqcup s_2 = s_2 \sqcup s_1$
\end{lemma}
\begin{proof}
    Follows directly from the total ordering $(\Sigma, \sqsubset)$
\end{proof}

\begin{lemma}[Joining]
  \label{lem:Joining}
  For all graphs $G$ and $G'$ and all graph updates $s_1 = (\alpha_1, \e_1)$ and  $s_2 = (\alpha_2, \e_2)$, we have
  $G \overset{s_1 s_2}{\longrightarrow} G'$
  iff $G[\e_1 \mapsto \alpha_1 \sqcup G(\e_1)][\e_2 \mapsto \alpha_2 G(\e_2)] = G'$
\end{lemma}
\begin{proof}
  If $G \overset{\alpha_1\alpha_2}{\longrightarrow} G'$
  There is a graph $G''$ such that $G[\e_1 \mapsto \alpha_1 \sqcup G(\e_1)]= G''$ and $G''[\e_2 \mapsto \alpha_2 \sqcup G(\e_2)] = G'$. For any edge \e, if $\e = \e_1$, then $G'(\e)$ = $\alpha_1 \sqcup G''(\e_1)$, and if $\e = \e_2$, then $G'(\e)$ = $\alpha_2 \sqcup G''(\e_2)$ = $\alpha_2 \sqcup G(\e_2)$, and otherwise $G'(\e) = G(\e)$.
\end{proof}

\begin{theorem}[Commutativity]
  \label{thm:Commutativity}
  For all graphs $G$ and $G'$ and all graph updates $s_1$ and $s_2$, 
  if $G \overset{s_1 s_2}{\longrightarrow} G'$,
  then $G \overset{s_2 s_1}{\longrightarrow} G'$
\end{theorem}
\begin{proof}
  By \autoref{lem:Joining}, if $\e_1 \ne \e_2$, we have $G'(\e_1) = \alpha_1 \sqcup G'(\e_1)$ and $G'(\e_2) = \alpha_2 \sqcup G(\e_2)$, and if $\e_1 = \e_2 = \e$, we have $G'(\e) = \alpha_1 \sqcup \alpha_2 \sqcup G(\e)$. Hence, for any edge \e, $\e_1$ and $\e_2$ are not equal we have $G'(\e) = G[\e_1 \mapsto \alpha_1 \sqcup G(\e_1)][\e_2 \mapsto \alpha \sqcup G(\e_2)](\e) = G[\e_2 \mapsto \alpha \sqcup G(\e_2)][\e_1 \mapsto \alpha_1 \sqcup G(\e_1)](\e)$ and otherwise by \autoref{lem:Join Commutativity}, we have $G'(\e) = \alpha_1 \sqcup \alpha_2 \sqcup G(\e) = \alpha_2 \sqcup \alpha_1 \sqcup G(\e)$
\end{proof}

\subsubsection{Agda Mechanization}%
\label{sub:Agda Mechanization}

\textbf{TODO}

\subsection{Groves}%
\label{sub:Groves}

\figureTermSyntax

% TODO: check for \hole --> \emptyHole

Our source language (\autoref{fig:Syntax}) is a simply typed lambda calculus with references, empty holes, and conflict holes.
The syntax of terms is sorted into
  expressions $e$,
  types $\tau$,
  and patterns $q$.
% These can include
%   graph annotations $G$,
%   edge annotations $\e$,
%   or source annotations $(v, p)$,
%   which we introduce in \autoref{sub:Convergent Graphs}.
Expressions include
  variables $\eVar{G}{x}$,
  lambda abstractions with pattern arguments $\eFun{G}{q}{\tau}{e}$,
  applications $\eApp{G}{e_1}{e_2}$,
  and numbers $\eNum{G}{n}$ with addition $\ePlus{G}{e_1}{e_2}$ and multiplication $\eTimes{G}{e_1}{e_2}$.
Types include
  arrows $\tArrow{G}{\tau_1}{\tau_2}$
  and the type of numbers $\tNum{G}$.
Patterns include variables $\pVar{G}{x}$.

We also extend each sort with 
empty holes $\hole_{(v, p)}$,
multiparent references $\multiVertex{\e}$,
unicycle references $\cycleVertex{\e}$,
and conflict holes $\conflictHole{t_i}_{i < n}$.
For conflict holes, it must be the case that $n \geq 2$.
We identify conflict holes up to reordering.

Multiparent references correspond to edges that target
  \emph{multiparent roots}---vertices targeted by multiple edges.
Unicycle references correspond to edges that target
  \emph{unicycle roots}---vertices that are their own least ancestor.
  Empty holes, which denote the absence of any corresponding structure,
are annotated with the source needed to construct terms in their place.

% Each term is annotated with a graph $G$, an edge $\e$, or a source $(\v,\p)$,

Each term is annotated with a graph $G$, an edge $\e$, or a source ($v$,$p$), which we introduced in \autoref{sub:Convergent Graphs}. Each term that is not a conflict hole carries an annotation.
For terms that are not holes, the annotation corresponds to the graph or edge it recomposes to. For constructable terms, the annotation corresponds to the graph it \emph{recomposes} to, as we explain in \autoref{sub:Graph Decomposition}. Multiparent and unicycle references are annotated with the edge whose destination is the referenced vertex. Empty holes are annotated with the source of their local parents. Since a conflict hole corresponds to all of the edges with a particular common source, they cannot be nested or contain empty holes.

% and the combined annotations correspond to the graph the entire conflict hole recomposes to.

\subsection{Graph Decomposition}%
\label{sub:Graph Decomposition}

% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
% and multiparents

% NOTE: this whole section is for describing the 4-tuple

Trees are often easier to work with than arbitrary graphs directly, for example to present the graph to users or for type checking.
A graph can be \emph{decomposed} into a collection of terms governed by the grammar in \autoref{fig:Syntax}, which we organize into a structure called a grove that readily \emph{recomposes} into the original graph.

\figureDecompExample

% TODO: make sure we're not claiming to invent the term "unicycular graph"
% TODO: maybe add a forward reference at end here to Implementation section
% TODO: make ID of y lower than the unicycle root, to show what we're minimizing

A grove is a 3-tuple of the following form:
    \figureDecompositionDefSets
where
\begin{align*}
  \figureDecompositionDefDecompComponents
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
given in \autoref{fig:Decomposition example grove}.

Let us now examine each of the three components in more detail.
%
The first component $\Set[NP]$ is the set of expressions rooted at a vertex with no parents, including the main root vertex.
% TODO: change vRoot to dot w/ subscript
In the example, $e_r$ is the expression corresponding to the graph root, $\vRoot$ / $._{\id{0}}$.
% TODO: change z_... to macro use w/ same appearance as in the graph
Expression, $z_{\id{52}}$ has been deleted and has no parents, so it also appears in $\Set[NP]$.

The first component $e_r$ is the expression derived from the root vertex $v_0$.
% \autoref{fig:Decomposition expr} defines $\expr(v)$ which determines the expression derived from $v$.
There is one rule for each term constructor.
For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr({+}_{\id{38}}, \PlusLeft) \text{+}^ {\id{38}} \expr({+}_{\id{38}}, \PlusRight)$.
In each child position, $\expr'(v, p)$ determines the subterm for the children of $v$ at position $p$.
An empty hole $\hole$ appears if there are no children at the given position.
A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
If there is one child, the expression is determined recursively unless that child appears in $MP$ or $U$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$, respectively.
For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.

The second component, $\Set[MP]$, is the set of expressions derived from a vertex with multiple parents. Wherever an edge to such a vertex appears, we place a reference to it $\multiVertex{u}$.
In the example, $x_{\id{42}}$ is referenced twice in $e_r$.

The third component $\Set[U]$ is the set of expressions derived from the remaining vertices, which necessarily have exactly one parent.
These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
Every unicycle consists of a single cycle, and each vertex may have additional children that are not part of the cycle. Breaking the cycle turns a unicyclic graph into a tree.
% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
Formally, we choose the vertex $v$ such that $v = \min{\ancestors{v}}$ because
only a vertex on the cycle can be its own ancestor.
An edge to this unicycle root decomposes into a unicycle reference $\cycleVertex{u}$.
In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.

Finally, we discover the unicycles of $\Set[U]$ by traversing any remaining edges, which necessarily constitute the elements of unicycular graph or another. Unicycular graph traversal occurs in two directions. We begin by choosing a a vertex as an arbitrary starting point, traverse its parents, and marking all of its vertices as seen. Since a vertex that has been seen twice must be on the cycle, we take the first such vertex as the root of the unicycle. We then traverse its children to produce a tree that covers the entire unicycular graph. Any remaining single-parented vertices must be parts of other unicycular graphs and are subsequently \emph{unicycle}-traversed until there are no more single-parented vertices left.

\begin{lemma}
    Unicycular graph traversal of a well-sorted graph produces a well-sorted tree
\end{lemma}

We say a term is \emph{well-sorted} when it recomposes into a well-sorted graph. The following theorem makes this intuition precise.

\begin{theorem}
    If $G$ is a well-sorted graph which decomposes into a grove $\Grove = (\Set[NP], \Set[MP], \Set[U])$, then any term $t \in \Set[NP] \cup \Set[MP] \cup \Set[U]$ recomposes into a well-sorted graph, and the composite of all such recompositions $\bigcup_{t\in \Set[NP] \cup \Set[MP] \cup \Set[U]}$ $\recomp{t}$ (viewed as sets of mappings from edges to edge states) is a graph $G'$ isomorphic to $G$.
\end{theorem}

\begin{proof}
    By rule induction on the derivations of $\Grove$ and $G'$
\end{proof}

That is, a well-sorted graph decomposes into a collection of terms that, when recomposed, produce the original graph. We give detailed information on decomposition and recomposition in \autoref{sec:Grove Formalism}.

% XXX

% decomposition produces sets of trees. As long as everything is well sorted, we
% can use them to produce expressions in the main grammar. (alt: define wrt the
% graph instead, once we establish the graph can be turned into expressions)

% *** well-sortedness judgment on trees: the tree only has valid positions for the
% constructors, and the child at each position has the sort corresponding to its
% position, and edges in position maps all start at the given constructor.

% Theorem: well-sorted graphs produce decompositions of well-sorted trees

% ``A graph has this property iff these conditions hold''

% then we can define a total mapping from well-sorted trees to expressions
% (because the tree structure is very generic, so this ties them back to the
% earlier grammar.)

% forest = 4-tuple

% forest validity: each of the trees in the forest are well sorted, and all Refs
% point to roots of other trees of the correct sort.

% every valid forest corresponds to a well behaved grove (i.e., ``expression
% forest'')

% soundness: if you start with a valid graph, you get a well behaved grove that
% connects every vertex (and only those vertices) in the graph. (i.e., if you turn
% it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
% Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
% add Root sort

% OR:

% use edge IDs instead of edges and build a mapping from vertices to
% sorts---change the $\e$s to $u$s)

% % TODO: maybe add a forward reference here to Implementation section

\figureUserActionSyntax

\subsection{User Actions}%
\label{sub:User Actions}

Given the grammar for user actions in \autoref{fig:User Action Syntax}, we want to show that user actions on terms correspond to equivalent graph updates. This can be done in either of two directions: from graph to grove or vice versa. Showing that operating on a grove is equivalent to operating on the underlying graph enables a more efficient implementation than the other way around, as then the editor is free to leverage term annotations to avoid a full decomposition pass after each user action, but ensuring the soundness of the resulting graph requires proving an additional theorem. For simplicity, we chose to translate all user edits directly to graph updates and then perform a decomposition pass over the entire graph after each user action.

Here, we give intuitive explanations for the user actions of \autoref{fig:User Action Syntax}. For a more formal treatment, see \autoref{sec:Grove Formalism}.

$\Construct{k}$ constructs a new edge whose destination is determined by the cursor's location. If the cursor is on an empty hole, the hole is filled by constructing a fresh vertex with constructor $k$. If the cursor is on an existing term and the constructor $k$ has a child position of the same sort, we \emph{wrap} the existing term by deleting it from its parent, constructing the new term in its place, then re-connecting the original term as a child of the new term. Otherwise, the editor may choose whether to apply updates that would produce conflicts.

Depending on the cursor's location, $\Delete$ deletes one or more edges. If the cursor is on a constructable term of reference, its corresponding edge is deleted. If the cursor is on a conflict hole, all the edges of the conflict are deleted. Otherwise, the cursor is on an empty hole in which case the the action is a no-op.

$\Reposition{v}{p}$ combines edge construction and deletion into a single \emph{atomic} operation. If the cursor is not an empty hole, the corresponding edge(s), as indicated above, are simultaneously deleted and then reconstructed as child(ren) with edge source $(v,p)$.

% TODO: define a grammar for user actions --> mapping to sequences of underlying graph edits
% TODO: define well-sorted graph & theorem: graph is well sorted => exists a decomposition

% move = delete + add

% other composite edits

% remember commutativity diagram: we want edits on trees to be equivalent to edits on graphs

% given a semantics for user actions, we can show that operating on the grove is equivalent to operating on the graph.
% This way is more efficient (less annoying?) because then we don't have to deal with decomposition for every edit.
% (we probably want to do it this way in Hazel)

% OR

% We could go the indirect route and avoid having to prove the theorem.
% (this way is easier to start with, and it's how GRV works)

% Mapping to b actions

% \[
%   \begin{array}{l}
%     \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
%     \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
%     \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
%     \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
%     \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
%     \CC(G) = \ldots                                                                                                                                    \\
%     \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
%     \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
%     \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
%     \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
%     \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
%     \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
%   \end{array}
% \]

### 2024-11-22 sections/related-work.tex
directory restructuring and hiding appendices

--- 2024-10-25 related-work.tex
@@
 In the context of collaborative coding, patch languages can differ in the data structures they operate on (e.g.
-line-based text~\cite{DBLP:conf/haskell/Roundy05}, character-based text~\cite{DBLP:conf/sigmod/EllisG89,DBLP:journals/pacmhci/LittLKH22}, tree-structured data).
+line-based text~\cite{DBLP:conf/haskell/Roundy05}, character-based text~\cite{DBLP:conf/sigmod/EllisG89,DBLP:journals/pacmhci/LittLKH22}, tree-structured data~\cite{DBLP:conf/esa/Klein98}).
 They can also differ in how they identify locations within the data (e.g. by using numeric offsets~\cite{DBLP:conf/sigmod/EllisG89}, unique identifiers, or paths through a tree).
@@
 The most common approach to merging concurrently developed patches is to deploy an operational transform~\cite{DBLP:conf/sigmod/EllisG89} whereby locations in a remote patch are modified based on the action of a local patch.
-Standard three-way merge algorithms in version control systems deploy this approach, as do real-time collaborative editors.
+Standard three-way merge algorithms in version control systems deploy this approach, as do real-time collaborative rich text editors~\cite{DBLP:journals/concurrency/IgnatAO21}.
+
 There have been a number of papers studying the algebraic properties of merging patches in this style.
-
 For example, the Darcs~\cite{DBLP:conf/haskell/Roundy05} version control system, like Grove, represents repositories using sets of patches.
@@
 
-Other examples of CRDT-based editors include and Peritext~\cite{DBLP:journals/pacmhci/LittLKH22} and Zed~\cite{zed-blog}.
-These systems differ from Grove by targeting synchronous editing rather than version control and by operating on sequential text rather than tree structures.
+Much of the prior work on CRDT-based collaborative editing has focused on text editing 
+\cite{DBLP:conf/icdcs/PreguicaMSL09, 
+DBLP:conf/cscw/OsterUMI06,
+DBLP:journals/jpdc/RohJKL11,
+DBLP:conf/wikis/Grishchenko10,
+DBLP:conf/icdcs/WeissUM09,
+DBLP:conf/doceng/NedelecMMD13,
+DBLP:conf/group/NicolaescuJDK16,ahmed2011evaluating,kleppmann2020moving}, typically with text tagged with unique IDs.
+These techniques have been used to develop collaborative text editors, e.g.
+Zed~\cite{zed-blog}.
 
-There have been other recent efforts to develop CmRDTs for tree data structures~\cite{DBLP:conf/icdcs/PreguicaMSL09, Highly-Available}.
+Peritext develops a CRDT-based approach for collaboratively editing rich text, which is structured as text with hierarchically marked regions~\cite{DBLP:journals/pacmhci/LittLKH22}.
+
+There have been other recent efforts to develop CmRDTs for richer tree data structures~\cite{DBLP:conf/icdcs/PreguicaMSL09, Highly-Available}.
 However, the focus in this work has been on avoiding conflicts and cycles entirely by applying \emph{ad hoc} heuristics for conflict resolution at merge-time, e.g.

### 2024-11-22 sections/discussion-and-conclusion.tex
directory restructuring and hiding appendices
add artifact citation

--- 2024-10-25 discussion-and-conclusion.tex
@@
 string literals.
-We also leave to future work the problem of combining existing work on sequence CRDTs~\cite{ahmed2011evaluating,kleppmann2020moving} with our work on tree/graph CRDTs.
+We also leave to future work the problem of combining existing work on sequence CRDTs with our work on tree/graph CRDTs.
 
@@
 % GUI for string conflicts: use popups
+
+\section*{Data Availability Statement}
+The artifact \cite{grove-artifact} that accompanies this work contains the Agda mechanization of the theorems discussed in section~\ref{sub:marking-agda} and section~\ref{sec:formalism-agda} and contains the implementation of the Grove Workbench discussed in section~\ref{sec:Implementation}.

### 2024-11-22 sections/formalism.tex
add artifact citation

@@
 \subsection{Mechanized Metatheory}
+\label{sec:formalism-agda}
 
-The definitions and theorems in \autoref{sub:Convergent Graphs} and \autoref{subsub:Graph Decomposition} have been mechanized in the Agda proof assistant, with the exception that termination is justified separately for some definitions and proofs.
+The definitions and theorems in \autoref{sub:Convergent Graphs} and \autoref{subsub:Graph Decomposition} have been mechanized~\cite{grove-artifact} in the Agda proof assistant, with the exception that termination is justified separately for some definitions and proofs.
 In particular, \autoref{thm:Commutativity} and \autoref{thm:Recomposability} are formalized and proven.

### 2024-11-22 sections/marking-system.tex
add artifact citation

@@
 
-We have mechanized our extension of the marked lambda calculus and these metatheorems using the Agda proof assistant, taking the standard approach of modeling judgments as inductive datatypes and inference rules as constructors.
+We have mechanized~\cite{grove-artifact} our extension of the marked lambda calculus and these metatheorems using the Agda proof assistant, taking the standard approach of modeling judgments as inductive datatypes and inference rules as constructors.
 Marked terms are intrinsically typed, allowing $L$ to be computed as a function of the term.

### 2024-12-06 figures/figure-different-parts-nested-parts.tex
Updates from Overleaf
Thomas: updated acmart.cls

@@
 \begin{minipage}[t]{0.45\linewidth}
-%\hfill
-\hskip0.12\columnwidth
+\hfill
+% \hskip0.12\columnwidth
 %%%%%%%%
@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Move:b}}}};
+\path (-5em,0cm) node (node/Move/bx) [gray,anchor=base west] {{\textbf{\small \ref*{fig:Move:b}}}};
 \path (0cm,0cm) node (node/DifferentParts/a) [anchor=base] {\strut};
-\path [draw,->,alice step] (node/Move/bx) to (-2em, 0cm |- node/DifferentParts/a);
+\path [draw,->,alice step] (node/Move/bx) to (-1.5em, 0cm |- node/DifferentParts/a);
 \end{tikzpicture}
@@
 % \begin{figure}
-\begin{minipage}[t]{0.45\linewidth}
+\begin{minipage}[t]{0.48\linewidth}
 \hfill
@@
 %%%%%%%%
-\begin{subfigure}[t]{0.25\linewidth}
+\begin{subfigure}[t]{0.3\linewidth}
 \centering
@@
 \label{fig:NestedParts:c}
-\figureCode{w * v + \hole{}}%
+{\small \figureCode{w * v + \hole{}}}%
 \begin{tikzpicture}

### 2024-12-06 figures/figure-multi-child.tex
Updates from Overleaf
Thomas: updated acmart.cls

@@
 \begin{tikzpicture}[remember picture, overlay]
-\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:NestedParts:c}}}};
+\path (-7.5em,0cm) node (node/NestedParts/cx) [gray,anchor=base west] {{\textbf{\small \ref*{fig:NestedParts:c}}}};
 %%%%
@@
 (c) When the corresponding patches are merged, there are two edges in the \li{R} position of the addition.
-Decomposition turns these into a \emph{local conflict}, leaving it to the users to resolve the problem by performing normal edits that eliminate .}%
+Decomposition turns these into a \emph{local conflict}, leaving it to the users to resolve the problem by performing normal edits.}%
 \Description{This figure shows an example of multi-child conflicts}
